### 2013-03-14 documentation/table-specification.tex
Initial commit

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage[hyphens]{url}
\usepackage[multiple]{footmisc}
\usepackage{longtable}
\usepackage[top=1.5in, bottom=1.0in, left=0.5in, right=0.5in]{geometry}

\newcommand{\tbd}{{\tiny \textbf{TBD}}}

\title{IPHAS DR2: Table specification}
\date{\today}
\author{Geert Barentsen}

\begin{document}

\maketitle

This document describes the proposed set of tables and columns which are to be 
included in IPHAS Data Release 2. In brief, it is proposed to create four 
tables:
\begin{itemize}
\item \textbf{iphasDetection}: the complete collection of all objects detected 
in those IPHAS exposures which pass the essential quality criteria. Each row 
represents a single-epoch, single-band detection, and hence this table 
contains duplicate detections of unique objects.
\item \textbf{iphasSource}: band-merged version of the detections table, in 
which detections have been cross-matched on a field-by-field basis. Unique 
sources may have duplicate entries in this table, but the `best' entry for 
each unique source will be flagged. 
\item \textbf{iphasExposure}: meta-data on the observing conditions for each 
run.
\item \textbf{iphasField}: meta-data on the observing conditions, based on 
iphasExposure but collated on a field-by-field basis with extra information added (cf. the existing iphas-qc.fits table).
\end{itemize}

On top of these four tables we may define `views' which define subsets of 
certain rows and columns for the purpose of being user-friendly. For example, 
there might be a view called \textbf{iphasPointSource} which consists of those 
entries from iphasSource which are marked `primary detections', hence avoiding duplicates. In what follows we concentrate on defining the four main tables however.

It is important to note that the naming conventions and column descriptions 
used below are largely based on two existing tables in the UKIDSS/GPS survey, 
namely `gpsDetection' and `gpsSource' (see the WFAU website for the 
definitions of these tables\footnote{\url{http://surveys.roe.ac.uk/wsa/www/WSA_TABLE_gpsSourceSchema.html}}
\footnote{\url{http://surveys.roe.ac.uk/wsa/www/WSA_TABLE_gpsDetectionSchema.html}}). 
There are several reasons to adopt the UKIDSS format:
\begin{itemize}
\item a consistent format will encourage UKIDSS users to adopt IPHAS;
\item UKIDSS have an existing algorithm in place to merge the single-band 
morphological classifications into `mergedClass'/`mergedClassStat' columns, as 
explained in their schema browser
\footnote{\url{http://surveys.roe.ac.uk/wsa/www/gloss_m.html#gpssource_mergedclass}}
\footnote{\url{http://surveys.roe.ac.uk/wsa/www/gloss_m.html#gpssource_mergedclassstat}};
\item UKIDSS have an existing method to flag duplicate detections and mark the 
best one\footnote{\url{http://surveys.roe.ac.uk/wsa/www/gloss_p.html#gpssource_priorsec}}.
\end{itemize}

\newpage
\section{Table definitions}
\subsection{iphasDetection}

\begin{center}
\begin{longtable}{llclp{10cm}}
\caption[short]{Definition of the \textbf{iphasDetection} table.} \label{iphasSource} \\

\hline \textbf{Name} & \textbf{Type} & \textbf{Length} & \textbf{Unit} & \textbf{Description} \\ \hline
\endfirsthead

\hline \textbf{Name} & \textbf{Type} & \textbf{Length} & \textbf{Unit} & \textbf{Description} \\ \hline
\endhead

\hline \multicolumn{5}{r}{{\it Continued on next page}} \\ 
\endfoot

\hline \hline
\endlastfoot

detectionID & int & 8 & & Unique identification number formed by concatenating runID (6 or 7 digits), ccd (1 digit), seqNum (6 digits). \\
runID & int & 4 & & Original identifier of the run (i.e. exposure) in the observing logs of the Isaac Newton Telescope (INT). 
References to the iphasExposure table. \\
ccd & int & 1 & & Extension number of the CCD of the Wide-Field Camera (WFC). \\
seqNum & int & 4 & & Unique running number of this detection (unique withing the given runID/ccd combination, assigned by the image detection pipeline). \\
band & char & 2 & & Filter used. One of `r', `i', `ha'. \\ 
x & real & 4 & pixel & X coordinate of the detection in the CCD reference frame. \\
y & real & 4 & pixel & Y coordinate of the detection in the CCD reference frame. \\
ra & real & 8 & degrees & J2000 right ascension with respect to the 2MASS PSC reference frame (which is consistent with ICRS to within 0.1 arcsec). \\
dec & real & 8 & degrees & J2000 declination with respect to the 2MASS PSC reference frame (which is consistent with ICRS to within 0.1 arcsec). \\
posErr & real & 4 & arcsec & Astrometric fit error (RMS). \\

gauSig & real & 4 & pixels & RMS of axes of ellipse fit, i.e. sqrt(a**2 + b**2). \\
ell & real & 4 & & Ellipticity (defined as 1-b/a, where a/b=semi-major/semi-minor axes).  \\
pa & real & 4 & & Position angle of ellipse major axis with respect to the x-axis. \\

peakMag & real & 4 & mag & Aperture corrected magnitude in the Vega system 
(based on the peak pixel height.) \\
peakMagErr & real & 4 & mag & Uncertainty in peakMag. Does not include systematics. \\ 
aperMag1 & real & 4 & mag & Aperture corrected magnitude in the Vega system (1.17 arcsec aperture diameter, i.e. 1/2 x rcore radius). \\ % 1/2 x 7 pixels
aperMag1err & real & 4 & mag & Uncertainty in aperMag1. Does not include systematics. \\ 
aperMag2 & real & 4 & mag & Aperture corrected magnitude in the Vega system (2.33 arcsec aperture diameter, i.e. rcore radius). \\ % 7 pixels
aperMag2err & real & 4 & mag & Uncertainty in aperMag2. Does not include systematics. \\ 
aperMag3 & real & 4 & mag & Aperture corrected magnitude in the Vega system (3.30 arcsec aperture diameter, i.e. sqrt(2) x rcore radius). \\ % sqrt(2) x 7 pixels
aperMag3err & real & 4 & mag & Uncertainty in aperMag3. Does not include systematics. \\ 
aperMag4 & real & 4 & mag & Aperture corrected magnitude in the Vega system (4.67 arcsec aperture diameter, i.e. 2 x rcore radius). \\ % 2 x 7 pixels
aperMag4err & real & 4 & mag & Uncertainty in aperMag4. Does not include systematics. \\ 
aperMag5 & real & 4 & mag & Aperture corrected magnitude in the Vega system (6.60 arcsec aperture diameter, i.e. 2 sqrt(2) x rcore radius). \\ % 2 x sqrt(2) x 7 pixels
aperMag5err & real & 4 & mag & Uncertainty in aperMag5. Does not include systematics. \\ 
sky & real & 4 & {\sc ADU} & Local interpolated sky level from background tracker. \\
skyVar & real & 4 & {\sc ADU} & Local estimate of variation in sky level around 
image. \\
class & int & 2 & & Flag indicating the most probable morphological classification (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated).\\
classStat & real & 4 & & N(0,1) stellarness-of-profile statistic. \\
badPix & real & 4 &  & Number of bad pixels. \\
deblend & bool & 1 &  & True if the object had to be deblended, i.e. if it was part of overlapping (blended) images. \\ 
saturated & bool & 1 &  & True if the peak pixel is saturated. \\ 
truncated & bool & 1 &  & True if the object is truncated (too close to an image boundary, defined as being within 4 arcsec from the CCD edge). \\ 
brightNeighb & bool & 1 &  & True if the object is within 10 arcmin of a star brighter than V < 4.5. \\ 
reliable & bool & 1 & & True if the source is stellar, not deblended, not saturated, not truncated, not near a bright neighbour, and has less than 1 bad pixel. \\

seeing & real & 4 & & Average seeing in the exposure. \\
mjd & real & 4 & & Modified Julian Date at start of exposure. \\

\end{longtable}
\end{center}

\newpage

\subsection{iphasSource}

\begin{center}
\begin{longtable}{llclp{9cm}}
\caption[short]{Definition of the \textbf{iphasSource} table.} \label{iphasSource} \\

\hline \textbf{Name} & \textbf{Type} & \textbf{Length} & \textbf{Unit} & \textbf{Description} \\ \hline
\endfirsthead

\hline \textbf{Name} & \textbf{Type} & \textbf{Length} & \textbf{Unit} & \textbf{Description} \\ \hline
\endhead

\hline \multicolumn{5}{r}{{\it Continued on next page}} \\ 
\endfoot

\hline \hline
\endlastfoot

sourceID & int & 8 &  & Unique identification number assigned by the merge algorithm. \\
name & char & 25 &  & Official designation of the form `IPHAS JHHMMSS.ss+DDMMSS.s'. \\
ra & real & 8 & degrees & J2000 right ascension with respect to the 2MASS PSC reference frame (which is consistent with ICRS to within 0.1 arcsec). \\
dec & real & 8 & degrees & J2000 declination with respect to the 2MASS PSC reference frame (which is consistent with ICRS to within 0.1 arcsec). \\
l & real & 8 & degrees & Galactic longitude converted from ra/dec (IAU 1958). \\
b & real & 8 & degrees & Galactic latitude converted from ra/dec (IAU 1958). \\
pos\_err & real & 4 & arcsec & Astrometric fit error (RMS). \\
mergedClassStat & real & 4 &  & Merged N(0,1) stellarness-of-profile statistic. \\
mergedClass & int & 2 &  & Combined class flag from three bands (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). \\
pStar & real & 4 &  & Probability that the source is a star. \\
pGalaxy & real & 4 &  & Probability that the source is a galaxy. \\
pNoise & real & 4 &  & Probability that the source is noise. \\
pSaturated & real & 4 &  & Probability that the source is saturated. \\

r & real & 4 & mag & Default r-band magnitude in the Vega system (using aperMag2, i.e. 2.3 arcsec aperture diameter). \\ 
rErr & real & 4 & mag & Uncertainty for r. Does not include systematic errors. \\ 
rPeakMag & real & 4 & mag & r-band magnitude in the Vega system (based on the peak pixel height.) \\ 
rPeakMagErr & real & 4 & mag & Uncertainty in rPeakMag. Does not include systematics. \\ 
rAperMag1 & real & 4 & mag & r-band magnitude in the Vega system (using aperMag1, i.e. 1.2 arcsec aperture diameter) \\ 
rAperMag1err & real & 4 & mag & Uncertainty in rAperMag1. Does not include systematics. \\ 
rAperMag3 & real & 4 & mag & r-band magnitude in the Vega system (using aperMag1, i.e. 3.3 arcsec aperture diameter) \\ 
rAperMag3err & real & 4 & mag & Uncertainty in rAperMag3. Does not include systematics. \\ 
rGauSig & real & 4 & pixels & RMS of axes of ellipse fit in r. \\ 
rEll & real & 4 & & Ellipticity in the r-band. \\
rPA & real & 4 & & Position angle in the r-band. \\
rDeblend & bool &  &  & True if the object had to be deblended in the r band, i.e. if it was part of overlapping images. \\ 
rClass & int & 2 & & Discrete image classification flag. \\
rClassStat & real & 4 & & N(0,1) stellarness-of-profile statistic. \\
rSaturated & bool &  &  & True if the peak pixel is saturated in the r-band. \\ 
rOnEdge & bool &  &  & True if the detection is located within 4 arcsec (12 pixels) from the edge of the CCD. \\ 
rReliable & bool & & & True if the source is stellar, not deblended, not saturated, not near the edge, and detected at SNR $>$ 5. \\
rSeeing & real & 4 & & Average seeing in the r-band exposure. \\
rDetectionID & int & 8 & & Reference to the r-band detection of this source in the iphasDetection table. \\

i & real & 4 & mag & Default i-band magnitude in the Vega system (using aperMag2, i.e. 2.3 arcsec aperture diameter). \\ 
rErr & real & 4 & mag & Uncertainty for i. Does not include systematic errors. \\ 
iPeakMag & real & 4 & mag & i-band magnitude in the Vega system (based on the peak pixel height.) \\ 
iPeakMagErr & real & 4 & mag & Uncertainty in iPeakMag. Does not include systematics. \\ 
iAperMag1 & real & 4 & mag & i-band magnitude in the Vega system (using aperMag1, i.e. 1.2 arcsec aperture diameter) \\ 
iAperMag1err & real & 4 & mag & Uncertainty in iAperMag1. Does not include systematics. \\ 
iAperMag3 & real & 4 & mag & i-band magnitude in the Vega system (using aperMag1, i.e. 3.3 arcsec aperture diameter) \\ 
iAperMag3err & real & 4 & mag & Uncertainty in iAperMag3. Does not include systematics. \\ 
iGauSig & real & 4 & pixels & RMS of axes of ellipse fit in i. \\ 
iEll & real & 4 & & Ellipticity in the i-band. \\
iPA & real & 4 & & Position angle in the i-band. \\
iDeblend & bool &  &  & True if the object had to be deblended in the i-band, i.e. if it was part of overlapping images. \\ 
iClass & int & 2 & & Discrete image classification flag. \\
iClassStat & real & 4 & & N(0,1) stellarness-of-profile statistic. \\
iSaturated & bool &  &  & True if the peak pixel is saturated in the i-band. \\ 
iOnEdge & bool &  &  & True if the detection is located within 4 arcsec (12 pixels) from the edge of the CCD. \\ 
iReliable & bool & & & True if the source is stellar, not deblended, not saturated, not near the edge, and detected at SNR $>$ 5. \\
iSeeing & real & 4 & & Average seeing in the i-band exposure. \\
iDetectionID & int & 8 & & Reference to the i-band detection of this source in the iphasDetection table. \\

ha & real & 4 & mag & Default H-alpha magnitude in the Vega system (using aperMag2, i.e. 2.3 arcsec aperture diameter). \\ 
rErr & real & 4 & mag & Uncertainty for ha. Does not include systematic errors. \\ 
iPeakMag & real & 4 & mag & H-alpha magnitude in the Vega system (based on the peak pixel height.) \\ 
iPeakMagErr & real & 4 & mag & Uncertainty in haPeakMag. Does not include systematics. \\ 
iAperMag1 & real & 4 & mag & H-alpha magnitude in the Vega system (using aperMag1, i.e. 1.2 arcsec aperture diameter) \\ 
iAperMag1err & real & 4 & mag & Uncertainty in haAperMag1. Does not include systematics. \\ 
iAperMag3 & real & 4 & mag & H-alpha magnitude in the Vega system (using aperMag1, i.e. 3.3 arcsec aperture diameter) \\ 
iAperMag3err & real & 4 & mag & Uncertainty in haAperMag3. Does not include systematics. \\ 
iGauSig & real & 4 & pixels & RMS of axes of ellipse fit in H-alpha. \\ 
iEll & real & 4 & & Ellipticity in H-alpha. \\
iPA & real & 4 & & Position angle in H-alpha. \\
iDeblend & bool &  &  & True if the object had to be deblended in H-alpha, i.e. if it was part of overlapping images. \\ 
iClass & int & 2 & & Discrete image classification flag. \\
iClassStat & real & 4 & & N(0,1) stellarness-of-profile statistic. \\
iSaturated & bool &  &  & True if the peak pixel is saturated in H-alpha. \\ 
iOnEdge & bool &  &  & True if the detection is located within 4 arcsec (12 pixels) from the edge of the CCD. \\ 
iReliable & bool & & & True if the source is stellar, not deblended, not saturated, not near the edge, and detected at SNR $>$ 5. \\
iSeeing & real & 4 & & Average seeing in the H-alpha exposure. \\
iDetectionID & int & 8 & & Reference to the H-alpha detection of this source in the iphasDetection table. \\

date & char & 10 & yyyy-mm-dd & Date of the night in which this source was 
observed. \\
fieldID & int & \tbd & & Unique identifier of the field observation. \\

r2 & real & 4 & mag & Alternative magnitude obtained in the same night as the default magnitude, if available. \\ 
i2 & real & 4 & mag & Alternative magnitude obtained in the same night as the default magnitude, if available. \\
ha2 & real & 4 & mag & Alternative magnitude obtained in the same night as the default magnitude, if available. \\ 
r2\_err & real & 4 & mag & Uncertainty for r2. Does not include systematic errors. \\ 
i2\_err & real & 4 & mag & Uncertainty for i2. Does not include systematic errors. \\
ha2\_err & real & 4 & mag & Uncertainty for ha2. Does not include systematic errors. \\ 
r2\_detectionID & \tbd & & & Identifier in the detection table. \\
i2\_detectionID & \tbd &  & & Identifier in the detection table. \\
ha2\_detectionID & \tbd & & & Identifier in the detection table. \\
nobs & int & 1 & & Total number of detections of this source available in the iphasDetection table. 
\\

\end{longtable}
\end{center}

\subsection{iphasRun}

Indexed by runID.

\tbd

\subsection{iphasField}

Indexed by fieldID.

\tbd

\section{TODO}
\begin{itemize}
\item How merge onEdge/Saturated/Reliable columns?
\item Show example queries
\item Bright star flag?
\item Columns for H-alpha are named i.
\item Change terminology: alternative magnitude into secondary magnitude.
\item Offset matching distances between bands, i.e. 4 columns i_dra, i_ddec, 
h_dra, h_ddec.  (UKIDSS has terminology... but maybe just iRA, haRA, etc is 
better?)
\item Promise a max time-interval for r2/i2/ha2, and promise that it's the 
partner.
\end{itemize}

\end{document}

### 2013-03-14 scripts/create-iphasDetections.py
Initial commit

@@ -40,7 +40,7 @@ else:
     DESTINATION = "/car-data/gb/iphasDetection"
 
 # Yale Bright Star Catalogue (Vizier V50), filtered for IPHAS area and V < 4.5
-BRIGHTCAT = fits.getdata('BrightStarCat-iphas.fits', 1)
+BRIGHTCAT = fits.getdata('lib/BrightStarCat-iphas.fits', 1)
 
 EXTS = [1, 2, 3, 4]  # Which extensions to expect in the fits catalogues?
 PXSCALE = 0.333  # Arcsec/pix of the CCD

### 2013-03-14 scripts/create-iphasDetection.py
Initial commit

@@ -695,7 +695,7 @@ def run_all(ncores=4):
 
 if __name__ == '__main__':
     logging.basicConfig(level=logging.ERROR)
-    run_all(6)
+    run_all(8)
 
     #run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/iphas_jul2007/r571033_cat.fits')
     #run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/iphas_jun2005/r455464_cat.fits')

### 2013-03-15 scripts/create-iphasDetection.py
WCS problem fix

@@ -18,7 +18,8 @@ TODO
 """
 
 from astropy.io import fits
-from astropy import wcs
+#from astropy import wcs
+from astLib import astWCS
 import numpy as np
 import logging
 import os
@@ -159,7 +160,7 @@ class CatalogueConverter():
                 self.fits[ccd].header['PV2_1'] = self.hdr('PROJP1', ccd)
             if self.hdr('PV2_3', ccd) != self.hdr('PROJP3', ccd):
                 self.fits[ccd].header['PV2_3'] = self.hdr('PROJP3', ccd)
-
+        
         # Some runs do not have date/time stored due to a glitch in the
         # Telescope Control System
         for ccd in EXTS:
@@ -328,13 +329,24 @@ class CatalogueConverter():
         dec = np.array([])
         for ccd in EXTS:
             # Bug? wcslib complains because the units are given as pixels!
-            self.fits[ccd].header['CUNIT1'] = 'deg'
-            self.fits[ccd].header['CUNIT2'] = 'deg'
+            #self.fits[ccd].header['CUNIT1'] = 'deg'
+            #self.fits[ccd].header['CUNIT2'] = 'deg'
+
+            """
             mywcs = wcs.WCS(self.fits[ccd].header, relax=True)
             myra, mydec = mywcs.wcs_pix2world(
                             self.fits[ccd].data.field('X_coordinate'),
                             self.fits[ccd].data.field('Y_coordinate'),
                             1)
+            """
+
+            mywcs = astWCS.WCS(self.fits[ccd].header, mode='pyfits')
+            radec = np.array(mywcs.pix2wcs(
+                                self.fits[ccd].data.field('X_coordinate'),
+                                self.fits[ccd].data.field('Y_coordinate')))
+            myra = radec[:, 0]
+            mydec = radec[:, 1]
+
             ra = np.concatenate((ra, myra))
             dec = np.concatenate((dec, mydec))
         return (ra, dec)
@@ -357,10 +369,12 @@ class CatalogueConverter():
 
     def get_csv_summary(self):
         """ Returns a CSV-formatted summary line """
+        # Average seeing and ellipticity across CCDs
[…]

### 2013-03-15 scripts/create-iphasDetection.py
Corrected seeing unit

@@ -18,7 +18,7 @@ TODO
 """
 
 from astropy.io import fits
-#from astropy import wcs
+#from astropy import wcs  # astropy.wcs doesn't like IPHAS catalogues!
 from astLib import astWCS
 import numpy as np
 import logging
@@ -117,10 +117,10 @@ class CatalogueConverter():
             self.fits = fits.open(self.path)
         except IOError, e:
             raise CatalogueException('IOError: %s' % e)
+        self.validate()  # Raises CatalogueException for dodgy catalogues
         self.cat_path = self.strip_basedir(path)
         self.image_path = self.get_image_path()
         self.conf_path = self.get_conf_path()
-        self.validate()  # Raises CatalogueException for dodgy catalogues
 
     def hdr(self, field, ext=1):
         """Return the value of a header field from extension `ext`."""
@@ -160,7 +160,7 @@ class CatalogueConverter():
                 self.fits[ccd].header['PV2_1'] = self.hdr('PROJP1', ccd)
             if self.hdr('PV2_3', ccd) != self.hdr('PROJP3', ccd):
                 self.fits[ccd].header['PV2_3'] = self.hdr('PROJP3', ccd)
-        
+
         # Some runs do not have date/time stored due to a glitch in the
         # Telescope Control System
         for ccd in EXTS:
@@ -168,6 +168,8 @@ class CatalogueConverter():
                 raise CatalogueException('UTSTART keyword missing')
             if not 'DATE-OBS' in self.fits[ccd].header:
                 raise CatalogueException('DATE-OBS keyword missing')
+            if not 'MJD-OBS' in self.fits[ccd].header:
+                raise CatalogueException('MJD-OBS keyword missing')
 
     def get_image_path(self):
         candidate = os.path.join(self.directory,
@@ -589,7 +591,7 @@ class CatalogueConverter():
         col_mjd = fits.Column(name='mjd', format='D', unit='Julian days',
                               array=mjd)
 
-        seeing = np.concatenate([[self.hdr('SEEING', ccd)]
+        seeing = np.concatenate([[PXSCALE*self.hdr('SEEING', ccd)]
                                  * self.fits[ccd].data.size
                                  for ccd in EXTS])
[…]

### 2013-03-17 scripts/create-iphasDetection.py
-

@@ -714,15 +714,8 @@ if __name__ == '__main__':
     #run_all(4)
 
     #Testcases:
-    #run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/iphas_jul2007/r571033_cat.fits')
-    #run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/iphas_jun2005/r455464_cat.fits')
-    #run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/uvex_sep2010/r755575_cat.fits')
-    #run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/iphas_dec2004/r434839_cat.fits')
-    #run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/iphas_dec2005/r486966_cat.fits')
     #run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/iphas_nov2003b/r375399_cat.fits')
+    #run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/iphas_nov2003b/r375400_cat.fits')
+    #run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/iphas_nov2003b/r375401_cat.fits')
 
-    run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/iphas_nov2003b/r375399_cat.fits')
-    run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/iphas_nov2003b/r375400_cat.fits')
-    run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/iphas_nov2003b/r375401_cat.fits')
-
-    #run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/iphas_jun2005/r455355_cat.fits')
+    run_one('/media/0133d764-0bfe-4007-a9cc-a7b1f61c4d1d/iphas/iphas_aug2004a/r414199_cat.fits')

### 2014-03-11 paper/iphas-dr2.tex
Initial commit of paper

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{amssymb,amsmath}
\usepackage[pdftex]{graphicx}
\usepackage{longtable} 
%\usepackage{lscape}
\usepackage{float}
\usepackage{booktabs}
\usepackage{aas_macros}

\def\ha{\mbox{H$\rm \alpha$}}
\def\arcsec{$''$}
\def\arcmin{$'$}
\def\deg{$^{\circ}$}
\def\micron{\mbox{$\mu$m}}

\title[IPHAS Data Release 2]{The Second Data Release 
of the INT Photometric H$\alpha$ Survey 
of the Northern Galactic Plane (IPHAS DR2)}

\author[G. Barentsen
et. al]{G. Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
H. J. Farnhill$^{1}$,
J. E. Drew$^{1}$,
M. J. Irwin$^{4}$, \newauthor
E. A. Gonz$\acute{\rm{a}}$lez-Solares$^{4}$,
R. Greimel$^{2}$,
B. Mizalski$^{3}$,
C. Ruhland$^{1}$, \newauthor plus many friends.
\newauthor\\
$^{1}$Centre for Astrophysics Research, STRI, University of Hertfordshire, College Lane Campus, Hatfield AL10 9AB\\
$^{2}$Institute for Geophysics, Astrophysics and Meteorology, Institute of Physics, Karl-Franzens-Universit$\ddot{a}$t Graz, Universit$\ddot{a}$tsplatz 5, 8010 Graz, Austria\\
$^{3}$Building, Institute, Street Address, City, Code, Country\\
$^{4}$Institute of Astronomy, Madingley Road, Cambridge CB3 0HA}

\begin{document}
\date{Current draft typeset \today}
\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2013}

\maketitle

\label{firstpage}

\begin{abstract} % BACKGROUND, OBJECTIVE, METHODS, RESULTS, CONCLUSIONS
The INT/WFC Photometric H$\alpha$ Survey 
of the Northern Galactic Plane (IPHAS)
is a 1800 deg$^2$ imaging survey
covering the entire northern Milky Way at $|b| < 5^\circ$
in the $r$, $i$ and \ha\ filters 
using the Wide Field Camera (WFC) 
on the 2.5-meter Isaac Newton Telescope (INT).
This data release presents the first 
uniformly-calibrated source catalogue
to have been extracted from the survey,
providing single-epoch photometry
for 219 million unique sources
across 92\% of the survey.
The observations were carried out between 2003 and 2012
at a median seeing of 1.1 arcsec
to a depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$ and \ha$=20.3\pm 0.3$
($5\sigma$ limits, Vega system).
We explain the data reduction 
and quality control procedures,
describe and test the new uniform photometric calibration,
and detail the construction of the source catalogue
and its quality warning flags.
We find that the new calibration is accurate to
$\sigma=0.03$ mag.
Finally, we demonstrate the ability of the 
catalogue's unique
$r$-$i$/$r$-\ha\ colour-colour diagram to
(i) characterise stellar populations and extinction regimes
towards different Galactic sightlines
and (ii) select reliable candidate \ha-emission line objects.
The catalogue which accompanies this paper
provides the much-needed visible-light counterpart
to several infrared surveys of the Galactic Plane,
including many poorly-studied star-forming regions and ``bubbles'',
and provides images for some of the most
crowded regions to be faced by Gaia.
\end{abstract}

\begin{keywords}
catalogues, surveys, stars: emission line, Be, Galaxy: stellar content
\end{keywords}

\section{Introduction}

Since the first data release in 2008, 
the INT/WFC Photometric H$\alpha$ Survey 
of the Northern Galactic Plane (IPHAS)
has provided new insights into the contents and structure 
of our own backyard, the Milky Way. 
The original motivation for undertaking 
this large-scale programme of observation
-- spanning almost a decade 
and using more than 300 nights 
at the Isaac Newton Telescope (INT) in La Palma -- 
was to provide the digital update 
to the photographic northern H$\alpha$ surveys 
of the mid-20th century. 
By increasing the sensitivity 
with respect to these previous surveys 
by a factor $\sim$1000 (7 magnitudes), 
it was envisaged that IPHAS would allow 
the limited bright samples of Galactic emission line objects 
available at the outset \citep[e.g.][]{Kohoutek1999}, 
to be extended into larger, deeper, more statistically-robust samples 
that in turn could better inform our understanding 
of the early and late stages of stellar evolution. 
This aim has begun to be realised through a
range of published studies including: 
a preliminary catalogue of candidate emission line objects \citep{Witham2008};
discoveries of new northern symbiotic stars \citep{Corradi2008, Corradi2010}; 
new cataclysmic variables \citep{Witham2007}; 
new groups of young stellar objects \citep{Vink2008,Barentsen2011a};
along with discoveries of new and remarkable planetary nebulae 
\citep{Mampaso2006, Corradi2011, Viironen2011}.

Over the years it has become apparent that the legacy of IPHAS 
will extend beyond the traditional \ha\ applications 
of identifying emission line stars and nebulae. 
Through the provision of $r$, $i$ broadband photometry 
alongside H$\alpha$ data,
IPHAS has created the opportunity 
to study Galactic Plane populations 
in a new way.
For example, the survey’s unique $r-H\alpha$ colour, 
when combined with $r-i$,
has been shown to provide simultaneous constraints 
on intrinsic stellar colour and interstellar extinction \citep{Drew2008}. 
This has opened the door 
to a wide range of Galactic science applications, 
including the mapping of extinction across the Plane in three dimensions
and the probabilistic inference of stellar properties
\citep{Sale2009, Sale2010, Giammanco2011, Sale2012, Barentsen2013}. 
In effect, the availability of narrowband H$\alpha$ 
alongside $r, i$ magnitudes
provides coarse spectral information for huge samples of stars 
which are otherwise too faint or numerous 
to be targeted by spectroscopic surveys.
For such science applications to succeed however, 
it is vital that the imaging data is transformed 
into a homogeneously calibrated photometric catalogue, 
in which quality problems 
and duplicate detections of the same source 
are flagged.

The first release of IPHAS data, 
covering roughly half the survey footprint,
was made in late 2007 \citep{Gonzalez-Solares2008}. 
At the time the data were insufficiently complete 
to support a homogeneously calibrated source catalogue.
The goal of this paper is to present the next release 
that takes the coverage up to over 90 percent of the survey area 
and includes a uniform calibration.
In this work we aim to
(i) explain the data reduction 
and quality control procedures that were applied,
(ii) describe and test the new global photometric calibration, and 
(iii) detail the construction and demonstrate the use 
of the source catalogue that has been extracted 
from the newly re-calibrated data.

In \S\ref{sec:observations} we start by recapitulating the key points
of the survey design and strategy.
In \S\ref{sec:reduction} we describe the data reduction
and quality control procedures.
In \S\ref{sec:calibration} we explain the uniform re-calibration
and test our results against the Sloan Digital Sky Survey (SDSS).
In \S\ref{sec:catalogue} we explain how the source catalogue was compiled.
In \S\ref{sec:discussion} we discuss the properties of the catalogue,
and finally in \S\ref{sec:demonstration} we demonstrate
the health of the release by demonstrating its scientific exploitation.
In \S\ref{sec:conclusions} we conclude and outline
our future ambitions.

\section{Observations}
\label{sec:observations}

The detailed properties of the IPHAS observing programme 
have been presented before 
by \citet{Drew2005} and \citet{Gonzalez-Solares2008}. 
To set the stage for this release, we briefly remind of some key points.
IPHAS is a 1800~sq. deg. imaging survey of the northern Galactic Plane, 
providing images and photometry in Sloan $r, i$ 
along with narrowband H$\alpha$. 
It is carried out using the Wide Field Camera (WFC) 
on the 2.5-meter Isaac Newton Telescope (INT) in La Palma. 
It is the first digital survey to offer comprehensive CCD photometry
of point sources in the Galactic Plane at visible wavelengths, 
and does so down to a limiting magnitude of $\sim$20th.
The IPHAS footprint on the northern sky spans a box 
of roughly 180 by 10 degrees, 
taking in the entire northern Galactic Plane 
at latitudes $-5^{\circ} < b < +5^{\circ}$ 
and longitudes $30^{\circ} < l < 215^{\circ}$.

The Wide Field Camera is a mosaic of 4 CCDs 
that captures a sky area of close to 0.29 square degrees.
To cover the entire northern Plane with some overlap,
the survey area was divided into 7635 telescope pointings.
Each of these pointings is accompanied by an offset position
at a displacement of $+$5 arcmin in declination 
and $+$5 arcmin in right ascension,
to deal with inter-CCD gaps, detector imperfections,
and to enable quality checks. 
The basic unit of observation hence
amounts to $2 \times 3$ exposures, 
in which each of the 3 survey filters is exposed at 2 offset sky positions, 
within an elapsed time of 10 minutes.
We shall refer to the unit of 3 exposures at the same position 
as a \emph{field},
and the combination of two fields at a small offset as a \emph{field pair}.
The survey hence contains 15270 fields grouped into 7635 field pairs.
To achieve the desired survey depth of 20th magnitude or fainter, 
the filter exposure times were set at 120 sec (narrowband \ha), 
30 sec ($r$) and 10 sec ($i$)
in the majority of the survey observations.\footnote{In 2003 
the $r$-band exposure time was 10~sec instead of 30~sec,
and since Oct 2010 the $i$-band exposure time 
has been increased from 10~sec to 20~sec.}

Data-taking began in the second half of 2003, 
and every field had been observed at least once by the end of 2008.
At that time only 76 percent of the field pairs 
satisfied our minimum quality criteria however,  
often due to the effects of clouds, poor seeing, or technical faults
(the quality criteria will be detailed in the next section). 
Since then, a programme of repeat observations has been in place 
to improve data quality. 
As a result, 92\% of the survey 
now benefits from quality-approved data.
The most recent observations which are part of this release
were obtained in November 2012.

\begin{figure*}
        \includegraphics[width=1\linewidth]{./plots/footprint/footprint_small.png}
        \caption{Survey area showing the footprints
        of all the quality-approved IPHAS fields
        which have been included in this data release.
        The area covered by each field has been coloured black
        with a semi-transparent opacity of 20\%,
        such that regions where fields overlap are darker.
        The IPHAS strategy is to observe each field twice
        with a small offset,
        and hence the vast majority of the area 
        is covered twice (dominant gray colour).
        There are small overlaps between all the neighbouring fields
        which can be seen as a honeycomb-style pattern
        of dark gray lines across the survey area.
        Regions with incomplete data are apparent as white gaps (no data) 
        or as the lightest shade of gray
        (denoting that only the offset position is missing).}
        \label{fig:footprint}
\end{figure*}

Figure~\ref{fig:footprint} shows the footprint 
of the quality-approved observations included in this work. 
The fields which remain missing 
-- covering 7 percent of the survey area --
are predominantly located towards the Galactic anti-center 
at $l > 120^o$.
Fields at these longitudes can only be accessed from La Palma 
in the months of November-December,
which is when the weather and seeing conditions are often poor
at the INT and observing attempts have failed repeatedly.
To enable the survey to be brought to completion, 
a decision was made recently to limit repeats in this area 
to individual fields requiring replacement,
i.e. fresh observations in all 3 filters may only be obtained 
at one of the two offset positions, 
if the data for the partner offset has already passed quality control.  
The catalogue is structured such that it is clear 
where a contemporaneous observation of both halves of a field pair
is not available.

\section{Data reduction and quality control}
\label{sec:reduction}

\subsection{Initial pipeline processing}

All raw data obtained with the INT were transferred
to the Cambridge Astronomical Survey Unit (CASU) 
for initial processing and archival.
The procedures used by CASU were originally devised
for the INT Wide Field imaging Survey \citep[WFS;][]{McMahon2001,Irwin2005},
which was a 200 deg$^2$ survey programme carried out 
between 1998 and 2003 after the WFC was commissioned.
Because IPHAS uses the same telescope and camera combination,
we have been able to benefit from the existing WFS pipeline.
A detailed description of the processing steps 
is found in \citet{Irwin2001}.
Its application to IPHAS has previously been described
by \citet{Drew2005} and \citet{Gonzalez-Solares2008}
and much of the source code is available 
on line\footnote{http://casu.ast.cam.ac.uk/surveys-projects/software-release}. 
In brief, the pipeline takes care of bias subtraction,
linearity correction, flat-fielding,
gain correction and de-fringing.

The reduced images are then stored in a multi-extension FITS file 
with a primary header describing the characteristics
(position, filter, exposure time, etc.) 
and four compressed image extensions 
corresponding to each of the four CCDs.
Source detection and characterisation is then carried out 
using the \textsc{imcore} tool \citep{Irwin1985,Irwin1997}.
The flux of each source is measured using both
the peak pixel height (i.e. a square 0.33$\times$0.33\arcsec\ aperture)
as well as a series of circular apertures of increasing diameter 
(1.2\arcsec, 2.3\arcsec, 3.3\arcsec, 4.6\arcsec\ and 6.6\arcsec).

The local background levels are estimated 
by computing the sigma-clipped median
flux in a grid of 64$\times$64 pixels (21$\times$21\arcsec)
across the image,
which is then interpolated to obtain an estimate 
of the background level at each pixel.
These sky levels are subtracted from the aperture photometry and
-- when required --
a deblending routine is applied which also attempts to remove
the contamination from any other nearby sources.
Whilst this approach works very well 
across the vast majority of the survey area,
the Galactic Plane unavoidably contains crowded regions 
with large numbers of overlapping sources
or rapidly spatially-varying nebulosity,
in which case aperture photometry must always be interpreted 
with caution.
In \S\ref{sec:catalogue} we will explain that overlapping
sources to which the deblending routine was applied 
are flagged in the catalogue using the \emph{deblend} warning flag.

Finally, an astrometric solution is determined
based on the 2MASS point source catalog \citep{Skrutskie2006},
which itself is calibrated 
in the International Celestial Reference System (ICRS).
A provisional photometric calibration is also provided 
based on the average zeropoint
determined from a set of standard stars observed in the same night.
Sources are classified morphologically
-- stellar, galaxy or noise --
based on the curve-of-growth determined
from measuring the source intensity in a series of growing apertures.
Finally, the resulting source detection tables are also stored 
in multi-extension FITS files.

At the time of preparing DR2,
the CASU pipeline had processed
74\,195 IPHAS exposures 
in which a total of 1.9~billion \emph{candidate sources} were detected 
at the sensitive default detection level of 1.25\,$\sigma$
-- unavoidably including spurious detections, artefacts and
duplicate detections 
(in \S\ref{sec:catalogue} we will explain
how these have been removed or flagged).
The pipelined data set -- comprising 2.5~terabyte of FITS files --
was then transferred to the University of Hertfordshire
for the purpose of transforming the raw
detection tables into a reliable source catalogue which is 
(i) quality-controlled,
(ii) homogeneously calibrated, and 
(iii) contains user-friendly columns and warning flags.
It is these post-processing steps which are explained below.

\subsection{Quality control}
\label{sec:qc}

Observing time for IPHAS was obtained
on a semester-by-semester basis
through the traditional time allocation committees 
of the Isaac Newton Group of telescopes,
which are competitive and invariably over-subscribed.
For this reason, we attempted to utilise 
\emph{all} the nights allocated to IPHAS,
even those which were partially or entirely non-photometric
or otherwise affected by technical problems 
(e.g. electronic noise or telescope tracking problems).
Any unsuitable data that was taken as a result
of this strategy was subsequently flagged and rejected
using a series of seven quality criteria,
which ensure a reliable and homogeneous level of quality
across the data release:

\begin{figure}
    \begin{minipage}[b]{\linewidth}
        \includegraphics[width=\textwidth]{./plots/depth_r.pdf} 
    \end{minipage}
    \begin{minipage}[b]{\linewidth}
        \includegraphics[width=\textwidth]{./plots/depth_i.pdf} 
    \end{minipage}
    \begin{minipage}[b]{\linewidth}
        \includegraphics[width=\textwidth]{./plots/depth_h.pdf} 
    \end{minipage}
    \caption{Distribution of the 5$\sigma$ limiting magnitude
             across all quality-approved survey fields
             for $r$ (top), $i$ (middle) and \ha\ (bottom).
             Fields with a limiting magnitude brighter than
             20th ($r$) or 19th (\ha/$i$) were rejected
             from the data release.
             The $r$-band depth is most sensitive 
             to the presence of the moon above the horizon, 
             which is evidenced by the wide and bi-model shape
             of its distribution.}
    \label{fig:depth}
\end{figure}

(1) \emph{Depth.} 
We discarded any exposures for which the $5\sigma$ limiting magnitude 
was worse than 20th magnitude in the $r$-band
or worse than 19th in $i$ or \ha. 
Such data were typically obtained during poor weather or full moon.
Most observations fared significantly better than these limits.
Figure~\ref{fig:depth} presents the distribution of limiting magnitudes
for all quality-approved fields,
which shows a mean depth of 
$21.2\pm0.5$ ($r$), $20.0\pm0.3$ ($i$) and $20.3\pm0.3$ (\ha).
We found that the depth achieved depended 
most strongly on the presence of the moon,
which was above the horizon during 62 percent 
of our observations and explains the wide and bi-modal shape
of the $r$-band limiting magnitude distribution 
(top panel in Fig.~\ref{fig:depth}).
In contrast, the depth attained in $i$ and \ha\ 
is less sensitive to moonlight
and the distribution of their depths
is hence more narrow
(middle and bottom panel in Fig.~\ref{fig:depth}).

Our minimum limiting magnitude criteria
have led us to exclude 9\% of the pipelined data.
We note that much of the excluded data may nevertheless be useful
for e.g. time-domain studies of bright stars.
The detection tables for any such discarded data are made
available through our website (www.iphas.org)
but are ignored in what follows.

(2) \emph{Ellipticity.} 
The ellipticity of a point source,
defined as $e = 1 - b / a$ 
with $b$ the semi-minor and $a$ the semi-major axis,
is a morphological measure of the elongation of the point spread function.
It is expected to be zero (circular) across the field 
in a perfect imaging system,
but it is slightly non-zero in any real telescope data 
due to optical distortions and tracking errors.
The mean ellipticity across a field in the IPHAS data set 
is $0.09\pm0.04$.
There have been sporadic episodes with higher ellipticities however
due to mechanical glitches in the telescope tracking system.
For this reason, we rejected exposures in which the mean ellipticity
across the detector exceeded $e > 0.3$,
which is when the photometric measurements delivered by the pipeline
were found to become degraded.
Only 0.4\% of the exposures were discarded on this basis.

(3) \emph{Seeing.} 
The survey originally aimed to obtain data 
at seeing better than 1.7 arcsec.
This target is currently attained across 86\% of the footprint,
in particular at early longitudes,
e.g. 92\% of the fields at $l<120^\circ$ are better than 1.7 arcsec.
Figure~\ref{fig:seeing} presents the distribution
of the mean seeing for all the quality-approved fields.
We find a median value of 1.1~arcsec in $r$/\ha\
and 1.0~arcsec in $i$.
In the $r$-band, 90\% of the data is better than 1.5~arcsec
and 10\% is better than 0.8~arcsec.
To improve the area covered by this data release,
we have decided to include the small fraction of data
that was obtained under seeing up to 2.5 arcsec,
so that only 1 per cent of the pipelined exposures
had to be excluded.
In \S\ref{sec:catalogue} we will explain
that the catalogue is structured
such that the information on the seeing is included,
and that the photometry listed
is based on the exposures with the best-available seeing.

\begin{figure}
    \begin{minipage}[b]{\linewidth}
        \includegraphics[width=\textwidth]{./plots/seeing_r.pdf} 
    \end{minipage}
    \begin{minipage}[b]{\linewidth}
        \includegraphics[width=\textwidth]{./plots/seeing_i.pdf} 
    \end{minipage}
    \begin{minipage}[b]{\linewidth}
        \includegraphics[width=\textwidth]{./plots/seeing_ha.pdf} 
    \end{minipage}
    \caption{Seeing distribution across all
             quality-approved survey fields
             for $r$ (top), $i$ (middle) and \ha\ (bottom).}
    \label{fig:seeing}
\end{figure}

(4) \emph{Photometric repeatability.} 
The IPHAS field-pair observing strategy 
ensures that every pointing is immediately followed 
by an offset pointing at a displacement of $+$5 arcmin in Dec 
and $+$5 arcmin in RA.
This allows pairs of images to be checked 
for the presence of clouds or electronic noise.
To exploit this information,
the overlap regions of all field pairs were systematically cross-matched
to verify the consistency of the photometric measurements
for stars observed in both pointings.
We rejected field pairs in which more than 2\% of the stars 
showed an inconsistent measurement at the level of 0.2 mag,
or more than 26\% were inconsistent at the level of 0.1 mag.
These limits were determined empirically by inspecting
the images and photometry by eye.
11\% of the data was rejected as part of this step.

(5) \emph{Source density mapping.}
Spatial maps showing the number density of the detected sources
down to 20th magnitude were created to verify the health
of the data and to check for unexpected artefacts.
%(e.g. bright satellite trails)
In particular, we created density maps
which showed the number of \emph{unique} sources
obtained by cross-matching the detection tables of
all three bands with a maximum matching distance of 1 arcsec.
The success of such a cross-matching procedure crucially depends
on the accuracy of the astrometric solution in each band,
and hence we were able to detect and correct
exposures with poor astrometry by inspecting the density map
for spurious density variations.

(6) \emph{Visual examination.}
All images and their associated photometric colour/magnitude diagrams
were inspected by a team of 20 survey team members, 
such that each image in the data release 
was looked at by at least three different pairs of eyes.
Images deemed unsuitable were flagged, investigated and
excluded from the release if necessary. 
6\% of the attempts to observe a field
were placed on a \emph{black-list}
for this reason, most commonly due to the obvious presence
of clouds or extreme levels of scattered moonlight.
Such fields were often rejected as part of 
one or more of the above quality criteria as well.

(7) \emph{Contemporaneous field data.} 
Finally, only exposures which are part of a sequence 
of three consecutive images of the same field (H$\alpha$/$r$/$i$) 
were considered for inclusion in the release. 
This ensures that the three bands for a given field
are observed at nearly the same time --  
essentially always within 5 minutes.
An exception was made for 9 fields where the three exposures 
could not be obtained within the same night
but for which the time gap did not exceed 48 hours.
We note that the exact epoch of the magnitude in each band
is included in the source catalogue
(columns \emph{rMJD}, \emph{iMJD}, \emph{haMJD}).

The above criteria were satisfied 
for 14115 out of the 15270 fields (92\%).
In a few cases more than one successful attempt to observe
a field was available due to the fact that slightly stricter
quality criteria were adopted in the initial years of the survey.
In such cases, only the attempt 
with the best seeing and depth was selected
for inclusion in the release, because the focus 
of this work is to deliver the most reliable
measurement at a single epoch.
Those interested in any of the rejected data 
may nevertheless access the full set of detection tables 
on line.

\section{Photometric calibration}
\label{sec:calibration}

Having obtained a quality-approved set of observations,
we now turn to the problem of placing the data
onto a uniform photometric scale.

\subsection{Provisional nightly calibration}

For the purpose of providing an initial calibration 
of the $r$ and $i$ broadband fluxes,
photometric standard fields were observed every night.
The standards were chosen from a list based on 
the \cite{Landolt1992} and Stetson (http://cadcwww.dao.nrc.ca/standards) 
objects.
Two or three standard fields were observed 
during the evening and morning twilight,
and at intervals of 2-3 hours throughout the night.
The CASU pipeline automatically identified the observed standards 
and used them to determine a sigma-clipped average zeropoint \textsc{magzpt}
for each night and filter,
such that the number counts $DN$ 
in the pipeline-corrected CCD frames
relates to a magnitude $m$ as:
\begin{equation}
\begin{split}
   m  = & \textsc{magzpt} - 2.5 \log_{10}( DN / \textsc{exptime} ) \\
 &  - \textsc{extinct}\cdot(\textsc{airmass}-1) - \textsc{apcor} - \textsc{percorr}
\label{eqn:mag}
\end{split}
\end{equation}
where \textsc{exptime} is the exposure time in seconds,
\textsc{extinct} is the atmospheric extinction coefficient 
(typically 0.09 for $r$ and 0.05 for $i$ in La Palma),
\textsc{airmass} is the normalised optical path length 
through the atmosphere and
\textsc{apcor} is a correction for the flux
lost outside of the aperture used.
Finally, \textsc{percorr} is a correction based on the difference
between the median dark sky for a CCD against the median for all the CCDs 
and as such is an ancillary correction 
to account for sporadic gain variations. 
All these quantities correspond to header keywords in the 
multi-extension FITS files produced by the CASU pipeline.

The zeropoint was determined such that the resulting magnitude system
is in the WFC system that uses the SED of Vega 
as the zero colour reference object. 
Colour equations were used to transform between the IPHAS passbands 
and the Johnson-Cousins system 
of the published standard star photometry.
The entire procedure has been found to deliver zeropoints which 
are accurate at the level of 1-2\% 
in photometric conditions \citep{Gonzalez-Solares2011}.

Unlike the broadbands, 
standard star photometry is not available in the literature 
for the H$\alpha$ passband
and hence there is no formally recognised flux scale 
for the narrowband.
We can specify here however 
that the integrated in-band energy flux for Vega 
in the IPHAS \ha\ filter 
is $1.52 \times 10^{-7}$ ergs\,cm$^{-2}$\,s$^{-1}$ 
at the top of the Earth's atmosphere,
which is the flux obtained by folding 
Vega's SED with the filter throughput curve 
corrected for atmosphere and detector quantum efficiency
\citep[using the method explained by][]{Drew2005}.
This is 3.08 magnitudes less than the flux captured 
by the much broader $r$ band
which includes the \ha\ band.
Hence to assure zero colour relative to the broadbands,
we set the default zeropoint for the narrowband to be:
\begin{equation}
\textsc{magzpt}_{H\alpha} = \textsc{magzpt}_r - 3.08.
\label{eqn:zpha}
\end{equation}

\subsection{Uniform re-calibration}

Despite the best efforts made to obtain a nightly calibration,
large surveys naturally possess field-to-field variations
at the level of 0.1 mag
due to atmospheric fluctuations during the night
and imperfections in the pipeline or the instrument
(e.g. the WFC is known to suffer from sporadic glitches
in the timing of exposures).
Such variations need to be corrected for 
during a global re-calibration procedure.
Notable past examples include the global re-calibration 
of the Two Micron All Sky Survey \citep[2MASS;][]{Nikolaev2000},
the Sloan Digital Sky Survey \citep[SDSS;][]{Padmanabhan2008}
and the Panoramic Survey Telescope 
and Rapid Response System survey \citep[Pan-STARRS;][]{Schlafly2012},
which all achieved photometry 
that is globally consistent to within 1--2\%.

Surveys which observe identical stars at different epochs
can use the repeat measurements to ensure a uniform calibration.
For example, 2MASS attained its global calibration
by observing six standard fields each hour, 
allowing zeropoint variations to be tracked 
over very short timescales \citep{Nikolaev2000}.
Alternatively, the SDSS and PanSTARRS surveys could benefit
from revisiting regions in their footprint to 
carry out a so-called \emph{ubercalibration} procedure,
in which multiple measurements of stars at different epochs
are used to fit the calibration parameters
\citep{Padmanabhan2008,Schlafly2012}.

Unfortunately these schemes cannot be applied
directly to IPHAS
for two reasons. 
Firstly, the survey was carried out 
in competitive observing time
on a non-dedicated telescope, 
rendering the 2MASS approach 
of observing standards at a very high frequency
prohibitively expensive
-- in part because standard fields 
are very scarce within the Galactic Plane.
Secondly, the aim of IPHAS is to obtain magnitudes at a single epoch
and hence stars are not normally observed at more than one epoch,
unless they happen to fall on the narrow overlap region 
between two neighbouring fields.

Although the IPHAS data does contain a significant number of 
inter-field repeat measurements
due to the small overlaps
between neighbouring fields,
we have found the information contained
in these regions to be insufficient
to constrain the calibration parameters.
This is because photometry at the extreme edges of the WFC
-- where neighbouring fields overlap -- 
is itself prone to systematics at the level of 1-2\%.
The cause of these errors is thought to include 
the use of twilight sky flats in the pipeline,
which are known to be imperfect for calibrating stellar photometry 
due to stray light and vignetting \citep[e.g.][]{Manfroid1995}.
Moreover, the illumination correction in the overlap regions
is affected by a radial geometric distortion,
which causes the pixel scale towards the edges 
to increase \citep{Gonzalez-Solares2011}.
Although these systematics are reasonably small within a single field,
they can combine to cause artificial zeropoint gradients 
across the survey
when they are used to constrain a global calibration
without external constraints.

For these reasons, we have decided not to depend
on an ubercalibration-type scheme alone,
and have instead opted to exploit an external reference survey
-- where available --
to bring the majority of our data onto a uniform calibration.
This is explained next.

\subsubsection{Correcting zeropoints using APASS}

We have been able to benefit from the
AAVSO Photometric All-Sky Survey
(APASS; http://www.aavso.org/apass)
to bring the vast majority of the survey 
onto a uniform scale.
Since 2009,
APASS has been using two 20~cm-astrographs
to survey the entire sky down to $\sim$17th magnitude
in five filters which include Sloan $r$ and $i$.
The most recent catalogue available 
at the time of preparing this work was APASS DR7,
which provided data across roughly half of the IPHAS footprint (Fig.~\ref{fig:apass_r}).
The photometric accuracy of APASS is currently estimated 
to be at the level of 3\% (Henden, private communication),
which is significantly better 
than the provisional calibration of IPHAS
for which we estimate the $1\sigma$-error to be $\sim$10\%.
APASS achieves its uniform accuracy 
by measuring each star at least two times in photometric conditions
-- along with ample standard fields --
using the large $3\times3$ square degrees field of view of its detectors.

With the aim of bringing IPHAS to a similar accuracy of $\sim$3\%,
we used the APASS catalogue to identify and adjust all IPHAS fields 
which showed a magnitude offset larger than 3\% against APASS.
For this purpose,
the $r$- and $i$-band detection tables of each IPHAS field
were cross-matched against the APASS DR7 catalogue 
using a maximum matching distance of 1~arcsec.
The magnitude range was limited to
$13<r_{\rm APASS}<16.5$ and $12.5<i_{\rm APASS}<16.0$
in order to avoid sources 
brighter than the IPHAS saturation limit on one hand, 
and to avoid sources near the faint detection limit of APASS 
on the other hand.

The resulting set of $\sim$220\,000 cross-matched stars were then used 
to derive APASS-to-IPHAS magnitude transformations
using a linear least-squares fitting routine, 
which iteratively removed $3\sigma$-outliers to improve the fit.
The solution converged to:
\begin{align} 
r_{\rm IPHAS} = r_{\rm APASS} - 0.121 + 0.032(r-i)_{\rm APASS} \label{eqn:apass_r} \\
i_{\rm IPHAS} = i_{\rm APASS} - 0.364 + 0.006(r-i)_{\rm APASS} \label{eqn:apass_i}
\end{align}
The root mean square (rms) residuals of these transformations 
are 0.041 and 0.051, respectively.
The small colour terms in the equations
indicate that the $r$ and $i$ filters 
are very similar in both surveys.
The transformations include a large fixed offset
which is simply due to the fact that 
APASS magnitudes are given in the AB system
and IPHAS uses magnitudes relative to Vega.
Separate transformations were derived for sightlines 
with varying extinction properties to investigate the robustness
of the transformations with respect to different reddening regimes,
but the variations at these different
sightlines were found to be insignificant.
This is not surprising because heavily reddened objects 
are naturally scarce at $r<16$.

Having transformed APASS magnitudes into the IPHAS system,
we then computed the median magnitude offset 
for each field which contained at least 30 cross-matched stars.
This was the case for 48\% of the fields.
The mean offset was found to be
$0.014\pm0.104$ for $r$ and $0.007\pm0.108$ for $i$
(Table~\ref{tbl:offsets_before}).
A total of 4596 fields showed a median offset
exceeding $\pm$0.03~mag in either $r$ or $i$.

We then applied the most important step in our re-calibration scheme,
which is to adjust the provisional zeropoints of the 4596 aberrant fields
such that their offset is brought to zero.
This allowed the mean IPHAS-to-APASS offset 
to be brought down to $0.000\pm0.011$ in both $r$ and $i$
(Table~\ref{tbl:offsets_after}).
The procedure of fitting magnitude transformations and
correcting the zeropoints was repeated a few times to ensure 
convergence, which was essentially reached after the first iteration.

\begin{figure*}
    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_rshift.pdf} 
    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_r.pdf} 
    \caption{Median magnitude offset in the $r$ band between IPHAS and APASS,
             plotted on a field-by-field basis
             prior to the re-calibration procedure.
             Each square represents an IPHAS field
             which contains at least 30 stars with a counterpart
             in the APASS DR7 catalogue.
             The colours denote the median
             IPHAS-APASS magnitude offset in each field,
             which was computed after applying the APASS-to-IPHAS
             transformation to the APASS magnitudes (Eqn.~\ref{eqn:apass_r}).}
        \label{fig:apass_r}
    \vspace{1cm}
    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_rshift.pdf}
    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_r.pdf} 
    \caption{Median magnitude offset in the $r$ band
             between IPHAS and SDSS after the re-calibration
             procedure was applied.
             Again, each square represents an IPHAS field
             which contains at least 30 stars
             with a counterpart in the SDSS catalogue.
             The colours denote the median IPHAS-SDSS magnitude offset
             in each field,
             which was computed after applying the SDSS-to-IPHAS
             transformation to the SDSS magnitudes (Eqn.~\ref{eqn:sdss_r}).}
    \label{fig:sdss_r}
\end{figure*}

\begin{figure*}
	\vspace{2cm}
    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_ishift.pdf} 
    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_i.pdf} 
    \caption{Same as Figure~\ref{fig:apass_r} for the $i$-band.}
        \label{fig:apass_i}
    \vspace{1cm}
    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_ishift.pdf}
    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_i.pdf} 
    \caption{Same as Figure~\ref{fig:sdss_i} for the $i$-band.}
    \label{fig:sdss_i}
\end{figure*}

\begin{table}
        \begin{center}
                \begin{tabular}{lcc}
                        \toprule
                         {\bf Before re-calibration} & Mean & $\sigma$  \\
                        \midrule
                        $r$ (IPHAS - APASS) & +0.014 & 0.104 \\
                        $i$ (IPHAS - APASS) & +0.007 & 0.108 \\
                        $r$ (IPHAS - SDSS) & +0.016 & 0.088 \\
                        $i$ (IPHAS - SDSS) & +0.010 & 0.089 \\
                        \bottomrule
                \end{tabular}
        \caption{Mean magnitude offsets for objects
                 crossmatched between IPHAS and APASS/SDSS
                 before the uniform re-calibration.
                 Eqns.~\ref{eqn:apass_r}-\ref{eqn:sdss_i} were applied
                 to the APASS/SDSS magnitudes to bring them into the
                 Vega-based IPHAS system prior to computing
                 the offsets.)
                 }
        \label{tbl:offsets_before}

                \begin{tabular}{lcc}
                        \toprule
                         {\bf After re-calibration} & Mean & $\sigma$ \\
                        \midrule
                        $r$ (IPHAS - APASS) & +0.000 & 0.011\\
                        $i$ (IPHAS - APASS) & +0.000 & 0.011 \\
                        $r$ (IPHAS - SDSS)  & -0.001 & 0.029\\
                        $i$ (IPHAS - SDSS) & -0.002 & 0.032 \\
                        \bottomrule
                \end{tabular}
        \caption{Same as Table~\ref{tbl:offsets_before}
                 but computed after the uniform re-calibration
                 was carried out.}
        \label{tbl:offsets_after}
        \end{center}
\end{table}

\subsubsection{Adjusting fields not covered by APASS}

At the time of writing, the APASS catalogue did not provide 
sufficient coverage for 7359 of the fields in our data release.
Fortunately, these fields are pre-dominantly located 
in the early part of the Galactic Plane (Fig.~\ref{fig:apass_r}),
which were typically observed during the summer months
when photometric conditions are more prevalent at the telescope.
These remaining fields have nevertheless 
been brought onto the same uniform scale 
by employing an ubercalibration-style scheme, explained below,
which minimises the magnitude offsets between stars
located in the overlap regions with neighbouring fields.
Although we explained earlier that these overlap regions
are prone to systematics, the use of APASS enabled us to
keep the zeropoints of roughly half of the fields fixed,
which avoids these systematics from combining to introduce
artificial gradients across the survey.
In the next section we will show this to be true by validating
our calibration against SDSS.

A general solution to the problem of minimising the magnitude
differences between overlapping frames has previously been
described by~\citet{Glazebrook1994}.
In brief, there are two fundamental quantities to be
minimised between each pair of overlapping exposures, 
denoted by the indices $i$ and $j$. 
Firstly, the mean magnitude difference between stars in the overlap
region $\Delta_{ij}=\langle m_i-m_j\rangle=-\Delta_{ji}$ is a local
constraint. 
Secondly, to ensure the solution does not stray far 
from the existing calibration, 
the difference in zeropoints 
$\Delta\mathrm{ZP}_{ij}=-\Delta\mathrm{ZP}_{ji}$ 
between each pair of exposures must also be minimised.

Minimisation of these two quantities is a linear least squares problem 
because the magnitude $m$ depends linearly on the ZP (Eqn.~\ref{eqn:mag}).
Hence we can find the ZP shift to be applied to each field 
by minimising the sum:
\begin{equation}
   S = \sum_{i=1}^N \sum_{j=1}^N w_{ij} \theta_{ij} (\Delta_{ij} + a_i - a_j)^2
   \label{eqn:chi2}
\end{equation}
where $i$ denotes the exposure of interest, 
$j$ an overlapping exposure, 
$N$ the number of exposures,
$a_i$ the ZP to solve for 
and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$), 
$w_{ij}$ are weights set to the uncertainty in $\Delta_{ij}$
and $\theta_{ij}$ is an overlap function 
equal to either 1 if exposures $i$ and $j$ overlap or 0 otherwise. 
Solving for $a_i$ is equivalent to solving $\partial
S/\partial a_i=0$ which gives the matrix equation:
\begin{equation}
   \sum_{j=1}^N A_{ij} a_j = b_j
   \label{eqn:matrix}
\end{equation}
where 
\begin{eqnarray}
   A_{ij} &=& \delta_{ij} \sum_{k=1}^N w_{jk}\theta_{jk} - w_{ij} \theta_{ij},\label{eqn:aij}\\
   b_i &=& \sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ji} = -\sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ij}.\label{eqn:bi}
\end{eqnarray}
This prescription is essentially identical to \citet{Glazebrook1994}.
%,although we changed the formulation to ensure that $A_{ij}$ is positive definite.

As explained above, we enforce a strong external constraint
on the solution by keeping the zeropoint fixed 
for the 6756 fields which have been compared
and calibrated against APASS earlier.
We will hereafter refer to these fields as \emph{anchors}.
It is asserted that the zeropoints $a_i$ of these anchor fields 
are known and not solved for,
though they do appear in the vector $b_j$ as constraints.
In addition to the APASS-based anchors, 
we selected 3273 additional anchor fields by hand
to provide additional constraints in regions not covered by APASS.
These extra anchors were deemed to have reliable zeropoints 
based on 
(i) the information contained in the observing logs,
(ii) the stability of the standard star zeropoints during the night, and
(iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
which is located at $\sim$500~meter from the INT.

We then solved Eqn.~\ref{eqn:matrix} for the $r$ and $i$ bands
separately using the least-squares routine 
in Python's {\sc scipy.sparse} module for sparse matrix algebra.
This provided us with corrected zeropoints for the remaining fields,
which were shifted on average by $+0.02\pm0.11$ in $r$ 
and $+0.01\pm0.12$ in $i$ compared to their provisional calibration.

We then turned to the uniform calibration of the \ha\ data.
It is not possible to re-calibrate the narrowband 
in the same way as the broadbands,
because the APASS survey does not offer \ha\ photometry.
We can reasonably assume however,
that the corrections required for $r$ and \ha\ are identical
because the \ha\ zeropoints have been derived directly from the
$r$-band zeropoints during the provisional calibration (Eqn.~\ref{eqn:zpha}).
Moreover, the IPHAS data-taking pattern ensured 
that a field's \ha\ and $r$-band exposures
were taken at essentially the same time, 
separated only by the $\sim$30-second read-out time required by the WFC.
We have hence corrected the \ha\ zeropoints 
by re-using the zeropoint adjustements that were derived for the $r$ band
in the earlier steps.
An exception was made for 3101 fields
for which our quality-control routines revealed
zeropoint variations which exceeded 0.03~mag 
between consecutive fields,
which indicates non-photometric conditions.
For good practice, the \ha\ zeropoints of these fields
were adjusted by solving Eqn.~\ref{eqn:matrix}
rather than linking them directly to the $r$-band shift.

\subsection{Testing the calibration against SDSS}

Having re-calibrated all fields to an expected accuracy of 3\%,
we then used an independent survey to validate the results.
We have been able to exploit SDSS Data Release 9 \citep{Ahn2012}
for this purpose.
SDSS DR9 provides several strips at low
Galactic latitudes,
providing data across 18\% of the fields in our data release.
We cross-matched IPHAS fields against 
stars marked as reliable in the SDSS catalogue\footnote{
We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
to obtain photometry from the SDSS {\sc photoprimary} table 
with criteria {\sc type = star}, {\sc clean = 1} and {\sc score $>$ 0.7}.}
in the same way as we did for APASS earlier,
with the exception of using fainter magnitude ranges of 
$15<r_{\rm SDSS}<18.0$ and $14.5<i_{\rm SDSS}<17.5$.
This provided us with a set of 1.2 million cross-matched stars.

Colour transformations were obtained using a sigma-clipped linear least squares fit:
\begin{eqnarray}
r_{\rm IPHAS} = r_{\rm SDSS} - 0.093 - 0.044(r-i)_{\rm SDSS} \label{eqn:sdss_r} \\
i_{\rm IPHAS} = i_{\rm SDSS} - 0.318 - 0.095(r-i)_{\rm SDSS}. \label{eqn:sdss_i}
\end{eqnarray}
The rms residuals of these transformations are 0.045 and 0.073, respectively.
The equations are similar to the ones
previously determined for APASS,
although the colour terms are slightly larger;
the throughput curve of the SDSS $i$-band filter 
appears to be somewhat more sensitive at longer wavelengths
compared to both the IPHAS and APASS filters.

These global transformations were deemed adequate
for the purpose of validating our uniform calibration in a statistical sense.
Separate equations were derived towards different sightlines
to investigate the effects of varying reddening regimes.
The transformation coefficients were found 
to show some variation towards lowly reddened areas,
which have relatively few numbers
of (intrinsically) red objects at $r-i > 1$ 
which can skew the colour term.
The vast majority of red objects in the global sample
are those in highly reddened areas however,
which agree well with the global transformations
and dominate the statistical appraisal of our calibration.
%The small number of intrinsically red objects in lowly reddened fields
%have no impact on the statistical appraisal of our calibration.

Having transformed SDSS magnitudes into the IPHAS system,
we then computed the median magnitude offset for each IPHAS field
which contained at least 30 objects with a counterpart
in the SDSS catalogue.
This was the case for 2602 fields.
The median offsets for each of these fields
are shown in Figs.~\ref{fig:sdss_r}-\ref{fig:sdss_i}.
The mean offset and standard deviation found 
was $-0.001\pm0.029$ for $r$
and $-0.002\pm0.032$ for $i$ (Table~\ref{tbl:offsets_after}).
In comparison, offsets computed in the identical way
\emph{before} our re-calibration showed means
of $+0.016\pm0.088$ and $+0.010\pm0.089$ (Table~\ref{tbl:offsets_after}).
We conclude that our re-calibration procedure has
been successful in improving the
uniformity of the calibration by a factor three
and has achieved our aim of bringing the
accuracy down to the level of $\sigma=0.03$~mag.

We warn that the SDSS comparison reveals a number of fields with offsets
exceeding 0.05~mag (523 fields) or even 0.1~mag (18 fields).
Such sporadic outliers are consistent with the tails of a Gaussian
with mean $\sim0$ and $\sigma=0.03$.

In future work, we hope to draw upon
the PanSTARRS survey \citep{Schlafly2012}
to further improve the accuracy of our calibration.
At the time of preparing this work data from PanSTARRS
had not been made public yet.

\section{Source catalogue generation}
\label{sec:catalogue}

Having obtained a quality-checked 
and re-calibrated data set, 
we now turn to the problem
of transforming the observations 
into a user-friendly catalogue.
The aim of this catalogue is to detail
the best-available information for each unique source
in a convenient format,
including flags to warn about quality issues 
such as source blending and saturation.
Compiling the catalogue essentially required four steps:
\begin{enumerate}
\item the single-band detection tables 
produced by the CASU pipeline 
were augmented with new columns
and warning flags;
\item the detection tables were merged into multi-band field catalogues;
\item the overlap regions of the field catalogues 
were cross-matched to flag duplicate measurements 
and identify the best detection 
of each unique source; and
\item these primary detections
were compiled into the final source catalogue.
\end{enumerate}
Each of these four steps are now explained.

\subsection{User-friendly columns and warning flags}

As the first step, the detection tables 
were enhanced by creating new columns.
This is necessary because the tables 
generated by the CASU pipeline 
summarise the detections 
in their original CCD units,
e.g. source positions are given in pixel coordinates 
and photometry in number counts.
To transform these measurements into
user-friendly fields,
we have largely adopted the units and naming conventions
which are in use at the 
WFCAM Science Archive \citep[WSA;][]{Hambly2008}
and the VISTA Science Archive \citep[VSA;][]{Cross2012}.
These archives curate the near-infrared data from both
the UKIDSS Galactic Plane Survey \citep[GPS;][]{Lucas2008}
and the 
VISTA Variables in the Via Lactea survey \cite[VVV;][]{Minniti2010}.
Both these surveys provide high-resolution JHK photometry
in the Galactic Plane.
There is a significant degree of overlap
between the footprints of UKIDSS/GPS and IPHAS,
and hence by adopting a similar catalogue format
we hope to encourage scientific applications
which combine both data sets.

A detailed description of each column in our source catalogue
is given in Appendix~\ref{app:columns}.
In the remainder of this section we highlight the main features.

Firstly, we note that each source is uniquely identified by an
IAU-style designation of the form ``IPHAS2\ JHHMMSS.ss+DDMMSS.s''
(cf. column \emph{name} in Appendix~\ref{app:columns}),
where ``IPHAS2'' refers to the present
data release and the remainder of the string
denotes the J2000 coordinates in sexagesimal format.
For convenience, the coordinates
are also included in decimal degrees
(columns \emph{ra} and \emph{dec})
and in the Galactic coordinate system
(columns \emph{l} and \emph{b}).
We have also included an internal object identifier string 
of the form ``\#run-\#ccd-\#detection''
(e.g. ``64738-3-6473''),
which documents the INT exposure number (\#run),
the CCD number (\#ccd),
and the row number in the CASU detection table (\#detection)
-- see columns \emph{rDetectionID},
\emph{iDetectionID}, \emph{haDetectionID}.

Photometry is provided based on the 2.3-arcsec diameter circular aperture
by default (columns \emph{r}, \emph{i}, \emph{ha}).
The choice of this aperture size as the default 
is based on a trade-off between concerns 
about small number statistics and centroiding errors
for small apertures on one hand,
and diminishing signal-to-noise ratios and source confusion
for large apertures on the other hand.
The user is not restricted to this choice because
the catalogue also provides magnitudes
using three alternative aperture sizes:
the peak pixel height 
(columns \emph{rPeakMag}, \emph{iPeakMag}, \emph{haPeakMag}),
a circular 1.2-arcsec-diameter aperture 
(\emph{rAperMag1}, \emph{iAperMag1},
 \emph{haAperMag1}) and
a 3.3-arcsec-diameter aperture 
(\emph{rAperMag3}, \emph{iAperMag3},
 \emph{haAperMag3}).

Each of these magnitude measurements have been
corrected for the flux lost outside of their respective apertures,
using a correction term which is inferred from the
mean shape of the PSF measured locally in the CCD frame.
In the case of a point source,
the four alternative magnitudes are expected
to be consistent with each other
within the photon noise uncertainties
(which are given in columns \emph{rErr}, \emph{rPeakMagErr},
\emph{rAperMag1Err}, \emph{rAperMag3Err}, etc).
When this is not the case,
it is likely that the source is either
an extended object or that it has
been incorrectly measured as a result of
source blending or a rapidly spatially-varying nebulous background.
In \S\ref{sec:qualitycriteria} we will explain that the consistency
of the magnitude measurements in the different apertures
can be used as a criterion for selecting stellar objects
with reliable photometry from the catalogue.

The brightness of each object as a function of increasing
aperture size is also used by the CASU pipeline to provide
a discrete star/galaxy/noise classification flag
(\emph{rClass}, \emph{iClass}, \emph{haClass})
and a continuous stellarness-of-profile statistic
(\emph{rClassStat}, \emph{iClassStat}, \emph{haClassStat}).
For convenience, we have combined
these single-band morphological measures
into band-merged class probabilities and flags
(\emph{pStar}, \emph{pGalaxy}, \emph{pNoise},
\emph{mergedClass}, \emph{mergedClassStat})
using the merging scheme in use at the WSA\footnote{Explained at
http://surveys.roe.ac.uk/wsa/www/gloss\_m.html \#gpssource\_mergedclass
}.

Information on the quality of each detection is included 
in a series of additional columns.
We draw attention to three useful flags
which warn about the likely presence of systematics:
\begin{enumerate}
\item The \emph{saturated} column is used to flag sources
for which the peak pixel height exceeds 55000 counts,
which is typically the case for stars brighter than 12-13th magnitude.
Although the pipeline attempts to extrapolate the brightness of
saturated stars based on the shape of their PSF,
such extrapolation is prone to errors
and we do not recommend their use.
\item The \emph{deblend} column is used to flag sources 
which partially overlap with a nearby neighbour.
Although the pipeline applies a deblending procedure
to such objects, the procedure is currently applied separately
in each band and hence the $r$-$i$ and $r$-\ha\ colours
of such objects are prone to errors.
\item The \emph{brightNeighb} column is used to flag sources which are located
within 5 arcmin from an object brighter than $V=7$ 
according to the Bright Star Catalogue (BSC; Hoffleit et al. 1991), 
or within 10 arcmin if the neighbour is brighter than $V=4$.
Such very bright stars are known to cause systematic errors
and spurious detections as a result of stray light 
and diffraction spikes.
\end{enumerate}
In addition to the above, we also created warning flags for internal bookkeeping.
For example, we flagged detections which fell in the strongly vignetted regions of the focal plane,
which were truncated by CCD edges,
or which were otherwise affected by bad pixels in the detector.
We will explain below that none of such detections 
have been included in the catalogue
-- an alternative detection was available in essentially all these situations
because of the IPHAS field pair strategy --
and hence these internal warning flags do not appear
in the final source catalogue.

Finally, we note that basic information on the observing conditions
is included (\emph{fieldID}, \emph{fieldGrade}, \emph{night}, \emph{seeing}).
A table containing more detailed quality control information,
indexed by \emph{fieldID}, is made available on our website.

\subsection{Band-merging the detection tables}

The second step in compiling the source catalogue
is to merge the contemporaneous trios
of $r$, $i$, \ha\ detection tables
into multi-band field catalogues.
This required a positional matching procedure 
to link sources between the three bands
based on their position on the sky.
We used the \textsc{tmatchn} function 
of the \textsc{stilts} software for this purpose,
which allows rows from multiple tables to be matched \citep{Taylor2006}.
The result of the procedure is a band-merged catalogue
in which each row corresponds to a group of linked $r$/$i$/\ha\ detections
which satisfy a maximum matching distance criterion in a pair-wise sense.
Sources for which no counterpart was identified
are retained in the catalogue as single-band detections.

We employed a maximum matching distance of 1~arcsec,
which was chosen based on a trade-off between 
completeness and reliability.
On one hand, a matching distance larger than 1~arcsec 
was found to allow too many spurious and unrelated sources 
to be linked. 
On the other hand, a value smaller than 1~arcsec 
would pose problems for very faint sources 
with large centroiding errors, 
and would occasionally fail to link detections near CCD corners
where the astrometric solution can 
show systematic errors which exceed 0.5~arcsec.
The position offsets between the $r$ and $i$/\ha\ detections
have been included in the catalogue 
and can hence be tightened by the user if necessary
(columns \emph{iXi}, \emph{iEta}, \emph{haXi}, \emph{haEta}).
We note that also UKIDSS/GPS adopted 
a maximum matching distance of 1 arcsecond 
for similar reasons \citep{Hambly2008}.

The resulting band-merged catalogues were found
to be reliable for the vast majority of fields.
We warn that source blending and confusion is unavoidable
for faint objects in the Galactic Plane however;
in \S\ref{sec:discussion} we will show
that 19\% of the sources in our catalogue
are flagged as blended objects (column \emph{deblend})
and their band-merged data should be treated with care
because they may have fallen victim to source confusion
during the band-merging step.

\subsection{Selecting the primary detections}

We explained earlier that the survey contains repeat observations
of identical sources as a result of overlaps in the data-taking pattern.
Amongst all sources in the reliable magnitude range $13<r<19$,
we find that 65\% were detected twice and 25\% were detected three times or more.
Only 9\% were detected once.
Unsurprisingly, their spatial distribution traces
the inter-CCD gaps and footprint edges.

The principal aim of this data release is to provide 
reliable photometry at a single epoch,
and hence we have decided
to focus on providing the magnitudes
and coordinates using only the \emph{best-available} 
detection of each object -- 
hereafter referred to as the \emph{primary} detection.
Although overlapping fields could have been co-added 
to gain a small improvement in depth, 
we have decided against this for two reasons.
Firstly, combining the information from multiple epochs
would make the photometry of variable stars difficult to interpret.
Secondly, co-adding would cause the image quality to degrade towards the mean,
which is a draw-back for crowded fields.

Anyone interested in the alternative detections of a source
-- hereafter called the \emph{secondary} detections --
can nevertheless obtain this information in two ways.
To begin with, whenever a secondary detection was observed 
within 10 minutes of the primary,
the magnitudes of that secondary detection
have been included in the catalogue
(columns \emph{sourceID2}, \emph{fieldID2}, 
\emph{r2}, \emph{i2}, \emph{ha2},
\emph{rErr2}, \emph{iErr2}, \emph{haErr2}, \emph{errBits2}.
This is the case for 66\% of the sources brighter than $r<20$
due to the IPHAS field pair observing pattern.
In addition, the complete set of detection tables -- one for each exposure -- 
is made accessibly on our website to allow other uses of the data.
A user-friendly catalogue of secondary detections 
has not been compiled at present 
but may be part of a future data release.

The primary detection is defined as the entry in the 
set of band-merged field catalogues which provides 
the most reliable information for a unique source.
Primary detections have been selected using a so-called \emph{seaming} procedure
which has been adapted from the algorithm developed for the WSA\footnote{http://surveys.roe.ac.uk/wsa/dboverview.html\#merge}.
In brief, the first step of the procedure is to identify all the duplicate detections
by cross-matching the overlap regions of all field catalogues,
again using a maximum matching distance of 1\arcsec.
The duplicate detections for each unique source are then ranked according to
to (i) filter coverage, (ii) quality score (column \emph{errBits}),
and (iii) the average seeing of stars in the CCD frame rounded to 0.2~arcsec.
If this ranking scheme reveals multiple `winners' of identical quality,
then the one that was observed closest to the optical axis of the camera is chosen.

\subsection{Compiling the final source catalogue}

As the final step, the primary detections that have been
selected above are compiled
into the 98-column source catalogue
that is described in Appendix~\ref{app:columns}
and made available on line.
The entire list of sources naturally includes 
a significant number of spurious entries
as a result of the very sensitive detection levels
that are employed by the CASU pipeline by default.
To limit the size of the source catalogue,
we have decided to enforce three basic criteria
which must be met for a candidate source
to be included in the catalogue:
\begin{enumerate}
\item the source must have been detected at SNR$>5$ in at least
one of the bands, i.e. it is required that at least one of
\emph{rErr}, \emph{iErr} or \emph{haErr} is smaller
than 0.2 mag;
\item the shape of the source must not be an obvious
cosmic ray or noise artefact, i.e. we require
either \emph{pStar} or \emph{pGalaxy} to be
greater than 20\%;
\item the source must not have been detected in one of the strongly
vignetted corners of the detector, 
not have had any known bad pixels in the aperture,
and not have been on the edge of one of the CCDs,
i.e. we require the \emph{errBits} quality score
to be smaller than 64.
\end{enumerate}

A total of 219 million primary detections satisfied
the above criteria and have been included in the catalogue.
Amongst these, 158 million objects 
are detected in all three bands (72\%),
30 million are detected in two bands (14\%),
and 31 million entries are single-band detections (14\%).
Roughly half of the single-band detections were made in the $i$-band.
This is likely explained by the fact that the $i$-band is least
affected by interstellar extinction and can occasionally pick up
highly-reddened objects which are otherwise lost in $r$/\ha.

\section{Discussion}
\label{sec:discussion}

Having explained how the catalogue was generated,
we now offer an overview of its properties
by discussing  
(i) the recommended quality criteria,
(ii) the typical photometric uncertainties,
and (iii) the source densities and the frequency of source blending.

\subsection{Recommended quality criteria}
\label{sec:qualitycriteria}

Like any other photometric survey,
the majority of the objects in our catalogue
are faint sources observed near the detection limits:
55\% of the entries in the catalogue
are fainter than $r > 20$
and 18\% are even fainter than $r > 21$.
The measurements of such faint objects
are naturally prone to large
random and systematic uncertainties,
for example, an inaccurately subtracted background
will introduce a proportionally larger systematic error
in a faint object.
Most scientific applications will require a set of
quality criteria to be enforced for the purpose
of removing faint and low-quality objects.

The choice of quality criteria is often a complicated
trade-off between completeness on one hand
and accuracy on the other.
To aid users we have listed two sets of
recommended quality criteria 
in Tables~\ref{tab:reliable} and \ref{tab:veryreliable}.

Firstly, Table~\ref{tab:reliable} details
a set of minimum quality criteria
which should benefit most applications
which require reliable $r-i$ and $r-$\ha\ colours
without removing more than $\sim$80\%
of the sources brighter than $r < 19$.
The listed criteria are designed to 
(i) remove low-SNR sources, 
(ii) remove saturated sources,
and (iii) remove objects for which the 2.3~arcsec diameter
aperture magnitudes are inconsistent 
with the alternative 1.2~arcsec diameter aperture measurements.
The last criterion is a proxy
for identifying objects which are affected
by inaccurate background subtraction
or failed source deblending.
A total of 86 out of 219 million sources 
(39\%) satisfy all the criteria listed in Table~\ref{tab:reliable}
and are hereafter referred to as ``reliable''.
For convenience, the catalogue contains a boolean column
named \emph{reliable} which flags these objects
and makes their selection easy.

\begin{table*}
\vspace{2cm}

\begin{tabular}{p{8cm}lp{6cm}}
  \hline
  Quality criterion & Rows passed & Description \\
  \hline
  rErr\,$<$\,0.1 {\sc and} iErr\,$<$\,0.1 {\sc and} haErr\,$<$\,0.1 &
  109 million (50\%) &
  Require the photon noise to be less than 
  0.1 mag in all bands (i.e. SNR$>$10).
  This implicitly requires a detection in all three bands.  \\
  $r>13$ {\sc and} $i>12$ {\sc and} \ha\,$>12.5$ {\sc and not} \emph{saturated} &
  158 million (72\%) &
  The brightness must not exceed the nominal saturation limit
  and the peak pixel height must not exceed 55\,000 counts.
  Again, this implicitly requires a detection in all three bands.
  \\
  $|\mathrm{r}- \mathrm{rAperMag1}| 
  < 3\sqrt{\mathrm{rErr}^2 + \mathrm{rAperMag1Err}^2} 
  + 0.03$ &
  176 million (80\%) &
  Require the $r$ magnitude measured 
  in the default 2.3\arcsec\ diameter aperture
  to be consistent with the measurement 
  made in the smaller 1.2\arcsec\ aperture,
  albeit tolerating a 0.03 mag systematic error.
  This will reject sources for which the background
  subtraction or the deblending procedure
  was not performed reliably. \\
  $|\mathrm{i}- \mathrm{iAperMag1}| 
  < 3\sqrt{\mathrm{iErr}^2 + \mathrm{iAperMag1Err}^2} 
  + 0.03$ &
  183 million (84\%) &
  Same as above for $i$. \\
  $|\mathrm{ha}- \mathrm{haAperMag1}| < 
  3\sqrt{\mathrm{haErr}^2 + \mathrm{haAperMag1Err}^2} 
  + 0.03$ &
  158 million (72\%) &
  Same as above for \ha. \\
  \hline
  All of the above (flagged as {\bf\emph{reliable}}) &
  86 million (39\%) & \\
  \hline
\end{tabular}
\caption{Recommended minimum quality criteria 
for selecting objects with reliable colours 
from the IPHAS DR2 source catalogue. 
86~million entries in the catalogue (39\%)
satisfy all the criteria listed in this table.
For convenience, these have been flagged in the catalogue
using the column named \emph{reliable}.}
\label{tab:reliable}

\vspace{2cm}

\begin{tabular}{p{8cm}lp{6cm}}
  \hline
  Quality criterion & Rows passed & Description \\
  \hline
   reliable &
   86 million (39\%) &
   The object must satisfy the criteria listed in Table~\ref{tab:reliable}. \\

   pStar $>$ 0.9 &
   145 million (66\%) &
   % pStar > 0.89 is identical to mergedClass == -1
   The object must appears as a perfect point source,
   as inferred from comparing its Point Spread Function (PSF)
   with the average PSF measured in the same CCD. \\

   {\sc not} \emph{deblend} &
   177 million (81\%) &
   The source must appear as a single, unconfused object. \\

   {\sc not} \emph{brightNeighb} &
   216 million (99\%) &
   There is no star brighter than $V < 4$ within 10 arcmin, 
   or brighter than $V < 7$ within 5 arcmin.
   Such very bright stars cause scattered light and diffraction spikes,
   which may add systematic errors to the photometry
   or even trigger spurious detections. \\  
  \hline

  All of the above (flagged as {\bf\emph{veryReliable}}) &
  59 million (27\%) & \\
  \hline
\end{tabular}
\caption{Additional quality criteria which are recommended
for applications which require very reliable colours
at the expense of completeness. 
For convenience, the sources which satisfy the criteria listed
in this table have been flagged in the catalogue
using the column named \emph{veryReliable}.}
\label{tab:veryreliable}

\vspace{2cm}
\end{table*}

For applications which require
an even higher standard of reliability,
a further set of additional quality criteria
are listed in Table~\ref{tab:veryreliable}.
These criteria are designed to ensure that
(i) the object appeared as a perfect point source,
(ii) the object was not blended with a nearby neighbour,
and (ii) the object was not located near a very bright star.
59 million sources (27\%) satisfy
these additional criteria 
and are hereafter referred to as ``very reliable''.
Again, the catalogue contains a boolean column
named \emph{veryReliable} which flags these objects.

Figure~\ref{fig:magdist} compares the $r$-band magnitude
distribution for objects with and without the 
\emph{reliable} and \emph{veryReliable} criteria applied. 
We find that 81\% of the sources 
in the magnitude range $13 < r < 19$
are considered \emph{reliable},
which drops to 72\% in the range $19 < r < 20$
and 9\% at $r>20$.
In contrast, only 54\% of the sources 
in the magnitude range $13 < r < 19$
are considered \emph{veryReliable}.
The stricter criteria filter out a lot
of objects at early Galactic longitudes 
where source blending is a common problem
(we will demonstrate this in \S\ref{sec:densities}).
The \emph{veryReliable} flag should hence
only be used in applications which require very reliable photometry
at the expense of completeness,
which might be the case for e.g. spectroscopic target selection.

\begin{figure}
    \includegraphics[width=0.5\textwidth]{./plots/magdist/magdist-r.pdf} 
    \caption{r-band magnitude distribution for all objects in the catalogue 
    (light grey), for objects flagged as \emph{reliable} 
    according to the criteria set out in Table~\ref{tab:reliable} (grey),
    and for objects flagged as \emph{veryReliable} 
    following Table~\ref{tab:veryreliable} (dark grey).
    The magnitude distributions for $i$ and \ha\
    look identical, apart from being shifted
    by about 1 and 0.5 mag towards brighter magnitudes,
    respectively.}
    \label{fig:magdist}
\end{figure}

It is easy to see how the quality criteria
may be adapted to be more tolerant.
For example, by raising the allowed photometric uncertainties
from 0.1 mag to 0.2 mag one can retrieve 42 million candidate sources.

\subsection{Photometric uncertainties}

Figure~\ref{fig:uncertainties} shows the mean photometric
uncertainties as a function of magnitude for each band.
We find that the uncertainties typically
reach 0.1 mag near 20.5 in $r$ 
and 19.5 in $i$/\ha when the default 2.3\arcsec\ aperture is used.
At this point we note that the average colour
of objects in the survey is
$1.06\pm0.12$ for ($r$-$i$) and $0.44\pm0.03$ for ($r$-\ha).
The better depth of $r$ is hence compensated
by the fact that stars tend to have 
brighter magnitudes in $i$ and \ha.

\begin{figure}
    \includegraphics[width=0.5\textwidth]{./plots/errors/uncertainties.pdf} 
    \caption{Mean photometric uncertainties
             for $r$ (top), $i$ (middle) and \ha\ (bottom).
             Data points shown are the average values of
             columns \emph{rErr}, \emph{iErr} and \emph{haErr}
             in the catalogue, 
             and the errorbars show the standard deviations.
             The dashed and solid lines indicate 
             the 10$\sigma$ and 5$\sigma$ limits, respectively.
             These uncertainties are based only on the (Poissonian)
             photon noise and hence this figure does not show
             systematic or calibration uncertainties.}
    \label{fig:uncertainties}
\end{figure}

The uncertainties shown in Fig.~\ref{fig:uncertainties}
are the random errors based on the expected Poissonian photon noise.
Systematics, such as calibration and deblending errors,
are not included.
To appraise the extent to which our photometry is affected
by such systematics, we can exploit the
secondary measurements which were made as part of the
field-pair observing strategy and are available for 51\%
of the sources in the catalogue.

Figure~\ref{fig:pairmag} shows the mean residuals between
the primary and secondary magnitudes
-- i.e. the average difference between catalogue columns \emph{r}-\emph{r2},
\emph{i}-\emph{i2}, \emph{ha}-\emph{ha2} -- as a function of magnitude.
We find that sources across the magnitude range 
$13 < r < 17$ are consistent at the level of 5\%
(i.e. $\sigma_{r-r2} \le 0.05$ mag),
with the best repeatability
reached at $r=14$ ($\sigma_{r-r2}~=~0.041$~mag).
We draw attention to the fact that brighter stars
tend to show significantly larger residuals
-- e.g. $\sigma_{r-r2}~=~0.14$~mag at $r=12$ --
which is due to saturation effects.
At the faint end we find residuals which
show significantly more scatter than would be expected
from photon noise alone, that is,
the effects of source blending and background subtraction
appear to dominate from $\sim$20th magnitude onwards.

\begin{figure*}
    \vspace{1cm}
    \includegraphics[width=0.5\textwidth]{./plots/pairmag/pairmag.pdf} 
    \caption{Photometric repeatability illustrated by plotting
             the mean residuals between all the primary and secondary detections
             in the catalogue as a function of magnitude.
             The best photometric repeatability is reached at $r=14$
             with $\sigma_{r-r2}~=~0.041$~mag.
             Note that bright stars at $r<13$ and $i<12$ 
             show increasing uncertainties due to saturation effects.}
    \label{fig:pairmag}
    \vspace{1cm}
    \includegraphics[width=0.5\textwidth]{./plots/pairmag/pairmag-reliable.pdf} 
    \caption{Same as Figure~\ref{fig:pairmag},
    except that only the subset of sources
    flagged as \emph{veryReliable} are now included.
    We find that applying the quality criteria
    has improved the photometric repeatability significantly.
    The best repeatability is again reached at $r=14$
    but has reduced to $\sigma=0.028$ mag.
    The quality criteria have also been successful
    at removing objects with large systematics at the bright
    and faint ends.}
    \label{fig:pairmag_reliable}
\end{figure*}

In Figure~\ref{fig:pairmag_reliable} we show 
a similar comparison of the primary and secondary detections,
but this time we have only include sources which are flagged
as \emph{veryReliable} in the catalogue
(i.e. not saturated, not confused, not near bright stars, etc.)
We find that the average residuals are significantly better
for this subset of the catalogue. Sources across the magnitude range 
$13 < r < 17$ are consistent at the level of 0.03 mag,
and the best repeatability is again reached at $r=14$
with $\sigma_{r-r2}~=~0.028$~mag.
We conclude that the \emph{veryReliable} quality criteria are effective
in reducing the systematic errors to
the same level as the accuracy of the
global photometric calibration (cf. \S\ref{sec:calibration}).
Moreover, the large systematics at the bright and faint end
have disappeared.

\subsection{Source densities and blending problems}
\label{sec:densities}

The mean source density as a function of Galactic longitude
is shown in Figure~\ref{fig:density} (thick blue line).
The densities were computed by counting the 
number of sources in $1^\circ$-wide longitude bins
across which covered the entire latitude range $-5^\circ<b<+5^\circ$.
Unsurprisingly, we find the average source densities to increase
towards the Galactic centre.
For example, the average source density near $l\simeq 30^\circ$
is roughly 30\,000 objects per square degree,
which is five times more than the density
found near $l\simeq 180^\circ$.

In addition to the global trend, 
there are significant variations in the source density on smaller scales.
For example,  we find a significant drop near the constellations 
of Cygnus ($l\simeq 80$) and Aquila ($l\simeq 40$),
which are regions known to be affected
by high levels of foreground extinction.
Dark clouds are visible towards these constellation by eye,
and they are often referred to as ``the Great Rift''.

We warn however that the densities reported here 
have not been corrected for survey completeness
or differences in the observing conditions across the survey.
For example, the dip in the density near $l\simeq140^\circ$
is an artificial feature caused by gaps
in the footprint coverage (which are apparent in Fig.~\ref{fig:footprint}).
In a forthcoming paper,
we aim to calibrate the IPHAS-based source densities
by injecting artificial stars into the IPHAS images
and measuring their recovery rate (Farnhill et al., in preparation).
Indeed IPHAS has the potential to offer
calibrated, two-dimensional stellar density maps
which can be used to constrain detailed models of our Galaxy,
but it is beyond the scope of the present work.

\begin{figure*}
    \includegraphics[width=\textwidth]{./plots/density/density.pdf} 
    \caption{Mean number density of sources in the catalogue 
    as a function of Galactic longitude, 
    with and without blended sources included. 
    The densities shown were computed by counting the sources 
    at each longitude between $-5^\circ<b<+5^\circ$ (upper blue line).
    We also show the densities based on only counting those sources 
    for which the \emph{deblend} flag is {\sc false}, 
    i.e. unconfused sources for which the CASU pipeline did not have to apply 
    a deblending procedure (lower red line). 
    }
    \label{fig:density}
\end{figure*}

In Figure~\ref{fig:density} we also shows the density
of non-blended sources (thin red line).
These are sources for which the \emph{deblend} flag is {\sc false},
i.e. sources for which the CASU pipeline did not have to apply 
a deblending procedure to separate the flux
originating from two or more overlapping objects.
We find a strong correlation between the source density
and the fraction of sources affected by source blending.
For example, only NN\% of the sources are blended
at $l>90^\circ$, whereas NN\% are blended at $l<90^\circ$.

As we explained earlier, blended sources must be used with caution.
Firstly, the deblending-procedure crucially depends
on the local PSF being measured accurately. 
Secondly, blended sources may are likely candidates to have fallen victim
to source confusion during the band-merging procedure.
In future work we hope to investigate the use of
more advanced PSF-fitting routines
in which sources are measured simultaneously across all bands,
perhaps guided by an external list of sources provided
by near-infrared surveys or Gaia data.

\section{Demonstration}
\label{sec:demonstration}

We conclude this paper by demonstrating how the unique
$r$-$i$/$r$-\ha\ colour-colour diagram offered by this catalogue
can readily be used to
(i) characterise the extinction regime at different sightlines, and
(ii) identify \ha\ emission-line objects.

\subsection{Colour-colour and colour-magnitude diagrams}

The survey's unique $r$-\ha\ colour,
when combined with $r$-$i$,
has been shown to provide simultaneous constraints 
on intrinsic stellar colour and interstellar extinction \citep{Drew2008}. 
That is, the main sequence in the $r$-$i$/$r$-\ha\ diagram
runs in a direction that is different from the reddening vector,
because the $r$-\ha\ colour tends to act as a coarse proxy for spectral type
and is less sensitive to reddening than $r$-$i$.
As a result, the distribution of a stellar population
in the IPHAS colour-colour diagram
can offer a handle on the properties and extinction regime
along a line of sight.

This is demonstrated in Figures \ref{fig:l180}, \ref{fig:l45} \& \ref{fig:l30},
where we present three sets of IPHAS colour/magnitude diagrams
towards three distinct sightlines
located at Galactic longitudes $180^\circ$, $45^\circ$ and $30^\circ$,
respectively.
Each figure contains all the sources which are flagged as \emph{veryReliable}
and are located in a region of one square degree 
centred on the coordinates indicated in the diagram
(i.e. within a radius of $0.564^\circ$ from the indicated sightline).
For clarity, we have imposed the additional criterion
for photometric uncertainties to be smaller than 0.05 mag in each band
(corresponding to a magnitude limit near $\sim$19th magnitude, effectively).

Each of the diagrams reveals a well-defined locus,
which demonstrates the health of the catalogue and the global calibration
for investigating stellar populations across wide areas.
We have annotated the colour-colour diagrams
by showing the position of the unreddened main sequence (thin solid line),
the unreddened giant branch (thick solid line)
and the reddening track for an A0V-type star (dashed line)
-- all three are based on the \cite{Pickles1998} library of empirical spectra
tabulated for IPHAS by \cite{Drew2005}.
In the colour-magnitude diagrams we only show the reddening vector
together with the unreddened 1~Gyr isochrone due to \cite{Bressan2012},
which are made available for the IPHAS filter system through a
popular on line tool hosted by the Observatory of Padova (http://stev.oapd.inaf.it/cmd).
The isochrone and reddening vector has been placed
at an arbitrary distance of 2~kpc.

Each of the sightlines reveals a stellar population
with distinct characteristics.
Firstly, towards the Galactic anti-centre at $l=180^\circ$ (Fig.~\ref{fig:l180}) 
we find a population dominated by lowly-reddened 
main sequence stars and an apparent absence of reddened giants.
%consistent with the sightline extinction of XX from Schlegel (cite). 
In contrast, closer towards the galactic
centre at $l=45^\circ$ (Fig.~\ref{fig:l45})
we find a wealth of reddened objects
which appear to be separated into a lowly
and a highly reddened component,
perhaps revealing two distinct parts of the Galaxy.
Finally, in one of our earliest sightlines at $l=30^\circ$
we find a very high number of extremely reddened giants
in addition to an unreddened population
of foreground dwarfs.

The number density of stars in the colour-colour
and colour-magnitude space
can be compared against population synthesis models
to create three-dimensional maps of the extinction
across several kpc \citep{Sale2010,Sale2012}.
Such an extinction map based on our catalogue
is to be released in a separate paper
that accompanies this data release (Sale et al., in preparation).

\begin{figure*}
    \begin{minipage}[b]{\linewidth}
        \includegraphics[width=0.5\textwidth]{./plots/ccd-180-3.pdf} 
        \includegraphics[width=0.5\textwidth]{./plots/cmd-180-3.pdf}
    \end{minipage}
    \caption{Colour-colour and colour-magnitude diagram (left and right panel)
    showing sources flagged as \emph{veryReliable}
    located in an area of one square degree
    centred near the Galactic anti-centre 
    at $(l,b)=(180^\circ,+3^\circ)$.
    The colour-colour diagram shows the
    position of the main sequence (thin solid line),
    giant stars (thick solid line)
    and the reddening track for an A0V-type star (dashed line)
    based on the \citet{Pickles1998} library of empirical spectra.
    The colour-magnitude only show the reddening vector
    along with the unreddened 1~Gyr isochrone due to \citet{Bressan2012},
    which has been placed at an arbitrary distance of 2~kpc for reference.
    This is one of the least reddened sightlines
    in the survey %(Av XX, Schlegel et al. XXXX)
    and hence the observed stellar population appears to be dominated 
    by lowly reddened main sequence stars.}
    \label{fig:l180}
    \begin{minipage}[b]{\linewidth}
        \includegraphics[width=0.5\textwidth]{./plots/ccd-45-2.pdf}
        \includegraphics[width=0.5\textwidth]{./plots/cmd-45-2.pdf}
    \end{minipage}
    \caption{Same as above for $(l,b)=(45^\circ,+2^\circ)$,
    which is one of the highest-density sightlines in the survey,
    revealing a population of stars with a reddening distribution
    that appears to be bi-model.}
    \label{fig:l45}
    \begin{minipage}[b]{\linewidth}
        \includegraphics[width=0.5\textwidth]{./plots/ccd-30-0.pdf}
        \includegraphics[width=0.5\textwidth]{./plots/cmd-30-0.pdf} 
    \end{minipage}
    \caption{Same as above for $(l,b)=(30^\circ,0^\circ)$.
    This is one of the most reddened sightlines in the survey.
    %(XXX, Schlegel et al XXXX).
    }
    \label{fig:l30}
\end{figure*}

\subsection{Identifying \ha\ emission-line objects}

A primary motivation for carrying
out the survey 
was to enable the discovery of 
new emission-line objects across the Galactic Plane.
\ha\ in emission is a well-known tracer
for stars in the short-lived pre- or
post-main sequence stages of their evolution,
and hence IPHAS aims to allow larger, deeper
and more statistically robust samples of such rare objects
to be established.
The survey-wide identification and analysis 
of such stars is beyond the scope of the present work,
but in this section we demonstrate how the
catalogue may be used for this purpose.

An initial list of candidate \ha-emitters
based on the first IPHAS data release has previously
been presented by \cite{Witham2008}. 
Because no global uniform calibration was available
at the time, \citeauthor{Witham2008} employed 
a sigma-clipping technique to select objects with
large, outlying $r$-\ha\ colours.
In contrast, the new catalogue
allows objects to be picked out
from the $r$-$i$/$r$-\ha\ colour-colour diagram
using model-based colour criteria
rather than a statistical procedure.
In what follows we demonstrate this ability 
by selecting candidate emission-line objects
towards a small region in the sky.

The target of our demonstration is Sh 2-82:
a 5~arcmin-wide H{\sc ii} region located near $(l,b)=(53.55^\circ, 0.00^\circ)$
in the constellation of Sagitta.
Nicknamed by amateur astronomers as the ``Little Cocoon Nebula'',
Sh 2-82 is ionised by 
the $\sim$10th magnitude star HD\,231616
% Georgelin1973 claims B0III, others claim B0V
with spectral type B0V/III
\citep{Georgelin1973,Mayer1973,Hunter1990}.
The ionising star has been placed at a likely distance of 1.5-1.7 kpc
based on the photometric parallax
\citep{Mayer1973,Lahulla1985,Hunter1990}.

Figure~\ref{fig:mosaic_iphas} shows a 20-by-15 arcmin
colour mosaic centred on Sh 2-82,
composed of our \ha\ (red channel),
$r$ (green channel),
and $i$ (blue channel) images.
The ionising star can be seen as the bright object
near the centre of the H{\sc ii} region,
which is surrounded by a faint reflection nebula
and several dark cloud filaments.
For comparison, Figure~\ref{fig:mosaic_spitzer} shows
a mosaic of the identical region 
as seen by the Spitzer Space Telescope
in the mid-infrared. The Spitzer image
reveals a bubble-shaped structure of warm dust
which surrounds the entire H{\sc ii} region.

\begin{figure*}
    \begin{minipage}[b]{0.8\linewidth}
        \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-iphas.pdf} 
    \end{minipage}
\caption{IPHAS-based mosaic of H{\sc ii} region Sh 2-82,
composed of \ha\ (red channel), $r$ (green channel) and $i$ (blue channel). Yellow triangles show the position of candidate \ha-emitters
which have been selected from the colour-colour diagram
in Figure~\ref{fig:emitters}. Note that the H{\sc ii} region is surrounded by a faint blue/green reflection nebula
and dark cloud filaments.}
\label{fig:mosaic_iphas}
    \begin{minipage}[b]{0.8\linewidth}
        \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-spitzer.pdf} 
    \end{minipage}
    \caption{Star-forming region Sh 2-82 as seen in the mid-infrared
    by the Spitzer Space Telescope. The mosaic is composed of the 24\,\micron\ (red), 8.0\,\micron\ (green) and 4.5\,\micron\ (blue) bands.
    The image reveals a bubble-shaped structure which surrounds the {\sc Hii} region that is seen in the IPHAS mosaic of the same region (Figure~\ref{fig:mosaic_iphas}).}
    \label{fig:mosaic_spitzer}
\end{figure*}

Figure~\ref{fig:emitters} presents
the IPHAS colour-colour diagram for the region covered by the
mosaic images.
Gray circles show all objects in the region
which are brighter than $r<20$
and have been flagged as \emph{reliable}
in our catalogue.
The diagram also shows the unreddened main sequence (solid line)
and the expected position of unredded main-sequence stars
with \ha\ in emission at a strength of EW=$-10{\rm \AA}$ (dashed line),
taken from the colour simulations due to \citet{Barentsen2011a}.
Six stars are found to lie confidently above the 
dashed line at the level of $3\sigma$ 
(i.e. the distance between the dashed line is larger than
three times the uncertainty on the $r$-\ha\ colour).
These reliable candidate \ha-emitters
are marked by red triangles in the colour-colour diagram
and their photometry is detailed in Appendix~\ref{app:emitters}.

\begin{figure}
  \includegraphics[width=0.45\textwidth]{./plots/sh2-82-ccd.pdf}
    \caption{$r$-$i$/$r$-\ha\ diagram for the rectangular region of 
    20-by-15 arcmin centred on the H{\sc ii} region Sh 2-82,
    which is the area shown in Figure~\ref{fig:mosaic_iphas}.    
    The diagram shows all objects in the catalogue
    which have been flagged as \emph{reliable} and are brighter
    than $r<20$ (grey circles).
    The unreddened main sequence is indicated by a solid line,
    while the main sequence for stars with an \ha\ emission line
    strength of $-10\,\rm{\AA}$ EW is indicated by a dashed line,
    following the colour simulations due to \citet{Barentsen2011a}.
    Red triangles indicate objects which have been identified as
    as highly likely \ha-emitters (see text).}
    \label{fig:emitters}
\end{figure}

The spatial distribution of our six candidate emission-line objects
is marked by yellow triangles in the colour mosaic (Fig.~\ref{fig:mosaic_iphas}).
They are likely to be genuine young stars for two reasons.
Firstly, two of our candidates have recently been identified as likely
candidate Young Stellar Objects (YSO)
in an investigation of the region by \cite{Yu2012}.
In their study, the authors used 2MASS and Spitzer data
to select likely young stars by selecting objects with
circumstellar disks based on the infrared colour excess.
Secondly, we find that the four remaining objects are 
also detected in the Spitzer-based image,
although their colours are less extreme
than those identified as likely YSOs by \citeauthor{Yu2012}.

Prior to IPHAS, this region was essentially unstudied
at faint magnitudes in visible light.
\cite{Lahulla1985} reported magnitudes for 8 stars in the optical at 
$V < 15$. In contrast, the IPHAS catalogue offers photometry
for NN stars in the region down to $r<20$.
This demonstrates the ability of IPHAS to providing complimentary
for the wealth of poorly-studied star-forming regions
at low Galactic latitudes,
which have been unveiled in recent years
by the observed wealth of star-forming ``bubbles''
at mid-infrared wave%\cite{Bica2003} was the first to identify a
%near-infrared cluster
%associated with this region using near-infrared 2MASS data,
%and only recently has this cluster been confirmed
%in a short investigation which combined 2MASS
%and Spitzer data \cite{Yu2012}.
%Many of the young stars identified by \citeauthor{Yu2012}
%can be seen as pink and red stars in Fig.~\ref{fig:mosaic_spitzer}.lengths \cite[e.g.][]{Churchwell2006,Simpson2012}.
%In fact the cluster of young stars associated with Sh 2-82
%was only first reported by \cite{Bica2003},
%remained unstudied until the study by \cite{Yu2012}.

%Amongst the six emitters, N are detected at SNR X
%in the Spitzer X micron band.

%This demonstrates how IPHAS is providing complimentary
%data for some of the thousands of star-forming ``bubbles'' have been unveiled at mid-infrared wavelengths \cite[e.g.][]{Churchwell2006,Simpson2012}.
%Optically-unveiled members of such star-forming regions
%may help to constrain extinctions and distances towards such regions \citep{Barentsen2013}.
%Moreover, in \cite{Barentsen2011a} we showed
%how the history of star formation in a region,
%and the possible triggered nature, may be discussed.
%This is particularly true

%TODO: annotate -10 EW text on plot?
%TODO: mention Gaia, e.g. IPHAS for transient identification?

%\subsection{Caveats and lessons learnt}
%\cite{Bica2003} was the first to identify a
%near-infrared cluster
%associated with this region using near-infrared 2MASS data,
%and only recently has this cluster been confirmed
%in a short investigation which combined 2MASS
%and Spitzer data \cite{Yu2012}.
%Many of the young stars identified by \citeauthor{Yu2012}
%can be seen as pink and red stars in Fig.~\ref{fig:mosaic_spitzer}.

\section{Conclusions and future work}
\label{sec:conclusions}

A new data release for the IPHAS survey was presented,
taking the coverage up to over 90\% of the Northern Galactic Plane 
at $|b|<5^\circ$
and providing a uniform photometric calibration
for the first time.
We explained the data reduction and quality control procedures that
were applied, described and tested the new global photometric calibration,
and detailed the construction of the user-friendly source catalogue.

The observations included in this release
were found to achieve a median seeing of 1.1 arcsec
and a $5\sigma-$depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$, \ha$=20.3\pm 0.3$.
The global calibration and photometric repeatability
is accurate at the level of $\sigma=0.03$ mag,
providing a significant improvement over the 
previous data release.
The source catalogue provides the best-available
single-epoch astrometry and photometry
for 219~million unique sources.

The data-taking strategy developed for IPHAS
have since been reapplied to carry out a companion INT/WFC survey called UVEX
in U/$g$/$r$ \citep{Groot2009},
and also southern counterpart to IPHAS and UVEX 
is being carried out in $u$/$g$/$r$/$i$/\ha\ 
called VPHAS+ (Drew et al, in press).
We hope to re-use the experience gained by this data release
to create similar releases for these companion surveys.

In future work, we aim to draw upon the PanSTARRS photometric
survey to further improve the accuracy of our global calibration.
We will also aim to correct the photometry for the
radial field distortions.
%We may also produce a catalogue which details all detections.

%TODO: mention Gaia?

\section*{Data access and source code}
\label{sec:dataaccess}

The catalogue is made available through the Vizier
catalogue tool (http://vizier.u-strasbg.fr),
where it is known as the ``IPHAS DR2 Source Catalogue''
(catalogue ID ???).
The full catalogue can also be downloaded in its entirety
from the IPHAS website (www.iphas.org) as a collection 
of binary FITS tables which comprise 50\,GB,
which is accompanied by a script
to ingest the data into a PostgreSQL database.
Our website also provides access to the pipeline-processed
imaging data, which we have updated to include
the re-calibrated DR2 zeropoint in the image headers.

The source code that was used to generate
the catalogue is available at
https://github.com/barentsen/iphas-dr2

\section*{Acknowledgments}

The IPHAS survey was carried out 
at the Isaac Newton Telescope (INT).
The INT is operated on the island of La Palma
by the Isaac Newton Group
in the Spanish Observatorio del Roque de los Muchachos
of the Instituto de Astrofisica de Canarias.
All data were processed 
by the Cambridge Astronomical Survey Unit,
at the Institute of Astronomy in Cambridge.

Preparation of the catalogue was eased greatly
by a number of software packages,
including the Python modules
AstroPy \citep{Astropy},
APLpy, NumPy and SciPy,
the PostgreSQL database software,
the TOPCAT and STILTS packages \citep{Taylor2005,Taylor2006},
and the Montage software maintained by NASA/IPAC.
We also made use of the SIMBAD, Vizier and Aladin \citep{Aladin} tools operated at CDS, Strasbourg, France.

This research made extensive use of
several complementary photometric surveys.
Our global calibration was aided
by the AAVSO Photometric All-Sky Survey (APASS),
funded by the Robert Martin Ayers Sciences Fund.
The calibration was tested against the
Sloan Digitized Sky Survey (SDSS),
funded by the Alfred P. Sloan Foundation, the Participating Institutions, the National Science Foundation, the U.S. Department of Energy, the National Aeronautics and Space Administration, the Japanese Monbukagakusho, the Max Planck Society, and the Higher Education Funding Council for England.
The astrometric pipeline reduction made
significant use of the Two Micron All Sky Survey (2MASS),
which is a joint project 
of the University of Massachusetts
and the Infrared Processing and Analysis Center/
California Institute of Technology,
funded by NASA and the NSF.

GB and JED acknowledge the support of a grant
from the Science \& Technology Facilities Council
of the UK (STFC, ref ST/J001335/1).
HJF is in receipt of an STFC postgraduate studentship.

\bibliographystyle{mn2e}
\bibliography{dr2}

\appendix
\input{app_columns.tex}

\section{Candidate emission-line objects towards Sh 2-82}
\label{app:emitters}
\begin{table*}
    \begin{tabular}{lccc}
    \toprule
    Name & $r$ & $i$ & \ha  \\
    \midrule
IPHAS2 J192954.40+181026.1& $17.69\pm0.01$ & $16.12\pm0.01$ & $16.19\pm0.01$ \\
IPHAS2 J193011.01+182051.2& $18.55\pm0.02$ & $16.95\pm0.02$ & $17.31\pm0.02$ \\
IPHAS2 J193021.52+181954.5& $19.72\pm0.05$ & $17.94\pm0.03$ & $18.47\pm0.04$ \\
IPHAS2 J193024.45+181938.3& $19.31\pm0.04$ & $17.55\pm0.02$ & $17.99\pm0.03$ \\
IPHAS2 J193033.00+181609.3& $18.25\pm0.01$ & $16.91\pm0.01$ & $16.92\pm0.01$ \\
IPHAS2 J193042.48+182317.4& $19.96\pm0.03$ & $18.11\pm0.03$ & $18.48\pm0.03$ \\
    \bottomrule
    \end{tabular}
    \caption{Candidate \ha-emitters towards Sh 2-82.}
    \label{tbl:emitters}
\end{table*}

\label{lastpage}

\end{document}

### 2014-03-12 paper/iphas-dr2.tex
Comment on (180, 3)

@@
 Firstly, towards the Galactic anti-centre at $l=180^\circ$ (Fig.~\ref{fig:l180}) 
-we find a population dominated by lowly-reddened 
-main sequence stars and an apparent absence of reddened giants.
-%consistent with the sightline extinction of XX from Schlegel (cite). 
+we find a population dominated by lowly-reddened main sequence stars.
+This is consistent with the estimated total sightline extinction of $E(B-V)=0.49$
+given by \cite{Schlegel1998}, following the correction of \cite{Schlafly2011}.
+Looking in more detail we can see that the stellar locus is narrower for M-type dwarfs than for earlier types;
+we do not observe M dwarfs expereincing the strongest reddening possible for this sightline.
+This implies that extinction is still increasing at distances beyond $\sim 2$~kpc,
+where M dwarfs are too faint to be contained in the IPHAS catalogue.
+It is also clear that there are no unreddened stars earlier than $\sim$K0 visible,
+such stars would be saturated if within a few hundred parsecs.
+This therefore suggests that there is a measurable increase in extinction locally.
+We also note an relative absence of late type giants which,
+due to the relative brevity of the corresponding phase of stellar evolution,
+would only account for a small proportion of a volume limited sample.
+
 In contrast, closer towards the galactic
@@
 perhaps revealing two distinct parts of the Galaxy.
+
 Finally, in one of our earliest sightlines at $l=30^\circ$
@@
 to create three-dimensional maps of the extinction
-across several kpc \citep{Sale2010,Sale2012}.
+across several kpc \citep{Sale2009,Sale2012}.
 Such an extinction map based on our catalogue
@@
 
+\end{document}\documentclass[useAMS,usenatbib]{mn2e}
+\usepackage{amssymb,amsmath}
+\usepackage[pdftex]{graphicx}
+\usepackage{longtable} 
+%\usepackage{lscape}
+\usepackage{float}
+\usepackage{booktabs}
+\usepackage{aas_macros}
+
+\def\ha{\mbox{H$\rm \alpha$}}
+\def\arcsec{$''$}
+\def\arcmin{$'$}
+\def\deg{$^{\circ}$}
+\def\micron{\mbox{$\mu$m}}
+
+\title[IPHAS Data Release 2]{The Second Data Release 
+of the INT Photometric H$\alpha$ Survey 
+of the Northern Galactic Plane (IPHAS DR2)}
+
+\author[G. Barentsen
+et. al]{G. Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
+H. J. Farnhill$^{1}$,
+J. E. Drew$^{1}$,
+M. J. Irwin$^{4}$, \newauthor
+E. A. Gonz$\acute{\rm{a}}$lez-Solares$^{4}$,
+R. Greimel$^{2}$,
+B. Mizalski$^{3}$,
+C. Ruhland$^{1}$, \newauthor plus many friends.
+\newauthor\\
+$^{1}$Centre for Astrophysics Research, STRI, University of Hertfordshire, College Lane Campus, Hatfield AL10 9AB\\
+$^{2}$Institute for Geophysics, Astrophysics and Meteorology, Institute of Physics, Karl-Franzens-Universit$\ddot{a}$t Graz, Universit$\ddot{a}$tsplatz 5, 8010 Graz, Austria\\
+$^{3}$Building, Institute, Street Address, City, Code, Country\\
+$^{4}$Institute of Astronomy, Madingley Road, Cambridge CB3 0HA}
+
+\begin{document}
+\date{Current draft typeset \today}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2013}
+
+\maketitle
+
+\label{firstpage}
+
+\begin{abstract} % BACKGROUND, OBJECTIVE, METHODS, RESULTS, CONCLUSIONS
+The INT/WFC Photometric H$\alpha$ Survey 
+of the Northern Galactic Plane (IPHAS)
+is a 1800 deg$^2$ imaging survey
+covering the entire northern Milky Way at $|b| < 5^\circ$
+in the $r$, $i$ and \ha\ filters 
+using the Wide Field Camera (WFC) 
+on the 2.5-meter Isaac Newton Telescope (INT).
+This data release presents the first 
+uniformly-calibrated source catalogue
+to have been extracted from the survey,
+providing single-epoch photometry
+for 219 million unique sources
+across 92\% of the survey.
+The observations were carried out between 2003 and 2012
+at a median seeing of 1.1 arcsec
+to a depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$ and \ha$=20.3\pm 0.3$
+($5\sigma$ limits, Vega system).
+We explain the data reduction 
+and quality control procedures,
+describe and test the new uniform photometric calibration,
+and detail the construction of the source catalogue
+and its quality warning flags.
+We find that the new calibration is accurate to
+$\sigma=0.03$ mag.
+Finally, we demonstrate the ability of the 
+catalogue's unique
+$r$-$i$/$r$-\ha\ colour-colour diagram to
+(i) characterise stellar populations and extinction regimes
+towards different Galactic sightlines
+and (ii) select reliable candidate \ha-emission line objects.
+The catalogue which accompanies this paper
+provides the much-needed visible-light counterpart
+to several infrared surveys of the Galactic Plane,
+including many poorly-studied star-forming regions and ``bubbles'',
+and provides images for some of the most
+crowded regions to be faced by Gaia.
+\end{abstract}
+
+\begin{keywords}
+catalogues, surveys, stars: emission line, Be, Galaxy: stellar content
+\end{keywords}
+
+\section{Introduction}
+
+Since the first data release in 2008, 
+the INT/WFC Photometric H$\alpha$ Survey 
+of the Northern Galactic Plane (IPHAS)
+has provided new insights into the contents and structure 
+of our own backyard, the Milky Way. 
+The original motivation for undertaking 
+this large-scale programme of observation
+-- spanning almost a decade 
+and using more than 300 nights 
+at the Isaac Newton Telescope (INT) in La Palma -- 
+was to provide the digital update 
+to the photographic northern H$\alpha$ surveys 
+of the mid-20th century. 
+By increasing the sensitivity 
+with respect to these previous surveys 
+by a factor $\sim$1000 (7 magnitudes), 
+it was envisaged that IPHAS would allow 
+the limited bright samples of Galactic emission line objects 
+available at the outset \citep[e.g.][]{Kohoutek1999}, 
+to be extended into larger, deeper, more statistically-robust samples 
+that in turn could better inform our understanding 
+of the early and late stages of stellar evolution. 
+This aim has begun to be realised through a
+range of published studies including: 
+a preliminary catalogue of candidate emission line objects \citep{Witham2008};
+discoveries of new northern symbiotic stars \citep{Corradi2008, Corradi2010}; 
+new cataclysmic variables \citep{Witham2007}; 
+new groups of young stellar objects \citep{Vink2008,Barentsen2011a};
+along with discoveries of new and remarkable planetary nebulae 
+\citep{Mampaso2006, Corradi2011, Viironen2011}.
+
+Over the years it has become apparent that the legacy of IPHAS 
+will extend beyond the traditional \ha\ applications 
+of identifying emission line stars and nebulae. 
+Through the provision of $r$, $i$ broadband photometry 
+alongside H$\alpha$ data,
+IPHAS has created the opportunity 
+to study Galactic Plane populations 
+in a new way.
+For example, the survey’s unique $r-H\alpha$ colour, 
+when combined with $r-i$,
+has been shown to provide simultaneous constraints 
+on intrinsic stellar colour and interstellar extinction \citep{Drew2008}. 
+This has opened the door 
+to a wide range of Galactic science applications, 
+including the mapping of extinction across the Plane in three dimensions
+and the probabilistic inference of stellar properties
+\citep{Sale2009, Sale2010, Giammanco2011, Sale2012, Barentsen2013}. 
+In effect, the availability of narrowband H$\alpha$ 
+alongside $r, i$ magnitudes
+provides coarse spectral information for huge samples of stars 
+which are otherwise too faint or numerous 
+to be targeted by spectroscopic surveys.
+For such science applications to succeed however, 
+it is vital that the imaging data is transformed 
+into a homogeneously calibrated photometric catalogue, 
+in which quality problems 
+and duplicate detections of the same source 
+are flagged.
+
+The first release of IPHAS data, 
+covering roughly half the survey footprint,
+was made in late 2007 \citep{Gonzalez-Solares2008}. 
+At the time the data were insufficiently complete 
+to support a homogeneously calibrated source catalogue.
+The goal of this paper is to present the next release 
+that takes the coverage up to over 90 percent of the survey area 
+and includes a uniform calibration.
+In this work we aim to
+(i) explain the data reduction 
+and quality control procedures that were applied,
+(ii) describe and test the new global photometric calibration, and 
+(iii) detail the construction and demonstrate the use 
+of the source catalogue that has been extracted 
+from the newly re-calibrated data.
+
+In \S\ref{sec:observations} we start by recapitulating the key points
+of the survey design and strategy.
+In \S\ref{sec:reduction} we describe the data reduction
+and quality control procedures.
+In \S\ref{sec:calibration} we explain the uniform re-calibration
+and test our results against the Sloan Digital Sky Survey (SDSS).
+In \S\ref{sec:catalogue} we explain how the source catalogue was compiled.
+In \S\ref{sec:discussion} we discuss the properties of the catalogue,
+and finally in \S\ref{sec:demonstration} we demonstrate
+the health of the release by demonstrating its scientific exploitation.
+In \S\ref{sec:conclusions} we conclude and outline
+our future ambitions.
+
+\section{Observations}
+\label{sec:observations}
+
+The detailed properties of the IPHAS observing programme 
+have been presented before 
+by \citet{Drew2005} and \citet{Gonzalez-Solares2008}. 
+To set the stage for this release, we briefly remind of some key points.
+IPHAS is a 1800~sq. deg. imaging survey of the northern Galactic Plane, 
+providing images and photometry in Sloan $r, i$ 
+along with narrowband H$\alpha$. 
+It is carried out using the Wide Field Camera (WFC) 
+on the 2.5-meter Isaac Newton Telescope (INT) in La Palma. 
+It is the first digital survey to offer comprehensive CCD photometry
+of point sources in the Galactic Plane at visible wavelengths, 
+and does so down to a limiting magnitude of $\sim$20th.
+The IPHAS footprint on the northern sky spans a box 
+of roughly 180 by 10 degrees, 
+taking in the entire northern Galactic Plane 
+at latitudes $-5^{\circ} < b < +5^{\circ}$ 
+and longitudes $30^{\circ} < l < 215^{\circ}$.
+
+The Wide Field Camera is a mosaic of 4 CCDs 
+that captures a sky area of close to 0.29 square degrees.
+To cover the entire northern Plane with some overlap,
+the survey area was divided into 7635 telescope pointings.
+Each of these pointings is accompanied by an offset position
+at a displacement of $+$5 arcmin in declination 
+and $+$5 arcmin in right ascension,
+to deal with inter-CCD gaps, detector imperfections,
+and to enable quality checks. 
+The basic unit of observation hence
+amounts to $2 \times 3$ exposures, 
+in which each of the 3 survey filters is exposed at 2 offset sky positions, 
+within an elapsed time of 10 minutes.
+We shall refer to the unit of 3 exposures at the same position 
+as a \emph{field},
+and the combination of two fields at a small offset as a \emph{field pair}.
+The survey hence contains 15270 fields grouped into 7635 field pairs.
+To achieve the desired survey depth of 20th magnitude or fainter, 
+the filter exposure times were set at 120 sec (narrowband \ha), 
+30 sec ($r$) and 10 sec ($i$)
+in the majority of the survey observations.\footnote{In 2003 
+the $r$-band exposure time was 10~sec instead of 30~sec,
+and since Oct 2010 the $i$-band exposure time 
+has been increased from 10~sec to 20~sec.}
+
+Data-taking began in the second half of 2003, 
+and every field had been observed at least once by the end of 2008.
+At that time only 76 percent of the field pairs 
+satisfied our minimum quality criteria however,  
+often due to the effects of clouds, poor seeing, or technical faults
+(the quality criteria will be detailed in the next section). 
+Since then, a programme of repeat observations has been in place 
+to improve data quality. 
+As a result, 92\% of the survey 
+now benefits from quality-approved data.
+The most recent observations which are part of this release
+were obtained in November 2012.
+
+\begin{figure*}
+        \includegraphics[width=1\linewidth]{./plots/footprint/footprint_small.png}
+        \caption{Survey area showing the footprints
+        of all the quality-approved IPHAS fields
+        which have been included in this data release.
+        The area covered by each field has been coloured black
+        with a semi-transparent opacity of 20\%,
+        such that regions where fields overlap are darker.
+        The IPHAS strategy is to observe each field twice
+        with a small offset,
+        and hence the vast majority of the area 
+        is covered twice (dominant gray colour).
+        There are small overlaps between all the neighbouring fields
+        which can be seen as a honeycomb-style pattern
+        of dark gray lines across the survey area.
+        Regions with incomplete data are apparent as white gaps (no data) 
+        or as the lightest shade of gray
+        (denoting that only the offset position is missing).}
+        \label{fig:footprint}
+\end{figure*}
+
+Figure~\ref{fig:footprint} shows the footprint 
+of the quality-approved observations included in this work. 
+The fields which remain missing 
+-- covering 7 percent of the survey area --
+are predominantly located towards the Galactic anti-center 
+at $l > 120^o$.
+Fields at these longitudes can only be accessed from La Palma 
+in the months of November-December,
+which is when the weather and seeing conditions are often poor
+at the INT and observing attempts have failed repeatedly.
+To enable the survey to be brought to completion, 
+a decision was made recently to limit repeats in this area 
+to individual fields requiring replacement,
+i.e. fresh observations in all 3 filters may only be obtained 
+at one of the two offset positions, 
+if the data for the partner offset has already passed quality control.  
+The catalogue is structured such that it is clear 
+where a contemporaneous observation of both halves of a field pair
+is not available.
+
+\section{Data reduction and quality control}
+\label{sec:reduction}
+
+\subsection{Initial pipeline processing}
+
+All raw data obtained with the INT were transferred
+to the Cambridge Astronomical Survey Unit (CASU) 
+for initial processing and archival.
+The procedures used by CASU were originally devised
+for the INT Wide Field imaging Survey \citep[WFS;][]{McMahon2001,Irwin2005},
+which was a 200 deg$^2$ survey programme carried out 
+between 1998 and 2003 after the WFC was commissioned.
+Because IPHAS uses the same telescope and camera combination,
+we have been able to benefit from the existing WFS pipeline.
+A detailed description of the processing steps 
+is found in \citet{Irwin2001}.
+Its application to IPHAS has previously been described
+by \citet{Drew2005} and \citet{Gonzalez-Solares2008}
+and much of the source code is available 
+on line\footnote{http://casu.ast.cam.ac.uk/surveys-projects/software-release}. 
+In brief, the pipeline takes care of bias subtraction,
+linearity correction, flat-fielding,
+gain correction and de-fringing.
+
+The reduced images are then stored in a multi-extension FITS file 
+with a primary header describing the characteristics
+(position, filter, exposure time, etc.) 
+and four compressed image extensions 
+corresponding to each of the four CCDs.
+Source detection and characterisation is then carried out 
+using the \textsc{imcore} tool \citep{Irwin1985,Irwin1997}.
+The flux of each source is measured using both
+the peak pixel height (i.e. a square 0.33$\times$0.33\arcsec\ aperture)
+as well as a series of circular apertures of increasing diameter 
+(1.2\arcsec, 2.3\arcsec, 3.3\arcsec, 4.6\arcsec\ and 6.6\arcsec).
+
+The local background levels are estimated 
+by computing the sigma-clipped median
+flux in a grid of 64$\times$64 pixels (21$\times$21\arcsec)
+across the image,
+which is then interpolated to obtain an estimate 
+of the background level at each pixel.
+These sky levels are subtracted from the aperture photometry and
+-- when required --
+a deblending routine is applied which also attempts to remove
+the contamination from any other nearby sources.
+Whilst this approach works very well 
+across the vast majority of the survey area,
+the Galactic Plane unavoidably contains crowded regions 
+with large numbers of overlapping sources
+or rapidly spatially-varying nebulosity,
+in which case aperture photometry must always be interpreted 
+with caution.
+In \S\ref{sec:catalogue} we will explain that overlapping
+sources to which the deblending routine was applied 
+are flagged in the catalogue using the \emph{deblend} warning flag.
+
+Finally, an astrometric solution is determined
+based on the 2MASS point source catalog \citep{Skrutskie2006},
+which itself is calibrated 
+in the International Celestial Reference System (ICRS).
+A provisional photometric calibration is also provided 
+based on the average zeropoint
+determined from a set of standard stars observed in the same night.
+Sources are classified morphologically
+-- stellar, galaxy or noise --
+based on the curve-of-growth determined
+from measuring the source intensity in a series of growing apertures.
+Finally, the resulting source detection tables are also stored 
+in multi-extension FITS files.
+
+At the time of preparing DR2,
+the CASU pipeline had processed
+74\,195 IPHAS exposures 
+in which a total of 1.9~billion \emph{candidate sources} were detected 
+at the sensitive default detection level of 1.25\,$\sigma$
+-- unavoidably including spurious detections, artefacts and
+duplicate detections 
+(in \S\ref{sec:catalogue} we will explain
+how these have been removed or flagged).
+The pipelined data set -- comprising 2.5~terabyte of FITS files --
+was then transferred to the University of Hertfordshire
+for the purpose of transforming the raw
+detection tables into a reliable source catalogue which is 
+(i) quality-controlled,
+(ii) homogeneously calibrated, and 
+(iii) contains user-friendly columns and warning flags.
+It is these post-processing steps which are explained below.
+
+\subsection{Quality control}
+\label{sec:qc}
+
+Observing time for IPHAS was obtained
+on a semester-by-semester basis
+through the traditional time allocation committees 
+of the Isaac Newton Group of telescopes,
+which are competitive and invariably over-subscribed.
+For this reason, we attempted to utilise 
+\emph{all} the nights allocated to IPHAS,
+even those which were partially or entirely non-photometric
+or otherwise affected by technical problems 
+(e.g. electronic noise or telescope tracking problems).
+Any unsuitable data that was taken as a result
+of this strategy was subsequently flagged and rejected
+using a series of seven quality criteria,
+which ensure a reliable and homogeneous level of quality
+across the data release:
+
+\begin{figure}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=\textwidth]{./plots/depth_r.pdf} 
+    \end{minipage}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=\textwidth]{./plots/depth_i.pdf} 
+    \end{minipage}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=\textwidth]{./plots/depth_h.pdf} 
+    \end{minipage}
+    \caption{Distribution of the 5$\sigma$ limiting magnitude
+             across all quality-approved survey fields
+             for $r$ (top), $i$ (middle) and \ha\ (bottom).
+             Fields with a limiting magnitude brighter than
+             20th ($r$) or 19th (\ha/$i$) were rejected
+             from the data release.
+             The $r$-band depth is most sensitive 
+             to the presence of the moon above the horizon, 
+             which is evidenced by the wide and bi-model shape
+             of its distribution.}
+    \label{fig:depth}
+\end{figure}
+
+(1) \emph{Depth.} 
+We discarded any exposures for which the $5\sigma$ limiting magnitude 
+was worse than 20th magnitude in the $r$-band
+or worse than 19th in $i$ or \ha. 
+Such data were typically obtained during poor weather or full moon.
+Most observations fared significantly better than these limits.
+Figure~\ref{fig:depth} presents the distribution of limiting magnitudes
+for all quality-approved fields,
+which shows a mean depth of 
+$21.2\pm0.5$ ($r$), $20.0\pm0.3$ ($i$) and $20.3\pm0.3$ (\ha).
+We found that the depth achieved depended 
+most strongly on the presence of the moon,
+which was above the horizon during 62 percent 
+of our observations and explains the wide and bi-modal shape
+of the $r$-band limiting magnitude distribution 
+(top panel in Fig.~\ref{fig:depth}).
+In contrast, the depth attained in $i$ and \ha\ 
+is less sensitive to moonlight
+and the distribution of their depths
+is hence more narrow
+(middle and bottom panel in Fig.~\ref{fig:depth}).
+
+Our minimum limiting magnitude criteria
+have led us to exclude 9\% of the pipelined data.
+We note that much of the excluded data may nevertheless be useful
+for e.g. time-domain studies of bright stars.
+The detection tables for any such discarded data are made
+available through our website (www.iphas.org)
+but are ignored in what follows.
+
+(2) \emph{Ellipticity.} 
+The ellipticity of a point source,
+defined as $e = 1 - b / a$ 
+with $b$ the semi-minor and $a$ the semi-major axis,
+is a morphological measure of the elongation of the point spread function.
+It is expected to be zero (circular) across the field 
+in a perfect imaging system,
+but it is slightly non-zero in any real telescope data 
+due to optical distortions and tracking errors.
+The mean ellipticity across a field in the IPHAS data set 
+is $0.09\pm0.04$.
+There have been sporadic episodes with higher ellipticities however
+due to mechanical glitches in the telescope tracking system.
+For this reason, we rejected exposures in which the mean ellipticity
+across the detector exceeded $e > 0.3$,
+which is when the photometric measurements delivered by the pipeline
+were found to become degraded.
+Only 0.4\% of the exposures were discarded on this basis.
+
+(3) \emph{Seeing.} 
+The survey originally aimed to obtain data 
+at seeing better than 1.7 arcsec.
+This target is currently attained across 86\% of the footprint,
+in particular at early longitudes,
+e.g. 92\% of the fields at $l<120^\circ$ are better than 1.7 arcsec.
+Figure~\ref{fig:seeing} presents the distribution
+of the mean seeing for all the quality-approved fields.
+We find a median value of 1.1~arcsec in $r$/\ha\
+and 1.0~arcsec in $i$.
+In the $r$-band, 90\% of the data is better than 1.5~arcsec
+and 10\% is better than 0.8~arcsec.
+To improve the area covered by this data release,
+we have decided to include the small fraction of data
+that was obtained under seeing up to 2.5 arcsec,
+so that only 1 per cent of the pipelined exposures
+had to be excluded.
+In \S\ref{sec:catalogue} we will explain
+that the catalogue is structured
+such that the information on the seeing is included,
+and that the photometry listed
+is based on the exposures with the best-available seeing.
+
+\begin{figure}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=\textwidth]{./plots/seeing_r.pdf} 
+    \end{minipage}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=\textwidth]{./plots/seeing_i.pdf} 
+    \end{minipage}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=\textwidth]{./plots/seeing_ha.pdf} 
+    \end{minipage}
+    \caption{Seeing distribution across all
+             quality-approved survey fields
+             for $r$ (top), $i$ (middle) and \ha\ (bottom).}
+    \label{fig:seeing}
+\end{figure}
+
+(4) \emph{Photometric repeatability.} 
+The IPHAS field-pair observing strategy 
+ensures that every pointing is immediately followed 
+by an offset pointing at a displacement of $+$5 arcmin in Dec 
+and $+$5 arcmin in RA.
+This allows pairs of images to be checked 
+for the presence of clouds or electronic noise.
+To exploit this information,
+the overlap regions of all field pairs were systematically cross-matched
+to verify the consistency of the photometric measurements
+for stars observed in both pointings.
+We rejected field pairs in which more than 2\% of the stars 
+showed an inconsistent measurement at the level of 0.2 mag,
+or more than 26\% were inconsistent at the level of 0.1 mag.
+These limits were determined empirically by inspecting
+the images and photometry by eye.
+11\% of the data was rejected as part of this step.
+
+(5) \emph{Source density mapping.}
+Spatial maps showing the number density of the detected sources
+down to 20th magnitude were created to verify the health
+of the data and to check for unexpected artefacts.
+%(e.g. bright satellite trails)
+In particular, we created density maps
+which showed the number of \emph{unique} sources
+obtained by cross-matching the detection tables of
+all three bands with a maximum matching distance of 1 arcsec.
+The success of such a cross-matching procedure crucially depends
+on the accuracy of the astrometric solution in each band,
+and hence we were able to detect and correct
+exposures with poor astrometry by inspecting the density map
+for spurious density variations.
+
+(6) \emph{Visual examination.}
+All images and their associated photometric colour/magnitude diagrams
+were inspected by a team of 20 survey team members, 
+such that each image in the data release 
+was looked at by at least three different pairs of eyes.
+Images deemed unsuitable were flagged, investigated and
+excluded from the release if necessary. 
+6\% of the attempts to observe a field
+were placed on a \emph{black-list}
+for this reason, most commonly due to the obvious presence
+of clouds or extreme levels of scattered moonlight.
+Such fields were often rejected as part of 
+one or more of the above quality criteria as well.
+
+(7) \emph{Contemporaneous field data.} 
+Finally, only exposures which are part of a sequence 
+of three consecutive images of the same field (H$\alpha$/$r$/$i$) 
+were considered for inclusion in the release. 
+This ensures that the three bands for a given field
+are observed at nearly the same time --  
+essentially always within 5 minutes.
+An exception was made for 9 fields where the three exposures 
+could not be obtained within the same night
+but for which the time gap did not exceed 48 hours.
+We note that the exact epoch of the magnitude in each band
+is included in the source catalogue
+(columns \emph{rMJD}, \emph{iMJD}, \emph{haMJD}).
+
+The above criteria were satisfied 
+for 14115 out of the 15270 fields (92\%).
+In a few cases more than one successful attempt to observe
+a field was available due to the fact that slightly stricter
+quality criteria were adopted in the initial years of the survey.
+In such cases, only the attempt 
+with the best seeing and depth was selected
+for inclusion in the release, because the focus 
+of this work is to deliver the most reliable
+measurement at a single epoch.
+Those interested in any of the rejected data 
+may nevertheless access the full set of detection tables 
+on line.
+
+\section{Photometric calibration}
+\label{sec:calibration}
+
+Having obtained a quality-approved set of observations,
+we now turn to the problem of placing the data
+onto a uniform photometric scale.
+
+\subsection{Provisional nightly calibration}
+
+For the purpose of providing an initial calibration 
+of the $r$ and $i$ broadband fluxes,
+photometric standard fields were observed every night.
+The standards were chosen from a list based on 
+the \cite{Landolt1992} and Stetson (http://cadcwww.dao.nrc.ca/standards) 
+objects.
+Two or three standard fields were observed 
+during the evening and morning twilight,
+and at intervals of 2-3 hours throughout the night.
+The CASU pipeline automatically identified the observed standards 
+and used them to determine a sigma-clipped average zeropoint \textsc{magzpt}
+for each night and filter,
+such that the number counts $DN$ 
+in the pipeline-corrected CCD frames
+relates to a magnitude $m$ as:
+\begin{equation}
+\begin{split}
+   m  = & \textsc{magzpt} - 2.5 \log_{10}( DN / \textsc{exptime} ) \\
+ &  - \textsc{extinct}\cdot(\textsc{airmass}-1) - \textsc{apcor} - \textsc{percorr}
+\label{eqn:mag}
+\end{split}
+\end{equation}
+where \textsc{exptime} is the exposure time in seconds,
+\textsc{extinct} is the atmospheric extinction coefficient 
+(typically 0.09 for $r$ and 0.05 for $i$ in La Palma),
+\textsc{airmass} is the normalised optical path length 
+through the atmosphere and
+\textsc{apcor} is a correction for the flux
+lost outside of the aperture used.
+Finally, \textsc{percorr} is a correction based on the difference
+between the median dark sky for a CCD against the median for all the CCDs 
+and as such is an ancillary correction 
+to account for sporadic gain variations. 
+All these quantities correspond to header keywords in the 
+multi-extension FITS files produced by the CASU pipeline.
+
+The zeropoint was determined such that the resulting magnitude system
+is in the WFC system that uses the SED of Vega 
+as the zero colour reference object. 
+Colour equations were used to transform between the IPHAS passbands 
+and the Johnson-Cousins system 
+of the published standard star photometry.
+The entire procedure has been found to deliver zeropoints which 
+are accurate at the level of 1-2\% 
+in photometric conditions \citep{Gonzalez-Solares2011}.
+
+Unlike the broadbands, 
+standard star photometry is not available in the literature 
+for the H$\alpha$ passband
+and hence there is no formally recognised flux scale 
+for the narrowband.
+We can specify here however 
+that the integrated in-band energy flux for Vega 
+in the IPHAS \ha\ filter 
+is $1.52 \times 10^{-7}$ ergs\,cm$^{-2}$\,s$^{-1}$ 
+at the top of the Earth's atmosphere,
+which is the flux obtained by folding 
+Vega's SED with the filter throughput curve 
+corrected for atmosphere and detector quantum efficiency
+\citep[using the method explained by][]{Drew2005}.
+This is 3.08 magnitudes less than the flux captured 
+by the much broader $r$ band
+which includes the \ha\ band.
+Hence to assure zero colour relative to the broadbands,
+we set the default zeropoint for the narrowband to be:
+\begin{equation}
+\textsc{magzpt}_{H\alpha} = \textsc{magzpt}_r - 3.08.
+\label{eqn:zpha}
+\end{equation}
+
+\subsection{Uniform re-calibration}
+
+Despite the best efforts made to obtain a nightly calibration,
+large surveys naturally possess field-to-field variations
+at the level of 0.1 mag
+due to atmospheric fluctuations during the night
+and imperfections in the pipeline or the instrument
+(e.g. the WFC is known to suffer from sporadic glitches
+in the timing of exposures).
+Such variations need to be corrected for 
+during a global re-calibration procedure.
+Notable past examples include the global re-calibration 
+of the Two Micron All Sky Survey \citep[2MASS;][]{Nikolaev2000},
+the Sloan Digital Sky Survey \citep[SDSS;][]{Padmanabhan2008}
+and the Panoramic Survey Telescope 
+and Rapid Response System survey \citep[Pan-STARRS;][]{Schlafly2012},
+which all achieved photometry 
+that is globally consistent to within 1--2\%.
+
+Surveys which observe identical stars at different epochs
+can use the repeat measurements to ensure a uniform calibration.
+For example, 2MASS attained its global calibration
+by observing six standard fields each hour, 
+allowing zeropoint variations to be tracked 
+over very short timescales \citep{Nikolaev2000}.
+Alternatively, the SDSS and PanSTARRS surveys could benefit
+from revisiting regions in their footprint to 
+carry out a so-called \emph{ubercalibration} procedure,
+in which multiple measurements of stars at different epochs
+are used to fit the calibration parameters
+\citep{Padmanabhan2008,Schlafly2012}.
+
+Unfortunately these schemes cannot be applied
+directly to IPHAS
+for two reasons. 
+Firstly, the survey was carried out 
+in competitive observing time
+on a non-dedicated telescope, 
+rendering the 2MASS approach 
+of observing standards at a very high frequency
+prohibitively expensive
+-- in part because standard fields 
+are very scarce within the Galactic Plane.
+Secondly, the aim of IPHAS is to obtain magnitudes at a single epoch
+and hence stars are not normally observed at more than one epoch,
+unless they happen to fall on the narrow overlap region 
+between two neighbouring fields.
+
+Although the IPHAS data does contain a significant number of 
+inter-field repeat measurements
+due to the small overlaps
+between neighbouring fields,
+we have found the information contained
+in these regions to be insufficient
+to constrain the calibration parameters.
+This is because photometry at the extreme edges of the WFC
+-- where neighbouring fields overlap -- 
+is itself prone to systematics at the level of 1-2\%.
+The cause of these errors is thought to include 
+the use of twilight sky flats in the pipeline,
+which are known to be imperfect for calibrating stellar photometry 
+due to stray light and vignetting \citep[e.g.][]{Manfroid1995}.
+Moreover, the illumination correction in the overlap regions
+is affected by a radial geometric distortion,
+which causes the pixel scale towards the edges 
+to increase \citep{Gonzalez-Solares2011}.
+Although these systematics are reasonably small within a single field,
+they can combine to cause artificial zeropoint gradients 
+across the survey
+when they are used to constrain a global calibration
+without external constraints.
+
+For these reasons, we have decided not to depend
+on an ubercalibration-type scheme alone,
+and have instead opted to exploit an external reference survey
+-- where available --
+to bring the majority of our data onto a uniform calibration.
+This is explained next.
+
+\subsubsection{Correcting zeropoints using APASS}
+
+We have been able to benefit from the
+AAVSO Photometric All-Sky Survey
+(APASS; http://www.aavso.org/apass)
+to bring the vast majority of the survey 
+onto a uniform scale.
+Since 2009,
+APASS has been using two 20~cm-astrographs
+to survey the entire sky down to $\sim$17th magnitude
+in five filters which include Sloan $r$ and $i$.
+The most recent catalogue available 
+at the time of preparing this work was APASS DR7,
+which provided data across roughly half of the IPHAS footprint (Fig.~\ref{fig:apass_r}).
+The photometric accuracy of APASS is currently estimated 
+to be at the level of 3\% (Henden, private communication),
+which is significantly better 
+than the provisional calibration of IPHAS
+for which we estimate the $1\sigma$-error to be $\sim$10\%.
+APASS achieves its uniform accuracy 
+by measuring each star at least two times in photometric conditions
+-- along with ample standard fields --
+using the large $3\times3$ square degrees field of view of its detectors.
+
+With the aim of bringing IPHAS to a similar accuracy of $\sim$3\%,
+we used the APASS catalogue to identify and adjust all IPHAS fields 
+which showed a magnitude offset larger than 3\% against APASS.
+For this purpose,
+the $r$- and $i$-band detection tables of each IPHAS field
+were cross-matched against the APASS DR7 catalogue 
+using a maximum matching distance of 1~arcsec.
+The magnitude range was limited to
+$13<r_{\rm APASS}<16.5$ and $12.5<i_{\rm APASS}<16.0$
+in order to avoid sources 
+brighter than the IPHAS saturation limit on one hand, 
+and to avoid sources near the faint detection limit of APASS 
+on the other hand.
+
+The resulting set of $\sim$220\,000 cross-matched stars were then used 
+to derive APASS-to-IPHAS magnitude transformations
+using a linear least-squares fitting routine, 
+which iteratively removed $3\sigma$-outliers to improve the fit.
+The solution converged to:
+\begin{align} 
+r_{\rm IPHAS} = r_{\rm APASS} - 0.121 + 0.032(r-i)_{\rm APASS} \label{eqn:apass_r} \\
+i_{\rm IPHAS} = i_{\rm APASS} - 0.364 + 0.006(r-i)_{\rm APASS} \label{eqn:apass_i}
+\end{align}
+The root mean square (rms) residuals of these transformations 
+are 0.041 and 0.051, respectively.
+The small colour terms in the equations
+indicate that the $r$ and $i$ filters 
+are very similar in both surveys.
+The transformations include a large fixed offset
+which is simply due to the fact that 
+APASS magnitudes are given in the AB system
+and IPHAS uses magnitudes relative to Vega.
+Separate transformations were derived for sightlines 
+with varying extinction properties to investigate the robustness
+of the transformations with respect to different reddening regimes,
+but the variations at these different
+sightlines were found to be insignificant.
+This is not surprising because heavily reddened objects 
+are naturally scarce at $r<16$.
+
+Having transformed APASS magnitudes into the IPHAS system,
+we then computed the median magnitude offset 
+for each field which contained at least 30 cross-matched stars.
+This was the case for 48\% of the fields.
+The mean offset was found to be
+$0.014\pm0.104$ for $r$ and $0.007\pm0.108$ for $i$
+(Table~\ref{tbl:offsets_before}).
+A total of 4596 fields showed a median offset
+exceeding $\pm$0.03~mag in either $r$ or $i$.
+
+We then applied the most important step in our re-calibration scheme,
+which is to adjust the provisional zeropoints of the 4596 aberrant fields
+such that their offset is brought to zero.
+This allowed the mean IPHAS-to-APASS offset 
+to be brought down to $0.000\pm0.011$ in both $r$ and $i$
+(Table~\ref{tbl:offsets_after}).
+The procedure of fitting magnitude transformations and
+correcting the zeropoints was repeated a few times to ensure 
+convergence, which was essentially reached after the first iteration.
+
+\begin{figure*}
+    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_rshift.pdf} 
+    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_r.pdf} 
+    \caption{Median magnitude offset in the $r$ band between IPHAS and APASS,
+             plotted on a field-by-field basis
+             prior to the re-calibration procedure.
+             Each square represents an IPHAS field
+             which contains at least 30 stars with a counterpart
+             in the APASS DR7 catalogue.
+             The colours denote the median
+             IPHAS-APASS magnitude offset in each field,
+             which was computed after applying the APASS-to-IPHAS
+             transformation to the APASS magnitudes (Eqn.~\ref{eqn:apass_r}).}
+        \label{fig:apass_r}
+    \vspace{1cm}
+    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_rshift.pdf}
+    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_r.pdf} 
+    \caption{Median magnitude offset in the $r$ band
+             between IPHAS and SDSS after the re-calibration
+             procedure was applied.
+             Again, each square represents an IPHAS field
+             which contains at least 30 stars
+             with a counterpart in the SDSS catalogue.
+             The colours denote the median IPHAS-SDSS magnitude offset
+             in each field,
+             which was computed after applying the SDSS-to-IPHAS
+             transformation to the SDSS magnitudes (Eqn.~\ref{eqn:sdss_r}).}
+    \label{fig:sdss_r}
+\end{figure*}
+
+\begin{figure*}
+	\vspace{2cm}
+    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_ishift.pdf} 
+    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_i.pdf} 
+    \caption{Same as Figure~\ref{fig:apass_r} for the $i$-band.}
+        \label{fig:apass_i}
+    \vspace{1cm}
+    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_ishift.pdf}
+    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_i.pdf} 
+    \caption{Same as Figure~\ref{fig:sdss_i} for the $i$-band.}
+    \label{fig:sdss_i}
+\end{figure*}
+
+\begin{table}
+        \begin{center}
+                \begin{tabular}{lcc}
+                        \toprule
+                         {\bf Before re-calibration} & Mean & $\sigma$  \\
+                        \midrule
+                        $r$ (IPHAS - APASS) & +0.014 & 0.104 \\
+                        $i$ (IPHAS - APASS) & +0.007 & 0.108 \\
+                        $r$ (IPHAS - SDSS) & +0.016 & 0.088 \\
+                        $i$ (IPHAS - SDSS) & +0.010 & 0.089 \\
+                        \bottomrule
+                \end{tabular}
+        \caption{Mean magnitude offsets for objects
+                 crossmatched between IPHAS and APASS/SDSS
+                 before the uniform re-calibration.
+                 Eqns.~\ref{eqn:apass_r}-\ref{eqn:sdss_i} were applied
+                 to the APASS/SDSS magnitudes to bring them into the
+                 Vega-based IPHAS system prior to computing
+                 the offsets.)
+                 }
+        \label{tbl:offsets_before}
+
+                \begin{tabular}{lcc}
+                        \toprule
+                         {\bf After re-calibration} & Mean & $\sigma$ \\
+                        \midrule
+                        $r$ (IPHAS - APASS) & +0.000 & 0.011\\
+                        $i$ (IPHAS - APASS) & +0.000 & 0.011 \\
+                        $r$ (IPHAS - SDSS)  & -0.001 & 0.029\\
+                        $i$ (IPHAS - SDSS) & -0.002 & 0.032 \\
+                        \bottomrule
+                \end{tabular}
+        \caption{Same as Table~\ref{tbl:offsets_before}
+                 but computed after the uniform re-calibration
+                 was carried out.}
+        \label{tbl:offsets_after}
+        \end{center}
+\end{table}
+
+\subsubsection{Adjusting fields not covered by APASS}
+
+At the time of writing, the APASS catalogue did not provide 
+sufficient coverage for 7359 of the fields in our data release.
+Fortunately, these fields are pre-dominantly located 
+in the early part of the Galactic Plane (Fig.~\ref{fig:apass_r}),
+which were typically observed during the summer months
+when photometric conditions are more prevalent at the telescope.
+These remaining fields have nevertheless 
+been brought onto the same uniform scale 
+by employing an ubercalibration-style scheme, explained below,
+which minimises the magnitude offsets between stars
+located in the overlap regions with neighbouring fields.
+Although we explained earlier that these overlap regions
+are prone to systematics, the use of APASS enabled us to
+keep the zeropoints of roughly half of the fields fixed,
+which avoids these systematics from combining to introduce
+artificial gradients across the survey.
+In the next section we will show this to be true by validating
+our calibration against SDSS.
+
+A general solution to the problem of minimising the magnitude
+differences between overlapping frames has previously been
+described by~\citet{Glazebrook1994}.
+In brief, there are two fundamental quantities to be
+minimised between each pair of overlapping exposures, 
+denoted by the indices $i$ and $j$. 
+Firstly, the mean magnitude difference between stars in the overlap
+region $\Delta_{ij}=\langle m_i-m_j\rangle=-\Delta_{ji}$ is a local
+constraint. 
+Secondly, to ensure the solution does not stray far 
+from the existing calibration, 
+the difference in zeropoints 
+$\Delta\mathrm{ZP}_{ij}=-\Delta\mathrm{ZP}_{ji}$ 
+between each pair of exposures must also be minimised.
+
+Minimisation of these two quantities is a linear least squares problem 
+because the magnitude $m$ depends linearly on the ZP (Eqn.~\ref{eqn:mag}).
+Hence we can find the ZP shift to be applied to each field 
+by minimising the sum:
+\begin{equation}
+   S = \sum_{i=1}^N \sum_{j=1}^N w_{ij} \theta_{ij} (\Delta_{ij} + a_i - a_j)^2
+   \label{eqn:chi2}
+\end{equation}
+where $i$ denotes the exposure of interest, 
+$j$ an overlapping exposure, 
+$N$ the number of exposures,
+$a_i$ the ZP to solve for 
+and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$), 
+$w_{ij}$ are weights set to the uncertainty in $\Delta_{ij}$
+and $\theta_{ij}$ is an overlap function 
+equal to either 1 if exposures $i$ and $j$ overlap or 0 otherwise. 
+Solving for $a_i$ is equivalent to solving $\partial
+S/\partial a_i=0$ which gives the matrix equation:
+\begin{equation}
+   \sum_{j=1}^N A_{ij} a_j = b_j
+   \label{eqn:matrix}
+\end{equation}
+where 
+\begin{eqnarray}
+   A_{ij} &=& \delta_{ij} \sum_{k=1}^N w_{jk}\theta_{jk} - w_{ij} \theta_{ij},\label{eqn:aij}\\
+   b_i &=& \sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ji} = -\sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ij}.\label{eqn:bi}
+\end{eqnarray}
+This prescription is essentially identical to \citet{Glazebrook1994}.
+%,although we changed the formulation to ensure that $A_{ij}$ is positive definite.
+
+As explained above, we enforce a strong external constraint
+on the solution by keeping the zeropoint fixed 
+for the 6756 fields which have been compared
+and calibrated against APASS earlier.
+We will hereafter refer to these fields as \emph{anchors}.
+It is asserted that the zeropoints $a_i$ of these anchor fields 
+are known and not solved for,
+though they do appear in the vector $b_j$ as constraints.
+In addition to the APASS-based anchors, 
+we selected 3273 additional anchor fields by hand
+to provide additional constraints in regions not covered by APASS.
+These extra anchors were deemed to have reliable zeropoints 
+based on 
+(i) the information contained in the observing logs,
+(ii) the stability of the standard star zeropoints during the night, and
+(iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
+which is located at $\sim$500~meter from the INT.
+
+We then solved Eqn.~\ref{eqn:matrix} for the $r$ and $i$ bands
+separately using the least-squares routine 
+in Python's {\sc scipy.sparse} module for sparse matrix algebra.
+This provided us with corrected zeropoints for the remaining fields,
+which were shifted on average by $+0.02\pm0.11$ in $r$ 
+and $+0.01\pm0.12$ in $i$ compared to their provisional calibration.
+
+We then turned to the uniform calibration of the \ha\ data.
+It is not possible to re-calibrate the narrowband 
+in the same way as the broadbands,
+because the APASS survey does not offer \ha\ photometry.
+We can reasonably assume however,
+that the corrections required for $r$ and \ha\ are identical
+because the \ha\ zeropoints have been derived directly from the
+$r$-band zeropoints during the provisional calibration (Eqn.~\ref{eqn:zpha}).
+Moreover, the IPHAS data-taking pattern ensured 
+that a field's \ha\ and $r$-band exposures
+were taken at essentially the same time, 
+separated only by the $\sim$30-second read-out time required by the WFC.
+We have hence corrected the \ha\ zeropoints 
+by re-using the zeropoint adjustements that were derived for the $r$ band
+in the earlier steps.
+An exception was made for 3101 fields
+for which our quality-control routines revealed
+zeropoint variations which exceeded 0.03~mag 
+between consecutive fields,
+which indicates non-photometric conditions.
+For good practice, the \ha\ zeropoints of these fields
+were adjusted by solving Eqn.~\ref{eqn:matrix}
+rather than linking them directly to the $r$-band shift.
+
+\subsection{Testing the calibration against SDSS}
+
+Having re-calibrated all fields to an expected accuracy of 3\%,
+we then used an independent survey to validate the results.
+We have been able to exploit SDSS Data Release 9 \citep{Ahn2012}
+for this purpose.
+SDSS DR9 provides several strips at low
+Galactic latitudes,
+providing data across 18\% of the fields in our data release.
+We cross-matched IPHAS fields against 
+stars marked as reliable in the SDSS catalogue\footnote{
+We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
+to obtain photometry from the SDSS {\sc photoprimary} table 
+with criteria {\sc type = star}, {\sc clean = 1} and {\sc score $>$ 0.7}.}
+in the same way as we did for APASS earlier,
+with the exception of using fainter magnitude ranges of 
+$15<r_{\rm SDSS}<18.0$ and $14.5<i_{\rm SDSS}<17.5$.
+This provided us with a set of 1.2 million cross-matched stars.
+
+Colour transformations were obtained using a sigma-clipped linear least squares fit:
+\begin{eqnarray}
+r_{\rm IPHAS} = r_{\rm SDSS} - 0.093 - 0.044(r-i)_{\rm SDSS} \label{eqn:sdss_r} \\
+i_{\rm IPHAS} = i_{\rm SDSS} - 0.318 - 0.095(r-i)_{\rm SDSS}. \label{eqn:sdss_i}
+\end{eqnarray}
+The rms residuals of these transformations are 0.045 and 0.073, respectively.
+The equations are similar to the ones
+previously determined for APASS,
+although the colour terms are slightly larger;
+the throughput curve of the SDSS $i$-band filter 
+appears to be somewhat more sensitive at longer wavelengths
+compared to both the IPHAS and APASS filters.
+
+These global transformations were deemed adequate
+for the purpose of validating our uniform calibration in a statistical sense.
+Separate equations were derived towards different sightlines
+to investigate the effects of varying reddening regimes.
+The transformation coefficients were found 
+to show some variation towards lowly reddened areas,
+which have relatively few numbers
+of (intrinsically) red objects at $r-i > 1$ 
+which can skew the colour term.
+The vast majority of red objects in the global sample
+are those in highly reddened areas however,
+which agree well with the global transformations
+and dominate the statistical appraisal of our calibration.
+%The small number of intrinsically red objects in lowly reddened fields
+%have no impact on the statistical appraisal of our calibration.
+
+Having transformed SDSS magnitudes into the IPHAS system,
+we then computed the median magnitude offset for each IPHAS field
+which contained at least 30 objects with a counterpart
+in the SDSS catalogue.
+This was the case for 2602 fields.
+The median offsets for each of these fields
+are shown in Figs.~\ref{fig:sdss_r}-\ref{fig:sdss_i}.
+The mean offset and standard deviation found 
+was $-0.001\pm0.029$ for $r$
+and $-0.002\pm0.032$ for $i$ (Table~\ref{tbl:offsets_after}).
+In comparison, offsets computed in the identical way
+\emph{before} our re-calibration showed means
+of $+0.016\pm0.088$ and $+0.010\pm0.089$ (Table~\ref{tbl:offsets_after}).
+We conclude that our re-calibration procedure has
+been successful in improving the
+uniformity of the calibration by a factor three
+and has achieved our aim of bringing the
+accuracy down to the level of $\sigma=0.03$~mag.
+
+We warn that the SDSS comparison reveals a number of fields with offsets
+exceeding 0.05~mag (523 fields) or even 0.1~mag (18 fields).
+Such sporadic outliers are consistent with the tails of a Gaussian
+with mean $\sim0$ and $\sigma=0.03$.
+
+In future work, we hope to draw upon
+the PanSTARRS survey \citep{Schlafly2012}
+to further improve the accuracy of our calibration.
+At the time of preparing this work data from PanSTARRS
+had not been made public yet.
+
+\section{Source catalogue generation}
+\label{sec:catalogue}
+
+Having obtained a quality-checked 
+and re-calibrated data set, 
+we now turn to the problem
+of transforming the observations 
+into a user-friendly catalogue.
+The aim of this catalogue is to detail
+the best-available information for each unique source
+in a convenient format,
+including flags to warn about quality issues 
+such as source blending and saturation.
+Compiling the catalogue essentially required four steps:
+\begin{enumerate}
+\item the single-band detection tables 
+produced by the CASU pipeline 
+were augmented with new columns
+and warning flags;
+\item the detection tables were merged into multi-band field catalogues;
+\item the overlap regions of the field catalogues 
+were cross-matched to flag duplicate measurements 
+and identify the best detection 
+of each unique source; and
+\item these primary detections
+were compiled into the final source catalogue.
+\end{enumerate}
+Each of these four steps are now explained.
+
+\subsection{User-friendly columns and warning flags}
+
+As the first step, the detection tables 
+were enhanced by creating new columns.
+This is necessary because the tables 
+generated by the CASU pipeline 
+summarise the detections 
+in their original CCD units,
+e.g. source positions are given in pixel coordinates 
+and photometry in number counts.
+To transform these measurements into
+user-friendly fields,
+we have largely adopted the units and naming conventions
+which are in use at the 
+WFCAM Science Archive \citep[WSA;][]{Hambly2008}
+and the VISTA Science Archive \citep[VSA;][]{Cross2012}.
+These archives curate the near-infrared data from both
+the UKIDSS Galactic Plane Survey \citep[GPS;][]{Lucas2008}
+and the 
+VISTA Variables in the Via Lactea survey \cite[VVV;][]{Minniti2010}.
+Both these surveys provide high-resolution JHK photometry
+in the Galactic Plane.
+There is a significant degree of overlap
+between the footprints of UKIDSS/GPS and IPHAS,
+and hence by adopting a similar catalogue format
+we hope to encourage scientific applications
+which combine both data sets.
+
+A detailed description of each column in our source catalogue
+is given in Appendix~\ref{app:columns}.
+In the remainder of this section we highlight the main features.
+
+Firstly, we note that each source is uniquely identified by an
+IAU-style designation of the form ``IPHAS2\ JHHMMSS.ss+DDMMSS.s''
+(cf. column \emph{name} in Appendix~\ref{app:columns}),
+where ``IPHAS2'' refers to the present
+data release and the remainder of the string
+denotes the J2000 coordinates in sexagesimal format.
+For convenience, the coordinates
+are also included in decimal degrees
+(columns \emph{ra} and \emph{dec})
+and in the Galactic coordinate system
+(columns \emph{l} and \emph{b}).
+We have also included an internal object identifier string 
+of the form ``\#run-\#ccd-\#detection''
+(e.g. ``64738-3-6473''),
+which documents the INT exposure number (\#run),
+the CCD number (\#ccd),
+and the row number in the CASU detection table (\#detection)
+-- see columns \emph{rDetectionID},
+\emph{iDetectionID}, \emph{haDetectionID}.
+
+Photometry is provided based on the 2.3-arcsec diameter circular aperture
+by default (columns \emph{r}, \emph{i}, \emph{ha}).
+The choice of this aperture size as the default 
+is based on a trade-off between concerns 
+about small number statistics and centroiding errors
+for small apertures on one hand,
+and diminishing signal-to-noise ratios and source confusion
+for large apertures on the other hand.
+The user is not restricted to this choice because
+the catalogue also provides magnitudes
+using three alternative aperture sizes:
+the peak pixel height 
+(columns \emph{rPeakMag}, \emph{iPeakMag}, \emph{haPeakMag}),
+a circular 1.2-arcsec-diameter aperture 
+(\emph{rAperMag1}, \emph{iAperMag1},
+ \emph{haAperMag1}) and
+a 3.3-arcsec-diameter aperture 
+(\emph{rAperMag3}, \emph{iAperMag3},
+ \emph{haAperMag3}).
+
+Each of these magnitude measurements have been
+corrected for the flux lost outside of their respective apertures,
+using a correction term which is inferred from the
+mean shape of the PSF measured locally in the CCD frame.
+In the case of a point source,
+the four alternative magnitudes are expected
+to be consistent with each other
+within the photon noise uncertainties
+(which are given in columns \emph{rErr}, \emph{rPeakMagErr},
+\emph{rAperMag1Err}, \emph{rAperMag3Err}, etc).
+When this is not the case,
+it is likely that the source is either
+an extended object or that it has
+been incorrectly measured as a result of
+source blending or a rapidly spatially-varying nebulous background.
+In \S\ref{sec:qualitycriteria} we will explain that the consistency
+of the magnitude measurements in the different apertures
+can be used as a criterion for selecting stellar objects
+with reliable photometry from the catalogue.
+
+The brightness of each object as a function of increasing
+aperture size is also used by the CASU pipeline to provide
+a discrete star/galaxy/noise classification flag
+(\emph{rClass}, \emph{iClass}, \emph{haClass})
+and a continuous stellarness-of-profile statistic
+(\emph{rClassStat}, \emph{iClassStat}, \emph{haClassStat}).
+For convenience, we have combined
+these single-band morphological measures
+into band-merged class probabilities and flags
+(\emph{pStar}, \emph{pGalaxy}, \emph{pNoise},
+\emph{mergedClass}, \emph{mergedClassStat})
+using the merging scheme in use at the WSA\footnote{Explained at
+http://surveys.roe.ac.uk/wsa/www/gloss\_m.html \#gpssource\_mergedclass
+}.
+
+Information on the quality of each detection is included 
+in a series of additional columns.
+We draw attention to three useful flags
+which warn about the likely presence of systematics:
+\begin{enumerate}
+\item The \emph{saturated} column is used to flag sources
+for which the peak pixel height exceeds 55000 counts,
+which is typically the case for stars brighter than 12-13th magnitude.
+Although the pipeline attempts to extrapolate the brightness of
+saturated stars based on the shape of their PSF,
+such extrapolation is prone to errors
+and we do not recommend their use.
+\item The \emph{deblend} column is used to flag sources 
+which partially overlap with a nearby neighbour.
+Although the pipeline applies a deblending procedure
+to such objects, the procedure is currently applied separately
+in each band and hence the $r$-$i$ and $r$-\ha\ colours
+of such objects are prone to errors.
+\item The \emph{brightNeighb} column is used to flag sources which are located
+within 5 arcmin from an object brighter than $V=7$ 
+according to the Bright Star Catalogue (BSC; Hoffleit et al. 1991), 
+or within 10 arcmin if the neighbour is brighter than $V=4$.
+Such very bright stars are known to cause systematic errors
+and spurious detections as a result of stray light 
+and diffraction spikes.
+\end{enumerate}
+In addition to the above, we also created warning flags for internal bookkeeping.
+For example, we flagged detections which fell in the strongly vignetted regions of the focal plane,
+which were truncated by CCD edges,
+or which were otherwise affected by bad pixels in the detector.
+We will explain below that none of such detections 
+have been included in the catalogue
+-- an alternative detection was available in essentially all these situations
+because of the IPHAS field pair strategy --
+and hence these internal warning flags do not appear
+in the final source catalogue.
+
+Finally, we note that basic information on the observing conditions
+is included (\emph{fieldID}, \emph{fieldGrade}, \emph{night}, \emph{seeing}).
+A table containing more detailed quality control information,
+indexed by \emph{fieldID}, is made available on our website.
+
+\subsection{Band-merging the detection tables}
+
+The second step in compiling the source catalogue
+is to merge the contemporaneous trios
+of $r$, $i$, \ha\ detection tables
+into multi-band field catalogues.
+This required a positional matching procedure 
+to link sources between the three bands
+based on their position on the sky.
+We used the \textsc{tmatchn} function 
+of the \textsc{stilts} software for this purpose,
+which allows rows from multiple tables to be matched \citep{Taylor2006}.
+The result of the procedure is a band-merged catalogue
+in which each row corresponds to a group of linked $r$/$i$/\ha\ detections
+which satisfy a maximum matching distance criterion in a pair-wise sense.
+Sources for which no counterpart was identified
+are retained in the catalogue as single-band detections.
+
+We employed a maximum matching distance of 1~arcsec,
+which was chosen based on a trade-off between 
+completeness and reliability.
+On one hand, a matching distance larger than 1~arcsec 
+was found to allow too many spurious and unrelated sources 
+to be linked. 
+On the other hand, a value smaller than 1~arcsec 
+would pose problems for very faint sources 
+with large centroiding errors, 
+and would occasionally fail to link detections near CCD corners
+where the astrometric solution can 
+show systematic errors which exceed 0.5~arcsec.
+The position offsets between the $r$ and $i$/\ha\ detections
+have been included in the catalogue 
+and can hence be tightened by the user if necessary
+(columns \emph{iXi}, \emph{iEta}, \emph{haXi}, \emph{haEta}).
+We note that also UKIDSS/GPS adopted 
+a maximum matching distance of 1 arcsecond 
+for similar reasons \citep{Hambly2008}.
+
+The resulting band-merged catalogues were found
+to be reliable for the vast majority of fields.
+We warn that source blending and confusion is unavoidable
+for faint objects in the Galactic Plane however;
+in \S\ref{sec:discussion} we will show
+that 19\% of the sources in our catalogue
+are flagged as blended objects (column \emph{deblend})
+and their band-merged data should be treated with care
+because they may have fallen victim to source confusion
+during the band-merging step.
+
+\subsection{Selecting the primary detections}
+
+We explained earlier that the survey contains repeat observations
+of identical sources as a result of overlaps in the data-taking pattern.
+Amongst all sources in the reliable magnitude range $13<r<19$,
+we find that 65\% were detected twice and 25\% were detected three times or more.
+Only 9\% were detected once.
+Unsurprisingly, their spatial distribution traces
+the inter-CCD gaps and footprint edges.
+
+The principal aim of this data release is to provide 
+reliable photometry at a single epoch,
+and hence we have decided
+to focus on providing the magnitudes
+and coordinates using only the \emph{best-available} 
+detection of each object -- 
+hereafter referred to as the \emph{primary} detection.
+Although overlapping fields could have been co-added 
+to gain a small improvement in depth, 
+we have decided against this for two reasons.
+Firstly, combining the information from multiple epochs
+would make the photometry of variable stars difficult to interpret.
+Secondly, co-adding would cause the image quality to degrade towards the mean,
+which is a draw-back for crowded fields.
+
+Anyone interested in the alternative detections of a source
+-- hereafter called the \emph{secondary} detections --
+can nevertheless obtain this information in two ways.
+To begin with, whenever a secondary detection was observed 
+within 10 minutes of the primary,
+the magnitudes of that secondary detection
+have been included in the catalogue
+(columns \emph{sourceID2}, \emph{fieldID2}, 
+\emph{r2}, \emph{i2}, \emph{ha2},
+\emph{rErr2}, \emph{iErr2}, \emph{haErr2}, \emph{errBits2}.
+This is the case for 66\% of the sources brighter than $r<20$
+due to the IPHAS field pair observing pattern.
+In addition, the complete set of detection tables -- one for each exposure -- 
+is made accessibly on our website to allow other uses of the data.
+A user-friendly catalogue of secondary detections 
+has not been compiled at present 
+but may be part of a future data release.
+
+The primary detection is defined as the entry in the 
+set of band-merged field catalogues which provides 
+the most reliable information for a unique source.
+Primary detections have been selected using a so-called \emph{seaming} procedure
+which has been adapted from the algorithm developed for the WSA\footnote{http://surveys.roe.ac.uk/wsa/dboverview.html\#merge}.
+In brief, the first step of the procedure is to identify all the duplicate detections
+by cross-matching the overlap regions of all field catalogues,
+again using a maximum matching distance of 1\arcsec.
+The duplicate detections for each unique source are then ranked according to
+to (i) filter coverage, (ii) quality score (column \emph{errBits}),
+and (iii) the average seeing of stars in the CCD frame rounded to 0.2~arcsec.
+If this ranking scheme reveals multiple `winners' of identical quality,
+then the one that was observed closest to the optical axis of the camera is chosen.
+
+\subsection{Compiling the final source catalogue}
+
+As the final step, the primary detections that have been
+selected above are compiled
+into the 98-column source catalogue
+that is described in Appendix~\ref{app:columns}
+and made available on line.
+The entire list of sources naturally includes 
+a significant number of spurious entries
+as a result of the very sensitive detection levels
+that are employed by the CASU pipeline by default.
+To limit the size of the source catalogue,
+we have decided to enforce three basic criteria
+which must be met for a candidate source
+to be included in the catalogue:
+\begin{enumerate}
+\item the source must have been detected at SNR$>5$ in at least
+one of the bands, i.e. it is required that at least one of
+\emph{rErr}, \emph{iErr} or \emph{haErr} is smaller
+than 0.2 mag;
+\item the shape of the source must not be an obvious
+cosmic ray or noise artefact, i.e. we require
+either \emph{pStar} or \emph{pGalaxy} to be
+greater than 20\%;
+\item the source must not have been detected in one of the strongly
+vignetted corners of the detector, 
+not have had any known bad pixels in the aperture,
+and not have been on the edge of one of the CCDs,
+i.e. we require the \emph{errBits} quality score
+to be smaller than 64.
+\end{enumerate}
+
+A total of 219 million primary detections satisfied
+the above criteria and have been included in the catalogue.
+Amongst these, 158 million objects 
+are detected in all three bands (72\%),
+30 million are detected in two bands (14\%),
+and 31 million entries are single-band detections (14\%).
+Roughly half of the single-band detections were made in the $i$-band.
+This is likely explained by the fact that the $i$-band is least
+affected by interstellar extinction and can occasionally pick up
+highly-reddened objects which are otherwise lost in $r$/\ha.
+
+\section{Discussion}
+\label{sec:discussion}
+
+Having explained how the catalogue was generated,
+we now offer an overview of its properties
+by discussing  
+(i) the recommended quality criteria,
+(ii) the typical photometric uncertainties,
+and (iii) the source densities and the frequency of source blending.
+
+\subsection{Recommended quality criteria}
+\label{sec:qualitycriteria}
+
+Like any other photometric survey,
+the majority of the objects in our catalogue
+are faint sources observed near the detection limits:
+55\% of the entries in the catalogue
+are fainter than $r > 20$
+and 18\% are even fainter than $r > 21$.
+The measurements of such faint objects
+are naturally prone to large
+random and systematic uncertainties,
+for example, an inaccurately subtracted background
+will introduce a proportionally larger systematic error
+in a faint object.
+Most scientific applications will require a set of
+quality criteria to be enforced for the purpose
+of removing faint and low-quality objects.
+
+The choice of quality criteria is often a complicated
+trade-off between completeness on one hand
+and accuracy on the other.
+To aid users we have listed two sets of
+recommended quality criteria 
+in Tables~\ref{tab:reliable} and \ref{tab:veryreliable}.
+
+Firstly, Table~\ref{tab:reliable} details
+a set of minimum quality criteria
+which should benefit most applications
+which require reliable $r-i$ and $r-$\ha\ colours
+without removing more than $\sim$80\%
+of the sources brighter than $r < 19$.
+The listed criteria are designed to 
+(i) remove low-SNR sources, 
+(ii) remove saturated sources,
+and (iii) remove objects for which the 2.3~arcsec diameter
+aperture magnitudes are inconsistent 
+with the alternative 1.2~arcsec diameter aperture measurements.
+The last criterion is a proxy
+for identifying objects which are affected
+by inaccurate background subtraction
+or failed source deblending.
+A total of 86 out of 219 million sources 
+(39\%) satisfy all the criteria listed in Table~\ref{tab:reliable}
+and are hereafter referred to as ``reliable''.
+For convenience, the catalogue contains a boolean column
+named \emph{reliable} which flags these objects
+and makes their selection easy.
+
+\begin{table*}
+\vspace{2cm}
+
+\begin{tabular}{p{8cm}lp{6cm}}
+  \hline
+  Quality criterion & Rows passed & Description \\
+  \hline
+  rErr\,$<$\,0.1 {\sc and} iErr\,$<$\,0.1 {\sc and} haErr\,$<$\,0.1 &
+  109 million (50\%) &
+  Require the photon noise to be less than 
+  0.1 mag in all bands (i.e. SNR$>$10).
+  This implicitly requires a detection in all three bands.  \\
+  $r>13$ {\sc and} $i>12$ {\sc and} \ha\,$>12.5$ {\sc and not} \emph{saturated} &
+  158 million (72\%) &
+  The brightness must not exceed the nominal saturation limit
+  and the peak pixel height must not exceed 55\,000 counts.
+  Again, this implicitly requires a detection in all three bands.
+  \\
+  $|\mathrm{r}- \mathrm{rAperMag1}| 
+  < 3\sqrt{\mathrm{rErr}^2 + \mathrm{rAperMag1Err}^2} 
+  + 0.03$ &
+  176 million (80\%) &
+  Require the $r$ magnitude measured 
+  in the default 2.3\arcsec\ diameter aperture
+  to be consistent with the measurement 
+  made in the smaller 1.2\arcsec\ aperture,
+  albeit tolerating a 0.03 mag systematic error.
+  This will reject sources for which the background
+  subtraction or the deblending procedure
+  was not performed reliably. \\
+  $|\mathrm{i}- \mathrm{iAperMag1}| 
+  < 3\sqrt{\mathrm{iErr}^2 + \mathrm{iAperMag1Err}^2} 
+  + 0.03$ &
+  183 million (84\%) &
+  Same as above for $i$. \\
+  $|\mathrm{ha}- \mathrm{haAperMag1}| < 
+  3\sqrt{\mathrm{haErr}^2 + \mathrm{haAperMag1Err}^2} 
+  + 0.03$ &
+  158 million (72\%) &
+  Same as above for \ha. \\
+  \hline
+  All of the above (flagged as {\bf\emph{reliable}}) &
+  86 million (39\%) & \\
+  \hline
+\end{tabular}
+\caption{Recommended minimum quality criteria 
+for selecting objects with reliable colours 
+from the IPHAS DR2 source catalogue. 
+86~million entries in the catalogue (39\%)
+satisfy all the criteria listed in this table.
+For convenience, these have been flagged in the catalogue
+using the column named \emph{reliable}.}
+\label{tab:reliable}
+
+\vspace{2cm}
+
+\begin{tabular}{p{8cm}lp{6cm}}
+  \hline
+  Quality criterion & Rows passed & Description \\
+  \hline
+   reliable &
+   86 million (39\%) &
+   The object must satisfy the criteria listed in Table~\ref{tab:reliable}. \\
+
+   pStar $>$ 0.9 &
+   145 million (66\%) &
+   % pStar > 0.89 is identical to mergedClass == -1
+   The object must appears as a perfect point source,
+   as inferred from comparing its Point Spread Function (PSF)
+   with the average PSF measured in the same CCD. \\
+
+   {\sc not} \emph{deblend} &
+   177 million (81\%) &
+   The source must appear as a single, unconfused object. \\
+
+   {\sc not} \emph{brightNeighb} &
+   216 million (99\%) &
+   There is no star brighter than $V < 4$ within 10 arcmin, 
+   or brighter than $V < 7$ within 5 arcmin.
+   Such very bright stars cause scattered light and diffraction spikes,
+   which may add systematic errors to the photometry
+   or even trigger spurious detections. \\  
+  \hline
+
+  All of the above (flagged as {\bf\emph{veryReliable}}) &
+  59 million (27\%) & \\
+  \hline
+\end{tabular}
+\caption{Additional quality criteria which are recommended
+for applications which require very reliable colours
+at the expense of completeness. 
+For convenience, the sources which satisfy the criteria listed
+in this table have been flagged in the catalogue
+using the column named \emph{veryReliable}.}
+\label{tab:veryreliable}
+
+\vspace{2cm}
+\end{table*}
+
+For applications which require
+an even higher standard of reliability,
+a further set of additional quality criteria
+are listed in Table~\ref{tab:veryreliable}.
+These criteria are designed to ensure that
+(i) the object appeared as a perfect point source,
+(ii) the object was not blended with a nearby neighbour,
+and (ii) the object was not located near a very bright star.
+59 million sources (27\%) satisfy
+these additional criteria 
+and are hereafter referred to as ``very reliable''.
+Again, the catalogue contains a boolean column
+named \emph{veryReliable} which flags these objects.
+
+Figure~\ref{fig:magdist} compares the $r$-band magnitude
+distribution for objects with and without the 
+\emph{reliable} and \emph{veryReliable} criteria applied. 
+We find that 81\% of the sources 
+in the magnitude range $13 < r < 19$
+are considered \emph{reliable},
+which drops to 72\% in the range $19 < r < 20$
+and 9\% at $r>20$.
+In contrast, only 54\% of the sources 
+in the magnitude range $13 < r < 19$
+are considered \emph{veryReliable}.
+The stricter criteria filter out a lot
+of objects at early Galactic longitudes 
+where source blending is a common problem
+(we will demonstrate this in \S\ref{sec:densities}).
+The \emph{veryReliable} flag should hence
+only be used in applications which require very reliable photometry
+at the expense of completeness,
+which might be the case for e.g. spectroscopic target selection.
+
+\begin{figure}
+    \includegraphics[width=0.5\textwidth]{./plots/magdist/magdist-r.pdf} 
+    \caption{r-band magnitude distribution for all objects in the catalogue 
+    (light grey), for objects flagged as \emph{reliable} 
+    according to the criteria set out in Table~\ref{tab:reliable} (grey),
+    and for objects flagged as \emph{veryReliable} 
+    following Table~\ref{tab:veryreliable} (dark grey).
+    The magnitude distributions for $i$ and \ha\
+    look identical, apart from being shifted
+    by about 1 and 0.5 mag towards brighter magnitudes,
+    respectively.}
+    \label{fig:magdist}
+\end{figure}
+
+It is easy to see how the quality criteria
+may be adapted to be more tolerant.
+For example, by raising the allowed photometric uncertainties
+from 0.1 mag to 0.2 mag one can retrieve 42 million candidate sources.
+
+\subsection{Photometric uncertainties}
+
+Figure~\ref{fig:uncertainties} shows the mean photometric
+uncertainties as a function of magnitude for each band.
+We find that the uncertainties typically
+reach 0.1 mag near 20.5 in $r$ 
+and 19.5 in $i$/\ha when the default 2.3\arcsec\ aperture is used.
+At this point we note that the average colour
+of objects in the survey is
+$1.06\pm0.12$ for ($r$-$i$) and $0.44\pm0.03$ for ($r$-\ha).
+The better depth of $r$ is hence compensated
+by the fact that stars tend to have 
+brighter magnitudes in $i$ and \ha.
+
+\begin{figure}
+    \includegraphics[width=0.5\textwidth]{./plots/errors/uncertainties.pdf} 
+    \caption{Mean photometric uncertainties
+             for $r$ (top), $i$ (middle) and \ha\ (bottom).
+             Data points shown are the average values of
+             columns \emph{rErr}, \emph{iErr} and \emph{haErr}
+             in the catalogue, 
+             and the errorbars show the standard deviations.
+             The dashed and solid lines indicate 
+             the 10$\sigma$ and 5$\sigma$ limits, respectively.
+             These uncertainties are based only on the (Poissonian)
+             photon noise and hence this figure does not show
+             systematic or calibration uncertainties.}
+    \label{fig:uncertainties}
+\end{figure}
+
+The uncertainties shown in Fig.~\ref{fig:uncertainties}
+are the random errors based on the expected Poissonian photon noise.
+Systematics, such as calibration and deblending errors,
+are not included.
+To appraise the extent to which our photometry is affected
+by such systematics, we can exploit the
+secondary measurements which were made as part of the
+field-pair observing strategy and are available for 51\%
+of the sources in the catalogue.
+
+Figure~\ref{fig:pairmag} shows the mean residuals between
+the primary and secondary magnitudes
+-- i.e. the average difference between catalogue columns \emph{r}-\emph{r2},
+\emph{i}-\emph{i2}, \emph{ha}-\emph{ha2} -- as a function of magnitude.
+We find that sources across the magnitude range 
+$13 < r < 17$ are consistent at the level of 5\%
+(i.e. $\sigma_{r-r2} \le 0.05$ mag),
+with the best repeatability
+reached at $r=14$ ($\sigma_{r-r2}~=~0.041$~mag).
+We draw attention to the fact that brighter stars
+tend to show significantly larger residuals
+-- e.g. $\sigma_{r-r2}~=~0.14$~mag at $r=12$ --
+which is due to saturation effects.
+At the faint end we find residuals which
+show significantly more scatter than would be expected
+from photon noise alone, that is,
+the effects of source blending and background subtraction
+appear to dominate from $\sim$20th magnitude onwards.
+
+\begin{figure*}
+    \vspace{1cm}
+    \includegraphics[width=0.5\textwidth]{./plots/pairmag/pairmag.pdf} 
+    \caption{Photometric repeatability illustrated by plotting
+             the mean residuals between all the primary and secondary detections
+             in the catalogue as a function of magnitude.
+             The best photometric repeatability is reached at $r=14$
+             with $\sigma_{r-r2}~=~0.041$~mag.
+             Note that bright stars at $r<13$ and $i<12$ 
+             show increasing uncertainties due to saturation effects.}
+    \label{fig:pairmag}
+    \vspace{1cm}
+    \includegraphics[width=0.5\textwidth]{./plots/pairmag/pairmag-reliable.pdf} 
+    \caption{Same as Figure~\ref{fig:pairmag},
+    except that only the subset of sources
+    flagged as \emph{veryReliable} are now included.
+    We find that applying the quality criteria
+    has improved the photometric repeatability significantly.
+    The best repeatability is again reached at $r=14$
+    but has reduced to $\sigma=0.028$ mag.
+    The quality criteria have also been successful
+    at removing objects with large systematics at the bright
+    and faint ends.}
+    \label{fig:pairmag_reliable}
+\end{figure*}
+
+In Figure~\ref{fig:pairmag_reliable} we show 
+a similar comparison of the primary and secondary detections,
+but this time we have only include sources which are flagged
+as \emph{veryReliable} in the catalogue
+(i.e. not saturated, not confused, not near bright stars, etc.)
+We find that the average residuals are significantly better
+for this subset of the catalogue. Sources across the magnitude range 
+$13 < r < 17$ are consistent at the level of 0.03 mag,
+and the best repeatability is again reached at $r=14$
+with $\sigma_{r-r2}~=~0.028$~mag.
+We conclude that the \emph{veryReliable} quality criteria are effective
+in reducing the systematic errors to
+the same level as the accuracy of the
+global photometric calibration (cf. \S\ref{sec:calibration}).
+Moreover, the large systematics at the bright and faint end
+have disappeared.
+
+\subsection{Source densities and blending problems}
+\label{sec:densities}
+
+The mean source density as a function of Galactic longitude
+is shown in Figure~\ref{fig:density} (thick blue line).
+The densities were computed by counting the 
+number of sources in $1^\circ$-wide longitude bins
+across which covered the entire latitude range $-5^\circ<b<+5^\circ$.
+Unsurprisingly, we find the average source densities to increase
+towards the Galactic centre.
+For example, the average source density near $l\simeq 30^\circ$
+is roughly 30\,000 objects per square degree,
+which is five times more than the density
+found near $l\simeq 180^\circ$.
+
+In addition to the global trend, 
+there are significant variations in the source density on smaller scales.
+For example,  we find a significant drop near the constellations 
+of Cygnus ($l\simeq 80$) and Aquila ($l\simeq 40$),
+which are regions known to be affected
+by high levels of foreground extinction.
+Dark clouds are visible towards these constellation by eye,
+and they are often referred to as ``the Great Rift''.
+
+We warn however that the densities reported here 
+have not been corrected for survey completeness
+or differences in the observing conditions across the survey.
+For example, the dip in the density near $l\simeq140^\circ$
+is an artificial feature caused by gaps
+in the footprint coverage (which are apparent in Fig.~\ref{fig:footprint}).
+In a forthcoming paper,
+we aim to calibrate the IPHAS-based source densities
+by injecting artificial stars into the IPHAS images
+and measuring their recovery rate (Farnhill et al., in preparation).
+Indeed IPHAS has the potential to offer
+calibrated, two-dimensional stellar density maps
+which can be used to constrain detailed models of our Galaxy,
+but it is beyond the scope of the present work.
+
+\begin{figure*}
+    \includegraphics[width=\textwidth]{./plots/density/density.pdf} 
+    \caption{Mean number density of sources in the catalogue 
+    as a function of Galactic longitude, 
+    with and without blended sources included. 
+    The densities shown were computed by counting the sources 
+    at each longitude between $-5^\circ<b<+5^\circ$ (upper blue line).
+    We also show the densities based on only counting those sources 
+    for which the \emph{deblend} flag is {\sc false}, 
+    i.e. unconfused sources for which the CASU pipeline did not have to apply 
+    a deblending procedure (lower red line). 
+    }
+    \label{fig:density}
+\end{figure*}
+
+In Figure~\ref{fig:density} we also shows the density
+of non-blended sources (thin red line).
+These are sources for which the \emph{deblend} flag is {\sc false},
+i.e. sources for which the CASU pipeline did not have to apply 
+a deblending procedure to separate the flux
+originating from two or more overlapping objects.
+We find a strong correlation between the source density
+and the fraction of sources affected by source blending.
+For example, only NN\% of the sources are blended
+at $l>90^\circ$, whereas NN\% are blended at $l<90^\circ$.
+
+As we explained earlier, blended sources must be used with caution.
+Firstly, the deblending-procedure crucially depends
+on the local PSF being measured accurately. 
+Secondly, blended sources may are likely candidates to have fallen victim
+to source confusion during the band-merging procedure.
+In future work we hope to investigate the use of
+more advanced PSF-fitting routines
+in which sources are measured simultaneously across all bands,
+perhaps guided by an external list of sources provided
+by near-infrared surveys or Gaia data.
+
+\section{Demonstration}
+\label{sec:demonstration}
+
+We conclude this paper by demonstrating how the unique
+$r$-$i$/$r$-\ha\ colour-colour diagram offered by this catalogue
+can readily be used to
+(i) characterise the extinction regime at different sightlines, and
+(ii) identify \ha\ emission-line objects.
+
+\subsection{Colour-colour and colour-magnitude diagrams}
+
+The survey's unique $r$-\ha\ colour,
+when combined with $r$-$i$,
+has been shown to provide simultaneous constraints 
+on intrinsic stellar colour and interstellar extinction \citep{Drew2008}. 
+That is, the main sequence in the $r$-$i$/$r$-\ha\ diagram
+runs in a direction that is different from the reddening vector,
+because the $r$-\ha\ colour tends to act as a coarse proxy for spectral type
+and is less sensitive to reddening than $r$-$i$.
+As a result, the distribution of a stellar population
+in the IPHAS colour-colour diagram
+can offer a handle on the properties and extinction regime
+along a line of sight.
+
+This is demonstrated in Figures \ref{fig:l180}, \ref{fig:l45} \& \ref{fig:l30},
+where we present three sets of IPHAS colour/magnitude diagrams
+towards three distinct sightlines
+located at Galactic longitudes $180^\circ$, $45^\circ$ and $30^\circ$,
+respectively.
+Each figure contains all the sources which are flagged as \emph{veryReliable}
+and are located in a region of one square degree 
+centred on the coordinates indicated in the diagram
+(i.e. within a radius of $0.564^\circ$ from the indicated sightline).
+For clarity, we have imposed the additional criterion
+for photometric uncertainties to be smaller than 0.05 mag in each band
+(corresponding to a magnitude limit near $\sim$19th magnitude, effectively).
+
+Each of the diagrams reveals a well-defined locus,
+which demonstrates the health of the catalogue and the global calibration
+for investigating stellar populations across wide areas.
+We have annotated the colour-colour diagrams
+by showing the position of the unreddened main sequence (thin solid line),
+the unreddened giant branch (thick solid line)
+and the reddening track for an A0V-type star (dashed line)
+-- all three are based on the \cite{Pickles1998} library of empirical spectra
+tabulated for IPHAS by \cite{Drew2005}.
+In the colour-magnitude diagrams we only show the reddening vector
+together with the unreddened 1~Gyr isochrone due to \cite{Bressan2012},
+which are made available for the IPHAS filter system through a
+popular on line tool hosted by the Observatory of Padova (http://stev.oapd.inaf.it/cmd).
+The isochrone and reddening vector has been placed
+at an arbitrary distance of 2~kpc.
+
+Each of the sightlines reveals a stellar population
+with distinct characteristics.
+Firstly, towards the Galactic anti-centre at $l=180^\circ$ (Fig.~\ref{fig:l180}) 
+we find a population dominated by lowly-reddened 
+main sequence stars and an apparent absence of reddened giants.
+%consistent with the sightline extinction of XX from Schlegel (cite). 
+In contrast, closer towards the galactic
+centre at $l=45^\circ$ (Fig.~\ref{fig:l45})
+we find a wealth of reddened objects
+which appear to be separated into a lowly
+and a highly reddened component,
+perhaps revealing two distinct parts of the Galaxy.
+Finally, in one of our earliest sightlines at $l=30^\circ$
+we find a very high number of extremely reddened giants
+in addition to an unreddened population
+of foreground dwarfs.
+
+The number density of stars in the colour-colour
+and colour-magnitude space
+can be compared against population synthesis models
+to create three-dimensional maps of the extinction
+across several kpc \citep{Sale2010,Sale2012}.
+Such an extinction map based on our catalogue
+is to be released in a separate paper
+that accompanies this data release (Sale et al., in preparation).
+
+\begin{figure*}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=0.5\textwidth]{./plots/ccd-180-3.pdf} 
+        \includegraphics[width=0.5\textwidth]{./plots/cmd-180-3.pdf}
+    \end{minipage}
+    \caption{Colour-colour and colour-magnitude diagram (left and right panel)
+    showing sources flagged as \emph{veryReliable}
+    located in an area of one square degree
+    centred near the Galactic anti-centre 
+    at $(l,b)=(180^\circ,+3^\circ)$.
+    The colour-colour diagram shows the
+    position of the main sequence (thin solid line),
+    giant stars (thick solid line)
+    and the reddening track for an A0V-type star (dashed line)
+    based on the \citet{Pickles1998} library of empirical spectra.
+    The colour-magnitude only show the reddening vector
+    along with the unreddened 1~Gyr isochrone due to \citet{Bressan2012},
+    which has been placed at an arbitrary distance of 2~kpc for reference.
+    This is one of the least reddened sightlines
+    in the survey %(Av XX, Schlegel et al. XXXX)
+    and hence the observed stellar population appears to be dominated 
+    by lowly reddened main sequence stars.}
+    \label{fig:l180}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=0.5\textwidth]{./plots/ccd-45-2.pdf}
+        \includegraphics[width=0.5\textwidth]{./plots/cmd-45-2.pdf}
+    \end{minipage}
+    \caption{Same as above for $(l,b)=(45^\circ,+2^\circ)$,
+    which is one of the highest-density sightlines in the survey,
+    revealing a population of stars with a reddening distribution
+    that appears to be bi-model.}
+    \label{fig:l45}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=0.5\textwidth]{./plots/ccd-30-0.pdf}
+        \includegraphics[width=0.5\textwidth]{./plots/cmd-30-0.pdf} 
+    \end{minipage}
+    \caption{Same as above for $(l,b)=(30^\circ,0^\circ)$.
+    This is one of the most reddened sightlines in the survey.
+    %(XXX, Schlegel et al XXXX).
+    }
+    \label{fig:l30}
+\end{figure*}
+
+\subsection{Identifying \ha\ emission-line objects}
+
+A primary motivation for carrying
+out the survey 
+was to enable the discovery of 
+new emission-line objects across the Galactic Plane.
+\ha\ in emission is a well-known tracer
+for stars in the short-lived pre- or
+post-main sequence stages of their evolution,
+and hence IPHAS aims to allow larger, deeper
+and more statistically robust samples of such rare objects
+to be established.
+The survey-wide identification and analysis 
+of such stars is beyond the scope of the present work,
+but in this section we demonstrate how the
+catalogue may be used for this purpose.
+
+An initial list of candidate \ha-emitters
+based on the first IPHAS data release has previously
+been presented by \cite{Witham2008}. 
+Because no global uniform calibration was available
+at the time, \citeauthor{Witham2008} employed 
+a sigma-clipping technique to select objects with
+large, outlying $r$-\ha\ colours.
+In contrast, the new catalogue
+allows objects to be picked out
+from the $r$-$i$/$r$-\ha\ colour-colour diagram
+using model-based colour criteria
+rather than a statistical procedure.
+In what follows we demonstrate this ability 
+by selecting candidate emission-line objects
+towards a small region in the sky.
+
+The target of our demonstration is Sh 2-82:
+a 5~arcmin-wide H{\sc ii} region located near $(l,b)=(53.55^\circ, 0.00^\circ)$
+in the constellation of Sagitta.
+Nicknamed by amateur astronomers as the ``Little Cocoon Nebula'',
+Sh 2-82 is ionised by 
+the $\sim$10th magnitude star HD\,231616
+% Georgelin1973 claims B0III, others claim B0V
+with spectral type B0V/III
+\citep{Georgelin1973,Mayer1973,Hunter1990}.
+The ionising star has been placed at a likely distance of 1.5-1.7 kpc
+based on the photometric parallax
+\citep{Mayer1973,Lahulla1985,Hunter1990}.
+
+Figure~\ref{fig:mosaic_iphas} shows a 20-by-15 arcmin
+colour mosaic centred on Sh 2-82,
+composed of our \ha\ (red channel),
+$r$ (green channel),
+and $i$ (blue channel) images.
+The ionising star can be seen as the bright object
+near the centre of the H{\sc ii} region,
+which is surrounded by a faint reflection nebula
+and several dark cloud filaments.
+For comparison, Figure~\ref{fig:mosaic_spitzer} shows
+a mosaic of the identical region 
+as seen by the Spitzer Space Telescope
+in the mid-infrared. The Spitzer image
+reveals a bubble-shaped structure of warm dust
+which surrounds the entire H{\sc ii} region.
+
+\begin{figure*}
+    \begin{minipage}[b]{0.8\linewidth}
+        \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-iphas.pdf} 
+    \end{minipage}
+\caption{IPHAS-based mosaic of H{\sc ii} region Sh 2-82,
+composed of \ha\ (red channel), $r$ (green channel) and $i$ (blue channel). Yellow triangles show the position of candidate \ha-emitters
+which have been selected from the colour-colour diagram
+in Figure~\ref{fig:emitters}. Note that the H{\sc ii} region is surrounded by a faint blue/green reflection nebula
+and dark cloud filaments.}
+\label{fig:mosaic_iphas}
+    \begin{minipage}[b]{0.8\linewidth}
+        \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-spitzer.pdf} 
+    \end{minipage}
+    \caption{Star-forming region Sh 2-82 as seen in the mid-infrared
+    by the Spitzer Space Telescope. The mosaic is composed of the 24\,\micron\ (red), 8.0\,\micron\ (green) and 4.5\,\micron\ (blue) bands.
+    The image reveals a bubble-shaped structure which surrounds the {\sc Hii} region that is seen in the IPHAS mosaic of the same region (Figure~\ref{fig:mosaic_iphas}).}
+    \label{fig:mosaic_spitzer}
+\end{figure*}
+
+Figure~\ref{fig:emitters} presents
+the IPHAS colour-colour diagram for the region covered by the
+mosaic images.
+Gray circles show all objects in the region
+which are brighter than $r<20$
+and have been flagged as \emph{reliable}
+in our catalogue.
+The diagram also shows the unreddened main sequence (solid line)
+and the expected position of unredded main-sequence stars
+with \ha\ in emission at a strength of EW=$-10{\rm \AA}$ (dashed line),
+taken from the colour simulations due to \citet{Barentsen2011a}.
+Six stars are found to lie confidently above the 
+dashed line at the level of $3\sigma$ 
+(i.e. the distance between the dashed line is larger than
+three times the uncertainty on the $r$-\ha\ colour).
+These reliable candidate \ha-emitters
+are marked by red triangles in the colour-colour diagram
+and their photometry is detailed in Appendix~\ref{app:emitters}.
+
+\begin{figure}
+  \includegraphics[width=0.45\textwidth]{./plots/sh2-82-ccd.pdf}
+    \caption{$r$-$i$/$r$-\ha\ diagram for the rectangular region of 
+    20-by-15 arcmin centred on the H{\sc ii} region Sh 2-82,
+    which is the area shown in Figure~\ref{fig:mosaic_iphas}.    
+    The diagram shows all objects in the catalogue
+    which have been flagged as \emph{reliable} and are brighter
+    than $r<20$ (grey circles).
+    The unreddened main sequence is indicated by a solid line,
+    while the main sequence for stars with an \ha\ emission line
+    strength of $-10\,\rm{\AA}$ EW is indicated by a dashed line,
+    following the colour simulations due to \citet{Barentsen2011a}.
+    Red triangles indicate objects which have been identified as
+    as highly likely \ha-emitters (see text).}
+    \label{fig:emitters}
+\end{figure}
+
+The spatial distribution of our six candidate emission-line objects
+is marked by yellow triangles in the colour mosaic (Fig.~\ref{fig:mosaic_iphas}).
+They are likely to be genuine young stars for two reasons.
+Firstly, two of our candidates have recently been identified as likely
+candidate Young Stellar Objects (YSO)
+in an investigation of the region by \cite{Yu2012}.
+In their study, the authors used 2MASS and Spitzer data
+to select likely young stars by selecting objects with
+circumstellar disks based on the infrared colour excess.
+Secondly, we find that the four remaining objects are 
+also detected in the Spitzer-based image,
+although their colours are less extreme
+than those identified as likely YSOs by \citeauthor{Yu2012}.
+
+Prior to IPHAS, this region was essentially unstudied
+at faint magnitudes in visible light.
+\cite{Lahulla1985} reported magnitudes for 8 stars in the optical at 
+$V < 15$. In contrast, the IPHAS catalogue offers photometry
+for NN stars in the region down to $r<20$.
+This demonstrates the ability of IPHAS to providing complimentary
+for the wealth of poorly-studied star-forming regions
+at low Galactic latitudes,
+which have been unveiled in recent years
+by the observed wealth of star-forming ``bubbles''
+at mid-infrared wave%\cite{Bica2003} was the first to identify a
+%near-infrared cluster
+%associated with this region using near-infrared 2MASS data,
+%and only recently has this cluster been confirmed
+%in a short investigation which combined 2MASS
+%and Spitzer data \cite{Yu2012}.
+%Many of the young stars identified by \citeauthor{Yu2012}
+%can be seen as pink and red stars in Fig.~\ref{fig:mosaic_spitzer}.lengths \cite[e.g.][]{Churchwell2006,Simpson2012}.
+%In fact the cluster of young stars associated with Sh 2-82
+%was only first reported by \cite{Bica2003},
+%remained unstudied until the study by \cite{Yu2012}.
+
+%Amongst the six emitters, N are detected at SNR X
+%in the Spitzer X micron band.
+
+%This demonstrates how IPHAS is providing complimentary
+%data for some of the thousands of star-forming ``bubbles'' have been unveiled at mid-infrared wavelengths \cite[e.g.][]{Churchwell2006,Simpson2012}.
+%Optically-unveiled members of such star-forming regions
+%may help to constrain extinctions and distances towards such regions \citep{Barentsen2013}.
+%Moreover, in \cite{Barentsen2011a} we showed
+%how the history of star formation in a region,
+%and the possible triggered nature, may be discussed.
+%This is particularly true
+
+%TODO: annotate -10 EW text on plot?
+%TODO: mention Gaia, e.g. IPHAS for transient identification?
+
+%\subsection{Caveats and lessons learnt}
+%\cite{Bica2003} was the first to identify a
+%near-infrared cluster
+%associated with this region using near-infrared 2MASS data,
+%and only recently has this cluster been confirmed
+%in a short investigation which combined 2MASS
+%and Spitzer data \cite{Yu2012}.
+%Many of the young stars identified by \citeauthor{Yu2012}
+%can be seen as pink and red stars in Fig.~\ref{fig:mosaic_spitzer}.
+
+\section{Conclusions and future work}
+\label{sec:conclusions}
+
+A new data release for the IPHAS survey was presented,
+taking the coverage up to over 90\% of the Northern Galactic Plane 
+at $|b|<5^\circ$
+and providing a uniform photometric calibration
+for the first time.
+We explained the data reduction and quality control procedures that
+were applied, described and tested the new global photometric calibration,
+and detailed the construction of the user-friendly source catalogue.
+
+The observations included in this release
+were found to achieve a median seeing of 1.1 arcsec
+and a $5\sigma-$depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$, \ha$=20.3\pm 0.3$.
+The global calibration and photometric repeatability
+is accurate at the level of $\sigma=0.03$ mag,
+providing a significant improvement over the 
+previous data release.
+The source catalogue provides the best-available
+single-epoch astrometry and photometry
+for 219~million unique sources.
+
+The data-taking strategy developed for IPHAS
+have since been reapplied to carry out a companion INT/WFC survey called UVEX
+in U/$g$/$r$ \citep{Groot2009},
+and also southern counterpart to IPHAS and UVEX 
+is being carried out in $u$/$g$/$r$/$i$/\ha\ 
+called VPHAS+ (Drew et al, in press).
+We hope to re-use the experience gained by this data release
+to create similar releases for these companion surveys.
+
+In future work, we aim to draw upon the PanSTARRS photometric
+survey to further improve the accuracy of our global calibration.
+We will also aim to correct the photometry for the
+radial field distortions.
+%We may also produce a catalogue which details all detections.
+
+%TODO: mention Gaia?
+
+\section*{Data access and source code}
+\label{sec:dataaccess}
+
+The catalogue is made available through the Vizier
+catalogue tool (http://vizier.u-strasbg.fr),
+where it is known as the ``IPHAS DR2 Source Catalogue''
+(catalogue ID ???).
+The full catalogue can also be downloaded in its entirety
+from the IPHAS website (www.iphas.org) as a collection 
+of binary FITS tables which comprise 50\,GB,
+which is accompanied by a script
+to ingest the data into a PostgreSQL database.
+Our website also provides access to the pipeline-processed
+imaging data, which we have updated to include
+the re-calibrated DR2 zeropoint in the image headers.
+
+The source code that was used to generate
+the catalogue is available at
+https://github.com/barentsen/iphas-dr2
+
+\section*{Acknowledgments}
+
+The IPHAS survey was carried out 
+at the Isaac Newton Telescope (INT).
+The INT is operated on the island of La Palma
+by the Isaac Newton Group
+in the Spanish Observatorio del Roque de los Muchachos
+of the Instituto de Astrofisica de Canarias.
+All data were processed 
+by the Cambridge Astronomical Survey Unit,
+at the Institute of Astronomy in Cambridge.
+
+Preparation of the catalogue was eased greatly
+by a number of software packages,
+including the Python modules
+AstroPy \citep{Astropy},
+APLpy, NumPy and SciPy,
+the PostgreSQL database software,
+the TOPCAT and STILTS packages \citep{Taylor2005,Taylor2006},
+and the Montage software maintained by NASA/IPAC.
+We also made use of the SIMBAD, Vizier and Aladin \citep{Aladin} tools operated at CDS, Strasbourg, France.
+
+This research made extensive use of
+several complementary photometric surveys.
+Our global calibration was aided
+by the AAVSO Photometric All-Sky Survey (APASS),
+funded by the Robert Martin Ayers Sciences Fund.
+The calibration was tested against the
+Sloan Digitized Sky Survey (SDSS),
+funded by the Alfred P.
+Sloan Foundation, the Participating Institutions, the National Science Foundation, the U.S.
+Department of Energy, the National Aeronautics and Space Administration, the Japanese Monbukagakusho, the Max Planck Society, and the Higher Education Funding Council for England.
+The astrometric pipeline reduction made
+significant use of the Two Micron All Sky Survey (2MASS),
+which is a joint project 
+of the University of Massachusetts
+and the Infrared Processing and Analysis Center/
+California Institute of Technology,
+funded by NASA and the NSF.
+
+GB and JED acknowledge the support of a grant
+from the Science \& Technology Facilities Council
+of the UK (STFC, ref ST/J001335/1).
+HJF is in receipt of an STFC postgraduate studentship.
+
+\bibliographystyle{mn2e}
+\bibliography{dr2}
+
+\appendix
+\input{app_columns.tex}
+
+\section{Candidate emission-line objects towards Sh 2-82}
+\label{app:emitters}
+\begin{table*}
+    \begin{tabular}{lccc}
+    \toprule
+    Name & $r$ & $i$ & \ha  \\
+    \midrule
+IPHAS2 J192954.40+181026.1& $17.69\pm0.01$ & $16.12\pm0.01$ & $16.19\pm0.01$ \\
+IPHAS2 J193011.01+182051.2& $18.55\pm0.02$ & $16.95\pm0.02$ & $17.31\pm0.02$ \\
+IPHAS2 J193021.52+181954.5& $19.72\pm0.05$ & $17.94\pm0.03$ & $18.47\pm0.04$ \\
+IPHAS2 J193024.45+181938.3& $19.31\pm0.04$ & $17.55\pm0.02$ & $17.99\pm0.03$ \\
+IPHAS2 J193033.00+181609.3& $18.25\pm0.01$ & $16.91\pm0.01$ & $16.92\pm0.01$ \\
+IPHAS2 J193042.48+182317.4& $19.96\pm0.03$ & $18.11\pm0.03$ & $18.48\pm0.03$ \\
+    \bottomrule
+    \end{tabular}
+    \caption{Candidate \ha-emitters towards Sh 2-82.}
+    \label{tbl:emitters}
+\end{table*}
+
+\label{lastpage}
+
 \end{document}

### 2014-03-12 paper/iphas-dr2.tex
Now (45,2)

@@
 
-In contrast, closer towards the galactic
-centre at $l=45^\circ$ (Fig.~\ref{fig:l45})
-we find a wealth of reddened objects
-which appear to be separated into a lowly
-and a highly reddened component,
-perhaps revealing two distinct parts of the Galaxy.
+In contrast, 
+closer towards the galactic centre at $l=45^\circ$ (Fig.~\ref{fig:l45}) we find a wealth of reddened objects.
+In the colour magnitude diagram it is clear that the stars are split into two distinct groups,
+with one significantly redder than the other.
+The bluer group is composed of main sequence stars,
+with the slope of this group in the colour magnitude diagram attributable to the significantly increasing extinction.
+Meanwhile the redder group is principally composed of red giant stars.
+As these stars are intrinsicly brighter,
+at some given apparent magnitude they will be substantially further away than their main sequence counterparts.
+Given that, along this sightline,
+extinction continues to increase with distance,
+the red giants we observe will be subject to more reddening than a main sequence star exhibiting the same apparent magnitude.
+Therefore, the clear split between the two groups arises due to the combination of increasing extinction and the distinct absolute magnitude ranges of main sequnce and giant branch stars.
 

### 2014-03-12 paper/iphas-dr2.tex
finally (30,0)

@@
 In contrast, 
-closer towards the galactic centre at $l=45^\circ$ (Fig.~\ref{fig:l45}) we find a wealth of reddened objects.
+closer towards the galactic centre at $l=45^{\circ}$ (Fig.~\ref{fig:l45}) we find a wealth of reddened objects.
 In the colour magnitude diagram it is clear that the stars are split into two distinct groups,
@@
 
-Finally, in one of our earliest sightlines at $l=30^\circ$
+Finally, in one of our earliest sightlines at $l=30^{\circ}$
 we find a very high number of extremely reddened giants
@@
 of foreground dwarfs.
+In contrast to the sightline at $l=45^{\circ}$,
+there is no clear group of giant stars visible in the colour magnitude diagram of Fig.~\ref{fig:l30}.
+Though the red clump stars are manifested as a slight over density which sits roughly 0.4~mags redder than the A0V reddening track.
+At $(l,b)=(45^{\circ}, 2^{\circ})$ the giant stars observed exhibited a relativly narrow range of reddenings as the lay beyond most or essentially all of the Galactic dust.
+At $(l,b)=(30^{\circ}, 0^{\circ})$ this is not the case:
+even at the substantial distances at which we can observe reddened giant stars extinction is continuing to rise and we still find ourselves well within the Galactic dust disc.
+It is also apparent that the $(r-i)$ width of both the M dawrfs and early A dwarfs is greater than that in Fig.~\ref{fig:l45}.
+This is indicative of a steeper rise in reddening,
+both within several hundred parsecs (M dwarfs)
+and within a few kiloparsecs (early A dwarfs).
 

### 2014-03-12 paper/iphas-dr2.tex
Detailed applications

@@
 
-The number density of stars in the colour-colour
-and colour-magnitude space
-can be compared against population synthesis models
-to create three-dimensional maps of the extinction
-across several kpc \citep{Sale2009,Sale2012}.
-Such an extinction map based on our catalogue
-is to be released in a separate paper
+Here we have presented only a short qualitative description of the information obtainable from IPHAS colour-colour and colour-magnitude plots.
+A more rigourous quantitative analysis of the IPHAS catalogue
+can be undertaken to estimate both the stellar density distribution in the Milky Way \citep{Sale2010}
+and to create detailed three-dimensional maps of the extinction across several kpc \citep{Sale2009,Sale2012}.
+An extinction map based on our catalogue is to be released in a separate paper
 that accompanies this data release (Sale et al., in preparation).

### 2014-03-12 paper/iphas-dr2.tex
Updated caption to Fig 13

@@
 closer towards the galactic centre at $l=45^{\circ}$ (Fig.~\ref{fig:l45}) we find a wealth of reddened objects.
-In the colour magnitude diagram it is clear that the stars are split into two distinct groups,
+In the colour-magnitude diagram it is clear that the stars are split into two distinct groups,
 with one significantly redder than the other.
 The bluer group is composed of main sequence stars,
-with the slope of this group in the colour magnitude diagram attributable to the significantly increasing extinction.
+with the slope of this group in the colour-magnitude diagram attributable to the significantly increasing extinction.
 Meanwhile the redder group is principally composed of red giant stars.
@@
 In contrast to the sightline at $l=45^{\circ}$,
-there is no clear group of giant stars visible in the colour magnitude diagram of Fig.~\ref{fig:l30}.
-Though the red clump stars are manifested as a slight over density which sits roughly 0.4~mags redder than the A0V reddening track.
-At $(l,b)=(45^{\circ}, 2^{\circ})$ the giant stars observed exhibited a relativly narrow range of reddenings as the lay beyond most or essentially all of the Galactic dust.
+there is no clear group of giant stars visible in the colour-magnitude diagram of Fig.~\ref{fig:l30},
+though the red clump stars are manifested as a slight over density which sits roughly 0.4~mags redder than the A0V reddening track.
+At $(l,b)=(45^{\circ}, 2^{\circ})$ the giant stars observed exhibit a relativly narrow range of reddenings as the lie beyond most or essentially all of the Galactic dust.
 At $(l,b)=(30^{\circ}, 0^{\circ})$ this is not the case:
@@
     which is one of the highest-density sightlines in the survey,
-    revealing a population of stars with a reddening distribution
-    that appears to be bi-model.}
+    revealing two groups of stars in colour-magnitude space.}
     \label{fig:l45}

### 2014-03-12 paper/iphas-dr2.tex
Spell check

@@
 Looking in more detail we can see that the stellar locus is narrower for M-type dwarfs than for earlier types;
-we do not observe M dwarfs expereincing the strongest reddening possible for this sightline.
+we do not observe M dwarfs experiencing the strongest reddening possible for this sightline.
 This implies that extinction is still increasing at distances beyond $\sim 2$~kpc,
@@
 Meanwhile the redder group is principally composed of red giant stars.
-As these stars are intrinsicly brighter,
+As these stars are intrinsically brighter,
 at some given apparent magnitude they will be substantially further away than their main sequence counterparts.
-Given that, along this sightline,
+Given that, along this sight-line,
 extinction continues to increase with distance,
 the red giants we observe will be subject to more reddening than a main sequence star exhibiting the same apparent magnitude.
-Therefore, the clear split between the two groups arises due to the combination of increasing extinction and the distinct absolute magnitude ranges of main sequnce and giant branch stars.
+Therefore, the clear split between the two groups arises due to the combination of increasing extinction and the distinct absolute magnitude ranges of main sequence and giant branch stars.
 
@@
 though the red clump stars are manifested as a slight over density which sits roughly 0.4~mags redder than the A0V reddening track.
-At $(l,b)=(45^{\circ}, 2^{\circ})$ the giant stars observed exhibit a relativly narrow range of reddenings as the lie beyond most or essentially all of the Galactic dust.
+At $(l,b)=(45^{\circ}, 2^{\circ})$ the giant stars observed exhibit a relatively narrow range of reddenings as the lie beyond most or essentially all of the Galactic dust.
 At $(l,b)=(30^{\circ}, 0^{\circ})$ this is not the case:
 even at the substantial distances at which we can observe reddened giant stars extinction is continuing to rise and we still find ourselves well within the Galactic dust disc.
-It is also apparent that the $(r-i)$ width of both the M dawrfs and early A dwarfs is greater than that in Fig.~\ref{fig:l45}.
+It is also apparent that the $(r-i)$ width of both the M dwarfs and early A dwarfs is greater than that in Fig.~\ref{fig:l45}.
 This is indicative of a steeper rise in reddening,
 both within several hundred parsecs (M dwarfs)
-and within a few kiloparsecs (early A dwarfs).
+and within a few kpc (early A dwarfs).
 
 Here we have presented only a short qualitative description of the information obtainable from IPHAS colour-colour and colour-magnitude plots.
-A more rigourous quantitative analysis of the IPHAS catalogue
+A more rigorous quantitative analysis of the IPHAS catalogue
 can be undertaken to estimate both the stellar density distribution in the Milky Way \citep{Sale2010}

### 2014-03-12 paper/iphas-dr2.tex
removing strange duplication of tex

@@
 \end{document}\documentclass[useAMS,usenatbib]{mn2e}
-\usepackage{amssymb,amsmath}
-\usepackage[pdftex]{graphicx}
-\usepackage{longtable} 
-%\usepackage{lscape}
-\usepackage{float}
-\usepackage{booktabs}
-\usepackage{aas_macros}
-
-\def\ha{\mbox{H$\rm \alpha$}}
-\def\arcsec{$''$}
-\def\arcmin{$'$}
-\def\deg{$^{\circ}$}
-\def\micron{\mbox{$\mu$m}}
-
-\title[IPHAS Data Release 2]{The Second Data Release 
-of the INT Photometric H$\alpha$ Survey 
-of the Northern Galactic Plane (IPHAS DR2)}
-
-\author[G. Barentsen
-et. al]{G. Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
-H. J. Farnhill$^{1}$,
-J. E. Drew$^{1}$,
-M. J. Irwin$^{4}$, \newauthor
-E. A. Gonz$\acute{\rm{a}}$lez-Solares$^{4}$,
-R. Greimel$^{2}$,
-B. Mizalski$^{3}$,
-C. Ruhland$^{1}$, \newauthor plus many friends.
-\newauthor\\
-$^{1}$Centre for Astrophysics Research, STRI, University of Hertfordshire, College Lane Campus, Hatfield AL10 9AB\\
-$^{2}$Institute for Geophysics, Astrophysics and Meteorology, Institute of Physics, Karl-Franzens-Universit$\ddot{a}$t Graz, Universit$\ddot{a}$tsplatz 5, 8010 Graz, Austria\\
-$^{3}$Building, Institute, Street Address, City, Code, Country\\
-$^{4}$Institute of Astronomy, Madingley Road, Cambridge CB3 0HA}
-
-\begin{document}
-\date{Current draft typeset \today}
-\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2013}
-
-\maketitle
-
-\label{firstpage}
-
-\begin{abstract} % BACKGROUND, OBJECTIVE, METHODS, RESULTS, CONCLUSIONS
-The INT/WFC Photometric H$\alpha$ Survey 
-of the Northern Galactic Plane (IPHAS)
-is a 1800 deg$^2$ imaging survey
-covering the entire northern Milky Way at $|b| < 5^\circ$
-in the $r$, $i$ and \ha\ filters 
-using the Wide Field Camera (WFC) 
-on the 2.5-meter Isaac Newton Telescope (INT).
-This data release presents the first 
-uniformly-calibrated source catalogue
-to have been extracted from the survey,
-providing single-epoch photometry
-for 219 million unique sources
-across 92\% of the survey.
-The observations were carried out between 2003 and 2012
-at a median seeing of 1.1 arcsec
-to a depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$ and \ha$=20.3\pm 0.3$
-($5\sigma$ limits, Vega system).
-We explain the data reduction 
-and quality control procedures,
-describe and test the new uniform photometric calibration,
-and detail the construction of the source catalogue
-and its quality warning flags.
-We find that the new calibration is accurate to
-$\sigma=0.03$ mag.
-Finally, we demonstrate the ability of the 
-catalogue's unique
-$r$-$i$/$r$-\ha\ colour-colour diagram to
-(i) characterise stellar populations and extinction regimes
-towards different Galactic sightlines
-and (ii) select reliable candidate \ha-emission line objects.
-The catalogue which accompanies this paper
-provides the much-needed visible-light counterpart
-to several infrared surveys of the Galactic Plane,
-including many poorly-studied star-forming regions and ``bubbles'',
-and provides images for some of the most
-crowded regions to be faced by Gaia.
-\end{abstract}
-
-\begin{keywords}
-catalogues, surveys, stars: emission line, Be, Galaxy: stellar content
-\end{keywords}
-
-\section{Introduction}
-
-Since the first data release in 2008, 
-the INT/WFC Photometric H$\alpha$ Survey 
-of the Northern Galactic Plane (IPHAS)
-has provided new insights into the contents and structure 
-of our own backyard, the Milky Way. 
-The original motivation for undertaking 
-this large-scale programme of observation
--- spanning almost a decade 
-and using more than 300 nights 
-at the Isaac Newton Telescope (INT) in La Palma -- 
-was to provide the digital update 
-to the photographic northern H$\alpha$ surveys 
-of the mid-20th century. 
-By increasing the sensitivity 
-with respect to these previous surveys 
-by a factor $\sim$1000 (7 magnitudes), 
-it was envisaged that IPHAS would allow 
-the limited bright samples of Galactic emission line objects 
-available at the outset \citep[e.g.][]{Kohoutek1999}, 
-to be extended into larger, deeper, more statistically-robust samples 
-that in turn could better inform our understanding 
-of the early and late stages of stellar evolution. 
-This aim has begun to be realised through a
-range of published studies including: 
-a preliminary catalogue of candidate emission line objects \citep{Witham2008};
-discoveries of new northern symbiotic stars \citep{Corradi2008, Corradi2010}; 
-new cataclysmic variables \citep{Witham2007}; 
-new groups of young stellar objects \citep{Vink2008,Barentsen2011a};
-along with discoveries of new and remarkable planetary nebulae 
-\citep{Mampaso2006, Corradi2011, Viironen2011}.
-
-Over the years it has become apparent that the legacy of IPHAS 
-will extend beyond the traditional \ha\ applications 
-of identifying emission line stars and nebulae. 
-Through the provision of $r$, $i$ broadband photometry 
-alongside H$\alpha$ data,
-IPHAS has created the opportunity 
-to study Galactic Plane populations 
-in a new way.
-For example, the survey’s unique $r-H\alpha$ colour, 
-when combined with $r-i$,
-has been shown to provide simultaneous constraints 
-on intrinsic stellar colour and interstellar extinction \citep{Drew2008}. 
-This has opened the door 
-to a wide range of Galactic science applications, 
-including the mapping of extinction across the Plane in three dimensions
-and the probabilistic inference of stellar properties
-\citep{Sale2009, Sale2010, Giammanco2011, Sale2012, Barentsen2013}. 
-In effect, the availability of narrowband H$\alpha$ 
-alongside $r, i$ magnitudes
-provides coarse spectral information for huge samples of stars 
-which are otherwise too faint or numerous 
-to be targeted by spectroscopic surveys.
-For such science applications to succeed however, 
-it is vital that the imaging data is transformed 
-into a homogeneously calibrated photometric catalogue, 
-in which quality problems 
-and duplicate detections of the same source 
-are flagged.
-
-The first release of IPHAS data, 
-covering roughly half the survey footprint,
-was made in late 2007 \citep{Gonzalez-Solares2008}. 
-At the time the data were insufficiently complete 
-to support a homogeneously calibrated source catalogue.
-The goal of this paper is to present the next release 
-that takes the coverage up to over 90 percent of the survey area 
-and includes a uniform calibration.
-In this work we aim to
-(i) explain the data reduction 
-and quality control procedures that were applied,
-(ii) describe and test the new global photometric calibration, and 
-(iii) detail the construction and demonstrate the use 
-of the source catalogue that has been extracted 
-from the newly re-calibrated data.
-
-In \S\ref{sec:observations} we start by recapitulating the key points
-of the survey design and strategy.
-In \S\ref{sec:reduction} we describe the data reduction
-and quality control procedures.
-In \S\ref{sec:calibration} we explain the uniform re-calibration
-and test our results against the Sloan Digital Sky Survey (SDSS).
-In \S\ref{sec:catalogue} we explain how the source catalogue was compiled.
-In \S\ref{sec:discussion} we discuss the properties of the catalogue,
-and finally in \S\ref{sec:demonstration} we demonstrate
-the health of the release by demonstrating its scientific exploitation.
-In \S\ref{sec:conclusions} we conclude and outline
-our future ambitions.
-
-\section{Observations}
-\label{sec:observations}
-
-The detailed properties of the IPHAS observing programme 
-have been presented before 
-by \citet{Drew2005} and \citet{Gonzalez-Solares2008}. 
-To set the stage for this release, we briefly remind of some key points.
-IPHAS is a 1800~sq. deg. imaging survey of the northern Galactic Plane, 
-providing images and photometry in Sloan $r, i$ 
-along with narrowband H$\alpha$. 
-It is carried out using the Wide Field Camera (WFC) 
-on the 2.5-meter Isaac Newton Telescope (INT) in La Palma. 
-It is the first digital survey to offer comprehensive CCD photometry
-of point sources in the Galactic Plane at visible wavelengths, 
-and does so down to a limiting magnitude of $\sim$20th.
-The IPHAS footprint on the northern sky spans a box 
-of roughly 180 by 10 degrees, 
-taking in the entire northern Galactic Plane 
-at latitudes $-5^{\circ} < b < +5^{\circ}$ 
-and longitudes $30^{\circ} < l < 215^{\circ}$.
-
-The Wide Field Camera is a mosaic of 4 CCDs 
-that captures a sky area of close to 0.29 square degrees.
-To cover the entire northern Plane with some overlap,
-the survey area was divided into 7635 telescope pointings.
-Each of these pointings is accompanied by an offset position
-at a displacement of $+$5 arcmin in declination 
-and $+$5 arcmin in right ascension,
-to deal with inter-CCD gaps, detector imperfections,
-and to enable quality checks. 
-The basic unit of observation hence
-amounts to $2 \times 3$ exposures, 
-in which each of the 3 survey filters is exposed at 2 offset sky positions, 
-within an elapsed time of 10 minutes.
-We shall refer to the unit of 3 exposures at the same position 
-as a \emph{field},
-and the combination of two fields at a small offset as a \emph{field pair}.
-The survey hence contains 15270 fields grouped into 7635 field pairs.
-To achieve the desired survey depth of 20th magnitude or fainter, 
-the filter exposure times were set at 120 sec (narrowband \ha), 
-30 sec ($r$) and 10 sec ($i$)
-in the majority of the survey observations.\footnote{In 2003 
-the $r$-band exposure time was 10~sec instead of 30~sec,
-and since Oct 2010 the $i$-band exposure time 
-has been increased from 10~sec to 20~sec.}
-
-Data-taking began in the second half of 2003, 
-and every field had been observed at least once by the end of 2008.
-At that time only 76 percent of the field pairs 
-satisfied our minimum quality criteria however,  
-often due to the effects of clouds, poor seeing, or technical faults
-(the quality criteria will be detailed in the next section). 
-Since then, a programme of repeat observations has been in place 
-to improve data quality. 
-As a result, 92\% of the survey 
-now benefits from quality-approved data.
-The most recent observations which are part of this release
-were obtained in November 2012.
-
-\begin{figure*}
-        \includegraphics[width=1\linewidth]{./plots/footprint/footprint_small.png}
-        \caption{Survey area showing the footprints
-        of all the quality-approved IPHAS fields
-        which have been included in this data release.
-        The area covered by each field has been coloured black
-        with a semi-transparent opacity of 20\%,
-        such that regions where fields overlap are darker.
-        The IPHAS strategy is to observe each field twice
-        with a small offset,
-        and hence the vast majority of the area 
-        is covered twice (dominant gray colour).
-        There are small overlaps between all the neighbouring fields
-        which can be seen as a honeycomb-style pattern
-        of dark gray lines across the survey area.
-        Regions with incomplete data are apparent as white gaps (no data) 
-        or as the lightest shade of gray
-        (denoting that only the offset position is missing).}
-        \label{fig:footprint}
-\end{figure*}
-
-Figure~\ref{fig:footprint} shows the footprint 
-of the quality-approved observations included in this work. 
-The fields which remain missing 
--- covering 7 percent of the survey area --
-are predominantly located towards the Galactic anti-center 
-at $l > 120^o$.
-Fields at these longitudes can only be accessed from La Palma 
-in the months of November-December,
-which is when the weather and seeing conditions are often poor
-at the INT and observing attempts have failed repeatedly.
-To enable the survey to be brought to completion, 
-a decision was made recently to limit repeats in this area 
-to individual fields requiring replacement,
-i.e. fresh observations in all 3 filters may only be obtained 
-at one of the two offset positions, 
-if the data for the partner offset has already passed quality control.  
-The catalogue is structured such that it is clear 
-where a contemporaneous observation of both halves of a field pair
-is not available.
-
-\section{Data reduction and quality control}
-\label{sec:reduction}
-
-\subsection{Initial pipeline processing}
-
-All raw data obtained with the INT were transferred
-to the Cambridge Astronomical Survey Unit (CASU) 
-for initial processing and archival.
-The procedures used by CASU were originally devised
-for the INT Wide Field imaging Survey \citep[WFS;][]{McMahon2001,Irwin2005},
-which was a 200 deg$^2$ survey programme carried out 
-between 1998 and 2003 after the WFC was commissioned.
-Because IPHAS uses the same telescope and camera combination,
-we have been able to benefit from the existing WFS pipeline.
-A detailed description of the processing steps 
-is found in \citet{Irwin2001}.
-Its application to IPHAS has previously been described
-by \citet{Drew2005} and \citet{Gonzalez-Solares2008}
-and much of the source code is available 
-on line\footnote{http://casu.ast.cam.ac.uk/surveys-projects/software-release}. 
-In brief, the pipeline takes care of bias subtraction,
-linearity correction, flat-fielding,
-gain correction and de-fringing.
-
-The reduced images are then stored in a multi-extension FITS file 
-with a primary header describing the characteristics
-(position, filter, exposure time, etc.) 
-and four compressed image extensions 
-corresponding to each of the four CCDs.
-Source detection and characterisation is then carried out 
-using the \textsc{imcore} tool \citep{Irwin1985,Irwin1997}.
-The flux of each source is measured using both
-the peak pixel height (i.e. a square 0.33$\times$0.33\arcsec\ aperture)
-as well as a series of circular apertures of increasing diameter 
-(1.2\arcsec, 2.3\arcsec, 3.3\arcsec, 4.6\arcsec\ and 6.6\arcsec).
-
-The local background levels are estimated 
-by computing the sigma-clipped median
-flux in a grid of 64$\times$64 pixels (21$\times$21\arcsec)
-across the image,
-which is then interpolated to obtain an estimate 
-of the background level at each pixel.
-These sky levels are subtracted from the aperture photometry and
--- when required --
-a deblending routine is applied which also attempts to remove
-the contamination from any other nearby sources.
-Whilst this approach works very well 
-across the vast majority of the survey area,
-the Galactic Plane unavoidably contains crowded regions 
-with large numbers of overlapping sources
-or rapidly spatially-varying nebulosity,
-in which case aperture photometry must always be interpreted 
-with caution.
-In \S\ref{sec:catalogue} we will explain that overlapping
-sources to which the deblending routine was applied 
-are flagged in the catalogue using the \emph{deblend} warning flag.
-
-Finally, an astrometric solution is determined
-based on the 2MASS point source catalog \citep{Skrutskie2006},
-which itself is calibrated 
-in the International Celestial Reference System (ICRS).
-A provisional photometric calibration is also provided 
-based on the average zeropoint
-determined from a set of standard stars observed in the same night.
-Sources are classified morphologically
--- stellar, galaxy or noise --
-based on the curve-of-growth determined
-from measuring the source intensity in a series of growing apertures.
-Finally, the resulting source detection tables are also stored 
-in multi-extension FITS files.
-
-At the time of preparing DR2,
-the CASU pipeline had processed
-74\,195 IPHAS exposures 
-in which a total of 1.9~billion \emph{candidate sources} were detected 
-at the sensitive default detection level of 1.25\,$\sigma$
--- unavoidably including spurious detections, artefacts and
-duplicate detections 
-(in \S\ref{sec:catalogue} we will explain
-how these have been removed or flagged).
-The pipelined data set -- comprising 2.5~terabyte of FITS files --
-was then transferred to the University of Hertfordshire
-for the purpose of transforming the raw
-detection tables into a reliable source catalogue which is 
-(i) quality-controlled,
-(ii) homogeneously calibrated, and 
-(iii) contains user-friendly columns and warning flags.
-It is these post-processing steps which are explained below.
-
-\subsection{Quality control}
-\label{sec:qc}
-
-Observing time for IPHAS was obtained
-on a semester-by-semester basis
-through the traditional time allocation committees 
-of the Isaac Newton Group of telescopes,
-which are competitive and invariably over-subscribed.
-For this reason, we attempted to utilise 
-\emph{all} the nights allocated to IPHAS,
-even those which were partially or entirely non-photometric
-or otherwise affected by technical problems 
-(e.g. electronic noise or telescope tracking problems).
-Any unsuitable data that was taken as a result
-of this strategy was subsequently flagged and rejected
-using a series of seven quality criteria,
-which ensure a reliable and homogeneous level of quality
-across the data release:
-
-\begin{figure}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/depth_r.pdf} 
-    \end{minipage}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/depth_i.pdf} 
-    \end{minipage}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/depth_h.pdf} 
-    \end{minipage}
-    \caption{Distribution of the 5$\sigma$ limiting magnitude
-             across all quality-approved survey fields
-             for $r$ (top), $i$ (middle) and \ha\ (bottom).
-             Fields with a limiting magnitude brighter than
-             20th ($r$) or 19th (\ha/$i$) were rejected
-             from the data release.
-             The $r$-band depth is most sensitive 
-             to the presence of the moon above the horizon, 
-             which is evidenced by the wide and bi-model shape
-             of its distribution.}
-    \label{fig:depth}
-\end{figure}
-
-(1) \emph{Depth.} 
-We discarded any exposures for which the $5\sigma$ limiting magnitude 
-was worse than 20th magnitude in the $r$-band
-or worse than 19th in $i$ or \ha. 
-Such data were typically obtained during poor weather or full moon.
-Most observations fared significantly better than these limits.
-Figure~\ref{fig:depth} presents the distribution of limiting magnitudes
-for all quality-approved fields,
-which shows a mean depth of 
-$21.2\pm0.5$ ($r$), $20.0\pm0.3$ ($i$) and $20.3\pm0.3$ (\ha).
-We found that the depth achieved depended 
-most strongly on the presence of the moon,
-which was above the horizon during 62 percent 
-of our observations and explains the wide and bi-modal shape
-of the $r$-band limiting magnitude distribution 
-(top panel in Fig.~\ref{fig:depth}).
-In contrast, the depth attained in $i$ and \ha\ 
-is less sensitive to moonlight
-and the distribution of their depths
-is hence more narrow
-(middle and bottom panel in Fig.~\ref{fig:depth}).
-
-Our minimum limiting magnitude criteria
-have led us to exclude 9\% of the pipelined data.
-We note that much of the excluded data may nevertheless be useful
-for e.g. time-domain studies of bright stars.
-The detection tables for any such discarded data are made
-available through our website (www.iphas.org)
-but are ignored in what follows.
-
-(2) \emph{Ellipticity.} 
-The ellipticity of a point source,
-defined as $e = 1 - b / a$ 
-with $b$ the semi-minor and $a$ the semi-major axis,
-is a morphological measure of the elongation of the point spread function.
-It is expected to be zero (circular) across the field 
-in a perfect imaging system,
-but it is slightly non-zero in any real telescope data 
-due to optical distortions and tracking errors.
-The mean ellipticity across a field in the IPHAS data set 
-is $0.09\pm0.04$.
-There have been sporadic episodes with higher ellipticities however
-due to mechanical glitches in the telescope tracking system.
-For this reason, we rejected exposures in which the mean ellipticity
-across the detector exceeded $e > 0.3$,
-which is when the photometric measurements delivered by the pipeline
-were found to become degraded.
-Only 0.4\% of the exposures were discarded on this basis.
-
-(3) \emph{Seeing.} 
-The survey originally aimed to obtain data 
-at seeing better than 1.7 arcsec.
-This target is currently attained across 86\% of the footprint,
-in particular at early longitudes,
-e.g. 92\% of the fields at $l<120^\circ$ are better than 1.7 arcsec.
-Figure~\ref{fig:seeing} presents the distribution
-of the mean seeing for all the quality-approved fields.
-We find a median value of 1.1~arcsec in $r$/\ha\
-and 1.0~arcsec in $i$.
-In the $r$-band, 90\% of the data is better than 1.5~arcsec
-and 10\% is better than 0.8~arcsec.
-To improve the area covered by this data release,
-we have decided to include the small fraction of data
-that was obtained under seeing up to 2.5 arcsec,
-so that only 1 per cent of the pipelined exposures
-had to be excluded.
-In \S\ref{sec:catalogue} we will explain
-that the catalogue is structured
-such that the information on the seeing is included,
-and that the photometry listed
-is based on the exposures with the best-available seeing.
-
-\begin{figure}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/seeing_r.pdf} 
-    \end{minipage}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/seeing_i.pdf} 
-    \end{minipage}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/seeing_ha.pdf} 
-    \end{minipage}
-    \caption{Seeing distribution across all
-             quality-approved survey fields
-             for $r$ (top), $i$ (middle) and \ha\ (bottom).}
-    \label{fig:seeing}
-\end{figure}
-
-(4) \emph{Photometric repeatability.} 
-The IPHAS field-pair observing strategy 
-ensures that every pointing is immediately followed 
-by an offset pointing at a displacement of $+$5 arcmin in Dec 
-and $+$5 arcmin in RA.
-This allows pairs of images to be checked 
-for the presence of clouds or electronic noise.
-To exploit this information,
-the overlap regions of all field pairs were systematically cross-matched
-to verify the consistency of the photometric measurements
-for stars observed in both pointings.
-We rejected field pairs in which more than 2\% of the stars 
-showed an inconsistent measurement at the level of 0.2 mag,
-or more than 26\% were inconsistent at the level of 0.1 mag.
-These limits were determined empirically by inspecting
-the images and photometry by eye.
-11\% of the data was rejected as part of this step.
-
-(5) \emph{Source density mapping.}
-Spatial maps showing the number density of the detected sources
-down to 20th magnitude were created to verify the health
-of the data and to check for unexpected artefacts.
-%(e.g. bright satellite trails)
-In particular, we created density maps
-which showed the number of \emph{unique} sources
-obtained by cross-matching the detection tables of
-all three bands with a maximum matching distance of 1 arcsec.
-The success of such a cross-matching procedure crucially depends
-on the accuracy of the astrometric solution in each band,
-and hence we were able to detect and correct
-exposures with poor astrometry by inspecting the density map
-for spurious density variations.
-
-(6) \emph{Visual examination.}
-All images and their associated photometric colour/magnitude diagrams
-were inspected by a team of 20 survey team members, 
-such that each image in the data release 
-was looked at by at least three different pairs of eyes.
-Images deemed unsuitable were flagged, investigated and
-excluded from the release if necessary. 
-6\% of the attempts to observe a field
-were placed on a \emph{black-list}
-for this reason, most commonly due to the obvious presence
-of clouds or extreme levels of scattered moonlight.
-Such fields were often rejected as part of 
-one or more of the above quality criteria as well.
-
-(7) \emph{Contemporaneous field data.} 
-Finally, only exposures which are part of a sequence 
-of three consecutive images of the same field (H$\alpha$/$r$/$i$) 
-were considered for inclusion in the release. 
-This ensures that the three bands for a given field
-are observed at nearly the same time --  
-essentially always within 5 minutes.
-An exception was made for 9 fields where the three exposures 
-could not be obtained within the same night
-but for which the time gap did not exceed 48 hours.
-We note that the exact epoch of the magnitude in each band
-is included in the source catalogue
-(columns \emph{rMJD}, \emph{iMJD}, \emph{haMJD}).
-
-The above criteria were satisfied 
-for 14115 out of the 15270 fields (92\%).
-In a few cases more than one successful attempt to observe
-a field was available due to the fact that slightly stricter
-quality criteria were adopted in the initial years of the survey.
-In such cases, only the attempt 
-with the best seeing and depth was selected
-for inclusion in the release, because the focus 
-of this work is to deliver the most reliable
-measurement at a single epoch.
-Those interested in any of the rejected data 
-may nevertheless access the full set of detection tables 
-on line.
-
-\section{Photometric calibration}
-\label{sec:calibration}
-
-Having obtained a quality-approved set of observations,
-we now turn to the problem of placing the data
-onto a uniform photometric scale.
-
-\subsection{Provisional nightly calibration}
-
-For the purpose of providing an initial calibration 
-of the $r$ and $i$ broadband fluxes,
-photometric standard fields were observed every night.
-The standards were chosen from a list based on 
-the \cite{Landolt1992} and Stetson (http://cadcwww.dao.nrc.ca/standards) 
-objects.
-Two or three standard fields were observed 
-during the evening and morning twilight,
-and at intervals of 2-3 hours throughout the night.
-The CASU pipeline automatically identified the observed standards 
-and used them to determine a sigma-clipped average zeropoint \textsc{magzpt}
-for each night and filter,
-such that the number counts $DN$ 
-in the pipeline-corrected CCD frames
-relates to a magnitude $m$ as:
-\begin{equation}
-\begin{split}
-   m  = & \textsc{magzpt} - 2.5 \log_{10}( DN / \textsc{exptime} ) \\
- &  - \textsc{extinct}\cdot(\textsc{airmass}-1) - \textsc{apcor} - \textsc{percorr}
-\label{eqn:mag}
-\end{split}
-\end{equation}
-where \textsc{exptime} is the exposure time in seconds,
-\textsc{extinct} is the atmospheric extinction coefficient 
-(typically 0.09 for $r$ and 0.05 for $i$ in La Palma),
-\textsc{airmass} is the normalised optical path length 
-through the atmosphere and
-\textsc{apcor} is a correction for the flux
-lost outside of the aperture used.
-Finally, \textsc{percorr} is a correction based on the difference
-between the median dark sky for a CCD against the median for all the CCDs 
-and as such is an ancillary correction 
-to account for sporadic gain variations. 
-All these quantities correspond to header keywords in the 
-multi-extension FITS files produced by the CASU pipeline.
-
-The zeropoint was determined such that the resulting magnitude system
-is in the WFC system that uses the SED of Vega 
-as the zero colour reference object. 
-Colour equations were used to transform between the IPHAS passbands 
-and the Johnson-Cousins system 
-of the published standard star photometry.
-The entire procedure has been found to deliver zeropoints which 
-are accurate at the level of 1-2\% 
-in photometric conditions \citep{Gonzalez-Solares2011}.
-
-Unlike the broadbands, 
-standard star photometry is not available in the literature 
-for the H$\alpha$ passband
-and hence there is no formally recognised flux scale 
-for the narrowband.
-We can specify here however 
-that the integrated in-band energy flux for Vega 
-in the IPHAS \ha\ filter 
-is $1.52 \times 10^{-7}$ ergs\,cm$^{-2}$\,s$^{-1}$ 
-at the top of the Earth's atmosphere,
-which is the flux obtained by folding 
-Vega's SED with the filter throughput curve 
-corrected for atmosphere and detector quantum efficiency
-\citep[using the method explained by][]{Drew2005}.
-This is 3.08 magnitudes less than the flux captured 
-by the much broader $r$ band
-which includes the \ha\ band.
-Hence to assure zero colour relative to the broadbands,
-we set the default zeropoint for the narrowband to be:
-\begin{equation}
-\textsc{magzpt}_{H\alpha} = \textsc{magzpt}_r - 3.08.
-\label{eqn:zpha}
-\end{equation}
-
-\subsection{Uniform re-calibration}
-
-Despite the best efforts made to obtain a nightly calibration,
-large surveys naturally possess field-to-field variations
-at the level of 0.1 mag
-due to atmospheric fluctuations during the night
-and imperfections in the pipeline or the instrument
-(e.g. the WFC is known to suffer from sporadic glitches
-in the timing of exposures).
-Such variations need to be corrected for 
-during a global re-calibration procedure.
-Notable past examples include the global re-calibration 
-of the Two Micron All Sky Survey \citep[2MASS;][]{Nikolaev2000},
-the Sloan Digital Sky Survey \citep[SDSS;][]{Padmanabhan2008}
-and the Panoramic Survey Telescope 
-and Rapid Response System survey \citep[Pan-STARRS;][]{Schlafly2012},
-which all achieved photometry 
-that is globally consistent to within 1--2\%.
-
-Surveys which observe identical stars at different epochs
-can use the repeat measurements to ensure a uniform calibration.
-For example, 2MASS attained its global calibration
-by observing six standard fields each hour, 
-allowing zeropoint variations to be tracked 
-over very short timescales \citep{Nikolaev2000}.
-Alternatively, the SDSS and PanSTARRS surveys could benefit
-from revisiting regions in their footprint to 
-carry out a so-called \emph{ubercalibration} procedure,
-in which multiple measurements of stars at different epochs
-are used to fit the calibration parameters
-\citep{Padmanabhan2008,Schlafly2012}.
-
-Unfortunately these schemes cannot be applied
-directly to IPHAS
-for two reasons. 
-Firstly, the survey was carried out 
-in competitive observing time
-on a non-dedicated telescope, 
-rendering the 2MASS approach 
-of observing standards at a very high frequency
-prohibitively expensive
--- in part because standard fields 
-are very scarce within the Galactic Plane.
-Secondly, the aim of IPHAS is to obtain magnitudes at a single epoch
-and hence stars are not normally observed at more than one epoch,
-unless they happen to fall on the narrow overlap region 
-between two neighbouring fields.
-
-Although the IPHAS data does contain a significant number of 
-inter-field repeat measurements
-due to the small overlaps
-between neighbouring fields,
-we have found the information contained
-in these regions to be insufficient
-to constrain the calibration parameters.
-This is because photometry at the extreme edges of the WFC
--- where neighbouring fields overlap -- 
-is itself prone to systematics at the level of 1-2\%.
-The cause of these errors is thought to include 
-the use of twilight sky flats in the pipeline,
-which are known to be imperfect for calibrating stellar photometry 
-due to stray light and vignetting \citep[e.g.][]{Manfroid1995}.
-Moreover, the illumination correction in the overlap regions
-is affected by a radial geometric distortion,
-which causes the pixel scale towards the edges 
-to increase \citep{Gonzalez-Solares2011}.
-Although these systematics are reasonably small within a single field,
-they can combine to cause artificial zeropoint gradients 
-across the survey
-when they are used to constrain a global calibration
-without external constraints.
-
-For these reasons, we have decided not to depend
-on an ubercalibration-type scheme alone,
-and have instead opted to exploit an external reference survey
--- where available --
-to bring the majority of our data onto a uniform calibration.
-This is explained next.
-
-\subsubsection{Correcting zeropoints using APASS}
-
-We have been able to benefit from the
-AAVSO Photometric All-Sky Survey
-(APASS; http://www.aavso.org/apass)
-to bring the vast majority of the survey 
-onto a uniform scale.
-Since 2009,
-APASS has been using two 20~cm-astrographs
-to survey the entire sky down to $\sim$17th magnitude
-in five filters which include Sloan $r$ and $i$.
-The most recent catalogue available 
-at the time of preparing this work was APASS DR7,
-which provided data across roughly half of the IPHAS footprint (Fig.~\ref{fig:apass_r}).
-The photometric accuracy of APASS is currently estimated 
-to be at the level of 3\% (Henden, private communication),
-which is significantly better 
-than the provisional calibration of IPHAS
-for which we estimate the $1\sigma$-error to be $\sim$10\%.
-APASS achieves its uniform accuracy 
-by measuring each star at least two times in photometric conditions
--- along with ample standard fields --
-using the large $3\times3$ square degrees field of view of its detectors.
-
-With the aim of bringing IPHAS to a similar accuracy of $\sim$3\%,
-we used the APASS catalogue to identify and adjust all IPHAS fields 
-which showed a magnitude offset larger than 3\% against APASS.
-For this purpose,
-the $r$- and $i$-band detection tables of each IPHAS field
-were cross-matched against the APASS DR7 catalogue 
-using a maximum matching distance of 1~arcsec.
-The magnitude range was limited to
-$13<r_{\rm APASS}<16.5$ and $12.5<i_{\rm APASS}<16.0$
-in order to avoid sources 
-brighter than the IPHAS saturation limit on one hand, 
-and to avoid sources near the faint detection limit of APASS 
-on the other hand.
-
-The resulting set of $\sim$220\,000 cross-matched stars were then used 
-to derive APASS-to-IPHAS magnitude transformations
-using a linear least-squares fitting routine, 
-which iteratively removed $3\sigma$-outliers to improve the fit.
-The solution converged to:
-\begin{align} 
-r_{\rm IPHAS} = r_{\rm APASS} - 0.121 + 0.032(r-i)_{\rm APASS} \label{eqn:apass_r} \\
-i_{\rm IPHAS} = i_{\rm APASS} - 0.364 + 0.006(r-i)_{\rm APASS} \label{eqn:apass_i}
-\end{align}
-The root mean square (rms) residuals of these transformations 
-are 0.041 and 0.051, respectively.
-The small colour terms in the equations
-indicate that the $r$ and $i$ filters 
-are very similar in both surveys.
-The transformations include a large fixed offset
-which is simply due to the fact that 
-APASS magnitudes are given in the AB system
-and IPHAS uses magnitudes relative to Vega.
-Separate transformations were derived for sightlines 
-with varying extinction properties to investigate the robustness
-of the transformations with respect to different reddening regimes,
-but the variations at these different
-sightlines were found to be insignificant.
-This is not surprising because heavily reddened objects 
-are naturally scarce at $r<16$.
-
-Having transformed APASS magnitudes into the IPHAS system,
-we then computed the median magnitude offset 
-for each field which contained at least 30 cross-matched stars.
-This was the case for 48\% of the fields.
-The mean offset was found to be
-$0.014\pm0.104$ for $r$ and $0.007\pm0.108$ for $i$
-(Table~\ref{tbl:offsets_before}).
-A total of 4596 fields showed a median offset
-exceeding $\pm$0.03~mag in either $r$ or $i$.
-
-We then applied the most important step in our re-calibration scheme,
-which is to adjust the provisional zeropoints of the 4596 aberrant fields
-such that their offset is brought to zero.
-This allowed the mean IPHAS-to-APASS offset 
-to be brought down to $0.000\pm0.011$ in both $r$ and $i$
-(Table~\ref{tbl:offsets_after}).
-The procedure of fitting magnitude transformations and
-correcting the zeropoints was repeated a few times to ensure 
-convergence, which was essentially reached after the first iteration.
-
-\begin{figure*}
-    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_rshift.pdf} 
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_r.pdf} 
-    \caption{Median magnitude offset in the $r$ band between IPHAS and APASS,
-             plotted on a field-by-field basis
-             prior to the re-calibration procedure.
-             Each square represents an IPHAS field
-             which contains at least 30 stars with a counterpart
-             in the APASS DR7 catalogue.
-             The colours denote the median
-             IPHAS-APASS magnitude offset in each field,
-             which was computed after applying the APASS-to-IPHAS
-             transformation to the APASS magnitudes (Eqn.~\ref{eqn:apass_r}).}
-        \label{fig:apass_r}
-    \vspace{1cm}
-    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_rshift.pdf}
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_r.pdf} 
-    \caption{Median magnitude offset in the $r$ band
-             between IPHAS and SDSS after the re-calibration
-             procedure was applied.
-             Again, each square represents an IPHAS field
-             which contains at least 30 stars
-             with a counterpart in the SDSS catalogue.
-             The colours denote the median IPHAS-SDSS magnitude offset
-             in each field,
-             which was computed after applying the SDSS-to-IPHAS
-             transformation to the SDSS magnitudes (Eqn.~\ref{eqn:sdss_r}).}
-    \label{fig:sdss_r}
-\end{figure*}
-
-\begin{figure*}
-	\vspace{2cm}
-    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_ishift.pdf} 
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_i.pdf} 
-    \caption{Same as Figure~\ref{fig:apass_r} for the $i$-band.}
-        \label{fig:apass_i}
-    \vspace{1cm}
-    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_ishift.pdf}
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_i.pdf} 
-    \caption{Same as Figure~\ref{fig:sdss_i} for the $i$-band.}
-    \label{fig:sdss_i}
-\end{figure*}
-
-\begin{table}
-        \begin{center}
-                \begin{tabular}{lcc}
-                        \toprule
-                         {\bf Before re-calibration} & Mean & $\sigma$  \\
-                        \midrule
-                        $r$ (IPHAS - APASS) & +0.014 & 0.104 \\
-                        $i$ (IPHAS - APASS) & +0.007 & 0.108 \\
-                        $r$ (IPHAS - SDSS) & +0.016 & 0.088 \\
-                        $i$ (IPHAS - SDSS) & +0.010 & 0.089 \\
-                        \bottomrule
-                \end{tabular}
-        \caption{Mean magnitude offsets for objects
-                 crossmatched between IPHAS and APASS/SDSS
-                 before the uniform re-calibration.
-                 Eqns.~\ref{eqn:apass_r}-\ref{eqn:sdss_i} were applied
-                 to the APASS/SDSS magnitudes to bring them into the
-                 Vega-based IPHAS system prior to computing
-                 the offsets.)
-                 }
-        \label{tbl:offsets_before}
-
-                \begin{tabular}{lcc}
-                        \toprule
-                         {\bf After re-calibration} & Mean & $\sigma$ \\
-                        \midrule
-                        $r$ (IPHAS - APASS) & +0.000 & 0.011\\
-                        $i$ (IPHAS - APASS) & +0.000 & 0.011 \\
-                        $r$ (IPHAS - SDSS)  & -0.001 & 0.029\\
-                        $i$ (IPHAS - SDSS) & -0.002 & 0.032 \\
-                        \bottomrule
-                \end{tabular}
-        \caption{Same as Table~\ref{tbl:offsets_before}
-                 but computed after the uniform re-calibration
-                 was carried out.}
-        \label{tbl:offsets_after}
-        \end{center}
-\end{table}
-
-\subsubsection{Adjusting fields not covered by APASS}
-
-At the time of writing, the APASS catalogue did not provide 
-sufficient coverage for 7359 of the fields in our data release.
-Fortunately, these fields are pre-dominantly located 
-in the early part of the Galactic Plane (Fig.~\ref{fig:apass_r}),
-which were typically observed during the summer months
-when photometric conditions are more prevalent at the telescope.
-These remaining fields have nevertheless 
-been brought onto the same uniform scale 
-by employing an ubercalibration-style scheme, explained below,
-which minimises the magnitude offsets between stars
-located in the overlap regions with neighbouring fields.
-Although we explained earlier that these overlap regions
-are prone to systematics, the use of APASS enabled us to
-keep the zeropoints of roughly half of the fields fixed,
-which avoids these systematics from combining to introduce
-artificial gradients across the survey.
-In the next section we will show this to be true by validating
-our calibration against SDSS.
-
-A general solution to the problem of minimising the magnitude
-differences between overlapping frames has previously been
-described by~\citet{Glazebrook1994}.
-In brief, there are two fundamental quantities to be
-minimised between each pair of overlapping exposures, 
-denoted by the indices $i$ and $j$. 
-Firstly, the mean magnitude difference between stars in the overlap
-region $\Delta_{ij}=\langle m_i-m_j\rangle=-\Delta_{ji}$ is a local
-constraint. 
-Secondly, to ensure the solution does not stray far 
-from the existing calibration, 
-the difference in zeropoints 
-$\Delta\mathrm{ZP}_{ij}=-\Delta\mathrm{ZP}_{ji}$ 
-between each pair of exposures must also be minimised.
-
-Minimisation of these two quantities is a linear least squares problem 
-because the magnitude $m$ depends linearly on the ZP (Eqn.~\ref{eqn:mag}).
-Hence we can find the ZP shift to be applied to each field 
-by minimising the sum:
-\begin{equation}
-   S = \sum_{i=1}^N \sum_{j=1}^N w_{ij} \theta_{ij} (\Delta_{ij} + a_i - a_j)^2
-   \label{eqn:chi2}
-\end{equation}
-where $i$ denotes the exposure of interest, 
-$j$ an overlapping exposure, 
-$N$ the number of exposures,
-$a_i$ the ZP to solve for 
-and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$), 
-$w_{ij}$ are weights set to the uncertainty in $\Delta_{ij}$
-and $\theta_{ij}$ is an overlap function 
-equal to either 1 if exposures $i$ and $j$ overlap or 0 otherwise. 
-Solving for $a_i$ is equivalent to solving $\partial
-S/\partial a_i=0$ which gives the matrix equation:
-\begin{equation}
-   \sum_{j=1}^N A_{ij} a_j = b_j
-   \label{eqn:matrix}
-\end{equation}
-where 
-\begin{eqnarray}
-   A_{ij} &=& \delta_{ij} \sum_{k=1}^N w_{jk}\theta_{jk} - w_{ij} \theta_{ij},\label{eqn:aij}\\
-   b_i &=& \sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ji} = -\sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ij}.\label{eqn:bi}
-\end{eqnarray}
-This prescription is essentially identical to \citet{Glazebrook1994}.
-%,although we changed the formulation to ensure that $A_{ij}$ is positive definite.
-
-As explained above, we enforce a strong external constraint
-on the solution by keeping the zeropoint fixed 
-for the 6756 fields which have been compared
-and calibrated against APASS earlier.
-We will hereafter refer to these fields as \emph{anchors}.
-It is asserted that the zeropoints $a_i$ of these anchor fields 
-are known and not solved for,
-though they do appear in the vector $b_j$ as constraints.
-In addition to the APASS-based anchors, 
-we selected 3273 additional anchor fields by hand
-to provide additional constraints in regions not covered by APASS.
-These extra anchors were deemed to have reliable zeropoints 
-based on 
-(i) the information contained in the observing logs,
-(ii) the stability of the standard star zeropoints during the night, and
-(iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
-which is located at $\sim$500~meter from the INT.
-
-We then solved Eqn.~\ref{eqn:matrix} for the $r$ and $i$ bands
-separately using the least-squares routine 
-in Python's {\sc scipy.sparse} module for sparse matrix algebra.
-This provided us with corrected zeropoints for the remaining fields,
-which were shifted on average by $+0.02\pm0.11$ in $r$ 
-and $+0.01\pm0.12$ in $i$ compared to their provisional calibration.
-
-We then turned to the uniform calibration of the \ha\ data.
-It is not possible to re-calibrate the narrowband 
-in the same way as the broadbands,
-because the APASS survey does not offer \ha\ photometry.
-We can reasonably assume however,
-that the corrections required for $r$ and \ha\ are identical
-because the \ha\ zeropoints have been derived directly from the
-$r$-band zeropoints during the provisional calibration (Eqn.~\ref{eqn:zpha}).
-Moreover, the IPHAS data-taking pattern ensured 
-that a field's \ha\ and $r$-band exposures
-were taken at essentially the same time, 
-separated only by the $\sim$30-second read-out time required by the WFC.
-We have hence corrected the \ha\ zeropoints 
-by re-using the zeropoint adjustements that were derived for the $r$ band
-in the earlier steps.
-An exception was made for 3101 fields
-for which our quality-control routines revealed
-zeropoint variations which exceeded 0.03~mag 
-between consecutive fields,
-which indicates non-photometric conditions.
-For good practice, the \ha\ zeropoints of these fields
-were adjusted by solving Eqn.~\ref{eqn:matrix}
-rather than linking them directly to the $r$-band shift.
-
-\subsection{Testing the calibration against SDSS}
-
-Having re-calibrated all fields to an expected accuracy of 3\%,
-we then used an independent survey to validate the results.
-We have been able to exploit SDSS Data Release 9 \citep{Ahn2012}
-for this purpose.
-SDSS DR9 provides several strips at low
-Galactic latitudes,
-providing data across 18\% of the fields in our data release.
-We cross-matched IPHAS fields against 
-stars marked as reliable in the SDSS catalogue\footnote{
-We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
-to obtain photometry from the SDSS {\sc photoprimary} table 
-with criteria {\sc type = star}, {\sc clean = 1} and {\sc score $>$ 0.7}.}
-in the same way as we did for APASS earlier,
-with the exception of using fainter magnitude ranges of 
-$15<r_{\rm SDSS}<18.0$ and $14.5<i_{\rm SDSS}<17.5$.
-This provided us with a set of 1.2 million cross-matched stars.
-
-Colour transformations were obtained using a sigma-clipped linear least squares fit:
-\begin{eqnarray}
-r_{\rm IPHAS} = r_{\rm SDSS} - 0.093 - 0.044(r-i)_{\rm SDSS} \label{eqn:sdss_r} \\
-i_{\rm IPHAS} = i_{\rm SDSS} - 0.318 - 0.095(r-i)_{\rm SDSS}. \label{eqn:sdss_i}
-\end{eqnarray}
-The rms residuals of these transformations are 0.045 and 0.073, respectively.
-The equations are similar to the ones
-previously determined for APASS,
-although the colour terms are slightly larger;
-the throughput curve of the SDSS $i$-band filter 
-appears to be somewhat more sensitive at longer wavelengths
-compared to both the IPHAS and APASS filters.
-
-These global transformations were deemed adequate
-for the purpose of validating our uniform calibration in a statistical sense.
-Separate equations were derived towards different sightlines
-to investigate the effects of varying reddening regimes.
-The transformation coefficients were found 
-to show some variation towards lowly reddened areas,
-which have relatively few numbers
-of (intrinsically) red objects at $r-i > 1$ 
-which can skew the colour term.
-The vast majority of red objects in the global sample
-are those in highly reddened areas however,
-which agree well with the global transformations
-and dominate the statistical appraisal of our calibration.
-%The small number of intrinsically red objects in lowly reddened fields
-%have no impact on the statistical appraisal of our calibration.
-
-Having transformed SDSS magnitudes into the IPHAS system,
-we then computed the median magnitude offset for each IPHAS field
-which contained at least 30 objects with a counterpart
-in the SDSS catalogue.
-This was the case for 2602 fields.
-The median offsets for each of these fields
-are shown in Figs.~\ref{fig:sdss_r}-\ref{fig:sdss_i}.
-The mean offset and standard deviation found 
-was $-0.001\pm0.029$ for $r$
-and $-0.002\pm0.032$ for $i$ (Table~\ref{tbl:offsets_after}).
-In comparison, offsets computed in the identical way
-\emph{before} our re-calibration showed means
-of $+0.016\pm0.088$ and $+0.010\pm0.089$ (Table~\ref{tbl:offsets_after}).
-We conclude that our re-calibration procedure has
-been successful in improving the
-uniformity of the calibration by a factor three
-and has achieved our aim of bringing the
-accuracy down to the level of $\sigma=0.03$~mag.
-
-We warn that the SDSS comparison reveals a number of fields with offsets
-exceeding 0.05~mag (523 fields) or even 0.1~mag (18 fields).
-Such sporadic outliers are consistent with the tails of a Gaussian
-with mean $\sim0$ and $\sigma=0.03$.
-
-In future work, we hope to draw upon
-the PanSTARRS survey \citep{Schlafly2012}
-to further improve the accuracy of our calibration.
-At the time of preparing this work data from PanSTARRS
-had not been made public yet.
-
-\section{Source catalogue generation}
-\label{sec:catalogue}
-
-Having obtained a quality-checked 
-and re-calibrated data set, 
-we now turn to the problem
-of transforming the observations 
-into a user-friendly catalogue.
-The aim of this catalogue is to detail
-the best-available information for each unique source
-in a convenient format,
-including flags to warn about quality issues 
-such as source blending and saturation.
-Compiling the catalogue essentially required four steps:
-\begin{enumerate}
-\item the single-band detection tables 
-produced by the CASU pipeline 
-were augmented with new columns
-and warning flags;
-\item the detection tables were merged into multi-band field catalogues;
-\item the overlap regions of the field catalogues 
-were cross-matched to flag duplicate measurements 
-and identify the best detection 
-of each unique source; and
-\item these primary detections
-were compiled into the final source catalogue.
-\end{enumerate}
-Each of these four steps are now explained.
-
-\subsection{User-friendly columns and warning flags}
-
-As the first step, the detection tables 
-were enhanced by creating new columns.
-This is necessary because the tables 
-generated by the CASU pipeline 
-summarise the detections 
-in their original CCD units,
-e.g. source positions are given in pixel coordinates 
-and photometry in number counts.
-To transform these measurements into
-user-friendly fields,
-we have largely adopted the units and naming conventions
-which are in use at the 
-WFCAM Science Archive \citep[WSA;][]{Hambly2008}
-and the VISTA Science Archive \citep[VSA;][]{Cross2012}.
-These archives curate the near-infrared data from both
-the UKIDSS Galactic Plane Survey \citep[GPS;][]{Lucas2008}
-and the 
-VISTA Variables in the Via Lactea survey \cite[VVV;][]{Minniti2010}.
-Both these surveys provide high-resolution JHK photometry
-in the Galactic Plane.
-There is a significant degree of overlap
-between the footprints of UKIDSS/GPS and IPHAS,
-and hence by adopting a similar catalogue format
-we hope to encourage scientific applications
-which combine both data sets.
-
-A detailed description of each column in our source catalogue
-is given in Appendix~\ref{app:columns}.
-In the remainder of this section we highlight the main features.
-
-Firstly, we note that each source is uniquely identified by an
-IAU-style designation of the form ``IPHAS2\ JHHMMSS.ss+DDMMSS.s''
-(cf. column \emph{name} in Appendix~\ref{app:columns}),
-where ``IPHAS2'' refers to the present
-data release and the remainder of the string
-denotes the J2000 coordinates in sexagesimal format.
-For convenience, the coordinates
-are also included in decimal degrees
-(columns \emph{ra} and \emph{dec})
-and in the Galactic coordinate system
-(columns \emph{l} and \emph{b}).
-We have also included an internal object identifier string 
-of the form ``\#run-\#ccd-\#detection''
-(e.g. ``64738-3-6473''),
-which documents the INT exposure number (\#run),
-the CCD number (\#ccd),
-and the row number in the CASU detection table (\#detection)
--- see columns \emph{rDetectionID},
-\emph{iDetectionID}, \emph{haDetectionID}.
-
-Photometry is provided based on the 2.3-arcsec diameter circular aperture
-by default (columns \emph{r}, \emph{i}, \emph{ha}).
-The choice of this aperture size as the default 
-is based on a trade-off between concerns 
-about small number statistics and centroiding errors
-for small apertures on one hand,
-and diminishing signal-to-noise ratios and source confusion
-for large apertures on the other hand.
-The user is not restricted to this choice because
-the catalogue also provides magnitudes
-using three alternative aperture sizes:
-the peak pixel height 
-(columns \emph{rPeakMag}, \emph{iPeakMag}, \emph{haPeakMag}),
-a circular 1.2-arcsec-diameter aperture 
-(\emph{rAperMag1}, \emph{iAperMag1},
- \emph{haAperMag1}) and
-a 3.3-arcsec-diameter aperture 
-(\emph{rAperMag3}, \emph{iAperMag3},
- \emph{haAperMag3}).
-
-Each of these magnitude measurements have been
-corrected for the flux lost outside of their respective apertures,
-using a correction term which is inferred from the
-mean shape of the PSF measured locally in the CCD frame.
-In the case of a point source,
-the four alternative magnitudes are expected
-to be consistent with each other
-within the photon noise uncertainties
-(which are given in columns \emph{rErr}, \emph{rPeakMagErr},
-\emph{rAperMag1Err}, \emph{rAperMag3Err}, etc).
-When this is not the case,
-it is likely that the source is either
-an extended object or that it has
-been incorrectly measured as a result of
-source blending or a rapidly spatially-varying nebulous background.
-In \S\ref{sec:qualitycriteria} we will explain that the consistency
-of the magnitude measurements in the different apertures
-can be used as a criterion for selecting stellar objects
-with reliable photometry from the catalogue.
-
-The brightness of each object as a function of increasing
-aperture size is also used by the CASU pipeline to provide
-a discrete star/galaxy/noise classification flag
-(\emph{rClass}, \emph{iClass}, \emph{haClass})
-and a continuous stellarness-of-profile statistic
-(\emph{rClassStat}, \emph{iClassStat}, \emph{haClassStat}).
-For convenience, we have combined
-these single-band morphological measures
-into band-merged class probabilities and flags
-(\emph{pStar}, \emph{pGalaxy}, \emph{pNoise},
-\emph{mergedClass}, \emph{mergedClassStat})
-using the merging scheme in use at the WSA\footnote{Explained at
-http://surveys.roe.ac.uk/wsa/www/gloss\_m.html \#gpssource\_mergedclass
-}.
-
-Information on the quality of each detection is included 
-in a series of additional columns.
-We draw attention to three useful flags
-which warn about the likely presence of systematics:
-\begin{enumerate}
-\item The \emph{saturated} column is used to flag sources
-for which the peak pixel height exceeds 55000 counts,
-which is typically the case for stars brighter than 12-13th magnitude.
-Although the pipeline attempts to extrapolate the brightness of
-saturated stars based on the shape of their PSF,
-such extrapolation is prone to errors
-and we do not recommend their use.
-\item The \emph{deblend} column is used to flag sources 
-which partially overlap with a nearby neighbour.
-Although the pipeline applies a deblending procedure
-to such objects, the procedure is currently applied separately
-in each band and hence the $r$-$i$ and $r$-\ha\ colours
-of such objects are prone to errors.
-\item The \emph{brightNeighb} column is used to flag sources which are located
-within 5 arcmin from an object brighter than $V=7$ 
-according to the Bright Star Catalogue (BSC; Hoffleit et al. 1991), 
-or within 10 arcmin if the neighbour is brighter than $V=4$.
-Such very bright stars are known to cause systematic errors
-and spurious detections as a result of stray light 
-and diffraction spikes.
-\end{enumerate}
-In addition to the above, we also created warning flags for internal bookkeeping.
-For example, we flagged detections which fell in the strongly vignetted regions of the focal plane,
-which were truncated by CCD edges,
-or which were otherwise affected by bad pixels in the detector.
-We will explain below that none of such detections 
-have been included in the catalogue
--- an alternative detection was available in essentially all these situations
-because of the IPHAS field pair strategy --
-and hence these internal warning flags do not appear
-in the final source catalogue.
-
-Finally, we note that basic information on the observing conditions
-is included (\emph{fieldID}, \emph{fieldGrade}, \emph{night}, \emph{seeing}).
-A table containing more detailed quality control information,
-indexed by \emph{fieldID}, is made available on our website.
-
-\subsection{Band-merging the detection tables}
-
-The second step in compiling the source catalogue
-is to merge the contemporaneous trios
-of $r$, $i$, \ha\ detection tables
-into multi-band field catalogues.
-This required a positional matching procedure 
-to link sources between the three bands
-based on their position on the sky.
-We used the \textsc{tmatchn} function 
-of the \textsc{stilts} software for this purpose,
-which allows rows from multiple tables to be matched \citep{Taylor2006}.
-The result of the procedure is a band-merged catalogue
-in which each row corresponds to a group of linked $r$/$i$/\ha\ detections
-which satisfy a maximum matching distance criterion in a pair-wise sense.
-Sources for which no counterpart was identified
-are retained in the catalogue as single-band detections.
-
-We employed a maximum matching distance of 1~arcsec,
-which was chosen based on a trade-off between 
-completeness and reliability.
-On one hand, a matching distance larger than 1~arcsec 
-was found to allow too many spurious and unrelated sources 
-to be linked. 
-On the other hand, a value smaller than 1~arcsec 
-would pose problems for very faint sources 
-with large centroiding errors, 
-and would occasionally fail to link detections near CCD corners
-where the astrometric solution can 
-show systematic errors which exceed 0.5~arcsec.
-The position offsets between the $r$ and $i$/\ha\ detections
-have been included in the catalogue 
-and can hence be tightened by the user if necessary
-(columns \emph{iXi}, \emph{iEta}, \emph{haXi}, \emph{haEta}).
-We note that also UKIDSS/GPS adopted 
-a maximum matching distance of 1 arcsecond 
-for similar reasons \citep{Hambly2008}.
-
-The resulting band-merged catalogues were found
-to be reliable for the vast majority of fields.
-We warn that source blending and confusion is unavoidable
-for faint objects in the Galactic Plane however;
-in \S\ref{sec:discussion} we will show
-that 19\% of the sources in our catalogue
-are flagged as blended objects (column \emph{deblend})
-and their band-merged data should be treated with care
-because they may have fallen victim to source confusion
-during the band-merging step.
-
-\subsection{Selecting the primary detections}
-
-We explained earlier that the survey contains repeat observations
-of identical sources as a result of overlaps in the data-taking pattern.
-Amongst all sources in the reliable magnitude range $13<r<19$,
-we find that 65\% were detected twice and 25\% were detected three times or more.
-Only 9\% were detected once.
-Unsurprisingly, their spatial distribution traces
-the inter-CCD gaps and footprint edges.
-
-The principal aim of this data release is to provide 
-reliable photometry at a single epoch,
-and hence we have decided
-to focus on providing the magnitudes
-and coordinates using only the \emph{best-available} 
-detection of each object -- 
-hereafter referred to as the \emph{primary} detection.
-Although overlapping fields could have been co-added 
-to gain a small improvement in depth, 
-we have decided against this for two reasons.
-Firstly, combining the information from multiple epochs
-would make the photometry of variable stars difficult to interpret.
-Secondly, co-adding would cause the image quality to degrade towards the mean,
-which is a draw-back for crowded fields.
-
-Anyone interested in the alternative detections of a source
--- hereafter called the \emph{secondary} detections --
-can nevertheless obtain this information in two ways.
-To begin with, whenever a secondary detection was observed 
-within 10 minutes of the primary,
-the magnitudes of that secondary detection
-have been included in the catalogue
-(columns \emph{sourceID2}, \emph{fieldID2}, 
-\emph{r2}, \emph{i2}, \emph{ha2},
-\emph{rErr2}, \emph{iErr2}, \emph{haErr2}, \emph{errBits2}.
-This is the case for 66\% of the sources brighter than $r<20$
-due to the IPHAS field pair observing pattern.
-In addition, the complete set of detection tables -- one for each exposure -- 
-is made accessibly on our website to allow other uses of the data.
-A user-friendly catalogue of secondary detections 
-has not been compiled at present 
-but may be part of a future data release.
-
-The primary detection is defined as the entry in the 
-set of band-merged field catalogues which provides 
-the most reliable information for a unique source.
-Primary detections have been selected using a so-called \emph{seaming} procedure
-which has been adapted from the algorithm developed for the WSA\footnote{http://surveys.roe.ac.uk/wsa/dboverview.html\#merge}.
-In brief, the first step of the procedure is to identify all the duplicate detections
-by cross-matching the overlap regions of all field catalogues,
-again using a maximum matching distance of 1\arcsec.
-The duplicate detections for each unique source are then ranked according to
-to (i) filter coverage, (ii) quality score (column \emph{errBits}),
-and (iii) the average seeing of stars in the CCD frame rounded to 0.2~arcsec.
-If this ranking scheme reveals multiple `winners' of identical quality,
-then the one that was observed closest to the optical axis of the camera is chosen.
-
-\subsection{Compiling the final source catalogue}
-
-As the final step, the primary detections that have been
-selected above are compiled
-into the 98-column source catalogue
-that is described in Appendix~\ref{app:columns}
-and made available on line.
-The entire list of sources naturally includes 
-a significant number of spurious entries
-as a result of the very sensitive detection levels
-that are employed by the CASU pipeline by default.
-To limit the size of the source catalogue,
-we have decided to enforce three basic criteria
-which must be met for a candidate source
-to be included in the catalogue:
-\begin{enumerate}
-\item the source must have been detected at SNR$>5$ in at least
-one of the bands, i.e. it is required that at least one of
-\emph{rErr}, \emph{iErr} or \emph{haErr} is smaller
-than 0.2 mag;
-\item the shape of the source must not be an obvious
-cosmic ray or noise artefact, i.e. we require
-either \emph{pStar} or \emph{pGalaxy} to be
-greater than 20\%;
-\item the source must not have been detected in one of the strongly
-vignetted corners of the detector, 
-not have had any known bad pixels in the aperture,
-and not have been on the edge of one of the CCDs,
-i.e. we require the \emph{errBits} quality score
-to be smaller than 64.
-\end{enumerate}
-
-A total of 219 million primary detections satisfied
-the above criteria and have been included in the catalogue.
-Amongst these, 158 million objects 
-are detected in all three bands (72\%),
-30 million are detected in two bands (14\%),
-and 31 million entries are single-band detections (14\%).
-Roughly half of the single-band detections were made in the $i$-band.
-This is likely explained by the fact that the $i$-band is least
-affected by interstellar extinction and can occasionally pick up
-highly-reddened objects which are otherwise lost in $r$/\ha.
-
-\section{Discussion}
-\label{sec:discussion}
-
-Having explained how the catalogue was generated,
-we now offer an overview of its properties
-by discussing  
-(i) the recommended quality criteria,
-(ii) the typical photometric uncertainties,
-and (iii) the source densities and the frequency of source blending.
-
-\subsection{Recommended quality criteria}
-\label{sec:qualitycriteria}
-
-Like any other photometric survey,
-the majority of the objects in our catalogue
-are faint sources observed near the detection limits:
-55\% of the entries in the catalogue
-are fainter than $r > 20$
-and 18\% are even fainter than $r > 21$.
-The measurements of such faint objects
-are naturally prone to large
-random and systematic uncertainties,
-for example, an inaccurately subtracted background
-will introduce a proportionally larger systematic error
-in a faint object.
-Most scientific applications will require a set of
-quality criteria to be enforced for the purpose
-of removing faint and low-quality objects.
-
-The choice of quality criteria is often a complicated
-trade-off between completeness on one hand
-and accuracy on the other.
-To aid users we have listed two sets of
-recommended quality criteria 
-in Tables~\ref{tab:reliable} and \ref{tab:veryreliable}.
-
-Firstly, Table~\ref{tab:reliable} details
-a set of minimum quality criteria
-which should benefit most applications
-which require reliable $r-i$ and $r-$\ha\ colours
-without removing more than $\sim$80\%
-of the sources brighter than $r < 19$.
-The listed criteria are designed to 
-(i) remove low-SNR sources, 
-(ii) remove saturated sources,
-and (iii) remove objects for which the 2.3~arcsec diameter
-aperture magnitudes are inconsistent 
-with the alternative 1.2~arcsec diameter aperture measurements.
-The last criterion is a proxy
-for identifying objects which are affected
-by inaccurate background subtraction
-or failed source deblending.
-A total of 86 out of 219 million sources 
-(39\%) satisfy all the criteria listed in Table~\ref{tab:reliable}
-and are hereafter referred to as ``reliable''.
-For convenience, the catalogue contains a boolean column
-named \emph{reliable} which flags these objects
-and makes their selection easy.
-
-\begin{table*}
-\vspace{2cm}
-
-\begin{tabular}{p{8cm}lp{6cm}}
-  \hline
-  Quality criterion & Rows passed & Description \\
-  \hline
-  rErr\,$<$\,0.1 {\sc and} iErr\,$<$\,0.1 {\sc and} haErr\,$<$\,0.1 &
-  109 million (50\%) &
-  Require the photon noise to be less than 
-  0.1 mag in all bands (i.e. SNR$>$10).
-  This implicitly requires a detection in all three bands.  \\
-  $r>13$ {\sc and} $i>12$ {\sc and} \ha\,$>12.5$ {\sc and not} \emph{saturated} &
-  158 million (72\%) &
-  The brightness must not exceed the nominal saturation limit
-  and the peak pixel height must not exceed 55\,000 counts.
-  Again, this implicitly requires a detection in all three bands.
-  \\
-  $|\mathrm{r}- \mathrm{rAperMag1}| 
-  < 3\sqrt{\mathrm{rErr}^2 + \mathrm{rAperMag1Err}^2} 
-  + 0.03$ &
-  176 million (80\%) &
-  Require the $r$ magnitude measured 
-  in the default 2.3\arcsec\ diameter aperture
-  to be consistent with the measurement 
-  made in the smaller 1.2\arcsec\ aperture,
-  albeit tolerating a 0.03 mag systematic error.
-  This will reject sources for which the background
-  subtraction or the deblending procedure
-  was not performed reliably. \\
-  $|\mathrm{i}- \mathrm{iAperMag1}| 
-  < 3\sqrt{\mathrm{iErr}^2 + \mathrm{iAperMag1Err}^2} 
-  + 0.03$ &
-  183 million (84\%) &
-  Same as above for $i$. \\
-  $|\mathrm{ha}- \mathrm{haAperMag1}| < 
-  3\sqrt{\mathrm{haErr}^2 + \mathrm{haAperMag1Err}^2} 
-  + 0.03$ &
-  158 million (72\%) &
-  Same as above for \ha. \\
-  \hline
-  All of the above (flagged as {\bf\emph{reliable}}) &
-  86 million (39\%) & \\
-  \hline
-\end{tabular}
-\caption{Recommended minimum quality criteria 
-for selecting objects with reliable colours 
-from the IPHAS DR2 source catalogue. 
-86~million entries in the catalogue (39\%)
-satisfy all the criteria listed in this table.
-For convenience, these have been flagged in the catalogue
-using the column named \emph{reliable}.}
-\label{tab:reliable}
-
-\vspace{2cm}
-
-\begin{tabular}{p{8cm}lp{6cm}}
-  \hline
-  Quality criterion & Rows passed & Description \\
-  \hline
-   reliable &
-   86 million (39\%) &
-   The object must satisfy the criteria listed in Table~\ref{tab:reliable}. \\
-
-   pStar $>$ 0.9 &
-   145 million (66\%) &
-   % pStar > 0.89 is identical to mergedClass == -1
-   The object must appears as a perfect point source,
-   as inferred from comparing its Point Spread Function (PSF)
-   with the average PSF measured in the same CCD. \\
-
-   {\sc not} \emph{deblend} &
-   177 million (81\%) &
-   The source must appear as a single, unconfused object. \\
-
-   {\sc not} \emph{brightNeighb} &
-   216 million (99\%) &
-   There is no star brighter than $V < 4$ within 10 arcmin, 
-   or brighter than $V < 7$ within 5 arcmin.
-   Such very bright stars cause scattered light and diffraction spikes,
-   which may add systematic errors to the photometry
-   or even trigger spurious detections. \\  
-  \hline
-
-  All of the above (flagged as {\bf\emph{veryReliable}}) &
-  59 million (27\%) & \\
-  \hline
-\end{tabular}
-\caption{Additional quality criteria which are recommended
-for applications which require very reliable colours
-at the expense of completeness. 
-For convenience, the sources which satisfy the criteria listed
-in this table have been flagged in the catalogue
-using the column named \emph{veryReliable}.}
-\label{tab:veryreliable}
-
-\vspace{2cm}
-\end{table*}
-
-For applications which require
-an even higher standard of reliability,
-a further set of additional quality criteria
-are listed in Table~\ref{tab:veryreliable}.
-These criteria are designed to ensure that
-(i) the object appeared as a perfect point source,
-(ii) the object was not blended with a nearby neighbour,
-and (ii) the object was not located near a very bright star.
-59 million sources (27\%) satisfy
-these additional criteria 
-and are hereafter referred to as ``very reliable''.
-Again, the catalogue contains a boolean column
-named \emph{veryReliable} which flags these objects.
-
-Figure~\ref{fig:magdist} compares the $r$-band magnitude
-distribution for objects with and without the 
-\emph{reliable} and \emph{veryReliable} criteria applied. 
-We find that 81\% of the sources 
-in the magnitude range $13 < r < 19$
-are considered \emph{reliable},
-which drops to 72\% in the range $19 < r < 20$
-and 9\% at $r>20$.
-In contrast, only 54\% of the sources 
-in the magnitude range $13 < r < 19$
-are considered \emph{veryReliable}.
-The stricter criteria filter out a lot
-of objects at early Galactic longitudes 
-where source blending is a common problem
-(we will demonstrate this in \S\ref{sec:densities}).
-The \emph{veryReliable} flag should hence
-only be used in applications which require very reliable photometry
-at the expense of completeness,
-which might be the case for e.g. spectroscopic target selection.
-
-\begin{figure}
-    \includegraphics[width=0.5\textwidth]{./plots/magdist/magdist-r.pdf} 
-    \caption{r-band magnitude distribution for all objects in the catalogue 
-    (light grey), for objects flagged as \emph{reliable} 
-    according to the criteria set out in Table~\ref{tab:reliable} (grey),
-    and for objects flagged as \emph{veryReliable} 
-    following Table~\ref{tab:veryreliable} (dark grey).
-    The magnitude distributions for $i$ and \ha\
-    look identical, apart from being shifted
-    by about 1 and 0.5 mag towards brighter magnitudes,
-    respectively.}
-    \label{fig:magdist}
-\end{figure}
-
-It is easy to see how the quality criteria
-may be adapted to be more tolerant.
-For example, by raising the allowed photometric uncertainties
-from 0.1 mag to 0.2 mag one can retrieve 42 million candidate sources.
-
-\subsection{Photometric uncertainties}
-
-Figure~\ref{fig:uncertainties} shows the mean photometric
-uncertainties as a function of magnitude for each band.
-We find that the uncertainties typically
-reach 0.1 mag near 20.5 in $r$ 
-and 19.5 in $i$/\ha when the default 2.3\arcsec\ aperture is used.
-At this point we note that the average colour
-of objects in the survey is
-$1.06\pm0.12$ for ($r$-$i$) and $0.44\pm0.03$ for ($r$-\ha).
-The better depth of $r$ is hence compensated
-by the fact that stars tend to have 
-brighter magnitudes in $i$ and \ha.
-
-\begin{figure}
-    \includegraphics[width=0.5\textwidth]{./plots/errors/uncertainties.pdf} 
-    \caption{Mean photometric uncertainties
-             for $r$ (top), $i$ (middle) and \ha\ (bottom).
-             Data points shown are the average values of
-             columns \emph{rErr}, \emph{iErr} and \emph{haErr}
-             in the catalogue, 
-             and the errorbars show the standard deviations.
-             The dashed and solid lines indicate 
-             the 10$\sigma$ and 5$\sigma$ limits, respectively.
-             These uncertainties are based only on the (Poissonian)
-             photon noise and hence this figure does not show
-             systematic or calibration uncertainties.}
-    \label{fig:uncertainties}
-\end{figure}
-
-The uncertainties shown in Fig.~\ref{fig:uncertainties}
-are the random errors based on the expected Poissonian photon noise.
-Systematics, such as calibration and deblending errors,
-are not included.
-To appraise the extent to which our photometry is affected
-by such systematics, we can exploit the
-secondary measurements which were made as part of the
-field-pair observing strategy and are available for 51\%
-of the sources in the catalogue.
-
-Figure~\ref{fig:pairmag} shows the mean residuals between
-the primary and secondary magnitudes
--- i.e. the average difference between catalogue columns \emph{r}-\emph{r2},
-\emph{i}-\emph{i2}, \emph{ha}-\emph{ha2} -- as a function of magnitude.
-We find that sources across the magnitude range 
-$13 < r < 17$ are consistent at the level of 5\%
-(i.e. $\sigma_{r-r2} \le 0.05$ mag),
-with the best repeatability
-reached at $r=14$ ($\sigma_{r-r2}~=~0.041$~mag).
-We draw attention to the fact that brighter stars
-tend to show significantly larger residuals
--- e.g. $\sigma_{r-r2}~=~0.14$~mag at $r=12$ --
-which is due to saturation effects.
-At the faint end we find residuals which
-show significantly more scatter than would be expected
-from photon noise alone, that is,
-the effects of source blending and background subtraction
-appear to dominate from $\sim$20th magnitude onwards.
-
-\begin{figure*}
-    \vspace{1cm}
-    \includegraphics[width=0.5\textwidth]{./plots/pairmag/pairmag.pdf} 
-    \caption{Photometric repeatability illustrated by plotting
-             the mean residuals between all the primary and secondary detections
-             in the catalogue as a function of magnitude.
-             The best photometric repeatability is reached at $r=14$
-             with $\sigma_{r-r2}~=~0.041$~mag.
-             Note that bright stars at $r<13$ and $i<12$ 
-             show increasing uncertainties due to saturation effects.}
-    \label{fig:pairmag}
-    \vspace{1cm}
-    \includegraphics[width=0.5\textwidth]{./plots/pairmag/pairmag-reliable.pdf} 
-    \caption{Same as Figure~\ref{fig:pairmag},
-    except that only the subset of sources
-    flagged as \emph{veryReliable} are now included.
-    We find that applying the quality criteria
-    has improved the photometric repeatability significantly.
-    The best repeatability is again reached at $r=14$
-    but has reduced to $\sigma=0.028$ mag.
-    The quality criteria have also been successful
-    at removing objects with large systematics at the bright
-    and faint ends.}
-    \label{fig:pairmag_reliable}
-\end{figure*}
-
-In Figure~\ref{fig:pairmag_reliable} we show 
-a similar comparison of the primary and secondary detections,
-but this time we have only include sources which are flagged
-as \emph{veryReliable} in the catalogue
-(i.e. not saturated, not confused, not near bright stars, etc.)
-We find that the average residuals are significantly better
-for this subset of the catalogue. Sources across the magnitude range 
-$13 < r < 17$ are consistent at the level of 0.03 mag,
-and the best repeatability is again reached at $r=14$
-with $\sigma_{r-r2}~=~0.028$~mag.
-We conclude that the \emph{veryReliable} quality criteria are effective
-in reducing the systematic errors to
-the same level as the accuracy of the
-global photometric calibration (cf. \S\ref{sec:calibration}).
-Moreover, the large systematics at the bright and faint end
-have disappeared.
-
-\subsection{Source densities and blending problems}
-\label{sec:densities}
-
-The mean source density as a function of Galactic longitude
-is shown in Figure~\ref{fig:density} (thick blue line).
-The densities were computed by counting the 
-number of sources in $1^\circ$-wide longitude bins
-across which covered the entire latitude range $-5^\circ<b<+5^\circ$.
-Unsurprisingly, we find the average source densities to increase
-towards the Galactic centre.
-For example, the average source density near $l\simeq 30^\circ$
-is roughly 30\,000 objects per square degree,
-which is five times more than the density
-found near $l\simeq 180^\circ$.
-
-In addition to the global trend, 
-there are significant variations in the source density on smaller scales.
-For example,  we find a significant drop near the constellations 
-of Cygnus ($l\simeq 80$) and Aquila ($l\simeq 40$),
-which are regions known to be affected
-by high levels of foreground extinction.
-Dark clouds are visible towards these constellation by eye,
-and they are often referred to as ``the Great Rift''.
-
-We warn however that the densities reported here 
-have not been corrected for survey completeness
-or differences in the observing conditions across the survey.
-For example, the dip in the density near $l\simeq140^\circ$
-is an artificial feature caused by gaps
-in the footprint coverage (which are apparent in Fig.~\ref{fig:footprint}).
-In a forthcoming paper,
-we aim to calibrate the IPHAS-based source densities
-by injecting artificial stars into the IPHAS images
-and measuring their recovery rate (Farnhill et al., in preparation).
-Indeed IPHAS has the potential to offer
-calibrated, two-dimensional stellar density maps
-which can be used to constrain detailed models of our Galaxy,
-but it is beyond the scope of the present work.
-
-\begin{figure*}
-    \includegraphics[width=\textwidth]{./plots/density/density.pdf} 
-    \caption{Mean number density of sources in the catalogue 
-    as a function of Galactic longitude, 
-    with and without blended sources included. 
-    The densities shown were computed by counting the sources 
-    at each longitude between $-5^\circ<b<+5^\circ$ (upper blue line).
-    We also show the densities based on only counting those sources 
-    for which the \emph{deblend} flag is {\sc false}, 
-    i.e. unconfused sources for which the CASU pipeline did not have to apply 
-    a deblending procedure (lower red line). 
-    }
-    \label{fig:density}
-\end{figure*}
-
-In Figure~\ref{fig:density} we also shows the density
-of non-blended sources (thin red line).
-These are sources for which the \emph{deblend} flag is {\sc false},
-i.e. sources for which the CASU pipeline did not have to apply 
-a deblending procedure to separate the flux
-originating from two or more overlapping objects.
-We find a strong correlation between the source density
-and the fraction of sources affected by source blending.
-For example, only NN\% of the sources are blended
-at $l>90^\circ$, whereas NN\% are blended at $l<90^\circ$.
-
-As we explained earlier, blended sources must be used with caution.
-Firstly, the deblending-procedure crucially depends
-on the local PSF being measured accurately. 
-Secondly, blended sources may are likely candidates to have fallen victim
-to source confusion during the band-merging procedure.
-In future work we hope to investigate the use of
-more advanced PSF-fitting routines
-in which sources are measured simultaneously across all bands,
-perhaps guided by an external list of sources provided
-by near-infrared surveys or Gaia data.
-
-\section{Demonstration}
-\label{sec:demonstration}
-
-We conclude this paper by demonstrating how the unique
-$r$-$i$/$r$-\ha\ colour-colour diagram offered by this catalogue
-can readily be used to
-(i) characterise the extinction regime at different sightlines, and
-(ii) identify \ha\ emission-line objects.
-
-\subsection{Colour-colour and colour-magnitude diagrams}
-
-The survey's unique $r$-\ha\ colour,
-when combined with $r$-$i$,
-has been shown to provide simultaneous constraints 
-on intrinsic stellar colour and interstellar extinction \citep{Drew2008}. 
-That is, the main sequence in the $r$-$i$/$r$-\ha\ diagram
-runs in a direction that is different from the reddening vector,
-because the $r$-\ha\ colour tends to act as a coarse proxy for spectral type
-and is less sensitive to reddening than $r$-$i$.
-As a result, the distribution of a stellar population
-in the IPHAS colour-colour diagram
-can offer a handle on the properties and extinction regime
-along a line of sight.
-
-This is demonstrated in Figures \ref{fig:l180}, \ref{fig:l45} \& \ref{fig:l30},
-where we present three sets of IPHAS colour/magnitude diagrams
-towards three distinct sightlines
-located at Galactic longitudes $180^\circ$, $45^\circ$ and $30^\circ$,
-respectively.
-Each figure contains all the sources which are flagged as \emph{veryReliable}
-and are located in a region of one square degree 
-centred on the coordinates indicated in the diagram
-(i.e. within a radius of $0.564^\circ$ from the indicated sightline).
-For clarity, we have imposed the additional criterion
-for photometric uncertainties to be smaller than 0.05 mag in each band
-(corresponding to a magnitude limit near $\sim$19th magnitude, effectively).
-
-Each of the diagrams reveals a well-defined locus,
-which demonstrates the health of the catalogue and the global calibration
-for investigating stellar populations across wide areas.
-We have annotated the colour-colour diagrams
-by showing the position of the unreddened main sequence (thin solid line),
-the unreddened giant branch (thick solid line)
-and the reddening track for an A0V-type star (dashed line)
--- all three are based on the \cite{Pickles1998} library of empirical spectra
-tabulated for IPHAS by \cite{Drew2005}.
-In the colour-magnitude diagrams we only show the reddening vector
-together with the unreddened 1~Gyr isochrone due to \cite{Bressan2012},
-which are made available for the IPHAS filter system through a
-popular on line tool hosted by the Observatory of Padova (http://stev.oapd.inaf.it/cmd).
-The isochrone and reddening vector has been placed
-at an arbitrary distance of 2~kpc.
-
-Each of the sightlines reveals a stellar population
-with distinct characteristics.
-Firstly, towards the Galactic anti-centre at $l=180^\circ$ (Fig.~\ref{fig:l180}) 
-we find a population dominated by lowly-reddened 
-main sequence stars and an apparent absence of reddened giants.
-%consistent with the sightline extinction of XX from Schlegel (cite). 
-In contrast, closer towards the galactic
-centre at $l=45^\circ$ (Fig.~\ref{fig:l45})
-we find a wealth of reddened objects
-which appear to be separated into a lowly
-and a highly reddened component,
-perhaps revealing two distinct parts of the Galaxy.
-Finally, in one of our earliest sightlines at $l=30^\circ$
-we find a very high number of extremely reddened giants
-in addition to an unreddened population
-of foreground dwarfs.
-
-The number density of stars in the colour-colour
-and colour-magnitude space
-can be compared against population synthesis models
-to create three-dimensional maps of the extinction
-across several kpc \citep{Sale2010,Sale2012}.
-Such an extinction map based on our catalogue
-is to be released in a separate paper
-that accompanies this data release (Sale et al., in preparation).
-
-\begin{figure*}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=0.5\textwidth]{./plots/ccd-180-3.pdf} 
-        \includegraphics[width=0.5\textwidth]{./plots/cmd-180-3.pdf}
-    \end{minipage}
-    \caption{Colour-colour and colour-magnitude diagram (left and right panel)
-    showing sources flagged as \emph{veryReliable}
-    located in an area of one square degree
-    centred near the Galactic anti-centre 
-    at $(l,b)=(180^\circ,+3^\circ)$.
-    The colour-colour diagram shows the
-    position of the main sequence (thin solid line),
-    giant stars (thick solid line)
-    and the reddening track for an A0V-type star (dashed line)
-    based on the \citet{Pickles1998} library of empirical spectra.
-    The colour-magnitude only show the reddening vector
-    along with the unreddened 1~Gyr isochrone due to \citet{Bressan2012},
-    which has been placed at an arbitrary distance of 2~kpc for reference.
-    This is one of the least reddened sightlines
-    in the survey %(Av XX, Schlegel et al. XXXX)
-    and hence the observed stellar population appears to be dominated 
-    by lowly reddened main sequence stars.}
-    \label{fig:l180}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=0.5\textwidth]{./plots/ccd-45-2.pdf}
-        \includegraphics[width=0.5\textwidth]{./plots/cmd-45-2.pdf}
-    \end{minipage}
-    \caption{Same as above for $(l,b)=(45^\circ,+2^\circ)$,
-    which is one of the highest-density sightlines in the survey,
-    revealing a population of stars with a reddening distribution
-    that appears to be bi-model.}
-    \label{fig:l45}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=0.5\textwidth]{./plots/ccd-30-0.pdf}
-        \includegraphics[width=0.5\textwidth]{./plots/cmd-30-0.pdf} 
-    \end{minipage}
-    \caption{Same as above for $(l,b)=(30^\circ,0^\circ)$.
-    This is one of the most reddened sightlines in the survey.
-    %(XXX, Schlegel et al XXXX).
-    }
-    \label{fig:l30}
-\end{figure*}
-
-\subsection{Identifying \ha\ emission-line objects}
-
-A primary motivation for carrying
-out the survey 
-was to enable the discovery of 
-new emission-line objects across the Galactic Plane.
-\ha\ in emission is a well-known tracer
-for stars in the short-lived pre- or
-post-main sequence stages of their evolution,
-and hence IPHAS aims to allow larger, deeper
-and more statistically robust samples of such rare objects
-to be established.
-The survey-wide identification and analysis 
-of such stars is beyond the scope of the present work,
-but in this section we demonstrate how the
-catalogue may be used for this purpose.
-
-An initial list of candidate \ha-emitters
-based on the first IPHAS data release has previously
-been presented by \cite{Witham2008}. 
-Because no global uniform calibration was available
-at the time, \citeauthor{Witham2008} employed 
-a sigma-clipping technique to select objects with
-large, outlying $r$-\ha\ colours.
-In contrast, the new catalogue
-allows objects to be picked out
-from the $r$-$i$/$r$-\ha\ colour-colour diagram
-using model-based colour criteria
-rather than a statistical procedure.
-In what follows we demonstrate this ability 
-by selecting candidate emission-line objects
-towards a small region in the sky.
-
-The target of our demonstration is Sh 2-82:
-a 5~arcmin-wide H{\sc ii} region located near $(l,b)=(53.55^\circ, 0.00^\circ)$
-in the constellation of Sagitta.
-Nicknamed by amateur astronomers as the ``Little Cocoon Nebula'',
-Sh 2-82 is ionised by 
-the $\sim$10th magnitude star HD\,231616
-% Georgelin1973 claims B0III, others claim B0V
-with spectral type B0V/III
-\citep{Georgelin1973,Mayer1973,Hunter1990}.
-The ionising star has been placed at a likely distance of 1.5-1.7 kpc
-based on the photometric parallax
-\citep{Mayer1973,Lahulla1985,Hunter1990}.
-
-Figure~\ref{fig:mosaic_iphas} shows a 20-by-15 arcmin
-colour mosaic centred on Sh 2-82,
-composed of our \ha\ (red channel),
-$r$ (green channel),
-and $i$ (blue channel) images.
-The ionising star can be seen as the bright object
-near the centre of the H{\sc ii} region,
-which is surrounded by a faint reflection nebula
-and several dark cloud filaments.
-For comparison, Figure~\ref{fig:mosaic_spitzer} shows
-a mosaic of the identical region 
-as seen by the Spitzer Space Telescope
-in the mid-infrared. The Spitzer image
-reveals a bubble-shaped structure of warm dust
-which surrounds the entire H{\sc ii} region.
-
-\begin{figure*}
-    \begin{minipage}[b]{0.8\linewidth}
-        \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-iphas.pdf} 
-    \end{minipage}
-\caption{IPHAS-based mosaic of H{\sc ii} region Sh 2-82,
-composed of \ha\ (red channel), $r$ (green channel) and $i$ (blue channel). Yellow triangles show the position of candidate \ha-emitters
-which have been selected from the colour-colour diagram
-in Figure~\ref{fig:emitters}. Note that the H{\sc ii} region is surrounded by a faint blue/green reflection nebula
-and dark cloud filaments.}
-\label{fig:mosaic_iphas}
-    \begin{minipage}[b]{0.8\linewidth}
-        \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-spitzer.pdf} 
-    \end{minipage}
-    \caption{Star-forming region Sh 2-82 as seen in the mid-infrared
-    by the Spitzer Space Telescope. The mosaic is composed of the 24\,\micron\ (red), 8.0\,\micron\ (green) and 4.5\,\micron\ (blue) bands.
-    The image reveals a bubble-shaped structure which surrounds the {\sc Hii} region that is seen in the IPHAS mosaic of the same region (Figure~\ref{fig:mosaic_iphas}).}
-    \label{fig:mosaic_spitzer}
-\end{figure*}
-
-Figure~\ref{fig:emitters} presents
-the IPHAS colour-colour diagram for the region covered by the
-mosaic images.
-Gray circles show all objects in the region
-which are brighter than $r<20$
-and have been flagged as \emph{reliable}
-in our catalogue.
-The diagram also shows the unreddened main sequence (solid line)
-and the expected position of unredded main-sequence stars
-with \ha\ in emission at a strength of EW=$-10{\rm \AA}$ (dashed line),
-taken from the colour simulations due to \citet{Barentsen2011a}.
-Six stars are found to lie confidently above the 
-dashed line at the level of $3\sigma$ 
-(i.e. the distance between the dashed line is larger than
-three times the uncertainty on the $r$-\ha\ colour).
-These reliable candidate \ha-emitters
-are marked by red triangles in the colour-colour diagram
-and their photometry is detailed in Appendix~\ref{app:emitters}.
-
-\begin{figure}
-  \includegraphics[width=0.45\textwidth]{./plots/sh2-82-ccd.pdf}
-    \caption{$r$-$i$/$r$-\ha\ diagram for the rectangular region of 
-    20-by-15 arcmin centred on the H{\sc ii} region Sh 2-82,
-    which is the area shown in Figure~\ref{fig:mosaic_iphas}.    
-    The diagram shows all objects in the catalogue
-    which have been flagged as \emph{reliable} and are brighter
-    than $r<20$ (grey circles).
-    The unreddened main sequence is indicated by a solid line,
-    while the main sequence for stars with an \ha\ emission line
-    strength of $-10\,\rm{\AA}$ EW is indicated by a dashed line,
-    following the colour simulations due to \citet{Barentsen2011a}.
-    Red triangles indicate objects which have been identified as
-    as highly likely \ha-emitters (see text).}
-    \label{fig:emitters}
-\end{figure}
-
-The spatial distribution of our six candidate emission-line objects
-is marked by yellow triangles in the colour mosaic (Fig.~\ref{fig:mosaic_iphas}).
-They are likely to be genuine young stars for two reasons.
-Firstly, two of our candidates have recently been identified as likely
-candidate Young Stellar Objects (YSO)
-in an investigation of the region by \cite{Yu2012}.
-In their study, the authors used 2MASS and Spitzer data
-to select likely young stars by selecting objects with
-circumstellar disks based on the infrared colour excess.
-Secondly, we find that the four remaining objects are 
-also detected in the Spitzer-based image,
-although their colours are less extreme
-than those identified as likely YSOs by \citeauthor{Yu2012}.
-
-Prior to IPHAS, this region was essentially unstudied
-at faint magnitudes in visible light.
-\cite{Lahulla1985} reported magnitudes for 8 stars in the optical at 
-$V < 15$. In contrast, the IPHAS catalogue offers photometry
-for NN stars in the region down to $r<20$.
-This demonstrates the ability of IPHAS to providing complimentary
-for the wealth of poorly-studied star-forming regions
-at low Galactic latitudes,
-which have been unveiled in recent years
-by the observed wealth of star-forming ``bubbles''
-at mid-infrared wave%\cite{Bica2003} was the first to identify a
-%near-infrared cluster
-%associated with this region using near-infrared 2MASS data,
-%and only recently has this cluster been confirmed
-%in a short investigation which combined 2MASS
-%and Spitzer data \cite{Yu2012}.
-%Many of the young stars identified by \citeauthor{Yu2012}
-%can be seen as pink and red stars in Fig.~\ref{fig:mosaic_spitzer}.lengths \cite[e.g.][]{Churchwell2006,Simpson2012}.
-%In fact the cluster of young stars associated with Sh 2-82
-%was only first reported by \cite{Bica2003},
-%remained unstudied until the study by \cite{Yu2012}.
-
-%Amongst the six emitters, N are detected at SNR X
-%in the Spitzer X micron band.
-
-%This demonstrates how IPHAS is providing complimentary
-%data for some of the thousands of star-forming ``bubbles'' have been unveiled at mid-infrared wavelengths \cite[e.g.][]{Churchwell2006,Simpson2012}.
-%Optically-unveiled members of such star-forming regions
-%may help to constrain extinctions and distances towards such regions \citep{Barentsen2013}.
-%Moreover, in \cite{Barentsen2011a} we showed
-%how the history of star formation in a region,
-%and the possible triggered nature, may be discussed.
-%This is particularly true
-
-%TODO: annotate -10 EW text on plot?
-%TODO: mention Gaia, e.g. IPHAS for transient identification?
-
-%\subsection{Caveats and lessons learnt}
-%\cite{Bica2003} was the first to identify a
-%near-infrared cluster
-%associated with this region using near-infrared 2MASS data,
-%and only recently has this cluster been confirmed
-%in a short investigation which combined 2MASS
-%and Spitzer data \cite{Yu2012}.
-%Many of the young stars identified by \citeauthor{Yu2012}
-%can be seen as pink and red stars in Fig.~\ref{fig:mosaic_spitzer}.
-
-\section{Conclusions and future work}
-\label{sec:conclusions}
-
-A new data release for the IPHAS survey was presented,
-taking the coverage up to over 90\% of the Northern Galactic Plane 
-at $|b|<5^\circ$
-and providing a uniform photometric calibration
-for the first time.
-We explained the data reduction and quality control procedures that
-were applied, described and tested the new global photometric calibration,
-and detailed the construction of the user-friendly source catalogue.
-
-The observations included in this release
-were found to achieve a median seeing of 1.1 arcsec
-and a $5\sigma-$depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$, \ha$=20.3\pm 0.3$.
-The global calibration and photometric repeatability
-is accurate at the level of $\sigma=0.03$ mag,
-providing a significant improvement over the 
-previous data release.
-The source catalogue provides the best-available
-single-epoch astrometry and photometry
-for 219~million unique sources.
-
-The data-taking strategy developed for IPHAS
-have since been reapplied to carry out a companion INT/WFC survey called UVEX
-in U/$g$/$r$ \citep{Groot2009},
-and also southern counterpart to IPHAS and UVEX 
-is being carried out in $u$/$g$/$r$/$i$/\ha\ 
-called VPHAS+ (Drew et al, in press).
-We hope to re-use the experience gained by this data release
-to create similar releases for these companion surveys.
-
-In future work, we aim to draw upon the PanSTARRS photometric
-survey to further improve the accuracy of our global calibration.
-We will also aim to correct the photometry for the
-radial field distortions.
-%We may also produce a catalogue which details all detections.
-
-%TODO: mention Gaia?
-
-\section*{Data access and source code}
-\label{sec:dataaccess}
-
-The catalogue is made available through the Vizier
-catalogue tool (http://vizier.u-strasbg.fr),
-where it is known as the ``IPHAS DR2 Source Catalogue''
-(catalogue ID ???).
-The full catalogue can also be downloaded in its entirety
-from the IPHAS website (www.iphas.org) as a collection 
-of binary FITS tables which comprise 50\,GB,
-which is accompanied by a script
-to ingest the data into a PostgreSQL database.
-Our website also provides access to the pipeline-processed
-imaging data, which we have updated to include
-the re-calibrated DR2 zeropoint in the image headers.
-
-The source code that was used to generate
-the catalogue is available at
-https://github.com/barentsen/iphas-dr2
-
-\section*{Acknowledgments}
-
-The IPHAS survey was carried out 
-at the Isaac Newton Telescope (INT).
-The INT is operated on the island of La Palma
-by the Isaac Newton Group
-in the Spanish Observatorio del Roque de los Muchachos
-of the Instituto de Astrofisica de Canarias.
-All data were processed 
-by the Cambridge Astronomical Survey Unit,
-at the Institute of Astronomy in Cambridge.
-
-Preparation of the catalogue was eased greatly
-by a number of software packages,
-including the Python modules
-AstroPy \citep{Astropy},
-APLpy, NumPy and SciPy,
-the PostgreSQL database software,
-the TOPCAT and STILTS packages \citep{Taylor2005,Taylor2006},
-and the Montage software maintained by NASA/IPAC.
-We also made use of the SIMBAD, Vizier and Aladin \citep{Aladin} tools operated at CDS, Strasbourg, France.
-
-This research made extensive use of
-several complementary photometric surveys.
-Our global calibration was aided
-by the AAVSO Photometric All-Sky Survey (APASS),
-funded by the Robert Martin Ayers Sciences Fund.
-The calibration was tested against the
-Sloan Digitized Sky Survey (SDSS),
-funded by the Alfred P.
-Sloan Foundation, the Participating Institutions, the National Science Foundation, the U.S.
-Department of Energy, the National Aeronautics and Space Administration, the Japanese Monbukagakusho, the Max Planck Society, and the Higher Education Funding Council for England.
-The astrometric pipeline reduction made
-significant use of the Two Micron All Sky Survey (2MASS),
-which is a joint project 
-of the University of Massachusetts
-and the Infrared Processing and Analysis Center/
-California Institute of Technology,
-funded by NASA and the NSF.
-
-GB and JED acknowledge the support of a grant
-from the Science \& Technology Facilities Council
-of the UK (STFC, ref ST/J001335/1).
-HJF is in receipt of an STFC postgraduate studentship.
-
-\bibliographystyle{mn2e}
-\bibliography{dr2}
-
-\appendix
-\input{app_columns.tex}
-
-\section{Candidate emission-line objects towards Sh 2-82}
-\label{app:emitters}
-\begin{table*}
-    \begin{tabular}{lccc}
-    \toprule
-    Name & $r$ & $i$ & \ha  \\
-    \midrule
-IPHAS2 J192954.40+181026.1& $17.69\pm0.01$ & $16.12\pm0.01$ & $16.19\pm0.01$ \\
-IPHAS2 J193011.01+182051.2& $18.55\pm0.02$ & $16.95\pm0.02$ & $17.31\pm0.02$ \\
-IPHAS2 J193021.52+181954.5& $19.72\pm0.05$ & $17.94\pm0.03$ & $18.47\pm0.04$ \\
-IPHAS2 J193024.45+181938.3& $19.31\pm0.04$ & $17.55\pm0.02$ & $17.99\pm0.03$ \\
-IPHAS2 J193033.00+181609.3& $18.25\pm0.01$ & $16.91\pm0.01$ & $16.92\pm0.01$ \\
-IPHAS2 J193042.48+182317.4& $19.96\pm0.03$ & $18.11\pm0.03$ & $18.48\pm0.03$ \\
-    \bottomrule
-    \end{tabular}
-    \caption{Candidate \ha-emitters towards Sh 2-82.}
-    \label{tbl:emitters}
-\end{table*}
-
-\label{lastpage}
-
-\end{document}

### 2014-03-12 paper/iphas-dr2.tex
More tidying
-

--- 2014-03-12 paper/iphas-dr2.tex
@@
 \usepackage{booktabs}
-\usepackage{aas_macros}
+%\usepackage{aas_macros}
+
+\graphicspath{ {/home/gb/msc/Dropbox/papers/dr2/} }
 
@@
 \author[G. Barentsen
-et. al]{G. Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
-H. J. Farnhill$^{1}$,
-J. E. Drew$^{1}$,
-M. J. Irwin$^{4}$, \newauthor
-E. A. Gonz$\acute{\rm{a}}$lez-Solares$^{4}$,
-R. Greimel$^{2}$,
-B. Mizalski$^{3}$,
-C. Ruhland$^{1}$, \newauthor plus many friends.
+et. al]{Geert Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
+H. J. Farnhill$^1$,
+J. E. Drew$^1$,
+E. A. Gonz$\acute{\rm{a}}$lez-Solares$^2$, \newauthor
+M. J. Irwin$^2$,
+S. E. Sale$^3$,
+R. Greimel$^4$,
+P. Groot$^5$,
+J. Lewis$^2$,
+A. Mampaso$^6$, \newauthor
+B. Mizalski$^{99}$,
+C. Ruhland$^1$,
+plus many contributors.
 \newauthor\\
-$^{1}$Centre for Astrophysics Research, STRI, University of Hertfordshire, College Lane Campus, Hatfield AL10 9AB\\
-$^{2}$Institute for Geophysics, Astrophysics and Meteorology, Institute of Physics, Karl-Franzens-Universit$\ddot{a}$t Graz, Universit$\ddot{a}$tsplatz 5, 8010 Graz, Austria\\
-$^{3}$Building, Institute, Street Address, City, Code, Country\\
-$^{4}$Institute of Astronomy, Madingley Road, Cambridge CB3 0HA}
+$^{1}$School of Physics, Astronomy \& Mathematics, University of Hertfordshire, College Lane, Hatfield, Hertfordshire, AL10 9AB, U.K.\\
+$^{2}$Institute of Astronomy, Cambridge University, Madingley Road, Cambridge, CB3 OHA, U.K.\\
+$^{3}$Rudolf Peierls Centre for Theoretical Physics, Keble Road, Oxford, OX1 3NP, U.K.\\
+$^{4}$IGAM, Institute of Physics, University of Graz, Universit\"atsplatz 5, Graz, Austria.\\
+$^{5}$Afdeling Sterrenkunde, Radboud Universiteit Nijmegen, Faculteit NWI, Postbus 9010, 6500 GL Nijmegen, The Netherlands.\\
+$^{6}$Instituto de Astrof\'isica de Canarias, 38200 La Laguna, Tenerife, Spain.\\
+$^{99}$Please add your current affiliation.\\
+}
 
@@
 of the Northern Galactic Plane (IPHAS)
-is a 1800 deg$^2$ imaging survey
+is a 1800 square degrees imaging survey
 covering the entire northern Milky Way at $|b| < 5^\circ$
@@
 uniformly-calibrated source catalogue
-to have been extracted from the survey,
+to have been extracted from the images,
 providing single-epoch photometry
@@
 and its quality warning flags.
-We find that the new calibration is accurate to
-$\sigma=0.03$ mag.
+We find that the new calibration
+and the photometric repeatability are accurate to
+$\sigma=0.03$~mag.
 Finally, we demonstrate the ability of the 
@@
 towards different Galactic sightlines
-and (ii) select reliable candidate \ha-emission line objects.
+and (ii) select candidate \ha-emission line objects.
 The catalogue which accompanies this paper
@@
 to several infrared surveys of the Galactic Plane,
-including many poorly-studied star-forming regions and ``bubbles'',
-and provides images for some of the most
+including many poorly-studied star-forming regions and ``bubbles''.
+It also provides images for some of the most
 crowded regions to be faced by Gaia.
@@
 The densities were computed by counting the 
-number of sources in $1^\circ$-wide longitude bins
-across which covered the entire latitude range $-5^\circ<b<+5^\circ$.
+number of sources in $1^\circ$-wide longitude bins,
+each covering the entire latitude range $-5^\circ<b<+5^\circ$.
 Unsurprisingly, we find the average source densities to increase
@@
 In addition to the global trend, 
-there are significant variations in the source density on smaller scales.
+there are significant variations
+in the source density on smaller scales.
 For example,  we find a significant drop near the constellations 
@@
 by high levels of foreground extinction.
-Dark clouds are visible towards these constellation by eye,
-and they are often referred to as ``the Great Rift''.
+In fact dark clouds are visible towards these constellation by eye,
+and they have often been referred to as ``the Great Rift''.
 
@@
 and measuring their recovery rate (Farnhill et al., in preparation).
-Indeed IPHAS has the potential to offer
+IPHAS has the potential to offer
 calibrated, two-dimensional stellar density maps
-which can be used to constrain detailed models of our Galaxy,
-but it is beyond the scope of the present work.
+which can be used to constrain detailed models of our Galaxy.
 
@@
 originating from two or more overlapping objects.
-We find a strong correlation between the source density
-and the fraction of sources affected by source blending.
-For example, only NN\% of the sources are blended
-at $l>90^\circ$, whereas NN\% are blended at $l<90^\circ$.
+We find that the fraction of sources affected by blending problems
+is strongly correlated with the local source density.
+For example, only 11\% of the sources are blended
+at $l>90^\circ$, whereas 24\% are blended at $l<90^\circ$.
 
 As we explained earlier, blended sources must be used with caution.
-Firstly, the deblending-procedure crucially depends
-on the local PSF being measured accurately. 
-Secondly, blended sources may are likely candidates to have fallen victim
+In particular, we warn that blended sources
+are likely candidates to have fallen victim
 to source confusion during the band-merging procedure.
@@
 in which sources are measured simultaneously across all bands,
-perhaps guided by an external list of sources provided
-by near-infrared surveys or Gaia data.
-
-\section{Demonstration}
-\label{sec:demonstration}
-
-We conclude this paper by demonstrating how the unique
-$r$-$i$/$r$-\ha\ colour-colour diagram offered by this catalogue
-can readily be used to
-(i) characterise the extinction regime at different sightlines, and
-(ii) identify \ha\ emission-line objects.
-
-\subsection{Colour-colour and colour-magnitude diagrams}
-
-The survey's unique $r$-\ha\ colour,
-when combined with $r$-$i$,
-has been shown to provide simultaneous constraints 
-on intrinsic stellar colour and interstellar extinction \citep{Drew2008}. 
-That is, the main sequence in the $r$-$i$/$r$-\ha\ diagram
-runs in a direction that is different from the reddening vector,
-because the $r$-\ha\ colour tends to act as a coarse proxy for spectral type
-and is less sensitive to reddening than $r$-$i$.
-As a result, the distribution of a stellar population
-in the IPHAS colour-colour diagram
-can offer a handle on the properties and extinction regime
-along a line of sight.
-
-This is demonstrated in Figures \ref{fig:l180}, \ref{fig:l45} \& \ref{fig:l30},
-where we present three sets of IPHAS colour/magnitude diagrams
-towards three distinct sightlines
-located at Galactic longitudes $180^\circ$, $45^\circ$ and $30^\circ$,
-respectively.
-Each figure contains all the sources which are flagged as \emph{veryReliable}
-and are located in a region of one square degree 
-centred on the coordinates indicated in the diagram
-(i.e. within a radius of $0.564^\circ$ from the indicated sightline).
-For clarity, we have imposed the additional criterion
-for photometric uncertainties to be smaller than 0.05 mag in each band
-(corresponding to a magnitude limit near $\sim$19th magnitude, effectively).
-
-Each of the diagrams reveals a well-defined locus,
-which demonstrates the health of the catalogue and the global calibration
-for investigating stellar populations across wide areas.
-We have annotated the colour-colour diagrams
-by showing the position of the unreddened main sequence (thin solid line),
-the unreddened giant branch (thick solid line)
-and the reddening track for an A0V-type star (dashed line)
--- all three are based on the \cite{Pickles1998} library of empirical spectra
-tabulated for IPHAS by \cite{Drew2005}.
-In the colour-magnitude diagrams we only show the reddening vector
-together with the unreddened 1~Gyr isochrone due to \cite{Bressan2012},
-which are made available for the IPHAS filter system through a
-popular on line tool hosted by the Observatory of Padova (http://stev.oapd.inaf.it/cmd).
-The isochrone and reddening vector has been placed
-at an arbitrary distance of 2~kpc.
-
-Each of the sightlines reveals a stellar population
-with distinct characteristics.
-Firstly, towards the Galactic anti-centre at $l=180^\circ$ (Fig.~\ref{fig:l180}) 
-we find a population dominated by lowly-reddened main sequence stars.
-This is consistent with the estimated total sightline extinction of $E(B-V)=0.49$
-given by \cite{Schlegel1998}, following the correction of \cite{Schlafly2011}.
-Looking in more detail we can see that the stellar locus is narrower for M-type dwarfs than for earlier types;
-we do not observe M dwarfs experiencing the strongest reddening possible for this sightline.
-This implies that extinction is still increasing at distances beyond $\sim 2$~kpc,
-where M dwarfs are too faint to be contained in the IPHAS catalogue.
-It is also clear that there are no unreddened stars earlier than $\sim$K0 visible,
-such stars would be saturated if within a few hundred parsecs.
-This therefore suggests that there is a measurable increase in extinction locally.
-We also note an relative absence of late type giants which,
-due to the relative brevity of the corresponding phase of stellar evolution,
-would only account for a small proportion of a volume limited sample.
-
-In contrast, 
-closer towards the galactic centre at $l=45^{\circ}$ (Fig.~\ref{fig:l45}) we find a wealth of reddened objects.
-In the colour-magnitude diagram it is clear that the stars are split into two distinct groups,
-with one significantly redder than the other.
-The bluer group is composed of main sequence stars,
-with the slope of this group in the colour-magnitude diagram attributable to the significantly increasing extinction.
-Meanwhile the redder group is principally composed of red giant stars.
-As these stars are intrinsically brighter,
-at some given apparent magnitude they will be substantially further away than their main sequence counterparts.
-Given that, along this sight-line,
-extinction continues to increase with distance,
-the red giants we observe will be subject to more reddening than a main sequence star exhibiting the same apparent magnitude.
-Therefore, the clear split between the two groups arises due to the combination of increasing extinction and the distinct absolute magnitude ranges of main sequence and giant branch stars.
-
-Finally, in one of our earliest sightlines at $l=30^{\circ}$
-we find a very high number of extremely reddened giants
-in addition to an unreddened population
-of foreground dwarfs.
-In contrast to the sightline at $l=45^{\circ}$,
-there is no clear group of giant stars visible in the colour-magnitude diagram of Fig.~\ref{fig:l30},
-though the red clump stars are manifested as a slight over density which sits roughly 0.4~mags redder than the A0V reddening track.
-At $(l,b)=(45^{\circ}, 2^{\circ})$ the giant stars observed exhibit a relatively narrow range of reddenings as the lie beyond most or essentially all of the Galactic dust.
-At $(l,b)=(30^{\circ}, 0^{\circ})$ this is not the case:
-even at the substantial distances at which we can observe reddened giant stars extinction is continuing to rise and we still find ourselves well within the Galactic dust disc.
-It is also apparent that the $(r-i)$ width of both the M dwarfs and early A dwarfs is greater than that in Fig.~\ref{fig:l45}.
-This is indicative of a steeper rise in reddening,
-both within several hundred parsecs (M dwarfs)
-and within a few kpc (early A dwarfs).
-
-Here we have presented only a short qualitative description of the information obtainable from IPHAS colour-colour and colour-magnitude plots.
-A more rigorous quantitative analysis of the IPHAS catalogue
-can be undertaken to estimate both the stellar density distribution in the Milky Way \citep{Sale2010}
-and to create detailed three-dimensional maps of the extinction across several kpc \citep{Sale2009,Sale2012}.
-An extinction map based on our catalogue is to be released in a separate paper
-that accompanies this data release (Sale et al., in preparation).
-
-\begin{figure*}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=0.5\textwidth]{./plots/ccd-180-3.pdf} 
-        \includegraphics[width=0.5\textwidth]{./plots/cmd-180-3.pdf}
-    \end{minipage}
-    \caption{Colour-colour and colour-magnitude diagram (left and right panel)
-    showing sources flagged as \emph{veryReliable}
-    located in an area of one square degree
-    centred near the Galactic anti-centre 
-    at $(l,b)=(180^\circ,+3^\circ)$.
-    The colour-colour diagram shows the
-    position of the main sequence (thin solid line),
-    giant stars (thick solid line)
-    and the reddening track for an A0V-type star (dashed line)
-    based on the \citet{Pickles1998} library of empirical spectra.
-    The colour-magnitude only show the reddening vector
-    along with the unreddened 1~Gyr isochrone due to \citet{Bressan2012},
-    which has been placed at an arbitrary distance of 2~kpc for reference.
-    This is one of the least reddened sightlines
-    in the survey %(Av XX, Schlegel et al. XXXX)
-    and hence the observed stellar population appears to be dominated 
-    by lowly reddened main sequence stars.}
-    \label{fig:l180}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=0.5\textwidth]{./plots/ccd-45-2.pdf}
-        \includegraphics[width=0.5\textwidth]{./plots/cmd-45-2.pdf}
-    \end{minipage}
-    \caption{Same as above for $(l,b)=(45^\circ,+2^\circ)$,
-    which is one of the highest-density sightlines in the survey,
-    revealing two groups of stars in colour-magnitude space.}
-    \label{fig:l45}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=0.5\textwidth]{./plots/ccd-30-0.pdf}
-        \includegraphics[width=0.5\textwidth]{./plots/cmd-30-0.pdf} 
-    \end{minipage}
-    \caption{Same as above for $(l,b)=(30^\circ,0^\circ)$.
-    This is one of the most reddened sightlines in the survey.
-    %(XXX, Schlegel et al XXXX).
-    }
-    \label{fig:l30}
-\end{figure*}
-
-\subsection{Identifying \ha\ emission-line objects}
-
-A primary motivation for carrying
-out the survey 
-was to enable the discovery of 
-new emission-line objects across the Galactic Plane.
-\ha\ in emission is a well-known tracer
-for stars in the short-lived pre- or
-post-main sequence stages of their evolution,
-and hence IPHAS aims to allow larger, deeper
-and more statistically robust samples of such rare objects
-to be established.
-The survey-wide identification and analysis 
-of such stars is beyond the scope of the present work,
-but in this section we demonstrate how the
-catalogue may be used for this purpose.
-
-An initial list of candidate \ha-emitters
-based on the first IPHAS data release has previously
-been presented by \cite{Witham2008}. 
-Because no global uniform calibration was available
-at the time, \citeauthor{Witham2008} employed 
-a sigma-clipping technique to select objects with
-large, outlying $r$-\ha\ colours.
-In contrast, the new catalogue
-allows objects to be picked out
-from the $r$-$i$/$r$-\ha\ colour-colour diagram
-using model-based colour criteria
-rather than a statistical procedure.
-In what follows we demonstrate this ability 
-by selecting candidate emission-line objects
-towards a small region in the sky.
-
-The target of our demonstration is Sh 2-82:
-a 5~arcmin-wide H{\sc ii} region located near $(l,b)=(53.55^\circ, 0.00^\circ)$
-in the constellation of Sagitta.
-Nicknamed by amateur astronomers as the ``Little Cocoon Nebula'',
-Sh 2-82 is ionised by 
-the $\sim$10th magnitude star HD\,231616
-% Georgelin1973 claims B0III, others claim B0V
-with spectral type B0V/III
-\citep{Georgelin1973,Mayer1973,Hunter1990}.
-The ionising star has been placed at a likely distance of 1.5-1.7 kpc
-based on the photometric parallax
-\citep{Mayer1973,Lahulla1985,Hunter1990}.
-
-Figure~\ref{fig:mosaic_iphas} shows a 20-by-15 arcmin
-colour mosaic centred on Sh 2-82,
-composed of our \ha\ (red channel),
-$r$ (green channel),
-and $i$ (blue channel) images.
-The ionising star can be seen as the bright object
-near the centre of the H{\sc ii} region,
-which is surrounded by a faint reflection nebula
-and several dark cloud filaments.
-For comparison, Figure~\ref{fig:mosaic_spitzer} shows
-a mosaic of the identical region 
-as seen by the Spitzer Space Telescope
-in the mid-infrared. The Spitzer image
-reveals a bubble-shaped structure of warm dust
-which surrounds the entire H{\sc ii} region.
-
-\begin{figure*}
-    \begin{minipage}[b]{0.8\linewidth}
-        \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-iphas.pdf} 
-    \end{minipage}
-\caption{IPHAS-based mosaic of H{\sc ii} region Sh 2-82,
-composed of \ha\ (red channel), $r$ (green channel) and $i$ (blue channel). Yellow triangles show the position of candidate \ha-emitters
-which have been selected from the colour-colour diagram
-in Figure~\ref{fig:emitters}. Note that the H{\sc ii} region is surrounded by a faint blue/green reflection nebula
-and dark cloud filaments.}
-\label{fig:mosaic_iphas}
-    \begin{minipage}[b]{0.8\linewidth}
-        \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-spitzer.pdf} 
-    \end{minipage}
-    \caption{Star-forming region Sh 2-82 as seen in the mid-infrared
-    by the Spitzer Space Telescope. The mosaic is composed of the 24\,\micron\ (red), 8.0\,\micron\ (green) and 4.5\,\micron\ (blue) bands.
-    The image reveals a bubble-shaped structure which surrounds the {\sc Hii} region that is seen in the IPHAS mosaic of the same region (Figure~\ref{fig:mosaic_iphas}).}
-    \label{fig:mosaic_spitzer}
-\end{figure*}
-
-Figure~\ref{fig:emitters} presents
-the IPHAS colour-colour diagram for the region covered by the
-mosaic images.
-Gray circles show all objects in the region
-which are brighter than $r<20$
-and have been flagged as \emph{reliable}
-in our catalogue.
-The diagram also shows the unreddened main sequence (solid line)
-and the expected position of unredded main-sequence stars
-with \ha\ in emission at a strength of EW=$-10{\rm \AA}$ (dashed line),
-taken from the colour simulations due to \citet{Barentsen2011a}.
-Six stars are found to lie confidently above the 
-dashed line at the level of $3\sigma$ 
-(i.e. the distance between the dashed line is larger than
-three times the uncertainty on the $r$-\ha\ colour).
-These reliable candidate \ha-emitters
-are marked by red triangles in the colour-colour diagram
-and their photometry is detailed in Appendix~\ref{app:emitters}.
-
-\begin{figure}
-  \includegraphics[width=0.45\textwidth]{./plots/sh2-82-ccd.pdf}
-    \caption{$r$-$i$/$r$-\ha\ diagram for the rectangular region of 
-    20-by-15 arcmin centred on the H{\sc ii} region Sh 2-82,
-    which is the area shown in Figure~\ref{fig:mosaic_iphas}.    
-    The diagram shows all objects in the catalogue
-    which have been flagged as \emph{reliable} and are brighter
-    than $r<20$ (grey circles).
-    The unreddened main sequence is indicated by a solid line,
-    while the main sequence for stars with an \ha\ emission line
-    strength of $-10\,\rm{\AA}$ EW is indicated by a dashed line,
-    following the colour simulations due to \citet{Barentsen2011a}.
-    Red triangles indicate objects which have been identified as
-    as highly likely \ha-emitters (see text).}
-    \label{fig:emitters}
-\end{figure}
-
-The spatial distribution of our six candidate emission-line objects
-is marked by yellow triangles in the colour mosaic (Fig.~\ref{fig:mosaic_iphas}).
-They are likely to be genuine young stars for two reasons.
-Firstly, two of our candidates have recently been identified as likely
-candidate Young Stellar Objects (YSO)
-in an investigation of the region by \cite{Yu2012}.
-In their study, the authors used 2MASS and Spitzer data
-to select likely young stars by selecting objects with
-circumstellar disks based on the infrared colour excess.
-Secondly, we find that the four remaining objects are 
-also detected in the Spitzer-based image,
-although their colours are less extreme
-than those identified as likely YSOs by \citeauthor{Yu2012}.
-
-Prior to IPHAS, this region was essentially unstudied
-at faint magnitudes in visible light.
-\cite{Lahulla1985} reported magnitudes for 8 stars in the optical at 
-$V < 15$. In contrast, the IPHAS catalogue offers photometry
-for NN stars in the region down to $r<20$.
-This demonstrates the ability of IPHAS to providing complimentary
-for the wealth of poorly-studied star-forming regions
-at low Galactic latitudes,
-which have been unveiled in recent years
-by the observed wealth of star-forming ``bubbles''
-at mid-infrared wave%\cite{Bica2003} was the first to identify a
-%near-infrared cluster
-%associated with this region using near-infrared 2MASS data,
-%and only recently has this cluster been confirmed
-%in a short investigation which combined 2MASS
-%and Spitzer data \cite{Yu2012}.
-%Many of the young stars identified by \citeauthor{Yu2012}
-%can be seen as pink and red stars in Fig.~\ref{fig:mosaic_spitzer}.lengths \cite[e.g.][]{Churchwell2006,Simpson2012}.
-%In fact the cluster of young stars associated with Sh 2-82
-%was only first reported by \cite{Bica2003},
-%remained unstudied until the study by \cite{Yu2012}.
-
-%Amongst the six emitters, N are detected at SNR X
-%in the Spitzer X micron band.
-
-%This demonstrates how IPHAS is providing complimentary
-%data for some of the thousands of star-forming ``bubbles'' have been unveiled at mid-infrared wavelengths \cite[e.g.][]{Churchwell2006,Simpson2012}.
-%Optically-unveiled members of such star-forming regions
-%may help to constrain extinctions and distances towards such regions \citep{Barentsen2013}.
-%Moreover, in \cite{Barentsen2011a} we showed
-%how the history of star formation in a region,
-%and the possible triggered nature, may be discussed.
-%This is particularly true
-
-%TODO: annotate -10 EW text on plot?
-%TODO: mention Gaia, e.g. IPHAS for transient identification?
-
-%\subsection{Caveats and lessons learnt}
-%\cite{Bica2003} was the first to identify a
-%near-infrared cluster
-%associated with this region using near-infrared 2MASS data,
-%and only recently has this cluster been confirmed
-%in a short investigation which combined 2MASS
-%and Spitzer data \cite{Yu2012}.
-%Many of the young stars identified by \citeauthor{Yu2012}
-%can be seen as pink and red stars in Fig.~\ref{fig:mosaic_spitzer}.
-
-\section{Conclusions and future work}
-\label{sec:conclusions}
-
-A new data release for the IPHAS survey was presented,
-taking the coverage up to over 90\% of the Northern Galactic Plane 
-at $|b|<5^\circ$
-and providing a uniform photometric calibration
-for the first time.
-We explained the data reduction and quality control procedures that
-were applied, described and tested the new global photometric calibration,
-and detailed the construction of the user-friendly source catalogue.
-
-The observations included in this release
-were found to achieve a median seeing of 1.1 arcsec
-and a $5\sigma-$depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$, \ha$=20.3\pm 0.3$.
-The global calibration and photometric repeatability
-is accurate at the level of $\sigma=0.03$ mag,
-providing a significant improvement over the 
-previous data release.
-The source catalogue provides the best-available
-single-epoch astrometry and photometry
-for 219~million unique sources.
-
-The data-taking strategy developed for IPHAS
-have since been reapplied to carry out a companion INT/WFC survey called UVEX
-in U/$g$/$r$ \citep{Groot2009},
-and also southern counterpart to IPHAS and UVEX 
-is being carried out in $u$/$g$/$r$/$i$/\ha\ 
-called VPHAS+ (Drew et al, in press).
-We hope to re-use the experience gained by this data release
-to create similar releases for these companion surveys.
-
-In future work, we aim to draw upon the PanSTARRS photometric
-survey to further improve the accuracy of our global calibration.
-We will also aim to correct the photometry for the
-radial field distortions.
-%We may also produce a catalogue which details all detections.
-
-%TODO: mention Gaia?
-
-\section*{Data access and source code}
-\label{sec:dataaccess}
-
-The catalogue is made available through the Vizier
-catalogue tool (http://vizier.u-strasbg.fr),
-where it is known as the ``IPHAS DR2 Source Catalogue''
-(catalogue ID ???).
-The full catalogue can also be downloaded in its entirety
-from the IPHAS website (www.iphas.org) as a collection 
-of binary FITS tables which comprise 50\,GB,
-which is accompanied by a script
-to ingest the data into a PostgreSQL database.
-Our website also provides access to the pipeline-processed
-imaging data, which we have updated to include
-the re-calibrated DR2 zeropoint in the image headers.
-
-The source code that was used to generate
-the catalogue is available at
-https://github.com/barentsen/iphas-dr2
-
-\section*{Acknowledgments}
-
-The IPHAS survey was carried out 
-at the Isaac Newton Telescope (INT).
-The INT is operated on the island of La Palma
-by the Isaac Newton Group
-in the Spanish Observatorio del Roque de los Muchachos
-of the Instituto de Astrofisica de Canarias.
-All data were processed 
-by the Cambridge Astronomical Survey Unit,
-at the Institute of Astronomy in Cambridge.
-
-Preparation of the catalogue was eased greatly
-by a number of software packages,
-including the Python modules
-AstroPy \citep{Astropy},
-APLpy, NumPy and SciPy,
-the PostgreSQL database software,
-the TOPCAT and STILTS packages \citep{Taylor2005,Taylor2006},
-and the Montage software maintained by NASA/IPAC.
-We also made use of the SIMBAD, Vizier and Aladin \citep{Aladin} tools operated at CDS, Strasbourg, France.
-
-This research made extensive use of
-several complementary photometric surveys.
-Our global calibration was aided
-by the AAVSO Photometric All-Sky Survey (APASS),
-funded by the Robert Martin Ayers Sciences Fund.
-The calibration was tested against the
-Sloan Digitized Sky Survey (SDSS),
-funded by the Alfred P.
-Sloan Foundation, the Participating Institutions, the National Science Foundation, the U.S.
-Department of Energy, the National Aeronautics and Space Administration, the Japanese Monbukagakusho, the Max Planck Society, and the Higher Education Funding Council for England.
-The astrometric pipeline reduction made
-significant use of the Two Micron All Sky Survey (2MASS),
-which is a joint project 
-of the University of Massachusetts
-and the Infrared Processing and Analysis Center/
-California Institute of Technology,
-funded by NASA and the NSF.
-
-GB and JED acknowledge the support of a grant
-from the Science \& Technology Facilities Council
-of the UK (STFC, ref ST/J001335/1).
-HJF is in receipt of an STFC postgraduate studentship.
-
-\bibliographystyle{mn2e}
-\bibliography{dr2}
-
-\appendix
-\input{app_columns.tex}
-
-\section{Candidate emission-line objects towards Sh 2-82}
-\label{app:emitters}
-\begin{table*}
-    \begin{tabular}{lccc}
-    \toprule
-    Name & $r$ & $i$ & \ha  \\
-    \midrule
-IPHAS2 J192954.40+181026.1& $17.69\pm0.01$ & $16.12\pm0.01$ & $16.19\pm0.01$ \\
-IPHAS2 J193011.01+182051.2& $18.55\pm0.02$ & $16.95\pm0.02$ & $17.31\pm0.02$ \\
-IPHAS2 J193021.52+181954.5& $19.72\pm0.05$ & $17.94\pm0.03$ & $18.47\pm0.04$ \\
-IPHAS2 J193024.45+181938.3& $19.31\pm0.04$ & $17.55\pm0.02$ & $17.99\pm0.03$ \\
-IPHAS2 J193033.00+181609.3& $18.25\pm0.01$ & $16.91\pm0.01$ & $16.92\pm0.01$ \\
-IPHAS2 J193042.48+182317.4& $19.96\pm0.03$ & $18.11\pm0.03$ & $18.48\pm0.03$ \\
-    \bottomrule
-    \end{tabular}
-    \caption{Candidate \ha-emitters towards Sh 2-82.}
-    \label{tbl:emitters}
-\end{table*}
-
-\label{lastpage}
-
-\end{document}\documentclass[useAMS,usenatbib]{mn2e}
-\usepackage{amssymb,amsmath}
-\usepackage[pdftex]{graphicx}
-\usepackage{longtable} 
-%\usepackage{lscape}
-\usepackage{float}
-\usepackage{booktabs}
-\usepackage{aas_macros}
-
-\def\ha{\mbox{H$\rm \alpha$}}
-\def\arcsec{$''$}
-\def\arcmin{$'$}
-\def\deg{$^{\circ}$}
-\def\micron{\mbox{$\mu$m}}
-
-\title[IPHAS Data Release 2]{The Second Data Release 
-of the INT Photometric H$\alpha$ Survey 
-of the Northern Galactic Plane (IPHAS DR2)}
-
-\author[G. Barentsen
-et. al]{G. Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
-H. J. Farnhill$^{1}$,
-J. E. Drew$^{1}$,
-M. J. Irwin$^{4}$, \newauthor
-E. A. Gonz$\acute{\rm{a}}$lez-Solares$^{4}$,
-R. Greimel$^{2}$,
-B. Mizalski$^{3}$,
-C. Ruhland$^{1}$, \newauthor plus many friends.
-\newauthor\\
-$^{1}$Centre for Astrophysics Research, STRI, University of Hertfordshire, College Lane Campus, Hatfield AL10 9AB\\
-$^{2}$Institute for Geophysics, Astrophysics and Meteorology, Institute of Physics, Karl-Franzens-Universit$\ddot{a}$t Graz, Universit$\ddot{a}$tsplatz 5, 8010 Graz, Austria\\
-$^{3}$Building, Institute, Street Address, City, Code, Country\\
-$^{4}$Institute of Astronomy, Madingley Road, Cambridge CB3 0HA}
-
-\begin{document}
-\date{Current draft typeset \today}
-\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2013}
-
-\maketitle
-
-\label{firstpage}
-
-\begin{abstract} % BACKGROUND, OBJECTIVE, METHODS, RESULTS, CONCLUSIONS
-The INT/WFC Photometric H$\alpha$ Survey 
-of the Northern Galactic Plane (IPHAS)
-is a 1800 deg$^2$ imaging survey
-covering the entire northern Milky Way at $|b| < 5^\circ$
-in the $r$, $i$ and \ha\ filters 
-using the Wide Field Camera (WFC) 
-on the 2.5-meter Isaac Newton Telescope (INT).
-This data release presents the first 
-uniformly-calibrated source catalogue
-to have been extracted from the survey,
-providing single-epoch photometry
-for 219 million unique sources
-across 92\% of the survey.
-The observations were carried out between 2003 and 2012
-at a median seeing of 1.1 arcsec
-to a depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$ and \ha$=20.3\pm 0.3$
-($5\sigma$ limits, Vega system).
-We explain the data reduction 
-and quality control procedures,
-describe and test the new uniform photometric calibration,
-and detail the construction of the source catalogue
-and its quality warning flags.
-We find that the new calibration is accurate to
-$\sigma=0.03$ mag.
-Finally, we demonstrate the ability of the 
-catalogue's unique
-$r$-$i$/$r$-\ha\ colour-colour diagram to
-(i) characterise stellar populations and extinction regimes
-towards different Galactic sightlines
-and (ii) select reliable candidate \ha-emission line objects.
-The catalogue which accompanies this paper
-provides the much-needed visible-light counterpart
-to several infrared surveys of the Galactic Plane,
-including many poorly-studied star-forming regions and ``bubbles'',
-and provides images for some of the most
-crowded regions to be faced by Gaia.
-\end{abstract}
-
-\begin{keywords}
-catalogues, surveys, stars: emission line, Be, Galaxy: stellar content
-\end{keywords}
-
-\section{Introduction}
-
-Since the first data release in 2008, 
-the INT/WFC Photometric H$\alpha$ Survey 
-of the Northern Galactic Plane (IPHAS)
-has provided new insights into the contents and structure 
-of our own backyard, the Milky Way. 
-The original motivation for undertaking 
-this large-scale programme of observation
--- spanning almost a decade 
-and using more than 300 nights 
-at the Isaac Newton Telescope (INT) in La Palma -- 
-was to provide the digital update 
-to the photographic northern H$\alpha$ surveys 
-of the mid-20th century. 
-By increasing the sensitivity 
-with respect to these previous surveys 
-by a factor $\sim$1000 (7 magnitudes), 
-it was envisaged that IPHAS would allow 
-the limited bright samples of Galactic emission line objects 
-available at the outset \citep[e.g.][]{Kohoutek1999}, 
-to be extended into larger, deeper, more statistically-robust samples 
-that in turn could better inform our understanding 
-of the early and late stages of stellar evolution. 
-This aim has begun to be realised through a
-range of published studies including: 
-a preliminary catalogue of candidate emission line objects \citep{Witham2008};
-discoveries of new northern symbiotic stars \citep{Corradi2008, Corradi2010}; 
-new cataclysmic variables \citep{Witham2007}; 
-new groups of young stellar objects \citep{Vink2008,Barentsen2011a};
-along with discoveries of new and remarkable planetary nebulae 
-\citep{Mampaso2006, Corradi2011, Viironen2011}.
-
-Over the years it has become apparent that the legacy of IPHAS 
-will extend beyond the traditional \ha\ applications 
-of identifying emission line stars and nebulae. 
-Through the provision of $r$, $i$ broadband photometry 
-alongside H$\alpha$ data,
-IPHAS has created the opportunity 
-to study Galactic Plane populations 
-in a new way.
-For example, the survey’s unique $r-H\alpha$ colour, 
-when combined with $r-i$,
-has been shown to provide simultaneous constraints 
-on intrinsic stellar colour and interstellar extinction \citep{Drew2008}. 
-This has opened the door 
-to a wide range of Galactic science applications, 
-including the mapping of extinction across the Plane in three dimensions
-and the probabilistic inference of stellar properties
-\citep{Sale2009, Sale2010, Giammanco2011, Sale2012, Barentsen2013}. 
-In effect, the availability of narrowband H$\alpha$ 
-alongside $r, i$ magnitudes
-provides coarse spectral information for huge samples of stars 
-which are otherwise too faint or numerous 
-to be targeted by spectroscopic surveys.
-For such science applications to succeed however, 
-it is vital that the imaging data is transformed 
-into a homogeneously calibrated photometric catalogue, 
-in which quality problems 
-and duplicate detections of the same source 
-are flagged.
-
-The first release of IPHAS data, 
-covering roughly half the survey footprint,
-was made in late 2007 \citep{Gonzalez-Solares2008}. 
-At the time the data were insufficiently complete 
-to support a homogeneously calibrated source catalogue.
-The goal of this paper is to present the next release 
-that takes the coverage up to over 90 percent of the survey area 
-and includes a uniform calibration.
-In this work we aim to
-(i) explain the data reduction 
-and quality control procedures that were applied,
-(ii) describe and test the new global photometric calibration, and 
-(iii) detail the construction and demonstrate the use 
-of the source catalogue that has been extracted 
-from the newly re-calibrated data.
-
-In \S\ref{sec:observations} we start by recapitulating the key points
-of the survey design and strategy.
-In \S\ref{sec:reduction} we describe the data reduction
-and quality control procedures.
-In \S\ref{sec:calibration} we explain the uniform re-calibration
-and test our results against the Sloan Digital Sky Survey (SDSS).
-In \S\ref{sec:catalogue} we explain how the source catalogue was compiled.
-In \S\ref{sec:discussion} we discuss the properties of the catalogue,
-and finally in \S\ref{sec:demonstration} we demonstrate
-the health of the release by demonstrating its scientific exploitation.
-In \S\ref{sec:conclusions} we conclude and outline
-our future ambitions.
-
-\section{Observations}
-\label{sec:observations}
-
-The detailed properties of the IPHAS observing programme 
-have been presented before 
-by \citet{Drew2005} and \citet{Gonzalez-Solares2008}. 
-To set the stage for this release, we briefly remind of some key points.
-IPHAS is a 1800~sq. deg. imaging survey of the northern Galactic Plane, 
-providing images and photometry in Sloan $r, i$ 
-along with narrowband H$\alpha$. 
-It is carried out using the Wide Field Camera (WFC) 
-on the 2.5-meter Isaac Newton Telescope (INT) in La Palma. 
-It is the first digital survey to offer comprehensive CCD photometry
-of point sources in the Galactic Plane at visible wavelengths, 
-and does so down to a limiting magnitude of $\sim$20th.
-The IPHAS footprint on the northern sky spans a box 
-of roughly 180 by 10 degrees, 
-taking in the entire northern Galactic Plane 
-at latitudes $-5^{\circ} < b < +5^{\circ}$ 
-and longitudes $30^{\circ} < l < 215^{\circ}$.
-
-The Wide Field Camera is a mosaic of 4 CCDs 
-that captures a sky area of close to 0.29 square degrees.
-To cover the entire northern Plane with some overlap,
-the survey area was divided into 7635 telescope pointings.
-Each of these pointings is accompanied by an offset position
-at a displacement of $+$5 arcmin in declination 
-and $+$5 arcmin in right ascension,
-to deal with inter-CCD gaps, detector imperfections,
-and to enable quality checks. 
-The basic unit of observation hence
-amounts to $2 \times 3$ exposures, 
-in which each of the 3 survey filters is exposed at 2 offset sky positions, 
-within an elapsed time of 10 minutes.
-We shall refer to the unit of 3 exposures at the same position 
-as a \emph{field},
-and the combination of two fields at a small offset as a \emph{field pair}.
-The survey hence contains 15270 fields grouped into 7635 field pairs.
-To achieve the desired survey depth of 20th magnitude or fainter, 
-the filter exposure times were set at 120 sec (narrowband \ha), 
-30 sec ($r$) and 10 sec ($i$)
-in the majority of the survey observations.\footnote{In 2003 
-the $r$-band exposure time was 10~sec instead of 30~sec,
-and since Oct 2010 the $i$-band exposure time 
-has been increased from 10~sec to 20~sec.}
-
-Data-taking began in the second half of 2003, 
-and every field had been observed at least once by the end of 2008.
-At that time only 76 percent of the field pairs 
-satisfied our minimum quality criteria however,  
-often due to the effects of clouds, poor seeing, or technical faults
-(the quality criteria will be detailed in the next section). 
-Since then, a programme of repeat observations has been in place 
-to improve data quality. 
-As a result, 92\% of the survey 
-now benefits from quality-approved data.
-The most recent observations which are part of this release
-were obtained in November 2012.
-
-\begin{figure*}
-        \includegraphics[width=1\linewidth]{./plots/footprint/footprint_small.png}
-        \caption{Survey area showing the footprints
-        of all the quality-approved IPHAS fields
-        which have been included in this data release.
-        The area covered by each field has been coloured black
-        with a semi-transparent opacity of 20\%,
-        such that regions where fields overlap are darker.
-        The IPHAS strategy is to observe each field twice
-        with a small offset,
-        and hence the vast majority of the area 
-        is covered twice (dominant gray colour).
-        There are small overlaps between all the neighbouring fields
-        which can be seen as a honeycomb-style pattern
-        of dark gray lines across the survey area.
-        Regions with incomplete data are apparent as white gaps (no data) 
-        or as the lightest shade of gray
-        (denoting that only the offset position is missing).}
-        \label{fig:footprint}
-\end{figure*}
-
-Figure~\ref{fig:footprint} shows the footprint 
-of the quality-approved observations included in this work. 
-The fields which remain missing 
--- covering 7 percent of the survey area --
-are predominantly located towards the Galactic anti-center 
-at $l > 120^o$.
-Fields at these longitudes can only be accessed from La Palma 
-in the months of November-December,
-which is when the weather and seeing conditions are often poor
-at the INT and observing attempts have failed repeatedly.
-To enable the survey to be brought to completion, 
-a decision was made recently to limit repeats in this area 
-to individual fields requiring replacement,
-i.e. fresh observations in all 3 filters may only be obtained 
-at one of the two offset positions, 
-if the data for the partner offset has already passed quality control.  
-The catalogue is structured such that it is clear 
-where a contemporaneous observation of both halves of a field pair
-is not available.
-
-\section{Data reduction and quality control}
-\label{sec:reduction}
-
-\subsection{Initial pipeline processing}
-
-All raw data obtained with the INT were transferred
-to the Cambridge Astronomical Survey Unit (CASU) 
-for initial processing and archival.
-The procedures used by CASU were originally devised
-for the INT Wide Field imaging Survey \citep[WFS;][]{McMahon2001,Irwin2005},
-which was a 200 deg$^2$ survey programme carried out 
-between 1998 and 2003 after the WFC was commissioned.
-Because IPHAS uses the same telescope and camera combination,
-we have been able to benefit from the existing WFS pipeline.
-A detailed description of the processing steps 
-is found in \citet{Irwin2001}.
-Its application to IPHAS has previously been described
-by \citet{Drew2005} and \citet{Gonzalez-Solares2008}
-and much of the source code is available 
-on line\footnote{http://casu.ast.cam.ac.uk/surveys-projects/software-release}. 
-In brief, the pipeline takes care of bias subtraction,
-linearity correction, flat-fielding,
-gain correction and de-fringing.
-
-The reduced images are then stored in a multi-extension FITS file 
-with a primary header describing the characteristics
-(position, filter, exposure time, etc.) 
-and four compressed image extensions 
-corresponding to each of the four CCDs.
-Source detection and characterisation is then carried out 
-using the \textsc{imcore} tool \citep{Irwin1985,Irwin1997}.
-The flux of each source is measured using both
-the peak pixel height (i.e. a square 0.33$\times$0.33\arcsec\ aperture)
-as well as a series of circular apertures of increasing diameter 
-(1.2\arcsec, 2.3\arcsec, 3.3\arcsec, 4.6\arcsec\ and 6.6\arcsec).
-
-The local background levels are estimated 
-by computing the sigma-clipped median
-flux in a grid of 64$\times$64 pixels (21$\times$21\arcsec)
-across the image,
-which is then interpolated to obtain an estimate 
-of the background level at each pixel.
-These sky levels are subtracted from the aperture photometry and
--- when required --
-a deblending routine is applied which also attempts to remove
-the contamination from any other nearby sources.
-Whilst this approach works very well 
-across the vast majority of the survey area,
-the Galactic Plane unavoidably contains crowded regions 
-with large numbers of overlapping sources
-or rapidly spatially-varying nebulosity,
-in which case aperture photometry must always be interpreted 
-with caution.
-In \S\ref{sec:catalogue} we will explain that overlapping
-sources to which the deblending routine was applied 
-are flagged in the catalogue using the \emph{deblend} warning flag.
-
-Finally, an astrometric solution is determined
-based on the 2MASS point source catalog \citep{Skrutskie2006},
-which itself is calibrated 
-in the International Celestial Reference System (ICRS).
-A provisional photometric calibration is also provided 
-based on the average zeropoint
-determined from a set of standard stars observed in the same night.
-Sources are classified morphologically
--- stellar, galaxy or noise --
-based on the curve-of-growth determined
-from measuring the source intensity in a series of growing apertures.
-Finally, the resulting source detection tables are also stored 
-in multi-extension FITS files.
-
-At the time of preparing DR2,
-the CASU pipeline had processed
-74\,195 IPHAS exposures 
-in which a total of 1.9~billion \emph{candidate sources} were detected 
-at the sensitive default detection level of 1.25\,$\sigma$
--- unavoidably including spurious detections, artefacts and
-duplicate detections 
-(in \S\ref{sec:catalogue} we will explain
-how these have been removed or flagged).
-The pipelined data set -- comprising 2.5~terabyte of FITS files --
-was then transferred to the University of Hertfordshire
-for the purpose of transforming the raw
-detection tables into a reliable source catalogue which is 
-(i) quality-controlled,
-(ii) homogeneously calibrated, and 
-(iii) contains user-friendly columns and warning flags.
-It is these post-processing steps which are explained below.
-
-\subsection{Quality control}
-\label{sec:qc}
-
-Observing time for IPHAS was obtained
-on a semester-by-semester basis
-through the traditional time allocation committees 
-of the Isaac Newton Group of telescopes,
-which are competitive and invariably over-subscribed.
-For this reason, we attempted to utilise 
-\emph{all} the nights allocated to IPHAS,
-even those which were partially or entirely non-photometric
-or otherwise affected by technical problems 
-(e.g. electronic noise or telescope tracking problems).
-Any unsuitable data that was taken as a result
-of this strategy was subsequently flagged and rejected
-using a series of seven quality criteria,
-which ensure a reliable and homogeneous level of quality
-across the data release:
-
-\begin{figure}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/depth_r.pdf} 
-    \end{minipage}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/depth_i.pdf} 
-    \end{minipage}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/depth_h.pdf} 
-    \end{minipage}
-    \caption{Distribution of the 5$\sigma$ limiting magnitude
-             across all quality-approved survey fields
-             for $r$ (top), $i$ (middle) and \ha\ (bottom).
-             Fields with a limiting magnitude brighter than
-             20th ($r$) or 19th (\ha/$i$) were rejected
-             from the data release.
-             The $r$-band depth is most sensitive 
-             to the presence of the moon above the horizon, 
-             which is evidenced by the wide and bi-model shape
-             of its distribution.}
-    \label{fig:depth}
-\end{figure}
-
-(1) \emph{Depth.} 
-We discarded any exposures for which the $5\sigma$ limiting magnitude 
-was worse than 20th magnitude in the $r$-band
-or worse than 19th in $i$ or \ha. 
-Such data were typically obtained during poor weather or full moon.
-Most observations fared significantly better than these limits.
-Figure~\ref{fig:depth} presents the distribution of limiting magnitudes
-for all quality-approved fields,
-which shows a mean depth of 
-$21.2\pm0.5$ ($r$), $20.0\pm0.3$ ($i$) and $20.3\pm0.3$ (\ha).
-We found that the depth achieved depended 
-most strongly on the presence of the moon,
-which was above the horizon during 62 percent 
-of our observations and explains the wide and bi-modal shape
-of the $r$-band limiting magnitude distribution 
-(top panel in Fig.~\ref{fig:depth}).
-In contrast, the depth attained in $i$ and \ha\ 
-is less sensitive to moonlight
-and the distribution of their depths
-is hence more narrow
-(middle and bottom panel in Fig.~\ref{fig:depth}).
-
-Our minimum limiting magnitude criteria
-have led us to exclude 9\% of the pipelined data.
-We note that much of the excluded data may nevertheless be useful
-for e.g. time-domain studies of bright stars.
-The detection tables for any such discarded data are made
-available through our website (www.iphas.org)
-but are ignored in what follows.
-
-(2) \emph{Ellipticity.} 
-The ellipticity of a point source,
-defined as $e = 1 - b / a$ 
-with $b$ the semi-minor and $a$ the semi-major axis,
-is a morphological measure of the elongation of the point spread function.
-It is expected to be zero (circular) across the field 
-in a perfect imaging system,
-but it is slightly non-zero in any real telescope data 
-due to optical distortions and tracking errors.
-The mean ellipticity across a field in the IPHAS data set 
-is $0.09\pm0.04$.
-There have been sporadic episodes with higher ellipticities however
-due to mechanical glitches in the telescope tracking system.
-For this reason, we rejected exposures in which the mean ellipticity
-across the detector exceeded $e > 0.3$,
-which is when the photometric measurements delivered by the pipeline
-were found to become degraded.
-Only 0.4\% of the exposures were discarded on this basis.
-
-(3) \emph{Seeing.} 
-The survey originally aimed to obtain data 
-at seeing better than 1.7 arcsec.
-This target is currently attained across 86\% of the footprint,
-in particular at early longitudes,
-e.g. 92\% of the fields at $l<120^\circ$ are better than 1.7 arcsec.
-Figure~\ref{fig:seeing} presents the distribution
-of the mean seeing for all the quality-approved fields.
-We find a median value of 1.1~arcsec in $r$/\ha\
-and 1.0~arcsec in $i$.
-In the $r$-band, 90\% of the data is better than 1.5~arcsec
-and 10\% is better than 0.8~arcsec.
-To improve the area covered by this data release,
-we have decided to include the small fraction of data
-that was obtained under seeing up to 2.5 arcsec,
-so that only 1 per cent of the pipelined exposures
-had to be excluded.
-In \S\ref{sec:catalogue} we will explain
-that the catalogue is structured
-such that the information on the seeing is included,
-and that the photometry listed
-is based on the exposures with the best-available seeing.
-
-\begin{figure}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/seeing_r.pdf} 
-    \end{minipage}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/seeing_i.pdf} 
-    \end{minipage}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/seeing_ha.pdf} 
-    \end{minipage}
-    \caption{Seeing distribution across all
-             quality-approved survey fields
-             for $r$ (top), $i$ (middle) and \ha\ (bottom).}
-    \label{fig:seeing}
-\end{figure}
-
-(4) \emph{Photometric repeatability.} 
-The IPHAS field-pair observing strategy 
-ensures that every pointing is immediately followed 
-by an offset pointing at a displacement of $+$5 arcmin in Dec 
-and $+$5 arcmin in RA.
-This allows pairs of images to be checked 
-for the presence of clouds or electronic noise.
-To exploit this information,
-the overlap regions of all field pairs were systematically cross-matched
-to verify the consistency of the photometric measurements
-for stars observed in both pointings.
-We rejected field pairs in which more than 2\% of the stars 
-showed an inconsistent measurement at the level of 0.2 mag,
-or more than 26\% were inconsistent at the level of 0.1 mag.
-These limits were determined empirically by inspecting
-the images and photometry by eye.
-11\% of the data was rejected as part of this step.
-
-(5) \emph{Source density mapping.}
-Spatial maps showing the number density of the detected sources
-down to 20th magnitude were created to verify the health
-of the data and to check for unexpected artefacts.
-%(e.g. bright satellite trails)
-In particular, we created density maps
-which showed the number of \emph{unique} sources
-obtained by cross-matching the detection tables of
-all three bands with a maximum matching distance of 1 arcsec.
-The success of such a cross-matching procedure crucially depends
-on the accuracy of the astrometric solution in each band,
-and hence we were able to detect and correct
-exposures with poor astrometry by inspecting the density map
-for spurious density variations.
-
-(6) \emph{Visual examination.}
-All images and their associated photometric colour/magnitude diagrams
-were inspected by a team of 20 survey team members, 
-such that each image in the data release 
-was looked at by at least three different pairs of eyes.
-Images deemed unsuitable were flagged, investigated and
-excluded from the release if necessary. 
-6\% of the attempts to observe a field
-were placed on a \emph{black-list}
-for this reason, most commonly due to the obvious presence
-of clouds or extreme levels of scattered moonlight.
-Such fields were often rejected as part of 
-one or more of the above quality criteria as well.
-
-(7) \emph{Contemporaneous field data.} 
-Finally, only exposures which are part of a sequence 
-of three consecutive images of the same field (H$\alpha$/$r$/$i$) 
-were considered for inclusion in the release. 
-This ensures that the three bands for a given field
-are observed at nearly the same time --  
-essentially always within 5 minutes.
-An exception was made for 9 fields where the three exposures 
-could not be obtained within the same night
-but for which the time gap did not exceed 48 hours.
-We note that the exact epoch of the magnitude in each band
-is included in the source catalogue
-(columns \emph{rMJD}, \emph{iMJD}, \emph{haMJD}).
-
-The above criteria were satisfied 
-for 14115 out of the 15270 fields (92\%).
-In a few cases more than one successful attempt to observe
-a field was available due to the fact that slightly stricter
-quality criteria were adopted in the initial years of the survey.
-In such cases, only the attempt 
-with the best seeing and depth was selected
-for inclusion in the release, because the focus 
-of this work is to deliver the most reliable
-measurement at a single epoch.
-Those interested in any of the rejected data 
-may nevertheless access the full set of detection tables 
-on line.
-
-\section{Photometric calibration}
-\label{sec:calibration}
-
-Having obtained a quality-approved set of observations,
-we now turn to the problem of placing the data
-onto a uniform photometric scale.
-
-\subsection{Provisional nightly calibration}
-
-For the purpose of providing an initial calibration 
-of the $r$ and $i$ broadband fluxes,
-photometric standard fields were observed every night.
-The standards were chosen from a list based on 
-the \cite{Landolt1992} and Stetson (http://cadcwww.dao.nrc.ca/standards) 
-objects.
-Two or three standard fields were observed 
-during the evening and morning twilight,
-and at intervals of 2-3 hours throughout the night.
-The CASU pipeline automatically identified the observed standards 
-and used them to determine a sigma-clipped average zeropoint \textsc{magzpt}
-for each night and filter,
-such that the number counts $DN$ 
-in the pipeline-corrected CCD frames
-relates to a magnitude $m$ as:
-\begin{equation}
-\begin{split}
-   m  = & \textsc{magzpt} - 2.5 \log_{10}( DN / \textsc{exptime} ) \\
- &  - \textsc{extinct}\cdot(\textsc{airmass}-1) - \textsc{apcor} - \textsc{percorr}
-\label{eqn:mag}
-\end{split}
-\end{equation}
-where \textsc{exptime} is the exposure time in seconds,
-\textsc{extinct} is the atmospheric extinction coefficient 
-(typically 0.09 for $r$ and 0.05 for $i$ in La Palma),
-\textsc{airmass} is the normalised optical path length 
-through the atmosphere and
-\textsc{apcor} is a correction for the flux
-lost outside of the aperture used.
-Finally, \textsc{percorr} is a correction based on the difference
-between the median dark sky for a CCD against the median for all the CCDs 
-and as such is an ancillary correction 
-to account for sporadic gain variations. 
-All these quantities correspond to header keywords in the 
-multi-extension FITS files produced by the CASU pipeline.
-
-The zeropoint was determined such that the resulting magnitude system
-is in the WFC system that uses the SED of Vega 
-as the zero colour reference object. 
-Colour equations were used to transform between the IPHAS passbands 
-and the Johnson-Cousins system 
-of the published standard star photometry.
-The entire procedure has been found to deliver zeropoints which 
-are accurate at the level of 1-2\% 
-in photometric conditions \citep{Gonzalez-Solares2011}.
-
-Unlike the broadbands, 
-standard star photometry is not available in the literature 
-for the H$\alpha$ passband
-and hence there is no formally recognised flux scale 
-for the narrowband.
-We can specify here however 
-that the integrated in-band energy flux for Vega 
-in the IPHAS \ha\ filter 
-is $1.52 \times 10^{-7}$ ergs\,cm$^{-2}$\,s$^{-1}$ 
-at the top of the Earth's atmosphere,
-which is the flux obtained by folding 
-Vega's SED with the filter throughput curve 
-corrected for atmosphere and detector quantum efficiency
-\citep[using the method explained by][]{Drew2005}.
-This is 3.08 magnitudes less than the flux captured 
-by the much broader $r$ band
-which includes the \ha\ band.
-Hence to assure zero colour relative to the broadbands,
-we set the default zeropoint for the narrowband to be:
-\begin{equation}
-\textsc{magzpt}_{H\alpha} = \textsc{magzpt}_r - 3.08.
-\label{eqn:zpha}
-\end{equation}
-
-\subsection{Uniform re-calibration}
-
-Despite the best efforts made to obtain a nightly calibration,
-large surveys naturally possess field-to-field variations
-at the level of 0.1 mag
-due to atmospheric fluctuations during the night
-and imperfections in the pipeline or the instrument
-(e.g. the WFC is known to suffer from sporadic glitches
-in the timing of exposures).
-Such variations need to be corrected for 
-during a global re-calibration procedure.
-Notable past examples include the global re-calibration 
-of the Two Micron All Sky Survey \citep[2MASS;][]{Nikolaev2000},
-the Sloan Digital Sky Survey \citep[SDSS;][]{Padmanabhan2008}
-and the Panoramic Survey Telescope 
-and Rapid Response System survey \citep[Pan-STARRS;][]{Schlafly2012},
-which all achieved photometry 
-that is globally consistent to within 1--2\%.
-
-Surveys which observe identical stars at different epochs
-can use the repeat measurements to ensure a uniform calibration.
-For example, 2MASS attained its global calibration
-by observing six standard fields each hour, 
-allowing zeropoint variations to be tracked 
-over very short timescales \citep{Nikolaev2000}.
-Alternatively, the SDSS and PanSTARRS surveys could benefit
-from revisiting regions in their footprint to 
-carry out a so-called \emph{ubercalibration} procedure,
-in which multiple measurements of stars at different epochs
-are used to fit the calibration parameters
-\citep{Padmanabhan2008,Schlafly2012}.
-
-Unfortunately these schemes cannot be applied
-directly to IPHAS
-for two reasons. 
-Firstly, the survey was carried out 
-in competitive observing time
-on a non-dedicated telescope, 
-rendering the 2MASS approach 
-of observing standards at a very high frequency
-prohibitively expensive
--- in part because standard fields 
-are very scarce within the Galactic Plane.
-Secondly, the aim of IPHAS is to obtain magnitudes at a single epoch
-and hence stars are not normally observed at more than one epoch,
-unless they happen to fall on the narrow overlap region 
-between two neighbouring fields.
-
-Although the IPHAS data does contain a significant number of 
-inter-field repeat measurements
-due to the small overlaps
-between neighbouring fields,
-we have found the information contained
-in these regions to be insufficient
-to constrain the calibration parameters.
-This is because photometry at the extreme edges of the WFC
--- where neighbouring fields overlap -- 
-is itself prone to systematics at the level of 1-2\%.
-The cause of these errors is thought to include 
-the use of twilight sky flats in the pipeline,
-which are known to be imperfect for calibrating stellar photometry 
-due to stray light and vignetting \citep[e.g.][]{Manfroid1995}.
-Moreover, the illumination correction in the overlap regions
-is affected by a radial geometric distortion,
-which causes the pixel scale towards the edges 
-to increase \citep{Gonzalez-Solares2011}.
-Although these systematics are reasonably small within a single field,
-they can combine to cause artificial zeropoint gradients 
-across the survey
-when they are used to constrain a global calibration
-without external constraints.
-
-For these reasons, we have decided not to depend
-on an ubercalibration-type scheme alone,
-and have instead opted to exploit an external reference survey
--- where available --
-to bring the majority of our data onto a uniform calibration.
-This is explained next.
-
-\subsubsection{Correcting zeropoints using APASS}
-
-We have been able to benefit from the
-AAVSO Photometric All-Sky Survey
-(APASS; http://www.aavso.org/apass)
-to bring the vast majority of the survey 
-onto a uniform scale.
-Since 2009,
-APASS has been using two 20~cm-astrographs
-to survey the entire sky down to $\sim$17th magnitude
-in five filters which include Sloan $r$ and $i$.
-The most recent catalogue available 
-at the time of preparing this work was APASS DR7,
-which provided data across roughly half of the IPHAS footprint (Fig.~\ref{fig:apass_r}).
-The photometric accuracy of APASS is currently estimated 
-to be at the level of 3\% (Henden, private communication),
-which is significantly better 
-than the provisional calibration of IPHAS
-for which we estimate the $1\sigma$-error to be $\sim$10\%.
-APASS achieves its uniform accuracy 
-by measuring each star at least two times in photometric conditions
--- along with ample standard fields --
-using the large $3\times3$ square degrees field of view of its detectors.
-
-With the aim of bringing IPHAS to a similar accuracy of $\sim$3\%,
-we used the APASS catalogue to identify and adjust all IPHAS fields 
-which showed a magnitude offset larger than 3\% against APASS.
-For this purpose,
-the $r$- and $i$-band detection tables of each IPHAS field
-were cross-matched against the APASS DR7 catalogue 
-using a maximum matching distance of 1~arcsec.
-The magnitude range was limited to
-$13<r_{\rm APASS}<16.5$ and $12.5<i_{\rm APASS}<16.0$
-in order to avoid sources 
-brighter than the IPHAS saturation limit on one hand, 
-and to avoid sources near the faint detection limit of APASS 
-on the other hand.
-
-The resulting set of $\sim$220\,000 cross-matched stars were then used 
-to derive APASS-to-IPHAS magnitude transformations
-using a linear least-squares fitting routine, 
-which iteratively removed $3\sigma$-outliers to improve the fit.
-The solution converged to:
-\begin{align} 
-r_{\rm IPHAS} = r_{\rm APASS} - 0.121 + 0.032(r-i)_{\rm APASS} \label{eqn:apass_r} \\
-i_{\rm IPHAS} = i_{\rm APASS} - 0.364 + 0.006(r-i)_{\rm APASS} \label{eqn:apass_i}
-\end{align}
-The root mean square (rms) residuals of these transformations 
-are 0.041 and 0.051, respectively.
-The small colour terms in the equations
-indicate that the $r$ and $i$ filters 
-are very similar in both surveys.
-The transformations include a large fixed offset
-which is simply due to the fact that 
-APASS magnitudes are given in the AB system
-and IPHAS uses magnitudes relative to Vega.
-Separate transformations were derived for sightlines 
-with varying extinction properties to investigate the robustness
-of the transformations with respect to different reddening regimes,
-but the variations at these different
-sightlines were found to be insignificant.
-This is not surprising because heavily reddened objects 
-are naturally scarce at $r<16$.
-
-Having transformed APASS magnitudes into the IPHAS system,
-we then computed the median magnitude offset 
-for each field which contained at least 30 cross-matched stars.
-This was the case for 48\% of the fields.
-The mean offset was found to be
-$0.014\pm0.104$ for $r$ and $0.007\pm0.108$ for $i$
-(Table~\ref{tbl:offsets_before}).
-A total of 4596 fields showed a median offset
-exceeding $\pm$0.03~mag in either $r$ or $i$.
-
-We then applied the most important step in our re-calibration scheme,
-which is to adjust the provisional zeropoints of the 4596 aberrant fields
-such that their offset is brought to zero.
-This allowed the mean IPHAS-to-APASS offset 
-to be brought down to $0.000\pm0.011$ in both $r$ and $i$
-(Table~\ref{tbl:offsets_after}).
-The procedure of fitting magnitude transformations and
-correcting the zeropoints was repeated a few times to ensure 
-convergence, which was essentially reached after the first iteration.
-
-\begin{figure*}
-    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_rshift.pdf} 
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_r.pdf} 
-    \caption{Median magnitude offset in the $r$ band between IPHAS and APASS,
-             plotted on a field-by-field basis
-             prior to the re-calibration procedure.
-             Each square represents an IPHAS field
-             which contains at least 30 stars with a counterpart
-             in the APASS DR7 catalogue.
-             The colours denote the median
-             IPHAS-APASS magnitude offset in each field,
-             which was computed after applying the APASS-to-IPHAS
-             transformation to the APASS magnitudes (Eqn.~\ref{eqn:apass_r}).}
-        \label{fig:apass_r}
-    \vspace{1cm}
-    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_rshift.pdf}
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_r.pdf} 
-    \caption{Median magnitude offset in the $r$ band
-             between IPHAS and SDSS after the re-calibration
-             procedure was applied.
-             Again, each square represents an IPHAS field
-             which contains at least 30 stars
-             with a counterpart in the SDSS catalogue.
-             The colours denote the median IPHAS-SDSS magnitude offset
-             in each field,
-             which was computed after applying the SDSS-to-IPHAS
-             transformation to the SDSS magnitudes (Eqn.~\ref{eqn:sdss_r}).}
-    \label{fig:sdss_r}
-\end{figure*}
-
-\begin{figure*}
-	\vspace{2cm}
-    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_ishift.pdf} 
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_i.pdf} 
-    \caption{Same as Figure~\ref{fig:apass_r} for the $i$-band.}
-        \label{fig:apass_i}
-    \vspace{1cm}
-    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_ishift.pdf}
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_i.pdf} 
-    \caption{Same as Figure~\ref{fig:sdss_i} for the $i$-band.}
-    \label{fig:sdss_i}
-\end{figure*}
-
-\begin{table}
-        \begin{center}
-                \begin{tabular}{lcc}
-                        \toprule
-                         {\bf Before re-calibration} & Mean & $\sigma$  \\
-                        \midrule
-                        $r$ (IPHAS - APASS) & +0.014 & 0.104 \\
-                        $i$ (IPHAS - APASS) & +0.007 & 0.108 \\
-                        $r$ (IPHAS - SDSS) & +0.016 & 0.088 \\
-                        $i$ (IPHAS - SDSS) & +0.010 & 0.089 \\
-                        \bottomrule
-                \end{tabular}
-        \caption{Mean magnitude offsets for objects
-                 crossmatched between IPHAS and APASS/SDSS
-                 before the uniform re-calibration.
-                 Eqns.~\ref{eqn:apass_r}-\ref{eqn:sdss_i} were applied
-                 to the APASS/SDSS magnitudes to bring them into the
-                 Vega-based IPHAS system prior to computing
-                 the offsets.)
-                 }
-        \label{tbl:offsets_before}
-
-                \begin{tabular}{lcc}
-                        \toprule
-                         {\bf After re-calibration} & Mean & $\sigma$ \\
-                        \midrule
-                        $r$ (IPHAS - APASS) & +0.000 & 0.011\\
-                        $i$ (IPHAS - APASS) & +0.000 & 0.011 \\
-                        $r$ (IPHAS - SDSS)  & -0.001 & 0.029\\
-                        $i$ (IPHAS - SDSS) & -0.002 & 0.032 \\
-                        \bottomrule
-                \end{tabular}
-        \caption{Same as Table~\ref{tbl:offsets_before}
-                 but computed after the uniform re-calibration
-                 was carried out.}
-        \label{tbl:offsets_after}
-        \end{center}
-\end{table}
-
-\subsubsection{Adjusting fields not covered by APASS}
-
-At the time of writing, the APASS catalogue did not provide 
-sufficient coverage for 7359 of the fields in our data release.
-Fortunately, these fields are pre-dominantly located 
-in the early part of the Galactic Plane (Fig.~\ref{fig:apass_r}),
-which were typically observed during the summer months
-when photometric conditions are more prevalent at the telescope.
-These remaining fields have nevertheless 
-been brought onto the same uniform scale 
-by employing an ubercalibration-style scheme, explained below,
-which minimises the magnitude offsets between stars
-located in the overlap regions with neighbouring fields.
-Although we explained earlier that these overlap regions
-are prone to systematics, the use of APASS enabled us to
-keep the zeropoints of roughly half of the fields fixed,
-which avoids these systematics from combining to introduce
-artificial gradients across the survey.
-In the next section we will show this to be true by validating
-our calibration against SDSS.
-
-A general solution to the problem of minimising the magnitude
-differences between overlapping frames has previously been
-described by~\citet{Glazebrook1994}.
-In brief, there are two fundamental quantities to be
-minimised between each pair of overlapping exposures, 
-denoted by the indices $i$ and $j$. 
-Firstly, the mean magnitude difference between stars in the overlap
-region $\Delta_{ij}=\langle m_i-m_j\rangle=-\Delta_{ji}$ is a local
-constraint. 
-Secondly, to ensure the solution does not stray far 
-from the existing calibration, 
-the difference in zeropoints 
-$\Delta\mathrm{ZP}_{ij}=-\Delta\mathrm{ZP}_{ji}$ 
-between each pair of exposures must also be minimised.
-
-Minimisation of these two quantities is a linear least squares problem 
-because the magnitude $m$ depends linearly on the ZP (Eqn.~\ref{eqn:mag}).
-Hence we can find the ZP shift to be applied to each field 
-by minimising the sum:
-\begin{equation}
-   S = \sum_{i=1}^N \sum_{j=1}^N w_{ij} \theta_{ij} (\Delta_{ij} + a_i - a_j)^2
-   \label{eqn:chi2}
-\end{equation}
-where $i$ denotes the exposure of interest, 
-$j$ an overlapping exposure, 
-$N$ the number of exposures,
-$a_i$ the ZP to solve for 
-and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$), 
-$w_{ij}$ are weights set to the uncertainty in $\Delta_{ij}$
-and $\theta_{ij}$ is an overlap function 
-equal to either 1 if exposures $i$ and $j$ overlap or 0 otherwise. 
-Solving for $a_i$ is equivalent to solving $\partial
-S/\partial a_i=0$ which gives the matrix equation:
-\begin{equation}
-   \sum_{j=1}^N A_{ij} a_j = b_j
-   \label{eqn:matrix}
-\end{equation}
-where 
-\begin{eqnarray}
-   A_{ij} &=& \delta_{ij} \sum_{k=1}^N w_{jk}\theta_{jk} - w_{ij} \theta_{ij},\label{eqn:aij}\\
-   b_i &=& \sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ji} = -\sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ij}.\label{eqn:bi}
-\end{eqnarray}
-This prescription is essentially identical to \citet{Glazebrook1994}.
-%,although we changed the formulation to ensure that $A_{ij}$ is positive definite.
-
-As explained above, we enforce a strong external constraint
-on the solution by keeping the zeropoint fixed 
-for the 6756 fields which have been compared
-and calibrated against APASS earlier.
-We will hereafter refer to these fields as \emph{anchors}.
-It is asserted that the zeropoints $a_i$ of these anchor fields 
-are known and not solved for,
-though they do appear in the vector $b_j$ as constraints.
-In addition to the APASS-based anchors, 
-we selected 3273 additional anchor fields by hand
-to provide additional constraints in regions not covered by APASS.
-These extra anchors were deemed to have reliable zeropoints 
-based on 
-(i) the information contained in the observing logs,
-(ii) the stability of the standard star zeropoints during the night, and
-(iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
-which is located at $\sim$500~meter from the INT.
-
-We then solved Eqn.~\ref{eqn:matrix} for the $r$ and $i$ bands
-separately using the least-squares routine 
-in Python's {\sc scipy.sparse} module for sparse matrix algebra.
-This provided us with corrected zeropoints for the remaining fields,
-which were shifted on average by $+0.02\pm0.11$ in $r$ 
-and $+0.01\pm0.12$ in $i$ compared to their provisional calibration.
-
-We then turned to the uniform calibration of the \ha\ data.
-It is not possible to re-calibrate the narrowband 
-in the same way as the broadbands,
-because the APASS survey does not offer \ha\ photometry.
-We can reasonably assume however,
-that the corrections required for $r$ and \ha\ are identical
-because the \ha\ zeropoints have been derived directly from the
-$r$-band zeropoints during the provisional calibration (Eqn.~\ref{eqn:zpha}).
-Moreover, the IPHAS data-taking pattern ensured 
-that a field's \ha\ and $r$-band exposures
-were taken at essentially the same time, 
-separated only by the $\sim$30-second read-out time required by the WFC.
-We have hence corrected the \ha\ zeropoints 
-by re-using the zeropoint adjustements that were derived for the $r$ band
-in the earlier steps.
-An exception was made for 3101 fields
-for which our quality-control routines revealed
-zeropoint variations which exceeded 0.03~mag 
-between consecutive fields,
-which indicates non-photometric conditions.
-For good practice, the \ha\ zeropoints of these fields
-were adjusted by solving Eqn.~\ref{eqn:matrix}
-rather than linking them directly to the $r$-band shift.
-
-\subsection{Testing the calibration against SDSS}
-
-Having re-calibrated all fields to an expected accuracy of 3\%,
-we then used an independent survey to validate the results.
-We have been able to exploit SDSS Data Release 9 \citep{Ahn2012}
-for this purpose.
-SDSS DR9 provides several strips at low
-Galactic latitudes,
-providing data across 18\% of the fields in our data release.
-We cross-matched IPHAS fields against 
-stars marked as reliable in the SDSS catalogue\footnote{
-We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
-to obtain photometry from the SDSS {\sc photoprimary} table 
-with criteria {\sc type = star}, {\sc clean = 1} and {\sc score $>$ 0.7}.}
-in the same way as we did for APASS earlier,
-with the exception of using fainter magnitude ranges of 
-$15<r_{\rm SDSS}<18.0$ and $14.5<i_{\rm SDSS}<17.5$.
-This provided us with a set of 1.2 million cross-matched stars.
-
-Colour transformations were obtained using a sigma-clipped linear least squares fit:
-\begin{eqnarray}
-r_{\rm IPHAS} = r_{\rm SDSS} - 0.093 - 0.044(r-i)_{\rm SDSS} \label{eqn:sdss_r} \\
-i_{\rm IPHAS} = i_{\rm SDSS} - 0.318 - 0.095(r-i)_{\rm SDSS}. \label{eqn:sdss_i}
-\end{eqnarray}
-The rms residuals of these transformations are 0.045 and 0.073, respectively.
-The equations are similar to the ones
-previously determined for APASS,
-although the colour terms are slightly larger;
-the throughput curve of the SDSS $i$-band filter 
-appears to be somewhat more sensitive at longer wavelengths
-compared to both the IPHAS and APASS filters.
-
-These global transformations were deemed adequate
-for the purpose of validating our uniform calibration in a statistical sense.
-Separate equations were derived towards different sightlines
-to investigate the effects of varying reddening regimes.
-The transformation coefficients were found 
-to show some variation towards lowly reddened areas,
-which have relatively few numbers
-of (intrinsically) red objects at $r-i > 1$ 
-which can skew the colour term.
-The vast majority of red objects in the global sample
-are those in highly reddened areas however,
-which agree well with the global transformations
-and dominate the statistical appraisal of our calibration.
-%The small number of intrinsically red objects in lowly reddened fields
-%have no impact on the statistical appraisal of our calibration.
-
-Having transformed SDSS magnitudes into the IPHAS system,
-we then computed the median magnitude offset for each IPHAS field
-which contained at least 30 objects with a counterpart
-in the SDSS catalogue.
-This was the case for 2602 fields.
-The median offsets for each of these fields
-are shown in Figs.~\ref{fig:sdss_r}-\ref{fig:sdss_i}.
-The mean offset and standard deviation found 
-was $-0.001\pm0.029$ for $r$
-and $-0.002\pm0.032$ for $i$ (Table~\ref{tbl:offsets_after}).
-In comparison, offsets computed in the identical way
-\emph{before} our re-calibration showed means
-of $+0.016\pm0.088$ and $+0.010\pm0.089$ (Table~\ref{tbl:offsets_after}).
-We conclude that our re-calibration procedure has
-been successful in improving the
-uniformity of the calibration by a factor three
-and has achieved our aim of bringing the
-accuracy down to the level of $\sigma=0.03$~mag.
-
-We warn that the SDSS comparison reveals a number of fields with offsets
-exceeding 0.05~mag (523 fields) or even 0.1~mag (18 fields).
-Such sporadic outliers are consistent with the tails of a Gaussian
-with mean $\sim0$ and $\sigma=0.03$.
-
-In future work, we hope to draw upon
-the PanSTARRS survey \citep{Schlafly2012}
-to further improve the accuracy of our calibration.
-At the time of preparing this work data from PanSTARRS
-had not been made public yet.
-
-\section{Source catalogue generation}
-\label{sec:catalogue}
-
-Having obtained a quality-checked 
-and re-calibrated data set, 
-we now turn to the problem
-of transforming the observations 
-into a user-friendly catalogue.
-The aim of this catalogue is to detail
-the best-available information for each unique source
-in a convenient format,
-including flags to warn about quality issues 
-such as source blending and saturation.
-Compiling the catalogue essentially required four steps:
-\begin{enumerate}
-\item the single-band detection tables 
-produced by the CASU pipeline 
-were augmented with new columns
-and warning flags;
-\item the detection tables were merged into multi-band field catalogues;
-\item the overlap regions of the field catalogues 
-were cross-matched to flag duplicate measurements 
-and identify the best detection 
-of each unique source; and
-\item these primary detections
-were compiled into the final source catalogue.
-\end{enumerate}
-Each of these four steps are now explained.
-
-\subsection{User-friendly columns and warning flags}
-
-As the first step, the detection tables 
-were enhanced by creating new columns.
-This is necessary because the tables 
-generated by the CASU pipeline 
-summarise the detections 
-in their original CCD units,
-e.g. source positions are given in pixel coordinates 
-and photometry in number counts.
-To transform these measurements into
-user-friendly fields,
-we have largely adopted the units and naming conventions
-which are in use at the 
-WFCAM Science Archive \citep[WSA;][]{Hambly2008}
-and the VISTA Science Archive \citep[VSA;][]{Cross2012}.
-These archives curate the near-infrared data from both
-the UKIDSS Galactic Plane Survey \citep[GPS;][]{Lucas2008}
-and the 
-VISTA Variables in the Via Lactea survey \cite[VVV;][]{Minniti2010}.
-Both these surveys provide high-resolution JHK photometry
-in the Galactic Plane.
-There is a significant degree of overlap
-between the footprints of UKIDSS/GPS and IPHAS,
-and hence by adopting a similar catalogue format
-we hope to encourage scientific applications
-which combine both data sets.
-
-A detailed description of each column in our source catalogue
-is given in Appendix~\ref{app:columns}.
-In the remainder of this section we highlight the main features.
-
-Firstly, we note that each source is uniquely identified by an
-IAU-style designation of the form ``IPHAS2\ JHHMMSS.ss+DDMMSS.s''
-(cf. column \emph{name} in Appendix~\ref{app:columns}),
-where ``IPHAS2'' refers to the present
-data release and the remainder of the string
-denotes the J2000 coordinates in sexagesimal format.
-For convenience, the coordinates
-are also included in decimal degrees
-(columns \emph{ra} and \emph{dec})
-and in the Galactic coordinate system
-(columns \emph{l} and \emph{b}).
-We have also included an internal object identifier string 
-of the form ``\#run-\#ccd-\#detection''
-(e.g. ``64738-3-6473''),
-which documents the INT exposure number (\#run),
-the CCD number (\#ccd),
-and the row number in the CASU detection table (\#detection)
--- see columns \emph{rDetectionID},
-\emph{iDetectionID}, \emph{haDetectionID}.
-
-Photometry is provided based on the 2.3-arcsec diameter circular aperture
-by default (columns \emph{r}, \emph{i}, \emph{ha}).
-The choice of this aperture size as the default 
-is based on a trade-off between concerns 
-about small number statistics and centroiding errors
-for small apertures on one hand,
-and diminishing signal-to-noise ratios and source confusion
-for large apertures on the other hand.
-The user is not restricted to this choice because
-the catalogue also provides magnitudes
-using three alternative aperture sizes:
-the peak pixel height 
-(columns \emph{rPeakMag}, \emph{iPeakMag}, \emph{haPeakMag}),
-a circular 1.2-arcsec-diameter aperture 
-(\emph{rAperMag1}, \emph{iAperMag1},
- \emph{haAperMag1}) and
-a 3.3-arcsec-diameter aperture 
-(\emph{rAperMag3}, \emph{iAperMag3},
- \emph{haAperMag3}).
-
-Each of these magnitude measurements have been
-corrected for the flux lost outside of their respective apertures,
-using a correction term which is inferred from the
-mean shape of the PSF measured locally in the CCD frame.
-In the case of a point source,
-the four alternative magnitudes are expected
-to be consistent with each other
-within the photon noise uncertainties
-(which are given in columns \emph{rErr}, \emph{rPeakMagErr},
-\emph{rAperMag1Err}, \emph{rAperMag3Err}, etc).
-When this is not the case,
-it is likely that the source is either
-an extended object or that it has
-been incorrectly measured as a result of
-source blending or a rapidly spatially-varying nebulous background.
-In \S\ref{sec:qualitycriteria} we will explain that the consistency
-of the magnitude measurements in the different apertures
-can be used as a criterion for selecting stellar objects
-with reliable photometry from the catalogue.
-
-The brightness of each object as a function of increasing
-aperture size is also used by the CASU pipeline to provide
-a discrete star/galaxy/noise classification flag
-(\emph{rClass}, \emph{iClass}, \emph{haClass})
-and a continuous stellarness-of-profile statistic
-(\emph{rClassStat}, \emph{iClassStat}, \emph{haClassStat}).
-For convenience, we have combined
-these single-band morphological measures
-into band-merged class probabilities and flags
-(\emph{pStar}, \emph{pGalaxy}, \emph{pNoise},
-\emph{mergedClass}, \emph{mergedClassStat})
-using the merging scheme in use at the WSA\footnote{Explained at
-http://surveys.roe.ac.uk/wsa/www/gloss\_m.html \#gpssource\_mergedclass
-}.
-
-Information on the quality of each detection is included 
-in a series of additional columns.
-We draw attention to three useful flags
-which warn about the likely presence of systematics:
-\begin{enumerate}
-\item The \emph{saturated} column is used to flag sources
-for which the peak pixel height exceeds 55000 counts,
-which is typically the case for stars brighter than 12-13th magnitude.
-Although the pipeline attempts to extrapolate the brightness of
-saturated stars based on the shape of their PSF,
-such extrapolation is prone to errors
-and we do not recommend their use.
-\item The \emph{deblend} column is used to flag sources 
-which partially overlap with a nearby neighbour.
-Although the pipeline applies a deblending procedure
-to such objects, the procedure is currently applied separately
-in each band and hence the $r$-$i$ and $r$-\ha\ colours
-of such objects are prone to errors.
-\item The \emph{brightNeighb} column is used to flag sources which are located
-within 5 arcmin from an object brighter than $V=7$ 
-according to the Bright Star Catalogue (BSC; Hoffleit et al. 1991), 
-or within 10 arcmin if the neighbour is brighter than $V=4$.
-Such very bright stars are known to cause systematic errors
-and spurious detections as a result of stray light 
-and diffraction spikes.
-\end{enumerate}
-In addition to the above, we also created warning flags for internal bookkeeping.
-For example, we flagged detections which fell in the strongly vignetted regions of the focal plane,
-which were truncated by CCD edges,
-or which were otherwise affected by bad pixels in the detector.
-We will explain below that none of such detections 
-have been included in the catalogue
--- an alternative detection was available in essentially all these situations
-because of the IPHAS field pair strategy --
-and hence these internal warning flags do not appear
-in the final source catalogue.
-
-Finally, we note that basic information on the observing conditions
-is included (\emph{fieldID}, \emph{fieldGrade}, \emph{night}, \emph{seeing}).
-A table containing more detailed quality control information,
-indexed by \emph{fieldID}, is made available on our website.
-
-\subsection{Band-merging the detection tables}
-
-The second step in compiling the source catalogue
-is to merge the contemporaneous trios
-of $r$, $i$, \ha\ detection tables
-into multi-band field catalogues.
-This required a positional matching procedure 
-to link sources between the three bands
-based on their position on the sky.
-We used the \textsc{tmatchn} function 
-of the \textsc{stilts} software for this purpose,
-which allows rows from multiple tables to be matched \citep{Taylor2006}.
-The result of the procedure is a band-merged catalogue
-in which each row corresponds to a group of linked $r$/$i$/\ha\ detections
-which satisfy a maximum matching distance criterion in a pair-wise sense.
-Sources for which no counterpart was identified
-are retained in the catalogue as single-band detections.
-
-We employed a maximum matching distance of 1~arcsec,
-which was chosen based on a trade-off between 
-completeness and reliability.
-On one hand, a matching distance larger than 1~arcsec 
-was found to allow too many spurious and unrelated sources 
-to be linked. 
-On the other hand, a value smaller than 1~arcsec 
-would pose problems for very faint sources 
-with large centroiding errors, 
-and would occasionally fail to link detections near CCD corners
-where the astrometric solution can 
-show systematic errors which exceed 0.5~arcsec.
-The position offsets between the $r$ and $i$/\ha\ detections
-have been included in the catalogue 
-and can hence be tightened by the user if necessary
-(columns \emph{iXi}, \emph{iEta}, \emph{haXi}, \emph{haEta}).
-We note that also UKIDSS/GPS adopted 
-a maximum matching distance of 1 arcsecond 
-for similar reasons \citep{Hambly2008}.
-
-The resulting band-merged catalogues were found
-to be reliable for the vast majority of fields.
-We warn that source blending and confusion is unavoidable
-for faint objects in the Galactic Plane however;
-in \S\ref{sec:discussion} we will show
-that 19\% of the sources in our catalogue
-are flagged as blended objects (column \emph{deblend})
-and their band-merged data should be treated with care
-because they may have fallen victim to source confusion
-during the band-merging step.
-
-\subsection{Selecting the primary detections}
-
-We explained earlier that the survey contains repeat observations
-of identical sources as a result of overlaps in the data-taking pattern.
-Amongst all sources in the reliable magnitude range $13<r<19$,
-we find that 65\% were detected twice and 25\% were detected three times or more.
-Only 9\% were detected once.
-Unsurprisingly, their spatial distribution traces
-the inter-CCD gaps and footprint edges.
-
-The principal aim of this data release is to provide 
-reliable photometry at a single epoch,
-and hence we have decided
-to focus on providing the magnitudes
-and coordinates using only the \emph{best-available} 
-detection of each object -- 
-hereafter referred to as the \emph{primary} detection.
-Although overlapping fields could have been co-added 
-to gain a small improvement in depth, 
-we have decided against this for two reasons.
-Firstly, combining the information from multiple epochs
-would make the photometry of variable stars difficult to interpret.
-Secondly, co-adding would cause the image quality to degrade towards the mean,
-which is a draw-back for crowded fields.
-
-Anyone interested in the alternative detections of a source
--- hereafter called the \emph{secondary} detections --
-can nevertheless obtain this information in two ways.
-To begin with, whenever a secondary detection was observed 
-within 10 minutes of the primary,
-the magnitudes of that secondary detection
-have been included in the catalogue
-(columns \emph{sourceID2}, \emph{fieldID2}, 
-\emph{r2}, \emph{i2}, \emph{ha2},
-\emph{rErr2}, \emph{iErr2}, \emph{haErr2}, \emph{errBits2}.
-This is the case for 66\% of the sources brighter than $r<20$
-due to the IPHAS field pair observing pattern.
-In addition, the complete set of detection tables -- one for each exposure -- 
-is made accessibly on our website to allow other uses of the data.
-A user-friendly catalogue of secondary detections 
-has not been compiled at present 
-but may be part of a future data release.
-
-The primary detection is defined as the entry in the 
-set of band-merged field catalogues which provides 
-the most reliable information for a unique source.
-Primary detections have been selected using a so-called \emph{seaming} procedure
-which has been adapted from the algorithm developed for the WSA\footnote{http://surveys.roe.ac.uk/wsa/dboverview.html\#merge}.
-In brief, the first step of the procedure is to identify all the duplicate detections
-by cross-matching the overlap regions of all field catalogues,
-again using a maximum matching distance of 1\arcsec.
-The duplicate detections for each unique source are then ranked according to
-to (i) filter coverage, (ii) quality score (column \emph{errBits}),
-and (iii) the average seeing of stars in the CCD frame rounded to 0.2~arcsec.
-If this ranking scheme reveals multiple `winners' of identical quality,
-then the one that was observed closest to the optical axis of the camera is chosen.
-
-\subsection{Compiling the final source catalogue}
-
-As the final step, the primary detections that have been
-selected above are compiled
-into the 98-column source catalogue
-that is described in Appendix~\ref{app:columns}
-and made available on line.
-The entire list of sources naturally includes 
-a significant number of spurious entries
-as a result of the very sensitive detection levels
-that are employed by the CASU pipeline by default.
-To limit the size of the source catalogue,
-we have decided to enforce three basic criteria
-which must be met for a candidate source
-to be included in the catalogue:
-\begin{enumerate}
-\item the source must have been detected at SNR$>5$ in at least
-one of the bands, i.e. it is required that at least one of
-\emph{rErr}, \emph{iErr} or \emph{haErr} is smaller
-than 0.2 mag;
-\item the shape of the source must not be an obvious
-cosmic ray or noise artefact, i.e. we require
-either \emph{pStar} or \emph{pGalaxy} to be
-greater than 20\%;
-\item the source must not have been detected in one of the strongly
-vignetted corners of the detector, 
-not have had any known bad pixels in the aperture,
-and not have been on the edge of one of the CCDs,
-i.e. we require the \emph{errBits} quality score
-to be smaller than 64.
-\end{enumerate}
-
-A total of 219 million primary detections satisfied
-the above criteria and have been included in the catalogue.
-Amongst these, 158 million objects 
-are detected in all three bands (72\%),
-30 million are detected in two bands (14\%),
-and 31 million entries are single-band detections (14\%).
-Roughly half of the single-band detections were made in the $i$-band.
-This is likely explained by the fact that the $i$-band is least
-affected by interstellar extinction and can occasionally pick up
-highly-reddened objects which are otherwise lost in $r$/\ha.
-
-\section{Discussion}
-\label{sec:discussion}
-
-Having explained how the catalogue was generated,
-we now offer an overview of its properties
-by discussing  
-(i) the recommended quality criteria,
-(ii) the typical photometric uncertainties,
-and (iii) the source densities and the frequency of source blending.
-
-\subsection{Recommended quality criteria}
-\label{sec:qualitycriteria}
-
-Like any other photometric survey,
-the majority of the objects in our catalogue
-are faint sources observed near the detection limits:
-55\% of the entries in the catalogue
-are fainter than $r > 20$
-and 18\% are even fainter than $r > 21$.
-The measurements of such faint objects
-are naturally prone to large
-random and systematic uncertainties,
-for example, an inaccurately subtracted background
-will introduce a proportionally larger systematic error
-in a faint object.
-Most scientific applications will require a set of
-quality criteria to be enforced for the purpose
-of removing faint and low-quality objects.
-
-The choice of quality criteria is often a complicated
-trade-off between completeness on one hand
-and accuracy on the other.
-To aid users we have listed two sets of
-recommended quality criteria 
-in Tables~\ref{tab:reliable} and \ref{tab:veryreliable}.
-
-Firstly, Table~\ref{tab:reliable} details
-a set of minimum quality criteria
-which should benefit most applications
-which require reliable $r-i$ and $r-$\ha\ colours
-without removing more than $\sim$80\%
-of the sources brighter than $r < 19$.
-The listed criteria are designed to 
-(i) remove low-SNR sources, 
-(ii) remove saturated sources,
-and (iii) remove objects for which the 2.3~arcsec diameter
-aperture magnitudes are inconsistent 
-with the alternative 1.2~arcsec diameter aperture measurements.
-The last criterion is a proxy
-for identifying objects which are affected
-by inaccurate background subtraction
-or failed source deblending.
-A total of 86 out of 219 million sources 
-(39\%) satisfy all the criteria listed in Table~\ref{tab:reliable}
-and are hereafter referred to as ``reliable''.
-For convenience, the catalogue contains a boolean column
-named \emph{reliable} which flags these objects
-and makes their selection easy.
-
-\begin{table*}
-\vspace{2cm}
-
-\begin{tabular}{p{8cm}lp{6cm}}
-  \hline
-  Quality criterion & Rows passed & Description \\
-  \hline
-  rErr\,$<$\,0.1 {\sc and} iErr\,$<$\,0.1 {\sc and} haErr\,$<$\,0.1 &
-  109 million (50\%) &
-  Require the photon noise to be less than 
-  0.1 mag in all bands (i.e. SNR$>$10).
-  This implicitly requires a detection in all three bands.  \\
-  $r>13$ {\sc and} $i>12$ {\sc and} \ha\,$>12.5$ {\sc and not} \emph{saturated} &
-  158 million (72\%) &
-  The brightness must not exceed the nominal saturation limit
-  and the peak pixel height must not exceed 55\,000 counts.
-  Again, this implicitly requires a detection in all three bands.
-  \\
-  $|\mathrm{r}- \mathrm{rAperMag1}| 
-  < 3\sqrt{\mathrm{rErr}^2 + \mathrm{rAperMag1Err}^2} 
-  + 0.03$ &
-  176 million (80\%) &
-  Require the $r$ magnitude measured 
-  in the default 2.3\arcsec\ diameter aperture
-  to be consistent with the measurement 
-  made in the smaller 1.2\arcsec\ aperture,
-  albeit tolerating a 0.03 mag systematic error.
-  This will reject sources for which the background
-  subtraction or the deblending procedure
-  was not performed reliably. \\
-  $|\mathrm{i}- \mathrm{iAperMag1}| 
-  < 3\sqrt{\mathrm{iErr}^2 + \mathrm{iAperMag1Err}^2} 
-  + 0.03$ &
-  183 million (84\%) &
-  Same as above for $i$. \\
-  $|\mathrm{ha}- \mathrm{haAperMag1}| < 
-  3\sqrt{\mathrm{haErr}^2 + \mathrm{haAperMag1Err}^2} 
-  + 0.03$ &
-  158 million (72\%) &
-  Same as above for \ha. \\
-  \hline
-  All of the above (flagged as {\bf\emph{reliable}}) &
-  86 million (39\%) & \\
-  \hline
-\end{tabular}
-\caption{Recommended minimum quality criteria 
-for selecting objects with reliable colours 
-from the IPHAS DR2 source catalogue. 
-86~million entries in the catalogue (39\%)
-satisfy all the criteria listed in this table.
-For convenience, these have been flagged in the catalogue
-using the column named \emph{reliable}.}
-\label{tab:reliable}
-
-\vspace{2cm}
-
-\begin{tabular}{p{8cm}lp{6cm}}
-  \hline
-  Quality criterion & Rows passed & Description \\
-  \hline
-   reliable &
-   86 million (39\%) &
-   The object must satisfy the criteria listed in Table~\ref{tab:reliable}. \\
-
-   pStar $>$ 0.9 &
-   145 million (66\%) &
-   % pStar > 0.89 is identical to mergedClass == -1
-   The object must appears as a perfect point source,
-   as inferred from comparing its Point Spread Function (PSF)
-   with the average PSF measured in the same CCD. \\
-
-   {\sc not} \emph{deblend} &
-   177 million (81\%) &
-   The source must appear as a single, unconfused object. \\
-
-   {\sc not} \emph{brightNeighb} &
-   216 million (99\%) &
-   There is no star brighter than $V < 4$ within 10 arcmin, 
-   or brighter than $V < 7$ within 5 arcmin.
-   Such very bright stars cause scattered light and diffraction spikes,
-   which may add systematic errors to the photometry
-   or even trigger spurious detections. \\  
-  \hline
-
-  All of the above (flagged as {\bf\emph{veryReliable}}) &
-  59 million (27\%) & \\
-  \hline
-\end{tabular}
-\caption{Additional quality criteria which are recommended
-for applications which require very reliable colours
-at the expense of completeness. 
-For convenience, the sources which satisfy the criteria listed
-in this table have been flagged in the catalogue
-using the column named \emph{veryReliable}.}
-\label{tab:veryreliable}
-
-\vspace{2cm}
-\end{table*}
-
-For applications which require
-an even higher standard of reliability,
-a further set of additional quality criteria
-are listed in Table~\ref{tab:veryreliable}.
-These criteria are designed to ensure that
-(i) the object appeared as a perfect point source,
-(ii) the object was not blended with a nearby neighbour,
-and (ii) the object was not located near a very bright star.
-59 million sources (27\%) satisfy
-these additional criteria 
-and are hereafter referred to as ``very reliable''.
-Again, the catalogue contains a boolean column
-named \emph{veryReliable} which flags these objects.
-
-Figure~\ref{fig:magdist} compares the $r$-band magnitude
-distribution for objects with and without the 
-\emph{reliable} and \emph{veryReliable} criteria applied. 
-We find that 81\% of the sources 
-in the magnitude range $13 < r < 19$
-are considered \emph{reliable},
-which drops to 72\% in the range $19 < r < 20$
-and 9\% at $r>20$.
-In contrast, only 54\% of the sources 
-in the magnitude range $13 < r < 19$
-are considered \emph{veryReliable}.
-The stricter criteria filter out a lot
-of objects at early Galactic longitudes 
-where source blending is a common problem
-(we will demonstrate this in \S\ref{sec:densities}).
-The \emph{veryReliable} flag should hence
-only be used in applications which require very reliable photometry
-at the expense of completeness,
-which might be the case for e.g. spectroscopic target selection.
-
-\begin{figure}
-    \includegraphics[width=0.5\textwidth]{./plots/magdist/magdist-r.pdf} 
-    \caption{r-band magnitude distribution for all objects in the catalogue 
-    (light grey), for objects flagged as \emph{reliable} 
-    according to the criteria set out in Table~\ref{tab:reliable} (grey),
-    and for objects flagged as \emph{veryReliable} 
-    following Table~\ref{tab:veryreliable} (dark grey).
-    The magnitude distributions for $i$ and \ha\
-    look identical, apart from being shifted
-    by about 1 and 0.5 mag towards brighter magnitudes,
-    respectively.}
-    \label{fig:magdist}
-\end{figure}
-
-It is easy to see how the quality criteria
-may be adapted to be more tolerant.
-For example, by raising the allowed photometric uncertainties
-from 0.1 mag to 0.2 mag one can retrieve 42 million candidate sources.
-
-\subsection{Photometric uncertainties}
-
-Figure~\ref{fig:uncertainties} shows the mean photometric
-uncertainties as a function of magnitude for each band.
-We find that the uncertainties typically
-reach 0.1 mag near 20.5 in $r$ 
-and 19.5 in $i$/\ha when the default 2.3\arcsec\ aperture is used.
-At this point we note that the average colour
-of objects in the survey is
-$1.06\pm0.12$ for ($r$-$i$) and $0.44\pm0.03$ for ($r$-\ha).
-The better depth of $r$ is hence compensated
-by the fact that stars tend to have 
-brighter magnitudes in $i$ and \ha.
-
-\begin{figure}
-    \includegraphics[width=0.5\textwidth]{./plots/errors/uncertainties.pdf} 
-    \caption{Mean photometric uncertainties
-             for $r$ (top), $i$ (middle) and \ha\ (bottom).
-             Data points shown are the average values of
-             columns \emph{rErr}, \emph{iErr} and \emph{haErr}
-             in the catalogue, 
-             and the errorbars show the standard deviations.
-             The dashed and solid lines indicate 
-             the 10$\sigma$ and 5$\sigma$ limits, respectively.
-             These uncertainties are based only on the (Poissonian)
-             photon noise and hence this figure does not show
-             systematic or calibration uncertainties.}
-    \label{fig:uncertainties}
-\end{figure}
-
-The uncertainties shown in Fig.~\ref{fig:uncertainties}
-are the random errors based on the expected Poissonian photon noise.
-Systematics, such as calibration and deblending errors,
-are not included.
-To appraise the extent to which our photometry is affected
-by such systematics, we can exploit the
-secondary measurements which were made as part of the
-field-pair observing strategy and are available for 51\%
-of the sources in the catalogue.
-
-Figure~\ref{fig:pairmag} shows the mean residuals between
-the primary and secondary magnitudes
--- i.e. the average difference between catalogue columns \emph{r}-\emph{r2},
-\emph{i}-\emph{i2}, \emph{ha}-\emph{ha2} -- as a function of magnitude.
-We find that sources across the magnitude range 
-$13 < r < 17$ are consistent at the level of 5\%
-(i.e. $\sigma_{r-r2} \le 0.05$ mag),
-with the best repeatability
-reached at $r=14$ ($\sigma_{r-r2}~=~0.041$~mag).
-We draw attention to the fact that brighter stars
-tend to show significantly larger residuals
--- e.g. $\sigma_{r-r2}~=~0.14$~mag at $r=12$ --
-which is due to saturation effects.
-At the faint end we find residuals which
-show significantly more scatter than would be expected
-from photon noise alone, that is,
-the effects of source blending and background subtraction
-appear to dominate from $\sim$20th magnitude onwards.
-
-\begin{figure*}
-    \vspace{1cm}
-    \includegraphics[width=0.5\textwidth]{./plots/pairmag/pairmag.pdf} 
-    \caption{Photometric repeatability illustrated by plotting
-             the mean residuals between all the primary and secondary detections
-             in the catalogue as a function of magnitude.
-             The best photometric repeatability is reached at $r=14$
-             with $\sigma_{r-r2}~=~0.041$~mag.
-             Note that bright stars at $r<13$ and $i<12$ 
-             show increasing uncertainties due to saturation effects.}
-    \label{fig:pairmag}
-    \vspace{1cm}
-    \includegraphics[width=0.5\textwidth]{./plots/pairmag/pairmag-reliable.pdf} 
-    \caption{Same as Figure~\ref{fig:pairmag},
-    except that only the subset of sources
-    flagged as \emph{veryReliable} are now included.
-    We find that applying the quality criteria
-    has improved the photometric repeatability significantly.
-    The best repeatability is again reached at $r=14$
-    but has reduced to $\sigma=0.028$ mag.
-    The quality criteria have also been successful
-    at removing objects with large systematics at the bright
-    and faint ends.}
-    \label{fig:pairmag_reliable}
-\end{figure*}
-
-In Figure~\ref{fig:pairmag_reliable} we show 
-a similar comparison of the primary and secondary detections,
-but this time we have only include sources which are flagged
-as \emph{veryReliable} in the catalogue
-(i.e. not saturated, not confused, not near bright stars, etc.)
-We find that the average residuals are significantly better
-for this subset of the catalogue. Sources across the magnitude range 
-$13 < r < 17$ are consistent at the level of 0.03 mag,
-and the best repeatability is again reached at $r=14$
-with $\sigma_{r-r2}~=~0.028$~mag.
-We conclude that the \emph{veryReliable} quality criteria are effective
-in reducing the systematic errors to
-the same level as the accuracy of the
-global photometric calibration (cf. \S\ref{sec:calibration}).
-Moreover, the large systematics at the bright and faint end
-have disappeared.
-
-\subsection{Source densities and blending problems}
-\label{sec:densities}
-
-The mean source density as a function of Galactic longitude
-is shown in Figure~\ref{fig:density} (thick blue line).
-The densities were computed by counting the 
-number of sources in $1^\circ$-wide longitude bins
-across which covered the entire latitude range $-5^\circ<b<+5^\circ$.
-Unsurprisingly, we find the average source densities to increase
-towards the Galactic centre.
-For example, the average source density near $l\simeq 30^\circ$
-is roughly 30\,000 objects per square degree,
-which is five times more than the density
-found near $l\simeq 180^\circ$.
-
-In addition to the global trend, 
-there are significant variations in the source density on smaller scales.
-For example,  we find a significant drop near the constellations 
-of Cygnus ($l\simeq 80$) and Aquila ($l\simeq 40$),
-which are regions known to be affected
-by high levels of foreground extinction.
-Dark clouds are visible towards these constellation by eye,
-and they are often referred to as ``the Great Rift''.
-
-We warn however that the densities reported here 
-have not been corrected for survey completeness
-or differences in the observing conditions across the survey.
-For example, the dip in the density near $l\simeq140^\circ$
-is an artificial feature caused by gaps
-in the footprint coverage (which are apparent in Fig.~\ref{fig:footprint}).
-In a forthcoming paper,
-we aim to calibrate the IPHAS-based source densities
-by injecting artificial stars into the IPHAS images
-and measuring their recovery rate (Farnhill et al., in preparation).
-Indeed IPHAS has the potential to offer
-calibrated, two-dimensional stellar density maps
-which can be used to constrain detailed models of our Galaxy,
-but it is beyond the scope of the present work.
-
-\begin{figure*}
-    \includegraphics[width=\textwidth]{./plots/density/density.pdf} 
-    \caption{Mean number density of sources in the catalogue 
-    as a function of Galactic longitude, 
-    with and without blended sources included. 
-    The densities shown were computed by counting the sources 
-    at each longitude between $-5^\circ<b<+5^\circ$ (upper blue line).
-    We also show the densities based on only counting those sources 
-    for which the \emph{deblend} flag is {\sc false}, 
-    i.e. unconfused sources for which the CASU pipeline did not have to apply 
-    a deblending procedure (lower red line). 
-    }
-    \label{fig:density}
-\end{figure*}
-
-In Figure~\ref{fig:density} we also shows the density
-of non-blended sources (thin red line).
-These are sources for which the \emph{deblend} flag is {\sc false},
-i.e. sources for which the CASU pipeline did not have to apply 
-a deblending procedure to separate the flux
-originating from two or more overlapping objects.
-We find a strong correlation between the source density
-and the fraction of sources affected by source blending.
-For example, only NN\% of the sources are blended
-at $l>90^\circ$, whereas NN\% are blended at $l<90^\circ$.
-
-As we explained earlier, blended sources must be used with caution.
-Firstly, the deblending-procedure crucially depends
-on the local PSF being measured accurately. 
-Secondly, blended sources may are likely candidates to have fallen victim
-to source confusion during the band-merging procedure.
-In future work we hope to investigate the use of
-more advanced PSF-fitting routines
-in which sources are measured simultaneously across all bands,
-perhaps guided by an external list of sources provided
-by near-infrared surveys or Gaia data.
+perhaps guided by an external catalogue such as the one
+that is to be provided by Gaia.
 
@@
 \appendix
-\input{app_columns.tex}
+\onecolumn
+\section{Catalogue format}
+\label{app:columns}
+
+\small
+\begin{longtable}{lllp{10cm}}
+\caption{\label{tab:columns} 
+Definition of columns in the IPHAS DR2 source catalogue.
+} \\
+\hline
+Column & Type & Unit & Description \\
+\hline
+\endfirsthead
+
+\multicolumn{3}{c}%
+{{\bfseries \tablename\ \thetable{} -- continued}} \\
+\hline
+Column & Type & Unit & Description \\
+\hline
+\endhead
+
+\hline \hline
+\endlastfoot
+\input{tables/columns.tex}
+\end{longtable}
+\normalsize
+\twocolumn
 
@@
 \label{app:emitters}
-\begin{table*}
+\begin{table}
     \begin{tabular}{lccc}
@@
     \label{tbl:emitters}
-\end{table*}
+\end{table}
 

### 2014-03-12 paper/iphas-dr2.tex

@@
 Firstly, towards the Galactic anti-centre at $l=180^\circ$ (Fig.~\ref{fig:l180}) 
-we find a population dominated by lowly-reddened 
-main sequence stars and an apparent absence of reddened giants.
-%consistent with the sightline extinction of XX from Schlegel (cite). 
-In contrast, closer towards the galactic
-centre at $l=45^\circ$ (Fig.~\ref{fig:l45})
-we find a wealth of reddened objects
-which appear to be separated into a lowly
-and a highly reddened component,
-perhaps revealing two distinct parts of the Galaxy.
-Finally, in one of our earliest sightlines at $l=30^\circ$
+we find a population dominated by lowly-reddened main sequence stars.
+This is consistent with the estimated total sightline extinction of $E(B-V)=0.49$
+given by \cite{Schlegel1998}, following the correction of \cite{Schlafly2011}.
+Looking in more detail we can see that the stellar locus is narrower for M-type dwarfs than for earlier types;
+we do not observe M dwarfs experiencing the strongest reddening possible for this sightline.
+This implies that extinction is still increasing at distances beyond $\sim 2$~kpc,
+where M dwarfs are too faint to be contained in the IPHAS catalogue.
+It is also clear that there are no unreddened stars earlier than $\sim$K0 visible,
+such stars would be saturated if within a few hundred parsecs.
+This therefore suggests that there is a measurable increase in extinction locally.
+We also note an relative absence of late type giants which,
+due to the relative brevity of the corresponding phase of stellar evolution,
+would only account for a small proportion of a volume limited sample.
+
+In contrast, 
+closer towards the galactic centre at $l=45^{\circ}$ (Fig.~\ref{fig:l45}) we find a wealth of reddened objects.
+In the colour-magnitude diagram it is clear that the stars are split into two distinct groups,
+with one significantly redder than the other.
+The bluer group is composed of main sequence stars,
+with the slope of this group in the colour-magnitude diagram attributable to the significantly increasing extinction.
+Meanwhile the redder group is principally composed of red giant stars.
+As these stars are intrinsically brighter,
+at some given apparent magnitude they will be substantially further away than their main sequence counterparts.
+Given that, along this sight-line,
+extinction continues to increase with distance,
+the red giants we observe will be subject to more reddening than a main sequence star exhibiting the same apparent magnitude.
+Therefore, the clear split between the two groups arises due to the combination of increasing extinction and the distinct absolute magnitude ranges of main sequence and giant branch stars.
+
+Finally, in one of our earliest sightlines at $l=30^{\circ}$
 we find a very high number of extremely reddened giants
@@
 of foreground dwarfs.
-
-The number density of stars in the colour-colour
-and colour-magnitude space
-can be compared against population synthesis models
-to create three-dimensional maps of the extinction
-across several kpc \citep{Sale2010,Sale2012}.
-Such an extinction map based on our catalogue
-is to be released in a separate paper
+In contrast to the sightline at $l=45^{\circ}$,
+there is no clear group of giant stars visible in the colour-magnitude diagram of Fig.~\ref{fig:l30},
+though the red clump stars are manifested as a slight over density which sits roughly 0.4~mags redder than the A0V reddening track.
+At $(l,b)=(45^{\circ}, 2^{\circ})$ the giant stars observed exhibit a relatively narrow range of reddenings as the lie beyond most or essentially all of the Galactic dust.
+At $(l,b)=(30^{\circ}, 0^{\circ})$ this is not the case:
+even at the substantial distances at which we can observe reddened giant stars extinction is continuing to rise and we still find ourselves well within the Galactic dust disc.
+It is also apparent that the $(r-i)$ width of both the M dwarfs and early A dwarfs is greater than that in Fig.~\ref{fig:l45}.
+This is indicative of a steeper rise in reddening,
+both within several hundred parsecs (M dwarfs)
+and within a few kpc (early A dwarfs).
+
+Here we have presented only a short qualitative description of the information obtainable from IPHAS colour-colour and colour-magnitude plots.
+A more rigorous quantitative analysis of the IPHAS catalogue
+can be undertaken to estimate both the stellar density distribution in the Milky Way \citep{Sale2010}
+and to create detailed three-dimensional maps of the extinction across several kpc \citep{Sale2009,Sale2012}.
+An extinction map based on our catalogue is to be released in a separate paper
 that accompanies this data release (Sale et al., in preparation).

### 2014-03-12 pipeline.py
revert to the default pipeline state

@@ -64,34 +64,34 @@ client[:].execute('reload(concatenating)', block=True)
 Pipeline starts here
 """
 # Create an index of all single-band catalogues
-#detections.save_metadata(cluster)  # produces 'metadata.fits'
+detections.save_metadata(cluster)  # produces 'metadata.fits'
 
 # Enforce zp(Halpha) = zp(r) - 3.14
-#detections.sanitise_zeropoints()  # produces 'zeropoints-precalibration.csv'
+detections.sanitise_zeropoints()  # produces 'zeropoints-precalibration.csv'
 
 # Convert the single-band catalogues from CASU into our own catalogue format
-#detections.convert_catalogues(cluster)  # produces 'detected/nnnnnnn_det.fits'
+detections.convert_catalogues(cluster)  # produces 'detected/nnnnnnn_det.fits'
 
 # Bandmerge all runs obtained at the same epoch and pointing
-#bandmerging.bandmerge(cluster)  # produces 'bandmerged/nnnn.fits'
+bandmerging.bandmerge(cluster)  # produces 'bandmerged/nnnn.fits'
 
 # Compute the magnitude offsets between all runs, which is a necessary
 # input to the re-calibration step. Executing this on too many cores has been 
 # found to result in # "[Errno 105] No buffer space available", 
 # so we run it on cluster_highmem defined earlier.
-#offsets.compute_offsets(cluster_highmem)  # produces 'offsets-{r|i|ha}.csv'
+offsets.compute_offsets(cluster_highmem)  # produces 'offsets-{r|i|ha}.csv'
 
 # Find the set of zeropoint shifts which minimize the offsets obtained above
-#calibration.calibrate()  # produces 'calibration/calibration-{r|i|ha}.csv'
+calibration.calibrate()  # produces 'calibration/calibration-{r|i|ha}.csv'
 
 # Apply the zeropoint shifts found above to the bandmerged catalogues
-#calibration.apply_calibration(cluster) # produces 'bandmerged-calibrated/nnnn.fits'
+calibration.apply_calibration(cluster) # produces 'bandmerged-calibrated/nnnn.fits'
 
 # Identify duplicate detections where multiple pointings overlap ('seams');
-#seaming.seam(cluster_highmem)  # produces 'seamed/nnnn.fits'
+seaming.seam(cluster_highmem)  # produces 'seamed/nnnn.fits'
 
 # Finally, concatenate the individual pointings into a single catalogue 
[…]

### 2014-03-13 paper/iphas-dr2.tex

@@
 \usepackage{booktabs}
-%\usepackage{aas_macros}
+\usepackage{aas_macros}
 
@@
 quality criteria to be enforced for the purpose
-of removing faint and low-quality objects.
+of removing such faint and low-quality objects.
 
@@
 
-\subsection{Photometric uncertainties}
+\subsection{Photometric uncertainties and repeatability}
 
 Figure~\ref{fig:uncertainties} shows the mean photometric
-uncertainties as a function of magnitude for each band.
-We find that the uncertainties typically
-reach 0.1 mag near 20.5 in $r$ 
+uncertainty as a function of magnitude for each band.
+We find that the uncertainty typically
+reaches 0.1 mag near 20.5 in $r$ 
 and 19.5 in $i$/\ha when the default 2.3\arcsec\ aperture is used.
 At this point we note that the average colour
-of objects in the survey is
+across all sources in the catalogue is
 $1.06\pm0.12$ for ($r$-$i$) and $0.44\pm0.03$ for ($r$-\ha).
@@
 by the fact that stars tend to have 
-brighter magnitudes in $i$ and \ha.
+brighter magnitudes in both $i$ and \ha.
 
@@
 \begin{figure*}
-    \vspace{1cm}
-    \includegraphics[width=0.5\textwidth]{./plots/pairmag/pairmag.pdf} 
-    \caption{Photometric repeatability illustrated by plotting
-             the mean residuals between all the primary and secondary detections
-             in the catalogue as a function of magnitude.
-             The best photometric repeatability is reached at $r=14$
+\begin{minipage}[b]{0.45\linewidth}
+\centering
+\includegraphics[width=\textwidth]{./plots/pairmag/pairmag.pdf}
+\label{fig:pairmag}
+\end{minipage}
+\hspace{0.5cm}
+\begin{minipage}[b]{0.45\linewidth}
+\centering
+\includegraphics[width=\textwidth]{./plots/pairmag/pairmag-reliable.pdf}
+\label{fig:pairmag_reliable}
+\end{minipage}
+\caption{Photometric repeatability as a function of magnitude
+         for both all sources in the catalogue (left panel)
+         and for the \emph{veryReliable} sources alone (right panel).
+         The values shown are the mean residuals
+         between the primary and the secondary detections,
+         while the errorbars show the standard deviations.
+         Left panel: 
+         the best photometric repeatability is reached at $r=14$
          with $\sigma_{r-r2}~=~0.041$~mag.
-             Note that bright stars at $r<13$ and $i<12$ 
-             show increasing uncertainties due to saturation effects.}
-    \label{fig:pairmag}
-    \vspace{1cm}
-    \includegraphics[width=0.5\textwidth]{./plots/pairmag/pairmag-reliable.pdf} 
-    \caption{Same as Figure~\ref{fig:pairmag},
-    except that only the subset of sources
-    flagged as \emph{veryReliable} are now included.
-    We find that applying the quality criteria
+         Bright stars at $r<13$ and $i<12$ 
+         show increasing uncertainties due to saturation effects.
+         Right panel: we find that applying the quality criteria
          has improved the photometric repeatability significantly.
@@
          at removing objects with large systematics at the bright
-    and faint ends.}
-    \label{fig:pairmag_reliable}
+         and faint ends.
+    }
 \end{figure*}
@@
 
-\subsection{Source densities and blending problems}
+\subsection{Source densities and blending}
 \label{sec:densities}
@@
 runs in a direction that is different from the reddening vector,
-because the $r$-\ha\ colour tends to act as a coarse proxy for spectral type
+because the $r$-\ha\ colour tends to act
+as a coarse proxy for spectral type
 and is less sensitive to reddening than $r$-$i$.
@@
 
-This is demonstrated in Figures \ref{fig:l180}, \ref{fig:l45} \& \ref{fig:l30},
+This is demonstrated 
+in Figures \ref{fig:l180}, \ref{fig:l45} \& \ref{fig:l30},
 where we present three sets of IPHAS colour/magnitude diagrams
 towards three distinct sightlines
-located at Galactic longitudes $180^\circ$, $45^\circ$ and $30^\circ$,
-respectively.
-Each figure contains all the sources which are flagged as \emph{veryReliable}
+located at Galactic longitudes 
+$180^\circ$, $45^\circ$ and $30^\circ$, respectively.
+Each figure contains all the sources
+which are flagged as \emph{veryReliable}
 and are located in a region of one square degree 
@@
 For clarity, we have imposed the additional criterion
-for photometric uncertainties to be smaller than 0.05 mag in each band
-(corresponding to a magnitude limit near $\sim$19th magnitude, effectively).
+that the photometric uncertainties
+must be smaller than 0.05 mag in each band
+(corresponding to a magnitude limit near $\sim$19th magnitude).
 
@@
 We have annotated the colour-colour diagrams
-by showing the position of the unreddened main sequence (thin solid line),
-the unreddened giant branch (thick solid line)
+by showing the position 
+of the unreddened main sequence (thin solid line),
+the unreddened giant branch (thick solid line),
 and the reddening track for an A0V-type star (dashed line)
--- all three are based on the \cite{Pickles1998} library of empirical spectra
-tabulated for IPHAS by \cite{Drew2005}.
+-- all three are based on the \cite{Pickles1998} library 
+of empirical spectra
+transformed into the Vega-based IPHAS system by \cite{Drew2005}.
 In the colour-magnitude diagrams we only show the reddening vector
 together with the unreddened 1~Gyr isochrone due to \cite{Bressan2012},
-which are made available for the IPHAS filter system through a
-popular on line tool hosted by the Observatory of Padova (http://stev.oapd.inaf.it/cmd).
-The isochrone and reddening vector has been placed
+which is made available for the IPHAS system through a
+popular on line tool hosted at the Observatory of Padova
+(http://stev.oapd.inaf.it/cmd).
+The isochrone and reddening vector have been placed
 at an arbitrary distance of 2~kpc.
 
-Each of the sightlines reveals a stellar population
+Each of the sightlines reveal a stellar population
 with distinct characteristics.
-Firstly, towards the Galactic anti-centre at $l=180^\circ$ (Fig.~\ref{fig:l180}) 
+Firstly, towards the Galactic anti-centre 
+at $l=180^\circ$ (Fig.~\ref{fig:l180})
 we find a population dominated by lowly-reddened main sequence stars.
-This is consistent with the estimated total sightline extinction of $E(B-V)=0.49$
-given by \cite{Schlegel1998}, following the correction of \cite{Schlafly2011}.
-Looking in more detail we can see that the stellar locus is narrower for M-type dwarfs than for earlier types;
-we do not observe M dwarfs experiencing the strongest reddening possible for this sightline.
-This implies that extinction is still increasing at distances beyond $\sim 2$~kpc,
+This is consistent with the estimated total sightline extinction 
+of $E(B-V)=0.49$ given by \cite{Schlegel1998},
+following the correction of \cite{Schlafly2011}.
+Looking in more detail we can see
+that the stellar locus is narrower for M-type dwarfs
+than for earlier types;
+we do not observe M dwarfs experiencing 
+the strongest reddening possible for this sightline.
+This implies that extinction is still increasing 
+at distances beyond $\sim$1-2~kpc,
 where M dwarfs are too faint to be contained in the IPHAS catalogue.
-It is also clear that there are no unreddened stars earlier than $\sim$K0 visible,
+It is also clear that there are no unreddened stars
+earlier than $\sim$K0 visible,
 such stars would be saturated if within a few hundred parsecs.
-This therefore suggests that there is a measurable increase in extinction locally.
-We also note an relative absence of late type giants which,
-due to the relative brevity of the corresponding phase of stellar evolution,
+This therefore suggests 
+that there is a measurable increase in extinction locally.
+We also note a relative absence of late type giants which,
+due to the relative brevity 
+of the corresponding phase of stellar evolution,
 would only account for a small proportion of a volume limited sample.
@@
 In contrast, 
-closer towards the galactic centre at $l=45^{\circ}$ (Fig.~\ref{fig:l45}) we find a wealth of reddened objects.
-In the colour-magnitude diagram it is clear that the stars are split into two distinct groups,
+closer towards the galactic centre
+at $l=45^{\circ}$ (Fig.~\ref{fig:l45})
+we find a wealth of reddened objects.
+In the colour-magnitude diagram it is clear
+that the stars are split into two distinct groups,
 with one significantly redder than the other.
 The bluer group is composed of main sequence stars,
-with the slope of this group in the colour-magnitude diagram attributable to the significantly increasing extinction.
+with the slope of this group in the colour-magnitude diagram
+attributable to the significantly increasing extinction.
 Meanwhile the redder group is principally composed of red giant stars.
 As these stars are intrinsically brighter,
-at some given apparent magnitude they will be substantially further away than their main sequence counterparts.
+at some given apparent magnitude they will be substantially further away
+than their main sequence counterparts.
 Given that, along this sight-line,
 extinction continues to increase with distance,
-the red giants we observe will be subject to more reddening than a main sequence star exhibiting the same apparent magnitude.
-Therefore, the clear split between the two groups arises due to the combination of increasing extinction and the distinct absolute magnitude ranges of main sequence and giant branch stars.
+the red giants we observe will be subject to more reddening
+than a main sequence star exhibiting the same apparent magnitude.
+Therefore, the clear split between the two groups arises
+due to the combination of increasing extinction
+and the distinct absolute magnitude ranges
+of main sequence and giant branch stars.
 
@@
 In contrast to the sightline at $l=45^{\circ}$,
-there is no clear group of giant stars visible in the colour-magnitude diagram of Fig.~\ref{fig:l30},
-though the red clump stars are manifested as a slight over density which sits roughly 0.4~mags redder than the A0V reddening track.
-At $(l,b)=(45^{\circ}, 2^{\circ})$ the giant stars observed exhibit a relatively narrow range of reddenings as the lie beyond most or essentially all of the Galactic dust.
+there is no clear group of giant stars visible
+in the colour-magnitude diagram of Fig.~\ref{fig:l30},
+though the red clump stars are manifested as a slight over density
+which sits roughly 0.4~mags redder than the A0V reddening track.
+At $(l,b)=(45^{\circ}, 2^{\circ})$ the giant stars observed
+exhibit a relatively narrow range of reddenings
+as they lie beyond most or essentially all of the Galactic dust.
 At $(l,b)=(30^{\circ}, 0^{\circ})$ this is not the case:
-even at the substantial distances at which we can observe reddened giant stars extinction is continuing to rise and we still find ourselves well within the Galactic dust disc.
-It is also apparent that the $(r-i)$ width of both the M dwarfs and early A dwarfs is greater than that in Fig.~\ref{fig:l45}.
+even at the substantial distances
+at which we can observe reddened giant stars,
+extinction is continuing to rise
+and we still find ourselves well within the Galactic dust disc.
+It is also apparent that the $(r-i)$ width of both the M dwarfs
+and early A dwarfs is greater than that in Fig.~\ref{fig:l45}.
 This is indicative of a steeper rise in reddening,
@@
 
-Here we have presented only a short qualitative description of the information obtainable from IPHAS colour-colour and colour-magnitude plots.
+Here we have presented only a short qualitative description
+of the information obtainable
+from IPHAS colour-colour and colour-magnitude plots.
 A more rigorous quantitative analysis of the IPHAS catalogue
-can be undertaken to estimate both the stellar density distribution in the Milky Way \citep{Sale2010}
-and to create detailed three-dimensional maps of the extinction across several kpc \citep{Sale2009,Sale2012}.
-An extinction map based on our catalogue is to be released in a separate paper
+can be undertaken to estimate both
+the stellar density distribution in the Milky Way \citep{Sale2010}
+and to create detailed three-dimensional maps
+of the extinction across several kpc \citep{Sale2009,Sale2012}.
+An extinction map based on our catalogue
+is to be released in a separate paper
 that accompanies this data release (Sale et al., in preparation).
@@
     \end{minipage}
-    \caption{Colour-colour and colour-magnitude diagram (left and right panel)
+    \caption{Colour-colour and colour-magnitude diagrams
+    (left and right panel)
     showing sources flagged as \emph{veryReliable}
@@
     at $(l,b)=(180^\circ,+3^\circ)$.
-    The colour-colour diagram shows the
-    position of the main sequence (thin solid line),
+    The diagrams are plotted as 2D-histograms
+    which show the density of objects
+    in bins of $\sim0.01\times0.01$~mag;
+    bins containing 1-10 objects are coloured red,
+    bins with 10-20 objects are orange,
+    and bins with 20-25 objects are yellow.
+    The left panel is annotated with
+    the position of the main sequence (thin solid line),
     giant stars (thick solid line)
-    and the reddening track for an A0V-type star (dashed line)
-    based on the \citet{Pickles1998} library of empirical spectra.
-    The colour-magnitude only show the reddening vector
+    and the reddening track for an A0V-type star (dashed line),
+    which are based on the \citet{Pickles1998} library
+    of observed spectra.
+    The right panel only shows the reddening vector
     along with the unreddened 1~Gyr isochrone due to \citet{Bressan2012},
-    which has been placed at an arbitrary distance of 2~kpc for reference.
+    which has been placed at an arbitrary distance of 2~kpc.
     This is one of the least reddened sightlines
-    in the survey %(Av XX, Schlegel et al. XXXX)
+    in the survey
     and hence the observed stellar population appears to be dominated 
-    by lowly reddened main sequence stars.}
+    by lowly reddened main sequence stars (see text).}
     \label{fig:l180}
@@
     \end{minipage}
-    \caption{Same as above for $(l,b)=(30^\circ,0^\circ)$.
-    This is one of the most reddened sightlines in the survey.
-    %(XXX, Schlegel et al XXXX).
+    \caption{Same as above for $(l,b)=(30^\circ,0^\circ)$,
+    showing one of the most reddened sightlines in the survey.
     }
@@
 and the Montage software maintained by NASA/IPAC.
-We also made use of the SIMBAD, Vizier and Aladin \citep{Aladin} tools operated at CDS, Strasbourg, France.
+We also made use of the SIMBAD, Vizier and Aladin \citep{Aladin}
+operated at CDS, Strasbourg, France.
 
@@
 Sloan Digitized Sky Survey (SDSS),
-funded by the Alfred P.
-Sloan Foundation, the Participating Institutions, the National Science Foundation, the U.S.
-Department of Energy, the National Aeronautics and Space Administration, the Japanese Monbukagakusho, the Max Planck Society, and the Higher Education Funding Council for England.
+funded by the Alfred P. Sloan Foundation,
+the Participating Institutions,
+the National Science Foundation,
+the U.S. Department of Energy,
+the National Aeronautics and Space Administration,
+the Japanese Monbukagakusho, the Max Planck Society,
+and the Higher Education Funding Council for England.
 The astrometric pipeline reduction made
@@
 \bibliographystyle{mn2e}
-\bibliography{dr2}
+\bibliography{iphas-dr2}
 

### 2014-03-13 paper/tables/columns.tex

name & string &  & Sexagesimal, equatorial position-based source name in the form: JHHMMSS.ss+DDMMSS.s. The full naming convention for IPHAS DR2 sources has the form "IPHAS2 JHHMMSS.ss+DDMMSS.s", where "J" indicates the position is J2000. The "IPHAS2" prefix is not included in the column. \\
ra & double & degrees & J2000 Right Ascension with respect to the 2MASS PSC reference frame (which is consistent with ICRS to within 0.1 arcsec). The coordinate given is obtained from the astrometric measurement in the r'-band exposure. If the source is undetected in r', then the i' or H$\alpha$-band coordinate is given. \\
dec & double & degrees & J2000 Declination. See comments above. \\
sourceID & string &  & Unique identification number of the detection. Identical to rDetectionID if the source was detected in the r-band; identical to iDetectionID or haDetectionID otherwise. \\
posErr & float & arcsec & Astrometric fit error (RMS). Be aware that the error might be significantly larger than the RMS near CCD edges. \\
l & double & degrees & Galactic longitude $\ell$ converted from ra/dec (IAU 1958 system). \\
b & double & degrees & Galactic latitude $b$ converted from ra/dec (IAU 1958 system). \\
mergedClass & short &  & Image classification flag based on all bands (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). Computed using the UKIDSS scheme. \\
mergedClassStat & float &  & Merged N(0,1) stellarness-of-profile statistic. Computed using the UKIDSS scheme. \\
pStar & float &  & Probability that the source is a star (value between 0 and 1). \\
pGalaxy & float &  & Probability that the source is a galaxy (value between 0 and 1). \\
pNoise & float &  & Probability that the source is noise (value between 0 and 1). \\
pSaturated & float &  & Probability that the source is saturated (value between 0 and 1). \\
rmi & float & mag & ($r$ - $i$) colour, formed by subtracting columns r and i. Included in the catalogue for convenience only. To obtain the uncertainty, take the root of the sum of the squares of columns rErr and iErr. \\
rmha & float & mag & ($r$ - H$\alpha$) colour, formed by subtracting columns r and ha. See comments above. \\
r & float & mag & Default r-band magnitude using a 2.3 arcsec diameter aperture. Calibrated in the Vega system. \\
rErr & float & mag & Uncertainty for r. Does not include systematic errors. \\
rPeakMag & float & mag & Alternative r-band magnitude derived from the peak pixel height (i.e. a 0.3x0.3 arcsec square aperture). Calibrated in the Vega system. \\
rPeakMagErr & float & mag & Uncertainty in rPeakMag. Does not include systematics. \\
rAperMag1 & float & mag & Alternative r-band magnitude using a 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
rAperMag1err & float & mag & Uncertainty in rAperMag1. Does not include systematics. \\
rAperMag3 & float & mag & Alternative r-band magnitude using a 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
rAperMag3err & float & mag & Uncertainty in rAperMag3. Does not include systematics. \\
rGauSig & float & pixels & RMS of axes of ellipse fit in r. \\
rEll & float &  & Ellipticity in the r-band. \\
rPA & float & degrees & Position angle in the r-band. \\
rClass & short &  & Discrete image classification flag (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). \\
rClassStat & float &  & N(0,1) stellarness-of-profile statistic. \\
rErrBits & short &  & Bitmask used to flag a bright neighbour (1), source blending (2) and saturation (8). \\
rMJD & double & days & Modified Julian Date at the start of the r-band exposure. \\
rSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
rDetectionID & string &  & Unique identifier of the r-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e. composed of the INT telescope run number, the CCD number and a sequential source detection number. \\
rX & float & pixels & Pixel coordinate of the source in the r-band exposure, in the coordinate system of the CCD. \\
rY & float & pixels & Pixel coordinate of the source in the r-band exposure, in the coordinate system of the CCD. \\
i & float & mag & Default i-band magnitude using a 2.3 arcsec diameter aperture. Calibrated in the Vega system. \\
iErr & float & mag & Uncertainty for i. Does not include systematic errors. \\
iPeakMag & float & mag & Alternative i-band magnitude derived from the peak pixel height (i.e. a 0.3x0.3 arcsec square aperture). Calibrated in the Vega system. \\
iPeakMagErr & float & mag & Uncertainty in iPeakMag. Does not include systematics. \\
iAperMag1 & float & mag & Alternative i-band magnitude using a 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
iAperMag1err & float & mag & Uncertainty in iAperMag1. Does not include systematics. \\
iAperMag3 & float & mag & Alternative i-band magnitude using a 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
iAperMag3err & float & mag & Uncertainty in iAperMag3. Does not include systematics. \\
iGauSig & float & pixels & RMS of axes of ellipse fit. \\
iEll & float &  & Ellipticity. \\
iPA & float & degrees & Position angle. \\
iClass & short &  & Discrete image classification flag (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). \\
iClassStat & float &  & N(0,1) stellarness-of-profile statistic. \\
iErrBits & short &  & Bitmask used to flag a bright neighbour (1), source blending (2) and saturation (8). \\
iMJD & double & days & Modified Julian Date at the start of the single-band exposure. \\
iSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
iDetectionID & string &  & Unique identifier of the r-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e. composed of the INT telescope run number, the CCD number and a sequential source detection number. \\
iX & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
iY & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
iXi & float & arcsec & Position offset of the i-band detection relative to the ra column. The original i-band coordinates can be obtained by computing (ra+iXi/3600, dec+iEta/3600). \\
iEta & float & arcsec & Position offset of the i-band detection relative to the dec column. See comments above. \\
ha & float & mag & Default H-alpha magnitude using a 2.3 arcsec aperture. Calibrated in the Vega system. \\
haErr & float & mag & Uncertainty for ha. Does not include systematic errors. \\
haPeakMag & float & mag & Alternative H-alpha magnitude derived from the peak pixel height (i.e. a 0.3x0.3 arcsec square aperture). Calibrated in the Vega system. \\
haPeakMagErr & float & mag & Uncertainty in haPeakMag. Does not include systematics. \\
haAperMag1 & float & mag & Alternative H-alpha magnitude using a 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
haAperMag1err & float & mag & Uncertainty in haAperMag1. Does not include systematics. \\
haAperMag3 & float & mag & Alternative H-alpha magnitude using a 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
haAperMag3err & float & mag & Uncertainty in haAperMag3. Does not include systematics. \\
haGauSig & float & pixels & RMS of axes of ellipse fit. \\
haEll & float &  & Ellipticity \\
haPA & float & degrees & Position angle. \\
haClass & short &  & Discrete image classification flag (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). \\
haClassStat & float &  & N(0,1) stellarness-of-profile statistic. \\
haErrBits & short &  & Bitmask used to flag a bright neighbour (1), source blending (2) and saturation (8). \\
haMJD & double & days & Modified Julian Date at the start of the single-band exposure. \\
haSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
haDetectionID & string &  & Unique identifier of the r-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e. composed of the INT telescope run number, the CCD number and a sequential source detection number. \\
haX & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
haY & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
haXi & float & arcsec & Position offset of the H-alpha detection relative to the ra column. The original Ha-band coordinates can be obtained by computing (ra+haXi/3600, dec+haEta/3600). \\
haEta & float & arcsec & Position offset of the H-alpha relative to the ra column. See comments above. \\
brightNeighb & boolean &  & True if a very bright star is nearby (i.e. errBits $\geq$ 1). This indicates that the source might be spurious, or the photometry unreliable. \\
deblend & boolean &  & True if the source was blended with a nearby neighbour due to crowding (i.e. errBits $\geq$ 2). Although a deblending procedure is applied when measuring the photometry, the result may be unreliable (e.g. colours should not be trusted). \\
saturated & boolean &  & True if the source is saturated (i.e. peak pixel $>$ 55000 counts) in one or more bands (i.e. errBits $\geq$ 8). The photometry of saturated stars is affected by systematic errors. \\
errBits & short &  & Maximum value of (rErrBits, iErrBits, haErrBits). \\
nBands & short &  & Number of bands in which the source is detected (equals 1, 2 or 3). \\
reliable & boolean &  & True if: errBits\,$\leq$\,2 \& nBands\,==\,3 \& r\,$>$\,13 \& i\,$>$\,12 \& ha\,$>$\,12.5 \& rErr\,$<$\,0.1 \& iErr\,$<$\,0.1 \& haErr\,$<$\,0.1 \& (abs(r-rAperMag1)\,$<$\,3*hypot(rErr,rAperMag1Err)+0.03) \& (abs(i-iAperMag1)\,$<$\,3*hypot(iErr,iAperMag1Err)+0.03) \& (abs(ha-haAperMag1)\,$<$\,3*hypot(haErr,haAperMag1Err)+0.03). \\
veryReliable & boolean &  & True if: reliable \& pStar\,$>$\,0.9 \& errBits\,=\,0. \\
fieldID & string &  & Human-readable IPHAS field number and observing run (e.g. 0001o\_aug2003). \\
fieldGrade & string &  & Internal quality control score of the field. One of A, B, C or D. \\
night & integer &  & Night of the observation (YYYYMMDD). Refers to the UT date at the start of the night. \\
seeing & float & arcsec & Maximum value of (rSeeing, iSeeing, haSeeing). \\
ccd & short &  & CCD-chip number on the Wide Field Camera (WFC) of the Isaac Newton Telescope (INT). 1, 2, 3 or 4. \\
nObs & short &  & Number of repeat observations of this source in the survey. \\
sourceID2 & string &  & SourceID of the object in the partner exposure (if obtained within 10 minutes of the primary detection). \\
fieldID2 & string &  & FieldID of the partner detection (e.g. 0001o\_aug2003). \\
r2 & float & mag & r-band magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
rErr2 & float & mag & Uncertainty for r2. \\
i2 & float & mag & i-band magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
iErr2 & float & mag & Uncertainty for i2. \\
ha2 & float & mag & H-alpha magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
haErr2 & float & mag & Uncertainty for ha2. \\
errBits2 & integer &  & Error bitmask for the partner detection. Used to flag a bright neighbour (1), source blending (2), saturation (8), vignetting (64), truncation (128) and bad pixels (32768). \\

### 2014-03-13 paper/iphas-dr2.tex
Improves Section 7

@@
 
-A primary motivation for carrying
-out the survey 
-was to enable the discovery of 
-new emission-line objects across the Galactic Plane.
+A primary motivation for carrying out the survey 
+was to enable the discovery 
+of new emission-line objects
+across the Galactic Plane.
 \ha\ in emission is a well-known tracer
@@
 post-main sequence stages of their evolution,
-and hence IPHAS aims to allow larger, deeper
-and more statistically robust samples of such rare objects
-to be established.
+and hence IPHAS allows larger, deeper
+and more statistically robust samples
+of such rare objects to be established.
 The survey-wide identification and analysis 
-of such stars is beyond the scope of the present work,
+of emission-line objects is beyond the scope
+of the present work,
 but in this section we demonstrate how the
@@
 \citep{Georgelin1973,Mayer1973,Hunter1990}.
-The ionising star has been placed at a likely distance of 1.5-1.7 kpc
-based on the photometric parallax
+The ionising star has been placed 
+at a likely distance of 1.5-1.7 kpc
+based on its photometric parallax
 \citep{Mayer1973,Lahulla1985,Hunter1990}.
@@
 and $i$ (blue channel) images.
-The ionising star can be seen as the bright object
-near the centre of the H{\sc ii} region,
+The ionising star can be seen as the brightest object
+in the centre of the H{\sc ii} region,
 which is surrounded by a faint reflection nebula
@@
 For comparison, Figure~\ref{fig:mosaic_spitzer} shows
-a mosaic of the identical region 
-as seen by the Spitzer Space Telescope
-in the mid-infrared. The Spitzer image
-reveals a bubble-shaped structure of warm dust
-which surrounds the entire H{\sc ii} region.
+a mosaic of the same region 
+as seen in the mid-infrared by the Spitzer Space Telescope
+The infrared image reveals a bubble-shaped structure
+of warm dust which surrounds the entire H{\sc ii} region,
+and is thought to contain a wealth of infrared-bright
+young stellar objects \citep{Yu2012}.
+Many of these young objects 
+appear in Fig.~\ref{fig:mosaic_spitzer}
+as red- and pink-coloured objects,
+located predominantly in the north-western part
+of the bubble-shaped structure.
 
@@
 Figure~\ref{fig:emitters} presents
-the IPHAS colour-colour diagram for the region covered by the
-mosaic images.
-Gray circles show all objects in the region
+the IPHAS colour-colour diagram for 
+the 20-by-15 arcmin region shown in the mosaics.
+Gray circles show all objects
 which are brighter than $r<20$
@@
 and the expected position of unredded main-sequence stars
-with \ha\ in emission at a strength of EW=$-10{\rm \AA}$ (dashed line),
+with \ha\ in emission
+at a strength of EW=$-10{\rm \AA}$ (dashed line),
 taken from the colour simulations due to \citet{Barentsen2011a}.
 Six stars are found to lie confidently above the 
-dashed line at the level of $3\sigma$ 
-(i.e. the distance between the dashed line is larger than
-three times the uncertainty on the $r$-\ha\ colour).
+dashed line at the level of $3\sigma$,
+that is, the distance between the objects and the dashed line
+is larger than three times the uncertainty on the $r$-\ha\ colour.
 These reliable candidate \ha-emitters
 are marked by red triangles in the colour-colour diagram
-and their photometry is detailed in Appendix~\ref{app:emitters}.
+and their photometry is detailed in Table~\ref{tbl:emitters}.
+The spatial distribution of the six candidate emission-line objects
+is marked by yellow triangles in the colour mosaic (Fig.~\ref{fig:mosaic_iphas}).
 
@@
 
-The spatial distribution of our six candidate emission-line objects
-is marked by yellow triangles in the colour mosaic (Fig.~\ref{fig:mosaic_iphas}).
-They are likely to be genuine young stars for two reasons.
-Firstly, two of our candidates have recently been identified as likely
-candidate Young Stellar Objects (YSO)
+\begin{table}
+\footnotesize
+\resizebox{0.45\textwidth}{!}{
+    \begin{tabular}{lccc}
+    \toprule
+    Name [IPHAS2 ...] & $r$ & $i$ & \ha  \\
+    \midrule
+J192954.40+181026.1& $17.69\pm0.01$ & $16.12\pm0.01$ & $16.19\pm0.01$ \\
+J193011.01+182051.2& $18.55\pm0.02$ & $16.95\pm0.02$ & $17.31\pm0.02$ \\
+J193021.52+181954.5& $19.72\pm0.05$ & $17.94\pm0.03$ & $18.47\pm0.04$ \\
+J193024.45+181938.3& $19.31\pm0.04$ & $17.55\pm0.02$ & $17.99\pm0.03$ \\
+J193033.00+181609.3& $18.25\pm0.01$ & $16.91\pm0.01$ & $16.92\pm0.01$ \\
+J193042.48+182317.4& $19.96\pm0.03$ & $18.11\pm0.03$ & $18.48\pm0.03$ \\
+    \bottomrule
+    \end{tabular}}
+    \caption{Candidate \ha-emitters towards Sh 2-82.}
+    \label{tbl:emitters}
+\end{table}
+
+In previous work, we have shown how the majority of
+\ha-emitters seen towards an H{\sc ii} region
+are likely to be Classical T Tauri Stars (CTTS),
+which are young objects which show \ha\ in emission
+due to the presence of hot, infalling gas
+which is accreting onto the star
+from a circumstellar disk \cite{Barentsen2011a}.
+This is likely to be the case in this demonstration
+as well. Two of our candidates have recently been identified
+as Young Stellar Objects (YSO)
 in an investigation of the region by \cite{Yu2012}.
@@
 to select likely young stars by selecting objects with
-circumstellar disks based on the infrared colour excess.
-Secondly, we find that the four remaining objects are 
-also detected in the Spitzer-based image,
-although their colours are less extreme
-than those identified as likely YSOs by \citeauthor{Yu2012}.
+an infrared colour excess, consistent with 
+the presence of a heated circumstellar disk.
+Although \citeauthor{Yu2012} did not identify
+four of our candidates as having
+an infrared excess, we note  that each of them
+is detected at S/N$>$? in the Spitzer 8.0\,\micron\ image,
+which is more likely for a circumstellar-bearing
+object than it is for a field star.
 
@@
 $V < 15$. In contrast, the IPHAS catalogue offers photometry
-for NN stars in the region down to $r<20$.
-This demonstrates the ability of IPHAS to providing complimentary
-for the wealth of poorly-studied star-forming regions
+for 2197 objects in the region down to $r<20$.
+
+Sh 2-82 is one of a wealth of poorly-studied star-forming regions
 at low Galactic latitudes,
-which have been unveiled in recent years
-by the observed wealth of star-forming ``bubbles''
-at mid-infrared wave%\cite{Bica2003} was the first to identify a
-%near-infrared cluster
-%associated with this region using near-infrared 2MASS data,
-%and only recently has this cluster been confirmed
-%in a short investigation which combined 2MASS
-%and Spitzer data \cite{Yu2012}.
-%Many of the young stars identified by \citeauthor{Yu2012}
-%can be seen as pink and red stars in Fig.~\ref{fig:mosaic_spitzer}.lengths \cite[e.g.][]{Churchwell2006,Simpson2012}.
-%In fact the cluster of young stars associated with Sh 2-82
-%was only first reported by \cite{Bica2003},
-%remained unstudied until the study by \cite{Yu2012}.
-
-%Amongst the six emitters, N are detected at SNR X
-%in the Spitzer X micron band.
-
-%This demonstrates how IPHAS is providing complimentary
-%data for some of the thousands of star-forming ``bubbles'' have been unveiled at mid-infrared wavelengths \cite[e.g.][]{Churchwell2006,Simpson2012}.
-%Optically-unveiled members of such star-forming regions
-%may help to constrain extinctions and distances towards such regions \citep{Barentsen2013}.
-%Moreover, in \cite{Barentsen2011a} we showed
-%how the history of star formation in a region,
-%and the possible triggered nature, may be discussed.
-%This is particularly true
+which have started to become revealed
+in recent years through catalogues
+of ``bubbles'' seen
+in mid-infrared surveys \cite[e.g.][]{Churchwell2006,Simpson2012}
+and previously unknown clusters seen 
+at near-infrared wavelengths \cite[e.g.][]{Bica2003}.
+The optical data provided by IPHAS offers a handle (or lower limit)
+on the extinction and distance towards many of these regions,
+and can provide information on the optically-unveiled
+members of the region, as we demonstrated here.
 
@@
 %TODO: mention Gaia, e.g. IPHAS for transient identification?
-
 %\subsection{Caveats and lessons learnt}
-%\cite{Bica2003} was the first to identify a
-%near-infrared cluster
-%associated with this region using near-infrared 2MASS data,
-%and only recently has this cluster been confirmed
-%in a short investigation which combined 2MASS
-%and Spitzer data \cite{Yu2012}.
-%Many of the young stars identified by \citeauthor{Yu2012}
-%can be seen as pink and red stars in Fig.~\ref{fig:mosaic_spitzer}.
+%\subsection{Comparison with Gaia}?
 
@@
 The observations included in this release
-were found to achieve a median seeing of 1.1 arcsec
+achieved a median seeing of 1.1 arcsec
 and a $5\sigma-$depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$, \ha$=20.3\pm 0.3$.
 The global calibration and photometric repeatability
-is accurate at the level of $\sigma=0.03$ mag,
+was found to be accurate at the level of $\sigma=0.03$ mag,
 providing a significant improvement over the 
@@
 
+In future work, we aim to draw upon the PanSTARRS photometric
+survey to further improve the accuracy of our global calibration.
+We will also aim to correct the IPHAS photometry for the
+radial field distortions,
+and aim to produce a catalogue
+which details all the secondary detections
+to aid time-domain studies.
+
 The data-taking strategy developed for IPHAS
 have since been reapplied to carry out a companion INT/WFC survey called UVEX
-in U/$g$/$r$ \citep{Groot2009},
-and also southern counterpart to IPHAS and UVEX 
-is being carried out in $u$/$g$/$r$/$i$/\ha\ 
-called VPHAS+ (Drew et al, in press).
-We hope to re-use the experience gained by this data release
-to create similar releases for these companion surveys.
-
-In future work, we aim to draw upon the PanSTARRS photometric
-survey to further improve the accuracy of our global calibration.
-We will also aim to correct the photometry for the
-radial field distortions.
-%We may also produce a catalogue which details all detections.
-
-%TODO: mention Gaia?
+in U/$g$/$r$/He {\sc i} \citep{Groot2009}
+and a southern companion survey in $u$/$g$/$r$/$i$/\ha\ 
+called VPHAS+ \citep{Drew2014}.
+We intend to use the present data release
+as a template for future releases from these companion surveys.
 
@@
 
-\section{Candidate emission-line objects towards Sh 2-82}
-\label{app:emitters}
-\begin{table}
-    \begin{tabular}{lccc}
-    \toprule
-    Name & $r$ & $i$ & \ha  \\
-    \midrule
-IPHAS2 J192954.40+181026.1& $17.69\pm0.01$ & $16.12\pm0.01$ & $16.19\pm0.01$ \\
-IPHAS2 J193011.01+182051.2& $18.55\pm0.02$ & $16.95\pm0.02$ & $17.31\pm0.02$ \\
-IPHAS2 J193021.52+181954.5& $19.72\pm0.05$ & $17.94\pm0.03$ & $18.47\pm0.04$ \\
-IPHAS2 J193024.45+181938.3& $19.31\pm0.04$ & $17.55\pm0.02$ & $17.99\pm0.03$ \\
-IPHAS2 J193033.00+181609.3& $18.25\pm0.01$ & $16.91\pm0.01$ & $16.92\pm0.01$ \\
-IPHAS2 J193042.48+182317.4& $19.96\pm0.03$ & $18.11\pm0.03$ & $18.48\pm0.03$ \\
-    \bottomrule
-    \end{tabular}
-    \caption{Candidate \ha-emitters towards Sh 2-82.}
-    \label{tbl:emitters}
-\end{table}
-
 \label{lastpage}

### 2014-03-13 dr2/detections.py
Improves documentation

@@ -1,32 +1,84 @@
 #!/usr/bin/env python
 # -*- coding: utf-8 -*-
-"""Generates user-friendly single-band IPHAS detection catalogues.
+"""Generates user-friendly single-band IPHAS DR2 detection catalogues.
 
-This script reads in catalogues produces by the 'imcore' image detection tool
-of the Cambridge Astronomical Survey Unit (CASU) pipeline, and transforms
-these catalogues into a more user-friendly format which is known in the
-data release as the 'iphasDetection' table.
-
-The transformation involves converting fluxes to magnitudes, computing
-celestial coordinates and adding flags to indicate quality problems.
-
-The script also checks the header for known errors and fixes the World
-Coordinate System (WCS) where necessary.
-
-This script can also produce a table called 'runs.csv' which
-provides a table of all available IPHAS exposures.
-
-Computing requirements: ~48h CPU (2012 hardware). Low on RAM.
-
-Possible improvements
----------------------
-- look up confidence value for each star in the confidence maps.
+Summary
+-------
+This module is used to transform the detection tables that are produced
+by the ``imcore`` source detection and measurement tool of the Cambridge
+Astronomical Survey Unit (CASU). Its main purpose is to convert the columns
+in the CASU detection tables from instrumental units into more user-friendly
+astronomical units (e.g. converting fluxes into magnitudes, converting pixel
+coordinates into celestial coordinates, and adding user-friendly warning flags
+to signal quality problems.) Whilst doing so, the script will corrects for a
+range of known issues present in the data, e.g. it will apply a custom fix
+to the World Coordinate System (WCS) where necessary.
+This module can also produce a table called 'runs.csv' which details the
+metadata of an entire directory of CASU detection tables, which is useful to
+carry out quality control.
 
 Usage
 -----
-detections.save_metadata()
-detections.sanitise_zeropoints()
-detections.save_detections()
+1. To create a table called 'metadata.fits' which tabulates the header
+   information for the entire set of IPHAS detection tables, run:
+
+    ``detections.save_metadata()``
+
+2. To prepare a list of zeropoints which enforces the fixed offset between
+   the r- and Halpha-band zeropoints:
+
+    ``detections.sanitise_zeropoints()``
+
[…]

### 2014-03-13 dr2/detections.py
Improves documentation

@@ -13,9 +13,9 @@ coordinates into celestial coordinates, and adding user-friendly warning flags
 to signal quality problems.) Whilst doing so, the script will corrects for a
 range of known issues present in the data, e.g. it will apply a custom fix
 to the World Coordinate System (WCS) where necessary.
-This module can also produce a table called 'runs.csv' which details the
-metadata of an entire directory of CASU detection tables, which is useful to
-carry out quality control.
+This module can also produce a table called 'metadata.fits' which details the
+metadata of an entire directory of CASU detection tables, which is useful for
+carrying out quality control.
 
 Usage
 -----
@@ -40,15 +40,15 @@ During the creation of IPHAS DR2, we encountered the following issues
 with the pipeline-processed data set from CASU:
 
  * In very crowded fields, the astrometric solution can be off by more than
-   0.5 arcsec towards the edges of a CCD. The problem is rare and has been
-   resolved ad-hoc using the code under scripts/tune-wcs (which details the
-   runs that have been fixed.) For a future data release, it might be worth
-   refining the astrometry on a systematic basis.
+   0.5 arcsec towards the edges of the CCDs. The problem is rare and has been
+   resolved ad-hoc using the code under ``scripts/tune-wcs``.
+   Prior to a future data release, it may be worth refining the astrometry
+   across the data set.
  * The WCS standards have evolved over the years, and the WCS in some of the
-   older data can still be inconsistent with modern conventions.
-   The ``fix_wcs()`` function in the ``dr2.detections`` must be used to fix
-   faulty WCS keywords in the data set.
- * Spurious sources frequently appear in the vignetted corners of CCDs 3 and 1,
+   older data is inconsistent with modern conventions. The function
+   ``fix_wcs()`` in this module must be used before converting pixel to world
+   coordinates.
+ * Spurious sources frequently appear in the vignetted corners of CCDs 1 and 3,
    and in general near CCD edges. It is very important to mask these out during
    catalogue generation.
  * Data from December 2003 are affected by unusual bad columns appearing in
@@ -56,7 +56,7 @@ with the pipeline-processed data set from CASU:
    confidence map. A significant number of spurious detections were found
    near these columns. Detections made near these columns in said month are
    currently flagged as containing bad pixels by the ``column_badPix()``
-   method in this module, which is a bit of a hack really.
+   method in this module, which is a bit of a hack.
  * The 'badpix' column is missing from the CASU pipeline-produced detection
[…]

### 2014-03-13 pipeline.py
Improves documentation

@@ -1,15 +1,25 @@
 #!/usr/bin/env python
 # -*- coding: utf-8 -*-
-"""Pipeline script to produce IPHAS Data Release 2 using an MPI cluster.
+"""Pipeline script used to produce IPHAS Data Release 2 on a computing cluster.
 
-This script drives the entire pipeline for converting the single-band 
-catalogues produced by the Cambridge Astronomical Surveys Unit (CASU) 
-into a band-merged and homogeneously calibrated source catalogue 
-of unique objects.
+Summary
+-------
+This script drives the entire creation of the IPHAS DR2 source catalogue.
+Its purpose is to convert the IPHAS single-band detection tables, which are
+kindly created by the Cambridge Astronomical Surveys Unit (CASU),
+into a band-merged and globally calibrated source catalogue which lists only
+the best-available measurement for each unique source.
 
-This script relies on the presence of a running computing cluster setup with 
-IPython.parallel (using the "ipcluster" tool); i.e. before running the pipeline,
-execute 'qsub scripts/pipeline-cluster.pbs'.
+Dependencies
+------------
+This script relies on the presence of a running instance of an IPython.parallel
+cluster, which can be started on a local machine using:
+    
+    ``ipcluster start``
+
+or can be started on a PBS-controlled computing cluster by submitting the job:
+
+   ``qsub scripts/pipeline-cluster.pbs``.
 """
 from IPython import parallel
 from astropy import log
@@ -46,10 +56,11 @@ with client[:].sync_imports():
     from dr2 import bandmerging
     from dr2 import seaming
     from dr2 import concatenating
-    #from dr2 import images
+    from dr2 import images
 
-# While in development, reload every module by default
-# This is to make sure that the latest version gets used
+# While in development, reload every module by default,
+# to make sure that the latest version gets used. If the pipeline were ever
+# to be used in production (never), this could be removed. 
 client[:].execute('reload(constants)', block=True)
 client[:].execute('reload(util)', block=True)
 client[:].execute('reload(detections)', block=True)
@@ -58,7 +69,7 @@ client[:].execute('reload(calibration)', block=True)
 client[:].execute('reload(bandmerging)', block=True)
 client[:].execute('reload(seaming)', block=True)
 client[:].execute('reload(concatenating)', block=True)
-#client[:].execute('reload(images)', block=True)
+client[:].execute('reload(images)', block=True)
 
 """
 Pipeline starts here

### 2014-03-13 paper/iphas-dr2.tex
Typos

@@
 E. A. Gonz$\acute{\rm{a}}$lez-Solares$^2$, \newauthor
+R. Greimel$^3$,
 M. J. Irwin$^2$,
-S. E. Sale$^3$,
-R. Greimel$^4$,
-P. Groot$^5$,
-J. Lewis$^2$,
-A. Mampaso$^6$, \newauthor
 B. Mizalski$^{99}$,
 C. Ruhland$^1$,
+P. Groot$^4$,
+A. Mampaso$^5$, \newauthor
+S. E. Sale$^6$,
 plus many contributors.
@@
 $^{2}$Institute of Astronomy, Cambridge University, Madingley Road, Cambridge, CB3 OHA, U.K.\\
-$^{3}$Rudolf Peierls Centre for Theoretical Physics, Keble Road, Oxford, OX1 3NP, U.K.\\
-$^{4}$IGAM, Institute of Physics, University of Graz, Universit\"atsplatz 5, Graz, Austria.\\
-$^{5}$Afdeling Sterrenkunde, Radboud Universiteit Nijmegen, Faculteit NWI, Postbus 9010, 6500 GL Nijmegen, The Netherlands.\\
-$^{6}$Instituto de Astrof\'isica de Canarias, 38200 La Laguna, Tenerife, Spain.\\
+$^{3}$IGAM, Institute of Physics, University of Graz, Universit\"atsplatz 5, Graz, Austria.\\
+$^{4}$Afdeling Sterrenkunde, Radboud Universiteit Nijmegen, Faculteit NWI, Postbus 9010, 6500 GL Nijmegen, The Netherlands.\\
+$^{5}$Instituto de Astrof\'isica de Canarias, 38200 La Laguna, Tenerife, Spain.\\
+$^{6}$Rudolf Peierls Centre for Theoretical Physics, Keble Road, Oxford, OX1 3NP, U.K.\\
 $^{99}$Please add your current affiliation.\\
@@
 \citep[using the method explained by][]{Drew2005}.
-This is 3.08 magnitudes less than the flux captured 
+This is 3.14 magnitudes less than the flux captured 
 by the much broader $r$ band
@@
 \begin{equation}
-\textsc{magzpt}_{H\alpha} = \textsc{magzpt}_r - 3.08.
+\textsc{magzpt}_{H\alpha} = \textsc{magzpt}_r - 3.14.
 \label{eqn:zpha}
@@
 All data were processed 
-by the Cambridge Astronomical Survey Unit,
+by the Cambridge Astronomical Survey Unit
 at the Institute of Astronomy in Cambridge.
@@
 by a number of software packages,
-including the Python modules
+including the PostgreSQL database software,
+the TOPCAT and STILTS packages \citep{Taylor2005,Taylor2006},
+and the Python modules
 AstroPy \citep{Astropy},
-APLpy, NumPy and SciPy,
-the PostgreSQL database software,
-the TOPCAT and STILTS packages \citep{Taylor2005,Taylor2006},
-and the Montage software maintained by NASA/IPAC.
-We also made use of the SIMBAD, Vizier and Aladin \citep{Aladin}
+NumPy/SciPy \citep{Numpy},
+Matplotlib \citep{Matplotlib},
+IPython \citep{IPython},
+and APLpy.
+We also made use of the Montage software maintained by NASA/IPAC,
+and the SIMBAD, Vizier and Aladin \citep{Aladin} services
 operated at CDS, Strasbourg, France.

### 2014-03-13 paper/iphas-dr2.tex
Update iphas-dr2.tex

@@
 S. E. Sale$^6$,
-plus many contributors.
+M. J. Barlow,
+R. L. M Corradi,
+J. J. Drake,
+J. Fabregat,
+B. T. Gaensicke,
+A. S. Hales,
+J. Irwin,
+C. Knigge,
+D. J. Lennon,
+J. R. Lewis,
+M. Mohr-Smith,
+R. A. H. Morris,
+T. Naylor,
+Q. A. Parker,
+S. Phillipps,
+R. Raddi,
+P. Rodriguez-Gil,
+L. Sabin,
+D. Steeghs,
+Y. C. Unruh,
+K. Viironen,
+J. S. Vink,
+N. A. Walton,
+N. J. Wright,
+A. A. Zijlstra,
+plus all those we forgot.
 \newauthor\\
@@
 on the 2.5-meter Isaac Newton Telescope (INT).
-This data release presents the first 
+It is the first survey to offer comprehensive CCD photometry
+of point sources across the Galactic Plane at visible wavelengths.
+We present the first quality-controlled and
 uniformly-calibrated source catalogue
@@
 for 219 million unique sources
-across 92\% of the survey.
-The observations were carried out between 2003 and 2012
+across 92\% of the survey area.
+The observations were carried out between 2003 and 2013
 at a median seeing of 1.1 arcsec
@@
 and quality control procedures,
-describe and test the new uniform photometric calibration,
+describe and test the new uniform calibration,
 and detail the construction of the source catalogue
 and its quality warning flags.
-We find that the new calibration
+We find that the calibration
 and the photometric repeatability are accurate to
@@
 towards different Galactic sightlines
-and (ii) select candidate \ha-emission line objects.
+and (ii) select candidate \ha\ emission-line objects.
 The catalogue which accompanies this paper
@@
 to several infrared surveys of the Galactic Plane,
-including many poorly-studied star-forming regions and ``bubbles''.
-It also provides images for some of the most
+and provides images for some of the most
 crowded regions to be faced by Gaia.
@@
 available at the outset \citep[e.g.][]{Kohoutek1999}, 
-to be extended into larger, deeper, more statistically-robust samples 
-that in turn could better inform our understanding 
+to be extended into larger, deeper, more statistically-robust samples.
+This in turn could better inform our understanding 
 of the early and late stages of stellar evolution. 
@@
 will extend beyond the traditional \ha\ applications 
-of identifying emission line stars and nebulae. 
+of identifying emission-line stars and nebulae. 
 Through the provision of $r$, $i$ broadband photometry 
@@
 in which quality problems 
-and duplicate detections of the same source 
-are flagged.
+and duplicate detections are flagged.
 
@@
 (ii) describe and test the new global photometric calibration, and 
-(iii) detail the construction and demonstrate the use 
-of the source catalogue that has been extracted 
-from the newly re-calibrated data.
+(iii) detail the construction of the source catalogue
+and demonstrate its use.
 
 In \S\ref{sec:observations} we start by recapitulating the key points
-of the survey design and strategy.
+of the survey observing strategy.
 In \S\ref{sec:reduction} we describe the data reduction
@@
 and finally in \S\ref{sec:demonstration} we demonstrate
-the health of the release by demonstrating its scientific exploitation.
+the scientific exploitation of the colour/magnitude diagrams
+using the new catalogue.
 In \S\ref{sec:conclusions} we conclude and outline
@@
 that captures a sky area of close to 0.29 square degrees.
-To cover the entire northern Plane with some overlap,
+To cover the entire Northern Plane with some overlap,
 the survey area was divided into 7635 telescope pointings.
@@
 amounts to $2 \times 3$ exposures, 
-in which each of the 3 survey filters is exposed at 2 offset sky positions, 
+in which each of the 3 survey filters is exposed at 2 offset sky positions 
 within an elapsed time of 10 minutes.
@@
 and the combination of two fields at a small offset as a \emph{field pair}.
-The survey hence contains 15270 fields grouped into 7635 field pairs.
+The survey hence contains 15\,270 fields grouped into 7\,635 field pairs.
 To achieve the desired survey depth of 20th magnitude or fainter, 
@@
 and every field had been observed at least once by the end of 2008.
-At that time only 76 percent of the field pairs 
+At that time only 76\% of the field pairs 
 satisfied our minimum quality criteria however,  
@@
         Regions with incomplete data are apparent as white gaps (no data) 
-        or as the lightest shade of gray
-        (denoting that only the offset position is missing).}
+        or in light gray (denoting that only the offset position is missing).}
         \label{fig:footprint}
@@
 The fields which remain missing 
--- covering 7 percent of the survey area --
+-- covering 8\% of the survey area --
 are predominantly located towards the Galactic anti-center 
@@
 i.e. fresh observations in all 3 filters may only be obtained 
-at one of the two offset positions, 
-if the data for the partner offset has already passed quality control.  
+at one of the two offset positions.  
 The catalogue is structured such that it is clear 
@@
 
-The reduced images are then stored in a multi-extension FITS file 
+The reduced images are then stored in multi-extension FITS files
 with a primary header describing the characteristics
 (position, filter, exposure time, etc.) 
-and four compressed image extensions 
+and four image extensions 
 corresponding to each of the four CCDs.
@@
 -- when required --
-a deblending routine is applied which also attempts to remove
-the contamination from any other nearby sources.
+a deblending routine is applied which attempts to remove
+the contamination from any nearby sources.
 Whilst this approach works very well 
@@
 with caution.
-In \S\ref{sec:catalogue} we will explain that overlapping
-sources to which the deblending routine was applied 
+In \S\ref{sec:catalogue} we will explain that objects
+which suffer from this problem 
 are flagged in the catalogue using the \emph{deblend} warning flag.
@@
 in which a total of 1.9~billion \emph{candidate sources} were detected 
-at the sensitive default detection level of 1.25\,$\sigma$
--- unavoidably including spurious detections, artefacts and
-duplicate detections 
-(in \S\ref{sec:catalogue} we will explain
-how these have been removed or flagged).
+at the sensitive default detection level of 1.25\,$\sigma$.
+This unavoidably includes spurious detections, artefacts and
+duplicate detections,
+in \S\ref{sec:catalogue} we will explain
+how these have been removed or flagged in the final catalogue.
 The pipelined data set -- comprising 2.5~terabyte of FITS files --
@@
 (iii) contains user-friendly columns and warning flags.
-It is these post-processing steps which are explained below.
+It is these post-processing steps which are explained next.
 
@@
 of the Isaac Newton Group of telescopes,
-which are competitive and invariably over-subscribed.
+which are invariably over-subscribed.
 For this reason, we attempted to utilise 
@@
 Any unsuitable data that was taken as a result
-of this strategy was subsequently flagged and rejected
+of this strategy has subsequently been flagged and removed
 using a series of seven quality criteria,
@@
 for all quality-approved fields,
-which shows a mean depth of 
+showing a mean depth of 
 $21.2\pm0.5$ ($r$), $20.0\pm0.3$ ($i$) and $20.3\pm0.3$ (\ha).
-We found that the depth achieved depended 
+The depth achieved depended 
 most strongly on the presence of the moon,
-which was above the horizon during 62 percent 
-of our observations and explains the wide and bi-modal shape
+which was above the horizon during 62\% 
+of the observations and explains the wide and bi-modal shape
 of the $r$-band limiting magnitude distribution 
@@
 is less sensitive to moonlight
-and the distribution of their depths
-is hence more narrow
-(middle and bottom panel in Fig.~\ref{fig:depth}).
-
-Our minimum limiting magnitude criteria
-have led us to exclude 9\% of the pipelined data.
-We note that much of the excluded data may nevertheless be useful
+and their magnitude limit distributions
+are hence more narrow
+(middle and bottom panels in Fig.~\ref{fig:depth}).
+
+The minimum depth criterion
+has led us to exclude 9\% of the pipelined data,
+which were often obtained under cloudy conditions.
+We note that some of the excluded data may nevertheless be useful
 for e.g. time-domain studies of bright stars.
-The detection tables for any such discarded data are made
-available through our website (www.iphas.org)
+The detection tables of the discarded data are made
+available through our website for this reason (www.iphas.org),
 but are ignored in what follows.
@@
 is $0.09\pm0.04$.
-There have been sporadic episodes with higher ellipticities however
+There have been sporadic episodes with higher ellipticities
 due to mechanical glitches in the telescope tracking system.
@@
 that was obtained under seeing up to 2.5 arcsec,
-so that only 1 per cent of the pipelined exposures
-had to be excluded.
+so that only 1\%of the pipelined exposures
+had to be excluded due to poor seeing.
 In \S\ref{sec:catalogue} we will explain
-that the catalogue is structured
-such that the information on the seeing is included,
-and that the photometry listed
-is based on the exposures with the best-available seeing.
+that the photometry listed in the catalogue
+is predominantly based on the images with the
+best-available seeing, and that the seeing estimate
+is included as a column in the catalogue.
 

### 2014-03-14 paper/iphas-dr2.tex
Update iphas-dr2.tex

@@
 the overlap regions of all field pairs were systematically cross-matched
-to verify the consistency of the photometric measurements
+to verify the consistency of the photometry
 for stars observed in both pointings.
@@
 the images and photometry by eye.
-11\% of the data was rejected as part of this step.
+We rejected 11\% of the data as part of this step
+(many of which had already been rejected as part of the limiting magnitude criterion).
 
@@
 all three bands with a maximum matching distance of 1 arcsec.
-The success of such a cross-matching procedure crucially depends
-on the accuracy of the astrometric solution in each band,
-and hence we were able to detect and correct
-exposures with poor astrometry by inspecting the density map
-for spurious density variations.
+This was particularly effective for revealing
+fields with an inaccurate astrometric solution in one of the bands.
+In the majority of the cases we were able to correct the astrometry
+of such fields by hand.
 
@@
 All images and their associated photometric colour/magnitude diagrams
-were inspected by a team of 20 survey team members, 
+were inspected by a team of 20 survey members, 
 such that each image in the data release 
@@
 excluded from the release if necessary. 
-6\% of the attempts to observe a field
-were placed on a \emph{black-list}
-for this reason, most commonly due to the obvious presence
+6\% of the observations were placed on a \emph{black-list}
+as part of this procedure, most commonly due to the obvious presence
 of clouds or extreme levels of scattered moonlight.
-Such fields were often rejected as part of 
-one or more of the above quality criteria as well.
 
@@
 This ensures that the three bands for a given field
-are observed at nearly the same time --  
-essentially always within 5 minutes.
+are observed at the $\sim$same time --  
+essentially always within an elapsed time of 5 minutes.
 An exception was made for 9 fields where the three exposures 
@@
 
-The above criteria were satisfied 
-for 14115 out of the 15270 fields (92\%).
-In a few cases more than one successful attempt to observe
-a field was available due to the fact that slightly stricter
+The above criteria were satisfied by at least one observing attempt
+for 14\,115 out of the 15\,270 fields (92\%).
+In some cases more than one successful attempt to observe
+a field was available due to the fact that stricter
 quality criteria were adopted in the initial years of the survey.
@@
 with the best seeing and depth was selected
-for inclusion in the release, because the focus 
-of this work is to deliver the most reliable
+for inclusion in the catalogue, because the focus 
+of this release is to deliver the most reliable
 measurement at a single epoch.
 Those interested in any of the rejected data 
-may nevertheless access the full set of detection tables 
-on line.
+can nevertheless data-mine the full set of detection tables 
+via our website.
 
@@
 in the pipeline-corrected CCD frames
-relates to a magnitude $m$ as:
+relate to a magnitude $m$ as:
 \begin{equation}
@@
 Finally, \textsc{percorr} is a correction based on the difference
-between the median dark sky for a CCD against the median for all the CCDs 
+between the median dark sky for a CCD against the median for all the CCDs, 
 and as such is an ancillary correction 
@@
 Vega's SED with the filter throughput curve 
-corrected for atmosphere and detector quantum efficiency
-\citep[using the method explained by][]{Drew2005}.
+corrected for the atmosphere and detector quantum efficiency
+\citep[following the method explained by][]{Drew2005}.
 This is 3.14 magnitudes less than the flux captured 
@@
 at the level of 0.1 mag
-due to atmospheric fluctuations during the night
+due to atmospheric changes during the night
 and imperfections in the pipeline or the instrument
-(e.g. the WFC is known to suffer from sporadic glitches
+(e.g. the WFC is known to suffer from sporadic errors
 in the timing of exposures).
@@
 which all achieved photometry 
-that is globally consistent to within 1--2\%.
+that is globally consistent to within 0.01--0.02~mag.
 
@@
 are used to fit the calibration parameters
-\citep{Padmanabhan2008,Schlafly2012}.
+\citep[e.g.][]{Padmanabhan2008,Schlafly2012}.
 
@@
 and hence stars are not normally observed at more than one epoch,
-unless they happen to fall on the narrow overlap region 
+unless they happen to fall within a narrow overlap region 
 between two neighbouring fields.
@@
 Although the IPHAS data does contain a significant number of 
-inter-field repeat measurements
-due to the small overlaps
-between neighbouring fields,
+such inter-field repeat measurements,
 we have found the information contained
@@
 -- where neighbouring fields overlap -- 
-is itself prone to systematics at the level of 1-2\%.
+is prone to systematics at the level of 1--2\%.
 The cause of these errors is thought to include 
@@
 Moreover, the illumination correction in the overlap regions
-is affected by a radial geometric distortion,
+is most strongly affected by a radial geometric distortion in the WFC,
 which causes the pixel scale towards the edges 
@@
 when they are used to constrain a global calibration
-without external constraints.
+without other external constraints.
 
@@
 on an ubercalibration-type scheme alone,
-and have instead opted to exploit an external reference survey
+and have opted to involve an external reference survey
 -- where available --
@@
 Since 2009,
-APASS has been using two 20~cm-astrographs
+APASS has been using two 20\,cm-astrographs
 to survey the entire sky down to $\sim$17th magnitude
@@
 at the time of preparing this work was APASS DR7,
-which provided data across roughly half of the IPHAS footprint (Fig.~\ref{fig:apass_r}).
+which provides data across $\sim$half of the IPHAS footprint
+The overlap regions are shown in Fig.~\ref{fig:apass_r}.
 The photometric accuracy of APASS is currently estimated 
@@
 than the provisional calibration of IPHAS
-for which we estimate the $1\sigma$-error to be $\sim$10\%.
+for which we estimate the $1\sigma$-uncertainty to be $\sim$10\%.
 APASS achieves its uniform accuracy 
@@
 With the aim of bringing IPHAS to a similar accuracy of $\sim$3\%,
-we used the APASS catalogue to identify and adjust all IPHAS fields 
+we used the APASS catalogue to identify and adjust the calibration of all IPHAS fields 
 which showed a magnitude offset larger than 3\% against APASS.
@@
 are very similar in both surveys.
-The transformations include a large fixed offset
-which is simply due to the fact that 
+The transformations include a large fixed offset,
+but this is simply due to the fact that 
 APASS magnitudes are given in the AB system
@@
 with varying extinction properties to investigate the robustness
-of the transformations with respect to different reddening regimes,
-but the variations at these different
+of the transformations with respect to different reddening regimes.
+The variations at these different
 sightlines were found to be insignificant.
@@
 for each field which contained at least 30 cross-matched stars.
-This was the case for 48\% of the fields.
+This was the case for 48\% of our fields.
 The mean offset was found to be
-$0.014\pm0.104$ for $r$ and $0.007\pm0.108$ for $i$
+$0.014\pm0.104$~mag in $r$ and $0.007\pm0.108$~mag in $i$
 (Table~\ref{tbl:offsets_before}).
@@
 
-We then applied the most important step in our re-calibration scheme,
-which is to adjust the provisional zeropoints of the 4596 aberrant fields
+We then applied the most important step in our calibration scheme,
+which is to adjust the provisional zeropoints of these 4596 aberrant fields
 such that their offset is brought to zero.
 This allowed the mean IPHAS-to-APASS offset 
-to be brought down to $0.000\pm0.011$ in both $r$ and $i$
+to be brought down to $0.000\pm0.011$~mag in both $r$ and $i$
 (Table~\ref{tbl:offsets_after}).
 The procedure of fitting magnitude transformations and
-correcting the zeropoints was repeated a few times to ensure 
+correcting the IPHAS zeropoints was repeated a few times to ensure 
 convergence, which was essentially reached after the first iteration.
@@
              prior to the re-calibration procedure.
-             Each square represents an IPHAS field
+             Each square represents the footprint of an IPHAS field
              which contains at least 30 stars with a counterpart
@@
              which was computed after applying the APASS-to-IPHAS
-             transformation to the APASS magnitudes (Eqn.~\ref{eqn:apass_r}).}
+             transformation to the APASS magnitudes (Eqn.~\ref{eqn:apass_r}).
+             For clarity, we do not show the fields at the offset positions.}
         \label{fig:apass_r}
@@
              procedure was applied.
-             Again, each square represents an IPHAS field
+             Each square represents the footprint of an IPHAS field
              which contains at least 30 stars
@@
              which was computed after applying the SDSS-to-IPHAS
-             transformation to the SDSS magnitudes (Eqn.~\ref{eqn:sdss_r}).}
+             transformation to the SDSS magnitudes (Eqn.~\ref{eqn:sdss_r}).
+             For clarity, we do not show the fields at the offset positions.}
     \label{fig:sdss_r}
@@
 by employing an ubercalibration-style scheme, explained below,
-which minimises the magnitude offsets between stars
+which minimised the magnitude offsets between stars
 located in the overlap regions with neighbouring fields.
@@
 $a_i$ the ZP to solve for 
-and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$), 
+and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$). 
 $w_{ij}$ are weights set to the uncertainty in $\Delta_{ij}$
@@
 on the solution by keeping the zeropoint fixed 
-for the 6756 fields which have been compared
+for the 6\,756 fields which have been compared
 and calibrated against APASS earlier.
-We will hereafter refer to these fields as \emph{anchors}.
+We hereafter refer to these fields as \emph{anchors}.
 It is asserted that the zeropoints $a_i$ of these anchor fields 
@@
 In addition to the APASS-based anchors, 
-we selected 3273 additional anchor fields by hand
+we selected 3\,273 additional anchor fields by hand
 to provide additional constraints in regions not covered by APASS.
@@
 (iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
-which is located at $\sim$500~meter from the INT.
+which is located at $\sim$500\,meter from the INT.
 
@@
 and $+0.01\pm0.12$ in $i$ compared to their provisional calibration.
+The outcome is validated in the next section.
 
@@
 that the corrections required for $r$ and \ha\ are identical
-because the \ha\ zeropoints have been derived directly from the
-$r$-band zeropoints during the provisional calibration (Eqn.~\ref{eqn:zpha}).
-Moreover, the IPHAS data-taking pattern ensured 
+because the IPHAS data-taking pattern ensured 
 that a field's \ha\ and $r$-band exposures
 were taken at essentially the same time, 
-separated only by the $\sim$30-second read-out time required by the WFC.
+separated only by the $\sim$30\,sec read-out time required by the WFC.
 We have hence corrected the \ha\ zeropoints 
@@
 in the earlier steps.
-An exception was made for 3101 fields
+An exception was made for 3\,101 fields
 for which our quality-control routines revealed
@@
 between consecutive fields,
-which indicates non-photometric conditions.
+suggesting strongly non-photometric conditions.
 For good practice, the \ha\ zeropoints of these fields
@@
 
-Having re-calibrated all fields to an expected accuracy of 3\%,
-we then used an independent survey to validate the results.
+Having re-calibrated all fields to the expected APASS-accuracy of 3\%,
+we then used a different survey to validate the results.
 We have been able to exploit SDSS Data Release 9 \citep{Ahn2012}
@@
 providing data across 18\% of the fields in our data release.
-We cross-matched IPHAS fields against 
-stars marked as reliable in the SDSS catalogue\footnote{
+We cross-matched the IPHAS fields against the subset of
+objects marked as reliable stars in the SDSS catalogue\footnote{
 We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
@@
 $15<r_{\rm SDSS}<18.0$ and $14.5<i_{\rm SDSS}<17.5$.
-This provided us with a set of 1.2 million cross-matched stars.
-
-Colour transformations were obtained using a sigma-clipped linear least squares fit:
+This provided us with a set of 1.2~million cross-matched stars.
+
+Colour transformations were again obtained using a sigma-clipped linear least squares fit:
 \begin{eqnarray}
@@
 previously determined for APASS,
-although the colour terms are slightly larger;
-the throughput curve of the SDSS $i$-band filter 
+although the colour terms are slightly larger.
+The throughput curve of the SDSS $i$-band filter 
 appears to be somewhat more sensitive at longer wavelengths
-compared to both the IPHAS and APASS filters.
+than both the IPHAS and APASS filters.
 
@@
 to investigate the effects of varying reddening regimes.
-The transformation coefficients were found 
+The colour term was found 
 to show some variation towards lowly reddened areas,
-which have relatively few numbers
-of (intrinsically) red objects at $r-i > 1$ 
-which can skew the colour term.
+where red objects at $r-i > 1$ are rare.
 The vast majority of red objects in the global sample
@@
 and dominate the statistical appraisal of our calibration.
-%The small number of intrinsically red objects in lowly reddened fields
-%have no impact on the statistical appraisal of our calibration.
 
@@
 we then computed the median magnitude offset for each IPHAS field
-which contained at least 30 objects with a counterpart
-in the SDSS catalogue.
-This was the case for 2602 fields.
+which contained at least 30 objects with a cross-matched counterpart
+in SDSS.
+This was the case for 2\,602 fields.
 The median offsets for each of these fields
 are shown in Figs.~\ref{fig:sdss_r}-\ref{fig:sdss_i}.
-The mean offset and standard deviation found 
-was $-0.001\pm0.029$ for $r$
-and $-0.002\pm0.032$ for $i$ (Table~\ref{tbl:offsets_after}).
+Importantly, the mean offset and standard deviation found 
+is $-0.001\pm0.029$~mag in $r$
+and $-0.002\pm0.032$~mag in $i$ (Table~\ref{tbl:offsets_after}).
 In comparison, offsets computed in the identical way
-\emph{before} our re-calibration showed means
-of $+0.016\pm0.088$ and $+0.010\pm0.089$ (Table~\ref{tbl:offsets_after}).
+\emph{before} carrying out our re-calibration showed means
+of $+0.016\pm0.088$~mag in $r$ and $+0.010\pm0.089$~mag in $i$ (Table~\ref{tbl:offsets_after}).
 We conclude that our re-calibration procedure has
@@
 
-We warn that the SDSS comparison reveals a number of fields with offsets
-exceeding 0.05~mag (523 fields) or even 0.1~mag (18 fields).
+We warn that the SDSS comparison revealed a number of fields where the offsets
+exceeded 0.05~mag (523 fields) or even 0.1~mag (18 fields).
 Such sporadic outliers are consistent with the tails of a Gaussian
-with mean $\sim0$ and $\sigma=0.03$.
+with mean $\approx0$ and $\sigma=0.03$.
+Moreover, the SDSS calibration itself is not a perfect scale itself.
 
@@
 were cross-matched to flag duplicate measurements 
-and identify the best detection 
+and identify the primary (best) detection 
 of each unique source; and
@@
 \end{enumerate}
-Each of these four steps are now explained.
+Each of these four steps are explained next.
 
@@
 (columns \emph{ra} and \emph{dec})
-and in the Galactic coordinate system
+and in Galactic coordinates
 (columns \emph{l} and \emph{b}).
@@
 and the row number in the CASU detection table (\#detection)
--- see columns \emph{rDetectionID},
+-- column names \emph{rDetectionID},
 \emph{iDetectionID}, \emph{haDetectionID}.
@@
 for large apertures on the other hand.
-The user is not restricted to this choice because
+The user is not restricted to this choice, because
 the catalogue also provides magnitudes
@@
 (columns \emph{rPeakMag}, \emph{iPeakMag}, \emph{haPeakMag}),
-a circular 1.2-arcsec-diameter aperture 
+the circular 1.2-arcsec-diameter aperture 
 (\emph{rAperMag1}, \emph{iAperMag1},
  \emph{haAperMag1}) and
-a 3.3-arcsec-diameter aperture 
+the 3.3-arcsec-diameter aperture 
 (\emph{rAperMag3}, \emph{iAperMag3},
@@
 to be consistent with each other
-within the photon noise uncertainties
-(which are given in columns \emph{rErr}, \emph{rPeakMagErr},
+within the photon noise uncertainty
+(which is given in columns \emph{rErr}, \emph{rPeakMagErr},
 \emph{rAperMag1Err}, \emph{rAperMag3Err}, etc).
@@
 it is likely that the source is either
-an extended object or that it has
-been incorrectly measured as a result of
+an extended object for which the aperture correction is invalid,
+or that the object has been incorrectly measured as a result of
 source blending or a rapidly spatially-varying nebulous background.
 In \S\ref{sec:qualitycriteria} we will explain that the consistency
-of the magnitude measurements in the different apertures
+of the different-aperture magnitude measurements
 can be used as a criterion for selecting stellar objects
@@
 into band-merged class probabilities and flags
-(\emph{pStar}, \emph{pGalaxy}, \emph{pNoise},
-\emph{mergedClass}, \emph{mergedClassStat})
 using the merging scheme in use at the WSA\footnote{Explained at
 http://surveys.roe.ac.uk/wsa/www/gloss\_m.html \#gpssource\_mergedclass
-}.
+} (\emph{pStar}, \emph{pGalaxy}, \emph{pNoise},
+\emph{mergedClass}, \emph{mergedClassStat}).
 
@@
 We draw attention to three useful flags
-which warn about the likely presence of systematics:
+which warn about the likely presence of a systematic error:
 \begin{enumerate}
@@
 Although the pipeline attempts to extrapolate the brightness of
-saturated stars based on the shape of their PSF,
-such extrapolation is prone to errors
-and we do not recommend their use.
+saturated stars based on the shape of the PSF,
+such extrapolation is prone to error
+and we do not recommend its use.
 \item The \emph{deblend} column is used to flag sources 
@@
 in each band and hence the $r$-$i$ and $r$-\ha\ colours
-of such objects are prone to errors.
+may be inaccurate if the deblending proceeded differently in each band.
 \item The \emph{brightNeighb} column is used to flag sources which are located
@@
 We will explain below that none of such detections 
-have been included in the catalogue
--- an alternative detection was available in essentially all these situations
-because of the IPHAS field pair strategy --
-and hence these internal warning flags do not appear
-in the final source catalogue.
+have been included in the catalogue.
+An alternative detection was available in essentially all these situations
+as a result of the IPHAS field pair strategy,
+and hence it was not necessary to include these internal warning flags
+in the source catalogue.
 
@@
 with large centroiding errors, 
-and would occasionally fail to link detections near CCD corners
+and would occasionally fail near CCD corners,
 where the astrometric solution can 
@@
 (columns \emph{iXi}, \emph{iEta}, \emph{haXi}, \emph{haEta}).
-We note that also UKIDSS/GPS adopted 
-a maximum matching distance of 1 arcsecond 
+We note that UKIDSS/GPS adopted 
+the same maximum matching distance of 1 arcsec
 for similar reasons \citep{Hambly2008}.
@@
 to be reliable for the vast majority of fields.
-We warn that source blending and confusion is unavoidable
-for faint objects in the Galactic Plane however;
+We do warn that source blending and confusion is unavoidable
+for faint objects in the most crowded regions of the Galactic Plane;
 in \S\ref{sec:discussion} we will show
@@
 of identical sources as a result of overlaps in the data-taking pattern.
-Amongst all sources in the reliable magnitude range $13<r<19$,
+Amongst all sources in the magnitude range $13<r<19$,
 we find that 65\% were detected twice and 25\% were detected three times or more.
@@
 within 10 minutes of the primary,
-the magnitudes of that secondary detection
-have been included in the catalogue
+we have included the identifier and the default photometry
+of that secondary detection
+in the catalogue for convenience
 (columns \emph{sourceID2}, \emph{fieldID2}, 
 \emph{r2}, \emph{i2}, \emph{ha2},
-\emph{rErr2}, \emph{iErr2}, \emph{haErr2}, \emph{errBits2}.
-This is the case for 66\% of the sources brighter than $r<20$
-due to the IPHAS field pair observing pattern.
+\emph{rErr2}, \emph{iErr2}, \emph{haErr2}, \emph{errBits2}).
+Such a secondary detection is available for 66\% of the sources brighter than $r<20$
+due to the IPHAS field pair strategy.
 In addition, the complete set of detection tables -- one for each exposure -- 
@@
 the most reliable information for a unique source.
-Primary detections have been selected using a so-called \emph{seaming} procedure
-which has been adapted from the algorithm developed for the WSA\footnote{http://surveys.roe.ac.uk/wsa/dboverview.html\#merge}.
+Primary detections have been selected using a so-called \emph{seaming} procedure,
+which we adapted from the algorithm developed for the WSA\footnote{http://surveys.roe.ac.uk/wsa/dboverview.html\#merge}.
 In brief, the first step of the procedure is to identify all the duplicate detections
@@
 The duplicate detections for each unique source are then ranked according to
-to (i) filter coverage, (ii) quality score (column \emph{errBits}),
+(i) filter coverage, (ii) quality score (column \emph{errBits}),
 and (iii) the average seeing of stars in the CCD frame rounded to 0.2~arcsec.
@@
 
-As the final step, the primary detections that have been
-selected above are compiled
+As the final step, the primary detections
+selected above were compiled
 into the 98-column source catalogue
 that is described in Appendix~\ref{app:columns}
-and made available on line.
+and is made available on our website and through Vizier.
 The entire list of sources naturally includes 
 a significant number of spurious entries
-as a result of the very sensitive detection levels
-that are employed by the CASU pipeline by default.
+as a result of the very sensitive default detection settings
+that are employed by the CASU pipeline.
 To limit the size of the source catalogue,
@@
 and 31 million entries are single-band detections (14\%).
-Roughly half of the single-band detections were made in the $i$-band.
-This is likely explained by the fact that the $i$-band is least
+Half of the single-band detections were made in the $i$-band.
+This is likely explained by the fact that $i$ is least
 affected by interstellar extinction and can occasionally pick up
@@
 
-Having explained how the catalogue was generated,
+Having explained how the catalogue was created,
 we now offer an overview of its properties
@@
 (i) the recommended quality criteria,
-(ii) the typical photometric uncertainties,
+(ii) the photometric uncertainties and repeatability,
 and (iii) the source densities and the frequency of source blending.
@@
 55\% of the entries in the catalogue
-are fainter than $r > 20$
-and 18\% are even fainter than $r > 21$.
-The measurements of such faint objects
-are naturally prone to large
+are fainter than $r > 20$.
+%and 18\% are even fainter than $r > 21$.
+The measurements of faint objects
+are naturally prone to larger
 random and systematic uncertainties,
@@
 will introduce a proportionally larger systematic error
-in a faint object.
-Most scientific applications will require a set of
+for a faint object.
+Most scientific applications will hence require a set of
 quality criteria to be enforced for the purpose
-of removing such faint and low-quality objects.
-
-The choice of quality criteria is often a complicated
+of removing faint and low-quality objects.
+
+The choice of quality criteria is a complicated
 trade-off between completeness on one hand
@@
 which require reliable $r-i$ and $r-$\ha\ colours
-without removing more than $\sim$80\%
+without removing more than 80\%
 of the sources brighter than $r < 19$.
 The listed criteria are designed to 
-(i) remove low-SNR sources, 
+(i) remove low-S/N sources, 
 (ii) remove saturated sources,
-and (iii) remove objects for which the 2.3~arcsec diameter
-aperture magnitudes are inconsistent 
-with the alternative 1.2~arcsec diameter aperture measurements.
-The last criterion is a proxy
+and (iii) remove objects for which the 2.3-arcsec diameter
+aperture magnitude is inconsistent 
+with its alternative 1.2-arcsec diameter measurement.
+This last criterion is a proxy
 for identifying objects which are affected
@@
 a further set of additional quality criteria
-are listed in Table~\ref{tab:veryreliable}.
+are suggested in Table~\ref{tab:veryreliable}.
 These criteria are designed to ensure that
@@
 (ii) the object was not blended with a nearby neighbour,
-and (ii) the object was not located near a very bright star.
+and (iii) the object was not located near a very bright star.
 59 million sources (27\%) satisfy
@@
 are considered \emph{veryReliable}.
-The stricter criteria filter out a lot
+In \S\ref{sec:densities} we will show that these
+stricter criteria remove a lot
 of objects at early Galactic longitudes 
-where source blending is a common problem
-(we will demonstrate this in \S\ref{sec:densities}).
+where source blending can affect more than a quarter of the objects.
 The \emph{veryReliable} flag should hence
@@
 at the expense of completeness,
-which might be the case for e.g. spectroscopic target selection.
+which is often desired for spectroscopic target selection.
 
@@
 For example, by raising the allowed photometric uncertainties
-from 0.1 mag to 0.2 mag one can retrieve 42 million candidate sources.
+from 0.1 mag to 0.2 mag, 42 million additional candidate sources can be added.
 
@@
 Figure~\ref{fig:uncertainties} shows the mean photometric
-uncertainty as a function of magnitude for each band.
+uncertainty as a function of magnitude in each band.
 We find that the uncertainty typically
-reaches 0.1 mag near 20.5 in $r$ 
-and 19.5 in $i$/\ha when the default 2.3\arcsec\ aperture is used.
+reaches 0.1~mag near 20.5 in $r$ 
+and 19.5 in $i$/\ha\ for the default 2.3\arcsec\ aperture.
 At this point we note that the average colour
 across all sources in the catalogue is
-$1.06\pm0.12$ for ($r$-$i$) and $0.44\pm0.03$ for ($r$-\ha).
+$1.06\pm0.12$ for $r$-$i$ and $0.44\pm0.03$ for $r$-\ha.
 The better depth of $r$ is hence compensated
@@
 by such systematics, we can exploit the
-secondary measurements which were made as part of the
-field-pair observing strategy and are available for 51\%
-of the sources in the catalogue.
-
+secondary measurements which are present in the catalogue.
 Figure~\ref{fig:pairmag} shows the mean residuals between
@@
 reached at $r=14$ ($\sigma_{r-r2}~=~0.041$~mag).
-We draw attention to the fact that brighter stars
-tend to show significantly larger residuals
+The brightest stars tend to show large residuals
 -- e.g. $\sigma_{r-r2}~=~0.14$~mag at $r=12$ --
@@
 show significantly more scatter than would be expected
-from photon noise alone, that is,
+from photon noise alone;
 the effects of source blending and background subtraction
@@
 a similar comparison of the primary and secondary detections,
-but this time we have only include sources which are flagged
+but this time we have only included sources which are flagged
 as \emph{veryReliable} in the catalogue
@@
 We find that the average residuals are significantly better
-for this subset of the catalogue. Sources across the magnitude range 
-$13 < r < 17$ are consistent at the level of 0.03 mag,
-and the best repeatability is again reached at $r=14$
-with $\sigma_{r-r2}~=~0.028$~mag.
+for this subset. Sources across the magnitude range 
+$13 < r < 17$ are now consistent at the level of 0.03 instead of 0.05~mag.
+The best repeatability is again reached at $r=14$,
+but this time with $\sigma_{r-r2}~=~0.028$~mag.
 We conclude that the \emph{veryReliable} quality criteria are effective
 in reducing the systematic errors to
-the same level as the accuracy of the
+the same level as the estimated accuracy of the
 global photometric calibration (cf. \S\ref{sec:calibration}).
@@
 by high levels of foreground extinction.
-In fact dark clouds are visible towards these constellation by eye,
+Dark clouds are readily visible towards these constellation by eye,
 and they have often been referred to as ``the Great Rift''.
 
-We warn however that the densities reported here 
+We warn that the relative densities reported here 
 have not been corrected for survey completeness
-or differences in the observing conditions across the survey.
+or differences in the observing conditions across the survey fields.
 For example, the dip in the density near $l\simeq140^\circ$
 is an artificial feature caused by gaps
-in the footprint coverage (which are apparent in Fig.~\ref{fig:footprint}).
+in the footprint coverage (seen in Fig.~\ref{fig:footprint}).
 In a forthcoming paper,
-we aim to calibrate the IPHAS-based source densities
+we will calibrate the IPHAS-based source densities
 by injecting artificial stars into the IPHAS images
@@
 As we explained earlier, blended sources must be used with caution.
-In particular, we warn that blended sources
+In particular, we warn that blended objects
 are likely candidates to have fallen victim
 to source confusion during the band-merging procedure.
-In future work we hope to investigate the use of
+We have started investigating the use of
 more advanced PSF-fitting routines
 in which sources are measured simultaneously across all bands,
-perhaps guided by an external catalogue such as the one
+perhaps guided by an external astrometric catalogue such as the one
 that is to be provided by Gaia.
@@
 on intrinsic stellar colour and interstellar extinction \citep{Drew2008}. 
-That is, the main sequence in the $r$-$i$/$r$-\ha\ diagram
+Put differently, the main sequence in the $r$-$i$/$r$-\ha\ diagram
 runs in a direction that is different from the reddening vector,
@@
 in the IPHAS colour-colour diagram
-can offer a handle on the properties and extinction regime
-along a line of sight.
+can offer a handle on the properties of the population
+and the extinction along a line of sight.
 
@@
 that the photometric uncertainties
-must be smaller than 0.05 mag in each band
-(corresponding to a magnitude limit near $\sim$19th magnitude).
-
-Each of the diagrams reveals a well-defined locus,
+must be smaller than 0.05 mag in each band,
+corresponding to a magnitude limit near $\sim$19th magnitude.
+
+Each of the diagrams reveal a well-defined locus,
 which demonstrates the health of the catalogue and the global calibration
@@
 which is made available for the IPHAS system through a
-popular on line tool hosted at the Observatory of Padova
+on line tool hosted at the Observatory of Padova
 (http://stev.oapd.inaf.it/cmd).
@@
 
-Finally, in one of our earliest sightlines at $l=30^{\circ}$
+Finally, in one of our earliest sightlines at $l=30^{\circ}$,
 we find a very high number of extremely reddened giants
@@
 which sits roughly 0.4~mags redder than the A0V reddening track.
-At $(l,b)=(45^{\circ}, 2^{\circ})$ the giant stars observed
+At $(l,b)=(45^{\circ}, +2^{\circ})$ the giant stars observed
 exhibit a relatively narrow range of reddenings
@@
 An initial list of candidate \ha-emitters
-based on the first IPHAS data release has previously
-been presented by \cite{Witham2008}. 
-Because no global uniform calibration was available
+based on the first IPHAS data release was previously
+presented by \cite{Witham2008}. 
+Because no uniform calibration was available
 at the time, \citeauthor{Witham2008} employed 
@@
 and $i$ (blue channel) images.
-The ionising star can be seen as the brightest object
+The ionising star can be seen as the bright object
 in the centre of the H{\sc ii} region,
@@
 Many of these young objects 
-appear in Fig.~\ref{fig:mosaic_spitzer}
-as red- and pink-coloured objects,
+appear as red- and pink-coloured stars in the Spitzer mosaic --
 located predominantly in the north-western part
@@
 
-In previous work, we have shown how the majority of
+In previous work, we have shown that the majority of
 \ha-emitters seen towards an H{\sc ii} region
-are likely to be Classical T Tauri Stars (CTTS),
-which are young objects which show \ha\ in emission
+are likely to be Classical T Tauri Stars \citep{Barentsen2011a}.
+These are young objects which show \ha\ in emission
 due to the presence of hot, infalling gas
 which is accreting onto the star
-from a circumstellar disk \cite{Barentsen2011a}.
-This is likely to be the case in this demonstration
-as well. Two of our candidates have recently been identified
+from a circumstellar disk.
+This is likely to be the case for the candidate \ha-emitters
+we discovered towards Sh 2-82 as well.
+Two of our candidates have recently been identified
 as Young Stellar Objects (YSO)
-in an investigation of the region by \cite{Yu2012}.
+in an initial investigation of the region by \cite{Yu2012}.
 In their study, the authors used 2MASS and Spitzer data
@@
 four of our candidates as having
-an infrared excess, we note  that each of them
+an infrared excess, we note that each of these four objects
 is detected at S/N$>$? in the Spitzer 8.0\,\micron\ image,
-which is more likely for a circumstellar-bearing
-object than it is for a field star.
-
-Prior to IPHAS, this region was essentially unstudied
-at faint magnitudes in visible light.
-\cite{Lahulla1985} reported magnitudes for 8 stars in the optical at 
-$V < 15$. In contrast, the IPHAS catalogue offers photometry
-for 2197 objects in the region down to $r<20$.
-
-Sh 2-82 is one of a wealth of poorly-studied star-forming regions
-at low Galactic latitudes,
-which have started to become revealed
-in recent years through catalogues
-of ``bubbles'' seen
-in mid-infrared surveys \cite[e.g.][]{Churchwell2006,Simpson2012}
-and previously unknown clusters seen 
+which is consistent with an optically-unveiled Class II YSO.
+
+%Prior to IPHAS, this region was essentially unstudied
+%at faint magnitudes in visible light.
+%\cite{Lahulla1985} reported magnitudes for 8 stars in the optical at 
+%$V < 15$. In contrast, the IPHAS catalogue offers photometry
+%for 2197 objects in the region down to $r<20$.
+
+Sh 2-82 is one of a large population of poorly-studied star-forming regions
+located at low Galactic latitudes,
+which have only recently started to become revealed
+by efforts to catalogue the wealth of ``bubbles'' detected
+by mid-infrared surveys \cite[e.g.][]{Churchwell2006,Simpson2012},
+and by efforts to catalogue previously unknown clusters seen 
 at near-infrared wavelengths \cite[e.g.][]{Bica2003}.
-The optical data provided by IPHAS offers a handle (or lower limit)
-on the extinction and distance towards many of these regions,
-and can provide information on the optically-unveiled
-members of the region, as we demonstrated here.
+The optical data provided by IPHAS offers a handle -- or lower limit --
+on the extinction and distance towards many of these regions.
+When the extinction is sufficiently low,
+we have shown here that IPHAS can discover optically-unveiled 
+members of the region.
 
@@
 
-A new data release for the IPHAS survey was presented,
-taking the coverage up to over 90\% of the Northern Galactic Plane 
-at $|b|<5^\circ$
-and providing a uniform photometric calibration
-for the first time.
+A new data release for the INT/WFC Photometric \ha\ Survey
+of the Northern Galactic Plane (IPHAS) was presented.
+The new catalogue takes provides CCD photometry in visible light
+across 92\% of the survey area, which spans the entire Northern Milky Way 
+at $|b|<5^\circ$.
 We explained the data reduction and quality control procedures that
@@
 The global calibration and photometric repeatability
-was found to be accurate at the level of $\sigma=0.03$ mag,
+were found to be accurate at the level of $\sigma=0.03$ mag,
 providing a significant improvement over the 
@@
 for 219~million unique sources.
+For 86 million of these objects the catalogue
+provides photometry which can be considered to be reliable
+in all three bands.
 
@@
 The data-taking strategy developed for IPHAS
-have since been reapplied to carry out a companion INT/WFC survey called UVEX
+has since been reapplied to carry out a companion INT/WFC survey called UVEX
 in U/$g$/$r$/He {\sc i} \citep{Groot2009}
@@
 called VPHAS+ \citep{Drew2014}.
-We intend to use the present data release
-as a template for future releases from these companion surveys.
+We intend to use the present work
+as a template for future releases
+to be prepared for these companion surveys.
 
@@
 where it is known as the ``IPHAS DR2 Source Catalogue''
-(catalogue ID ???).
-The full catalogue can also be downloaded in its entirety
+and where it can be queried
+(catalogue ID ??? Conesearch/TAP URL??).
+The catalogue can also be downloaded in its entirety
 from the IPHAS website (www.iphas.org) as a collection 
-of binary FITS tables which comprise 50\,GB,
-which is accompanied by a script
+of binary FITS tables which comprise 50\,GB.
+The website also provides a script
 to ingest the data into a PostgreSQL database.
-Our website also provides access to the pipeline-processed
+Finally, the website also provides access to the pipeline-processed
 imaging data, which we have updated to include

### 2014-03-14 paper/iphas-dr2.tex
-

--- 2014-03-13 paper/iphas-dr2.tex
@@
 of emission-line objects is beyond the scope
-of the present work,
-but in this section we demonstrate how the
-catalogue may be used for this purpose.
+of the present work and will be the focus
+of a forthcoming paper (Barentsen et al, in preparation).
+In this section were merely aim to demonstrate
+how the catalogue may be used for this purpose.
 

### 2014-03-14 paper/iphas-dr2.tex
Spell check and added citations

@@
 \usepackage{longtable} 
-%\usepackage{lscape}
 \usepackage{float}
@@
 S. E. Sale$^6$,
-M. J. Barlow,
-R. L. M Corradi,
-J. J. Drake,
-J. Fabregat,
-B. T. Gaensicke,
-A. S. Hales,
-J. Irwin,
-C. Knigge,
-D. J. Lennon,
-J. R. Lewis,
-M. Mohr-Smith,
-R. A. H. Morris,
-T. Naylor,
-Q. A. Parker,
-S. Phillipps,
-R. Raddi,
-P. Rodriguez-Gil,
-L. Sabin,
-D. Steeghs,
-Y. C. Unruh,
-K. Viironen,
-J. S. Vink,
-N. A. Walton,
-N. J. Wright,
-A. A. Zijlstra,
-plus all those we forgot.
+M. J. Barlow$^{99}$,
+R. L. M Corradi$^{99}$,
+J. J. Drake$^{99}$,
+J. Fabregat$^{99}$, \newauthor
+B. T. Gaensicke$^{99}$,
+A. S. Hales$^{99}$,
+J. Irwin$^{99}$,
+C. Knigge$^{99}$,
+D. J. Lennon$^{99}$, \newauthor
+J. R. Lewis$^{99}$,
+M. Mohr-Smith$^{99}$,
+R. A. H. Morris$^{99}$,
+T. Naylor$^{99}$,
+Q. A. Parker$^{99}$, \newauthor
+S. Phillipps$^{99}$,
+R. Raddi$^{99}$,
+P. Rodriguez-Gil$^{99}$,
+L. Sabin$^{99}$,
+D. Steeghs$^{99}$, \newauthor
+Y. C. Unruh$^{99}$,
+K. Viironen$^{99}$,
+J. S. Vink$^{99}$,
+N. A. Walton$^{99}$,
+N. J. Wright$^{99}$, \newauthor
+A. A. Zijlstra$^{99}$,
+plus those I forgot (sorry).
 \newauthor\\
@@
 of the Northern Galactic Plane (IPHAS)
-is a 1800 square degrees imaging survey
+is a 1800 deg$^2$ imaging survey
 covering the entire northern Milky Way at $|b| < 5^\circ$
@@
 using the Wide Field Camera (WFC) 
-on the 2.5-meter Isaac Newton Telescope (INT).
-It is the first survey to offer comprehensive CCD photometry
-of point sources across the Galactic Plane at visible wavelengths.
+on the 2.5-meter Isaac Newton Telescope (INT) in La Palma.
 We present the first quality-controlled and
@@
 at a median seeing of 1.1 arcsec
-to a depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$ and \ha$=20.3\pm 0.3$
+and to a depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$ and \ha$=20.3\pm 0.3$
 ($5\sigma$ limits, Vega system).
@@
 describe and test the new uniform calibration,
-and detail the construction of the source catalogue
-and its quality warning flags.
+and detail the construction of the 97-column source catalogue.
 We find that the calibration
 and the photometric repeatability are accurate to
-$\sigma=0.03$~mag.
+0.03~mag (rms),
+and provide a series of recommended quality criteria
+which can be used to select the most reliable data
+from the catalogue.
 Finally, we demonstrate the ability of the 
 catalogue's unique
-$r$-$i$/$r$-\ha\ colour-colour diagram to
+$(r-i)$/($r-\ha)$ colour-colour diagram to
 (i) characterise stellar populations and extinction regimes
-towards different Galactic sightlines
+towards different Galactic sight-lines
 and (ii) select candidate \ha\ emission-line objects.
-The catalogue which accompanies this paper
-provides the much-needed visible-light counterpart
-to several infrared surveys of the Galactic Plane,
-and provides images for some of the most
+IPHAS is the first survey to offer comprehensive CCD photometry
+of point sources across the Galactic Plane at visible wavelengths,
+providing the much-needed counterpart
+to several existing near- and mid-infrared surveys,
+and providing images for some of the most
 crowded regions to be faced by Gaia.
@@
 in a new way.
-For example, the survey’s unique $r-H\alpha$ colour, 
-when combined with $r-i$,
+For example, the survey’s unique $(r-\ha)$ colour, 
+when combined with $(r-i)$,
 has been shown to provide simultaneous constraints 
@@
 In \S\ref{sec:catalogue} we explain how the source catalogue was compiled.
-In \S\ref{sec:discussion} we discuss the properties of the catalogue,
-and finally in \S\ref{sec:demonstration} we demonstrate
-the scientific exploitation of the colour/magnitude diagrams
-using the new catalogue.
-In \S\ref{sec:conclusions} we conclude and outline
-our future ambitions.
+In \S\ref{sec:discussion} we discuss the properties of the catalogue
+and in \S\ref{sec:demonstration} we demonstrate
+the scientific exploitation of the colour/magnitude diagrams.
+Finally, in \S\ref{sec:dataaccess} we discuss access
+to the data and source code,
+and in \S\ref{sec:conclusions} we conclude and outline
+our future ambition
 
@@
         and hence the vast majority of the area 
-        is covered twice (dominant gray colour).
+        is covered twice (dominant grey colour).
         There are small overlaps between all the neighbouring fields
         which can be seen as a honeycomb-style pattern
-        of dark gray lines across the survey area.
+        of dark grey lines across the survey area.
         Regions with incomplete data are apparent as white gaps (no data) 
-        or in light gray (denoting that only the offset position is missing).}
+        or in light grey (denoting that only the offset position is missing).}
         \label{fig:footprint}
@@
 -- covering 8\% of the survey area --
-are predominantly located towards the Galactic anti-center 
+are predominantly located towards the Galactic anti-centre 
 at $l > 120^o$.
@@
 Finally, an astrometric solution is determined
-based on the 2MASS point source catalog \citep{Skrutskie2006},
+based on the 2MASS point source catalogue \citep{Skrutskie2006},
 which itself is calibrated 
@@
 are used to fit the calibration parameters
-\citep[e.g.][]{Padmanabhan2008,Schlafly2012}.
+\citep[e.g.][]{Ivezic2007,Padmanabhan2008,Schlafly2012}.
 
@@
 to survey the entire sky down to $\sim$17th magnitude
-in five filters which include Sloan $r$ and $i$.
+in five filters which include Sloan $r$ and $i$ \citep{Henden2012}.
 The most recent catalogue available 
 at the time of preparing this work was APASS DR7,
-which provides data across $\sim$half of the IPHAS footprint
+which provides a good coverage across $\sim$half
+of the IPHAS footprint.
 The overlap regions are shown in Fig.~\ref{fig:apass_r}.
@@
 and IPHAS uses magnitudes relative to Vega.
-Separate transformations were derived for sightlines 
+Separate transformations were derived for sight-lines 
 with varying extinction properties to investigate the robustness
@@
 The variations at these different
-sightlines were found to be insignificant.
+sight-lines were found to be insignificant.
 This is not surprising because heavily reddened objects 
@@
         \caption{Mean magnitude offsets for objects
-                 crossmatched between IPHAS and APASS/SDSS
+                 cross-matched between IPHAS and APASS/SDSS
                  before the uniform re-calibration.
@@
 We have hence corrected the \ha\ zeropoints 
-by re-using the zeropoint adjustements that were derived for the $r$ band
+by re-using the zeropoint adjustments that were derived for the $r$ band
 in the earlier steps.
@@
 for the purpose of validating our uniform calibration in a statistical sense.
-Separate equations were derived towards different sightlines
+Separate equations were derived towards different sight-lines
 to investigate the effects of varying reddening regimes.
@@
 to such objects, the procedure is currently applied separately
-in each band and hence the $r$-$i$ and $r$-\ha\ colours
+in each band and hence the $(r-i)$ and $(r-\ha)$ colours
 may be inaccurate if the deblending proceeded differently in each band.
@@
 \begin{enumerate}
-\item the source must have been detected at SNR$>5$ in at least
+\item the source must have been detected at S/N$>5$ in at least
 one of the bands, i.e. it is required that at least one of
@@
 which should benefit most applications
-which require reliable $r-i$ and $r-$\ha\ colours
+which require reliable $(r-i)$ and $(r-\ha)$ colours
 without removing more than 80\%
@@
   Require the photon noise to be less than 
-  0.1 mag in all bands (i.e. SNR$>$10).
+  0.1 mag in all bands (i.e. S/N$>$10).
   This implicitly requires a detection in all three bands.  \\
@@
 across all sources in the catalogue is
-$1.06\pm0.12$ for $r$-$i$ and $0.44\pm0.03$ for $r$-\ha.
+$1.06\pm0.12$ for $(r-i)$ and $0.44\pm0.03$ for $(r-\ha)$.
 The better depth of $r$ is hence compensated
@@
              in the catalogue, 
-             and the errorbars show the standard deviations.
+             and the error-bars show the standard deviations.
              The dashed and solid lines indicate 
@@
 the effects of source blending and background subtraction
-appear to dominate from $\sim$20th magnitude onwards.
+appear to dominate from $\sim$20th magnitude onward.
 
@@
          between the primary and the secondary detections,
-         while the errorbars show the standard deviations.
+         while the error-bars show the standard deviations.
          Left panel: 
@@
 We conclude this paper by demonstrating how the unique
-$r$-$i$/$r$-\ha\ colour-colour diagram offered by this catalogue
+$(r-i)$/$(r-\ha)$ colour-colour diagram offered by this catalogue
 can readily be used to
-(i) characterise the extinction regime at different sightlines, and
+(i) characterise the extinction regime at different sight-lines, and
 (ii) identify \ha\ emission-line objects.
@@
 
-The survey's unique $r$-\ha\ colour,
-when combined with $r$-$i$,
+The survey's unique $(r-\ha)$ colour,
+when combined with $(r-i)$,
 has been shown to provide simultaneous constraints 
 on intrinsic stellar colour and interstellar extinction \citep{Drew2008}. 
-Put differently, the main sequence in the $r$-$i$/$r$-\ha\ diagram
+Put differently, the main sequence in the $(r-i)$/$(r-\ha)$ diagram
 runs in a direction that is different from the reddening vector,
@@
 as a coarse proxy for spectral type
-and is less sensitive to reddening than $r$-$i$.
+and is less sensitive to reddening than $(r-i)$.
 As a result, the distribution of a stellar population
@@
 where we present three sets of IPHAS colour/magnitude diagrams
-towards three distinct sightlines
+towards three distinct sight-lines
 located at Galactic longitudes 
@@
 centred on the coordinates indicated in the diagram
-(i.e. within a radius of $0.564^\circ$ from the indicated sightline).
+(i.e. within a radius of $0.564^\circ$ from the indicated sight-line).
 For clarity, we have imposed the additional criterion
@@
 
-Each of the sightlines reveal a stellar population
+Each of the sight-lines reveal a stellar population
 with distinct characteristics.
@@
 we find a population dominated by lowly-reddened main sequence stars.
-This is consistent with the estimated total sightline extinction 
+This is consistent with the estimated total sight-line extinction 
 of $E(B-V)=0.49$ given by \cite{Schlegel1998},
@@
 we do not observe M dwarfs experiencing 
-the strongest reddening possible for this sightline.
+the strongest reddening possible for this sight-line.
 This implies that extinction is still increasing 
@@
 
-Finally, in one of our earliest sightlines at $l=30^{\circ}$,
+Finally, in one of our earliest sight-lines at $l=30^{\circ}$,
 we find a very high number of extremely reddened giants
@@
 of foreground dwarfs.
-In contrast to the sightline at $l=45^{\circ}$,
+In contrast to the sight-line at $l=45^{\circ}$,
 there is no clear group of giant stars visible
@@
     which has been placed at an arbitrary distance of 2~kpc.
-    This is one of the least reddened sightlines
+    This is one of the least reddened sight-lines
     in the survey
@@
     \caption{Same as above for $(l,b)=(45^\circ,+2^\circ)$,
-    which is one of the highest-density sightlines in the survey,
+    which is one of the highest-density sight-lines in the survey,
     revealing two groups of stars in colour-magnitude space.}
@@
     \caption{Same as above for $(l,b)=(30^\circ,0^\circ)$,
-    showing one of the most reddened sightlines in the survey.
+    showing one of the most reddened sight-lines in the survey.
     }
@@
 allows objects to be picked out
-from the $r$-$i$/$r$-\ha\ colour-colour diagram
+from the $(r-i)$/$(r-\ha)$ colour-colour diagram
 using model-based colour criteria
@@
 The diagram also shows the unreddened main sequence (solid line)
-and the expected position of unredded main-sequence stars
+and the expected position of unreddened main-sequence stars
 with \ha\ in emission
@@
   \includegraphics[width=0.45\textwidth]{./plots/sh2-82-ccd.pdf}
-    \caption{$r$-$i$/$r$-\ha\ diagram for the rectangular region of 
+    \caption{$(r-i)$/$(r-\ha)$ diagram for the rectangular region of 
     20-by-15 arcmin centred on the H{\sc ii} region Sh 2-82,
@@
 
-%Prior to IPHAS, this region was essentially unstudied
-%at faint magnitudes in visible light.
-%\cite{Lahulla1985} reported magnitudes for 8 stars in the optical at 
-%$V < 15$. In contrast, the IPHAS catalogue offers photometry
+%The IPHAS catalogue offers photometry
 %for 2197 objects in the region down to $r<20$.
@@
 
+\section{Data access and source code}
+\label{sec:dataaccess}
+
+The catalogue is made available through the Vizier
+catalogue tool (http://vizier.u-strasbg.fr),
+where it is known as the ``IPHAS DR2 Source Catalogue''
+and where it can be queried
+(catalogue ID ??? Conesearch/TAP URL??).
+The catalogue can also be downloaded in its entirety
+from the IPHAS website (www.iphas.org) as a collection 
+of binary FITS tables which comprise 50\,GB,
+together with a script
+to ingest these tables into a PostgreSQL database.
+Finally, the website also provides access to the pipeline-processed
+imaging data, which we have updated to include
+the re-calibrated DR2 zeropoint in the image headers.
+
+In the spirit of reproducibility,
+the source code that was used to generate
+the catalogue is made available at
+https://github.com/barentsen/iphas-dr2
+
 \section{Conclusions and future work}
@@
 of the Northern Galactic Plane (IPHAS) was presented.
-The new catalogue takes provides CCD photometry in visible light
-across 92\% of the survey area, which spans the entire Northern Milky Way 
-at $|b|<5^\circ$.
+It is the first survey to offer comprehensive CCD photometry
+of point sources across the Galactic Plane at visible wavelengths,
+taking in the entire Northern Milky Way at $|b|<5^\circ$.
+The new 97-column catalogue provides single-epoch photometry
+across 92\% of the survey area,
+and is the first quality-controlled and uniformly calibrated
+catalogue to have been constructed from the imaging data.
 We explained the data reduction and quality control procedures that
-were applied, described and tested the new global photometric calibration,
-and detailed the construction of the user-friendly source catalogue.
+were applied,
+described and tested the global photometric calibration,
+and detailed the construction of the source catalogue.
 
@@
 as a template for future releases
-to be prepared for these companion surveys.
-
-\section*{Data access and source code}
-\label{sec:dataaccess}
-
-The catalogue is made available through the Vizier
-catalogue tool (http://vizier.u-strasbg.fr),
-where it is known as the ``IPHAS DR2 Source Catalogue''
-and where it can be queried
-(catalogue ID ??? Conesearch/TAP URL??).
-The catalogue can also be downloaded in its entirety
-from the IPHAS website (www.iphas.org) as a collection 
-of binary FITS tables which comprise 50\,GB.
-The website also provides a script
-to ingest the data into a PostgreSQL database.
-Finally, the website also provides access to the pipeline-processed
-imaging data, which we have updated to include
-the re-calibrated DR2 zeropoint in the image headers.
-
-The source code that was used to generate
-the catalogue is available at
-https://github.com/barentsen/iphas-dr2
-
-\section*{Acknowledgments}
+to be prepared for these companion surveys,
+allowing new insights into the contents and structure 
+of our own backyard, the Milky Way, to be revealed.
+
+\section*{Acknowledgements}
 

### 2014-03-14 paper/iphas-dr2.tex
Typos

@@
 \author[G. Barentsen
-et. al]{Geert Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
+et. al]{G. Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
 H. J. Farnhill$^1$,
 J. E. Drew$^1$,
-E. A. Gonz$\acute{\rm{a}}$lez-Solares$^2$, \newauthor
-R. Greimel$^3$,
+E. A. Gonz$\acute{\rm{a}}$lez-Solares$^2$, 
+R. Greimel$^3$, \newauthor
 M. J. Irwin$^2$,
-B. Mizalski$^{99}$,
+B. Mizalski$^{4}$,
 C. Ruhland$^1$,
-P. Groot$^4$,
-A. Mampaso$^5$, \newauthor
-S. E. Sale$^6$,
-M. J. Barlow$^{99}$,
-R. L. M Corradi$^{99}$,
-J. J. Drake$^{99}$,
-J. Fabregat$^{99}$, \newauthor
-B. T. Gaensicke$^{99}$,
-A. S. Hales$^{99}$,
-J. Irwin$^{99}$,
-C. Knigge$^{99}$,
-D. J. Lennon$^{99}$, \newauthor
-J. R. Lewis$^{99}$,
-M. Mohr-Smith$^{99}$,
-R. A. H. Morris$^{99}$,
-T. Naylor$^{99}$,
-Q. A. Parker$^{99}$, \newauthor
-S. Phillipps$^{99}$,
-R. Raddi$^{99}$,
-P. Rodriguez-Gil$^{99}$,
-L. Sabin$^{99}$,
-D. Steeghs$^{99}$, \newauthor
-Y. C. Unruh$^{99}$,
-K. Viironen$^{99}$,
-J. S. Vink$^{99}$,
-N. A. Walton$^{99}$,
-N. J. Wright$^{99}$, \newauthor
-A. A. Zijlstra$^{99}$,
-plus those I forgot (sorry).
+P. Groot$^5$,
+A. Mampaso$^6$, 
+S. E. Sale$^7$, \newauthor
+M. J. Barlow$^{8}$,
+R. L. M Corradi$^{6}$,
+J. J. Drake$^{9}$,
+J. Fabregat$^{10}$, 
+B. T. Gaensicke$^{11}$, \newauthor
+A. S. Hales$^{12}$,
+J. Irwin$^{9}$,
+C. Knigge$^{13}$,
+D. J. Lennon$^{14}$, 
+J. R. Lewis$^{2}$,
+M. Mohr-Smith$^{1}$, \newauthor
+R. A. H. Morris$^{15}$,
+T. Naylor$^{16}$,
+Q. A. Parker$^{17}$, 
+S. Phillipps$^{13}$,
+R. Raddi$^{11}$, \newauthor
+P. Rodriguez-Gil$^{6}$,
+L. Sabin$^{18}$,
+D. Steeghs$^{11}$, 
+Y. C. Unruh$^{19}$,
+K. Viironen$^{20}$, \newauthor
+J. S. Vink$^{21}$, 
+N. A. Walton$^{2}$,
+N. J. Wright$^{1}$,
+A. A. Zijlstra$^{22}$.
 \newauthor\\
@@
 $^{3}$IGAM, Institute of Physics, University of Graz, Universit\"atsplatz 5, Graz, Austria.\\
-$^{4}$Afdeling Sterrenkunde, Radboud Universiteit Nijmegen, Faculteit NWI, Postbus 9010, 6500 GL Nijmegen, The Netherlands.\\
-$^{5}$Instituto de Astrof\'isica de Canarias, 38200 La Laguna, Tenerife, Spain.\\
-$^{6}$Rudolf Peierls Centre for Theoretical Physics, Keble Road, Oxford, OX1 3NP, U.K.\\
-$^{99}$Please add your current affiliation.\\
-}
+$^{4}$South African Astronomical Observatory, P.O. Box 9, Observatory, 7935 Cape Town, South Africa.\\
+$^{5}$Afdeling Sterrenkunde, Radboud Universiteit Nijmegen, Faculteit NWI, Postbus 9010, 6500 GL Nijmegen, The Netherlands.\\
+$^{6}$Instituto de Astrof\'isica de Canarias, 38200 La Laguna, Tenerife, Spain.\\
+$^{7}$Rudolf Peierls Centre for Theoretical Physics, Keble Road, Oxford, OX1 3NP, U.K.\\
+$^{8}$University College London, Department of Physics \& Astronomy, 
+Gower Street, London WC1E 6BT, U.K.\\
+$^{9}$Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, 
+Cambridge, MA 02138, U.S.A. \\
+$^{10}$Observatorio Astr\'onomico, Universidad de Valencia,
+Catedr\'atico Jos\'e Beltr\'an 2, 46980 Paterna, Spain\\
+$^{11}$Department of Physics, University of Warwick, Gibbet Hill Road, Coventry, CV4 7AL, U.K.\\
+$^{12}$ Joint ALMA Observatory, Alonso de Córdova 3107, Vitacura 763-0355, Santiago, Chile. \\
+$^{13}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
+$^{14}$European Space Astronomy Centre (ESAC), Villafranca del Castillo, Villanueva de la Canada, E-28692 Madrid, Spain.\\
+$^{15}$School of Physics, Bristol University, Tyndall Avenue, Bristol, BS8 1TL, U.K.\\
+$^{16}$School of Physics, University of Exeter, Stocker Road, Exeter, EX4 4QL, U.K.\\
+$^{17}$Department of Physics \& Astronomy, Macquarie University, NSW 2109, Australia\\
+$^{18}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
+$^{19}$Department of Physics, Blackett Laboratory, Imperial College
+London, Prince Consort Road, London, SW7 2AZ, U.K.\\
+$^{20}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
+$^{21}$Armagh Observatory, College Hill, Armagh, Northern Ireland,
+BT61 9DG, U.K.\\
+$^{22}$Jodrell Bank Centre for Astrophysics, School of Physics \&
+Astronomy, University of Manchester, Oxford Road, Manchester M13 9PL, U.K.}
 
@@
 \date{Current draft typeset \today}
-\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2013}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2014}
 
@@
 and detail the construction of the 97-column source catalogue.
-We find that the calibration
+We show that the calibration
 and the photometric repeatability are accurate to
@@
 providing the much-needed counterpart
-to several existing near- and mid-infrared surveys,
-and providing images for some of the most
+to infrared surveys,
+whilst also providing images for some of the most
 crowded regions to be faced by Gaia.
@@
 imaging data, which we have updated to include
-the re-calibrated DR2 zeropoint in the image headers.
+the re-calibrated zeropoints and the best-available
+astrometric solution in the image headers.
 
@@
 The global calibration and photometric repeatability
-were found to be accurate at the level of $\sigma=0.03$ mag,
+were found to be accurate at the level of $0.03$ mag (rms),
 providing a significant improvement over the 
@@
 for 219~million unique sources.
-For 86 million of these objects the catalogue
-provides photometry which can be considered to be reliable
-in all three bands.
-
-In future work, we aim to draw upon the PanSTARRS photometric
+We detailed a list of recommended quality criteria
+to select objects with reliable colours from the catalogue.
+Finally, we demonstrated the use of the survey's unique
+$(r-i)$/$(r-\ha)$ diagram for characterising stellar populations
+and selecting emission-line objects.
+
+In future work, we intend to draw upon the PanSTARRS photometric
 survey to further improve the accuracy of our global calibration.
@@
 radial field distortions,
-and aim to produce a catalogue
+and plan to produce a catalogue 
 which details all the secondary detections
@@
 in U/$g$/$r$/He {\sc i} \citep{Groot2009}
-and a southern companion survey in $u$/$g$/$r$/$i$/\ha\ 
+and a southern survey in $u$/$g$/$r$/$i$/\ha\ 
 called VPHAS+ \citep{Drew2014}.
 We intend to use the present work
-as a template for future releases
-to be prepared for these companion surveys,
-allowing new insights into the contents and structure 
-of our own backyard, the Milky Way, to be revealed.
-
-\section*{Acknowledgements}
+as a template for generating catalogues
+from these companion surveys.
+When combined, these surveys
+will offer new insights
+into the contents and structure of the Milky Way
+across 360$^\circ$.
+
+\section*{Acknowledgments}
 
@@
 
+\label{lastpage}
+
 \bibliographystyle{mn2e}
@@
 
-\label{lastpage}
-
 \end{document}

### 2014-03-14 paper/iphas-dr2.tex

@@
 at a median seeing of 1.1 arcsec
-and to a depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$ and \ha$=20.3\pm 0.3$
+and to a mean depth of $r=21.2$ / $i=20.0$ / $\ha=20.3$
 ($5\sigma$ limits, Vega system).
@@
 and detail the construction of the 97-column source catalogue.
-We show that the calibration
-and the photometric repeatability are accurate to
-0.03~mag (rms),
+We show that the global calibration is accurate to
+0.03~mag (rms) by validating against SDSS,
 and provide a series of recommended quality criteria
@@
 providing the much-needed counterpart
-to infrared surveys,
-whilst also providing images for some of the most
+to infrared surveys.
+We also provide images for some of the most
 crowded regions to be faced by Gaia.

### 2014-03-17 paper/iphas-dr2.tex
Assign subfigures to Fig. 10
The two plots in Fig. 10 were referenced separately in the text, but were appearing as Fig. 6.2 (the section containing the reference). The subcaption package has been added to give a/b subfigure lettering, and the caption package to set the letters not to display. \captionsetup can be commented out to have the letters appear.

This does have the unfortunate effect of (slightly) increasing the space between the figures and the caption.

@@
 \usepackage{aas_macros}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 using the Wide Field Camera (WFC) 
-on the 2.5-meter Isaac Newton Telescope (INT) in La Palma.
+on the 2.5-metre Isaac Newton Telescope (INT) in La Palma.
 We present the first quality-controlled and
@@
 It is carried out using the Wide Field Camera (WFC) 
-on the 2.5-meter Isaac Newton Telescope (INT) in La Palma. 
+on the 2.5-metre Isaac Newton Telescope (INT) in La Palma. 
 It is the first digital survey to offer comprehensive CCD photometry
@@
              which contains at least 30 stars
-             with a counterpart in the SDSS catalogue.
+             with a counterpart in the SDSS DR9 catalogue.
              The colours denote the median IPHAS-SDSS magnitude offset
@@
 (iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
-which is located at $\sim$500\,meter from the INT.
+which is located $\sim$500\,metres from the INT.
 
@@
 or which were otherwise affected by bad pixels in the detector.
-We will explain below that none of such detections 
+We will explain below that no such detections 
 have been included in the catalogue.
@@
    % pStar > 0.89 is identical to mergedClass == -1
-   The object must appears as a perfect point source,
+   The object must appear as a perfect point source,
    as inferred from comparing its Point Spread Function (PSF)
@@
 \begin{figure*}
-\begin{minipage}[b]{0.45\linewidth}
+\captionsetup[subfigure]{labelformat=empty}
+\begin{subfigure}[b]{0.45\linewidth}
 \centering
 \includegraphics[width=\textwidth]{./plots/pairmag/pairmag.pdf}
+\caption{}
 \label{fig:pairmag}
-\end{minipage}
+\end{subfigure}
 \hspace{0.5cm}
-\begin{minipage}[b]{0.45\linewidth}
+\begin{subfigure}[b]{0.45\linewidth}
 \centering
 \includegraphics[width=\textwidth]{./plots/pairmag/pairmag-reliable.pdf}
+\caption{}
 \label{fig:pairmag_reliable}
-\end{minipage}
+\end{subfigure}
 \caption{Photometric repeatability as a function of magnitude

### 2014-03-17 paper/iphas-dr2.tex
Various minor improvements

@@
 C. Knigge$^{13}$,
+T. Kupfer$^{5}$,
 D. J. Lennon$^{14}$, 
-J. R. Lewis$^{2}$,
-M. Mohr-Smith$^{1}$, \newauthor
+J. R. Lewis$^{2}$, \newauthor
+M. Mohr-Smith$^{1}$, 
 R. A. H. Morris$^{15}$,
@@
 Q. A. Parker$^{17}$, 
-S. Phillipps$^{13}$,
-R. Raddi$^{11}$, \newauthor
+S. Phillipps$^{13}$, \newauthor
+R. Raddi$^{11}$, 
 P. Rodriguez-Gil$^{6}$,
@@
 D. Steeghs$^{11}$, 
-Y. C. Unruh$^{19}$,
-K. Viironen$^{20}$, \newauthor
+Y. C. Unruh$^{19}$, \newauthor
+K. Viironen$^{20}$, 
 J. S. Vink$^{21}$, 
@@
 $^{22}$Jodrell Bank Centre for Astrophysics, School of Physics \&
-Astronomy, University of Manchester, Oxford Road, Manchester M13 9PL, U.K.}
+Astronomy, University of Manchester, Manchester M13 9PL, U.K.}
 
@@
 to infrared surveys.
-We also provide images for some of the most
+IPHAS also provide images for some of the most
 crowded regions to be faced by Gaia.
@@
 For example, by raising the allowed photometric uncertainties
-from 0.1 mag to 0.2 mag, 42 million additional candidate sources can be added.
+from 0.1 mag to 0.2 mag,
+42 million additional candidate sources can be added.
 
@@
 a sigma-clipping technique to select objects with
-large, outlying $r$-\ha\ colours.
+large, outlying $(r-\ha)$ colours.
 In contrast, the new catalogue
@@
 a mosaic of the same region 
-as seen in the mid-infrared by the Spitzer Space Telescope
+as seen in the mid-infrared by the Spitzer Space Telescope's
+Galactic Plane survey \citep[GLIMPSE;][]{Benjamin2003,Churchwell2009}.
 The infrared image reveals a bubble-shaped structure
-of warm dust which surrounds the entire H{\sc ii} region,
-and is thought to contain a wealth of infrared-bright
-young stellar objects \citep{Yu2012}.
+which is thought to originate from the
+mid-infrared emission of Poly-Aromatic Hydrocarbons (PAHs)
+-- i.e. warm dust --
+which is frequently observed
+at the interface between neutral regions of interstellar material
+and the ionising radiation from early-type stars \citep{Churchwell2006}.
+In this example, \cite{Yu2012}
+recently noted that the warm dust
+that surrounds Sh~2-82 
+appears to contain infrared-bright
+Young Stellar Objects (YSOs).
 Many of these young objects 
-appear as red- and pink-coloured stars in the Spitzer mosaic --
-located predominantly in the north-western part
-of the bubble-shaped structure.
+appear as red- and pink-coloured stars
+in Fig.~\ref{fig:mosaic_spitzer},
+located predominantly in the top-left part
+of the bubble.
 
@@
     \caption{Star-forming region Sh 2-82 as seen in the mid-infrared
-    by the Spitzer Space Telescope. The mosaic is composed of the 24\,\micron\ (red), 8.0\,\micron\ (green) and 4.5\,\micron\ (blue) bands.
-    The image reveals a bubble-shaped structure which surrounds the {\sc Hii} region that is seen in the IPHAS mosaic of the same region (Figure~\ref{fig:mosaic_iphas}).}
+    by the Spitzer Space Telescope's GLIMPSE survey. The mosaic is composed of the 24\,\micron\ (red), 8.0\,\micron\ (green) and 4.5\,\micron\ (blue) bands.
+    The image reveals a bubble-shaped structure which surrounds the {\sc Hii} region that is seen in the IPHAS mosaic of the same region (Figure~\ref{fig:mosaic_iphas}). 
+    This structure has previously been labelled as N115 in the 
+catalogue due to \citet{Churchwell2006}.}
     \label{fig:mosaic_spitzer}
@@
 that is, the distance between the objects and the dashed line
-is larger than three times the uncertainty on the $r$-\ha\ colour.
+is larger than three times the uncertainty
+in their $(r-\ha)$ colour.
 These reliable candidate \ha-emitters
@@
 are likely to be Classical T Tauri Stars \citep{Barentsen2011a}.
-These are young objects which show \ha\ in emission
+These are young objects
+which are thought to show \ha\ in emission
 due to the presence of hot, infalling gas
@@
 Two of our candidates have recently been identified
-as Young Stellar Objects (YSO)
-in an initial investigation of the region by \cite{Yu2012}.
+as YSOs in the initial investigation of the region by \cite{Yu2012}.
 In their study, the authors used 2MASS and Spitzer data
@@
 Although \citeauthor{Yu2012} did not identify
-four of our candidates as having
-an infrared excess, we note that each of these four objects
-is detected at S/N$>$? in the Spitzer 8.0\,\micron\ image,
-which is consistent with an optically-unveiled Class II YSO.
-
-%The IPHAS catalogue offers photometry
-%for 2197 objects in the region down to $r<20$.
+the other four candidate emitters in our sample
+as having an infrared excess,
+we note that each of these four objects
+are detected in the Spitzer 8.0\,\micron\ image
+at S/N$>$5,
+which is consistent with the objects
+being optically-unveiled Class II YSOs
+with a mild infrared excess.
 
@@
 by efforts to catalogue the wealth of ``bubbles'' detected
-by mid-infrared surveys \cite[e.g.][]{Churchwell2006,Simpson2012},
+at mid-infrared wavelengths \citep{Churchwell2006,Simpson2012},
 and by efforts to catalogue previously unknown clusters seen 
-at near-infrared wavelengths \cite[e.g.][]{Bica2003}.
-The optical data provided by IPHAS offers a handle -- or lower limit --
-on the extinction and distance towards many of these regions.
-When the extinction is sufficiently low,
-we have shown here that IPHAS can discover optically-unveiled 
-members of the region.
-
-%TODO: annotate -10 EW text on plot?
-%TODO: mention Gaia, e.g. IPHAS for transient identification?
-%\subsection{Caveats and lessons learnt}
-%\subsection{Comparison with Gaia}?
+in the near-infrared \cite[e.g.][]{Bica2003}.
+The data provided by IPHAS offers a handle
+or lower limit on the extinction and distance
+towards these regions,
+and can be used to discover optically-unveiled
+T~Tauri stars when the extinction is sufficiently low.
 
@@
 and where it can be queried
-(catalogue ID ??? Conesearch/TAP URL??).
+(catalogue ID TBD; Conesearch/TAP URL TBD?).
 The catalogue can also be downloaded in its entirety
@@
 survey to further improve the accuracy of our global calibration.
-We will also aim to correct the IPHAS photometry for the
+We also aim to correct the IPHAS photometry for the
 radial field distortions,
-and plan to produce a catalogue 
+plan to produce a catalogue 
 which details all the secondary detections
-to aid time-domain studies.
+to aid time-domain studies,
+and are investigating the use of photometry
+derived from PSF-fitting across the survey.
+%TODO: mention Gaia, e.g. IPHAS for transient identification?
 
@@
 of the Instituto de Astrofisica de Canarias.
+We are deeply indebted to the ING staff and students
+for their ongoing support of the INT.
 All data were processed 
@@
 funded by NASA and the NSF.
+This work includes observations made
+with the Spitzer Space Telescope,
+which is operated by the Jet Propulsion Laboratory,
+California Institute of Technology under a contract with NASA.
 

### 2014-03-18 paper/iphas-dr2.tex
Incorporating feedback from JED

@@
 L. Sabin$^{18}$,
+S. Scaringi$^{19}$,
 D. Steeghs$^{11}$, 
-Y. C. Unruh$^{19}$, \newauthor
-K. Viironen$^{20}$, 
-J. S. Vink$^{21}$, 
+Y. C. Unruh$^{20}$, \newauthor
+K. Viironen$^{21}$, 
+J. S. Vink$^{22}$, 
 N. A. Walton$^{2}$,
 N. J. Wright$^{1}$,
-A. A. Zijlstra$^{22}$.
-\newauthor\\
+A. A. Zijlstra$^{23}$.
+\\
 $^{1}$School of Physics, Astronomy \& Mathematics, University of Hertfordshire, College Lane, Hatfield, Hertfordshire, AL10 9AB, U.K.\\
@@
 $^{11}$Department of Physics, University of Warwick, Gibbet Hill Road, Coventry, CV4 7AL, U.K.\\
-$^{12}$ Joint ALMA Observatory, Alonso de Córdova 3107, Vitacura 763-0355, Santiago, Chile. \\
+$^{12}$Joint ALMA Observatory, Alonso de Córdova 3107, Vitacura 763-0355, Santiago, Chile. \\
 $^{13}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
@@
 $^{18}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
-$^{19}$Department of Physics, Blackett Laboratory, Imperial College
+$^{19}$Instituut voor Sterrenkunde, K.U. Leuven, Celestijnenlaan 200D, B-3001 Leuven, Belgium\\
+$^{20}$Department of Physics, Blackett Laboratory, Imperial College
 London, Prince Consort Road, London, SW7 2AZ, U.K.\\
-$^{20}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
-$^{21}$Armagh Observatory, College Hill, Armagh, Northern Ireland,
+$^{21}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
+$^{22}$Armagh Observatory, College Hill, Armagh, Northern Ireland,
 BT61 9DG, U.K.\\
-$^{22}$Jodrell Bank Centre for Astrophysics, School of Physics \&
+$^{23}$Jodrell Bank Centre for Astrophysics, School of Physics \&
 Astronomy, University of Manchester, Manchester M13 9PL, U.K.}
@@
 at a median seeing of 1.1 arcsec
-and to a mean depth of $r=21.2$ / $i=20.0$ / $\ha=20.3$
-($5\sigma$ limits, Vega system).
+and to a mean $5\sigma$-depth of 
+21.2 ($r$), 20.0 ($i$) and 20.3 ($\ha$)
+in the Vega magnitude system.
 We explain the data reduction 
@@
 describe and test the new uniform calibration,
-and detail the construction of the 97-column source catalogue.
+and detail the construction of the new 97-column source catalogue.
 We show that the global calibration is accurate to
@@
 catalogue's unique
-$(r-i)$/($r-\ha)$ colour-colour diagram to
+$(r-\ha,\ r-i)$ colour-colour diagram to
 (i) characterise stellar populations and extinction regimes
@@
 providing the much-needed counterpart
-to infrared surveys.
-IPHAS also provide images for some of the most
-crowded regions to be faced by Gaia.
+to recent infrared surveys.
 \end{abstract}
@@
 \section{Introduction}
-
-Since the first data release in 2008, 
-the INT/WFC Photometric H$\alpha$ Survey 
-of the Northern Galactic Plane (IPHAS)
-has provided new insights into the contents and structure 
+The INT/WFC Photometric H$\alpha$ Survey
+of the Northern Galactic Plane \citep[IPHAS;][]{Drew2005}
+is providing new insights into the contents and structure 
 of our own backyard, the Milky Way.
@@
 of the early and late stages of stellar evolution.
-This aim has begun to be realised through a
+Since the publication of the IPHAS Initial Data Release \citep[IDR;][]{Gonzalez-Solares2008},
+these aims have begun to be realised through a
 range of published studies including: 
@@
 
-The first release of IPHAS data, 
-covering roughly half the survey footprint,
-was made in late 2007 \citep{Gonzalez-Solares2008}. 
-At the time the data were insufficiently complete 
+When the initial data release was created in late 2007,
+only $\sim$half of the survey footprint was covered
+and the data were insufficiently complete 
 to support a homogeneously calibrated source catalogue.
@@
 where a contemporaneous observation of both halves of a field pair
-is not available.
+is available.
 
@@
 to show some variation towards lowly reddened areas,
-where red objects at $r-i > 1$ are rare.
+where red objects at $(r-i) > 1$ are rare.
 The vast majority of red objects in the global sample
@@
 and has achieved our aim of bringing the
-accuracy down to the level of $\sigma=0.03$~mag.
+accuracy down to the level of 0.03~mag (rms).
 
@@
 the same maximum matching distance of 1 arcsec
-for similar reasons \citep{Hambly2008}.
+for similar reasons \citep{Hambly2008},
+and warn that applications which require absolute
+astrometry with sub-arcsecond accuracy must
+use these catalogues with caution.
 
@@
 of objects at early Galactic longitudes 
-where source blending can affect more than a quarter of the objects.
+where source blending can sometimes affect
+more than a quarter of the objects.
 The \emph{veryReliable} flag should hence
 only be used in applications which require very reliable photometry
-at the expense of completeness,
-which is often desired for spectroscopic target selection.
+at the expense of completeness.
 
@@
 for this subset. Sources across the magnitude range 
-$13 < r < 17$ are now consistent at the level of 0.03 instead of 0.05~mag.
+$13 < r < 17$ now show a scatter of 0.03 instead of 0.05~mag.
 The best repeatability is again reached at $r=14$,
@@
 Moreover, the large systematics at the bright and faint end
-have disappeared.
+disappear.
 
@@
 We conclude this paper by demonstrating how the unique
-$(r-i)$/$(r-\ha)$ colour-colour diagram offered by this catalogue
+$(r-\ha,\ r-i)$ colour-colour diagram offered by this catalogue
 can readily be used to
@@
 on intrinsic stellar colour and interstellar extinction \citep{Drew2008}. 
-Put differently, the main sequence in the $(r-i)$/$(r-\ha)$ diagram
+Put differently, the main sequence in the $(r-\ha,\ r-i)$ diagram
 runs in a direction that is different from the reddening vector,
@@
 allows objects to be picked out
-from the $(r-i)$/$(r-\ha)$ colour-colour diagram
+from the $(r-\ha,\ r-i)$ colour-colour diagram
 using model-based colour criteria
@@
 \citep{Georgelin1973,Mayer1973,Hunter1990}.
-The ionising star has been placed 
+This ionising star has been placed 
 at a likely distance of 1.5-1.7 kpc
@@
 a mosaic of the same region 
-as seen in the mid-infrared by the Spitzer Space Telescope's
-Galactic Plane survey \citep[GLIMPSE;][]{Benjamin2003,Churchwell2009}.
+as seen in the mid-infrared by the Spitzer Space Telescope \citep[GLIMPSE survey;][]{Benjamin2003,Churchwell2009}.
 The infrared image reveals a bubble-shaped structure
@@
 and the ionising radiation from early-type stars \citep{Churchwell2006}.
-In this example, \cite{Yu2012}
+\cite{Yu2012}
 recently noted that the warm dust
@@
   \includegraphics[width=0.45\textwidth]{./plots/sh2-82-ccd.pdf}
-    \caption{$(r-i)$/$(r-\ha)$ diagram for the rectangular region of 
+    \caption{$(r-\ha,\ r-i)$ diagram for the rectangular region of 
     20-by-15 arcmin centred on the H{\sc ii} region Sh 2-82,
@@
 towards these regions,
-and can be used to discover optically-unveiled
-T~Tauri stars when the extinction is sufficiently low.
+and can be used to discover T~Tauri stars
+when the extinction is sufficiently low.
 
@@
 The catalogue is made available through the Vizier
-catalogue tool (http://vizier.u-strasbg.fr),
-where it is known as the ``IPHAS DR2 Source Catalogue''
-and where it can be queried
-(catalogue ID TBD; Conesearch/TAP URL TBD?).
+catalogue tool (http://vizier.u-strasbg.fr)
+where it can be queried
+(Conesearch/TAP URL TBD).
 The catalogue can also be downloaded in its entirety
@@
 Finally, we demonstrated the use of the survey's unique
-$(r-i)$/$(r-\ha)$ diagram for characterising stellar populations
+$(r-\ha,\ r-i)$ diagram for characterising stellar populations
 and selecting emission-line objects.
@@
 The data-taking strategy developed for IPHAS
-has since been reapplied to carry out a companion INT/WFC survey called UVEX
-in U/$g$/$r$/He {\sc i} \citep{Groot2009}
-and a southern survey in $u$/$g$/$r$/$i$/\ha\ 
+has since been reapplied to carry out 
+a companion INT/WFC Galactic Plane survey called UVEX
+in U, $g$, $r$, He {\sc i} \citep{Groot2009},
+a survey of the Kepler field 
+in $U$, $g$, $r$, $i$, \ha\ 
+\citep{Greiss2012},
+and a survey of the entire Southern Galactic Plane
+in $u$, $g$, $r$, $i$, \ha\ 
 called VPHAS+ \citep{Drew2014}.
@@
 into the contents and structure of the Milky Way
-across 360$^\circ$.
+across 360$^\circ$,
+and provide images across the most crowded
+regions to be faced by Gaia.
+
+%IPHAS also provide images for some of the most
+%crowded regions to be faced by Gaia.
 

### 2014-03-19 paper/iphas-dr2.tex
Adds citations to introduction + minor fixes

@@
 J. J. Drake$^{9}$,
-J. Fabregat$^{10}$, 
-B. T. Gaensicke$^{11}$, \newauthor
-A. S. Hales$^{12}$,
+J. Eisl\"offel$^{10}$,
+J. Fabregat$^{11}$, \newauthor
+B. T. Gaensicke$^{12}$,
+A. S. Hales$^{13}$,
 J. Irwin$^{9}$,
-C. Knigge$^{13}$,
+C. Knigge$^{14}$,
 T. Kupfer$^{5}$,
-D. J. Lennon$^{14}$, 
-J. R. Lewis$^{2}$, \newauthor
+D. J. Lennon$^{15}$, \newauthor
+J. R. Lewis$^{2}$, 
 M. Mohr-Smith$^{1}$, 
-R. A. H. Morris$^{15}$,
-T. Naylor$^{16}$,
-Q. A. Parker$^{17}$, 
-S. Phillipps$^{13}$, \newauthor
+R. A. H. Morris$^{16}$,
+T. Naylor$^{17}$,
+Q. A. Parker$^{18}$, \newauthor
+S. Phillipps$^{13}$, 
 R. Raddi$^{11}$, 
 P. Rodriguez-Gil$^{6}$,
-L. Sabin$^{18}$,
-S. Scaringi$^{19}$,
-D. Steeghs$^{11}$, 
-Y. C. Unruh$^{20}$, \newauthor
-K. Viironen$^{21}$, 
-J. S. Vink$^{22}$, 
+L. Sabin$^{19}$,
+S. Scaringi$^{20}$,
+D. Steeghs$^{11}$, \newauthor
+Y. C. Unruh$^{21}$, 
+K. Viironen$^{22}$, 
+J. S. Vink$^{23}$, 
 N. A. Walton$^{2}$,
 N. J. Wright$^{1}$,
-A. A. Zijlstra$^{23}$.
+A. A. Zijlstra$^{24}$.
 \\
@@
 Cambridge, MA 02138, U.S.A. \\
-$^{10}$Observatorio Astr\'onomico, Universidad de Valencia,
+$^{10}$Th\"uringer Landessternwarte, Sternwarte 5, 07778, Tautenburg, Germany \\
+$^{11}$Observatorio Astr\'onomico, Universidad de Valencia,
 Catedr\'atico Jos\'e Beltr\'an 2, 46980 Paterna, Spain\\
-$^{11}$Department of Physics, University of Warwick, Gibbet Hill Road, Coventry, CV4 7AL, U.K.\\
-$^{12}$Joint ALMA Observatory, Alonso de Córdova 3107, Vitacura 763-0355, Santiago, Chile. \\
-$^{13}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
-$^{14}$European Space Astronomy Centre (ESAC), Villafranca del Castillo, Villanueva de la Canada, E-28692 Madrid, Spain.\\
-$^{15}$School of Physics, Bristol University, Tyndall Avenue, Bristol, BS8 1TL, U.K.\\
-$^{16}$School of Physics, University of Exeter, Stocker Road, Exeter, EX4 4QL, U.K.\\
-$^{17}$Department of Physics \& Astronomy, Macquarie University, NSW 2109, Australia\\
-$^{18}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
-$^{19}$Instituut voor Sterrenkunde, K.U. Leuven, Celestijnenlaan 200D, B-3001 Leuven, Belgium\\
-$^{20}$Department of Physics, Blackett Laboratory, Imperial College
+$^{12}$Department of Physics, University of Warwick, Gibbet Hill Road, Coventry, CV4 7AL, U.K.\\
+$^{13}$Joint ALMA Observatory, Alonso de Córdova 3107, Vitacura 763-0355, Santiago, Chile. \\
+$^{14}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
+$^{15}$European Space Astronomy Centre (ESAC), Villafranca del Castillo, Villanueva de la Canada, E-28692 Madrid, Spain.\\
+$^{16}$School of Physics, Bristol University, Tyndall Avenue, Bristol, BS8 1TL, U.K.\\
+$^{17}$School of Physics, University of Exeter, Stocker Road, Exeter, EX4 4QL, U.K.\\
+$^{18}$Department of Physics \& Astronomy, Macquarie University, NSW 2109, Australia\\
+$^{19}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
+$^{20}$Instituut voor Sterrenkunde, K.U. Leuven, Celestijnenlaan 200D, B-3001 Leuven, Belgium\\
+$^{21}$Department of Physics, Blackett Laboratory, Imperial College
 London, Prince Consort Road, London, SW7 2AZ, U.K.\\
-$^{21}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
-$^{22}$Armagh Observatory, College Hill, Armagh, Northern Ireland,
+$^{22}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
+$^{23}$Armagh Observatory, College Hill, Armagh, Northern Ireland,
 BT61 9DG, U.K.\\
-$^{23}$Jodrell Bank Centre for Astrophysics, School of Physics \&
+$^{24}$Jodrell Bank Centre for Astrophysics, School of Physics \&
 Astronomy, University of Manchester, Manchester M13 9PL, U.K.}
@@
 catalogue's unique
-$(r-\ha,\ r-i)$ colour-colour diagram to
+$(r-\ha,\ r-i)$ diagram to
 (i) characterise stellar populations and extinction regimes
@@
 a preliminary catalogue of candidate emission line objects \citep{Witham2008};
-discoveries of new northern symbiotic stars \citep{Corradi2008, Corradi2010}; 
+discoveries of new symbiotic stars \citep{Corradi2008, Corradi2010, Corradi2011}; 
 new cataclysmic variables \citep{Witham2007}; 
-new groups of young stellar objects \citep{Vink2008,Barentsen2011a};
+new groups of young stellar objects
+\citep{Vink2008,Barentsen2011a,Wright2012};
+%new asymptotic giant branch stars \citep{Wright2008};
+new classical Be stars \citep{Raddi2013};
 along with discoveries of new and remarkable planetary nebulae 
-\citep{Mampaso2006, Corradi2011, Viironen2011}.
+\citep{Mampaso2006, Corradi2011, Viironen2011}
+and new supernova remnants \citep{Sabin2013}.
 
@@
 The survey originally aimed to obtain data 
-at seeing better than 1.7 arcsec.
+at seeing better than 1.7~arcsec.
 This target is currently attained across 86\% of the footprint,
 in particular at early longitudes,
-e.g. 92\% of the fields at $l<120^\circ$ are better than 1.7 arcsec.
+e.g. 92\% of the fields at $l<120^\circ$ are better than 1.7~arcsec.
 Figure~\ref{fig:seeing} presents the distribution
@@
 ensures that every pointing is immediately followed 
-by an offset pointing at a displacement of $+$5 arcmin in Dec 
-and $+$5 arcmin in RA.
+by an offset pointing at a displacement of $+$5~arcmin in Dec 
+and $+$5~arcmin in RA.
 This allows pairs of images to be checked 
@@
 obtained by cross-matching the detection tables of
-all three bands with a maximum matching distance of 1 arcsec.
+all three bands with a maximum matching distance of 1~arcsec.
 This was particularly effective for revealing
@@
 Finally, only exposures which are part of a sequence 
-of three consecutive images of the same field (H$\alpha$/$r$/$i$) 
+of three consecutive images of the same field
+(\ha, $r$, $i$)
 were considered for inclusion in the release. 
@@
 could not be obtained within the same night
-but for which the time gap did not exceed 48 hours.
-We note that the exact epoch of the magnitude in each band
-is included in the source catalogue
+but for which the time gap between exposures did not exceed 48 hours.
+We note that the source catalogue details the exact epoch
+at the start of each exposure
 (columns \emph{rMJD}, \emph{iMJD}, \emph{haMJD}).
@@
 and the Panoramic Survey Telescope 
-and Rapid Response System survey \citep[Pan-STARRS;][]{Schlafly2012},
+And Rapid Response System survey \citep[Pan-STARRS;][]{Schlafly2012},
 which all achieved photometry 
@@
 derived from PSF-fitting across the survey.
-%TODO: mention Gaia, e.g. IPHAS for transient identification?
 
@@
 regions to be faced by Gaia.
-
-%IPHAS also provide images for some of the most
-%crowded regions to be faced by Gaia.
 

### 2014-03-19 paper/iphas-dr2.tex
Incorporating feedback from V.

@@
 range of published studies including: 
-a preliminary catalogue of candidate emission line objects \citep{Witham2008};
+a preliminary catalogue of candidate emission-line objects \citep{Witham2008};
 discoveries of new symbiotic stars \citep{Corradi2008, Corradi2010, Corradi2011}; 
@@
 along with discoveries of new and remarkable planetary nebulae 
-\citep{Mampaso2006, Corradi2011, Viironen2011}
+\citep{Mampaso2006, Viironen2009a, Viironen2009b, Corradi2011, Viironen2011}
 and new supernova remnants \citep{Sabin2013}.
@@
 Over the years it has become apparent that the legacy of IPHAS 
-will extend beyond the traditional \ha\ applications 
+will also reach beyond these traditional \ha\ applications 
 of identifying emission-line stars and nebulae. 
@@
 in a new way.
-For example, the survey’s unique $(r-\ha)$ colour, 
-when combined with $(r-i)$,
-has been shown to provide simultaneous constraints 
+For example, the survey’s unique $(r-\ha,\ r-i)$ colour-colour
+diagram has been shown to provide simultaneous constraints 
 on intrinsic stellar colour and interstellar extinction \citep{Drew2008}. 

### 2014-04-03 paper/iphas-dr2.tex
Improved image keywords

@@
              Fields with a limiting magnitude brighter than
-             20th ($r$) or 19th (\ha/$i$) were rejected
+             20th ($r$) or 19th (\ha, $i$) were rejected
              from the data release.
@@
 of the mean seeing for all the quality-approved fields.
-We find a median value of 1.1~arcsec in $r$/\ha\
+We find a median value of 1.1~arcsec in $r$, \ha\
 and 1.0~arcsec in $i$.
@@
 The result of the procedure is a band-merged catalogue
-in which each row corresponds to a group of linked $r$/$i$/\ha\ detections
+in which each row corresponds to a group of linked $r$, $i$, and \ha\ detections
 which satisfy a maximum matching distance criterion in a pair-wise sense.
@@
 show systematic errors which exceed 0.5~arcsec.
-The position offsets between the $r$ and $i$/\ha\ detections
+The position offsets between the $r$ and $i$, \ha\ detections
 have been included in the catalogue 
@@
 
+Instructions for acknowledging IPHAS can be found on our website.
+
 \section{Conclusions and future work}
@@
 at the Institute of Astronomy in Cambridge.
+The catalogue presented in this work was assembled
+at the Centre for Astrophysics Research, University of Hertfordshire, supported by a grant from the Science \& Technology Facilities Council
+of the UK (STFC, ref ST/J001335/1).
 
@@
 
-GB and JED acknowledge the support of a grant
-from the Science \& Technology Facilities Council
-of the UK (STFC, ref ST/J001335/1).
+GB and JED acknowledge the support of the STFC grant ST/J001335/1.
 HJF is in receipt of an STFC postgraduate studentship.

### 2014-04-03 dr2/detections.py

@@ -4,15 +4,16 @@
 
 Summary
 -------
-This module is used to transform the detection tables that are produced
-by the ``imcore`` source detection and measurement tool of the Cambridge
-Astronomical Survey Unit (CASU). Its main purpose is to convert the columns
-in the CASU detection tables from instrumental units into more user-friendly
-astronomical units (e.g. converting fluxes into magnitudes, converting pixel
-coordinates into celestial coordinates, and adding user-friendly warning flags
-to signal quality problems.) Whilst doing so, the script will corrects for a
-range of known issues present in the data, e.g. it will apply a custom fix
-to the World Coordinate System (WCS) where necessary.
+This module is used to transform the IPHAS detection tables that are kindly
+provided by the ``imcore`` source detection and measurement tool of the
+Cambridge Astronomical Survey Unit (CASU). Its main purpose is to convert the
+columns in the CASU detection tables from instrumental units into more
+user-friendly astronomical units (e.g. converting fluxes into magnitudes,
+converting pixel coordinates into celestial coordinates, and adding
+user-friendly warning flags to signal quality problems.) Whilst doing so,
+the script will corrects for a range of known issues present in the data,
+e.g. it will apply a custom fix to the World Coordinate System (WCS) where
+necessary.
 This module can also produce a table called 'metadata.fits' which details the
 metadata of an entire directory of CASU detection tables, which is useful for
 carrying out quality control.

### 2014-04-03 dr2/images.py

@@ -1,13 +1,16 @@
 #!/usr/bin/env python
 # -*- coding: utf-8 -*-
-"""Prepare IPHAS images for data release.
+"""Prepares IPHAS images for public release.
 
 This script will edit all images in the data release to ensure they have
 the correct astrometric solution (WCS) and calibration information (PHOTZP)
-in the header.
+in the header, which follows the conventions from the "ESO External Data
+Products Standard" where possible
+(cf. http://www.eso.org/sci/observing/phase3/p3edpstd.pdf )
 
-We adopt the conventions from the "ESO External Data Products Standard"
-where possible (cf. http://www.eso.org/sci/observing/phase3/p3edpstd.pdf )
+Finally, the images are converted from Rice tile-compression into GZIP format,
+which is slightly larger and slower but somewhat less obscure and in line with
+many other archives.
 """
 from __future__ import division, print_function, unicode_literals
 from astropy.io import fits
@@ -47,16 +50,16 @@ METADATA = dict(zip(MD['run'], MD))
 class CalibrationDatabase(object):
     """Class to hold the calibration shifts."""
     def __init__(self):
-        r = ascii.read(os.path.join(constants.DESTINATION, 'calibration', 
+        r = ascii.read(os.path.join(constants.DESTINATION, 'calibration',
                                     'calibration-r.csv'))
-        i = ascii.read(os.path.join(constants.DESTINATION, 'calibration', 
+        i = ascii.read(os.path.join(constants.DESTINATION, 'calibration',
                                     'calibration-i.csv'))
-        ha = ascii.read(os.path.join(constants.DESTINATION, 'calibration', 
-                                    'calibration-ha.csv'))
+        ha = ascii.read(os.path.join(constants.DESTINATION, 'calibration',
+                                     'calibration-ha.csv'))
         self.shifts = dict(zip(
-                               np.concatenate((r['run'], i['run'], ha['run'])),
-                               np.concatenate((r['shift'], i['shift'], ha['shift']))
-                          ))
+                           np.concatenate((r['run'], i['run'], ha['run'])),
+                           np.concatenate((r['shift'], i['shift'], ha['shift']))
+                           ))
 
 
[…]

### 2014-04-03 dr2/sql/sql.py

@@ -8,43 +8,43 @@ from .. import constants
 
 
 # Numpy dtype for a row in the IPHAS DR2 full source catalogue
-mydtype = [('name', 'S19'), ('ra', '>f8'), ('dec', '>f8'), ('sourceID', 'S14'), 
-           ('posErr', '>f4'), ('l', '>f8'), ('b', '>f8'), 
-           ('mergedClass', '>i2'), ('mergedClassStat', '>f4'), ('pStar', '>f4'), 
-           ('pGalaxy', '>f4'), ('pNoise', '>f4'), ('pSaturated', '>f4'), 
-           ('rmi', '>f4'), ('rmha', '>f4'), 
-           ('r', '>f4'), ('rErr', '>f4'), 
-           ('rPeakMag', '>f4'), ('rPeakMagErr', '>f4'), 
-           ('rAperMag1', '>f4'), ('rAperMag1Err', '>f4'), 
-           ('rAperMag3', '>f4'), ('rAperMag3Err', '>f4'), 
-           ('rGauSig', '>f4'), ('rEll', '>f4'), ('rPA', '>f4'), 
-           ('rClass', '>i2'), ('rClassStat', '>f4'), 
-           ('rErrBits', '>i2'), ('rMJD', '>f8'), ('rSeeing', '>f4'), 
-           ('rDetectionID', 'S15'), ('rX', '>f4'), ('rY', '>f4'), 
-           ('i', '>f4'), ('iErr', '>f4'), 
-           ('iPeakMag', '>f4'), ('iPeakMagErr', '>f4'), 
-           ('iAperMag1', '>f4'), ('iAperMag1Err', '>f4'), 
-           ('iAperMag3', '>f4'), ('iAperMag3Err', '>f4'), 
-           ('iGauSig', '>f4'), ('iEll', '>f4'), ('iPA', '>f4'), 
-           ('iClass', '>i2'), ('iClassStat', '>f4'), 
-           ('iErrBits', '>i2'), ('iMJD', '>f8'), ('iSeeing', '>f4'), 
-           ('iDetectionID', 'S15'), ('iX', '>f4'), ('iY', '>f4'), 
-           ('iXi', '>f4'), ('iEta', '>f4'), 
-           ('ha', '>f4'), ('haErr', '>f4'), 
-           ('haPeakMag', '>f4'), ('haPeakMagErr', '>f4'), 
[…]

### 2014-04-03 pipeline.py

@@ -6,12 +6,12 @@ Summary
 -------
 This script drives the entire creation of the IPHAS DR2 source catalogue.
 Its purpose is to convert the IPHAS single-band detection tables, which are
-kindly created by the Cambridge Astronomical Surveys Unit (CASU),
+kindly provided by the Cambridge Astronomical Surveys Unit (CASU),
 into a band-merged and globally calibrated source catalogue which lists only
 the best-available measurement for each unique source.
 
-Dependencies
-------------
+Usage
+-----
 This script relies on the presence of a running instance of an IPython.parallel
 cluster, which can be started on a local machine using:
     
@@ -20,6 +20,12 @@ cluster, which can be started on a local machine using:
 or can be started on a PBS-controlled computing cluster by submitting the job:
 
    ``qsub scripts/pipeline-cluster.pbs``.
+
+Once an IPython.parallel cluster is running, you can simply run
+
+   ``python pipeline.py``
+
+...and the entire data release will be created within a few thousand CPU hours.
 """
 from IPython import parallel
 from astropy import log

### 2014-04-06 paper/iphas-dr2.tex
Update iphas-dr2.tex
Could you please add the following acknowledgment for me and Boris?

The research leading to these results has received funding from the
European Research Council under the European Union's Seventh Framework
Programme (FP/2007-2013) / ERC Grant Agreement n. 320964 (WDTracer).
BTG was supported in part by the UK’s Science and Technology
Facilities Council (ST/I001719/1)."

@@
 0.03~mag (rms) by validating against SDSS,
+% RR: better Sloan Digital Sky Survey (SDSS)?
 and provide a series of recommended quality criteria
@@
 at the Isaac Newton Telescope (INT) in La Palma -- 
+% RR: The INT acronym is defined before
 was to provide the digital update 
@@
 the limited bright samples of Galactic emission line objects 
-available at the outset \citep[e.g.][]{Kohoutek1999}, 
-to be extended into larger, deeper, more statistically-robust samples.
+available at the outset \citep[e.g.][]{Kohoutek1999} 
+to be extended into larger, deeper, and more statistically-robust samples.
 This in turn could better inform our understanding 
@@
 To set the stage for this release, we briefly remind of some key points.
-IPHAS is a 1800~sq. deg. imaging survey of the northern Galactic Plane, 
+IPHAS is a 1800~deg$^{2}$. imaging survey of the northern Galactic Plane, 
 providing images and photometry in Sloan $r, i$ 
 along with narrowband H$\alpha$. 
-It is carried out using the Wide Field Camera (WFC) 
-on the 2.5-metre Isaac Newton Telescope (INT) in La Palma. 
+It is carried out using the WFC 
+on the 2.5-metre INT in La Palma. 
 It is the first digital survey to offer comprehensive CCD photometry
@@
 and does so down to a limiting magnitude of $\sim$20th.
-The IPHAS footprint on the northern sky spans a box 
+The IPHAS footprint spans a box 
 of roughly 180 by 10 degrees, 
@@
 
-The Wide Field Camera is a mosaic of 4 CCDs 
-that captures a sky area of close to 0.29 square degrees.
+The WFC is a mosaic of 4 CCDs 
+that captures a sky area of close to 0.29~deg$^{2}$.
 To cover the entire Northern Plane with some overlap,
@@
 Each of these pointings is accompanied by an offset position
-at a displacement of $+$5 arcmin in declination 
+displaced by $+$5 arcmin in declination 
 and $+$5 arcmin in right ascension,
@@
 
-Figure~\ref{fig:footprint} shows the footprint 
+Fig.~\ref{fig:footprint} shows the footprint % Should replace all Figure with Fig., according to the MNRAS style I think
 of the quality-approved observations included in this work. 
@@
 The flux of each source is measured using both
-the peak pixel height (i.e. a square 0.33$\times$0.33\arcsec\ aperture)
+the peak pixel height (i.e. a square 0.33$\arcsec\times$0.33\arcsec\ aperture)
+% not sure if you need the \arcsec I added, but it would look nicer to me.
 as well as a series of circular apertures of increasing diameter 
@@
 by computing the sigma-clipped median
-flux in a grid of 64$\times$64 pixels (21$\times$21\arcsec)
+flux in a grid of 64$\times$64 pixels (21\arcsec$\times$21\arcsec)
 across the image,
@@
 Finally, an astrometric solution is determined
-based on the 2MASS point source catalogue \citep{Skrutskie2006},
+based on the Two-Micron All Sky Survey (2MASS) point source catalogue \citep{Skrutskie2006},
 which itself is calibrated 
@@
 that was obtained under seeing up to 2.5 arcsec,
-so that only 1\%of the pipelined exposures
+so that only 1\% of the pipelined exposures
 had to be excluded due to poor seeing.
@@
 The zeropoint was determined such that the resulting magnitude system
-is in the WFC system that uses the SED of Vega 
+is in the WFC system that uses the Spectral Energy Distribution (SED) of Vega 
 as the zero colour reference object. 
@@
 Notable past examples include the global re-calibration 
-of the Two Micron All Sky Survey \citep[2MASS;][]{Nikolaev2000},
-the Sloan Digital Sky Survey \citep[SDSS;][]{Padmanabhan2008}
+of the Two Micron All Sky Survey \citep[][]{Nikolaev2000},
+the Sloan Digital Sky Survey \citep[][]{Padmanabhan2008}
+% Define both 2MASS and SDSS before in the text
 and the Panoramic Survey Telescope 
@@
 \begin{table}
+% The caption of tables should be on top and not at the bottom
+        \caption{Mean magnitude offsets for objects
+                 cross-matched between IPHAS and APASS/SDSS
+                 before the uniform re-calibration.
+                 Eqns.~\ref{eqn:apass_r}-\ref{eqn:sdss_i} were applied
+                 to the APASS/SDSS magnitudes to bring them into the
+                 Vega-based IPHAS system prior to computing
+                 the offsets.)
+                 }
+        \label{tbl:offsets_before}
         \begin{center}
@@
                 \end{tabular}
-        \caption{Mean magnitude offsets for objects
-                 cross-matched between IPHAS and APASS/SDSS
-                 before the uniform re-calibration.
-                 Eqns.~\ref{eqn:apass_r}-\ref{eqn:sdss_i} were applied
-                 to the APASS/SDSS magnitudes to bring them into the
-                 Vega-based IPHAS system prior to computing
-                 the offsets.)
-                 }
-        \label{tbl:offsets_before}
-
+        \caption{Same as Table~\ref{tbl:offsets_before}
+                 but computed after the uniform re-calibration
+                 was carried out.}
+        \label{tbl:offsets_after}
                 \begin{tabular}{lcc}
@@
                 \end{tabular}
-        \caption{Same as Table~\ref{tbl:offsets_before}
-                 but computed after the uniform re-calibration
-                 was carried out.}
-        \label{tbl:offsets_after}
+
         \end{center}
@@
 sufficient coverage for 7359 of the fields in our data release.
-Fortunately, these fields are pre-dominantly located 
+Fortunately, these fields are predominantly located 
 in the early part of the Galactic Plane (Fig.~\ref{fig:apass_r}),
@@
 In addition, the complete set of detection tables -- one for each exposure -- 
-is made accessibly on our website to allow other uses of the data.
+is made accessible on our website to allow other uses of the data.
 A user-friendly catalogue of secondary detections 
@@
 \begin{table*}
+
 \vspace{2cm}
-
+\caption{Recommended minimum quality criteria 
+for selecting objects with reliable colours 
+from the IPHAS DR2 source catalogue. 
+86~million entries in the catalogue (39\%)
+satisfy all the criteria listed in this table.
+For convenience, these have been flagged in the catalogue
+using the column named \emph{reliable}.
 \begin{tabular}{p{8cm}lp{6cm}}
@@
 \end{tabular}
-\caption{Recommended minimum quality criteria 
-for selecting objects with reliable colours 
-from the IPHAS DR2 source catalogue. 
-86~million entries in the catalogue (39\%)
-satisfy all the criteria listed in this table.
-For convenience, these have been flagged in the catalogue
-using the column named \emph{reliable}.}
+}
 \label{tab:reliable}
@@
 \vspace{2cm}
-
+\caption{Additional quality criteria which are recommended
+for applications which require very reliable colours
+at the expense of completeness. 
+For convenience, the sources which satisfy the criteria listed
+in this table have been flagged in the catalogue
+using the column named \emph{veryReliable}.
 \begin{tabular}{p{8cm}lp{6cm}}
@@
 \end{tabular}
-\caption{Additional quality criteria which are recommended
-for applications which require very reliable colours
-at the expense of completeness. 
-For convenience, the sources which satisfy the criteria listed
-in this table have been flagged in the catalogue
-using the column named \emph{veryReliable}.}
+}
 \label{tab:veryreliable}
@@
 of Cygnus ($l\simeq 80$) and Aquila ($l\simeq 40$),
+% what's the low surface density seen at ~90 deg?
 which are regions known to be affected
@@
 runs in a direction that is different from the reddening vector,
-because the $r$-\ha\ colour tends to act
+because the ($r-\ha$) colour tends to act
 as a coarse proxy for spectral type
@@
 which is thought to originate from the
-mid-infrared emission of Poly-Aromatic Hydrocarbons (PAHs)
+mid-infrared emission of Polycyclic Aromatic Hydrocarbons (PAHs)
 -- i.e. warm dust --
@@
 \begin{table}
+    \caption{Candidate \ha-emitters towards Sh 2-82.}
+    \label{tbl:emitters}
 \footnotesize
@@
     \end{tabular}}
-    \caption{Candidate \ha-emitters towards Sh 2-82.}
-    \label{tbl:emitters}
 \end{table}
@@
 which is consistent with the objects
-being optically-unveiled Class II YSOs
+being optically-unveiled, or Class II, YSOs
+% add e reference for the YSO classification?
 with a mild infrared excess.

### 2014-04-06 paper/iphas-dr2.tex
Update iphas-dr2.tex
Could you please add the following acknowledgment for me and Boris?

The research leading to these results has received funding from the
European Research Council under the European Union's Seventh Framework
Programme (FP/2007-2013) / ERC Grant Agreement n. 320964 (WDTracer).
BTG was supported in part by the UK’s Science and Technology
Facilities Council (ST/I001719/1)."

@@
 $^{1}$School of Physics, Astronomy \& Mathematics, University of Hertfordshire, College Lane, Hatfield, Hertfordshire, AL10 9AB, U.K.\\
-$^{2}$Institute of Astronomy, Cambridge University, Madingley Road, Cambridge, CB3 OHA, U.K.\\
+$^{2}$Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge, CB3 OHA, U.K.\\
 $^{3}$IGAM, Institute of Physics, University of Graz, Universit\"atsplatz 5, Graz, Austria.\\
@@
 
-GB and JED acknowledge the support of the STFC grant ST/J001335/1.
+GB, JED and SES acknowledge support from the United Kingdom Science Technology and Facilities Council (STFC, JED and GB ST/J001333/1, SES ST/K00106X/1).
 HJF is in receipt of an STFC postgraduate studentship.

### 2014-04-08 paper/iphas-dr2.tex
Edited RRs comments

--- 2014-04-06 paper/iphas-dr2.tex
@@
 of the Northern Galactic Plane (IPHAS)
-is a 1800 deg$^2$ imaging survey
+is a 1800~deg$^2$ imaging survey
 covering the entire northern Milky Way at $|b| < 5^\circ$
@@
 We show that the global calibration is accurate to
-0.03~mag (rms) by validating against SDSS,
-% RR: better Sloan Digital Sky Survey (SDSS)?
+0.03~mag (rms) by validating against the Sloan Digital Sky Survey (SDSS),
 and provide a series of recommended quality criteria
@@
 and using more than 300 nights 
-at the Isaac Newton Telescope (INT) in La Palma -- 
-% RR: The INT acronym is defined before
+at the INT in La Palma --
 was to provide the digital update 
@@
 To set the stage for this release, we briefly remind of some key points.
-IPHAS is a 1800~deg$^{2}$. imaging survey of the northern Galactic Plane, 
+IPHAS is a 1800~deg$^2$. imaging survey of the northern Galactic Plane, 
 providing images and photometry in Sloan $r, i$ 
@@
 The WFC is a mosaic of 4 CCDs 
-that captures a sky area of close to 0.29~deg$^{2}$.
+that captures a sky area of close to 0.29~deg$^2$.
 To cover the entire Northern Plane with some overlap,
@@
 
-Fig.~\ref{fig:footprint} shows the footprint % Should replace all Figure with Fig., according to the MNRAS style I think
+Fig.~\ref{fig:footprint} shows the footprin
 of the quality-approved observations included in this work. 
@@
 the peak pixel height (i.e. a square 0.33$\arcsec\times$0.33\arcsec\ aperture)
-% not sure if you need the \arcsec I added, but it would look nicer to me.
 as well as a series of circular apertures of increasing diameter 
@@
 Most observations fared significantly better than these limits.
-Figure~\ref{fig:depth} presents the distribution of limiting magnitudes
+Fig.~\ref{fig:depth} presents the distribution of limiting magnitudes
 for all quality-approved fields,
@@
 e.g. 92\% of the fields at $l<120^\circ$ are better than 1.7~arcsec.
-Figure~\ref{fig:seeing} presents the distribution
+Fig.~\ref{fig:seeing} presents the distribution
 of the mean seeing for all the quality-approved fields.
@@
 Notable past examples include the global re-calibration 
-of the Two Micron All Sky Survey \citep[][]{Nikolaev2000},
-the Sloan Digital Sky Survey \citep[][]{Padmanabhan2008}
-% Define both 2MASS and SDSS before in the text
+of 2MASS \citep[][]{Nikolaev2000},
+SDSS \citep[][]{Padmanabhan2008}
 and the Panoramic Survey Telescope 
@@
 Since 2009,
-APASS has been using two 20\,cm-astrographs
+APASS has been using two 20~cm-astrographs
 to survey the entire sky down to $\sim$17th magnitude
@@
     \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_i.pdf} 
-    \caption{Same as Figure~\ref{fig:apass_r} for the $i$-band.}
+    \caption{Same as Fig.~\ref{fig:apass_r} for the $i$-band.}
         \label{fig:apass_i}
@@
     \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_i.pdf} 
-    \caption{Same as Figure~\ref{fig:sdss_i} for the $i$-band.}
+    \caption{Same as Fig.~\ref{fig:sdss_i} for the $i$-band.}
     \label{fig:sdss_i}
@@
 (iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
-which is located $\sim$500\,metres from the INT.
+which is located $\sim$500~m from the INT.
 
@@
 were taken at essentially the same time, 
-separated only by the $\sim$30\,sec read-out time required by the WFC.
+separated only by the $\sim$30~s read-out time required by the WFC.
 We have hence corrected the \ha\ zeropoints 
@@
 For convenience, these have been flagged in the catalogue
-using the column named \emph{reliable}.
+using the column named \emph{reliable}.}
 \begin{tabular}{p{8cm}lp{6cm}}
@@
 \end{tabular}
-}
 \label{tab:reliable}
@@
 in this table have been flagged in the catalogue
-using the column named \emph{veryReliable}.
+using the column named \emph{veryReliable}.}
 \begin{tabular}{p{8cm}lp{6cm}}
@@
 \end{tabular}
-}
+
 \label{tab:veryreliable}
@@
 
-Figure~\ref{fig:magdist} compares the $r$-band magnitude
+Fig.~\ref{fig:magdist} compares the $r$-band magnitude
 distribution for objects with and without the 
@@
 
-Figure~\ref{fig:uncertainties} shows the mean photometric
+Fig.~\ref{fig:uncertainties} shows the mean photometric
 uncertainty as a function of magnitude in each band.
@@
 secondary measurements which are present in the catalogue.
-Figure~\ref{fig:pairmag} shows the mean residuals between
+Fig.~\ref{fig:pairmag} shows the mean residuals between
 the primary and secondary magnitudes
@@
 
-In Figure~\ref{fig:pairmag_reliable} we show 
+In Fig.~\ref{fig:pairmag_reliable} we show 
 a similar comparison of the primary and secondary detections,
@@
 The mean source density as a function of Galactic longitude
-is shown in Figure~\ref{fig:density} (thick blue line).
+is shown in Fig.~\ref{fig:density} (thick blue line).
 The densities were computed by counting the 
@@
 For example,  we find a significant drop near the constellations 
-of Cygnus ($l\simeq 80$) and Aquila ($l\simeq 40$),
-% what's the low surface density seen at ~90 deg?
+of Aquila ($l\simeq40$) and Cygnus ($l\simeq80$ and $l\simeq90$),
 which are regions known to be affected
 by high levels of foreground extinction.
-Dark clouds are readily visible towards these constellation by eye,
-and they have often been referred to as ``the Great Rift''.
+In fact dark clouds are readily visible in the night sky
+towards these constellations,
+and they are often referred to as the `the Great Rift' \citep[e.g.][]{BokBok}.
 
@@
 
-In Figure~\ref{fig:density} we also shows the density
+In Fig.~\ref{fig:density} we also shows the density
 of non-blended sources (thin red line).
@@
 This is demonstrated 
-in Figures \ref{fig:l180}, \ref{fig:l45} \& \ref{fig:l30},
+in Fig.~\ref{fig:l180}, \ref{fig:l45} \& \ref{fig:l30},
 where we present three sets of IPHAS colour/magnitude diagrams
@@
 
-Figure~\ref{fig:mosaic_iphas} shows a 20-by-15 arcmin
+Fig.~\ref{fig:mosaic_iphas} shows a 20-by-15 arcmin
 colour mosaic centred on Sh 2-82,
@@
 and several dark cloud filaments.
-For comparison, Figure~\ref{fig:mosaic_spitzer} shows
+For comparison, Fig.~\ref{fig:mosaic_spitzer} shows
 a mosaic of the same region 
@@
 which have been selected from the colour-colour diagram
-in Figure~\ref{fig:emitters}. Note that the H{\sc ii} region is surrounded by a faint blue/green reflection nebula
+in Fig.~\ref{fig:emitters}. Note that the H{\sc ii} region is surrounded by a faint blue/green reflection nebula
 and dark cloud filaments.}
@@
     \caption{Star-forming region Sh 2-82 as seen in the mid-infrared
-    by the Spitzer Space Telescope's GLIMPSE survey. The mosaic is composed of the 24\,\micron\ (red), 8.0\,\micron\ (green) and 4.5\,\micron\ (blue) bands.
-    The image reveals a bubble-shaped structure which surrounds the {\sc Hii} region that is seen in the IPHAS mosaic of the same region (Figure~\ref{fig:mosaic_iphas}). 
+    by the Spitzer Space Telescope's GLIMPSE survey. The mosaic is composed of the 24~\micron\ (red), 8.0~\micron\ (green) and 4.5~\micron\ (blue) bands.
+    The image reveals a bubble-shaped structure which surrounds the {\sc Hii} region that is seen in the IPHAS mosaic of the same region (Fig.~\ref{fig:mosaic_iphas}). 
     This structure has previously been labelled as N115 in the 
@@
 
-Figure~\ref{fig:emitters} presents
+Fig.~\ref{fig:emitters} presents
 the IPHAS colour-colour diagram for 
@@
     20-by-15 arcmin centred on the H{\sc ii} region Sh 2-82,
-    which is the area shown in Figure~\ref{fig:mosaic_iphas}.    
+    which is the area shown in Fig.~\ref{fig:mosaic_iphas}.    
     The diagram shows all objects in the catalogue
@@
     while the main sequence for stars with an \ha\ emission line
-    strength of $-10\,\rm{\AA}$ EW is indicated by a dashed line,
+    strength of $-10~\rm{\AA}$ EW is indicated by a dashed line,
     following the colour simulations due to \citet{Barentsen2011a}.
@@
 we note that each of these four objects
-are detected in the Spitzer 8.0\,\micron\ image
+are detected in the Spitzer 8.0~\micron\ image
 at S/N$>$5,
 which is consistent with the objects
-being optically-unveiled, or Class II, YSOs
-% add e reference for the YSO classification?
-with a mild infrared excess.
+being optically-unveiled stars with a circumstellar disk,
+i.e. they are likely Class II YSOs with a mild infrared excess.
 
@@
 from the IPHAS website (www.iphas.org) as a collection 
-of binary FITS tables which comprise 50\,GB,
+of binary FITS tables which comprise 50~GB,
 together with a script

### 2014-04-08 paper/iphas-dr2.tex

@@
 
-GB, JED and SES acknowledge support from the United Kingdom Science Technology and Facilities Council (STFC, JED and GB ST/J001333/1, SES ST/K00106X/1).
+GB, JED, SES and BTG acknowledge support from STFC
+(JED and GB ST/J001333/1, SES ST/K00106X/1, BTG ST/I001719/1). 
 HJF is in receipt of an STFC postgraduate studentship.
+
+BTG acknowledges funding from the
+European Research Council under the European Union's Seventh Framework
+Programme (FP/2007-2013) / ERC Grant Agreement n. 320964 (WDTracer).
 

### 2014-04-09 paper/iphas-dr2.tex
Partial fixes for comments from YU

@@
 For such science applications to succeed however, 
-it is vital that the imaging data is transformed 
+it is vital that the imaging data are transformed 
 into a homogeneously calibrated photometric catalogue, 
@@
 by \citet{Drew2005} and \citet{Gonzalez-Solares2008}. 
-To set the stage for this release, we briefly remind of some key points.
+To set the stage for this release, we briefly recall some key points.
 IPHAS is a 1800~deg$^2$. imaging survey of the northern Galactic Plane, 
@@
 Sources are classified morphologically
--- stellar, galaxy or noise --
+-- stellar, extended or noise --
 based on the curve-of-growth determined
@@
 from revisiting regions in their footprint to 
-carry out a so-called \emph{ubercalibration} procedure,
+carry out a so-called \emph{ubercalibration}
+\footnote{The term ubercalibration refers to the name of the code
+first used to re-calibrate SDSS photometry using repeat observations,
+and is a term that has since been adopted by other surveys.
+The name is an anglicised version of the German word \"uberkalibration,
+which was reportedly chosen because authors Doug Finkbeiner and David Schlegel
+have German-sounding names (Finkbeiner 2010).} procedure,
 in which multiple measurements of stars at different epochs

### 2014-04-09 paper/iphas-dr2.tex
First set of fixes for comments from CK

@@
 to be targeted by spectroscopic surveys.
-For such science applications to succeed however, 
+For such science applications to succeed, however, 
 it is vital that the imaging data are transformed 
@@
 of point sources in the Galactic Plane at visible wavelengths, 
-and does so down to a limiting magnitude of $\sim$20th.
+and does so down to a limiting magnitude of $\sim$20.
 The IPHAS footprint spans a box 
@@
         Regions with incomplete data are apparent as white gaps (no data) 
-        or in light grey (denoting that only the offset position is missing).}
+        or in light grey (denoting that only the offset position is missing).
+        The dark vertical strip near $l \sim$125\deg
+        is an uninteresting consequence of the tiling pattern,
+        which overlaps in the middle of the survey area by design.}
         \label{fig:footprint}
@@
 Its application to IPHAS has previously been described
-by \citet{Drew2005} and \citet{Gonzalez-Solares2008}
+by \citet{Drew2005} and \citet{Gonzalez-Solares2008},
 and much of the source code is available 
@@
 (e.g. electronic noise or telescope tracking problems).
-Any unsuitable data that was taken as a result
+Any unsuitable data that were taken as a result
 of this strategy has subsequently been flagged and removed
@@
 (top panel in Fig.~\ref{fig:depth}).
-In contrast, the depth attained in $i$ and \ha\ 
-is less sensitive to moonlight
+In contrast, the depths attained in $i$ and \ha\ 
+are less sensitive to moonlight,
 and their magnitude limit distributions
@@
 and 1.0~arcsec in $i$.
-In the $r$-band, 90\% of the data is better than 1.5~arcsec
+In the $r$-band, 90\% of the data is better than 1.5~arcsec,
 and 10\% is better than 0.8~arcsec.
@@
 This ensures that the three bands for a given field
-are observed at the $\sim$same time --  
+are observed at roughly the same time --  
 essentially always within an elapsed time of 5 minutes.
@@
    m  = & \textsc{magzpt} - 2.5 \log_{10}( DN / \textsc{exptime} ) \\
- &  - \textsc{extinct}\cdot(\textsc{airmass}-1) - \textsc{apcor} - \textsc{percorr}
+ &  - \textsc{extinct}\cdot(\textsc{airmass}-1) - \textsc{apcor} - \textsc{percorr},
 \label{eqn:mag}
@@
 for the narrowband.
-We can specify here however 
+We can specify here, however, 
 that the integrated in-band energy flux for Vega 
@@
 are very scarce within the Galactic Plane.
-Secondly, the aim of IPHAS is to obtain magnitudes at a single epoch
+Secondly, the aim of IPHAS is to obtain magnitudes at a single epoch,
 and hence stars are not normally observed at more than one epoch,
@@
 which is significantly better 
-than the provisional calibration of IPHAS
+than the provisional calibration of IPHAS,
 for which we estimate the $1\sigma$-uncertainty to be $\sim$10\%.
@@
 and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$). 
-$w_{ij}$ are weights set to the uncertainty in $\Delta_{ij}$
+$w_{ij}$ are weights set to the inverse square of the uncertainty in $\Delta_{ij}$
 and $\theta_{ij}$ is an overlap function 
@@
 Solving for $a_i$ is equivalent to solving $\partial
-S/\partial a_i=0$ which gives the matrix equation:
+S/\partial a_i=0$, which gives the matrix equation:
 \begin{equation}
-   \sum_{j=1}^N A_{ij} a_j = b_j
+   \sum_{j=1}^N A_{ij} a_j = b_j,
    \label{eqn:matrix}
@@
 saturated stars based on the shape of the PSF,
-such extrapolation is prone to error
+such extrapolation is prone to error,
 and we do not recommend its use.
@@
 to such objects, the procedure is currently applied separately
-in each band and hence the $(r-i)$ and $(r-\ha)$ colours
+in each band, and hence the $(r-i)$ and $(r-\ha)$ colours
 may be inaccurate if the deblending proceeded differently in each band.
@@
 that 19\% of the sources in our catalogue
-are flagged as blended objects (column \emph{deblend})
+are flagged as blended objects (column \emph{deblend}),
 and their band-merged data should be treated with care
@@
 It is also clear that there are no unreddened stars
-earlier than $\sim$K0 visible,
+earlier than $\sim$K0 visible;
 such stars would be saturated if within a few hundred parsecs.
@@
 closer towards the galactic centre
-at $l=45^{\circ}$ (Fig.~\ref{fig:l45})
+at $l=45^{\circ}$ (Fig.~\ref{fig:l45}),
 we find a wealth of reddened objects.
-In the colour-magnitude diagram it is clear
+In the colour-magnitude diagram, it is clear
 that the stars are split into two distinct groups,

### 2014-04-10 paper/figures/sourcecount/plot.py
Initial commit of source count plot

@@ -1,0 +1,35 @@
+"""Plots IPHAS DR2 source counts as a function of longitude."""
+import matplotlib.pyplot as plt
+from matplotlib.ticker import MultipleLocator
+from astropy.table import Table
+
+t1 = Table.read('count-all.csv', format='ascii.csv')
+t2 = Table.read('count-not-deblend.csv', format='ascii.csv')
+
+mask = (t1['l'] >= 30) & (t1['l'] < 215)
+
+plt.figure(figsize=(7,2.5))
+plt.subplots_adjust(0.08, 0.17, 0.98, 0.92)
+
+plt.step(t1['l'][mask]+0.5, t1['count'][mask], 
+         linewidth=1.5, color='#2980b9',
+         label='All sources')
+plt.step(t2['l'][mask]+0.5, t2['count'][mask],
+         linewidth=1.0, color='#c0392b',
+         label='Non-blended sources')
+
+minorLocator = MultipleLocator(5.)
+plt.gca().xaxis.set_minor_locator(minorLocator)
+majorLocator = MultipleLocator(20.)
+plt.gca().xaxis.set_major_locator(majorLocator)
+#minorLocator = MultipleLocator(2.5)
+#plt.gca().yaxis.set_minor_locator(minorLocator)
+
+plt.gca().get_yaxis().get_major_formatter().set_powerlimits((0, 0))
+
+plt.legend(loc='upper left', fontsize=9)
+plt.xlabel('Galactic longitude ($l$)')
+plt.ylabel('Source count')
+plt.xlim([220, 20])
+plt.savefig('sourcecount.pdf')
+plt.close()

### 2014-04-10 paper/figures/sourcecount/query.sh

@@ -1,0 +1,5 @@
+PSQL="ionice -c3 nice -n19 psql -p 5433 -t -A -F',' -c"
+
+$PSQL "SELECT ROUND(l) AS lonbin, COUNT(*), AVG(r-i), AVG(r-ha) FROM iphas2 WHERE b BETWEEN -5 and +5 GROUP BY lonbin ORDER BY lonbin" > count-all.csv
+$PSQL "SELECT ROUND(l) AS lonbin, COUNT(*), AVG(r-i), AVG(r-ha) FROM iphas2 WHERE b BETWEEN -5 and +5 AND NOT deblend GROUP BY lonbin ORDER BY lonbin" > count-not-deblend.csv
+

### 2014-04-10 paper/iphas-dr2.tex

@@
 is a 1800~deg$^2$ imaging survey
-covering the entire northern Milky Way at $|b| < 5^\circ$
+covering the Galactic latitude range $|b| < 5^\circ$
 in the $r$, $i$ and \ha\ filters 
@@
 of the Northern Galactic Plane \citep[IPHAS;][]{Drew2005}
-is providing new insights into the contents and structure 
-of our own backyard, the Milky Way.
-The original motivation for undertaking 
-this large-scale programme of observation
+is providing new insights into the contents and structure of the disk of the Milky Way.
+This large-scale programme of observation
 -- spanning almost a decade 
@@
 at the INT in La Palma --
-was to provide the digital update 
+aims to provide the digital update 
 to the photographic northern H$\alpha$ surveys 
-of the mid-20th century. 
+of the mid-20th century \citep[see][]{Kohoutek1999}. 
 By increasing the sensitivity 
@@
 by a factor $\sim$1000 (7 magnitudes), 
-it was envisaged that IPHAS would allow 
+IPHAS can expand
 the limited bright samples of Galactic emission line objects 
-available at the outset \citep[e.g.][]{Kohoutek1999}
-to be extended into larger, deeper, and more statistically-robust samples.
-This in turn could better inform our understanding 
+previously available into larger, deeper, and more statistically-robust samples that will 
+better inform our understanding 
 of the early and late stages of stellar evolution.
@@
 catalogue to have been constructed from the imaging data.
-We explained the data reduction and quality control procedures that
-were applied,
-described and tested the global photometric calibration,
-and detailed the construction of the source catalogue.
+%We explain the data reduction and quality control procedures that
+%were applied.  We describe and test the global photometric 
+%calibration, 
+%and detailed the construction of the source catalogue.
 
 The observations included in this release
-achieved a median seeing of 1.1 arcsec
+achieve a median seeing of 1.1 arcsec
 and a $5\sigma-$depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$, \ha$=20.3\pm 0.3$.
 The global calibration and photometric repeatability
-were found to be accurate at the level of $0.03$ mag (rms),
+are found to be accurate at the level of $0.03$ mag (rms),
 providing a significant improvement over the 
 previous data release.
-The source catalogue provides the best-available
+The source catalogue specifies the best-available
 single-epoch astrometry and photometry
 for 219~million unique sources.
-We detailed a list of recommended quality criteria
-to select objects with reliable colours from the catalogue.
-Finally, we demonstrated the use of the survey's unique
-$(r-\ha,\ r-i)$ diagram for characterising stellar populations
-and selecting emission-line objects.
-
-In future work, we intend to draw upon the PanSTARRS photometric
-survey to further improve the accuracy of our global calibration.
-We also aim to correct the IPHAS photometry for the
-radial field distortions,
-plan to produce a catalogue 
-which details all the secondary detections
-to aid time-domain studies,
-and are investigating the use of photometry
-derived from PSF-fitting across the survey.
+To support its exploitation, we provide a list of recommended quality criteria
+that will permit the selection of objects with reliable colours from 
+the catalogue.  The closing demonstrations highlight the use of the 
+survey's unique $(r-\ha,\ r-i)$ diagram for characterising stellar populations
+and selecting emission-line objects.  More comprehensive applications of IPHAS
+can be found in the works of Sale et al (2014), which applies DR2 to the problem
+of 3D extinction mapping, and of Sabin et al (2014), where the results of a search
+of the image database for new planetary nebulae is presented.
+
+The current plan is to work toward one further major IPHAS source catalogue, in which 
+the remaining gaps in sky coverage will have been eliminated -- observations aimed at 
+replacing data not meeting the IPHAS minimum quality requirements continue.  We will also 
+examine options to further improve the global calibration, perhaps tightening the accuracy to
+better than 2\%.  For example, we will investigate the use of e.g. the PanSTARRS photometric 
+ladder (Magnier et al 2013) as a reference set, and explore improving source recovery in 
+the most dense fields via the implementation of PSF fitting in place of aperture photometry.  
+Finally, the next catalogue will detail all the secondary detections to aid time-domain studies.
 
@@
 \citep{Greiss2012},
-and a survey of the entire Southern Galactic Plane
+and a survey of the Southern Galactic Plane and Bulge
 in $u$, $g$, $r$, $i$, \ha\ 
-called VPHAS+ \citep{Drew2014}.
-We intend to use the present work
-as a template for generating catalogues
-from these companion surveys.
-When combined, these surveys
-will offer new insights
-into the contents and structure of the Milky Way
-across 360$^\circ$,
-and provide images across the most crowded
-regions to be faced by Gaia.
+called VPHAS+ \citep{Drew2014}.  The work presented here
+stands as a potential template for the catalogues that
+remain to be generated from these sibling surveys.  
+In prospect from them, whether they are mined separately or together,
+are the means to ask seamless questions on the contents and structure of 
+the most highly-populated components of the Milky Way.
 

### 2014-04-10 paper/iphas-dr2.tex
-

--- 2014-04-09 paper/iphas-dr2.tex
@@
 \centering
-\includegraphics[width=\textwidth]{./plots/pairmag/pairmag.pdf}
+\includegraphics[width=\textwidth]{figures/repeatability/pairmag.pdf}
 \caption{}
@@
 \centering
-\includegraphics[width=\textwidth]{./plots/pairmag/pairmag-reliable.pdf}
+\includegraphics[width=\textwidth]{figures/repeatability/pairmag-reliable.pdf}
 \caption{}
@@
          and for the \emph{veryReliable} sources alone (right panel).
-         The values shown are the mean residuals
+         The values shown are the mean absolute residuals
          between the primary and the secondary detections,
@@
 
-\subsection{Source densities and blending}
+\subsection{Source counts and blending}
 \label{sec:densities}
 
-The mean source density as a function of Galactic longitude
-is shown in Fig.~\ref{fig:density} (thick blue line).
-The densities were computed by counting the 
-number of sources in $1^\circ$-wide longitude bins,
-each covering the entire latitude range $-5^\circ<b<+5^\circ$.
-Unsurprisingly, we find the average source densities to increase
-towards the Galactic centre.
+The number of sources in our catalogue,
+counted in 1-by-10~deg$^2$ strips,
+is shown in Fig.~\ref{fig:sourcecount}
+as a function of Galactic longitude (thick blue line).	
+Unsurprisingly, we find the number of sources
+to increase towards the Galactic centre.
 For example, the average source density near $l\simeq 30^\circ$
-is roughly 30\,000 objects per square degree,
+is roughly 300\,000 objects per square degree,
 which is five times more than the density
@@
 In addition to the global trend,
-there are significant variations
-in the source density on smaller scales.
+variations are also apparent on smaller scales.
 For example, we find a significant drop near the constellations 
@@
 which are regions known to be affected
-by high levels of foreground extinction.
-In fact dark clouds are readily visible in the night sky
-towards these constellations,
-and they are often referred to as the `the Great Rift' \citep[e.g.][]{BokBok}.
-
-We warn that the relative densities reported here 
+by high levels of foreground extinction
+\citep[known as `the Great Rift', e.g.][]{BokBok}.
+
+We warn that the source counts reported here 
 have not been corrected for survey completeness
-or differences in the observing conditions across the survey fields.
+or variations in the depth of different survey fields.
 For example, the dip in the density near $l\simeq140^\circ$
@@
 we will calibrate the IPHAS-based source densities
-by injecting artificial stars into the IPHAS images
-and measuring their recovery rate (Farnhill et al., in preparation).
-IPHAS has the potential to offer
-calibrated, two-dimensional stellar density maps
-which can be used to constrain detailed models of our Galaxy.
+by injecting artificial stars into the images
+and measuring their recovery rates
+(Farnhill et al., in preparation).
+Such calibrated density maps may be used 
+to constrain detailed models of our Galaxy.
 
 \begin{figure*}
-    \includegraphics[width=\textwidth]{./plots/density/density.pdf} 
-    \caption{Mean number density of sources in the catalogue 
-    as a function of Galactic longitude, 
-    with and without blended sources included. 
-    The densities shown were computed by counting the sources 
-    at each longitude between $-5^\circ<b<+5^\circ$ (upper blue line).
-    We also show the densities based on only counting those sources 
+    \includegraphics[width=\textwidth]{figures/sourcecount/sourcecount.pdf} 
+    \caption{Number of entries in the IPHAS DR2 source catalogue
+    as a function of Galactic longitude.
+    The upper blue line shows the number of sources
+    counted in 1\deg-wide longitude bins.
+    The lower red line uses the same binning
+    but includes only those sources 
     for which the \emph{deblend} flag is {\sc false}, 
-    i.e. unconfused sources for which the CASU pipeline did not have to apply 
-    a deblending procedure (lower red line). 
+    i.e. unconfused sources for which the CASU pipeline
+    did not have to apply a deblending procedure.
+    In both cases only counted sources
+    in the latitude range $-5^\circ<b<+5^\circ$,
+    such that one may obtain approximate source densities
+    by dividing the counts by 10~deg$^2$.
+    The apparent variations in the source counts
+    traces the structure of the Galaxy
+    and the distribution of foreground extinction,
+    but also includes instrumental effects
+	such as variations in the survey depth
+	and completeness (see text).
    }
-    \label{fig:density}
+    \label{fig:sourcecount}
 \end{figure*}
 
-In Fig.~\ref{fig:density} we also shows the density
+In Fig.~\ref{fig:sourcecount} we also shows the number
 of non-blended sources (thin red line).
@@
 In particular, we warn that blended objects
-are likely candidates to have fallen victim
+are likely to have fallen victim
 to source confusion during the band-merging procedure.
@@
 in which sources are measured simultaneously across all bands,
-perhaps guided by an external astrometric catalogue such as the one
-that is to be provided by Gaia.
+which may be guided by an external astrometric catalogue
+such as the one that is to be provided by Gaia.
 
@@
 
-A new data release for the INT/WFC Photometric \ha\ Survey
-of the Northern Galactic Plane (IPHAS) was presented.
-It is the first survey to offer comprehensive CCD photometry
+A new catalogue of 219 million sources has been
+derived from the INT/WFC Photometric \ha\ Survey
+of the Northern Galactic Plane.
+It is the first catalogue to offer comprehensive CCD photometry
 of point sources across the Galactic Plane at visible wavelengths,
-taking in the entire Northern Milky Way at $|b|<5^\circ$.
+taking in the Northern Milky Way at $|b|<5^\circ$.
 The new 97-column catalogue provides single-epoch photometry
@@
 The observations included in this release
-achieved a median seeing of 1.1 arcsec
-and a $5\sigma-$depth of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$, \ha$=20.3\pm 0.3$.
+show a median seeing of 1.1 arcsec
+and $5\sigma-$depths of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$,
+and \ha$=20.3\pm 0.3$.
 The global calibration and photometric repeatability
@@
 survey to further improve the accuracy of our global calibration.
-We also aim to correct the IPHAS photometry for the
-radial field distortions,
-plan to produce a catalogue 
+We also plan to produce a catalogue 
 which details all the secondary detections

### 2014-04-10 paper/iphas-dr2.tex

@@
 catalogue to have been constructed from the imaging data.
-We explained the data reduction and quality control procedures that
-were applied,
-described and tested the global photometric calibration,
-and detailed the construction of the source catalogue.
+%We explain the data reduction and quality control procedures that
+%were applied.  We describe and test the global photometric 
+%calibration, 
+%and detailed the construction of the source catalogue.
 
 The observations included in this release
-show a median seeing of 1.1 arcsec
-and $5\sigma-$depths of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$,
-and \ha$=20.3\pm 0.3$.
+achieve a median seeing of 1.1 arcsec
+and $5\sigma-$depths of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$, and \ha$=20.3\pm 0.3$.
 The global calibration and photometric repeatability
-were found to be accurate at the level of $0.03$ mag (rms),
+are found to be accurate at the level of $0.03$ mag (rms),
 providing a significant improvement over the 
 previous data release.
-The source catalogue provides the best-available
+The source catalogue specifies the best-available
 single-epoch astrometry and photometry
 for 219~million unique sources.
-We detailed a list of recommended quality criteria
-to select objects with reliable colours from the catalogue.
-Finally, we demonstrated the use of the survey's unique
-$(r-\ha,\ r-i)$ diagram for characterising stellar populations
-and selecting emission-line objects.
-
-In future work, we intend to draw upon the PanSTARRS photometric
-survey to further improve the accuracy of our global calibration.
-We also plan to produce a catalogue 
-which details all the secondary detections
-to aid time-domain studies,
-and are investigating the use of photometry
-derived from PSF-fitting across the survey.
+To support its exploitation, we provide a list of recommended quality criteria
+that will permit the selection of objects with reliable colours from 
+the catalogue.  The closing demonstrations highlight the use of the 
+survey's unique $(r-\ha,\ r-i)$ diagram for characterising stellar populations
+and selecting emission-line objects.  More comprehensive applications of IPHAS
+can be found in the works of Sale et al (2014), which applies DR2 to the problem
+of 3D extinction mapping, and of Sabin et al (2014), where the results of a search
+of the image database for new planetary nebulae is presented.
+
+The current plan is to work toward one further major IPHAS source catalogue, in which 
+the remaining gaps in sky coverage will have been eliminated -- observations aimed at 
+replacing data not meeting the IPHAS minimum quality requirements continue.  We will also 
+examine options to further improve the global calibration, perhaps tightening the accuracy to
+better than 2\%.  For example, we will investigate the use of e.g. the PanSTARRS photometric 
+ladder (Magnier et al 2013) as a reference set, and explore improving source recovery in 
+the most dense fields via the implementation of PSF fitting in place of aperture photometry.  
+Finally, the next catalogue will detail all the secondary detections to aid time-domain studies.
 

### 2014-04-11 paper/figures/repeatability/data/query.sh
Check-in of repeatability plot

@@ -1,0 +1,8 @@
+PSQL="ionice -c3 nice -n19 psql -p 5433 -t -A -F',' -c"
+
+for band in r i ha
+do
+    echo "Now querying ${band}"
+    $PSQL "SELECT ROUND(${band}*2)/2 AS magbin, AVG(ABS(${band}-${band}2)), STDDEV(ABS(${band}-${band}2)) FROM iphas2 WHERE errBits < 16 AND errBits2 < 16 GROUP BY magbin ORDER BY magbin" > ${band}m${band}2_errbits16.csv
+    $PSQL "SELECT ROUND(${band}*2)/2 AS magbin, AVG(ABS(${band}-${band}2)), STDDEV(ABS(${band}-${band}2)) FROM iphas2 WHERE veryReliable AND errBits2 = 0 GROUP BY magbin ORDER BY magbin" > ${band}m${band}2_errbits0_veryreliable.csv
+done

### 2014-04-11 paper/figures/repeatability/plot.py

@@ -1,0 +1,78 @@
+"""Plots the mean Poissonion error as a function of magnitude."""
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.ticker import MultipleLocator
+from astropy.table import Table
+
+reliable = True
+
+if reliable:
+    version = "_errbits0_veryreliable.csv"
+else:
+    version = "_errbits16.csv"
+
+data_r = Table.read('data/rmr2'+version)
+data_i = Table.read('data/imi2'+version)
+data_ha = Table.read('data/hamha2'+version)
+
+xlim = [11.25, 22.75]
+ylim = [-0.04, 0.49]
+
+majorLocator = MultipleLocator(1.0)
+y_majorLocator = MultipleLocator(0.1)
+
+plt.figure(figsize=(3.5, 3.5))
+plt.subplots_adjust(0.17, 0.15, 0.95, 0.93, hspace=0, wspace=0)
+
+ax = plt.subplot(3, 1, 1)
+plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5, zorder=-5)
+plt.plot([10,30], [0.2,0.2], color='#666666', linewidth=0.5, zorder=-5)
+plt.errorbar(data_r['mag'], data_r['mean'], yerr=data_r['std'], linestyle='None', marker='o',
+             markersize=3.0, c='black', capsize=1.5, elinewidth=1.5)
+plt.xlim(xlim)
+plt.ylim(ylim)
+plt.ylabel('$|r-r_2|$')
+plt.gca().xaxis.set_ticklabels([])
+plt.gca().xaxis.set_major_locator(majorLocator)
+plt.gca().yaxis.set_major_locator(y_majorLocator)
+
+if reliable:
+    plt.text(0., 1.05, "(b) \emph{veryReliable} sources", fontsize=11,
+             ha='left', va='bottom', transform=ax.transAxes)
+else:
+    plt.text(0., 1.05, "(a) all sources", fontsize=11,
+             ha='left', va='bottom', transform=ax.transAxes)
+
+plt.subplot(3, 1, 2)
+plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5)
+plt.plot([10,30], [0.2,0.2], color='#666666', linewidth=0.5)
[…]

### 2014-04-11 paper/figures/uncertainties/data/query.sh
Check-in of uncertainties plot

@@ -1,0 +1,7 @@
+# Computes the mean Poissonian uncertainties as a function of magnitude
+PSQL="ionice -c3 nice -n19 psql -p 5433 -t -A -F',' -c"
+for band in r i ha
+do
+    echo "Now querying ${band}"
+    $PSQL "SELECT ROUND(${band}*2)/2 AS magbin, AVG(${band}Err), STDDEV(${band}Err) FROM iphas2 GROUP BY magbin ORDER BY magbin" > ${band}_uncertainties.csv
+done

### 2014-04-11 paper/figures/uncertainties/plot.py

@@ -1,0 +1,69 @@
+"""Plots the average Poissonian error as a function of magnitude."""
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.ticker import MultipleLocator
+from astropy.table import Table
+
+data_r = Table.read('data/r_uncertainties.csv')
+data_i = Table.read('data/i_uncertainties.csv')
+data_ha = Table.read('data/ha_uncertainties.csv')
+
+xlim = [11.25, 22.75]
+ylim = [-0.04, 0.49]
+
+majorLocator = MultipleLocator(1.0)
+minorLocator = MultipleLocator(0.5)
+y_minorLocator = MultipleLocator(0.05)
+y_majorLocator = MultipleLocator(0.1)
+
+plt.figure(figsize=(3.5, 3.5))
+plt.subplots_adjust(0.15, 0.15, 0.95, 0.95, hspace=0, wspace=0)
+
+plt.subplot(3, 1, 1)
+plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5)
+plt.plot([10,30], [0.2,0.2], color='#666666', linewidth=0.5)
+plt.errorbar(data_r['mag'], data_r['mean'], yerr=data_r['std'], linestyle='None', marker='o',
+             markersize=3.0, c='black', capsize=1.5, elinewidth=1.5)
+plt.xlim(xlim)
+plt.ylim(ylim)
+plt.ylabel('$\\overline{\\sigma_r}$')
+plt.gca().xaxis.set_ticklabels([])
+plt.gca().xaxis.set_major_locator(majorLocator)
+#plt.gca().xaxis.set_minor_locator(minorLocator)
+plt.gca().yaxis.set_major_locator(y_majorLocator)
+#plt.gca().yaxis.set_minor_locator(y_minorLocator)
+
+plt.subplot(3, 1, 2)
+plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5)
+plt.plot([10,30], [0.2,0.2], color='#666666', linewidth=0.5)
+plt.errorbar(data_i['mag'], data_i['mean'], yerr=data_i['std'], linestyle='None', marker='o',
+             markersize=3.0, c='black', capsize=1.5, elinewidth=1.5)
+plt.xlim(xlim)
+plt.ylim(ylim)
[…]

### 2014-04-11 paper/figures/repeatability/plot.py
removed overline

@@ -1,4 +1,5 @@
-"""Plots the mean Poissonion error as a function of magnitude."""
+"""Plots the mean absolute difference between magnitude pairs
+to gauge the level of systematics as a function of magnitude."""
 import os
 import numpy as np
 import matplotlib.pyplot as plt
@@ -71,8 +72,8 @@ plt.gca().yaxis.set_major_locator(y_majorLocator)
 plt.xlabel('Magnitude')
 
 if reliable:
-    plt.savefig('pairmag-reliable.pdf')
+    plt.savefig('repeatability-reliable.pdf')
 else:
-    plt.savefig('pairmag.pdf')
-    
+    plt.savefig('repeatability.pdf')
+
 plt.close()

### 2014-04-11 paper/figures/uncertainties/plot.py

@@ -13,8 +13,6 @@ xlim = [11.25, 22.75]
 ylim = [-0.04, 0.49]
 
 majorLocator = MultipleLocator(1.0)
-minorLocator = MultipleLocator(0.5)
-y_minorLocator = MultipleLocator(0.05)
 y_majorLocator = MultipleLocator(0.1)
 
 plt.figure(figsize=(3.5, 3.5))
@@ -27,12 +25,10 @@ plt.errorbar(data_r['mag'], data_r['mean'], yerr=data_r['std'], linestyle='None'
              markersize=3.0, c='black', capsize=1.5, elinewidth=1.5)
 plt.xlim(xlim)
 plt.ylim(ylim)
-plt.ylabel('$\\overline{\\sigma_r}$')
+plt.ylabel('$\\sigma_r$')
 plt.gca().xaxis.set_ticklabels([])
 plt.gca().xaxis.set_major_locator(majorLocator)
-#plt.gca().xaxis.set_minor_locator(minorLocator)
 plt.gca().yaxis.set_major_locator(y_majorLocator)
-#plt.gca().yaxis.set_minor_locator(y_minorLocator)
 
 plt.subplot(3, 1, 2)
 plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5)
@@ -41,12 +37,10 @@ plt.errorbar(data_i['mag'], data_i['mean'], yerr=data_i['std'], linestyle='None'
              markersize=3.0, c='black', capsize=1.5, elinewidth=1.5)
 plt.xlim(xlim)
 plt.ylim(ylim)
-plt.ylabel('$\\overline{\\sigma_i}$')
+plt.ylabel('$\\sigma_i$')
 plt.gca().xaxis.set_ticklabels([])
 plt.gca().xaxis.set_major_locator(majorLocator)
-#plt.gca().xaxis.set_minor_locator(minorLocator)
 plt.gca().yaxis.set_major_locator(y_majorLocator)
-#plt.gca().yaxis.set_minor_locator(y_minorLocator)
 
 plt.subplot(3, 1, 3)
 plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5)
@@ -55,13 +49,10 @@ plt.errorbar(data_ha['mag'], data_ha['mean'], yerr=data_ha['std'], linestyle='No
[…]

### 2014-04-11 paper/iphas-dr2.tex

@@
 uniformly-calibrated source catalogue
-to have been extracted from the images,
+derived from the survey,
 providing single-epoch photometry
 for 219 million unique sources
-across 92\% of the survey area.
+across 92\% of the 1800 square-degree footprint.
 The observations were carried out between 2003 and 2013
@@
 describe and test the new uniform calibration,
-and detail the construction of the new 97-column source catalogue.
+and detail the construction of the new 98-column source catalogue.
 We show that the global calibration is accurate to
-0.03~mag (rms) by validating against the Sloan Digital Sky Survey (SDSS),
+0.03~mag (rms) 
+%by validating against the Sloan Digital Sky Survey (SDSS),
 and provide a series of recommended quality criteria
@@
 and using more than 300 nights 
-at the INT in La Palma --
+at the Isaac Newton Telescope (INT) in La Palma --
 aims to provide the digital update 
@@
 Over the years it has become apparent that the legacy of IPHAS 
-will also reach beyond these traditional \ha\ applications 
+will reach beyond these traditional \ha\ applications 
 of identifying emission-line stars and nebulae. 
@@
 it is vital that the imaging data are transformed 
-into a homogeneously calibrated photometric catalogue, 
+into a uniformly-calibrated photometric catalogue, 
 in which quality problems 
@@
 When the initial data release was created in late 2007,
-only $\sim$half of the survey footprint was covered
+just over $\sim$half of the survey footprint was covered
 and the data were insufficiently complete 
@@
 The goal of this paper is to present the next release 
-that takes the coverage up to over 90 percent of the survey area 
+that takes the coverage up to 92 percent of the survey area 
 and includes a uniform calibration.
-In this work we aim to
+In this work we
 (i) explain the data reduction 
@@
 and quality control procedures.
-In \S\ref{sec:calibration} we explain the uniform re-calibration
+In \S\ref{sec:calibration} we explain the uniform re-calibration, in which
+we draw upon the AAVSO Photometric All-Sky Survey (APASS)
 and test our results against the Sloan Digital Sky Survey (SDSS).
@@
 Finally, in \S\ref{sec:dataaccess} we discuss access
-to the data and source code,
-and in \S\ref{sec:conclusions} we conclude and outline
-our future ambition
+to the catalogue, an online library of reduced images and relevant source code,
+The paper ends with conclusions in \S\ref{sec:conclusions} where we also outline
+our future ambitions.
 
@@
 by \citet{Drew2005} and \citet{Gonzalez-Solares2008}. 
-To set the stage for this release, we briefly recall some key points.
-IPHAS is a 1800~deg$^2$. imaging survey of the northern Galactic Plane, 
-providing images and photometry in Sloan $r, i$ 
-along with narrowband H$\alpha$. 
-It is carried out using the WFC 
+To set the stage for this release, we provide a reminder of some key points.
+IPHAS is an imaging survey of the Galactic Plane north of the celestial equator, 
+from which photometry in Sloan $r, i$ 
+is extracted along with narrowband H$\alpha$. 
+It is carried out using the Wide Field Camera (WFC) 
 on the 2.5-metre INT in La Palma. 
-It is the first digital survey to offer comprehensive CCD photometry
-of point sources in the Galactic Plane at visible wavelengths, 
+It is the first digital survey to offer comprehensive optical CCD photometry
+of point sources in the Galactic Plane, 
 and does so down to a limiting magnitude of $\sim$20.
@@
 of roughly 180 by 10 degrees, 
-taking in the entire northern Galactic Plane 
-at latitudes $-5^{\circ} < b < +5^{\circ}$ 
+spanning Galactic latitudes $-5^{\circ} < b < +5^{\circ}$ 
 and longitudes $30^{\circ} < l < 215^{\circ}$.
@@
 and to enable quality checks. 
-The basic unit of observation hence
+Hence, the basic unit of observation hence
 amounts to $2 \times 3$ exposures, 
 in which each of the 3 survey filters is exposed at 2 offset sky positions 
-within an elapsed time of 10 minutes.
+within, typically, an elapsed time of 10 minutes.
 We shall refer to the unit of 3 exposures at the same position 
@@
 and the combination of two fields at a small offset as a \emph{field pair}.
-The survey hence contains 15\,270 fields grouped into 7\,635 field pairs.
+Altogether, the survey contains 15\,270 fields grouped into 7\,635 field pairs.
 To achieve the desired survey depth of 20th magnitude or fainter, 
@@
 At that time only 76\% of the field pairs 
-satisfied our minimum quality criteria however,  
-often due to the effects of clouds, poor seeing, or technical faults
-(the quality criteria will be detailed in the next section). 
+satisfied our minimum quality criteria, however.  The problems affecting
+the 24\% falling below survey standard were, most commonly: variable cloud cover; 
+poor seeing; technical faults (the quality criteria will be detailed in the 
+next section). 
 Since then, a programme of repeat observations has been in place 
 to improve data quality. 
-As a result, 92\% of the survey 
+As a result, 92\% of the survey footprint
 now benefits from quality-approved data.
-The most recent observations which are part of this release
+The most recent observations included in this release
 were obtained in November 2012.
@@
         There are small overlaps between all the neighbouring fields
-        which can be seen as a honeycomb-style pattern
+        which can be seen as a honeycomb pattern
         of dark grey lines across the survey area.
         Regions with incomplete data are apparent as white gaps (no data) 
-        or in light grey (denoting that only the offset position is missing).
+        or in light grey (denoting that only one offset is missing).
         The dark vertical strip near $l \sim$125\deg
-        is an uninteresting consequence of the tiling pattern,
-        which overlaps in the middle of the survey area by design.}
+        is an arbitrary consequence of the tiling pattern,
+        which was populated starting from 0h in right ascension.}
         \label{fig:footprint}
@@
 at $l > 120^o$.
-Fields at these longitudes can only be accessed from La Palma 
+Fields at these longitudes are mainly accessed from La Palma 
 in the months of November-December,
-which is when the weather and seeing conditions are often poor
-at the INT and observing attempts have failed repeatedly.
+which is when the La Palma weather and seeing conditions are often poor
+forcing many (unsuccessful) repeat observations.
 To enable the survey to be brought to completion, 
@@
 
-All raw data obtained with the INT were transferred
+All raw IPHAS data were transferred
 to the Cambridge Astronomical Survey Unit (CASU) 
@@
 the peak pixel height (i.e. a square 0.33$\arcsec\times$0.33\arcsec\ aperture)
-as well as a series of circular apertures of increasing diameter 
+and a series of circular apertures of increasing diameter 
 (1.2\arcsec, 2.3\arcsec, 3.3\arcsec, 4.6\arcsec\ and 6.6\arcsec).
@@
 -- when required --
-a deblending routine is applied which attempts to remove
+a deblending routine is applied to remove
 the contamination from any nearby sources.
-Whilst this approach works very well 
-across the vast majority of the survey area,
-the Galactic Plane unavoidably contains crowded regions 
+This approach works very well 
+across the vast majority of the survey area.  Nevertheless,
+the Galactic Plane contains crowded regions 
 with large numbers of overlapping sources
 or rapidly spatially-varying nebulosity,
-in which case aperture photometry must always be interpreted 
-with caution.
-In \S\ref{sec:catalogue} we will explain that objects
-which suffer from this problem 
-are flagged in the catalogue using the \emph{deblend} warning flag.
+where aperture photometry will be compromised by frequent blending
+or poor background subtraction.
+In \S\ref{sec:catalogue} we will explain how affected sources are 
+flagged. 
+%we will explain that objects affected in this way 
+%are flagged in the catalogue using the \emph{deblend} warning flag.
 
@@
 based on the average zeropoint
-determined from a set of standard stars observed in the same night.
+determined from sets of standard stars observed within the same night.
 Sources are classified morphologically
@@
 at the sensitive default detection level of 1.25\,$\sigma$.
-This unavoidably includes spurious detections, artefacts and
+This total inevitably includes spurious detections, artefacts and
 duplicate detections,
@@
 on a semester-by-semester basis
-through the traditional time allocation committees 
+through the open time allocation committees 
 of the Isaac Newton Group of telescopes,
@@
 Any unsuitable data that were taken as a result
-of this strategy has subsequently been flagged and removed
-using a series of seven quality criteria,
+of this strategy have subsequently been flagged and removed
+using seven quality criteria,
 which ensure a reliable and homogeneous level of quality
@@
              The $r$-band depth is most sensitive 
-             to the presence of the moon above the horizon, 
-             which is evidenced by the wide and bi-model shape
+             to the presence of the moon above the horizon: 
+             this is the reason for the wide, bi-modal character
              of its distribution.}
@@
 Such data were typically obtained during poor weather or full moon.
-Most observations fared significantly better than these limits.
+Most observations were significantly better than these limits.
 Fig.~\ref{fig:depth} presents the distribution of limiting magnitudes
@@
 which was above the horizon during 62\% 
-of the observations and explains the wide and bi-modal shape
+of the observations.  The great range in sky brightness this
+produced is the reason for the wide and bi-modal shape
 of the $r$-band limiting magnitude distribution 
@@
 In contrast, the depths attained in $i$ and \ha\ 
-are less sensitive to moonlight,
-and their magnitude limit distributions
-are hence more narrow
+are less sensitive to moonlight, leading to
+narrower magnitude limit distributions
 (middle and bottom panels in Fig.~\ref{fig:depth}).
@@
 The minimum depth criterion
-has led us to exclude 9\% of the pipelined data,
-which were often obtained under cloudy conditions.
+led us to exclude 9\% of the pipelined data,
+which had typically been obtained under cloudy conditions.
 We note that some of the excluded data may nevertheless be useful
@@
 but it is slightly non-zero in any real telescope data 
-due to optical distortions and tracking errors.
-The mean ellipticity across a field in the IPHAS data set 
+due to optical distortions and tracking errors.  Indeed, it is worth noting
+that IPHAS data have been collected from unguided exposures that rely
+entirely on the INT's tracking capability.
+The mean ellipticity measured 
+%across a field 
+from exposures forming this release
+%in the IPHAS data set 
 is $0.09\pm0.04$.
@@
 due to mechanical glitches in the telescope tracking system.
-For this reason, we rejected exposures in which the mean ellipticity
-across the detector exceeded $e > 0.3$,
-which is when the photometric measurements delivered by the pipeline
-were found to become degraded.
+To exclude these, we rejected exposures in which the mean ellipticity
+across the detector exceeded $e > 0.3$. We have also found that, as
+this threshold is breached, the photometric measurements delivered by 
+the pipeline become degraded.
 Only 0.4\% of the exposures were discarded on this basis.
@@
 (3) \emph{Seeing.} 
-The survey originally aimed to obtain data 
+The original survey goal was to obtain data 
 at seeing better than 1.7~arcsec.
 This target is currently attained across 86\% of the footprint,
-in particular at early longitudes,
+in particular at lower Galactic longitudes,
 e.g. 92\% of the fields at $l<120^\circ$ are better than 1.7~arcsec.
@@
 and 10\% is better than 0.8~arcsec.
-To improve the area covered by this data release,
+To increase the sky area covered by this data release,
 we have decided to include the small fraction of data
-that was obtained under seeing up to 2.5 arcsec,
-so that only 1\% of the pipelined exposures
-had to be excluded due to poor seeing.
+obtained under seeing of up to 2.5 arcsec.  This limited
+exclusion on grounds of poor seeing to only 1\%\ of the pipelined 
+exposures. 
 In \S\ref{sec:catalogue} we will explain
@@
 (4) \emph{Photometric repeatability.} 
-The IPHAS field-pair observing strategy 
+The IPHAS field-pair observing strategy normally
 ensures that every pointing is immediately followed 
@@
 or more than 26\% were inconsistent at the level of 0.1 mag.
-These limits were determined empirically by inspecting
+These limits were set empirically after inspecting
 the images and photometry by eye.
@@
 6\% of the observations were placed on a \emph{black-list}
-as part of this procedure, most commonly due to the obvious presence
-of clouds or extreme levels of scattered moonlight.
+as a result of this -- most commonly because of the obvious 
+presence of clouds or extreme levels of scattered moonlight.
 
@@
 This ensures that the three bands for a given field
-are observed at roughly the same time --  
-essentially always within an elapsed time of 5 minutes.
+are observed contemporaneously --  
+typically within 5 minutes of each other.
 An exception was made for 9 fields where the three exposures 
@@
 
-The above criteria were satisfied by at least one observing attempt
+All of the above criteria were satisfied by at least one observing attempt
 for 14\,115 out of the 15\,270 fields (92\%).
+
 In some cases more than one successful attempt to observe
-a field was available due to the fact that stricter
-quality criteria were adopted in the initial years of the survey.
+a field was available due to stricter
+quality criteria being applied in the initial years of the survey.
 In such cases, only the attempt 
-with the best seeing and depth was selected
-for inclusion in the catalogue, because the focus 
-of this release is to deliver the most reliable
-measurement at a single epoch.
+with the best seeing and depth has been selected
+for inclusion in the catalogue, in order  
+to deliver the most reliable measurement at a single epoch.
 Those interested in any of the rejected data 
@@
 \textsc{extinct} is the atmospheric extinction coefficient 
-(typically 0.09 for $r$ and 0.05 for $i$ in La Palma),
+(set in the pipeline at 0.09 for $r$ and 0.05 for $i$ as representative
+averages for La Palma),
 \textsc{airmass} is the normalised optical path length 
@@
 
-The zeropoint was determined such that the resulting magnitude system
-is in the WFC system that uses the Spectral Energy Distribution (SED) of Vega 
-as the zero colour reference object. 
+The broadband zeropoints were determined such that the resulting magnitude system
+refers to the spectral energy distribution (SED) of Vega 
+as the zero colour object. 
 Colour equations were used to transform between the IPHAS passbands 
@@
 Unlike the broadbands, 
-standard star photometry is not available in the literature 
+standard-star photometry is not available in the literature 
 for the H$\alpha$ passband
 and hence there is no formally recognised flux scale 
-for the narrowband.
+for it.
 We can specify here, however, 
@@
 which is the flux obtained by folding 
-Vega's SED with the filter throughput curve 
-corrected for the atmosphere and detector quantum efficiency
-\citep[following the method explained by][]{Drew2005}.
+Vega's SED with the filter transmission curve 
+(uncorrected for the atmosphere and detector quantum efficiency,
+which would otherwise scale down the narrowband flux by 0.707).
 This is 3.14 magnitudes less than the flux captured 
@@
 \end{equation}
+Note that this implies that the in-band flux corresponding to
+zero magnitude is $1.56 \time 10^{-7}$ ergs\,cm$^{-2}$\,s$^{-1}$,
+when the \ha\ magnitude for Vega is set by convention to 0.03.
 

### 2014-04-11 paper/iphas-dr2.tex
-

--- 2014-04-10 paper/iphas-dr2.tex
@@
 \usepackage{subcaption}
+
+\restylefloat{table}
 
@@
 at a median seeing of 1.1 arcsec
-and to a mean $5\sigma$-depth of 
+and to mean $5\sigma$-depths of 
 21.2 ($r$), 20.0 ($i$) and 20.3 ($\ha$)
@@
 We show that the global calibration is accurate to
-0.03~mag (rms) by validating against the Sloan Digital Sky Survey (SDSS),
+0.03~mag (rms),
 and provide a series of recommended quality criteria
@@
 from revisiting regions in their footprint to 
-carry out a so-called \emph{ubercalibration}
-\footnote{The term ubercalibration refers to the name of the code
-first used to re-calibrate SDSS photometry using repeat observations,
-and is a term that has since been adopted by other surveys.
-The name is an anglicised version of the German word \"uberkalibration,
+carry out a so-called \emph{ubercalibration}\footnote{`Ubercalibration'
+refers to the name of the code used to re-calibrate SDSS photometry
+using repeat observations. 
+This name -- which has since been adopted by other surveys --
+is an anglicised version of the German word `\"uberkalibration',
 which was reportedly chosen because authors Doug Finkbeiner and David Schlegel
-have German-sounding names (Finkbeiner 2010).} procedure,
+both have German-sounding names (Finkbeiner 2010).} procedure,
 in which multiple measurements of stars at different epochs
@@
 \begin{figure}
-    \includegraphics[width=0.5\textwidth]{./plots/errors/uncertainties.pdf} 
+    \includegraphics[width=0.5\textwidth]{figures/uncertainties/uncertainties.pdf} 
     \caption{Mean photometric uncertainties
@@
              photon noise and hence this figure does not show
-             systematic or calibration uncertainties.}
+             the level of systematic uncertainties.}
     \label{fig:uncertainties}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/repeatability/pairmag.pdf}
+\includegraphics[width=\textwidth]{figures/repeatability/repeatability.pdf}
 \caption{}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/repeatability/pairmag-reliable.pdf}
+\includegraphics[width=\textwidth]{figures/repeatability/repeatability-reliable.pdf}
 \caption{}
@@
 the $\sim$10th magnitude star HD\,231616
-% Georgelin1973 claims B0III, others claim B0V
 with spectral type B0V/III
@@
 
+\newpage
 \appendix
+
+\section{Converting IPHAS magnitudes into fluxes}
+
+Mean photon wavelength of the passband following eqn.~A14 in Bessel \& Murphy (2012):
+\begin{equation}
+\lambda_0 = \frac{\int{\lambda S(\lambda) d \lambda}}
+{\int{S(\lambda)d\lambda}}
+\end{equation}
+
+Pivot wavelength of the passband following eqn.~A16 in Bessel \& Murphy (2012):
+\begin{equation}
+\lambda_p = \frac{\int{S(\lambda) \lambda d \lambda}}
+{\int{\frac{S(\lambda)}{\lambda} d\lambda}}
+\end{equation}
+
+Mean monochromatic flux,
+following eqn.~A11 in Bessel \& Murphy (2012):
+\begin{equation}
+f_\lambda = \frac{\int{f_\lambda(\lambda) S(\lambda) \lambda d \lambda}}
+{\int{S(\lambda)\lambda d \lambda}}
+\end{equation}
+where $S(\lambda)$ includes the filter, atmosphere and QE.
+
+\begin{table}[h]
+    \caption{Wavelength and mean monochromatic flux density for Vega in the IPHAS filter system.}
+    \label{tbl:flux}
+    \begin{tabular}{lrrrr}
+    \toprule
+     Band & $\lambda_0$ & $\lambda_p$ & EW  & $f_\lambda$   \\
+     & \AA & \AA & \AA & erg\,cm$^{-2}$\,s$^{-1}$\,\AA$^{-1}$  \\
+    \midrule
+$r$ &  6223  &   6211  &  785.6  &  2.47e-9  \\
+\ha &  6568  &   6568  &   59.6  &  1.81e-9  \\
+$i$ &  7674  &   7674  &  759.9  &  1.30e-9   \\
+    \bottomrule
+    \end{tabular}
+\end{table}
+
+\newpage
 \onecolumn

### 2014-04-11 paper/iphas-dr2.tex

--- 2014-04-11 paper/iphas-dr2.tex
@@
 0.03~mag (rms) 
-%by validating against the Sloan Digital Sky Survey (SDSS),
 and provide a series of recommended quality criteria

### 2014-04-11 paper/figures/repeatability/plot.py
Updated uncertainty plots

@@ -17,6 +17,11 @@ data_r = Table.read('data/rmr2'+version)
 data_i = Table.read('data/imi2'+version)
 data_ha = Table.read('data/hamha2'+version)
 
+rErr = Table.read('../uncertainties/data/r_uncertainties.csv')
+iErr = Table.read('../uncertainties/data/i_uncertainties.csv')
+haErr = Table.read('../uncertainties/data/ha_uncertainties.csv')
+
+
 xlim = [11.25, 22.75]
 ylim = [-0.04, 0.49]
 
@@ -27,8 +32,9 @@ plt.figure(figsize=(3.5, 3.5))
 plt.subplots_adjust(0.17, 0.15, 0.95, 0.93, hspace=0, wspace=0)
 
 ax = plt.subplot(3, 1, 1)
-plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5, zorder=-5)
-plt.plot([10,30], [0.2,0.2], color='#666666', linewidth=0.5, zorder=-5)
+plt.plot(rErr['mag'], rErr['mean'], zorder=-99, color='#ff6666')
+plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5, zorder=-100)
+plt.plot([10,30], [0.2,0.2], color='#666666', linewidth=0.5, zorder=-100)
 plt.errorbar(data_r['mag'], data_r['mean'], yerr=data_r['std'], linestyle='None', marker='o',
              markersize=3.0, c='black', capsize=1.5, elinewidth=1.5)
 plt.xlim(xlim)
@@ -46,8 +52,9 @@ else:
              ha='left', va='bottom', transform=ax.transAxes)
 
 plt.subplot(3, 1, 2)
-plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5)
-plt.plot([10,30], [0.2,0.2], color='#666666', linewidth=0.5)
+plt.plot(iErr['mag'], iErr['mean'], zorder=-99, color='#ff6666')
+plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5, zorder=-100)
+plt.plot([10,30], [0.2,0.2], color='#666666', linewidth=0.5, zorder=-100)
 plt.errorbar(data_i['mag'], data_i['mean'], yerr=data_i['std'], linestyle='None', marker='o',
[…]

### 2014-04-11 paper/figures/uncertainties/plot.py

@@ -19,8 +19,8 @@ plt.figure(figsize=(3.5, 3.5))
 plt.subplots_adjust(0.15, 0.15, 0.95, 0.95, hspace=0, wspace=0)
 
 plt.subplot(3, 1, 1)
-plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5)
-plt.plot([10,30], [0.2,0.2], color='#666666', linewidth=0.5)
+plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5, zorder=-100)
+plt.plot([10,30], [0.2,0.2], color='#666666', linewidth=0.5, zorder=-100)
 plt.errorbar(data_r['mag'], data_r['mean'], yerr=data_r['std'], linestyle='None', marker='o',
              markersize=3.0, c='black', capsize=1.5, elinewidth=1.5)
 plt.xlim(xlim)
@@ -31,8 +31,8 @@ plt.gca().xaxis.set_major_locator(majorLocator)
 plt.gca().yaxis.set_major_locator(y_majorLocator)
 
 plt.subplot(3, 1, 2)
-plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5)
-plt.plot([10,30], [0.2,0.2], color='#666666', linewidth=0.5)
+plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5, zorder=-100)
+plt.plot([10,30], [0.2,0.2], color='#666666', linewidth=0.5, zorder=-100)
 plt.errorbar(data_i['mag'], data_i['mean'], yerr=data_i['std'], linestyle='None', marker='o',
              markersize=3.0, c='black', capsize=1.5, elinewidth=1.5)
 plt.xlim(xlim)
@@ -43,8 +43,8 @@ plt.gca().xaxis.set_major_locator(majorLocator)
 plt.gca().yaxis.set_major_locator(y_majorLocator)
 
 plt.subplot(3, 1, 3)
-plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5)
-plt.plot([10,30], [0.2,0.2], color='#666666', linewidth=0.5)
+plt.plot([10,30], [0.1,0.1], linestyle='dashed', color='#666666', linewidth=0.5, zorder=-100)
[…]

### 2014-04-11 paper/iphas-dr2.tex
Update iphas-dr2.tex
The calibration section has now been gone through (catalogue next)

@@
 \end{equation}
-Note that this implies that the in-band flux corresponding to
+Note that the above implies that the in-band flux corresponding to
 zero magnitude is $1.56 \time 10^{-7}$ ergs\,cm$^{-2}$\,s$^{-1}$,
@@
 for two reasons. 
-Firstly, the survey was carried out 
-in competitive observing time
-on a non-dedicated telescope, 
+First, the survey was carried out 
+in competitively-allocated observing time on a
+common-user telescope, 
 rendering the 2MASS approach 
 of observing standards at a very high frequency
-prohibitively expensive
--- in part because standard fields 
-are very scarce within the Galactic Plane.
-Secondly, the aim of IPHAS is to obtain magnitudes at a single epoch,
-and hence stars are not normally observed at more than one epoch,
+prohibitively expensive (it does not help that 
+standard fields 
+are very scarce within the Galactic Plane).
+Second, IPHAS is not specified as a variability survey, with
+the consequence that stars are not normally observed at more 
+than one epoch,
 unless they happen to fall within a narrow overlap region 
-between two neighbouring fields.
-
-Although the IPHAS data does contain a significant number of 
+between two neighbouring field pairs.
+
+Although the IPHAS data do contain a significant number of 
 such inter-field repeat measurements,
@@
 Moreover, the illumination correction in the overlap regions
-is most strongly affected by a radial geometric distortion in the WFC,
-which causes the pixel scale towards the edges 
-to increase \citep{Gonzalez-Solares2011}.
+is more affected by a radial geometric distortion in the WFC,
+which causes the pixel scale to increase as the edges 
+are approached \citep{Gonzalez-Solares2011}.
 Although these systematics are reasonably small within a single field,
-they can combine to cause artificial zeropoint gradients 
-across the survey
-when they are used to constrain a global calibration
-without other external constraints.
+they can combine, during the global calibration routine, to cause artificial zeropoint gradients 
+across the survey unless controlled by
+%when they are used to constrain a global calibration
+%without 
+other external constraints.
 
@@
 to bring the majority of our data onto a uniform calibration.
-This is explained next.
+%This is explained next.
 
@@
 (APASS; http://www.aavso.org/apass)
-to bring the vast majority of the survey 
+to bring 
+%the vast majority - exaggerating a bit? 
+most of the survey 
 onto a uniform scale.
@@
 which is significantly better 
-than the provisional calibration of IPHAS,
+than the original nightly calibrations of IPHAS,
 for which we estimate the $1\sigma$-uncertainty to be $\sim$10\%.
@@
 we used the APASS catalogue to identify and adjust the calibration of all IPHAS fields 
-which showed a magnitude offset larger than 3\% against APASS.
-For this purpose,
+which showed a magnitude offset larger than 3\% against APASS.  Experience of re-running
+the calibration and testing the results showed us that it was inadvisable to more
+finely tune the match for IPHAS data obtained in what were generally the most photometric
+nights.  To this end,
 the $r$- and $i$-band detection tables of each IPHAS field
@@
 and to avoid sources near the faint detection limit of APASS 
-on the other hand.
+on the other.
 
@@
 The small colour terms in the equations
-indicate that the $r$ and $i$ filters 
-are very similar in both surveys.
+indicate that the IPHAS and APASS broadband filters 
+are very similar.
 The transformations include a large fixed offset,
@@
 APASS magnitudes are given in the AB system
-and IPHAS uses magnitudes relative to Vega.
+and IPHAS uses Vega magnitudes.
 Separate transformations were derived for sight-lines 
@@
 sight-lines were found to be insignificant.
-This is not surprising because heavily reddened objects 
+This is not surprising because heavily-reddened objects 
 are naturally scarce at $r<16$.
@@
 for each field which contained at least 30 cross-matched stars.
-This was the case for 48\% of our fields.
+This was achieved for 48\% of our fields.
 The mean offset was found to be
@@
 sufficient coverage for 7359 of the fields in our data release.
-Fortunately, these fields are predominantly located 
+Fortunately, these fields are located mainly 
 in the early part of the Galactic Plane (Fig.~\ref{fig:apass_r}),
@@
 been brought onto the same uniform scale 
-by employing an ubercalibration-style scheme, explained below,
-which minimised the magnitude offsets between stars
-located in the overlap regions with neighbouring fields.
-Although we explained earlier that these overlap regions
-are prone to systematics, the use of APASS enabled us to
-keep the zeropoints of roughly half of the fields fixed,
-which avoids these systematics from combining to introduce
-artificial gradients across the survey.
-In the next section we will show this to be true by validating
-our calibration against SDSS.
-
-A general solution to the problem of minimising the magnitude
-differences between overlapping frames has previously been
-described by~\citet{Glazebrook1994}.
+by employing an ubercalibration-style scheme,
+which minimises the magnitude offsets between stars
+located in the overlap regions between neighbouring fields.
+%Although we explained earlier that these overlap regions
+%are prone to systematics, the use of APASS enabled us to
+%keep the zeropoints of roughly half of the fields fixed,
+%which avoids these systematics from combining to introduce
+%artificial gradients across the survey.
+%In the next section we will show this to be true by validating
+%our calibration against SDSS.
+An algorithm for achieving this minimisation
+%of minimising the magnitude differences between overlapping frames 
+has previously been described by~\citet{Glazebrook1994}.
 In brief, there are two fundamental quantities to be
@@
 
-As explained above, we enforce a strong external constraint
+We enforce a strong external constraint
 on the solution by keeping the zeropoint fixed 
-for the 6\,756 fields which have been compared
-and calibrated against APASS earlier.
+for the 6\,756 fields which have already been compared
+and calibrated against APASS.
 We hereafter refer to these fields as \emph{anchors}.
-It is asserted that the zeropoints $a_i$ of these anchor fields 
-are known and not solved for,
-though they do appear in the vector $b_j$ as constraints.
-In addition to the APASS-based anchors, 
+It is asserted that the zeropoints $a_i$ of the anchor fields 
+are known and not solved for.  However they do appear in the vector 
+$b_j$ as constraints.  In addition to the APASS-based anchors, 
 we selected 3\,273 additional anchor fields by hand
@@
 We can reasonably assume however,
-that the corrections required for $r$ and \ha\ are identical
-because the IPHAS data-taking pattern ensured 
+that the corrections required for $r$ and \ha\ are identical,
+much of the time, because the IPHAS data-taking pattern ensured 
 that a field's \ha\ and $r$-band exposures
-were taken at essentially the same time, 
-separated only by the $\sim$30~s read-out time required by the WFC.
-We have hence corrected the \ha\ zeropoints 
+were taken consecutively -- 
+separated only by the $\sim$30~s read-out time.
+Hence, we have corrected the \ha\ zeropoints 
 by re-using the zeropoint adjustments that were derived for the $r$ band
@@
 suggesting strongly non-photometric conditions.
-For good practice, the \ha\ zeropoints of these fields
+In these cases, the \ha\ zeropoints of these fields
 were adjusted by solving Eqn.~\ref{eqn:matrix}
@@
 
-Having re-calibrated all fields to the expected APASS-accuracy of 3\%,
-we then used a different survey to validate the results.
-We have been able to exploit SDSS Data Release 9 \citep{Ahn2012}
-for this purpose.
-SDSS DR9 provides several strips at low
+Having re-calibrated all fields to the expected APASS accuracy of 3\%,
+we turned to a different survey, SDSS Data Release 9 \citep{Ahn2012},
+to validate the results.
+SDSS DR9 includes several strips at low
 Galactic latitudes,
@@
 with criteria {\sc type = star}, {\sc clean = 1} and {\sc score $>$ 0.7}.}
-in the same way as we did for APASS earlier,
-with the exception of using fainter magnitude ranges of 
+cross-matching in much the same way as for APASS,
+with the difference of selecting from the  fainter magnitude ranges of 
 $15<r_{\rm SDSS}<18.0$ and $14.5<i_{\rm SDSS}<17.5$.
@@
 which contained at least 30 objects with a cross-matched counterpart
-in SDSS.
-This was the case for 2\,602 fields.
+in SDSS.  This was the case for 2\,602 fields.
 The median offsets for each of these fields
@@
 exceeded 0.05~mag (523 fields) or even 0.1~mag (18 fields).
-Such sporadic outliers are consistent with the tails of a Gaussian
-with mean $\approx0$ and $\sigma=0.03$.
-Moreover, the SDSS calibration itself is not a perfect scale itself.
-
-In future work, we hope to draw upon
-the PanSTARRS survey \citep{Schlafly2012}
-to further improve the accuracy of our calibration.
-At the time of preparing this work data from PanSTARRS
-had not been made public yet.
+This pattern of outliers is consistent with the tails of a Gaussian
+with mean $\approx0$ and $\sigma=0.03$.  Furthermore, it should not
+be forgotten that both the SDSS and APASS calibrations are approximations to
+perfection and will not be entirely free of anomalies.  Indeed as we worked, 
+we noticed the occasional unsurprising examples of inconsistency between these 
+two surveys.  
+%, we hope to draw upon
+%the PanSTARRS survey \citep{Schlafly2012}
+%to further improve the accuracy of our calibration.
+%At the time of preparing this work data from PanSTARRS
+%had not been made public yet.
 

### 2014-04-12 paper/iphas-dr2.tex
Update iphas-dr2.tex
I've been through everything now, except the figure captions (which I probably will leave to you).  Note that I've numbered the Sh 2-82 candidate Ha emitters but don't know which two are the objects in common with Yu+Wang (over to you to resolve the 'x and y' now in the text).  In section 5, you'll find a couple of comments, that begin with GEERT that you need to address.  Also you now have a Parker05 reference to add to the bibliography.  ...still to do: expand section 8.

@@
 into a user-friendly catalogue.
-The aim of this catalogue is to detail
+The aim is to present
 the best-available information for each unique source
@@
 such as source blending and saturation.
-Compiling the catalogue essentially required four steps:
+Compiling the catalogue involved four steps:
 \begin{enumerate}
@@
 
-As the first step, the detection tables 
-were enhanced by creating new columns.
-This is necessary because the tables 
-generated by the CASU pipeline 
-summarise the detections 
-in their original CCD units,
-e.g. source positions are given in pixel coordinates 
-and photometry in number counts.
-To transform these measurements into
-user-friendly fields,
+Enhancement of the detection tables by creating new columns
+is the necessary first step because the tables generated by the CASU pipeline
+refer e.g. to source positions in pixel coordinates, to photometric measures 
+in number counts etc., rather than in directly useful astronomical terms.
+To transform these data into user-friendly quantities,
 we have largely adopted the units and naming conventions
@@
 and the VISTA Science Archive \citep[VSA;][]{Cross2012}.
-These archives curate the near-infrared data from both
+These archives curate the high-resolution near-infrared photometry from both
 the UKIDSS Galactic Plane Survey \citep[GPS;][]{Lucas2008}
@@
 VISTA Variables in the Via Lactea survey \cite[VVV;][]{Minniti2010}.
-Both these surveys provide high-resolution JHK photometry
-in the Galactic Plane.
 There is a significant degree of overlap
@@
 and hence by adopting a similar catalogue format
-we hope to encourage scientific applications
+we hope to facilitate scientific applications
 which combine both data sets.
@@
 
-Firstly, we note that each source is uniquely identified by an
+First, we note that each source is uniquely identified by an
 IAU-style designation of the form ``IPHAS2\ JHHMMSS.ss+DDMMSS.s''
@@
 the CCD number (\#ccd),
-and the row number in the CASU detection table (\#detection)
--- column names \emph{rDetectionID},
+and the row number in the CASU detection table (\#detection).
+These columns are named \emph{rDetectionID},
 \emph{iDetectionID}, \emph{haDetectionID}.
@@
 the catalogue also provides magnitudes
-using three alternative aperture sizes:
+using three alternatives:
 the peak pixel height 
@@
 for which the peak pixel height exceeds 55000 counts,
-which is typically the case for stars brighter than 12-13th magnitude.
-Although the pipeline attempts to extrapolate the brightness of
+which is typically the case for stars brighter than 12-13th magnitude in
+$r$.  Although the pipeline attempts to extrapolate the brightness of
 saturated stars based on the shape of the PSF,
 such extrapolation is prone to error,
-and we do not recommend its use.
+and shoud be viewed as indicative rather than as precise measurement.
+%and we do not recommend reliance on it.
 \item The \emph{deblend} column is used to flag sources 
@@
 \item The \emph{brightNeighb} column is used to flag sources which are located
-within 5 arcmin from an object brighter than $V=7$ 
+within a radius of 5 arcmin from an object brighter than $V=7$ 
 according to the Bright Star Catalogue (BSC; Hoffleit et al. 1991), 
 or within 10 arcmin if the neighbour is brighter than $V=4$.
-Such very bright stars are known to cause systematic errors
+These brightest stars are known to cause systematic errors
 and spurious detections as a result of stray light 
@@
 or which were otherwise affected by bad pixels in the detector.
-We will explain below that no such detections 
-have been included in the catalogue.
-An alternative detection was available in essentially all these situations
-as a result of the IPHAS field pair strategy,
-and hence it was not necessary to include these internal warning flags
-in the source catalogue.
+No such detections have had to be included in the catalogue, as 
+alternative detections were available in essentially all these situations
+thanks to the IPHAS field pair strategy. In turn, this has meant
+that it has not proved necessary to include these internal warning flags
+in the pubished source catalogue.
 
@@
 into multi-band field catalogues.
-This required a positional matching procedure 
-to link sources between the three bands
-based on their position on the sky.
+This required a position matching procedure 
+to link sources between the three bands.
 We used the \textsc{tmatchn} function 
@@
 We employed a maximum matching distance of 1~arcsec,
-which was chosen based on a trade-off between 
-completeness and reliability.
-On one hand, a matching distance larger than 1~arcsec 
+which was chosen, trading off completeness against reliability.
+On the one hand, a matching distance larger than 1~arcsec 
 was found to allow too many spurious and unrelated sources 
 to be linked. 
-On the other hand, a value smaller than 1~arcsec 
+On the other, a value smaller than 1~arcsec 
 would pose problems for very faint sources 
@@
 show systematic errors which exceed 0.5~arcsec.
-The position offsets between the $r$ and $i$, \ha\ detections
-have been included in the catalogue 
-and can hence be tightened by the user if necessary
+The position offsets between the $r$ detection and detections in $i$, and/or \ha\
+have been included in the catalogue, giving the user the option to tighten them 
+further if necessary
 (columns \emph{iXi}, \emph{iEta}, \emph{haXi}, \emph{haEta}).
@@
 We explained earlier that the survey contains repeat observations
-of identical sources as a result of overlaps in the data-taking pattern.
+of identical sources as a result of field offsetting and overlaps.
 Amongst all sources in the magnitude range $13<r<19$,
@@
 Only 9\% were detected once.
-Unsurprisingly, their spatial distribution traces
-the inter-CCD gaps and footprint edges.
-
-The principal aim of this data release is to provide 
+%Unsurprisingly, their spatial distribution traces
+%the inter-CCD gaps and footprint edges.
+
+Since the principal aim of this data release is to provide 
 reliable photometry at a single epoch,
-and hence we have decided
-to focus on providing the magnitudes
-and coordinates using only the \emph{best-available} 
+we have focused on providing the magnitudes
+and coordinates from the \emph{best-available} 
 detection of each object -- 
@@
 Secondly, co-adding would cause the image quality to degrade towards the mean,
-which is a draw-back for crowded fields.
+which is particularly a draw-back for crowded fields.
 
@@
 can nevertheless obtain this information in two ways.
-To begin with, whenever a secondary detection was observed 
+To begin with, whenever a secondary detection was collected 
 within 10 minutes of the primary,
@@
 is made accessible on our website to allow other uses of the data.
-A user-friendly catalogue of secondary detections 
-has not been compiled at present 
-but may be part of a future data release.
-
-The primary detection is defined as the entry in the 
-set of band-merged field catalogues which provides 
-the most reliable information for a unique source.
-Primary detections have been selected using a so-called \emph{seaming} procedure,
+%A user-friendly catalogue of secondary detections 
+%has not been compiled at present 
+%but may be part of a future data release.
+
+%The primary detection is defined as the entry in the 
+%set of band-merged field catalogues which provides 
+%the most reliable information for a unique source.
+Primary detections have been selected from all available detections using a so-called \emph{seaming} procedure,
 which we adapted from the algorithm developed for the WSA\footnote{http://surveys.roe.ac.uk/wsa/dboverview.html\#merge}.
-In brief, the first step of the procedure is to identify all the duplicate detections
+In brief, the first step is to identify all the duplicate detections
 by cross-matching the overlap regions of all field catalogues,
@@
 and (iii) the average seeing of stars in the CCD frame rounded to 0.2~arcsec.
-If this ranking scheme reveals multiple `winners' of identical quality,
+If this ranking scheme reveals multiple `winners' of seemingly identical quality,
 then the one that was observed closest to the optical axis of the camera is chosen.
@@
 and is made available on our website and through Vizier.
-The entire list of sources naturally includes 
+The original unweeded list of sources naturally included 
 a significant number of spurious entries
@@
 Half of the single-band detections were made in the $i$-band.
-This is likely explained by the fact that $i$ is least
+This is because $i$ is least
 affected by interstellar extinction and can occasionally pick up
-highly-reddened objects which are otherwise lost in $r$/\ha.
+highly-reddened objects which are otherwise lost in $r$ and \ha.
 
@@
 
-Having explained how the catalogue was created,
-we now offer an overview of its properties
-by discussing  
+%Having explained how the catalogue was created,
+We now offer an overview of the properties of the
+catalogue by discussing  
 (i) the recommended quality criteria,
@@
 are naturally prone to larger
-random and systematic uncertainties,
-for example, an inaccurately subtracted background
+random and systematic uncertainties:
+for example, an inaccurately-subtracted background
 will introduce a proportionally larger systematic error
@@
 quality criteria to be enforced for the purpose
-of removing faint and low-quality objects.
-
-The choice of quality criteria is a complicated
-trade-off between completeness on one hand
-and accuracy on the other.
+of removing lower-quality objects.
+
+The choice of quality criteria will always tension 
+completeness against accuracy.
 To aid users we have listed two sets of
@@
 
-Firstly, Table~\ref{tab:reliable} details
+Table~\ref{tab:reliable} specifies
 a set of minimum quality criteria
@@
 which require reliable $(r-i)$ and $(r-\ha)$ colours
-without removing more than 80\%
+(GEERT: the rest of the sentence looks very odd to me -- I can't rephrase
+because I'm not sure what you are trying to say) without removing more 
+than 80\%
 of the sources brighter than $r < 19$.
@@
 for identifying objects which are affected
-by inaccurate background subtraction
+by poor background subtraction
 or failed source deblending.
@@
 For convenience, the catalogue contains a boolean column
-named \emph{reliable} which flags these objects
-and makes their selection easy.
+named \emph{reliable} that directly flags these objects.
 
@@
 In \S\ref{sec:densities} we will show that these
-stricter criteria remove a lot
+stricter criteria remove large numbers
 of objects at early Galactic longitudes 
@@
 The \emph{veryReliable} flag should hence
-only be used in applications which require very reliable photometry
-at the expense of completeness.
+only be used in applications in which 
+completeness is not a concern.
 
@@
 from 0.1 mag to 0.2 mag,
-42 million additional candidate sources can be added.
+42 million candidate sources are added to the 109 
+million satisfying the tighter error bound (GEERT - is this what you mean?).
 
@@
 across all sources in the catalogue is
-$1.06\pm0.12$ for $(r-i)$ and $0.44\pm0.03$ for $(r-\ha)$.
-The better depth of $r$ is hence compensated
-by the fact that stars tend to have 
+$1.06\pm0.12$ for $(r-i)$ and $0.44\pm0.03$ for $(r-\ha)$. The
+fainter limiting magnitude in $r$ is hence compensated
+by the fact that most sources have 
 brighter magnitudes in both $i$ and \ha.
@@
 The number of sources in our catalogue,
-counted in 1-by-10~deg$^2$ strips,
+counted in 1-by-10~deg$^2$ constant-longitude strips,
 is shown in Fig.~\ref{fig:sourcecount}
@@
 is roughly 300\,000 objects per square degree,
-which is five times more than the density
+which is six times more than the density
 found near $l\simeq 180^\circ$.
@@
 In addition to the global trend,
-variations are also apparent on smaller scales.
+real variations are also apparent on smaller scales.
 For example, we find a significant drop near the constellations 
@@
 by high levels of foreground extinction
-\citep[known as `the Great Rift', e.g.][]{BokBok}.
-
-We warn that the source counts reported here 
-have not been corrected for survey completeness
-or variations in the depth of different survey fields.
+\citep[the extremities of `the Great Rift', e.g.][]{BokBok}.
+However, we warn that the source counts shown 
+have not been corrected for field pairs that have yet
+to be released or for variations in the depth across the included fields.
 For example, the dip in the density near $l\simeq140^\circ$
@@
 in the footprint coverage (seen in Fig.~\ref{fig:footprint}).
-In a forthcoming paper,
-we will calibrate the IPHAS-based source densities
-by injecting artificial stars into the images
-and measuring their recovery rates
-(Farnhill et al., in preparation).
-Such calibrated density maps may be used 
-to constrain detailed models of our Galaxy.
+
+In a forthcoming paper, properly calibrated maps of stellar density
+of the northern Galactic Plane will be presented (Farnhill
+et al., in preparation).  This will incorporate completeness
+corrections based on the statistics of artificial source
+recovery.  Such maps are of interest as tests of detailed models of 
+our Galaxy.
 
@@
 
-In Fig.~\ref{fig:sourcecount} we also shows the number
-of non-blended sources (thin red line).
+Fig.~\ref{fig:sourcecount} also shows the number counts
+for non-blended sources (thin red line).
 These are sources for which the \emph{deblend} flag is {\sc false},
@@
 originating from two or more overlapping objects.
-We find that the fraction of sources affected by blending problems
-is strongly correlated with the local source density.
-For example, only 11\% of the sources are blended
+This provides some insight into how the fraction of 
+blend-affected sources correlates with the local source 
+density.  In hard numbers, only 11\% of the sources are blended
 at $l>90^\circ$, whereas 24\% are blended at $l<90^\circ$.
 
-As we explained earlier, blended sources must be used with caution.
-In particular, we warn that blended objects
-are likely to have fallen victim
-to source confusion during the band-merging procedure.
-We have started investigating the use of
-more advanced PSF-fitting routines
-in which sources are measured simultaneously across all bands,
-which may be guided by an external astrometric catalogue
-such as the one that is to be provided by Gaia.
+%As we explained earlier, blended sources must be used with caution.
+Finally, we warn that blended objects
+are more likely than unblended objects to have fallen victim
+to source confusion during the band-merging procedure.  It is
+important to bear this in mind when appraising stars of seemingly
+unusual colours (such as candidate emission line stars) -- if 
+blending is flagged, the probability that the unusual colour is 
+spurious is greatly increased.
+%We have started investigating the use of
+%more advanced PSF-fitting routines
+%in which sources are measured simultaneously across all bands,
+%which may be guided by an external astrometric catalogue
+%such as the one that is to be provided by Gaia.
 
@@
 can readily be used to
-(i) characterise the extinction regime at different sight-lines, and
+(i) characterise the extinction regime along different sight-lines, and
 (ii) identify \ha\ emission-line objects.
@@
 Put differently, the main sequence in the $(r-\ha,\ r-i)$ diagram
-runs in a direction that is different from the reddening vector,
+runs in a direction that is at a large angle relative to the reddening vector,
 because the ($r-\ha$) colour tends to act
@@
 Each figure contains all the sources
-which are flagged as \emph{veryReliable}
-and are located in a region of one square degree 
+flagged as \emph{veryReliable} within a region of one square degree 
 centred on the coordinates indicated in the diagram
@@
 Each of the diagrams reveal a well-defined locus,
-which demonstrates the health of the catalogue and the global calibration
+which helps to demonstrate the health of the catalogue and the 
+global calibration
 for investigating stellar populations across wide areas.
@@
 together with the unreddened 1~Gyr isochrone due to \cite{Bressan2012},
-which is made available for the IPHAS system through a
+which is made available for the IPHAS system through the
 on line tool hosted at the Observatory of Padova
@@
 
-Each of the sight-lines reveal a stellar population
+Each of the sightlines reveal a stellar population
 with distinct characteristics.
-Firstly, towards the Galactic anti-centre 
+Towards the Galactic anti-centre 
 at $l=180^\circ$ (Fig.~\ref{fig:l180})
@@
 This implies that extinction is still increasing 
-at distances beyond $\sim$1-2~kpc,
-where M dwarfs are too faint to be contained in the IPHAS catalogue.
+at distances of $\sim$1-2~kpc,
+where M dwarfs become too faint to be contained in the IPHAS catalogue.
 It is also clear that there are no unreddened stars
@@
 of the corresponding phase of stellar evolution,
-would only account for a small proportion of a volume limited sample.
+would only account for a small proportion of the more nearly volume 
+limited sample seen in the Anticentre direction.
 
@@
 at $l=45^{\circ}$ (Fig.~\ref{fig:l45}),
-we find a wealth of reddened objects.
+there is  a wealth of reddened objects.
 In the colour-magnitude diagram, it is clear
@@
 Meanwhile the redder group is principally composed of red giant stars.
-As these stars are intrinsically brighter,
-at some given apparent magnitude they will be substantially further away
-than their main sequence counterparts.
-Given that, along this sight-line,
-extinction continues to increase with distance,
-the red giants we observe will be subject to more reddening
-than a main sequence star exhibiting the same apparent magnitude.
-Therefore, the clear split between the two groups arises
-due to the combination of increasing extinction
-and the distinct absolute magnitude ranges
-of main sequence and giant branch stars.
+As these stars are intrinsically brighter, they will be substantially 
+further away than their main sequence counterparts at the same 
+apparent magnitude.  Given that extinction continues to increase 
+with distance, along this sightline, 
+the red giants we observe will be subject to appreciably more reddening
+than the main sequence stars, pushing them to $r - i \sim 1.5$.
 
@@
 in the colour-magnitude diagram of Fig.~\ref{fig:l30},
-though the red clump stars are manifested as a slight over density
-which sits roughly 0.4~mags redder than the A0V reddening track.
+although the red clump stars are manifest as a track of slight 
+over density sitting roughly 0.4~mags redder than the A0V reddening track.
 At $(l,b)=(45^{\circ}, +2^{\circ})$ the giant stars observed
@@
 at which we can observe reddened giant stars,
-extinction is continuing to rise
-and we still find ourselves well within the Galactic dust disc.
+extinction is continuing to rise within the Galactic mid-plane.
 It is also apparent that the $(r-i)$ width of both the M dwarfs
@@
 
-Here we have presented only a short qualitative description
-of the information obtainable
+These are just qualitative vignettes of the information obtainable
 from IPHAS colour-colour and colour-magnitude plots.
 A more rigorous quantitative analysis of the IPHAS catalogue
-can be undertaken to estimate both
-the stellar density distribution in the Milky Way \citep{Sale2010}
-and to create detailed three-dimensional maps
-of the extinction across several kpc \citep{Sale2009,Sale2012}.
-An extinction map based on our catalogue
-is to be released in a separate paper
-that accompanies this data release (Sale et al., in preparation).
+can be undertaken to estimate both the stellar density distribution 
+in the Milky Way \citep{Sale2010} and to create detailed three-dimensional 
+maps of the extinction across several kpc \citep{Sale2009,Sale2012}.
+A 3-D extinction map based on the DR2 catalogue
+is being released in a separate paper (Sale et al., in preparation).
 
@@
 
-A primary motivation for carrying out the survey 
-was to enable the discovery 
-of new emission-line objects
+An aim of IPHAS is to enable the discovery 
+of new fainter emission-line objects
 across the Galactic Plane.
@@
 In this section were merely aim to demonstrate
-how the catalogue may be used for this purpose.
+a use of the catalogue for this purpose.
 
@@
 as seen in the mid-infrared by the Spitzer Space Telescope \citep[GLIMPSE survey;][]{Benjamin2003,Churchwell2009}.
-The infrared image reveals a bubble-shaped structure
+The infrared image reveals an enclosing fuzzy bubble (appearing
+green in Fig.~\ref{fig:mosaic_spitzer})
 which is thought to originate from the
@@
 taken from the colour simulations due to \citet{Barentsen2011a}.
-Six stars are found to lie confidently above the 
+Six stars are found to lie above the 
 dashed line at the level of $3\sigma$,
@@
 are marked by red triangles in the colour-colour diagram
-and their photometry is detailed in Table~\ref{tbl:emitters}.
+and their photometry is set out in Table~\ref{tbl:emitters}.
 The spatial distribution of the six candidate emission-line objects
@@
 \resizebox{0.45\textwidth}{!}{
-    \begin{tabular}{lccc}
+    \begin{tabular}{clccc}
     \toprule
-    Name [IPHAS2 ...] & $r$ & $i$ & \ha  \\
+    # & Name [IPHAS2 ...] & $r$ & $i$ & \ha  \\
     \midrule
-J192954.40+181026.1& $17.69\pm0.01$ & $16.12\pm0.01$ & $16.19\pm0.01$ \\
-J193011.01+182051.2& $18.55\pm0.02$ & $16.95\pm0.02$ & $17.31\pm0.02$ \\
-J193021.52+181954.5& $19.72\pm0.05$ & $17.94\pm0.03$ & $18.47\pm0.04$ \\
-J193024.45+181938.3& $19.31\pm0.04$ & $17.55\pm0.02$ & $17.99\pm0.03$ \\
-J193033.00+181609.3& $18.25\pm0.01$ & $16.91\pm0.01$ & $16.92\pm0.01$ \\
-J193042.48+182317.4& $19.96\pm0.03$ & $18.11\pm0.03$ & $18.48\pm0.03$ \\
+1 & J192954.40+181026.1& $17.69\pm0.01$ & $16.12\pm0.01$ & $16.19\pm0.01$ \\
+2 & J193011.01+182051.2& $18.55\pm0.02$ & $16.95\pm0.02$ & $17.31\pm0.02$ \\
+3 & J193021.52+181954.5& $19.72\pm0.05$ & $17.94\pm0.03$ & $18.47\pm0.04$ \\
+4 & J193024.45+181938.3& $19.31\pm0.04$ & $17.55\pm0.02$ & $17.99\pm0.03$ \\
+5 & J193033.00+181609.3& $18.25\pm0.01$ & $16.91\pm0.01$ & $16.92\pm0.01$ \\
+6 & J193042.48+182317.4& $19.96\pm0.03$ & $18.11\pm0.03$ & $18.48\pm0.03$ \\
     \bottomrule
@@
 we discovered towards Sh 2-82 as well.
-Two of our candidates have recently been identified
+Two of our candidates, x and y in Tab.~\ref{tab:emitters}, have recently 
+been identified
 as YSOs in the initial investigation of the region by \cite{Yu2012}.
@@
 as having an infrared excess,
-we note that each of these four objects
+we note that all four objects
 are detected in the Spitzer 8.0~\micron\ image
 at S/N$>$5,
-which is consistent with the objects
+which is consistent with their
 being optically-unveiled stars with a circumstellar disk,
-i.e. they are likely Class II YSOs with a mild infrared excess.
+i.e. they are likely to be Class II YSOs exhibiting a mild infrared excess.
 
@@
 and can be used to discover T~Tauri stars
-when the extinction is sufficiently low.
+when the extinction is not too high ($A_V < 10$).
 
@@
 from the IPHAS website (www.iphas.org) as a collection 
-of binary FITS tables which comprise 50~GB,
+of binary FITS tables occupying 50~GB,
 together with a script
 to ingest these tables into a PostgreSQL database.
-Finally, the website also provides access to the pipeline-processed
+
+The website also provides access to the pipeline-processed
 imaging data, which we have updated to include
@@
 
-\section{Conclusions and future work}
+\section{Concluding remarks and future work}
 \label{sec:conclusions}
@@
 It is the first catalogue to offer comprehensive CCD photometry
-of point sources across the Galactic Plane at visible wavelengths,
-taking in the Northern Milky Way at $|b|<5^\circ$.
-The new 97-column catalogue provides single-epoch photometry
+of point sources across the Northern Galactic Plane at visible wavelengths,
+taking in the latitude range $|b|<5^\circ$.
+The new 98-column catalogue provides single-epoch photometry
 across 92\% of the survey area,
 and is the first quality-controlled and uniformly calibrated
-catalogue to have been constructed from the imaging data.
+catalogue to have been constructed from the imaging data.  This
+now means that there is H$\alpha$ coverage, accessible online,
+of the entire Galactic Plane -- given that the 
+southern Plane is already available thanks to the UK
+Schmidt H$\alpha$ Survey \citep[SHS,][][Parker05], the last of the 
+photographic surveys carried out by that telescope.
 %We explain the data reduction and quality control procedures that
@@
 the remaining gaps in sky coverage will have been eliminated -- observations aimed at 
-replacing data not meeting the IPHAS minimum quality requirements continue.  We will also 
+replacing data not meeting the IPHAS minimum quality requirements are continuing.  We will also 
 examine options to further improve the global calibration, perhaps tightening the accuracy to
-better than 2\%.  For example, we will investigate the use of e.g. the PanSTARRS photometric 
-ladder (Magnier et al 2013) as a reference set, and explore improving source recovery in 
+better than 2\%.  For example, we have in mind investigating the use of the PanSTARRS photometric 
+ladder (Magnier et al 2013) as a reference set, when it becomes available and we will 
+explore improving source recovery in 
 the most dense fields via the implementation of PSF fitting in place of aperture photometry.  
-Finally, the next catalogue will detail all the secondary detections to aid time-domain studies.
+Finally, the next catalogue will detail all the secondary detections to aid time-domain 
+studies.
 
@@
 in $u$, $g$, $r$, $i$, \ha\ 
-called VPHAS+ \citep{Drew2014}.  The work presented here
+called VPHAS+ \citep{Drew2014}.  The last of these incorporates the
+digital update of the SHS, offering all the advantages of calibrated
+photometry across a little over half the SHS footprint.  
+The work presented here
 stands as a potential template for the catalogues that
@@
 
-GB, JED, SES and BTG acknowledge support from STFC
-(JED and GB ST/J001333/1, SES ST/K00106X/1, BTG ST/I001719/1). 
-HJF is in receipt of an STFC postgraduate studentship.
-
-BTG acknowledges funding from the
+GB, JED, SES and BTG acknowledge support from the Science \& Technology
+Facilities Council of the United Kingdom
+(grants: GB and JED ST/J001333/1, SES ST/K00106X/1, BTG ST/I001719/1). 
+HJF is in receipt of an STFC postgraduate studentship.  JED would also
+like to convey her thanks to the Physics Department of Imperial College
+London that hosted this project from its inception to 2007 and supported
+her via a sabbatical year in 2003-4.  BTG acknowledges funding from the
 European Research Council under the European Union's Seventh Framework
 Programme (FP/2007-2013) / ERC Grant Agreement n. 320964 (WDTracer).
+NJW is in receipt of a Fellowship funded by the Royal Astronomical Society
+of the United Kingdom.
 

### 2014-04-13 paper/tables/columns.tex
Number the columns

@@
-name & string &  & Sexagesimal, equatorial position-based source name in the form: JHHMMSS.ss+DDMMSS.s.
+1 & name & string &  & Sexagesimal, equatorial position-based source name in the form: JHHMMSS.ss+DDMMSS.s.
 The full naming convention for IPHAS DR2 sources has the form "IPHAS2 JHHMMSS.ss+DDMMSS.s", where "J" indicates the position is J2000.
@@
 \\
-ra & double & degrees & J2000 Right Ascension with respect to the 2MASS PSC reference frame (which is consistent with ICRS to within 0.1 arcsec).
+2 & ra & double & degrees & J2000 Right Ascension with respect to the 2MASS PSC reference frame (which is consistent with ICRS to within 0.1 arcsec).
 The coordinate given is obtained from the astrometric measurement in the r'-band exposure.
@@
 \\
-dec & double & degrees & J2000 Declination. See comments above. \\
-sourceID & string &  & Unique identification number of the detection. Identical to rDetectionID if the source was detected in the r-band; identical to iDetectionID or haDetectionID otherwise. \\
-posErr & float & arcsec & Astrometric fit error (RMS). Be aware that the error might be significantly larger than the RMS near CCD edges. \\
-l & double & degrees & Galactic longitude $\ell$ converted from ra/dec (IAU 1958 system). \\
-b & double & degrees & Galactic latitude $b$ converted from ra/dec (IAU 1958 system). \\
-mergedClass & short &  & Image classification flag based on all bands (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). Computed using the UKIDSS scheme. \\
-mergedClassStat & float &  & Merged N(0,1) stellarness-of-profile statistic. Computed using the UKIDSS scheme. \\
-pStar & float &  & Probability that the source is a star (value between 0 and 1). \\
-pGalaxy & float &  & Probability that the source is a galaxy (value between 0 and 1). \\
-pNoise & float &  & Probability that the source is noise (value between 0 and 1). \\
-pSaturated & float &  & Probability that the source is saturated (value between 0 and 1). \\
-rmi & float & mag & ($r$ - $i$) colour, formed by subtracting columns r and i.
+3 & dec & double & degrees & J2000 Declination. See comments above. \\
+4 & sourceID & string &  & Unique identification number of the detection. Identical to rDetectionID if the source was detected in the r-band; identical to iDetectionID or haDetectionID otherwise. \\
+5 & posErr & float & arcsec & Astrometric fit error (RMS). Be aware that the error might be significantly larger than the RMS near CCD edges. \\
+6 & l & double & degrees & Galactic longitude $\ell$ converted from ra/dec (IAU 1958 system). \\
+7 & b & double & degrees & Galactic latitude $b$ converted from ra/dec (IAU 1958 system). \\
+8 & mergedClass & short &  & Image classification flag based on all bands (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). Computed using the UKIDSS scheme. \\
+9 & mergedClassStat & float &  & Merged N(0,1) stellarness-of-profile statistic. Computed using the UKIDSS scheme. \\
+10 & pStar & float &  & Probability that the source is a star (value between 0 and 1). \\
+11 & pGalaxy & float &  & Probability that the source is a galaxy (value between 0 and 1). \\
+12 & pNoise & float &  & Probability that the source is noise (value between 0 and 1). \\
+13 & pSaturated & float &  & Probability that the source is saturated (value between 0 and 1). \\
+14 & rmi & float & mag & ($r$ - $i$) colour, formed by subtracting columns r and i.
 Included in the catalogue for convenience only.
@@
 \\
-rmha & float & mag & ($r$ - H$\alpha$) colour, formed by subtracting columns r and ha. See comments above. \\
-r & float & mag & Default r-band magnitude using a 2.3 arcsec diameter aperture. Calibrated in the Vega system. \\
-rErr & float & mag & Uncertainty for r. Does not include systematic errors. \\
-rPeakMag & float & mag & Alternative r-band magnitude derived from the peak pixel height (i.e. a 0.3x0.3 arcsec square aperture). Calibrated in the Vega system. \\
-rPeakMagErr & float & mag & Uncertainty in rPeakMag. Does not include systematics. \\
-rAperMag1 & float & mag & Alternative r-band magnitude using a 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
-rAperMag1err & float & mag & Uncertainty in rAperMag1. Does not include systematics. \\
-rAperMag3 & float & mag & Alternative r-band magnitude using a 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
-rAperMag3err & float & mag & Uncertainty in rAperMag3. Does not include systematics. \\
-rGauSig & float & pixels & RMS of axes of ellipse fit in r. \\
-rEll & float &  & Ellipticity in the r-band. \\
-rPA & float & degrees & Position angle in the r-band. \\
-rClass & short &  & Discrete image classification flag (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). \\
-rClassStat & float &  & N(0,1) stellarness-of-profile statistic. \\
-rErrBits & short &  & Bitmask used to flag a bright neighbour (1), source blending (2) and saturation (8). \\
-rMJD & double & days & Modified Julian Date at the start of the r-band exposure. \\
-rSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
-rDetectionID & string &  & Unique identifier of the r-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e.
+15 & rmha & float & mag & ($r$ - H$\alpha$) colour, formed by subtracting columns r and ha. See comments above. \\
+16 & r & float & mag & Default r-band magnitude using a 2.3 arcsec diameter aperture. Calibrated in the Vega system. \\
+17 & rErr & float & mag & Uncertainty for r. Does not include systematic errors. \\
+18 & rPeakMag & float & mag & Alternative r-band magnitude derived from the peak pixel height (i.e. a 0.3x0.3 arcsec square aperture). Calibrated in the Vega system. \\
+19 & rPeakMagErr & float & mag & Uncertainty in rPeakMag. Does not include systematics. \\
+20 & rAperMag1 & float & mag & Alternative r-band magnitude using a 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
+21 & rAperMag1err & float & mag & Uncertainty in rAperMag1. Does not include systematics. \\
+22 & rAperMag3 & float & mag & Alternative r-band magnitude using a 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
+23 & rAperMag3err & float & mag & Uncertainty in rAperMag3. Does not include systematics. \\
+24 & rGauSig & float & pixels & RMS of axes of ellipse fit in r. \\
+25 & rEll & float &  & Ellipticity in the r-band. \\
+26 & rPA & float & degrees & Position angle in the r-band. \\
+27 & rClass & short &  & Discrete image classification flag (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). \\
+28 & rClassStat & float &  & N(0,1) stellarness-of-profile statistic. \\
+29 & rErrBits & short &  & Bitmask used to flag a bright neighbour (1), source blending (2) and saturation (8). \\
+30 & rMJD & double & days & Modified Julian Date at the start of the r-band exposure. \\
+31 & rSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
+32 & rDetectionID & string &  & Unique identifier of the r-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e.
 composed of the INT telescope run number, the CCD number and a sequential source detection number.
 \\
-rX & float & pixels & Pixel coordinate of the source in the r-band exposure, in the coordinate system of the CCD. \\
-rY & float & pixels & Pixel coordinate of the source in the r-band exposure, in the coordinate system of the CCD. \\
-i & float & mag & Default i-band magnitude using a 2.3 arcsec diameter aperture. Calibrated in the Vega system. \\
-iErr & float & mag & Uncertainty for i. Does not include systematic errors. \\
-iPeakMag & float & mag & Alternative i-band magnitude derived from the peak pixel height (i.e. a 0.3x0.3 arcsec square aperture). Calibrated in the Vega system. \\
-iPeakMagErr & float & mag & Uncertainty in iPeakMag. Does not include systematics. \\
-iAperMag1 & float & mag & Alternative i-band magnitude using a 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
-iAperMag1err & float & mag & Uncertainty in iAperMag1. Does not include systematics. \\
-iAperMag3 & float & mag & Alternative i-band magnitude using a 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
-iAperMag3err & float & mag & Uncertainty in iAperMag3. Does not include systematics. \\
-iGauSig & float & pixels & RMS of axes of ellipse fit. \\
-iEll & float &  & Ellipticity. \\
-iPA & float & degrees & Position angle. \\
-iClass & short &  & Discrete image classification flag (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). \\
-iClassStat & float &  & N(0,1) stellarness-of-profile statistic. \\
-iErrBits & short &  & Bitmask used to flag a bright neighbour (1), source blending (2) and saturation (8). \\
-iMJD & double & days & Modified Julian Date at the start of the single-band exposure. \\
-iSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
-iDetectionID & string &  & Unique identifier of the r-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e.
+33 & rX & float & pixels & Pixel coordinate of the source in the r-band exposure, in the coordinate system of the CCD. \\
+34 & rY & float & pixels & Pixel coordinate of the source in the r-band exposure, in the coordinate system of the CCD. \\
+35 & i & float & mag & Default i-band magnitude using a 2.3 arcsec diameter aperture. Calibrated in the Vega system. \\
+36 & iErr & float & mag & Uncertainty for i. Does not include systematic errors. \\
+37 & iPeakMag & float & mag & Alternative i-band magnitude derived from the peak pixel height (i.e. a 0.3x0.3 arcsec square aperture). Calibrated in the Vega system. \\
+38 & iPeakMagErr & float & mag & Uncertainty in iPeakMag. Does not include systematics. \\
+39 & iAperMag1 & float & mag & Alternative i-band magnitude using a 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
+40 & iAperMag1err & float & mag & Uncertainty in iAperMag1. Does not include systematics. \\
+41 & iAperMag3 & float & mag & Alternative i-band magnitude using a 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
+42 & iAperMag3err & float & mag & Uncertainty in iAperMag3. Does not include systematics. \\
+43 & iGauSig & float & pixels & RMS of axes of ellipse fit. \\
+44 & iEll & float &  & Ellipticity. \\
+45 & iPA & float & degrees & Position angle. \\
+46 & iClass & short &  & Discrete image classification flag (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). \\
+47 & iClassStat & float &  & N(0,1) stellarness-of-profile statistic. \\
+48 & iErrBits & short &  & Bitmask used to flag a bright neighbour (1), source blending (2) and saturation (8). \\
+49 & iMJD & double & days & Modified Julian Date at the start of the single-band exposure. \\
+50 & iSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
+51 & iDetectionID & string &  & Unique identifier of the r-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e.
 composed of the INT telescope run number, the CCD number and a sequential source detection number.
 \\
-iX & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
-iY & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
-iXi & float & arcsec & Position offset of the i-band detection relative to the ra column. The original i-band coordinates can be obtained by computing (ra+iXi/3600, dec+iEta/3600). \\
-iEta & float & arcsec & Position offset of the i-band detection relative to the dec column. See comments above. \\
-ha & float & mag & Default H-alpha magnitude using a 2.3 arcsec aperture. Calibrated in the Vega system. \\
-haErr & float & mag & Uncertainty for ha. Does not include systematic errors. \\
-haPeakMag & float & mag & Alternative H-alpha magnitude derived from the peak pixel height (i.e. a 0.3x0.3 arcsec square aperture). Calibrated in the Vega system. \\
-haPeakMagErr & float & mag & Uncertainty in haPeakMag. Does not include systematics. \\
-haAperMag1 & float & mag & Alternative H-alpha magnitude using a 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
-haAperMag1err & float & mag & Uncertainty in haAperMag1. Does not include systematics. \\
-haAperMag3 & float & mag & Alternative H-alpha magnitude using a 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
-haAperMag3err & float & mag & Uncertainty in haAperMag3. Does not include systematics. \\
-haGauSig & float & pixels & RMS of axes of ellipse fit. \\
-haEll & float &  & Ellipticity \\
-haPA & float & degrees & Position angle. \\
-haClass & short &  & Discrete image classification flag (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). \\
-haClassStat & float &  & N(0,1) stellarness-of-profile statistic. \\
-haErrBits & short &  & Bitmask used to flag a bright neighbour (1), source blending (2) and saturation (8). \\
-haMJD & double & days & Modified Julian Date at the start of the single-band exposure. \\
-haSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
-haDetectionID & string &  & Unique identifier of the r-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e.
+52 & iX & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
+53 & iY & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
+54 & iXi & float & arcsec & Position offset of the i-band detection relative to the ra column. The original i-band coordinates can be obtained by computing (ra+iXi/3600, dec+iEta/3600). \\
+55 & iEta & float & arcsec & Position offset of the i-band detection relative to the dec column. See comments above. \\
+56 & ha & float & mag & Default H-alpha magnitude using a 2.3 arcsec aperture. Calibrated in the Vega system. \\
+57 & haErr & float & mag & Uncertainty for ha. Does not include systematic errors. \\
+58 & haPeakMag & float & mag & Alternative H-alpha magnitude derived from the peak pixel height (i.e. a 0.3x0.3 arcsec square aperture). Calibrated in the Vega system. \\
+59 & haPeakMagErr & float & mag & Uncertainty in haPeakMag. Does not include systematics. \\
+60 & haAperMag1 & float & mag & Alternative H-alpha magnitude using a 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
+61 & haAperMag1err & float & mag & Uncertainty in haAperMag1. Does not include systematics. \\
+62 & haAperMag3 & float & mag & Alternative H-alpha magnitude using a 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
+63 & haAperMag3err & float & mag & Uncertainty in haAperMag3. Does not include systematics. \\
+64 & haGauSig & float & pixels & RMS of axes of ellipse fit. \\
+65 & haEll & float &  & Ellipticity \\
+66 & haPA & float & degrees & Position angle. \\
+67 & haClass & short &  & Discrete image classification flag (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). \\
+68 & haClassStat & float &  & N(0,1) stellarness-of-profile statistic. \\
+69 & haErrBits & short &  & Bitmask used to flag a bright neighbour (1), source blending (2) and saturation (8). \\
+70 & haMJD & double & days & Modified Julian Date at the start of the single-band exposure. \\
+71 & haSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
+72 & haDetectionID & string &  & Unique identifier of the r-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e.
 composed of the INT telescope run number, the CCD number and a sequential source detection number.
 \\
-haX & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
-haY & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
-haXi & float & arcsec & Position offset of the H-alpha detection relative to the ra column. The original Ha-band coordinates can be obtained by computing (ra+haXi/3600, dec+haEta/3600). \\
-haEta & float & arcsec & Position offset of the H-alpha relative to the ra column. See comments above. \\
-brightNeighb & boolean &  & True if a very bright star is nearby (i.e. errBits $\geq$ 1). This indicates that the source might be spurious, or the photometry unreliable. \\
-deblend & boolean &  & True if the source was blended with a nearby neighbour due to crowding (i.e.
+73 & haX & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
+74 & haY & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
+75 & haXi & float & arcsec & Position offset of the H-alpha detection relative to the ra column. The original Ha-band coordinates can be obtained by computing (ra+haXi/3600, dec+haEta/3600). \\
+76 & haEta & float & arcsec & Position offset of the H-alpha relative to the ra column. See comments above. \\
+77 & brightNeighb & boolean &  & True if a very bright star is nearby (i.e. errBits $\geq$ 1). This indicates that the source might be spurious, or the photometry unreliable. \\
+78 & deblend & boolean &  & True if the source was blended with a nearby neighbour due to crowding (i.e.
 errBits $\geq$ 2).
@@
 \\
-saturated & boolean &  & True if the source is saturated (i.e.
+79 & saturated & boolean &  & True if the source is saturated (i.e.
 peak pixel $>$ 55000 counts) in one or more bands (i.e.
@@
 \\
-errBits & short &  & Maximum value of (rErrBits, iErrBits, haErrBits). \\
-nBands & short &  & Number of bands in which the source is detected (equals 1, 2 or 3). \\
-reliable & boolean &  & True if: errBits\,$\leq$\,2 \& nBands\,==\,3 \& r\,$>$\,13 \& i\,$>$\,12 \& ha\,$>$\,12.5 \& rErr\,$<$\,0.1 \& iErr\,$<$\,0.1 \& haErr\,$<$\,0.1 \& (abs(r-rAperMag1)\,$<$\,3*hypot(rErr,rAperMag1Err)+0.03) \& (abs(i-iAperMag1)\,$<$\,3*hypot(iErr,iAperMag1Err)+0.03) \& (abs(ha-haAperMag1)\,$<$\,3*hypot(haErr,haAperMag1Err)+0.03).
+80 & errBits & short &  & Maximum value of (rErrBits, iErrBits, haErrBits). \\
+81 & nBands & short &  & Number of bands in which the source is detected (equals 1, 2 or 3). \\
+82 & reliable & boolean &  & True if: errBits\,$\leq$\,2 \& nBands\,==\,3 \& r\,$>$\,13 \& i\,$>$\,12 \& ha\,$>$\,12.5 \& rErr\,$<$\,0.1 \& iErr\,$<$\,0.1 \& haErr\,$<$\,0.1 \& (abs(r-rAperMag1)\,$<$\,3*hypot(rErr,rAperMag1Err)+0.03) \& (abs(i-iAperMag1)\,$<$\,3*hypot(iErr,iAperMag1Err)+0.03) \& (abs(ha-haAperMag1)\,$<$\,3*hypot(haErr,haAperMag1Err)+0.03).
 \\
-veryReliable & boolean &  & True if: reliable \& pStar\,$>$\,0.9 \& errBits\,=\,0. \\
-fieldID & string &  & Human-readable IPHAS field number and observing run (e.g. 0001o\_aug2003). \\
-fieldGrade & string &  & Internal quality control score of the field. One of A, B, C or D. \\
-night & integer &  & Night of the observation (YYYYMMDD). Refers to the UT date at the start of the night. \\
-seeing & float & arcsec & Maximum value of (rSeeing, iSeeing, haSeeing). \\
-ccd & short &  & CCD-chip number on the Wide Field Camera (WFC) of the Isaac Newton Telescope (INT). 1, 2, 3 or 4. \\
-nObs & short &  & Number of repeat observations of this source in the survey. \\
-sourceID2 & string &  & SourceID of the object in the partner exposure (if obtained within 10 minutes of the primary detection). \\
-fieldID2 & string &  & FieldID of the partner detection (e.g. 0001o\_aug2003). \\
-r2 & float & mag & r-band magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
-rErr2 & float & mag & Uncertainty for r2. \\
-i2 & float & mag & i-band magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
-iErr2 & float & mag & Uncertainty for i2. \\
-ha2 & float & mag & H-alpha magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
-haErr2 & float & mag & Uncertainty for ha2. \\
-errBits2 & integer &  & Error bitmask for the partner detection. Used to flag a bright neighbour (1), source blending (2), saturation (8), vignetting (64), truncation (128) and bad pixels (32768). \\
+83 & veryReliable & boolean &  & True if: reliable \& pStar\,$>$\,0.9 \& errBits\,=\,0. \\
+84 & fieldID & string &  & Human-readable IPHAS field number and observing run (e.g. 0001o\_aug2003). \\
+85 & fieldGrade & string &  & Internal quality control score of the field. One of A, B, C or D. \\
+86 & night & integer &  & Night of the observation (YYYYMMDD). Refers to the UT date at the start of the night. \\
+87 & seeing & float & arcsec & Maximum value of (rSeeing, iSeeing, haSeeing). \\
+88 & ccd & short &  & CCD-chip number on the Wide Field Camera (WFC) of the Isaac Newton Telescope (INT). 1, 2, 3 or 4. \\
+89 & nObs & short &  & Number of repeat observations of this source in the survey. \\
+90 & sourceID2 & string &  & SourceID of the object in the partner exposure (if obtained within 10 minutes of the primary detection). \\
+91 & fieldID2 & string &  & FieldID of the partner detection (e.g. 0001o\_aug2003). \\
+92 & r2 & float & mag & r-band magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
+93 & rErr2 & float & mag & Uncertainty for r2. \\
+94 & i2 & float & mag & i-band magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
+95 & iErr2 & float & mag & Uncertainty for i2. \\
+96 & ha2 & float & mag & H-alpha magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
+97 & haErr2 & float & mag & Uncertainty for ha2. \\
+98 & errBits2 & integer &  & Error bitmask for the partner detection.
+Used to flag a bright neighbour (1), source blending (2), saturation (8), vignetting (64), truncation (128) and bad pixels (32768).
+\\

### 2014-04-13 paper/tables/columns.py

@@ -25,8 +25,9 @@ data = json.loads(open(filename).read())
 
 # Write TeX table
 out = open('columns.tex', 'w')
-for column in data:
+for i, column in enumerate(data):
+    column['n'] = i + 1
     column['desc'] = html2tex(column['desc'])
-    out.write("{name} & {type} & {unit} & {desc} \\\\\n".format(**column))
+    out.write("{n} & {name} & {type} & {unit} & {desc} \\\\\n".format(**column))
 out.close()
 

### 2014-04-13 paper/iphas-dr2.tex
Changes following on from edit by JD

@@
 \usepackage{subcaption}
-
-\restylefloat{table}
 
@@
 \pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2014}
-
 \maketitle
@@
 is a 1800~deg$^2$ imaging survey
-covering the Galactic latitude range $|b| < 5^\circ$
+covering Galactic longitudes $l=30^\circ$ to $215^\circ$
+and latitudes $b = -5^\circ$ to $+5^\circ$
 in the $r$, $i$ and \ha\ filters 
@@
 for 219 million unique sources
-across 92\% of the 1800 square-degree footprint.
+across 92\% of the footprint.
 The observations were carried out between 2003 and 2013
 at a median seeing of 1.1 arcsec
-and to mean $5\sigma$-depths of 
+and to a mean $5\sigma$-depth of 
 21.2 ($r$), 20.0 ($i$) and 20.3 ($\ha$)
@@
 0.03~mag (rms) 
-and provide a series of recommended quality criteria
-which can be used to select the most reliable data
+and recommend a series of quality criteria
+to select the most reliable data
 from the catalogue.
@@
 \begin{table}
-% The caption of tables should be on top and not at the bottom
     \caption{Mean magnitude offsets for objects
@@
             \end{tabular}
-
         \end{center}
@@
 the majority of the objects in our catalogue
-are faint sources observed near the detection limits:
+are faint sources observed near the detection limits;
 55\% of the entries in the catalogue
@@
 which should benefit most applications
-which require reliable $(r-i)$ and $(r-\ha)$ colours
-(GEERT: the rest of the sentence looks very odd to me -- I can't rephrase
-because I'm not sure what you are trying to say) without removing more 
-than 80\%
-of the sources brighter than $r < 19$.
+which desire reliable $(r-i)$ and $(r-\ha)$ colours
+as well as completeness.
 The listed criteria are designed to 
@@
 \begin{table*}
-
 \vspace{2cm}
@@
 
-\vspace{2cm}
+\vspace{1cm}
 \end{table*}
@@
 For applications which require
-an even higher standard of reliability,
+a higher standard of reliability at the expense of completeness,
 a further set of additional quality criteria
@@
 59 million sources (27\%) satisfy
-these additional criteria 
+these stricter criteria 
 and are hereafter referred to as ``very reliable''.
@@
 Fig.~\ref{fig:magdist} compares the $r$-band magnitude
-distribution for objects with and without the 
-\emph{reliable} and \emph{veryReliable} criteria applied. 
+distributions for \emph{reliable}, \emph{veryReliable}
+and unfiltered objects.
 We find that 81\% of the sources 
-in the magnitude range $13 < r < 19$
-are considered \emph{reliable},
-which drops to 72\% in the range $19 < r < 20$
-and 9\% at $r>20$.
-In contrast, only 54\% of the sources 
-in the magnitude range $13 < r < 19$
-are considered \emph{veryReliable}.
-In \S\ref{sec:densities} we will show that these
-stricter criteria remove large numbers
-of objects at early Galactic longitudes 
-where source blending can sometimes affect
-more than a quarter of the objects.
+are considered \emph{reliable}
+and 54\% are \emph{veryReliable}
+in the magnitude range $13 < r < 19$.
+We will explain below that the
+\emph{veryReliable} category is least complete
+at low Galactic longitudes,
+where source blending can affect
+up to a quarter of the objects.
 The \emph{veryReliable} flag should hence
@@
 For example, by raising the allowed photometric uncertainties
-from 0.1 mag to 0.2 mag,
-42 million candidate sources are added to the 109 
-million satisfying the tighter error bound (GEERT - is this what you mean?).
+from 0.1~mag to 0.2~mag,
+42 million candidate sources would be added to the 109 
+million satisfying the tighter error bound.
+Our choice to adopt 0.1~mag as the cut-off uncertainty
+for the \emph{reliable} category is a pragmatic trade-off
+which we found to suit many science applications,
+but certainly not all. Users of the data are encouraged 
+to revise the quality criteria according to their needs.
 
@@
 
-%As we explained earlier, blended sources must be used with caution.
 Finally, we warn that blended objects
@@
         \toprule
-    # & Name [IPHAS2 ...] & $r$ & $i$ & \ha  \\
+        \# & Name [IPHAS2 ...] & $r$ & $i$ & \ha  \\
         \midrule
@@
         \bottomrule
-    \end{tabular}}
+        \end{tabular}
+    }
 \end{table}
@@
 In previous work, we have shown that the majority of
-\ha-emitters seen towards an H{\sc ii} region
+\ha-emitters that are unveiled by IPHAS towards an H{\sc ii}
 are likely to be Classical T Tauri Stars \citep{Barentsen2011a}.
@@
 we discovered towards Sh 2-82 as well.
-Two of our candidates, x and y in Tab.~\ref{tab:emitters}, have recently 
-been identified
-as YSOs in the initial investigation of the region by \cite{Yu2012}.
-In their study, the authors used 2MASS and Spitzer data
-to select likely young stars by selecting objects with
-an infrared colour excess, consistent with 
-the presence of a heated circumstellar disk.
-Although \citeauthor{Yu2012} did not identify
-the other four candidate emitters in our sample
-as having an infrared excess,
-we note that all four objects
-are detected in the Spitzer 8.0~\micron\ image
-at S/N$>$5,
-which is consistent with their
-being optically-unveiled stars with a circumstellar disk,
+Two of our candidates, \#1 and \#4 in Tab.~\ref{tbl:emitters},
+have previously been identified
+as candidate YSOs by \cite{Robitaille2008}
+and \cite{Yu2012}, respectively.
+In their studies, the authors used Spitzer data
+to select intrinsically red objects,
+consistent with the presence of a circumstellar disk.
+Although the other four candidate emitters in our sample
+have not previously appeared in the literature,
+we note that all four are detected
+in the Spitzer 8.0~\micron\ image at S/N$>$5,
+which is consistent with being optically-unveiled stars with a circumstellar disk,
 i.e. they are likely to be Class II YSOs exhibiting a mild infrared excess.
@@
 southern Plane is already available thanks to the UK
-Schmidt H$\alpha$ Survey \citep[SHS,][][Parker05], the last of the 
+Schmidt H$\alpha$ Survey \citep[SHS,][]{Parker2005}, the last of the 
 photographic surveys carried out by that telescope.
-%We explain the data reduction and quality control procedures that
-%were applied.  We describe and test the global photometric 
-%calibration, 
-%and detailed the construction of the source catalogue.
 
@@
 
-The IPHAS survey was carried out 
-at the Isaac Newton Telescope (INT).
 The INT is operated on the island of La Palma
@@
 \small
-\begin{longtable}{lllp{10cm}}
+\begin{longtable}{llllp{10cm}}
 \caption{\label{tab:columns} 
@@
 \hline
-Column & Type & Unit & Description \\
+\# & Column & Type & Unit & Description \\
 \hline

### 2014-04-13 paper/iphas-dr2.tex
Corrections for comments by NW

@@
 \citep{Vink2008,Barentsen2011a,Wright2012};
-%new asymptotic giant branch stars \citep{Wright2008};
 new classical Be stars \citep{Raddi2013};
-along with discoveries of new and remarkable planetary nebulae 
-\citep{Mampaso2006, Viironen2009a, Viironen2009b, Corradi2011, Viironen2011}
-and new supernova remnants \citep{Sabin2013}.
+along with discoveries of new supernova remnants \citep{Sabin2013}
+and new and remarkable planetary nebulae 
+\citep{Mampaso2006, Viironen2009a, Viironen2009b, Sabin2010, Corradi2011, Viironen2011}.
 
@@
 It is the first digital survey to offer comprehensive optical CCD photometry
-of point sources in the Galactic Plane, 
-and does so down to a limiting magnitude of $\sim$20.
-The IPHAS footprint spans a box 
+of point sources in the Galactic Plane;
+the footprint spans a box 
 of roughly 180 by 10 degrees, 
-spanning Galactic latitudes $-5^{\circ} < b < +5^{\circ}$ 
+covering Galactic latitudes $-5^{\circ} < b < +5^{\circ}$ 
 and longitudes $30^{\circ} < l < 215^{\circ}$.
@@
 30 sec ($r$) and 10 sec ($i$)
-in the majority of the survey observations.\footnote{In 2003 
-the $r$-band exposure time was 10~sec instead of 30~sec,
-and since Oct 2010 the $i$-band exposure time 
-has been increased from 10~sec to 20~sec.}
+in the vast majority of the survey observations.\footnote{The $r$-band exposure time was 10~s instead of 30~s in the first year of survey observations (2003).
+Since October 2010, the $i$-band exposure time
+has been increased from 10~s to 20~s to by-pass a sporadic exposure timing issue that affects the WFC.}
 
@@
 using repeat observations. 
-This name -- which has since been adopted by other surveys --
-is an anglicised version of the German word `\"uberkalibration',
-which was reportedly chosen because authors Doug Finkbeiner and David Schlegel
-both have German-sounding names (Finkbeiner 2010).} procedure,
+It is an anglicised version of the German word `\"uberkalibration',
+which was reportedly chosen because the initial authors both
+had German-sounding names \citep{Finkbeiner2010}.} procedure,
 in which multiple measurements of stars at different epochs
@@
 attributable to the significantly increasing extinction.
-Meanwhile the redder group is principally composed of red giant stars.
+Meanwhile the redder group is principally composed of red giant stars \citep[see][]{Wright2008}.
 As these stars are intrinsically brighter, they will be substantially 
@@
 of a forthcoming paper (Barentsen et al, in preparation).
-In this section were merely aim to demonstrate
+In this section we merely aim to demonstrate
 a use of the catalogue for this purpose.
@@
 in the constellation of Sagitta.
-Nicknamed by amateur astronomers as the ``Little Cocoon Nebula'',
+Nicknamed by amateur astronomers as the `Little Cocoon Nebula',
 Sh 2-82 is ionised by 
@@
 \begin{figure*}
-    \begin{minipage}[b]{0.8\linewidth}
+    \begin{minipage}[b]{0.78\linewidth}
         \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-iphas.pdf} 
@@
 \label{fig:mosaic_iphas}
-    \begin{minipage}[b]{0.8\linewidth}
+    \begin{minipage}[b]{0.78\linewidth}
         \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-spitzer.pdf} 
@@
     \caption{Star-forming region Sh 2-82 as seen in the mid-infrared
-    by the Spitzer Space Telescope's GLIMPSE survey. The mosaic is composed of the 24~\micron\ (red), 8.0~\micron\ (green) and 4.5~\micron\ (blue) bands.
-    The image reveals a bubble-shaped structure which surrounds the {\sc Hii} region that is seen in the IPHAS mosaic of the same region (Fig.~\ref{fig:mosaic_iphas}). 
+    by the Spitzer Space Telescope. The mosaic is composed of the 24~\micron\ (red), 8.0~\micron\ (green) and 4.5~\micron\ (blue) bands.
+    The image reveals a bubble-shaped structure which surrounds the {\sc Hii} region that is seen in the IPHAS mosaic which spans the same region (Fig.~\ref{fig:mosaic_iphas}). 
     This structure has previously been labelled as N115 in the 
-catalogue due to \citet{Churchwell2006}.}
+catalogue due to \citet{Churchwell2006}, and could be a possible site of triggered star formation \citep{Thompson2012,Kendrew2012}.}
     \label{fig:mosaic_spitzer}
@@
 with \ha\ in emission
-at a strength of EW=$-10{\rm \AA}$ (dashed line),
-taken from the colour simulations due to \citet{Barentsen2011a}.
+at a strength of EW$=-10$~\AA\ (dashed line).
 Six stars are found to lie above the 
 dashed line at the level of $3\sigma$,
-that is, the distance between the objects and the dashed line
+i.e. the distance between the objects and the dashed line
 is larger than three times the uncertainty
 in their $(r-\ha)$ colour.
-These reliable candidate \ha-emitters
-are marked by red triangles in the colour-colour diagram
-and their photometry is set out in Table~\ref{tbl:emitters}.
-The spatial distribution of the six candidate emission-line objects
-is marked by yellow triangles in the colour mosaic (Fig.~\ref{fig:mosaic_iphas}).
+These candidate \ha-emitters
+are marked by red triangles in the colour-colour diagram,
+and yellow triangles in the image mosaic (Fig.~\ref{fig:mosaic_iphas}).
+Their details are listed in Table~\ref{tbl:emitters}.
 
@@
     while the main sequence for stars with an \ha\ emission line
-    strength of $-10~\rm{\AA}$ EW is indicated by a dashed line,
-    following the colour simulations due to \citet{Barentsen2011a}.
+    strength of $-10~\rm{\AA}$~EW is indicated by a dashed line
+    \citep[both based on the colour simulations by][]{Barentsen2011a}.
     Red triangles indicate objects which have been identified as
-    as highly likely \ha-emitters (see text).}
+    as likely \ha-emitters.}
     \label{fig:emitters}
@@
 In previous work, we have shown that the majority of
-\ha-emitters that are unveiled by IPHAS towards an H{\sc ii}
+\ha-emitters seen by IPHAS towards an H{\sc ii} region
 are likely to be Classical T Tauri Stars \citep{Barentsen2011a}.
@@
 and \cite{Yu2012}, respectively.
-In their studies, the authors used Spitzer data
+In these studies, the authors used Spitzer data
 to select intrinsically red objects,
-consistent with the presence of a circumstellar disk.
+which is consistent with the presence of a circumstellar disk.
 Although the other four candidate emitters in our sample
@@
 we note that all four are detected
-in the Spitzer 8.0~\micron\ image at S/N$>$5,
-which is consistent with being optically-unveiled stars with a circumstellar disk,
-i.e. they are likely to be Class II YSOs exhibiting a mild infrared excess.
+in the Spitzer 8.0~\micron\ image at S/N$>$5.
+They are likely to be Class II YSOs exhibiting a mild infrared excess.
 
@@
 which have only recently started to become revealed
-by efforts to catalogue the wealth of ``bubbles'' detected
+by efforts to catalogue the wealth of `bubbles' detected
 at mid-infrared wavelengths \citep{Churchwell2006,Simpson2012},
@@
 in the near-infrared \cite[e.g.][]{Bica2003}.
-The data provided by IPHAS offers a handle
-or lower limit on the extinction and distance
-towards these regions,
-and can be used to discover T~Tauri stars
-when the extinction is not too high ($A_V < 10$).
+IPHAS data can offer a handle
+on the extinction, distance
+and stellar contents of these unexplored regions.
 
@@
 \small
-\begin{longtable}{llllp{10cm}}
+\begin{longtable}{rlllp{10cm}}
 \caption{\label{tab:columns} 

### 2014-04-13 paper/iphas-dr2.tex
Processed feedback from YU

@@
 they can combine, during the global calibration routine, to cause artificial zeropoint gradients 
-across the survey unless controlled by
-%when they are used to constrain a global calibration
-%without 
-other external constraints.
+across the survey unless controlled by other external constraints.
 
@@
 to bring the majority of our data onto a uniform calibration.
-%This is explained next.
 
@@
 (APASS; http://www.aavso.org/apass)
-to bring 
-%the vast majority - exaggerating a bit? 
-most of the survey 
+to bring most of the survey 
 onto a uniform scale.
@@
 which is significantly better 
-than the original nightly calibrations of IPHAS,
-for which we estimate the $1\sigma$-uncertainty to be $\sim$10\%.
+than the original nightly calibrations of IPHAS
+which are accurate to $\sim$10\% (Tab.~\ref{tbl:offsets_before}).
 APASS achieves its uniform accuracy 
-by measuring each star at least two times in photometric conditions
--- along with ample standard fields --
-using the large $3\times3$ square degrees field of view of its detectors.
-
-With the aim of bringing IPHAS to a similar accuracy of $\sim$3\%,
+by measuring each star at least two times in photometric conditions,
+along with ample standard fields,
+benefiting from the large 3-by-3 degrees field of view of its detectors.
+
+With the aim of bringing IPHAS to a similar accuracy of 3\%,
 we used the APASS catalogue to identify and adjust the calibration of all IPHAS fields 
@@
 APASS magnitudes are given in the AB system
-and IPHAS uses Vega magnitudes.
+and IPHAS uses Vega-based magnitudes.
 Separate transformations were derived for sight-lines 
@@
 for each field which contained at least 30 cross-matched stars.
-This was achieved for 48\% of our fields.
+This was achieved for 48\% of our fields (shown in Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}).
 The mean offset was found to be
@@
     \vspace{1cm}
+    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_ishift.pdf} 
+    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_i.pdf} 
+    \caption{Same as Fig.~\ref{fig:apass_r} for the $i$-band.}
+    \label{fig:apass_i}
+\end{figure*}
+
+\begin{figure*}
     \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_rshift.pdf}
     \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_r.pdf} 
-    \caption{Median magnitude offset in the $r$ band
-             between IPHAS and SDSS after the re-calibration
-             procedure was applied.
+    \caption{Median magnitude offset between IPHAS and SDSS in the $r$ band
+             \emph{after} the re-calibration procedure was applied.
              Each square represents the footprint of an IPHAS field
@@
              which was computed after applying the SDSS-to-IPHAS
-             transformation to the SDSS magnitudes (Eqn.~\ref{eqn:sdss_r}).
-             For clarity, we do not show the fields at the offset positions.}
+             transformation to the SDSS magnitudes (Eqn.~\ref{eqn:sdss_r}).}
     \label{fig:sdss_r}
-\end{figure*}
-
-\begin{figure*}
-	\vspace{2cm}
-    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_ishift.pdf} 
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_i.pdf} 
-    \caption{Same as Fig.~\ref{fig:apass_r} for the $i$-band.}
-        \label{fig:apass_i}
     \vspace{1cm}
@@
     \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_i.pdf} 
-    \caption{Same as Fig.~\ref{fig:sdss_i} for the $i$-band.}
+    \caption{Same as Fig.~\ref{fig:sdss_r} for the $i$-band.}
     \label{fig:sdss_i}
@@
              cross-matched between IPHAS and APASS/SDSS
-             before the uniform re-calibration.
-             Eqns.~\ref{eqn:apass_r}-\ref{eqn:sdss_i} were applied
-             to the APASS/SDSS magnitudes to bring them into the
+             \emph{before} the uniform re-calibration was carried out.
+             Transformations were applied
+             to the APASS and SDSS magnitudes to bring them into the
              Vega-based IPHAS system prior to computing
-             the offsets.)
+             these statistics (see text).
              }
@@
        \caption{Same as Table~\ref{tbl:offsets_before}
-                but computed after the uniform re-calibration
+                but computed \emph{after} the uniform re-calibration
                  was carried out.}
@@
 At the time of writing, the APASS catalogue did not provide 
-sufficient coverage for 7359 of the fields in our data release.
+sufficient coverage for 7359 of our fields.
 Fortunately, these fields are located mainly 
-in the early part of the Galactic Plane (Fig.~\ref{fig:apass_r}),
+at low Galactic longitudes (cf. Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}),
 which were typically observed during the summer months
@@
 located in the overlap regions between neighbouring fields.
-%Although we explained earlier that these overlap regions
-%are prone to systematics, the use of APASS enabled us to
-%keep the zeropoints of roughly half of the fields fixed,
-%which avoids these systematics from combining to introduce
-%artificial gradients across the survey.
-%In the next section we will show this to be true by validating
-%our calibration against SDSS.
+
 An algorithm for achieving this minimisation
-%of minimising the magnitude differences between overlapping frames 
 has previously been described by~\citet{Glazebrook1994}.
@@
 \begin{equation}
-   S = \sum_{i=1}^N \sum_{j=1}^N w_{ij} \theta_{ij} (\Delta_{ij} + a_i - a_j)^2
+   S = \sum_{i=1}^N \sum_{j=1}^N w_{ij} \theta_{ij} (\Delta_{ij} + a_i - a_j)^2,
    \label{eqn:chi2}
 \end{equation}
-where $i$ denotes the exposure of interest, 
+where $i$ denotes an exposure, 
 $j$ an overlapping exposure, 
@@
 and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$). 
-$w_{ij}$ are weights set to the inverse square of the uncertainty in $\Delta_{ij}$
+$w_{ij}$ are weights set to the inverse square of the uncertainty in $\Delta_{ij}$,
 and $\theta_{ij}$ is an overlap function 
@@
 \end{eqnarray}
-This prescription is essentially identical to \citet{Glazebrook1994}.
-%,although we changed the formulation to ensure that $A_{ij}$ is positive definite.
 
@@
 The median offsets for each of these fields
-are shown in Figs.~\ref{fig:sdss_r}-\ref{fig:sdss_i}.
+are shown in Figs.~\ref{fig:sdss_r} and \ref{fig:sdss_i}.
 Importantly, the mean offset and standard deviation found 
@@
 exceeded 0.05~mag (523 fields) or even 0.1~mag (18 fields).
-This pattern of outliers is consistent with the tails of a Gaussian
-with mean $\approx0$ and $\sigma=0.03$.  Furthermore, it should not
+This pattern of outliers is consistent with the tails of a Gaussian distribution
+with $\sigma=0.03$.  Furthermore, it should not
 be forgotten that both the SDSS and APASS calibrations are approximations to
@@
 two surveys.
-%, we hope to draw upon
-%the PanSTARRS survey \citep{Schlafly2012}
-%to further improve the accuracy of our calibration.
-%At the time of preparing this work data from PanSTARRS
-%had not been made public yet.
 
@@
 
-\subsection{Photometric uncertainties and repeatability}
-
-Fig.~\ref{fig:uncertainties} shows the mean photometric
-uncertainty as a function of magnitude in each band.
-We find that the uncertainty typically
-reaches 0.1~mag near 20.5 in $r$ 
-and 19.5 in $i$/\ha\ for the default 2.3\arcsec\ aperture.
-At this point we note that the average colour
-across all sources in the catalogue is
-$1.06\pm0.12$ for $(r-i)$ and $0.44\pm0.03$ for $(r-\ha)$. The
-fainter limiting magnitude in $r$ is hence compensated
-by the fact that most sources have 
-brighter magnitudes in both $i$ and \ha.
+\subsection{Random and systematic uncertainties}
 
@@
              columns \emph{rErr}, \emph{iErr} and \emph{haErr}
-             in the catalogue, 
+             in the catalogue 
              and the error-bars show the standard deviations.
@@
              the 10$\sigma$ and 5$\sigma$ limits, respectively.
-             These uncertainties are based only on the (Poissonian)
-             photon noise and hence this figure does not show
-             the level of systematic uncertainties.}
+             These statistics show the average level of the Poissonian
+             photon noise and do not include systematic uncertainties.}
     \label{fig:uncertainties}
@@
 
-The uncertainties shown in Fig.~\ref{fig:uncertainties}
+Fig.~\ref{fig:uncertainties} shows the mean photometric
+uncertainties (\emph{rErr}, \emph{iErr}, \emph{haErr})
+as a function of magnitude.
+We find that the uncertainty typically
+reaches 0.1~mag near 20.5 in $r$ 
+and 19.5 in $i$ and \ha.
+We note that the fainter depth in $r$ is compensated
+by the fact that most stars have brighter magnitudes in $i$ and \ha;
+the average colours in the catalogue are
+$\overline{(r-i)}=1.06\pm0.12$ and $\overline{(r-\ha)}=0.44\pm0.03$.
+We warn that the statistics shown in Fig.~\ref{fig:uncertainties}
 are the random errors based on the expected Poissonian photon noise.
@@
 are not included.
-To appraise the extent to which our photometry is affected
-by such systematics, we can exploit the
-secondary measurements which are present in the catalogue.
-Fig.~\ref{fig:pairmag} shows the mean residuals between
-the primary and secondary magnitudes
--- i.e. the average difference between catalogue columns \emph{r}-\emph{r2},
-\emph{i}-\emph{i2}, \emph{ha}-\emph{ha2} -- as a function of magnitude.
-We find that sources across the magnitude range 
-$13 < r < 17$ are consistent at the level of 5\%
-(i.e. $\sigma_{r-r2} \le 0.05$ mag),
-with the best repeatability
-reached at $r=14$ ($\sigma_{r-r2}~=~0.041$~mag).
-The brightest stars tend to show large residuals
--- e.g. $\sigma_{r-r2}~=~0.14$~mag at $r=12$ --
-which is due to saturation effects.
-At the faint end we find residuals which
-show significantly more scatter than would be expected
-from photon noise alone;
-the effects of source blending and background subtraction
-appear to dominate from $\sim$20th magnitude onward.
+
+To appraise the level at which our photometry is affected by systematics,
+we can exploit the secondary measurements which are present in the catalogue
+(i.e. \emph{r} vs \emph{r2}, \emph{i} vs \emph{i2}, \emph{ha} vs \emph{ha2}).
+In Fig.~\ref{fig:pairmag} we show the mean absolute residuals between
+these primary and secondary magnitudes as a function of magnitude (black dots).
+We also plot the Poissonian uncertainties for comparison (solid red line).
+We find the mean residual to equal 0.03~mag
+across the magnitude ranges 13 to 18 ($r$)
+and 12 to 17 ($i$, \ha),
+consistent with the accuracy of our calibration estimated earlier.
+Stars fainter than this range appear to be dominated by photon noise (red line),
+while stars at the bright end appear to suffer from significant systematic
+errors due to saturation effects.
 
@@
 \caption{Photometric repeatability as a function of magnitude
-         for both all sources in the catalogue (left panel)
+         for all sources in the catalogue (left panel)
          and for the \emph{veryReliable} sources alone (right panel).
-         The values shown are the mean absolute residuals
-         between the primary and the secondary detections,
-         while the error-bars show the standard deviations.
-         Left panel: 
-         the best photometric repeatability is reached at $r=14$
-         with $\sigma_{r-r2}~=~0.041$~mag.
-         Bright stars at $r<13$ and $i<12$ 
-         show increasing uncertainties due to saturation effects.
-         Right panel: we find that applying the quality criteria
-         has improved the photometric repeatability significantly.
-         The best repeatability is again reached at $r=14$
-         but has reduced to $\sigma=0.028$ mag.
-         The quality criteria have also been successful
-         at removing objects with large systematics at the bright
-         and faint ends.
+         Black dots show the mean absolute residuals
+         between the primary and the secondary detections.
+         The error-bars show the standard deviations.
+         The red trend line shows the average Poissonian uncertainties
+         we derived in Fig.~\ref{fig:uncertainties}.
+         We find that the \emph{veryReliable} quality criteria are successful
+         at removing objects with large uncertainties.
     }
@@
 In Fig.~\ref{fig:pairmag_reliable} we show 
-a similar comparison of the primary and secondary detections,
+a similar comparison between the primary and secondary detections,
 but this time we have only included sources which are flagged
-as \emph{veryReliable} in the catalogue
-(i.e. not saturated, not confused, not near bright stars, etc.)
-We find that the average residuals are significantly better
-for this subset. Sources across the magnitude range 
-$13 < r < 17$ now show a scatter of 0.03 instead of 0.05~mag.
-The best repeatability is again reached at $r=14$,
-but this time with $\sigma_{r-r2}~=~0.028$~mag.
+as \emph{veryReliable} in the catalogue.
+We do not observe an improvement in the average residuals as a function of magnitude,
+but the number of outliers is significantly smaller
+(evidenced by the shorter error bars which denote the standard deviation of the absolute residuals).
 We conclude that the \emph{veryReliable} quality criteria are effective
-in reducing the systematic errors to
-the same level as the estimated accuracy of the
-global photometric calibration (cf. \S\ref{sec:calibration}).
-Moreover, the large systematics at the bright and faint end
-disappear.
+at reducing the level of the systematic errors,
+while also removing the unreliable data at the bright and faint end.
 
@@
 
-The number of sources in our catalogue,
-counted in 1-by-10~deg$^2$ constant-longitude strips,
-is shown in Fig.~\ref{fig:sourcecount}
+Fig.~\ref{fig:sourcecount} shows the number of sources
+in the catalogue counted in $1^\circ$-wide strips
 as a function of Galactic longitude (thick blue line).	
@@
 found near $l\simeq 180^\circ$.
-
 In addition to the global trend,
-real variations are also apparent on smaller scales.
+variations are also apparent on smaller scales.
 For example, we find a significant drop near the constellations 

### 2014-04-13 paper/iphas-dr2.tex
details

@@
 is a 1800~deg$^2$ imaging survey
-covering Galactic longitudes $l=30^\circ$ to $215^\circ$
-and latitudes $b = -5^\circ$ to $+5^\circ$
+covering Galactic latitudes $|b| < 5^\circ$
+and longitudes $l=30$ to $215^\circ$
 in the $r$, $i$ and \ha\ filters 
@@
 describe and test the new uniform calibration,
-and detail the construction of the new 98-column source catalogue.
+and detail the construction of the new catalogue.
 We show that the global calibration is accurate to
@@
 30 sec ($r$) and 10 sec ($i$)
-in the vast majority of the survey observations.\footnote{The $r$-band exposure time was 10~s instead of 30~s in the first year of survey observations (2003).
-Since October 2010, the $i$-band exposure time
-has been increased from 10~s to 20~s to by-pass a sporadic exposure timing issue that affects the WFC.}
+in the vast majority of the survey observations.\footnote{The $r$-band exposure time was 10~s instead of 30~s in the first year of data taking (2003). Since October 2010, the $i$-band exposure time 
+has been increased from 10~s to 20~s to by-pass a sporadic exposure timing bug that affects the WFC.}
 
@@
 as a function of magnitude.
-We find that the uncertainty typically
-reaches 0.1~mag near 20.5 in $r$ 
-and 19.5 in $i$ and \ha.
+We find the typical uncertainty to reach 0.1~mag near $r=$20.5 
+and $i$/\ha$=19.5$.
 We note that the fainter depth in $r$ is compensated
@@
 
-To appraise the level at which our photometry is affected by systematics,
+To appraise the level at which our photometry is affected by such systematics,
 we can exploit the secondary measurements which are present in the catalogue
@@
 We also plot the Poissonian uncertainties for comparison (solid red line).
-We find the mean residual to equal 0.03~mag
+We find the mean residual to equal $0.03\pm0.04$~mag
 across the magnitude ranges 13 to 18 ($r$)
 and 12 to 17 ($i$, \ha),
-consistent with the accuracy of our calibration estimated earlier.
+consistent with the accuracy of our calibration estimated above.
 Stars fainter than this range appear to be dominated by photon noise (red line),
-while stars at the bright end appear to suffer from significant systematic
+while stars at the bright end appear to suffer from large systematic
 errors due to saturation effects.
@@
 We do not observe an improvement in the average residuals as a function of magnitude,
-but the number of outliers is significantly smaller
+but the number of outliers has decreased markedly
 (evidenced by the shorter error bars which denote the standard deviation of the absolute residuals).
@@
 We are deeply indebted to the ING staff and students
-for their ongoing support of the INT.
+for their ongoing support.
 All data were processed 
@@
 by a number of software packages,
-including the PostgreSQL database software,
-the TOPCAT and STILTS packages \citep{Taylor2005,Taylor2006},
+including the {\sc postgresql} database software,
+the {\sc topcat} and {\sc stilts} packages \citep{Taylor2005,Taylor2006},
 and the Python modules
-AstroPy \citep{Astropy},
-NumPy/SciPy \citep{Numpy},
-Matplotlib \citep{Matplotlib},
-IPython \citep{IPython},
-and APLpy.
-We also made use of the Montage software maintained by NASA/IPAC,
-and the SIMBAD, Vizier and Aladin \citep{Aladin} services
-operated at CDS, Strasbourg, France.
-
-This research made extensive use of
+{\sc astropy} \citep{Astropy},
+{\sc numpy} and {\sc scipy} \citep{Numpy},
+{\sc matplotlib} \citep{Matplotlib},
+{\sc ipython} \citep{IPython},
+and {\sc aplpy}.
+We also made use of the {\sc montage} software maintained by NASA/IPAC,
+and the {\sc simbad}, {\sc vizier} and {\sc aladin} services
+operated at CDS, Strasbourg, France \citep{Aladin}.
+
+Our work made extensive use of
 several complementary photometric surveys.

### 2014-04-14 paper/figures/seeing/seeing.py
Check-in of the seeing distribution plots

@@ -1,0 +1,40 @@
+from astropy.io import fits
+import pylab as p
+
+d = fits.getdata('/home/gb/dev/iphas-qc/qcdata/iphas-qc.fits', 1)
+
+mask = d['is_dr2']
+
+
+for band in ['r', 'i', 'ha']:
+
+    p.figure()
+    p.minorticks_on()
+
+    n, bins, patches = p.hist(d['seeing_'+band][mask],
+                              range=[0.5, 2.5],
+                              bins=20, 
+                              #histtype='stepfilled',
+                              #cumulative=True,
+                              lw=1.0,
+                              color='#dddddd')
+
+    p.ylabel('Fields')
+    p.xlabel("PSF FWHM [arcsec]")
+    p.xlim([0.5, 2.5])
+
+    if band == 'ha':
+        label = '$\\rm H\\alpha$'
+    else:
+        label = '$'+band+'$'
+
+    p.text(0.87, 0.9, label,
+        horizontalalignment='center',
+        verticalalignment='top',
+        transform=p.axes().transAxes,
+        fontsize=16)
+
+    p.tight_layout(pad=0.5)
+    p.savefig('seeing_{0}.pdf'.format(band))
+    p.ylim(0, 2500)
+

### 2014-04-14 paper/iphas-dr2.tex
Clarified that seeing really means PSF FWHM

@@
 The flux of each source is measured using both
-the peak pixel height (i.e. a square 0.33$\arcsec\times$0.33\arcsec\ aperture)
+the peak pixel height
+(i.e. a square 0.33$\arcsec\times$0.33\arcsec\ aperture)
 and a series of circular apertures of increasing diameter 
@@
 for all quality-approved fields,
-showing a mean depth of 
+showing mean depths of 
 $21.2\pm0.5$ ($r$), $20.0\pm0.3$ ($i$) and $20.3\pm0.3$ (\ha).
@@
 
-The minimum depth criterion
-led us to exclude 9\% of the pipelined data,
-which had typically been obtained under cloudy conditions.
-We note that some of the excluded data may nevertheless be useful
-for e.g. time-domain studies of bright stars.
-The detection tables of the discarded data are made
-available through our website for this reason (www.iphas.org),
-but are ignored in what follows.
-
 (2) \emph{Ellipticity.} 
@@
 with $b$ the semi-minor and $a$ the semi-major axis,
-is a morphological measure of the elongation of the point spread function.
+is a morphological measure of the elongation of the point spread function (PSF).
 It is expected to be zero (circular) across the field 
@@
 but it is slightly non-zero in any real telescope data 
-due to optical distortions and tracking errors.  Indeed, it is worth noting
-that IPHAS data have been collected from unguided exposures that rely
-entirely on the INT's tracking capability.
+due to optical distortions and tracking errors.
+Indeed, it is worth noting
+that IPHAS data have been collected
+from unguided exposures that rely entirely
+on the INT's tracking capability.
 The mean ellipticity measured
-%across a field 
-from exposures forming this release
-%in the IPHAS data set 
-is $0.09\pm0.04$.
+in the data is $0.09\pm0.04$.
 There have been sporadic episodes with higher ellipticities
@@
 To exclude these, we rejected exposures in which the mean ellipticity
-across the detector exceeded $e > 0.3$. We have also found that, as
-this threshold is breached, the photometric measurements delivered by 
-the pipeline become degraded.
-Only 0.4\% of the exposures were discarded on this basis.
+across the detector exceeded $e > 0.3$.
+We have also found that, as this threshold is breached,
+the photometric measurements delivered by the pipeline become degraded.
 
@@
 The original survey goal was to obtain data 
-at seeing better than 1.7~arcsec.
+at a resolution better than 1.7~arcsec,
+as evaluated by measuring the average PSF Full Width at Half Maximum (FWHM)
+across an image,
+which is a proxy for the combined effect of atmospheric and dome seeing.
 This target is currently attained across 86\% of the footprint,
@@
 e.g. 92\% of the fields at $l<120^\circ$ are better than 1.7~arcsec.
+To slightly increase the sky area offered by this data release,
+we have decided to include data obtained under seeing of up to 2.5 arcsec.
 Fig.~\ref{fig:seeing} presents the distribution
-of the mean seeing for all the quality-approved fields.
-We find a median value of 1.1~arcsec in $r$, \ha\
-and 1.0~arcsec in $i$.
+of the PSF FWHM for the approved fields in the release.
 In the $r$-band, 90\% of the data is better than 1.5~arcsec,
+50\% is better than 1.1~arcsec,
 and 10\% is better than 0.8~arcsec.
-To increase the sky area covered by this data release,
-we have decided to include the small fraction of data
-obtained under seeing of up to 2.5 arcsec.  This limited
-exclusion on grounds of poor seeing to only 1\%\ of the pipelined 
-exposures. 
 In \S\ref{sec:catalogue} we will explain
-that the photometry listed in the catalogue
-is predominantly based on the images with the
-best-available seeing, and that the seeing estimate
+that the photometry compiled in the source catalogue
+is normally derived from the field with the
+best-available seeing for a given object,
+and that the PSF FWHM measurement
 is included as a column in the catalogue.
@@
     \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/seeing_r.pdf} 
+        \includegraphics[width=\textwidth]{figures/seeing/seeing_r.pdf} 
     \end{minipage}
     \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/seeing_i.pdf} 
+        \includegraphics[width=\textwidth]{figures/seeing/seeing_i.pdf} 
     \end{minipage}
     \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/seeing_ha.pdf} 
+        \includegraphics[width=\textwidth]{figures/seeing/seeing_ha.pdf} 
     \end{minipage}
-    \caption{Seeing distribution across all
-             quality-approved survey fields
-             for $r$ (top), $i$ (middle) and \ha\ (bottom).}
+    \caption{Distribution of the PSF FWHM
+             for all the quality-approved fields
+             included in the release,
+             measured in $r$ (top), $i$ (middle) and \ha\ (bottom).
+             The PSF FWHM measures the effective image resolution
+             that arises from the combination of atmospheric and dome seeing.}
     \label{fig:seeing}
@@
 ensures that every pointing is immediately followed 
-by an offset pointing at a displacement of $+$5~arcmin in Dec 
-and $+$5~arcmin in RA.
+by an offset pointing at a displacement
+of $+$5~arcmin in Dec and $+$5~arcmin in RA.
 This allows pairs of images to be checked 
@@
 the images and photometry by eye.
-We rejected 11\% of the data as part of this step
-(many of which had already been rejected as part of the limiting magnitude criterion).
 
@@
 of the data and to check for unexpected artefacts.
-%(e.g. bright satellite trails)
 In particular, we created density maps
@@
 This was particularly effective for revealing
-fields with an inaccurate astrometric solution in one of the bands.
-In the majority of the cases we were able to correct the astrometry
-of such fields by hand.
+fields with an inaccurate astrometric solution in one of the bands,
+which were subsequently corrected.
 
 (6) \emph{Visual examination.}
-All images and their associated photometric colour/magnitude diagrams
+Images, colour mosaics,
+and the associated photometric colour/magnitude diagrams
 were inspected by a team of 20 survey members, 
@@
 was looked at by at least three different pairs of eyes.
-Images deemed unsuitable were flagged, investigated and
-excluded from the release if necessary. 
-6\% of the observations were placed on a \emph{black-list}
-as a result of this -- most commonly because of the obvious 
-presence of clouds or extreme levels of scattered moonlight.
+Images affected by clouds or extreme levels of scattered moonlight
+were flagged, investigated,
+and excluded from the release 
+by placing them on a \emph{black-list}.
 
@@
 are observed contemporaneously --  
-typically within 5 minutes of each other.
+normally within 5 minutes of each other.
 An exception was made for 9 fields where the three exposures 
@@
 for 14\,115 out of the 15\,270 fields (92\%).
-
 In some cases more than one successful attempt to observe
@@
 to deliver the most reliable measurement at a single epoch.
-Those interested in any of the rejected data 
-can nevertheless data-mine the full set of detection tables 
-via our website.
+
+We note that some of the excluded data may nevertheless be useful
+for e.g. time-domain studies of bright stars.
+The images and detection tables of the discarded data are made
+available through our website for this reason (www.iphas.org),
+but will be ignored in the remainder of this work.
 
@@
    The object must appear as a perfect point source,
-   as inferred from comparing its Point Spread Function (PSF)
+   as inferred from comparing its PSF
    with the average PSF measured in the same CCD. \\

### 2014-04-14 paper/figures/diagrams/plot.py
Check-in of example colour/magnitude diagrams

@@ -1,0 +1,217 @@
+"""Plots example colour/magnitude diagrams along different sightlines."""
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+from matplotlib.ticker import MultipleLocator
+from matplotlib import colors
+import numpy as np
+from astropy.io import ascii
+from astropy import log
+
+from dr2.sql import sql
+db = sql.SourceCatalogueDB()
+
+COLORMAP = colors.LinearSegmentedColormap.from_list('mymap', 
+    ['#bd0026', '#f03b20', '#fd8d3c', '#fecc5c', '#ffffb2'])
+
+# Tracks from Drew et al. 2005
+drew2005_a0 = {'rmi': [0.029, 0.699, 1.352, 1.991, 2.616],
+               'rmha': [0.001, 0.199, 0.355, 0.468, 0.544]}
+
+drew2005_dwarfs = {'ebv0': {
+                    'rmi': [-0.176, 0.029, 0.212,
+                            0.445, 0.650, 0.903, 1.144,
+                            1.521], #, 1.829], #, 2.514],
+                    'rmha': [0.067, 0.001, 0.114,
+                             0.278, 0.390, 0.499, 0.624,
+                             0.820]} #, 0.889]}#, 1.060]}
+                    }
+
+drew2005_giants = {'ebv0': {
+                    'rmi': [-0.180, 0.013, 0.177, 0.378, 
+                            0.505, 0.808, 0.935, 1.040,
+                            1.311, 1.652], #, 1.941],
+                    'rmha': [0.075, 0.022, 0.103, 0.236,
+                             0.284, 0.381, 0.473, 0.512,
+                             0.564, 0.644]}, #, 0.725]},
+                   'ebv1': {
+                    'rmi': [0.489, 0.682, 0.844, 1.035,
+                            1.157, 1.456, 1.581, 1.679,
+                            1.957, 2.296, 2.581],
+                    'rmha': [0.282, 0.220, 0.290, 0.408,
+                             0.445, 0.525, 0.611, 0.641,
+                             0.687, 0.753, 0.821]},
+                   'ebv2': {
+                    'rmi': [1.142, 1.335, 1.497, 1.677,
+                            1.793, 2.088, 2.212, 2.304,
+                            2.586, 2.924, 3.205],
+                    'rmha': [0.446, 0.375, 0.435, 0.537,
[…]

### 2014-04-14 paper/iphas-dr2.tex
Changes based on feedback by SG, JE and DS.

@@
 \def\ha{\mbox{H$\rm \alpha$}}
-\def\arcsec{$''$}
-\def\arcmin{$'$}
-\def\deg{$^{\circ}$}
 \def\micron{\mbox{$\mu$m}}
@@
 M. J. Barlow$^{8}$,
+P. J. Carter$^{12}$,
 R. L. M Corradi$^{6}$,
 J. J. Drake$^{9}$,
-J. Eisl\"offel$^{10}$,
-J. Fabregat$^{11}$, \newauthor
-B. T. Gaensicke$^{12}$,
-A. S. Hales$^{13}$,
+J. Eisl\"offel$^{10}$, \newauthor
+J. Fabregat$^{11}$, 
+B. T. G\"ansicke$^{12}$,
+N. P. Gentile Fusillo$^{12}$,
+S. Greiss$^{12}$,
+A. S. Hales$^{13}$, \newauthor
+S. Hodgkin$^2$, 
 J. Irwin$^{9}$,
@@
 T. Kupfer$^{5}$,
-D. J. Lennon$^{15}$, \newauthor
-J. R. Lewis$^{2}$, 
+D. J. Lennon$^{15}$,
+J. R. Lewis$^{2}$,  \newauthor
 M. Mohr-Smith$^{1}$, 
@@
 T. Naylor$^{17}$,
-Q. A. Parker$^{18}$, \newauthor
-S. Phillipps$^{13}$, 
+Q. A. Parker$^{18}$, 
+S. Phillipps$^{13}$, \newauthor
 R. Raddi$^{11}$, 
@@
 S. Scaringi$^{20}$,
-D. Steeghs$^{11}$, \newauthor
-Y. C. Unruh$^{21}$, 
+D. Steeghs$^{11}$, 
+Y. C. Unruh$^{21}$, \newauthor
+J. van Roestel$^5$,
 K. Viironen$^{22}$, 
@@
 Cambridge, MA 02138, U.S.A. \\
-$^{10}$Th\"uringer Landessternwarte, Sternwarte 5, 07778, Tautenburg, Germany \\
-$^{11}$Observatorio Astr\'onomico, Universidad de Valencia,
+$^{10}$Th\"uringer Landessternwarte, Sternwarte 5, D-07778, Tautenburg, Germany \\
+$^{11}$Observatorio Astron\'omico, Universidad de Valencia,
 Catedr\'atico Jos\'e Beltr\'an 2, 46980 Paterna, Spain\\
@@
 is a 1800~deg$^2$ imaging survey
-covering Galactic latitudes $|b| < 5^\circ$
-and longitudes $l=30$ to $215^\circ$
+covering Galactic latitudes $|b| < 5\degr$
+and longitudes $\ell=30$ to 215\degr
 in the $r$, $i$ and \ha\ filters 
@@
 We present the first quality-controlled and
-uniformly-calibrated source catalogue
+globally calibrated source catalogue
 derived from the survey,
@@
 The observations were carried out between 2003 and 2013
-at a median seeing of 1.1 arcsec
+at a median seeing of 1.1~arcsec (sampled at 0.33~arcsec/pixel)
 and to a mean $5\sigma$-depth of 
-21.2 ($r$), 20.0 ($i$) and 20.3 ($\ha$)
+21.2~($r$), 20.0~($i$) and 20.3~($\ha$)
 in the Vega magnitude system.
@@
 and quality control procedures,
-describe and test the new uniform calibration,
+describe and test the global re-calibration,
 and detail the construction of the new catalogue.
-We show that the global calibration is accurate to
+We show that the new calibration is accurate to
 0.03~mag (rms) 
@@
 towards different Galactic sight-lines
-and (ii) select candidate \ha\ emission-line objects.
+and (ii) select \ha\ emission-line objects.
 IPHAS is the first survey to offer comprehensive CCD photometry
@@
 of the early and late stages of stellar evolution.
-Since the publication of the IPHAS Initial Data Release \citep[IDR;][]{Gonzalez-Solares2008},
+Since the publication of the Initial Data Release \citep[IDR;][]{Gonzalez-Solares2008},
 these aims have begun to be realised through a
@@
 discoveries of new symbiotic stars \citep{Corradi2008, Corradi2010, Corradi2011}; 
-new cataclysmic variables \citep{Witham2007}; 
+new cataclysmic variables \citep{Witham2007,Wesson2008,Aungwerojwit2012}; 
 new groups of young stellar objects
@@
 it is vital that the imaging data are transformed 
-into a uniformly-calibrated photometric catalogue, 
+into a homogeneously-calibrated photometric catalogue, 
 in which quality problems 
@@
 that takes the coverage up to 92 percent of the survey area 
-and includes a uniform calibration.
+and includes a global photometric re-calibration.
 In this work we
@@
 and quality control procedures that were applied,
-(ii) describe and test the new global photometric calibration, and 
+(ii) describe and test the new photometric calibration, and 
 (iii) detail the construction of the source catalogue
@@
 and quality control procedures.
-In \S\ref{sec:calibration} we explain the uniform re-calibration, in which
+In \S\ref{sec:calibration} we explain the global re-calibration, in which
 we draw upon the AAVSO Photometric All-Sky Survey (APASS)
@@
 of roughly 180 by 10 degrees, 
-covering Galactic latitudes $-5^{\circ} < b < +5^{\circ}$ 
-and longitudes $30^{\circ} < l < 215^{\circ}$.
+covering Galactic latitudes $-5\degr < b < +5\degr$ 
+and longitudes $30\degr < \ell < 215\degr$.
 
@@
 At that time only 76\% of the field pairs 
-satisfied our minimum quality criteria, however.  The problems affecting
-the 24\% falling below survey standard were, most commonly: variable cloud cover; 
-poor seeing; technical faults (the quality criteria will be detailed in the 
-next section). 
+satisfied our minimum quality criteria, however.
+The problems affecting the 24\% falling below survey standard were,
+most commonly: variable cloud cover;
+poor seeing; technical faults
+(the quality criteria will be detailed in the next section).
 Since then, a programme of repeat observations has been in place 
@@
         or in light grey (denoting that only one offset is missing).
-        The dark vertical strip near $l \sim$125\deg
+        The dark vertical strip near $\ell \sim$125\degr
         is an arbitrary consequence of the tiling pattern,
@@
 are predominantly located towards the Galactic anti-centre 
-at $l > 120^o$.
+at $\ell > 120\degr$.
 Fields at these longitudes are mainly accessed from La Palma 
@@
 by \citet{Drew2005} and \citet{Gonzalez-Solares2008},
-and much of the source code is available 
-on line\footnote{http://casu.ast.cam.ac.uk/surveys-projects/software-release}. 
+and the source code is available
+on-line\footnote{http://casu.ast.cam.ac.uk/surveys-projects/software-release}. 
 In brief, the pipeline takes care of bias subtraction,
@@
 the peak pixel height
-(i.e. a square 0.33$\arcsec\times$0.33\arcsec\ aperture)
+(i.e. a square 0\farcs 33$\times$0\farcs 33 aperture)
 and a series of circular apertures of increasing diameter 
-(1.2\arcsec, 2.3\arcsec, 3.3\arcsec, 4.6\arcsec\ and 6.6\arcsec).
+(1\farcs 2, 2\farcs 3, 3\farcs 3, 4\farcs 6 and 6\farcs 6).
 
@@
 from measuring the source intensity in a series of growing apertures.
-Finally, the resulting source detection tables are also stored 
+Finally, the resulting source detection tables are stored 
 in multi-extension FITS files.
@@
 in which a total of 1.9~billion \emph{candidate sources} were detected 
-at the sensitive default detection level of 1.25\,$\sigma$.
+at the default detection level of 1.25\,$\sigma$.
 This total inevitably includes spurious detections, artefacts and
-duplicate detections,
+duplicate detections;
 in \S\ref{sec:catalogue} we will explain
@@
 
-Observing time for IPHAS was obtained
+Observing time for IPHAS is obtained
 on a semester-by-semester basis
 through the open time allocation committees 
-of the Isaac Newton Group of telescopes,
-which are invariably over-subscribed.
-For this reason, we attempted to utilise 
-\emph{all} the nights allocated to IPHAS,
-even those which were partially or entirely non-photometric
-or otherwise affected by technical problems 
-(e.g. electronic noise or telescope tracking problems).
-Any unsuitable data that were taken as a result
-of this strategy have subsequently been flagged and removed
+of the Isaac Newton Group of telescopes.
+The survey is allocated specific observing dates
+rather than particular observing conditions.
+In consequence, data were acquired
+under a large range of atmospheric conditions.
+Data taken under unsuitable conditions
+have been flagged and removed
 using seven quality criteria,
@@
              to the presence of the moon above the horizon: 
-             this is the reason for the wide, bi-modal character
+             this is the main reason for the wide, bi-modal character
              of its distribution.}
@@
 Fig.~\ref{fig:depth} presents the distribution of limiting magnitudes
-for all quality-approved fields,
-showing mean depths of 
+for all quality-approved fields;
+the mean depths and standard deviations are 
 $21.2\pm0.5$ ($r$), $20.0\pm0.3$ ($i$) and $20.3\pm0.3$ (\ha).
@@
 which was above the horizon during 62\% 
-of the observations.  The great range in sky brightness this
-produced is the reason for the wide and bi-modal shape
+of the observations.
+The great range in sky brightness this
+produced is the main reason for the wide and bi-modal shape
 of the $r$-band limiting magnitude distribution 
@@
 (middle and bottom panels in Fig.~\ref{fig:depth}).
+To a lesser extent, the wide spread in the $r$-band depth
+is also explained by the shorter $r$-band exposure time
+that was used in the first year of data-taking.
 
@@
 To exclude these, we rejected exposures in which the mean ellipticity
-across the detector exceeded $e > 0.3$.
+\emph{across the detectors} exceeded $e > 0.3$.
 We have also found that, as this threshold is breached,
@@
 as evaluated by measuring the average PSF Full Width at Half Maximum (FWHM)
-across an image,
+across the detectors,
 which is a proxy for the combined effect of atmospheric and dome seeing.
@@
 in particular at lower Galactic longitudes,
-e.g. 92\% of the fields at $l<120^\circ$ are better than 1.7~arcsec.
-To slightly increase the sky area offered by this data release,
-we have decided to include data obtained under seeing of up to 2.5 arcsec.
+e.g. 92\% of the fields at $\ell<120\degr$ are better than 1.7~arcsec.
+To increase the sky area offered by the present release slightly,
+we have decided to accept data obtained with seeing up to 2.5~arcsec.
 Fig.~\ref{fig:seeing} presents the distribution
-of the PSF FWHM for the approved fields in the release.
-In the $r$-band, 90\% of the data is better than 1.5~arcsec,
+of the PSF FWHM for the approved fields.
+In the $r$-band, 90\% is better than 1.5~arcsec,
 50\% is better than 1.1~arcsec,
-and 10\% is better than 0.8~arcsec.
+and 10\% is better than 0.8~arcsec
+(the WFC pixel scale is 0.33~arcsec/px).
 In \S\ref{sec:catalogue} we will explain
@@
 Having obtained a quality-approved set of observations,
-we now turn to the problem of placing the data
+we now turn to the challenge of placing the data
 onto a uniform photometric scale.
@@
 \textsc{apcor} is a correction for the flux
-lost outside of the aperture used.
+lost outside of the aperture used
+(we adopt a 2.3~arcsec diameter circular aperture by default).
 Finally, \textsc{percorr} is a correction based on the difference
@@
 and as such is an ancillary correction 
-to account for sporadic gain variations. 
+to account for sporadic gain variations;
+it is zero for most exposures.
 All these quantities correspond to header keywords in the 
@@
 
-\subsection{Uniform re-calibration}
+\subsection{Global re-calibration}
 
@@
 Surveys which observe identical stars at different epochs
-can use the repeat measurements to ensure a uniform calibration.
+can use the repeat measurements to ensure a homogeneous calibration.
 For example, 2MASS attained its global calibration
@@
 
-Although the IPHAS data do contain a significant number of 
-such inter-field repeat measurements,
-we have found the information contained
-in these regions to be insufficient
+We have found the information contained
+in these overlap regions to be insufficient
 to constrain the calibration parameters.
@@
 -- where available --
-to bring the majority of our data onto a uniform calibration.
+to bring the majority of our data onto a homogeneous calibration.
 
@@
 This was achieved for 48\% of our fields (shown in Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}).
-The mean offset was found to be
-$0.014\pm0.104$~mag in $r$ and $0.007\pm0.108$~mag in $i$
+The offsets follow a near-Gaussian distribution
+centred around $0.014\pm0.104$~mag in $r$ and $0.007\pm0.108$~mag in $i$
 (Table~\ref{tbl:offsets_before}).
-A total of 4596 fields showed a median offset
+A total of 4596 fields showed a systematic offset
 exceeding $\pm$0.03~mag in either $r$ or $i$.
@@
     \caption{Median magnitude offset between IPHAS and SDSS in the $r$ band
-             \emph{after} the re-calibration procedure was applied.
+             \emph{after} the re-calibration procedure using APASS was applied.
              Each square represents the footprint of an IPHAS field
@@
 \begin{table}
-    \caption{Mean magnitude offsets for objects
+    \caption{Magnitude offsets for objects
              cross-matched between IPHAS and APASS/SDSS
-             \emph{before} the uniform re-calibration was carried out.
-             Transformations were applied
+             \emph{before} the global re-calibration was carried out.
+             We characterise the distribution of the offsets,
+             which is approximately Gaussian in each case,
+             by listing the mean and the standard deviation values.
+			 We remind the reader that transformations were applied
              to the APASS and SDSS magnitudes to bring them into the
-             Vega-based IPHAS system prior to computing
-             these statistics (see text).
+             Vega-based IPHAS system prior to computing the offsets.
              }
@@
        \end{tabular}
-       \caption{Same as Table~\ref{tbl:offsets_before}
-                but computed \emph{after} the uniform re-calibration
-                 was carried out.}
+       \vspace{1cm}
+       \caption{Same as Table~\ref{tbl:offsets_before},
+                but computed \emph{after} the global re-calibration
+                was carried out. The mean and standard deviation values
+                of the offsets have improved significantly.}
         \label{tbl:offsets_after}
@@
 
-We then turned to the uniform calibration of the \ha\ data.
+We then turned to the global calibration of the \ha\ data.
 It is not possible to re-calibrate the narrowband 
@@
 These global transformations were deemed adequate
-for the purpose of validating our uniform calibration in a statistical sense.
+for the purpose of validating our calibration in a statistical sense.
 Separate equations were derived towards different sight-lines
@@
 The vast majority of red objects in the global sample
-are those in highly reddened areas however,
+are those in highly reddened areas, however,
 which agree well with the global transformations
@@
 
-We warn that the SDSS comparison revealed a number of fields where the offsets
+The SDSS comparison revealed a number of fields where the offsets
 exceeded 0.05~mag (523 fields) or even 0.1~mag (18 fields).
@@
 and re-calibrated data set, 
-we now turn to the problem
+we now turn to the challenge
 of transforming the observations 
@@
 First, we note that each source is uniquely identified by an
-IAU-style designation of the form ``IPHAS2\ JHHMMSS.ss+DDMMSS.s''
+IAU-style designation of the form `IPHAS2\ JHHMMSS.ss+DDMMSS.s'
 (cf. column \emph{name} in Appendix~\ref{app:columns}),
-where ``IPHAS2'' refers to the present
+where `IPHAS2' refers to the present
 data release and the remainder of the string
@@
 We have also included an internal object identifier string 
-of the form ``\#run-\#ccd-\#detection''
-(e.g. ``64738-3-6473''),
+of the form `\#run-\#ccd-\#detection'
+(e.g. `64738-3-6473'),
 which documents the INT exposure number (\#run),
@@
 where the astrometric solution can 
-show systematic errors which exceed 0.5~arcsec.
+show local systematic errors which exceed 0.5~arcsec
+(even though the nominal rms accuracy
+-- estimated against 2MASS across the focal plane --
+tends to be better than 0.1~arcsec).
 The position offsets between the $r$ detection and detections in $i$, and/or \ha\
@@
 highly-reddened objects which are otherwise lost in $r$ and \ha.
+Moreover, late-type stars tend to be intrinsically brighter in $i$.
 
@@
 which should benefit most applications
-which desire reliable $(r-i)$ and $(r-\ha)$ colours
+which desire reliable colours
 as well as completeness.
@@
 (39\%) satisfy all the criteria listed in Table~\ref{tab:reliable}
-and are hereafter referred to as ``reliable''.
+and are hereafter referred to as `reliable'.
 For convenience, the catalogue contains a boolean column
@@
   Require the $r$ magnitude measured 
-  in the default 2.3\arcsec\ diameter aperture
+  in the default 2\farcs 3-diameter aperture
   to be consistent with the measurement 
-  made in the smaller 1.2\arcsec\ aperture,
+  made in the smaller 1\farcs 2 aperture,
   albeit tolerating a 0.03 mag systematic error.
@@
 these stricter criteria 
-and are hereafter referred to as ``very reliable''.
+and are hereafter referred to as `very reliable'.
 Again, the catalogue contains a boolean column
@@
 Fig.~\ref{fig:sourcecount} shows the number of sources
-in the catalogue counted in $1^\circ$-wide strips
+in the catalogue counted in 1\degr-wide strips
 as a function of Galactic longitude (thick blue line).	
@@
 to increase towards the Galactic centre.
-For example, the average source density near $l\simeq 30^\circ$
+For example, the average source density near $\ell\simeq 30\degr$
 is roughly 300\,000 objects per square degree,
 which is six times more than the density
-found near $l\simeq 180^\circ$.
+found near $\ell\simeq 180\degr$.
 In addition to the global trend,
@@
 For example, we find a significant drop near the constellations 
-of Aquila ($l\simeq40$) and Cygnus ($l\simeq80$ and $l\simeq90$),
+of Aquila ($\ell\simeq40$) and Cygnus ($\ell\simeq80$ and $\ell\simeq90$),
 which are regions known to be affected
@@
 to be released or for variations in the depth across the included fields.
-For example, the dip in the density near $l\simeq140^\circ$
+For example, the dip in the density near $\ell\simeq140\degr$
 is an artificial feature caused by gaps
@@
     The upper blue line shows the number of sources
-    counted in 1\deg-wide longitude bins.
+    counted in 1\degr-wide longitude bins.
     The lower red line uses the same binning
@@
     In both cases only counted sources
-    in the latitude range $-5^\circ<b<+5^\circ$,
+    in the latitude range $-5\degr<b<+5\degr$,
     such that one may obtain approximate source densities
@@
 density.  In hard numbers, only 11\% of the sources are blended
-at $l>90^\circ$, whereas 24\% are blended at $l<90^\circ$.
+at $\ell>90\degr$, whereas 24\% are blended at $\ell<90\degr$.
 
@@
 located at Galactic longitudes 
-$180^\circ$, $45^\circ$ and $30^\circ$, respectively.
+180\degr, 45\degr and 30\degr, respectively,
+which were chosen because they show different characteristics.
 Each figure contains all the sources
@@
 centred on the coordinates indicated in the diagram
-(i.e. within a radius of $0.564^\circ$ from the indicated sight-line).
+(i.e. within a radius of 0\fdg 564 from the indicated sight-line).
 For clarity, we have imposed the additional criterion
@@
 must be smaller than 0.05 mag in each band,
-corresponding to a magnitude limit near $\sim$19th magnitude.
-
-Each of the diagrams reveal a well-defined locus,
-which helps to demonstrate the health of the catalogue and the 
-global calibration
+corresponding to a cut-off near 19th magnitude.
+
+Each of the diagrams reveals a well-defined locus,
+which helps to demonstrate the health of the catalogue
+and the global calibration
 for investigating stellar populations across wide areas.
@@
 which is made available for the IPHAS system through the
-on line tool hosted at the Observatory of Padova
+on-line tool hosted at the Observatory of Padova
 (http://stev.oapd.inaf.it/cmd).
@@
 
-Each of the sightlines reveal a stellar population
+Each of the sightlines reveals a stellar population
 with distinct characteristics.
 Towards the Galactic anti-centre 
-at $l=180^\circ$ (Fig.~\ref{fig:l180})
+at $\ell=180\degr$ (Fig.~\ref{fig:l180})
 we find a population dominated by lowly-reddened main sequence stars.
@@
 of the corresponding phase of stellar evolution,
-would only account for a small proportion of the more nearly volume 
-limited sample seen in the Anticentre direction.
-
-In contrast, 
-closer towards the galactic centre
-at $l=45^{\circ}$ (Fig.~\ref{fig:l45}),
-there is  a wealth of reddened objects.
+would only account for a small proportion of the more nearly \emph{volume-limited} sample seen in the Anticentre direction.
+
+In contrast, closer towards the galactic centre,
+we encounter samples which are \emph{magnitude-limited} instead.
+For example, at $\ell=45\degr$ (Fig.~\ref{fig:l45}),
+there is  a wealth of reddened objects and late type giants.
 In the colour-magnitude diagram, it is clear
@@
 the red giants we observe will be subject to appreciably more reddening
-than the main sequence stars, pushing them to $r - i \sim 1.5$.
-
-Finally, in one of our earliest sight-lines at $l=30^{\circ}$,
+than the main sequence stars, pushing them to $(r - i) \sim 1.5$.
+
+Finally, in one of our earliest sight-lines at $\ell=30\degr$,
 we find a very high number of extremely reddened giants
@@
 of foreground dwarfs.
-In contrast to the sight-line at $l=45^{\circ}$,
+In contrast to the sight-line at $\ell=45\degr$,
 there is no clear group of giant stars visible
@@
 although the red clump stars are manifest as a track of slight 
-over density sitting roughly 0.4~mags redder than the A0V reddening track.
-At $(l,b)=(45^{\circ}, +2^{\circ})$ the giant stars observed
+over-density sitting roughly 0.4~mags redder than the A0V reddening track.
+At $(l,b)=(45\degr, +2\degr)$ the giant stars observed
 exhibit a relatively narrow range of reddenings
 as they lie beyond most or essentially all of the Galactic dust.
-At $(l,b)=(30^{\circ}, 0^{\circ})$ this is not the case:
+At $(l,b)=(30\degr, 0\degr)$ this is not the case:
 even at the substantial distances
@@
     \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=0.5\textwidth]{./plots/ccd-180-3.pdf} 
-        \includegraphics[width=0.5\textwidth]{./plots/cmd-180-3.pdf}
+        \includegraphics[width=0.5\textwidth]{figures/diagrams/ccd-180-3.pdf} 
+        \includegraphics[width=0.5\textwidth]{figures/diagrams/cmd-180-3.pdf}
     \end{minipage}
@@
     showing sources flagged as \emph{veryReliable}
-    located in an area of one square degree
+    located in an area of 1~deg$^2$
     centred near the Galactic anti-centre 
-    at $(l,b)=(180^\circ,+3^\circ)$.
+    at $(l,b)=(180\degr,+3\degr)$.
     The diagrams are plotted as 2D-histograms
     which show the density of objects
-    in bins of $\sim0.01\times0.01$~mag;
-    bins containing 1-10 objects are coloured red,
-    bins with 10-20 objects are orange,
-    and bins with 20-25 objects are yellow.
+    in bins of 0.01-by-0.01~mag;
+    bins containing 1 to 10 objects are coloured red,
+    bins with 10 to 20 objects are orange,
+    and bins with 20 to 25 objects are yellow.
     The left panel is annotated with
@@
     giant stars (thick solid line)
-    and the reddening track for an A0V-type star (dashed line),
-    which are based on the \citet{Pickles1998} library
-    of observed spectra.
-    The right panel only shows the reddening vector
-    along with the unreddened 1~Gyr isochrone due to \citet{Bressan2012},
-    which has been placed at an arbitrary distance of 2~kpc.
+    and the reddening track for an A0V-type star (dashed line). 
+    The right panel shows the unreddened 1~Gyr isochrone
+    from the models by \citet[][solid line]{Bressan2012}
+    along with the reddening vector for an A0V-type (dashed line),
+    both placed at an arbitrary distance of 2~kpc.
     This is one of the least reddened sight-lines
@@
     \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=0.5\textwidth]{./plots/ccd-45-2.pdf}
-        \includegraphics[width=0.5\textwidth]{./plots/cmd-45-2.pdf}
+        \includegraphics[width=0.5\textwidth]{figures/diagrams/ccd-45-2.pdf}
+        \includegraphics[width=0.5\textwidth]{figures/diagrams/cmd-45-2.pdf}
     \end{minipage}
-    \caption{Same as above for $(l,b)=(45^\circ,+2^\circ)$,
+    \caption{Same as above for $(l,b)=(45\degr,+2\degr)$,
     which is one of the highest-density sight-lines in the survey,
@@
     \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=0.5\textwidth]{./plots/ccd-30-0.pdf}
-        \includegraphics[width=0.5\textwidth]{./plots/cmd-30-0.pdf} 
+        \includegraphics[width=0.5\textwidth]{figures/diagrams/ccd-30-0.pdf}
+        \includegraphics[width=0.5\textwidth]{figures/diagrams/cmd-30-0.pdf} 
     \end{minipage}
-    \caption{Same as above for $(l,b)=(30^\circ,0^\circ)$,
+    \caption{Same as above for $(l,b)=(30\degr,0\degr)$,
     showing one of the most reddened sight-lines in the survey.
@@
 presented by \cite{Witham2008}. 
-Because no uniform calibration was available
+Because no global calibration was available
 at the time, \citeauthor{Witham2008} employed 
@@
 The target of our demonstration is Sh 2-82:
-a 5~arcmin-wide H{\sc ii} region located near $(l,b)=(53.55^\circ, 0.00^\circ)$
+a 5~arcmin-wide H{\sc ii} region located near $(l,b)=(53.55\degr, 0.00\degr)$
 in the constellation of Sagitta.
@@
 \begin{figure*}
-    \begin{minipage}[b]{0.78\linewidth}
+    \begin{minipage}[b]{0.8\linewidth}
         \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-iphas.pdf} 
@@
 which have been selected from the colour-colour diagram
-in Fig.~\ref{fig:emitters}. Note that the H{\sc ii} region is surrounded by a faint blue/green reflection nebula
-and dark cloud filaments.}
+in Fig.~\ref{fig:emitters}. 
+%The H{\sc ii} region is surrounded by a very faint blue/green reflection nebula and dark cloud filaments.
+}
 \label{fig:mosaic_iphas}
-    \begin{minipage}[b]{0.78\linewidth}
+    \begin{minipage}[b]{0.8\linewidth}
         \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-spitzer.pdf} 
@@
     This structure has previously been labelled as N115 in the 
-catalogue due to \citet{Churchwell2006}, and could be a possible site of triggered star formation \citep{Thompson2012,Kendrew2012}.}
+catalogue of \citet{Churchwell2006}, and could be a possible site of triggered star formation \citep{Thompson2012,Kendrew2012}.}
     \label{fig:mosaic_spitzer}
@@
 the 20-by-15 arcmin region shown in the mosaics.
-Gray circles show all objects
+Grey circles show all objects
 which are brighter than $r<20$
@@
 of point sources across the Northern Galactic Plane at visible wavelengths,
-taking in the latitude range $|b|<5^\circ$.
+taking in the latitude range $|b|<5\degr$.
 The new 98-column catalogue provides single-epoch photometry
 across 92\% of the survey area,
-and is the first quality-controlled and uniformly calibrated
+and is the first quality-controlled and globally calibrated
 catalogue to have been constructed from the imaging data.  This
@@
 achieve a median seeing of 1.1 arcsec
-and $5\sigma-$depths of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$, and \ha$=20.3\pm 0.3$.
+and $5\sigma$-depths of $r=21.2\pm 0.5$, $i=20.0\pm 0.3$, and \ha$=20.3\pm 0.3$.
 The global calibration and photometric repeatability
@@
 (grants: GB and JED ST/J001333/1, SES ST/K00106X/1, BTG ST/I001719/1). 
-HJF is in receipt of an STFC postgraduate studentship.  JED would also
-like to convey her thanks to the Physics Department of Imperial College
-London that hosted this project from its inception to 2007 and supported
-her via a sabbatical year in 2003-4.  BTG acknowledges funding from the
-European Research Council under the European Union's Seventh Framework
-Programme (FP/2007-2013) / ERC Grant Agreement n. 320964 (WDTracer).
+HJF is in receipt of an STFC postgraduate studentship.
+JED would also like to convey her thanks
+to the Physics Department of Imperial College London
+that hosted this project from its inception to 2007 and supported
+her via a sabbatical year in 2003-4.
+JF is supported by the Spanish Plan Nacional de I+D+i and FEDER under contract AYA2010-18352.
+BTG acknowledges funding from the European Research Council
+under the European Union's Seventh Framework Programme (FP/2007-2013) 
+/ ERC Grant Agreement n. 320964 (WDTracer).
 NJW is in receipt of a Fellowship funded by the Royal Astronomical Society

### 2014-04-14 paper/iphas-dr2.tex
Changes based on feedback by BG.

@@
 \author[G. Barentsen
-et. al]{G. Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
+et. al]{Geert Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
 H. J. Farnhill$^1$,
@@
 -- spanning almost a decade 
-and using more than 300 nights 
+and using more than 300 nights in competitive open time 
 at the Isaac Newton Telescope (INT) in La Palma --
@@
 that captures a sky area of close to 0.29~deg$^2$.
-To cover the entire Northern Plane with some overlap,
+To cover the Northern Plane with some overlap,
 the survey area was divided into 7635 telescope pointings.
@@
 
-Fig.~\ref{fig:footprint} shows the footprin
+Fig.~\ref{fig:footprint} shows the footprint
 of the quality-approved observations included in this work. 
@@
 the CASU pipeline had processed
-74\,195 IPHAS exposures 
+74\,195 single-band IPHAS exposures 
 in which a total of 1.9~billion \emph{candidate sources} were detected 
@@
 and excluded from the release 
-by placing them on a \emph{black-list}.
+by placing them on a black-list.
 
@@
 normally within 5 minutes of each other.
-An exception was made for 9 fields where the three exposures 
-could not be obtained within the same night
-but for which the time gap between exposures did not exceed 48 hours.
+%An exception was made for 9 fields where the three exposures 
+%could not be obtained within the same night
+%but for which the time gap between exposures did not exceed 48 hours.
 We note that the source catalogue details the exact epoch
@@
 large surveys naturally possess field-to-field variations
-at the level of 0.1 mag
 due to atmospheric changes during the night
@@
 on the solution by keeping the zeropoint fixed 
-for the 6\,756 fields which have already been compared
+for the 
+% 6\,756 
+fields which have already been compared
 and calibrated against APASS.
@@
 in the earlier steps.
-An exception was made for 3\,101 fields
+An exception was made for a small subset of fields
+% 3\,101 fields
 for which our quality-control routines revealed
-zeropoint variations which exceeded 0.03~mag 
-between consecutive fields,
-suggesting strongly non-photometric conditions.
-In these cases, the \ha\ zeropoints of these fields
+strong zeropoint variations during the night,
+suggesting non-photometric conditions.
+In these cases, the \ha\ zeropoints
 were adjusted by solving Eqn.~\ref{eqn:matrix}
-rather than linking them directly to the $r$-band shift.
+rather than being linked to the $r$-band shift.
 
@@
 to show some variation towards lowly reddened areas,
-where red objects at $(r-i) > 1$ are rare.
+where different stellar populations are observed.
 The vast majority of red objects in the global sample
@@
 within 10 minutes of the primary,
-we have included the identifier and the default photometry
+we have included the identifier and the photometry
 of that secondary detection
@@
 due to the IPHAS field pair strategy.
-In addition, the complete set of detection tables -- one for each exposure -- 
-is made accessible on our website to allow other uses of the data.
-%A user-friendly catalogue of secondary detections 
-%has not been compiled at present 
-%but may be part of a future data release.
-
-%The primary detection is defined as the entry in the 
-%set of band-merged field catalogues which provides 
-%the most reliable information for a unique source.
+
 Primary detections have been selected from all available detections using a so-called \emph{seaming} procedure,
@@
 not have had any known bad pixels in the aperture,
-and not have been on the edge of one of the CCDs,
-i.e. we require the \emph{errBits} quality score
-to be smaller than 64.
+and not have been on the edge of one of the CCDs
+(this information is contained in the \emph{errBits} quality score column, which is used for internal bookkeeping).
 \end{enumerate}
@@
 are fainter than $r > 20$.
-%and 18\% are even fainter than $r > 21$.
 The measurements of faint objects
@@
   \hline
-  rErr\,$<$\,0.1 {\sc and} iErr\,$<$\,0.1 {\sc and} haErr\,$<$\,0.1 &
+  $rErr\,<\,0.1$ {\sc and} $iErr\,<\,0.1$ {\sc and} $haErr\,<\,0.1$ &
   109 million (50\%) &
@@
   \\
-  $|\mathrm{r}- \mathrm{rAperMag1}| 
-  < 3\sqrt{\mathrm{rErr}^2 + \mathrm{rAperMag1Err}^2} 
+  $|r- rAperMag1| 
+  < 3\sqrt{rErr^2 + rAperMag1Err^2} 
   + 0.03$ &
@@
   was not performed reliably. \\
-  $|\mathrm{i}- \mathrm{iAperMag1}| 
-  < 3\sqrt{\mathrm{iErr}^2 + \mathrm{iAperMag1Err}^2} 
+  $|i- iAperMag1| 
+  < 3\sqrt{iErr^2 + iAperMag1Err^2} 
   + 0.03$ &
@@
   Same as above for $i$. \\
-  $|\mathrm{ha}- \mathrm{haAperMag1}| < 
-  3\sqrt{\mathrm{haErr}^2 + \mathrm{haAperMag1Err}^2} 
+  $|ha- haAperMag1| < 
+  3\sqrt{haErr^2 + haAperMag1Err^2} 
   + 0.03$ &
@@
   \hline
-   reliable &
+   \emph{reliable} &
    86 million (39\%) &
@@
 
-   pStar $>$ 0.9 &
+   $pStar > 0.9$ &
    145 million (66\%) &
@@
     did not have to apply a deblending procedure.
-    In both cases only counted sources
+    In both cases we counted only those sources
     in the latitude range $-5\degr<b<+5\degr$,
@@
     by dividing the counts by 10~deg$^2$.
-    The apparent variations in the source counts
+    The global variations in the source counts
     traces the structure of the Galaxy
     and the distribution of foreground extinction,
-    but also includes instrumental effects
-	such as variations in the survey depth
+    but is also affected by instrumental effects
+	such as variations in the observed depth
 	and completeness (see text).
@@
 JF is supported by the Spanish Plan Nacional de I+D+i and FEDER under contract AYA2010-18352.
-BTG acknowledges funding from the European Research Council
-under the European Union's Seventh Framework Programme (FP/2007-2013) 
-/ ERC Grant Agreement n. 320964 (WDTracer).
-NJW is in receipt of a Fellowship funded by the Royal Astronomical Society
-of the United Kingdom.
+BTG acknowledges funding from the
+European Research Council under the European Union's Seventh Framework
+Programme (FP/2007-2013) / ERC Grant Agreement n. 320964 (WDTracer).
+NJW is in receipt of a Fellowship funded by 
+the Royal Astronomical Society of the United Kingdom.
 

### 2014-04-14 paper/iphas-dr2.tex
Changes based on feedback by PL, QP, RM.

@@
 P. Groot$^5$,
-A. Mampaso$^6$, 
-S. E. Sale$^7$, \newauthor
-M. J. Barlow$^{8}$,
-P. J. Carter$^{12}$,
-R. L. M Corradi$^{6}$,
-J. J. Drake$^{9}$,
-J. Eisl\"offel$^{10}$, \newauthor
-J. Fabregat$^{11}$, 
-B. T. G\"ansicke$^{12}$,
-N. P. Gentile Fusillo$^{12}$,
-S. Greiss$^{12}$,
-A. S. Hales$^{13}$, \newauthor
+A. Mampaso$^{6,7}$, 
+S. E. Sale$^8$, \newauthor
+M. J. Barlow$^9$,
+P. J. Carter$^{13}$,
+R. L. M Corradi$^{6,7}$,
+J. J. Drake$^{10}$,
+J. Eisl\"offel$^{11}$, \newauthor
+J. Fabregat$^{12}$, 
+B. T. G\"ansicke$^{13}$,
+N. P. Gentile Fusillo$^{13}$,
+S. Greiss$^{13}$,
+A. S. Hales$^{14}$, \newauthor
 S. Hodgkin$^2$, 
 J. Irwin$^{9}$,
-C. Knigge$^{14}$,
+C. Knigge$^{15}$,
 T. Kupfer$^{5}$,
-D. J. Lennon$^{15}$,
+D. J. Lennon$^{16}$,
 J. R. Lewis$^{2}$,  \newauthor
 M. Mohr-Smith$^{1}$, 
-R. A. H. Morris$^{16}$,
-T. Naylor$^{17}$,
-Q. A. Parker$^{18}$, 
-S. Phillipps$^{13}$, \newauthor
-R. Raddi$^{11}$, 
-P. Rodriguez-Gil$^{6}$,
-L. Sabin$^{19}$,
-S. Scaringi$^{20}$,
-D. Steeghs$^{11}$, 
-Y. C. Unruh$^{21}$, \newauthor
+R. A. H. Morris$^{17}$,
+T. Naylor$^{18}$,
+Q. A. Parker$^{19,20,21}$, 
+S. Phillipps$^{17}$, \newauthor
+R. Raddi$^{13}$, 
+P. Rodr\'\i guez-Gil$^{6,7}$,
+L. Sabin$^{22}$,
+S. Scaringi$^{23,24}$,
+D. Steeghs$^{13}$, \newauthor
+Y. C. Unruh$^{25}$, 
 J. van Roestel$^5$,
-K. Viironen$^{22}$, 
-J. S. Vink$^{23}$, 
-N. A. Walton$^{2}$,
+K. Viironen$^{26}$, 
+J. S. Vink$^{27}$, 
+N. A. Walton$^{2}$, \newauthor
 N. J. Wright$^{1}$,
-A. A. Zijlstra$^{24}$.
+A. A. Zijlstra$^{28}$.
 \\
@@
 $^{5}$Afdeling Sterrenkunde, Radboud Universiteit Nijmegen, Faculteit NWI, Postbus 9010, 6500 GL Nijmegen, The Netherlands.\\
-$^{6}$Instituto de Astrof\'isica de Canarias, 38200 La Laguna, Tenerife, Spain.\\
-$^{7}$Rudolf Peierls Centre for Theoretical Physics, Keble Road, Oxford, OX1 3NP, U.K.\\
-$^{8}$University College London, Department of Physics \& Astronomy, 
+$^{6}$Instituto de Astrof\'\i sica de Canarias, V\'\i a L\'actea, s/n, La Laguna, E-38205, Santa Cruz de Tenerife, Spain.\\
+$^{7}$Departamento de Astrof\'\i sica, Universidad de La Laguna, La Laguna, E-38204, Santa Cruz de Tenerife, Spain.\\
+$^{8}$Rudolf Peierls Centre for Theoretical Physics, Keble Road, Oxford, OX1 3NP, U.K.\\
+$^{9}$University College London, Department of Physics \& Astronomy, 
 Gower Street, London WC1E 6BT, U.K.\\
-$^{9}$Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, 
+$^{10}$Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, 
 Cambridge, MA 02138, U.S.A. \\
-$^{10}$Th\"uringer Landessternwarte, Sternwarte 5, D-07778, Tautenburg, Germany \\
-$^{11}$Observatorio Astron\'omico, Universidad de Valencia,
-Catedr\'atico Jos\'e Beltr\'an 2, 46980 Paterna, Spain\\
-$^{12}$Department of Physics, University of Warwick, Gibbet Hill Road, Coventry, CV4 7AL, U.K.\\
-$^{13}$Joint ALMA Observatory, Alonso de Córdova 3107, Vitacura 763-0355, Santiago, Chile. \\
-$^{14}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
-$^{15}$European Space Astronomy Centre (ESAC), Villafranca del Castillo, Villanueva de la Canada, E-28692 Madrid, Spain.\\
-$^{16}$School of Physics, Bristol University, Tyndall Avenue, Bristol, BS8 1TL, U.K.\\
-$^{17}$School of Physics, University of Exeter, Stocker Road, Exeter, EX4 4QL, U.K.\\
-$^{18}$Department of Physics \& Astronomy, Macquarie University, NSW 2109, Australia\\
-$^{19}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
-$^{20}$Instituut voor Sterrenkunde, K.U. Leuven, Celestijnenlaan 200D, B-3001 Leuven, Belgium\\
-$^{21}$Department of Physics, Blackett Laboratory, Imperial College
-London, Prince Consort Road, London, SW7 2AZ, U.K.\\
-$^{22}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
-$^{23}$Armagh Observatory, College Hill, Armagh, Northern Ireland,
-BT61 9DG, U.K.\\
-$^{24}$Jodrell Bank Centre for Astrophysics, School of Physics \&
-Astronomy, University of Manchester, Manchester M13 9PL, U.K.}
+$^{11}$Th\"uringer Landessternwarte, Sternwarte 5, D-07778, Tautenburg, Germany. \\
+$^{12}$Observatorio Astron\'omico, Universidad de Valencia,
+Catedr\'atico Jos\'e Beltr\'an 2, 46980 Paterna, Spain.\\
+$^{13}$Department of Physics, University of Warwick, Gibbet Hill Road, Coventry, CV4 7AL, U.K.\\
+$^{14}$Joint ALMA Observatory, Alonso de Córdova 3107, Vitacura 763-0355, Santiago, Chile. \\
+$^{15}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
+(see the source file for affiliations 16 to 28)
+%$^{16}$European Space Astronomy Centre (ESAC), Villafranca del Castillo, Villanueva de la Canada, E-28692 Madrid, Spain.\\
+%$^{17}$School of Physics, Bristol University, Tyndall Avenue, Bristol, BS8 1TL, U.K.\\
+%$^{18}$School of Physics, University of Exeter, Stocker Road, Exeter, EX4 4QL, U.K.\\
+%$^{19}$Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia.\\
+%$^{20}$Australian Astronomical Observatory, PO Box 915, North Ryde, NSW 1670, Australia.\\
+%$^{21}$Research Centre for Astronomy, Astrophysics and Astrophotonics, Macquarie University, Sydney, NSW 2109 Australia.\\
+%$^{22}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
+%$^{23}$Instituut voor Sterrenkunde, K.U. Leuven, Celestijnenlaan 200D, B-3001 Leuven, Belgium.\\
+%$^{24}$Max-Planck-Institut f\"ur extraterrestrische Physik (MPE), Giessenbachstrasse 1, D-85748, Garching bei M\"unchen, Germany. \\
+%$^{25}$Department of Physics, Blackett Laboratory, Imperial College London, Prince Consort Road, London, SW7 2AZ, U.K.\\
+%$^{26}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
+%$^{27}$Armagh Observatory, College Hill, Armagh, Northern Ireland, BT61 9DG, U.K.\\
+%$^{28}$Jodrell Bank Centre for Astrophysics, School of Physics \& Astronomy, University of Manchester, Manchester M13 9PL, U.K.
+}
 
@@
 covering Galactic latitudes $|b| < 5\degr$
-and longitudes $\ell=30$ to 215\degr
+and longitudes $\ell=30$ to 215\degr\ 
 in the $r$, $i$ and \ha\ filters 
@@
 for 219 million unique sources
-across 92\% of the footprint.
+across 92 per cent of the footprint.
 The observations were carried out between 2003 and 2013
@@
 of identifying emission-line stars and nebulae. 
-Through the provision of $r$, $i$ broadband photometry 
-alongside H$\alpha$ data,
+Through the provision of $r$ and $i$ broadband photometry 
+alongside narrowband \ha\ data,
 IPHAS has created the opportunity 
@@
 in a new way.
-For example, the survey’s unique $(r-\ha,\ r-i)$ colour-colour
+For example, the survey's unique $(r-\ha,\ r-i)$ colour-colour
 diagram has been shown to provide simultaneous constraints 
@@
 \citep{Sale2009, Sale2010, Giammanco2011, Sale2012, Barentsen2013}. 
-In effect, the availability of narrowband H$\alpha$ 
-alongside $r, i$ magnitudes
+In effect, the availability of narrowband \ha\
+alongside $r$ and $i$ magnitudes
 provides coarse spectral information for huge samples of stars 
@@
 When the initial data release was created in late 2007,
-just over $\sim$half of the survey footprint was covered
+just over half of the survey footprint was covered
 and the data were insufficiently complete 
@@
 
+\begin{figure*}
+        \includegraphics[width=1\linewidth]{./plots/footprint/footprint_small.png}
+        \caption{Survey area showing the footprints
+        of all the quality-approved IPHAS fields
+        which have been included in this data release.
+        The area covered by each field has been coloured black
+        with a semi-transparent opacity of 20 per cent,
+        such that regions where fields overlap are darker.
+        The IPHAS strategy is to observe each field twice
+        with a small offset,
+        and hence the vast majority of the area 
+        is covered twice (dominant grey colour).
+        There are small overlaps between all the neighbouring fields
+        which can be seen as a honeycomb pattern
+        of dark grey lines across the survey area.
+        Regions with incomplete data are apparent as white gaps (no data) 
+        or in light grey (denoting that only one offset is missing).
+        The dark vertical strip near $\ell \sim$125\degr\ 
+        is an arbitrary consequence of the tiling pattern,
+        which was populated starting from 0h in right ascension.}
+        \label{fig:footprint}
+\end{figure*}
+
 The detailed properties of the IPHAS observing programme 
@@
 by \citet{Drew2005} and \citet{Gonzalez-Solares2008}. 
-To set the stage for this release, we provide a reminder of some key points.
+To set the stage for this release, we recap some key points.
 IPHAS is an imaging survey of the Galactic Plane north of the celestial equator, 
-from which photometry in Sloan $r, i$ 
-is extracted along with narrowband H$\alpha$. 
+from which photometry in Sloan $r$ and $i$ 
+is extracted along with narrowband \ha. 
 It is carried out using the Wide Field Camera (WFC) 
@@
 Altogether, the survey contains 15\,270 fields grouped into 7\,635 field pairs.
-To achieve the desired survey depth of 20th magnitude or fainter, 
-the filter exposure times were set at 120 sec (narrowband \ha), 
-30 sec ($r$) and 10 sec ($i$)
-in the vast majority of the survey observations.\footnote{The $r$-band exposure time was 10~s instead of 30~s in the first year of data taking (2003). Since October 2010, the $i$-band exposure time 
+To achieve the desired survey depth
+of 20th magnitude or fainter in each filter, 
+the exposure times were set at 120~s (\ha), 
+30~s ($r$) and 10~s ($i$)
+in the vast majority of the survey observations.\footnote{The $r$-band exposure time was 10~s instead of 30~s in the first months of data taking. Since October 2010, the $i$-band exposure time 
 has been increased from 10~s to 20~s to by-pass a sporadic exposure timing bug that affects the WFC.}
@@
 and every field had been observed at least once by the end of 2008.
-At that time only 76\% of the field pairs 
+At that time only 76 per cent of the field pairs 
 satisfied our minimum quality criteria, however.
-The problems affecting the 24\% falling below survey standard were,
+The problems affecting the 24 per cent falling below survey standard were,
 most commonly: variable cloud cover;
@@
 to improve data quality. 
-As a result, 92\% of the survey footprint
+As a result, 92 per cent of the survey footprint
 now benefits from quality-approved data.
@@
 
-\begin{figure*}
-        \includegraphics[width=1\linewidth]{./plots/footprint/footprint_small.png}
-        \caption{Survey area showing the footprints
-        of all the quality-approved IPHAS fields
-        which have been included in this data release.
-        The area covered by each field has been coloured black
-        with a semi-transparent opacity of 20\%,
-        such that regions where fields overlap are darker.
-        The IPHAS strategy is to observe each field twice
-        with a small offset,
-        and hence the vast majority of the area 
-        is covered twice (dominant grey colour).
-        There are small overlaps between all the neighbouring fields
-        which can be seen as a honeycomb pattern
-        of dark grey lines across the survey area.
-        Regions with incomplete data are apparent as white gaps (no data) 
-        or in light grey (denoting that only one offset is missing).
-        The dark vertical strip near $\ell \sim$125\degr
-        is an arbitrary consequence of the tiling pattern,
-        which was populated starting from 0h in right ascension.}
-        \label{fig:footprint}
-\end{figure*}
-
 Fig.~\ref{fig:footprint} shows the footprint
@@
 The fields which remain missing 
--- covering 8\% of the survey area --
+-- covering 8 per cent of the survey area --
 are predominantly located towards the Galactic anti-centre 
@@
 Data taken under unsuitable conditions
-have been flagged and removed
-using seven quality criteria,
+have been rejected using seven quality criteria,
 which ensure a reliable and homogeneous level of quality
@@
 We discarded any exposures for which the $5\sigma$ limiting magnitude 
-was worse than 20th magnitude in the $r$-band
-or worse than 19th in $i$ or \ha. 
+was brighter than 20th magnitude in the $r$-band
+or brighter than 19th in $i$ or \ha. 
 Such data were typically obtained during poor weather or full moon.
@@
 most strongly on the presence of the moon,
-which was above the horizon during 62\% 
+which was above the horizon during 62 per cent 
 of the observations.
@@
 which is a proxy for the combined effect of atmospheric and dome seeing.
-This target is currently attained across 86\% of the footprint,
-in particular at lower Galactic longitudes,
-e.g. 92\% of the fields at $\ell<120\degr$ are better than 1.7~arcsec.
-To increase the sky area offered by the present release slightly,
+This target is currently attained across 86 per cent of the footprint.
+To increase the sky area offered by the data release slightly,
 we have decided to accept data obtained with seeing up to 2.5~arcsec.
@@
 of the PSF FWHM for the approved fields.
-In the $r$-band, 90\% is better than 1.5~arcsec,
-50\% is better than 1.1~arcsec,
-and 10\% is better than 0.8~arcsec
+In the $r$-band, 90 per cent is better than 1.5~arcsec,
+50 per cent is better than 1.1~arcsec,
+and 10 per cent is better than 0.8~arcsec
 (the WFC pixel scale is 0.33~arcsec/px).
@@
 for stars observed in both pointings.
-We rejected field pairs in which more than 2\% of the stars 
-showed an inconsistent measurement at the level of 0.2 mag,
-or more than 26\% were inconsistent at the level of 0.1 mag.
+We rejected field pairs in which more than 2 per cent of the stars 
+showed an inconsistent measurement at the level of 0.2~mag,
+or more than 26 per cent were inconsistent at the level of 0.1~mag.
 These limits were set empirically after inspecting
@@
 All of the above criteria were satisfied by at least one observing attempt
-for 14\,115 out of the 15\,270 fields (92\%).
+for 14\,115 out of the 15\,270 fields (92 per cent).
 In some cases more than one successful attempt to observe
@@
 lost outside of the aperture used
-(we adopt a 2.3~arcsec diameter circular aperture by default).
+(we adopt a 2\farcs 3-diameter circular aperture by default).
 Finally, \textsc{percorr} is a correction based on the difference
@@
 The entire procedure has been found to deliver zeropoints which 
-are accurate at the level of 1-2\% 
+are accurate at the level of 1--2 per cent 
 in photometric conditions \citep{Gonzalez-Solares2011}.
@@
 in the IPHAS \ha\ filter 
-is $1.52 \times 10^{-7}$ ergs\,cm$^{-2}$\,s$^{-1}$ 
+is $1.52 \times 10^{-7}$ erg\,cm$^{-2}$\,s$^{-1}$ 
 at the top of the Earth's atmosphere,
@@
 Note that the above implies that the in-band flux corresponding to
-zero magnitude is $1.56 \time 10^{-7}$ ergs\,cm$^{-2}$\,s$^{-1}$,
+zero magnitude is $1.56 \time 10^{-7}$ erg\,cm$^{-2}$\,s$^{-1}$,
 when the \ha\ magnitude for Vega is set by convention to 0.03.
@@
 -- where neighbouring fields overlap -- 
-is prone to systematics at the level of 1--2\%.
+is prone to systematics at the level of 1--2~per cent.
 The cause of these errors is thought to include 
@@
 The photometric accuracy of APASS is currently estimated 
-to be at the level of 3\% (Henden, private communication),
+to be at the level of 3 per cent (Henden, private communication),
 which is significantly better 
 than the original nightly calibrations of IPHAS
-which are accurate to $\sim$10\% (Tab.~\ref{tbl:offsets_before}).
+which are accurate to $\sim$10 per cent (Tab.~\ref{tbl:offsets_before}).
 APASS achieves its uniform accuracy 
@@
 
-With the aim of bringing IPHAS to a similar accuracy of 3\%,
+With the aim of bringing IPHAS to a similar accuracy of 3 per cent,
 we used the APASS catalogue to identify and adjust the calibration of all IPHAS fields 
-which showed a magnitude offset larger than 3\% against APASS.  Experience of re-running
+which showed a magnitude offset larger than 0.03~mag against APASS.  Experience of re-running
 the calibration and testing the results showed us that it was inadvisable to more
@@
 for each field which contained at least 30 cross-matched stars.
-This was achieved for 48\% of our fields (shown in Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}).
+This was achieved for 48 per cent of our fields (shown in Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}).
 The offsets follow a near-Gaussian distribution
@@
 that a field's \ha\ and $r$-band exposures
-were taken consecutively -- 
-separated only by the $\sim$30~s read-out time.
+were taken consecutively,
+albeit separated by the $\sim$30~s read-out time.
 Hence, we have corrected the \ha\ zeropoints 
@@
 
-Having re-calibrated all fields to the expected APASS accuracy of 3\%,
+Having re-calibrated all fields to the expected APASS accuracy of 3 per cent,
 we turned to a different survey, SDSS Data Release 9 \citep{Ahn2012},
@@
 Galactic latitudes,
-providing data across 18\% of the fields in our data release.
+providing data across 18 per cent of the fields in our data release.
 We cross-matched the IPHAS fields against the subset of
@@
 The colour term was found 
-to show some variation towards lowly reddened areas,
+to show some variation towards weakly reddened areas,
 where different stellar populations are observed.
@@
 \item the overlap regions of the field catalogues 
-were cross-matched to flag duplicate measurements 
+were cross-matched to flag duplicate (repeat) detections 
 and identify the primary (best) detection 
@@
 These archives curate the high-resolution near-infrared photometry from both
-the UKIDSS Galactic Plane Survey \citep[GPS;][]{Lucas2008}
+the UKIDSS Galactic Plane Survey \citep[GPS;][]{Lawrence2007,Lucas2008}
 and the 
@@
 aperture size is also used by the CASU pipeline to provide
-a discrete star/galaxy/noise classification flag
+a discrete star/galaxy/noise\footnote{For consistency with the terminology used in the CASU pipeline and the WSA archive, extended objects are classified as `galaxies' in the catalogue.
+This class may however include other types of resolved objects.} classification flag
 (\emph{rClass}, \emph{iClass}, \emph{haClass})
@@
 for which the peak pixel height exceeds 55000 counts,
-which is typically the case for stars brighter than 12-13th magnitude in
-$r$.  Although the pipeline attempts to extrapolate the brightness of
+which is typically the case for stars
+brighter than 12-13th magnitude in $r$.
+Although the pipeline attempts to extrapolate the brightness of
 saturated stars based on the shape of the PSF,
 such extrapolation is prone to error,
-and shoud be viewed as indicative rather than as precise measurement.
+and shoud be viewed as indicative rather than as precise measurement
+(systematic uncertainties as a function of magnitude
+will be discussed in \S\ref{sec:uncertainties}).
 %and we do not recommend reliance on it.
@@
 in \S\ref{sec:discussion} we will show
-that 19\% of the sources in our catalogue
+that 19 per cent of the sources in our catalogue
 are flagged as blended objects (column \emph{deblend}),
@@
 Amongst all sources in the magnitude range $13<r<19$,
-we find that 65\% were detected twice and 25\% were detected three times or more.
-Only 9\% were detected once.
-%Unsurprisingly, their spatial distribution traces
-%the inter-CCD gaps and footprint edges.
+we find that 65 per cent were detected twice
+and 25 per cent were detected three times or more.
+Only 9 per cent were detected once.
 
@@
 Secondly, co-adding would cause the image quality to degrade towards the mean,
-which is particularly a draw-back for crowded fields.
+which is particularly a drawback for crowded fields.
 
@@
 \emph{rErr2}, \emph{iErr2}, \emph{haErr2}, \emph{errBits2}).
-Such a secondary detection is available for 66\% of the sources brighter than $r<20$
-due to the IPHAS field pair strategy.
+Second, detections not included in the catalogue
+are made available on our website.
 
@@
 either \emph{pStar} or \emph{pGalaxy} to be
-greater than 20\%;
+greater than 20 per cent;
 \item the source must not have been detected in one of the strongly
@@
 Amongst these, 158 million objects 
-are detected in all three bands (72\%),
-30 million are detected in two bands (14\%),
-and 31 million entries are single-band detections (14\%).
+are detected in all three bands (72 per cent),
+30 million are detected in two bands (14 per cent),
+and 31 million entries are single-band detections (14 per cent).
 Half of the single-band detections were made in the $i$-band.
@@
 are faint sources observed near the detection limits;
-55\% of the entries in the catalogue
+55 per cent of the entries in the catalogue
 are fainter than $r > 20$.
@@
 (ii) remove saturated sources,
-and (iii) remove objects for which the 2.3-arcsec diameter
+and (iii) remove objects for which the 2\farcs 3-diameter
 aperture magnitude is inconsistent 
-with its alternative 1.2-arcsec diameter measurement.
+with its alternative 1\farcs 2-diameter measurement
+within each filter.
 This last criterion is a proxy
@@
 or failed source deblending.
-A total of 86 out of 219 million sources 
-(39\%) satisfy all the criteria listed in Table~\ref{tab:reliable}
+A total of 86 million sources out of 219 million 
+(39 per cent) satisfy all the criteria listed in Table~\ref{tab:reliable}
 and are hereafter referred to as `reliable'.
@@
 \subsection{Random and systematic uncertainties}
+\label{sec:uncertainties}
 
@@
 For example, we find a significant drop near the constellations 
-of Aquila ($\ell\simeq40$) and Cygnus ($\ell\simeq80$ and $\ell\simeq90$),
+of Aquila ($\ell\simeq40\degr$) and Cygnus ($\ell\simeq80\degr$ and $\ell\simeq90\degr$),
 which are regions known to be affected
@@
 spurious is greatly increased.
-%We have started investigating the use of
-%more advanced PSF-fitting routines
-%in which sources are measured simultaneously across all bands,
-%which may be guided by an external astrometric catalogue
-%such as the one that is to be provided by Gaia.
 
@@
 located at Galactic longitudes 
-180\degr, 45\degr and 30\degr, respectively,
-which were chosen because they show different characteristics.
+180\degr, 45\degr\ and 30\degr, respectively,
+which were chosen because they show stellar populations
+with different characteristics.
 Each figure contains all the sources
@@
 although the red clump stars are manifest as a track of slight 
-over-density sitting roughly 0.4~mags redder than the A0V reddening track.
+over-density sitting roughly 0.4~mag redder than the A0V reddening track.
 At $(l,b)=(45\degr, +2\degr)$ the giant stars observed
@@
 a mosaic of the same region 
-as seen in the mid-infrared by the Spitzer Space Telescope \citep[GLIMPSE survey;][]{Benjamin2003,Churchwell2009}.
+as seen in the mid-infrared by the Spitzer Space Telescope \citep{Benjamin2003,Churchwell2009}.
 The infrared image reveals an enclosing fuzzy bubble (appearing
@@
 In these studies, the authors used Spitzer data
-to select intrinsically red objects,
+to find intrinsically red objects,
 which is consistent with the presence of a circumstellar disk.
@@
 in the Spitzer 8.0~\micron\ image at S/N$>$5.
-They are likely to be Class II YSOs exhibiting a mild infrared excess.
+They are likely to be YSOs exhibiting a mild infrared excess.
 
@@
 of point sources across the Northern Galactic Plane at visible wavelengths,
-taking in the latitude range $|b|<5\degr$.
+taking in the Galactic latitude range $|b|<5\degr$
+at longitudes $\ell=30$ to 215\degr.
 The new 98-column catalogue provides single-epoch photometry
@@
 and selecting emission-line objects.  More comprehensive applications of IPHAS
-can be found in the works of Sale et al (2014), which applies DR2 to the problem
-of 3D extinction mapping, and of Sabin et al (2014), where the results of a search
+can be found in the works of Sale et al (submitted), which applies DR2 to the problem
+of 3D extinction mapping, and of Sabin et al (in preparation), where the results of a search
 of the image database for new planetary nebulae is presented.
@@
 in the Spanish Observatorio del Roque de los Muchachos
-of the Instituto de Astrofisica de Canarias.
+of the Instituto de Astrof\'\i sica de Canarias.
 We are deeply indebted to the ING staff and students
@@
 (grants: GB and JED ST/J001333/1, SES ST/K00106X/1, BTG ST/I001719/1). 
-HJF is in receipt of an STFC postgraduate studentship.
+HJF and MM-S both acknowledge STFC postgraduate studentships.
 JED would also like to convey her thanks
@@
 Programme (FP/2007-2013) / ERC Grant Agreement n. 320964 (WDTracer).
+PRG is supported by a Ram\'on y Cajal fellowship (RYC-2010-05762), and acknowledges support provided by the Spanish MINECO AYA2012--38700 grant.
 NJW is in receipt of a Fellowship funded by 

### 2014-04-14 paper/tables/columns.tex
Type in two rows

@@
 50 & iSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
-51 & iDetectionID & string &  & Unique identifier of the r-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e.
+51 & iDetectionID & string &  & Unique identifier of the i-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e.
 composed of the INT telescope run number, the CCD number and a sequential source detection number.
@@
 71 & haSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
-72 & haDetectionID & string &  & Unique identifier of the r-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e.
+72 & haDetectionID & string &  & Unique identifier of the H-alpha detection in the format "$\#$run-$\#$ccd-$\#$number", i.e.
 composed of the INT telescope run number, the CCD number and a sequential source detection number.

### 2014-04-15 paper/iphas-dr2.tex
First short at detection stats and Vega fluxes

@@
 J. E. Drew$^1$,
-E. A. Gonz$\acute{\rm{a}}$lez-Solares$^2$, 
-R. Greimel$^3$, \newauthor
+E. A. Gonz$\acute{\rm{a}}$lez-Solares$^2$, \newauthor
+R. Greimel$^3$, 
 M. J. Irwin$^2$,
@@
 P. Groot$^5$,
-A. Mampaso$^{6,7}$, 
-S. E. Sale$^8$, \newauthor
+A. Mampaso$^{6,7}$, \newauthor
+S. E. Sale$^8$, 
 M. J. Barlow$^9$,
@@
 R. L. M Corradi$^{6,7}$,
-J. J. Drake$^{10}$,
-J. Eisl\"offel$^{11}$, \newauthor
+J. J. Drake$^{10}$, \newauthor
+J. Eisl\"offel$^{11}$, 
 J. Fabregat$^{12}$, 
@@
 N. P. Gentile Fusillo$^{13}$,
-S. Greiss$^{13}$,
-A. S. Hales$^{14}$, \newauthor
+S. Greiss$^{13}$, \newauthor
+A. S. Hales$^{14}$, 
 S. Hodgkin$^2$,
+L. Huckvale$^{15}$,
 J. Irwin$^{9}$,
-C. Knigge$^{15}$,
-T. Kupfer$^{5}$,
-D. J. Lennon$^{16}$,
-J. R. Lewis$^{2}$,  \newauthor
+C. Knigge$^{16}$,
+T. Kupfer$^{5}$, \newauthor
+D. J. Lennon$^{17}$,
+J. R. Lewis$^{2}$,  
 M. Mohr-Smith$^{1}$, 
-R. A. H. Morris$^{17}$,
-T. Naylor$^{18}$,
-Q. A. Parker$^{19,20,21}$, 
-S. Phillipps$^{17}$, \newauthor
+R. A. H. Morris$^{18}$,
+T. Naylor$^{19}$, \newauthor
+Q. A. Parker$^{20,21,22}$, 
+S. Phillipps$^{17}$, 
 R. Raddi$^{13}$, 
 P. Rodr\'\i guez-Gil$^{6,7}$,
-L. Sabin$^{22}$,
-S. Scaringi$^{23,24}$,
-D. Steeghs$^{13}$, \newauthor
-Y. C. Unruh$^{25}$, 
+L. Sabin$^{23}$, \newauthor
+S. Scaringi$^{24,25}$,
+D. Steeghs$^{13}$, 
+Y. C. Unruh$^{26}$, 
 J. van Roestel$^5$,
-K. Viironen$^{26}$, 
-J. S. Vink$^{27}$, 
-N. A. Walton$^{2}$, \newauthor
+K. Viironen$^{27}$, \newauthor
+J. S. Vink$^{28}$, 
+N. A. Walton$^{2}$, 
 N. J. Wright$^{1}$,
-A. A. Zijlstra$^{28}$.
+A. A. Zijlstra$^{15}$.
 \\
@@
 $^{14}$Joint ALMA Observatory, Alonso de Córdova 3107, Vitacura 763-0355, Santiago, Chile. \\
-$^{15}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
-(see the source file for affiliations 16 to 28)
-%$^{16}$European Space Astronomy Centre (ESAC), Villafranca del Castillo, Villanueva de la Canada, E-28692 Madrid, Spain.\\
-%$^{17}$School of Physics, Bristol University, Tyndall Avenue, Bristol, BS8 1TL, U.K.\\
-%$^{18}$School of Physics, University of Exeter, Stocker Road, Exeter, EX4 4QL, U.K.\\
-%$^{19}$Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia.\\
-%$^{20}$Australian Astronomical Observatory, PO Box 915, North Ryde, NSW 1670, Australia.\\
-%$^{21}$Research Centre for Astronomy, Astrophysics and Astrophotonics, Macquarie University, Sydney, NSW 2109 Australia.\\
-%$^{22}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
-%$^{23}$Instituut voor Sterrenkunde, K.U. Leuven, Celestijnenlaan 200D, B-3001 Leuven, Belgium.\\
-%$^{24}$Max-Planck-Institut f\"ur extraterrestrische Physik (MPE), Giessenbachstrasse 1, D-85748, Garching bei M\"unchen, Germany. \\
-%$^{25}$Department of Physics, Blackett Laboratory, Imperial College London, Prince Consort Road, London, SW7 2AZ, U.K.\\
-%$^{26}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
-%$^{27}$Armagh Observatory, College Hill, Armagh, Northern Ireland, BT61 9DG, U.K.\\
-%$^{28}$Jodrell Bank Centre for Astrophysics, School of Physics \& Astronomy, University of Manchester, Manchester M13 9PL, U.K.
+$^{15}$Jodrell Bank Centre for Astrophysics, School of Physics \& Astronomy, University of Manchester, Manchester M13 9PL, U.K.\\
+%$^{16}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
+%$^{17}$European Space Astronomy Centre (ESAC), Villafranca del Castillo, Villanueva de la Canada, E-28692 Madrid, Spain.\\
+%$^{18}$School of Physics, Bristol University, Tyndall Avenue, Bristol, BS8 1TL, U.K.\\
+%$^{19}$School of Physics, University of Exeter, Stocker Road, Exeter, EX4 4QL, U.K.\\
+%$^{20}$Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia.\\
+%$^{21}$Australian Astronomical Observatory, PO Box 915, North Ryde, NSW 1670, Australia.\\
+%$^{22}$Research Centre for Astronomy, Astrophysics and Astrophotonics, Macquarie University, Sydney, NSW 2109 Australia.\\
+%$^{23}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
+%$^{24}$Instituut voor Sterrenkunde, K.U. Leuven, Celestijnenlaan 200D, B-3001 Leuven, Belgium.\\
+%$^{25}$Max-Planck-Institut f\"ur extraterrestrische Physik (MPE), Giessenbachstrasse 1, D-85748, Garching bei M\"unchen, Germany. \\
+%$^{26}$Department of Physics, Blackett Laboratory, Imperial College London, Prince Consort Road, London, SW7 2AZ, U.K.\\
+%$^{27}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
+%$^{28}$Armagh Observatory, College Hill, Armagh, Northern Ireland, BT61 9DG, U.K.\\
+$^{16-28}$ (affiliations 16 to 28 are given in the source file)
 }
@@
 covering Galactic latitudes $|b| < 5\degr$
-and longitudes $\ell=30$ to 215\degr\ 
+and longitudes $\ell=30\degr$ to 215\degr\ 
 in the $r$, $i$ and \ha\ filters 
@@
 These archives curate the high-resolution near-infrared photometry from both
-the UKIDSS Galactic Plane Survey \citep[GPS;][]{Lawrence2007,Lucas2008}
+the UKIRT Infrared Deep Sky Survey \citep[UKIDSS;][]{Lawrence2007}
 and the 
@@
 There is a significant degree of overlap
-between the footprints of UKIDSS/GPS and IPHAS,
+between the footprints of UKIDSS Galactic Plane Survey \citep[GPS;][]{Lucas2008}
+and IPHAS,
 and hence by adopting a similar catalogue format
@@
 Moreover, late-type stars tend to be intrinsically brighter in $i$.
+
+\begin{table}
+    \caption{Detection statistics.}
+    \label{tbl:detections}
+        \begin{tabular}{lrcrrr}
+        \toprule
+        Detection & N & \%($nObs>1$) &
+        $\overline{pStar}$ & $\overline{(r-i)}$ & $\overline{(r-\ha)}$ \\
+        \midrule
+        $r$, $i$, \ha & 159 (73\%) & 91\% & 0.82 & 1.03 & 0.44  \\
+        $r$, $i$ (not \ha) & 25 (11\%) & 77\% & 0.70 &  1.26 & - \\
+        $i$, \ha\ (not $r$) & 3 (1\%) & 73\% & 0.49 &  & \\
+        $r$, \ha\ (not $i$) & 2 (1\%) & 65\% & 0.53 & - & 0.58 \\
+        only $i$ & 15 (7\%) & 43\% & 0.44 \\
+        only $r$ & 9 (4\%) & 27\% & 0.42 \\
+        only \ha & 6 (3\%) & 12\% & 0.11 \\
+        \midrule
+        Total & 219 & 81\% \\
+        \bottomrule
+        \end{tabular}
+\end{table}
+
+% riha: 158831809 / 144339044
+% ri: 24558323 / 18878201
+% iha: 3422905 / 2495655
+% rha: 1606867 / 1036607
+% i: 15463660 / 6612351
+% r: 9263422 / 2478774
+% ha: 5844538 / 727732
 
@@
 
-The catalogue is made available through the Vizier
-catalogue tool (http://vizier.u-strasbg.fr)
-where it can be queried
-(Conesearch/TAP URL TBD).
-The catalogue can also be downloaded in its entirety
-from the IPHAS website (www.iphas.org) as a collection 
-of binary FITS tables occupying 50~GB,
-together with a script
-to ingest these tables into a PostgreSQL database.
+The catalogue is made available on our website (www.iphas.org)
+as a collection of binary FITS tables,
+each covering a $5\degr\times5\degr$ tile
+of the survey footprint.
+The catalogue can be queried in two ways.
+First, it is made available through the
+Vizier on-line service (http://vizier.u-strasbg.fr)
+where it can be queried using their web interface
+or using Virtual Observatory (VO) protocols.
+Second, on our website we provide a script
+to ingest the binary FITS tables into a
+{\sc PostgreSQL} database on a user's computer.
 
@@
 imaging data, which we have updated to include
-the re-calibrated zeropoints and the best-available
-astrometric solution in the image headers.
+a new keyword {\sc photzp},
+which contains the re-calibrated zeropoint.
+It can be used to convert the number counts $DN$
+-- which are the pixel values -- 
+into Vega-based magnitudes $m$ using:
+\begin{equation}
+\begin{split}
+   m  = \textsc{photzp} - 2.5 \log_{10}(DN)
+\label{eqn:mag}
+\end{split}
+\end{equation}
+That is, the {\sc photzp} has been computed such that it
+absorbs the required corrections for atmospheric extinction,
+gain variations, exposure time, and the re-calibration shift
+derived in this work.
+
+To convert the magnitude $m$ into fluxes,
+we list in Table~\ref{tbl:flux} something.
+
+% and the best-available
+%astrometric solution in the image headers.
 
@@
 
-Instructions for acknowledging IPHAS can be found on our website.
+Users of the data are reminded to cite both \citet{Drew2005}
+and this paper.
+
+Mean photon wavelength of the passband following eqn.~A14 in Bessel \& Murphy (2012):
+\begin{equation}
+\lambda_0 = \frac{\int{\lambda S(\lambda) d \lambda}}
+{\int{S(\lambda)d\lambda}}
+\end{equation}
+
+Pivot wavelength of the passband following eqn.~A16 in Bessel \& Murphy (2012):
+\begin{equation}
+\lambda_p = \frac{\int{S(\lambda) \lambda d \lambda}}
+{\int{\frac{S(\lambda)}{\lambda} d\lambda}}
+\end{equation}
+
+Mean monochromatic flux,
+following eqn.~A11 in Bessel \& Murphy (2012):
+\begin{equation}
+f_\lambda = \frac{\int{f_\lambda(\lambda) S(\lambda) \lambda d \lambda}}
+{\int{S(\lambda)\lambda d \lambda}}
+\end{equation}
+where $S(\lambda)$ includes the filter, atmosphere and QE.
+
+\begin{table}
+    \caption{Wavelength and mean monochromatic flux density for Vega in the IPHAS filter system.}
+    \label{tbl:flux}
+    \begin{tabular}{lrrrr}
+    \toprule
+     Band & $\lambda_0$ & $\lambda_p$ & EW  & $f_\lambda$   \\
+     & \AA & \AA & \AA & erg\,cm$^{-2}$\,s$^{-1}$\,\AA$^{-1}$  \\
+    \midrule
+$r$ &  6223  &   6211  &  785.6  &  2.47e-9  \\
+\ha &  6568  &   6568  &   59.6  &  1.81e-9  \\
+$i$ &  7674  &   7674  &  759.9  &  1.30e-9   \\
+    \bottomrule
+    \end{tabular}
+\end{table}
 
@@
 
-\section{Converting IPHAS magnitudes into fluxes}
-
-Mean photon wavelength of the passband following eqn.~A14 in Bessel \& Murphy (2012):
-\begin{equation}
-\lambda_0 = \frac{\int{\lambda S(\lambda) d \lambda}}
-{\int{S(\lambda)d\lambda}}
-\end{equation}
-
-Pivot wavelength of the passband following eqn.~A16 in Bessel \& Murphy (2012):
-\begin{equation}
-\lambda_p = \frac{\int{S(\lambda) \lambda d \lambda}}
-{\int{\frac{S(\lambda)}{\lambda} d\lambda}}
-\end{equation}
-
-Mean monochromatic flux,
-following eqn.~A11 in Bessel \& Murphy (2012):
-\begin{equation}
-f_\lambda = \frac{\int{f_\lambda(\lambda) S(\lambda) \lambda d \lambda}}
-{\int{S(\lambda)\lambda d \lambda}}
-\end{equation}
-where $S(\lambda)$ includes the filter, atmosphere and QE.
-
-\begin{table}[h]
-    \caption{Wavelength and mean monochromatic flux density for Vega in the IPHAS filter system.}
-    \label{tbl:flux}
-    \begin{tabular}{lrrrr}
-    \toprule
-     Band & $\lambda_0$ & $\lambda_p$ & EW  & $f_\lambda$   \\
-     & \AA & \AA & \AA & erg\,cm$^{-2}$\,s$^{-1}$\,\AA$^{-1}$  \\
-    \midrule
-$r$ &  6223  &   6211  &  785.6  &  2.47e-9  \\
-\ha &  6568  &   6568  &   59.6  &  1.81e-9  \\
-$i$ &  7674  &   7674  &  759.9  &  1.30e-9   \\
-    \bottomrule
-    \end{tabular}
-\end{table}
-
 \newpage

### 2014-04-15 paper/iphas-dr2.tex
Messing with Table 5

@@
 A. S. Hales$^{14}$, 
+A. A. Henden$^{15}$,
 S. Hodgkin$^2$,
-L. Huckvale$^{15}$,
+L. Huckvale$^{16}$,
 J. Irwin$^{9}$,
-C. Knigge$^{16}$,
-T. Kupfer$^{5}$, \newauthor
-D. J. Lennon$^{17}$,
+C. Knigge$^{17}$, \newauthor
+T. Kupfer$^{5}$, 
+D. J. Lennon$^{18}$,
 J. R. Lewis$^{2}$,  
 M. Mohr-Smith$^{1}$, 
-R. A. H. Morris$^{18}$,
-T. Naylor$^{19}$, \newauthor
-Q. A. Parker$^{20,21,22}$, 
-S. Phillipps$^{17}$, 
+R. A. H. Morris$^{19}$, \newauthor
+T. Naylor$^{20}$, 
+Q. A. Parker$^{21,22,23}$,
+S. Phillipps$^{19}$, 
 R. Raddi$^{13}$, 
-P. Rodr\'\i guez-Gil$^{6,7}$,
-L. Sabin$^{23}$, \newauthor
-S. Scaringi$^{24,25}$,
+P. Rodr\'\i guez-Gil$^{6,7}$, \newauthor
+L. Sabin$^{24}$, 
+S. Scaringi$^{25,26}$,
 D. Steeghs$^{13}$, 
-Y. C. Unruh$^{26}$, 
-J. van Roestel$^5$,
-K. Viironen$^{27}$, \newauthor
-J. S. Vink$^{28}$, 
+Y. C. Unruh$^{27}$, 
+J. van Roestel$^5$, \newauthor
+K. Viironen$^{28}$, 
+J. S. Vink$^{29}$, 
 N. A. Walton$^{2}$, 
@@
 $^{14}$Joint ALMA Observatory, Alonso de Córdova 3107, Vitacura 763-0355, Santiago, Chile. \\
-$^{15}$Jodrell Bank Centre for Astrophysics, School of Physics \& Astronomy, University of Manchester, Manchester M13 9PL, U.K.\\
-%$^{16}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
-%$^{17}$European Space Astronomy Centre (ESAC), Villafranca del Castillo, Villanueva de la Canada, E-28692 Madrid, Spain.\\
-%$^{18}$School of Physics, Bristol University, Tyndall Avenue, Bristol, BS8 1TL, U.K.\\
-%$^{19}$School of Physics, University of Exeter, Stocker Road, Exeter, EX4 4QL, U.K.\\
-%$^{20}$Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia.\\
-%$^{21}$Australian Astronomical Observatory, PO Box 915, North Ryde, NSW 1670, Australia.\\
-%$^{22}$Research Centre for Astronomy, Astrophysics and Astrophotonics, Macquarie University, Sydney, NSW 2109 Australia.\\
-%$^{23}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
-%$^{24}$Instituut voor Sterrenkunde, K.U. Leuven, Celestijnenlaan 200D, B-3001 Leuven, Belgium.\\
-%$^{25}$Max-Planck-Institut f\"ur extraterrestrische Physik (MPE), Giessenbachstrasse 1, D-85748, Garching bei M\"unchen, Germany. \\
-%$^{26}$Department of Physics, Blackett Laboratory, Imperial College London, Prince Consort Road, London, SW7 2AZ, U.K.\\
-%$^{27}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
-%$^{28}$Armagh Observatory, College Hill, Armagh, Northern Ireland, BT61 9DG, U.K.\\
-$^{16-28}$ (affiliations 16 to 28 are given in the source file)
+$^{15}$AAVSO, 49 Bay State Road, Cambridge, MA 02138, USA.\\
+%$^{16}$Jodrell Bank Centre for Astrophysics, School of Physics \& Astronomy, University of Manchester, Manchester M13 9PL, U.K.\\
+%$^{17}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
+%$^{18}$European Space Astronomy Centre (ESAC), Villafranca del Castillo, Villanueva de la Canada, E-28692 Madrid, Spain.\\
+%$^{19}$School of Physics, Bristol University, Tyndall Avenue, Bristol, BS8 1TL, U.K.\\
+%$^{20}$School of Physics, University of Exeter, Stocker Road, Exeter, EX4 4QL, U.K.\\
+%$^{21}$Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia.\\
+%$^{22}$Australian Astronomical Observatory, PO Box 915, North Ryde, NSW 1670, Australia.\\
+%$^{23}$Research Centre for Astronomy, Astrophysics and Astrophotonics, Macquarie University, Sydney, NSW 2109 Australia.\\
+%$^{24}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
+%$^{25}$Instituut voor Sterrenkunde, K.U. Leuven, Celestijnenlaan 200D, B-3001 Leuven, Belgium.\\
+%$^{26}$Max-Planck-Institut f\"ur extraterrestrische Physik (MPE), Giessenbachstrasse 1, D-85748, Garching bei M\"unchen, Germany. \\
+%$^{27}$Department of Physics, Blackett Laboratory, Imperial College London, Prince Consort Road, London, SW7 2AZ, U.K.\\
+%$^{28}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
+%$^{29}$Armagh Observatory, College Hill, Armagh, Northern Ireland, BT61 9DG, U.K.\\
+$^{16-29}$ (affiliations 16 through 29 are given in the source file)
 }
@@
 the above criteria and have been included in the catalogue.
-Amongst these, 158 million objects 
-are detected in all three bands (72 per cent),
+Amongst these, 159 million are detected 
+in all three bands (73 per cent),
 30 million are detected in two bands (14 per cent),
-and 31 million entries are single-band detections (14 per cent).
+and the remaining 30 million are single-band detections.
+The detailed breakdown is presented in Table~\ref{tbl:detections},
+along with the fraction of `confirmed' sources,
+hich we define
+as objects which were detected in more than
+one field pointing (usually the alternative pointing of a field pair).
+Sources detected in all three 
+(Table~\ref{tbl:detections}).
 Half of the single-band detections were made in the $i$-band.
-This is because $i$ is least
+This is most likely because $i$ is least
 affected by interstellar extinction and can occasionally pick up
 highly-reddened objects which are otherwise lost in $r$ and \ha.
-Moreover, late-type stars tend to be intrinsically brighter in $i$.
+Table~\ref{tbl:detections} also shows the fraction of the
+sources that were `confirmed', w
+For the class of 
+in more than one field (usually the field observed at the small offset position)
 
 \begin{table}
-    \caption{Detection statistics.}
+    \caption{Breakdown of catalogue entries as a function of the band(s) in which the source was detected.
+We also show the fraction of `confirmed' objects, which we define as sources which were detected in more than one field pointing (usually the field pair partner).
+Also included are the mean magnitudes for each class.}
     \label{tbl:detections}
-        \begin{tabular}{lrcrrr}
+        \begin{tabular}{lrcccc}
         \toprule
-        Detection & N & \%($nObs>1$) &
-        $\overline{pStar}$ & $\overline{(r-i)}$ & $\overline{(r-\ha)}$ \\
+        Band(s) & Sources & Confirmed & $\overline{r}$ & $\overline{i}$ & $\overline{\ha}$ \\
+        & [$10^6$] & [\emph{nObs}$>$1] & [mag] & [mag] & [mag] \\
         \midrule
-        $r$, $i$, \ha & 159 (73\%) & 91\% & 0.82 & 1.03 & 0.44  \\
-        $r$, $i$ (not \ha) & 25 (11\%) & 77\% & 0.70 &  1.26 & - \\
-        $i$, \ha\ (not $r$) & 3 (1\%) & 73\% & 0.49 &  & \\
-        $r$, \ha\ (not $i$) & 2 (1\%) & 65\% & 0.53 & - & 0.58 \\
-        only $i$ & 15 (7\%) & 43\% & 0.44 \\
-        only $r$ & 9 (4\%) & 27\% & 0.42 \\
-        only \ha & 6 (3\%) & 12\% & 0.11 \\
+        $r$, $i$, \ha & 159 & 91\% & 19.3 & 18.2 & 18.8  \\
+        $r$, $i$ & 25 & 77\% & 21.2 & 19.9 & - \\
+        $i$, \ha & 3 & 73\% & - & 19.6 & 20.2 \\
+        $r$, \ha & 2 & 65\% & 20.6 & - & 20.1 \\
+        $i$ & 15 & 43\% & - & 19.8 & - \\
+        $r$ & 9 & 27\% & 21.1 & - & - \\
+        \ha & 6 & 12\% & - & - & 19.9 \\
         \midrule
-        Total & 219 & 81\% \\
+        Total  & 219 & 81\% & 19.6 & 18.6 & 18.9 \\
         \bottomrule
@@
 each covering a $5\degr\times5\degr$ tile
-of the survey footprint.
+of the footprint.
 The catalogue can be queried in two ways.
-First, it is made available through the
-Vizier on-line service (http://vizier.u-strasbg.fr)
-where it can be queried using their web interface
-or using Virtual Observatory (VO) protocols.
+First, will be made available through the
+Vizier on-line service (http://vizier.u-strasbg.fr),
+where it can be accessed using a web interface
+and using Virtual Observatory (VO) protocols.
 Second, on our website we provide a script
 to ingest the binary FITS tables into a
-{\sc PostgreSQL} database on a user's computer.
+{\sc PostgreSQL} database for users
+who wish to mine the data locally.
 
@@
 imaging data, which we have updated to include
-a new keyword {\sc photzp},
+a new keyword, called {\sc photzp},
 which contains the re-calibrated zeropoint.
-It can be used to convert the number counts $DN$
--- which are the pixel values -- 
+The keyword can be used to convert the number counts $DN$,
+which stored in the pixel values,
 into Vega-based magnitudes $m$ using:
@@
 \begin{split}
-   m  = \textsc{photzp} - 2.5 \log_{10}(DN)
+   m  = \textsc{photzp} - 2.5 \log_{10}(DN).
 \label{eqn:mag}
@@
 \end{equation}
-That is, the {\sc photzp} has been computed such that it
+That is, {\sc photzp} has been computed such that it
 absorbs the required corrections for atmospheric extinction,
-gain variations, exposure time, and the re-calibration shift
-derived in this work.
+gain variations, exposure time, and the re-calibration.
+
+A user will only have to apply a correction
+for the flux lost outside
 

### 2014-04-16 paper/iphas-dr2.tex
First attempt to incluto include Vega fluxes, data access section, and detection statistics paragraph properly.

@@
 standard-star photometry is not available in the literature 
-for the H$\alpha$ passband
+for the \ha\ passband
 and hence there is no formally recognised flux scale 
@@
 We can specify here, however, 
-that the integrated in-band energy flux for Vega 
-in the IPHAS \ha\ filter 
-is $1.52 \times 10^{-7}$ erg\,cm$^{-2}$\,s$^{-1}$ 
-at the top of the Earth's atmosphere,
-which is the flux obtained by folding 
-Vega's SED with the filter transmission curve 
-(uncorrected for the atmosphere and detector quantum efficiency,
-which would otherwise scale down the narrowband flux by 0.707).
-This is 3.14 magnitudes less than the flux captured 
-by the much broader $r$ band
-which includes the \ha\ band.
+that the detected flux for Vega in the
+IPHAS \ha\ filter is 3.14 magnitudes
+less than the flux captured by the much broader
+$r$ band which includes the \ha\ band.
 Hence to assure zero colour relative to the broadbands,
@@
 \end{equation}
-Note that the above implies that the in-band flux corresponding to
-zero magnitude is $1.56 \time 10^{-7}$ erg\,cm$^{-2}$\,s$^{-1}$,
-when the \ha\ magnitude for Vega is set by convention to 0.03.
+For reference, Table~\ref{tbl:flux} details the flux of Vega
+in the IPHAS filter system.
+We refer to \citet{Drew2005} for a more comprehensive
+discussion on the throughput curves of the bands.
+
+\begin{table}
+    \caption{Mean monochromatic flux of Vega in the IPHAS filter system,
+             defined as 
+             $\langle f_\lambda\rangle~=~\int{f_\lambda(\lambda) S(\lambda) \lambda d \lambda} / \int{S(\lambda)\lambda d \lambda}$,
+             where $S(\lambda)$ is the photon response function 
+             (which includes atmospheric transmission,
+             filter transmission
+             and CCD response)
+             and $f_\lambda(\lambda)$ is the model SED for Vega
+             by Kurucz (http://kurucz.harvard.edu/stars.html).
+             For reference, we also provide the 
+             filter equivalent width EW~=~$\int S(\lambda) d\lambda$,
+             the mean photon wavelength 
+             $\lambda_0~=~\int{S(\lambda)\lambda d\lambda} / \int{S(\lambda)d\lambda}$,
+             and the pivot wavelength
+             $\lambda_p~=~\sqrt{\int{S(\lambda)\lambda d\lambda} / \int{\frac{S(\lambda)}{\lambda} d\lambda}}$.
+             These notations follow the definitions by \citet{Bessell2012}.
+             After multiplying $\langle f_\lambda\rangle$ by the EW,
+             we find that the detected flux for Vega in \ha\ is 3.14~magnitudes less
+             than that received in $r$, 
+          }
+    \label{tbl:flux}
+    \begin{center}
+    \begin{tabular}{lccrc}
+    \toprule
+     Filter  & $\langle f_\lambda \rangle$  & \multicolumn{1}{c}{EW} &$\lambda_0$ & $\lambda_p$    \\
+             & [erg\,cm$^{-2}$\,s$^{-1}$\,\AA$^{-1}$] & [\AA] & [\AA] & [\AA] \\
+    \midrule
+$r$ &  $2.47 \times 10^{-9}$  &  785.6 &  6223  &   6211   \\
+\ha &  $1.81 \times 10^{-9}$ &   59.6 &  6568  &   6568    \\
+$i$  &  $1.30 \times 10^{-9}$  &  759.9  &  7674  &   7674   \\
+    \bottomrule
+    \end{tabular}
+    \end{center}
+\end{table}
 
@@
 
-A total of 219 million primary detections satisfied
-the above criteria and have been included in the catalogue.
-Amongst these, 159 million are detected 
-in all three bands (73 per cent),
-30 million are detected in two bands (14 per cent),
-and the remaining 30 million are single-band detections.
-The detailed breakdown is presented in Table~\ref{tbl:detections},
-along with the fraction of `confirmed' sources,
-hich we define
-as objects which were detected in more than
-one field pointing (usually the alternative pointing of a field pair).
-Sources detected in all three 
-(Table~\ref{tbl:detections}).
-Half of the single-band detections were made in the $i$-band.
-This is most likely because $i$ is least
-affected by interstellar extinction and can occasionally pick up
-highly-reddened objects which are otherwise lost in $r$ and \ha.
-Table~\ref{tbl:detections} also shows the fraction of the
-sources that were `confirmed', w
-For the class of 
-in more than one field (usually the field observed at the small offset position)
-
 \begin{table}
-    \caption{Breakdown of catalogue entries as a function of the band(s) in which the source was detected.
-We also show the fraction of `confirmed' objects, which we define as sources which were detected in more than one field pointing (usually the field pair partner).
-Also included are the mean magnitudes for each class.}
+    \caption{Breakdown of catalogue entries as a function of the band(s)
+             in which the source was detected.
+             We also show the fraction of `confirmed' objects,
+             which we define as those sources detected in more than one field
+             (usually the field pair partner).
+             We also show the average magnitudes in each class.}
     \label{tbl:detections}
+    \begin{center}
         \begin{tabular}{lrcccc}
@@
         \end{tabular}
-\end{table}
-
         % riha: 158831809 / 144339044
@@
 		% ha: 5844538 / 727732
+	\end{center}
+\end{table}
+
+A total of 219 million primary detections satisfied
+the above criteria and have been included in the catalogue.
+
+Table~\ref{tbl:detections} details the number of bands
+in which the catalogue entries are captured;
+159 million sources are detected in all three filters (73 per cent),
+30 million are detected in two filters (14 per cent),
+and the remaining 30 million are single-band detections.
+Table~\ref{tbl:detections} also presents the average magnitudes
+and the fraction of `confirmed' objects,
+which we define as those sources which have been detected in more than
+one field
+(recall that the majority of the survey area
+is observed twice due to the field pair strategy).
+We find that the single-band detections tend to show low confirmation rates,
+which suggests that a significant fraction of these entries may be spurious detections.
+Hence, we strongly recommend users not to rely on single-band objects
+without inspecting the image data by eye.
+
+Not all the single-band detections are spurious sources, however.
+We note that the confirmation rate for $i$-band detections
+is markedly better than for $r$ and \ha,
+which is likely explained by the fact that $i$ is least
+affected by interstellar extinction, and so the survey can occasionally pick up
+highly-reddened objects in $i$ which are otherwise lost in $r$ and \ha.
+Moreover, objects which are intrinsically very red may also be picked
+up in $i$ alone, while faint objects with strong Balmer emission
+may appear only in \ha.
 
@@
 
-%Having explained how the catalogue was created,
 We now offer an overview of the properties of the
@@
 
-\section{Data access and source code}
+\section{Catalogue and image access}
 \label{sec:dataaccess}
 
-The catalogue is made available on our website (www.iphas.org)
+The catalogue is made available through the
+Vizier service (http://vizier.u-strasbg.fr),
+where it can be queried using a web interface
+and using Virtual Observatory (VO) protocols.
+In addition, the catalogue can be downloaded
+in its entirety from our website (www.iphas.org)
 as a collection of binary FITS tables,
@@
 of the footprint.
-The catalogue can be queried in two ways.
-First, will be made available through the
-Vizier on-line service (http://vizier.u-strasbg.fr),
-where it can be accessed using a web interface
-and using Virtual Observatory (VO) protocols.
-Second, on our website we provide a script
-to ingest the binary FITS tables into a
-{\sc PostgreSQL} database for users
-who wish to mine the data locally.
-
-The website also provides access to the pipeline-processed
-imaging data, which we have updated to include
-a new keyword, called {\sc photzp},
+
+We do not recommend using the catalogue
+to study extended objects which
+are larger than the
+aperture diameters used in this work.
+To enable the analysis of diffuse sources,
+our website provides access to the pipeline-processed
+imaging data. These have been updated to include
+a new FITS header keyword, called {\sc photzp},
 which contains the re-calibrated zeropoint.
-The keyword can be used to convert the number counts $DN$,
-which stored in the pixel values,
+This keyword can be used to convert the number counts $DN$,
+i.e. the pixel values in the images,
 into Vega-based magnitudes $m$ using:
@@
 \end{equation}
-That is, {\sc photzp} has been computed such that it
+The {\sc photzp} value has been computed such that it
 absorbs the required corrections for atmospheric extinction,
-gain variations, exposure time, and the re-calibration.
-
-A user will only have to apply a correction
-for the flux lost outside
-
-To convert the magnitude $m$ into fluxes,
-we list in Table~\ref{tbl:flux} something.
-
-% and the best-available
-%astrometric solution in the image headers.
+gain variations, exposure time, and the re-calibration shift.
+Users will only need to apply an aperture correction, if necessary.
+
+To estimate absolute narrow-band \ha\ fluxes from the image data,
+we note that the integrated in-band energy flux for Vega 
+in the IPHAS \ha\ filter 
+is $1.52 \times 10^{-7}$~erg\,cm$^{-2}$\,s$^{-1}$ 
+at the top of the Earth's atmosphere,
+which is the flux obtained by folding the 
+Vega model SED from Kurucz with the filter transmission curve 
+(uncorrected for the atmosphere and detector quantum efficiency,
+which would otherwise scale down the narrowband flux by 0.707).
+This implies that the in-band flux corresponding to
+zero magnitude is $1.56 \time 10^{-7}$ erg\,cm$^{-2}$\,s$^{-1}$,
+when the \ha\ magnitude for Vega is set by convention to 0.03 \citep{Fukugita1996}.
+
+The image repository on our website
+is accompanied by an archive of detection tables
+-- one for each image --
+which may be used to retrieve secondary detections that are not included
+in the DR2 source catalogue.
+We warn that this set of original images and detection tables
+includes data that did not pass quality control,
+and has not been globally re-calibrated.
+Such data are flagged and must be used with great caution.
 
@@
 https://github.com/barentsen/iphas-dr2
-
-Users of the data are reminded to cite both \citet{Drew2005}
-and this paper.
-
-Mean photon wavelength of the passband following eqn.~A14 in Bessel \& Murphy (2012):
-\begin{equation}
-\lambda_0 = \frac{\int{\lambda S(\lambda) d \lambda}}
-{\int{S(\lambda)d\lambda}}
-\end{equation}
-
-Pivot wavelength of the passband following eqn.~A16 in Bessel \& Murphy (2012):
-\begin{equation}
-\lambda_p = \frac{\int{S(\lambda) \lambda d \lambda}}
-{\int{\frac{S(\lambda)}{\lambda} d\lambda}}
-\end{equation}
-
-Mean monochromatic flux,
-following eqn.~A11 in Bessel \& Murphy (2012):
-\begin{equation}
-f_\lambda = \frac{\int{f_\lambda(\lambda) S(\lambda) \lambda d \lambda}}
-{\int{S(\lambda)\lambda d \lambda}}
-\end{equation}
-where $S(\lambda)$ includes the filter, atmosphere and QE.
-
-\begin{table}
-    \caption{Wavelength and mean monochromatic flux density for Vega in the IPHAS filter system.}
-    \label{tbl:flux}
-    \begin{tabular}{lrrrr}
-    \toprule
-     Band & $\lambda_0$ & $\lambda_p$ & EW  & $f_\lambda$   \\
-     & \AA & \AA & \AA & erg\,cm$^{-2}$\,s$^{-1}$\,\AA$^{-1}$  \\
-    \midrule
-$r$ &  6223  &   6211  &  785.6  &  2.47e-9  \\
-\ha &  6568  &   6568  &   59.6  &  1.81e-9  \\
-$i$ &  7674  &   7674  &  759.9  &  1.30e-9   \\
-    \bottomrule
-    \end{tabular}
-\end{table}
 

### 2014-04-16 paper/iphas-dr2.tex
Adds 400 sq deg colour colour diagram to show calibration shifts

@@
 are accurate at the level of 1--2 per cent 
-in photometric conditions \citep{Gonzalez-Solares2011}.
+in stable photometric conditions \citep{Gonzalez-Solares2011}.
 
@@
 in the timing of exposures).
-Such variations need to be corrected for 
+This is demonstrated in Fig.~\ref{fig:ccd_before},
+where we show the combined colour-colour diagram
+for the set of 2\,801 quality-approved fields
+which are located towards the Galactic anti-centre
+($160\degr < \ell < 200\degr$).
+Despite the fact that all the fields included in the diagram
+have been quality-approved,
+the main locus of stars is poorly defined
+due to the presence of calibration errors,
+which regularly exceed 0.1~mag.
+For comparison, Fig.~\ref{fig:ccd_after}
+shows the identical diagram
+obtained after re-calibrating.
+
+\begin{figure}
+\captionsetup[subfigure]{labelformat=empty}
+\begin{subfigure}[b]{\linewidth}
+\centering
+\includegraphics[width=\textwidth]{figures/caldiagram/ccd-uncalibrated.pdf}
+\caption{}
+\label{fig:ccd_before}
+\end{subfigure}
+\begin{subfigure}[b]{\linewidth}
+\centering
+\includegraphics[width=\textwidth]{figures/caldiagram/ccd-calibrated.pdf}
+\caption{}
+\label{fig:ccd_after}
+\end{subfigure}
+\caption{IPHAS $(r-\ha,\ r-i)$ colour-colour diagram for an area of 400~deg$^2$, plotted before (panel a) and after (panel b)
+re-calibration.
+Both figures were created by combining the stars detected
+across the entire set of 2\,801 quality-approved fields which are located
+towards the Galactic anti-centre  ($160\degr < \ell < 200\degr$).
+The diagrams are plotted as 2D-histograms
+which show the density of sources
+in bins of 0.01-by-0.01~mag;
+bins containing 1 to 200 sources are coloured red,
+while bins containing more than 400 sources are bright yellow.
+We have included all stars which are brighter than $r<18$
+and are classified as `very reliable'
+(these criteria will be explained in \S\ref{sec:qualitycriteria}).}
+\end{figure}
+
+The calibration variations need to be corrected for 
 during a global re-calibration procedure.
@@
 exceeding $\pm$0.03~mag in either $r$ or $i$.
-
-We then applied the most important step in our calibration scheme,
-which is to adjust the provisional zeropoints of these 4596 aberrant fields
-such that their offset is brought to zero.
-This allowed the mean IPHAS-to-APASS offset 
-to be brought down to $0.000\pm0.011$~mag in both $r$ and $i$
-(Table~\ref{tbl:offsets_after}).
-The procedure of fitting magnitude transformations and
-correcting the IPHAS zeropoints was repeated a few times to ensure 
-convergence, which was essentially reached after the first iteration.
-
-\begin{figure*}
-    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_rshift.pdf} 
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_r.pdf} 
-    \caption{Median magnitude offset in the $r$ band between IPHAS and APASS,
-             plotted on a field-by-field basis
-             prior to the re-calibration procedure.
-             Each square represents the footprint of an IPHAS field
-             which contains at least 30 stars with a counterpart
-             in the APASS DR7 catalogue.
-             The colours denote the median
-             IPHAS-APASS magnitude offset in each field,
-             which was computed after applying the APASS-to-IPHAS
-             transformation to the APASS magnitudes (Eqn.~\ref{eqn:apass_r}).
-             For clarity, we do not show the fields at the offset positions.}
-        \label{fig:apass_r}
-    \vspace{1cm}
-    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_ishift.pdf} 
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_i.pdf} 
-    \caption{Same as Fig.~\ref{fig:apass_r} for the $i$-band.}
-    \label{fig:apass_i}
-\end{figure*}
-
-\begin{figure*}
-    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_rshift.pdf}
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_r.pdf} 
-    \caption{Median magnitude offset between IPHAS and SDSS in the $r$ band
-             \emph{after} the re-calibration procedure using APASS was applied.
-             Each square represents the footprint of an IPHAS field
-             which contains at least 30 stars
-             with a counterpart in the SDSS DR9 catalogue.
-             The colours denote the median IPHAS-SDSS magnitude offset
-             in each field,
-             which was computed after applying the SDSS-to-IPHAS
-             transformation to the SDSS magnitudes (Eqn.~\ref{eqn:sdss_r}).}
-    \label{fig:sdss_r}
-    \vspace{1cm}
-    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_ishift.pdf}
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_i.pdf} 
-    \caption{Same as Fig.~\ref{fig:sdss_r} for the $i$-band.}
-    \label{fig:sdss_i}
-\end{figure*}
 
@@
 \end{table}
+
+We then applied the most important step in our calibration scheme,
+which is to adjust the provisional zeropoints of these 4596 aberrant fields
+such that their offset is brought to zero.
+This allowed the mean IPHAS-to-APASS offset 
+to be brought down to $0.000\pm0.011$~mag in both $r$ and $i$
+(Table~\ref{tbl:offsets_after}).
+The procedure of fitting magnitude transformations and
+correcting the IPHAS zeropoints was repeated a few times to ensure 
+convergence, which was essentially reached after the first iteration.
+
+\begin{figure*}
+    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_rshift.pdf} 
+    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_r.pdf} 
+    \caption{Median magnitude offset in the $r$ band between IPHAS and APASS,
+             plotted on a field-by-field basis
+             prior to the re-calibration procedure.
+             Each square represents the footprint of an IPHAS field
+             which contains at least 30 stars with a counterpart
+             in the APASS DR7 catalogue.
+             The colours denote the median
+             IPHAS-APASS magnitude offset in each field,
+             which was computed after applying the APASS-to-IPHAS
+             transformation to the APASS magnitudes (Eqn.~\ref{eqn:apass_r}).
+             For clarity, we do not show the fields at the offset positions.}
+        \label{fig:apass_r}
+    \vspace{1cm}
+    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_ishift.pdf} 
+    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_i.pdf} 
+    \caption{Same as Fig.~\ref{fig:apass_r} for the $i$-band.}
+    \label{fig:apass_i}
+\end{figure*}
+
+\begin{figure*}
+    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_rshift.pdf}
+    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_r.pdf} 
+    \caption{Median magnitude offset between IPHAS and SDSS in the $r$ band
+             \emph{after} the re-calibration procedure using APASS was applied.
+             Each square represents the footprint of an IPHAS field
+             which contains at least 30 stars
+             with a counterpart in the SDSS DR9 catalogue.
+             The colours denote the median IPHAS-SDSS magnitude offset
+             in each field,
+             which was computed after applying the SDSS-to-IPHAS
+             transformation to the SDSS magnitudes (Eqn.~\ref{eqn:sdss_r}).}
+    \label{fig:sdss_r}
+    \vspace{1cm}
+    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_ishift.pdf}
+    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_i.pdf} 
+    \caption{Same as Fig.~\ref{fig:sdss_r} for the $i$-band.}
+    \label{fig:sdss_i}
+\end{figure*}
 
@@
     bins containing 1 to 10 objects are coloured red,
-    bins with 10 to 20 objects are orange,
-    and bins with 20 to 25 objects are yellow.
+    while bins with more than 20 objects are yellow.
     The left panel is annotated with

### 2014-04-16 paper/figures/caldiagram/plot.py
Adds 400 sq deg colour colour diagram to show calibration shifts

@@ -1,0 +1,82 @@
+"""Plots the calibrated and uncalibrated CCD over a large area."""
+import os
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from astropy.io import fits
+from astropy import log
+from dr2 import constants
+
+fields = constants.IPHASQC[constants.IPHASQC_COND_RELEASE]
+mask = (fields['l'] > 160.0) & (fields['l'] < 200.)
+ids = fields['id'][mask]
+log.info('Plotting {0} fields.'.format(len(ids)))
+
+calibrated = True
+
+if calibrated:
+    CATALOGUE_PATH = '/home/gb/tmp/iphas-dr2-rc6/bandmerged-calibrated/'
+else:
+    CATALOGUE_PATH = '/home/gb/tmp/iphas-dr2-rc6/bandmerged/'
+
+COLORMAP = matplotlib.colors.LinearSegmentedColormap.from_list('mymap', 
+    ['#bd0026', '#f03b20', '#fd8d3c', '#fecc5c', '#ffffb2'])
+
+
+
+rmi, rmha = [], []
+
+for field in ids:
+    filename = os.path.join(CATALOGUE_PATH, field+'.fits')
+    log.info(filename)
+    d = fits.getdata(filename, 1)
+    veryreliable = (d['reliable']
+                    & (d['pStar'] > 0.9)
+                    & -d['deblend']
+                    & -d['brightNeighb']
+                    & (d['r'] < 18)
+                    )
+    rmi.append(d['rmi'][veryreliable])
+    rmha.append(d['rmha'][veryreliable])
+    #plt.scatter(d['rmi'][veryreliable], d['rmha'][veryreliable])
+
+
+plt.figure()
+plt.subplots_adjust(0.16, 0.20, 0.98, 0.98)
+
+vmax = 500
+hist, xedges, yedges = np.histogram2d(np.concatenate(rmi),
+                                      np.concatenate(rmha),
+                     range=[[-0.3, 3.1], [-0.1, 2.1]], 
+                      bins=[320, 300])
+extent = [xedges[0], xedges[-1], yedges[0], yedges[-1] ]
+hist[hist == 0] = np.nan
+plt.imshow(hist.T, extent=extent, 
+           interpolation='nearest',
+           origin='lower',
+           aspect='auto',
+           vmin=0,
+           vmax=vmax,
+           cmap=COLORMAP)
+
[…]

### 2014-04-16 paper/iphas-dr2.tex
spellcheck

@@
 N. J. Wright$^{1}$,
-A. A. Zijlstra$^{15}$.
+A. A. Zijlstra$^{16}$.
 \\
@@
 $^{13}$Department of Physics, University of Warwick, Gibbet Hill Road, Coventry, CV4 7AL, U.K.\\
-$^{14}$Joint ALMA Observatory, Alonso de Córdova 3107, Vitacura 763-0355, Santiago, Chile. \\
+$^{14}$Joint ALMA Observatory, Alonso de C\'ordova 3107, Vitacura 763-0355, Santiago, Chile. \\
 $^{15}$AAVSO, 49 Bay State Road, Cambridge, MA 02138, USA.\\
@@
 such extrapolation is prone to error,
-and shoud be viewed as indicative rather than as precise measurement
+and should be viewed as indicative rather than as precise measurement
 (systematic uncertainties as a function of magnitude
@@
 that it has not proved necessary to include these internal warning flags
-in the pubished source catalogue.
+in the published source catalogue.
 

### 2014-04-16 paper/figures/footprint/plot-footprint.py
Check-in of footprint plot

@@ -1,0 +1,354 @@
+"""Plots the IPHAS survey footprint."""
+import abc
+import numpy as np
+import matplotlib.pyplot as plt
+from astropy import wcs
+from astropy.io import fits
+
+def equ2gal(equ):
+    """Convert single Equatorial J2000d to Galactic coordinates."""
+    import math as m
+    from math import sin, cos, atan, asin, floor
+    
+    ra, dec = equ
+    OB = m.radians(23.4333334);
+    dec = m.radians(dec)
+    ra = m.radians(ra)
+    a = 27.128251 # The RA of the North Galactic Pole
+    d = 192.859481 # The declination of the North Galactic Pole
+    l = 32.931918 # The ascending node of the Galactic plane on the equator
+    sdec = sin(dec)
+    cdec = cos(dec)
+    sa = sin(m.radians(a))
+    ca = cos(m.radians(a))
+
+    GT = asin(cdec * ca * cos(ra - m.radians(d)) + sdec * sa)
+    GL = m.degrees(atan((sdec - sin(GT) * sa) / (cdec * sin(ra - m.radians(d)) * ca)))
+    TP = sdec - sin(GT) * sa
+    BT = cdec * sin(ra - m.radians(d)) * ca
+    if (BT < 0):
+        GL += 180
+    else:
+        if (TP < 0):
+            GL += 360
+           
+    GL += l
+    if (GL > 360):
+        GL -= 360
+   
+    LG = floor(GL)
+    LM = floor((GL - floor(GL)) * 60)
+    LS = ((GL - floor(GL)) * 60 - LM) * 60
+    GT = m.degrees(GT)
+
+    D = abs(GT)
+    if (GT > 0):
+        BG = floor(D)
+    else:
+        BG = -1*floor(D)
+        
+    BM = floor((D - floor(D)) * 60)
+    BS = ((D - floor(D)) * 60 - BM) * 60
+    if (GT < 0):
+        BM = -BM
+        BS = -BS
+    
+    #if GL > 180:
+    #    GL -= 360
+    return [GL,GT]
+    
+def equ2gal_array(equ):
+    """Convert Equatorial J2000d to Galactic coordinates."""
+    out = []
+    for e in equ:
+        out.append( equ2gal(e) )
+    return np.array(out)
+
+
+class Survey(object):
+    """Common base class for photometric surveys."""
+    
+    def __init__(self, name):
+        self.name = name
+    
+    
+    @abc.abstractmethod
[…]

### 2014-04-16 paper/iphas-dr2.tex
Polished Sec 1+2+3

@@
 S. E. Sale$^8$, 
-M. J. Barlow$^9$,
-P. J. Carter$^{13}$,
-R. L. M Corradi$^{6,7}$,
-J. J. Drake$^{10}$, \newauthor
-J. Eisl\"offel$^{11}$, 
-J. Fabregat$^{12}$, 
-B. T. G\"ansicke$^{13}$,
-N. P. Gentile Fusillo$^{13}$,
-S. Greiss$^{13}$, \newauthor
-A. S. Hales$^{14}$, 
-A. A. Henden$^{15}$,
+A. A. Henden$^9$,
+M. J. Barlow$^{10}$,
+P. J. Carter$^{11}$,
+R. L. M Corradi$^{6,7}$, \newauthor
+J. J. Drake$^{12}$, 
+J. Eisl\"offel$^{13}$, 
+J. Fabregat$^{14}$, 
+B. T. G\"ansicke$^{11}$,
+N. P. Gentile Fusillo$^{11}$, \newauthor
+S. Greiss$^{11}$, 
+A. S. Hales$^{15}$, 
 S. Hodgkin$^2$,
 L. Huckvale$^{16}$,
-J. Irwin$^{9}$,
+J. Irwin$^{12}$, 
 C. Knigge$^{17}$, \newauthor
@@
 S. Phillipps$^{19}$, 
-R. Raddi$^{13}$, 
+R. Raddi$^{11}$, 
 P. Rodr\'\i guez-Gil$^{6,7}$, \newauthor
@@
 S. Scaringi$^{25,26}$,
-D. Steeghs$^{13}$, 
+D. Steeghs$^{11}$, 
 Y. C. Unruh$^{27}$, 
@@
 $^{8}$Rudolf Peierls Centre for Theoretical Physics, Keble Road, Oxford, OX1 3NP, U.K.\\
-$^{9}$University College London, Department of Physics \& Astronomy, 
+$^{9}$AAVSO, 49 Bay State Road, Cambridge, MA 02138, USA.\\
+$^{10}$University College London, Department of Physics \& Astronomy, 
 Gower Street, London WC1E 6BT, U.K.\\
-$^{10}$Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, 
+$^{11}$Department of Physics, University of Warwick, Gibbet Hill Road, Coventry, CV4 7AL, U.K.\\
+$^{12}$Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, 
 Cambridge, MA 02138, U.S.A. \\
-$^{11}$Th\"uringer Landessternwarte, Sternwarte 5, D-07778, Tautenburg, Germany. \\
-$^{12}$Observatorio Astron\'omico, Universidad de Valencia,
+$^{13}$Th\"uringer Landessternwarte, Sternwarte 5, D-07778, Tautenburg, Germany. \\
+$^{14}$Observatorio Astron\'omico, Universidad de Valencia,
 Catedr\'atico Jos\'e Beltr\'an 2, 46980 Paterna, Spain.\\
-$^{13}$Department of Physics, University of Warwick, Gibbet Hill Road, Coventry, CV4 7AL, U.K.\\
-$^{14}$Joint ALMA Observatory, Alonso de C\'ordova 3107, Vitacura 763-0355, Santiago, Chile. \\
-$^{15}$AAVSO, 49 Bay State Road, Cambridge, MA 02138, USA.\\
+$^{15}$Joint ALMA Observatory, Alonso de C\'ordova 3107, Vitacura 763-0355, Santiago, Chile. \\
 %$^{16}$Jodrell Bank Centre for Astrophysics, School of Physics \& Astronomy, University of Manchester, Manchester M13 9PL, U.K.\\
@@
 with respect to these previous surveys 
-by a factor $\sim$1000 (7 magnitudes), 
+by a factor $\sim$1000 (7~magnitudes), 
 IPHAS can expand
@@
 The goal of this paper is to present the next release 
-that takes the coverage up to 92 percent of the survey area 
+that takes the coverage up to 92 per cent of the survey area 
 and includes a global photometric re-calibration.
@@
 \begin{figure*}
-        \includegraphics[width=1\linewidth]{./plots/footprint/footprint_small.png}
+        \includegraphics[width=1\linewidth]{figures/footprint/footprint_large.jpg}
         \caption{Survey area showing the footprints
@@
         Regions with incomplete data are apparent as white gaps (no data) 
-        or in light grey (denoting that only one offset is missing).
-        The dark vertical strip near $\ell \sim$125\degr\ 
+        or in light grey (denoting that one offset is missing).
+        The dark vertical strip near $\ell \simeq$125\degr\ 
         is an arbitrary consequence of the tiling pattern,
@@
 To cover the Northern Plane with some overlap,
-the survey area was divided into 7635 telescope pointings.
+the survey area was divided into 7,635 telescope pointings.
 Each of these pointings is accompanied by an offset position
-displaced by $+$5 arcmin in declination 
-and $+$5 arcmin in right ascension,
+displaced by $+$5 arcmin in Dec 
+and $+$5 arcmin in RA,
 to deal with inter-CCD gaps, detector imperfections,
 and to enable quality checks. 
-Hence, the basic unit of observation hence
+Hence, the basic unit of observation
 amounts to $2 \times 3$ exposures, 
@@
 and the combination of two fields at a small offset as a \emph{field pair}.
-Altogether, the survey contains 15\,270 fields grouped into 7\,635 field pairs.
+Altogether, the survey contains 15,270 fields
+grouped into 7,635 field pairs.
 To achieve the desired survey depth
@@
 in the months of November-December,
-which is when the La Palma weather and seeing conditions are often poor
+which is when the La Palma weather and seeing conditions are often poor,
 forcing many (unsuccessful) repeat observations.
@@
 flagged.
-%we will explain that objects affected in this way 
-%are flagged in the catalogue using the \emph{deblend} warning flag.
 
@@
 the CASU pipeline had processed
-74\,195 single-band IPHAS exposures 
-in which a total of 1.9~billion \emph{candidate sources} were detected 
-at the default detection level of 1.25\,$\sigma$.
-This total inevitably includes spurious detections, artefacts and
+74,195 single-band IPHAS exposures 
+in which a total of 1.9~billion \emph{candidate} detections
+were made at the default detection level of 1.25\,$\sigma$.
+This total inevitably includes spurious objects, artefacts and
 duplicate detections;
@@
 for the purpose of transforming the raw
-detection tables into a reliable source catalogue which is 
+detection tables into a more reliable source catalogue which is 
 (i) quality-controlled,
@@
 To a lesser extent, the wide spread in the $r$-band depth
-is also explained by the shorter $r$-band exposure time
-that was used in the first year of data-taking.
+is also explained by the shorter exposure time
+that was used for this band during the first months of data-taking.
 
@@
 with $b$ the semi-minor and $a$ the semi-major axis,
-is a morphological measure of the elongation of the point spread function (PSF).
-It is expected to be zero (circular) across the field 
+is a morphological measure of the elongation
+of the Point Spread Function (PSF).
+It is expected to be zero (circular)
 in a perfect imaging system,
@@
 due to mechanical glitches in the telescope tracking system.
-To exclude these, we rejected exposures in which the mean ellipticity
-\emph{across the detectors} exceeded $e > 0.3$.
+To exclude these, we rejected exposures
+in which the \emph{average} ellipticity
+across the detectors exceeded $e > 0.3$.
 We have also found that, as this threshold is breached,
@@
 across the detectors,
-which is a proxy for the combined effect of atmospheric and dome seeing.
+which is a proxy for the image quality
+arising from the combined effects of atmospheric and dome seeing.
 This target is currently attained across 86 per cent of the footprint.
 To increase the sky area offered by the data release slightly,
-we have decided to accept data obtained with seeing up to 2.5~arcsec.
+we have decided to accept data obtained with FWHM up to 2.5~arcsec.
 Fig.~\ref{fig:seeing} presents the distribution
@@
 best-available seeing for a given object,
-and that the PSF FWHM measurement
-is included as a column in the catalogue.
+and that the FWHM measurement
+is available as a column in the catalogue.
 
@@
 for stars observed in both pointings.
-We rejected field pairs in which more than 2 per cent of the stars 
+We automatically rejected field pairs
+in which more than 2 per cent of the stars 
 showed an inconsistent measurement at the level of 0.2~mag,
@@
 
-(5) \emph{Source density mapping.}
-Spatial maps showing the number density of the detected sources
-down to 20th magnitude were created to verify the health
-of the data and to check for unexpected artefacts.
-In particular, we created density maps
-which showed the number of \emph{unique} sources
-obtained by cross-matching the detection tables of
-all three bands with a maximum matching distance of 1~arcsec.
-This was particularly effective for revealing
-fields with an inaccurate astrometric solution in one of the bands,
-which were subsequently corrected.
-
-(6) \emph{Visual examination.}
+(5) \emph{Visual examination.}
 Images, colour mosaics,
@@
 by placing them on a black-list.
+
+(6) \emph{Source density mapping.}
+Spatial maps showing the number density of the detected sources
+down to 20th magnitude were created to verify the health
+of the data and to check for unexpected artefacts.
+In particular, we created density maps
+which showed the number of \emph{unique} sources
+obtained by cross-matching the detection tables of
+all three bands with a maximum matching distance of 1~arcsec.
+This was effective for revealing
+fields with an inaccurate astrometric solution in one of the bands,
+which were subsequently corrected.
 
@@
 
-All of the above criteria were satisfied by at least one observing attempt
-for 14\,115 out of the 15\,270 fields (92 per cent).
+The above criteria were satisfied by at least one observing attempt
+for 14,115 out of the 15,270 planned fields (92 per cent).
 In some cases more than one successful attempt to observe
@@
 The images and detection tables of the discarded data are made
-available through our website for this reason (www.iphas.org),
+available through our website (www.iphas.org),
 but will be ignored in the remainder of this work.

### 2014-04-16 paper/iphas-dr2.tex
Polishing Sec 4

@@
 showed an inconsistent measurement at the level of 0.2~mag,
-or more than 26 per cent were inconsistent at the level of 0.1~mag.
+or more than 25 per cent were inconsistent at the level of 0.1~mag.
 These limits were set empirically after inspecting
@@
 \textsc{extinct} is the atmospheric extinction coefficient 
-(set in the pipeline at 0.09 for $r$ and 0.05 for $i$ as representative
-averages for La Palma),
+(set in the pipeline at 0.09 for $r$ and 0.05 for $i$
+as representative averages for La Palma),
 \textsc{airmass} is the normalised optical path length 
-through the atmosphere and
-\textsc{apcor} is a correction for the flux
-lost outside of the aperture used
-(we adopt a 2\farcs 3-diameter circular aperture by default).
+through the atmosphere,
+and \textsc{apcor} is a correction for the flux
+lost outside of the aperture used.
 Finally, \textsc{percorr} is a correction based on the difference
@@
 The broadband zeropoints were determined such that the resulting magnitude system
-refers to the spectral energy distribution (SED) of Vega 
+refers to the Spectral Energy Distribution (SED) of Vega 
 as the zero colour object. 
@@
              These notations follow the definitions by \citet{Bessell2012}.
-             After multiplying $\langle f_\lambda\rangle$ by the EW,
-             we find that the detected flux for Vega in \ha\ is 3.14~magnitudes less
-             than that received in $r$, 
+             When multiplying $\langle f_\lambda\rangle$ by EW,
+             we find that the flux received for Vega
+             is 3.14~mag less in \ha\ than in $r$. 
           }
@@
 where we show the combined colour-colour diagram
-for the set of 2\,801 quality-approved fields
+for all of the 2,801 quality-approved fields
 which are located towards the Galactic anti-centre
 ($160\degr < \ell < 200\degr$).
-Despite the fact that all the fields included in the diagram
-have been quality-approved,
-the main locus of stars is poorly defined
-due to the presence of calibration errors,
-which regularly exceed 0.1~mag.
+The main locus of stars is poorly defined
+in the diagram due to the presence of 
+fields which suffer from calibration errors,
+which can exceed 0.1~mag or more.
 For comparison, Fig.~\ref{fig:ccd_after}
 shows the identical diagram
-obtained after re-calibrating.
+constructed after re-calibrating.
 

### 2014-04-16 paper/iphas-dr2.tex
Polishing Sec 4

@@
 where we show the combined colour-colour diagram
-for all of the 2,801 quality-approved fields
+for all 2,801 quality-approved fields
 which are located towards the Galactic anti-centre
@@
 in the diagram due to the presence of
-fields which suffer from calibration errors,
-which can exceed 0.1~mag or more.
+incorrectly calibrated fields.
+The calibration errors can exceed 0.1~mag or more
+and need to be corrected for 
+during a global re-calibration procedure.
 For comparison, Fig.~\ref{fig:ccd_after}
-shows the identical diagram
-constructed after re-calibrating.
+shows the identical diagram which was constructed
+after our re-calibration.
 
@@
 \end{subfigure}
-\caption{IPHAS $(r-\ha,\ r-i)$ colour-colour diagram for an area of 400~deg$^2$, plotted before (panel a) and after (panel b)
-re-calibration.
-Both figures were created by combining the stars detected
-across the entire set of 2\,801 quality-approved fields which are located
-towards the Galactic anti-centre  ($160\degr < \ell < 200\degr$).
+\caption{IPHAS $(r-\ha,\ r-i)$ colour-colour diagram
+         covering an area of 400~deg$^2$,
+         shown before (panel a) and after (panel b)
+         our global re-calibration was applied.
+         Both figures were created by combining stars
+         detected across all 2,801 quality-approved fields
+         which are located towards the Galactic anti-centre
+         (here defined as $160\degr<\ell<200\degr$).
          The diagrams are plotted as 2D-histograms
@@
          while bins containing more than 400 sources are bright yellow.
-We have included all stars which are brighter than $r<18$
-and are classified as `very reliable'
-(these criteria will be explained in \S\ref{sec:qualitycriteria}).}
+         The diagrams include all stars
+         brighter than $r<18$ which were
+         classified by the pipeline as being `very reliable'
+         (to be explained in
+         \S\ref{sec:qualitycriteria}).}
 \end{figure}
 
-The calibration variations need to be corrected for 
-during a global re-calibration procedure.
-Notable past examples include the global re-calibration 
-of 2MASS \citep[][]{Nikolaev2000},
+Notable past examples of surveys which required
+global re-calibration include 
+2MASS \citep[][]{Nikolaev2000},
 SDSS \citep[][]{Padmanabhan2008}

### 2014-04-16 paper/iphas-dr2.tex
Finished polishing Sec 4

@@
 which are located towards the Galactic anti-centre
-($160\degr < \ell < 200\degr$).
+(defined here as $160\degr < \ell < 200\degr$).
 The main locus of stars is poorly defined
@@
 shows the identical diagram which was constructed
-after our re-calibration.
+after our re-calibration procedure.
 
@@
          which are located towards the Galactic anti-centre
-         (here defined as $160\degr<\ell<200\degr$).
+         ($160\degr<\ell<200\degr$).
          The diagrams are plotted as 2D-histograms
@@
          (to be explained in
-         \S\ref{sec:qualitycriteria}).}
+         \S\ref{sec:qualitycriteria}).
+         The objects seen to fall above the locus of stars
+         \emph{after} re-calibration 
+         are likely to be genuine \ha\ emission-line objects.}
 \end{figure}
@@
 that is globally consistent to within 0.01--0.02~mag.
-
 Surveys which observe identical stars at different epochs
@@
 are used to fit the calibration parameters
-\citep[e.g.][]{Ivezic2007,Padmanabhan2008,Schlafly2012}.
+\citep{Ivezic2007,Padmanabhan2008,Schlafly2012}.
 
@@
 We have found the information contained
-in these overlap regions to be insufficient
+in these narrow overlap regions to be insufficient
 to constrain the calibration parameters.
@@
 than the original nightly calibrations of IPHAS
-which are accurate to $\sim$10 per cent (Tab.~\ref{tbl:offsets_before}).
+which are only accurate to $\sim$10 per cent
+when compared to APASS (Table~\ref{tbl:offsets_before}).
 APASS achieves its uniform accuracy 
@@
 
-The resulting set of $\sim$220\,000 cross-matched stars were then used 
+The resulting set of 220,000 cross-matched stars were then used 
 to derive APASS-to-IPHAS magnitude transformations
@@
 \end{align}
-The root mean square (rms) residuals of these transformations 
+The Root Mean Square (rms) residuals of these transformations 
 are 0.041 and 0.051, respectively.
@@
 The offsets follow a near-Gaussian distribution
-centred around $0.014\pm0.104$~mag in $r$ and $0.007\pm0.108$~mag in $i$
+with mean and sigma $0.014\pm0.104$~mag in $r$
+and $0.007\pm0.108$~mag in $i$
 (Table~\ref{tbl:offsets_before}).
-A total of 4596 fields showed a systematic offset
-exceeding $\pm$0.03~mag in either $r$ or $i$.
+A total of 4,596 IPHAS fields showed a systematic offset
+exceeding $\pm$0.03~mag in either $r$ or $i$
+when compared to APASS.
 
@@
 
-We then applied the most important step in our calibration scheme,
-which is to adjust the provisional zeropoints of these 4596 aberrant fields
+We then applied the most important step
+in our re-calibration scheme,
+which is to adjust the provisional zeropoints
+of these 4,596 aberrant fields
 such that their offset is brought to zero.
@@
 
+\subsubsection{Adjusting fields not covered by APASS}
+
+At the time of writing, the APASS catalogue did not provide 
+sufficient coverage for 7,359 of our fields.
+Fortunately, these fields are located mainly 
+at low Galactic longitudes (cf. Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}),
+which were typically observed during the summer months
+when photometric conditions are more prevalent at the telescope.
+These remaining fields have nevertheless 
+been brought onto the same uniform scale 
+by employing an ubercalibration-style scheme,
+which minimises the magnitude offsets between stars
+located in the overlap regions between neighbouring fields.
+
+An algorithm for achieving this minimisation
+has previously been described by~\citet{Glazebrook1994}.
+In brief, there are two fundamental quantities to be
+minimised between each pair of overlapping exposures, 
+denoted by the indices $i$ and $j$. 
+First, the mean magnitude difference between stars in the overlap
+region $\Delta_{ij}=\langle m_i-m_j\rangle=-\Delta_{ji}$ is a local
+constraint. 
+Second, to ensure the solution does not stray far 
+from the existing calibration, 
+the difference in zeropoints 
+$\Delta\mathrm{ZP}_{ij}=-\Delta\mathrm{ZP}_{ji}$ 
+between each pair of exposures must also be minimised.
+
+Minimisation of these two quantities is a linear least squares problem 
+because the magnitude $m$ depends linearly on the ZP (Eqn.~\ref{eqn:mag}).
+Hence we can find the ZP shift to be applied to each field 
+by minimising the sum:
+\begin{equation}
+   S = \sum_{i=1}^N \sum_{j=1}^N w_{ij} \theta_{ij} (\Delta_{ij} + a_i - a_j)^2,
+   \label{eqn:chi2}
+\end{equation}
+where $i$ denotes an exposure, 
+$j$ an overlapping exposure, 
+$N$ the number of exposures,
+$a_i$ the ZP to solve for, 
+and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$). 
+$w_{ij}$ are weights set to the inverse square of the uncertainty in $\Delta_{ij}$,
+and $\theta_{ij}$ is an overlap function 
+equal to either 1 if exposures $i$ and $j$ overlap or 0 otherwise. 
+Solving for $a_i$ is equivalent to solving $\partial
+S/\partial a_i=0$, which gives the matrix equation:
+\begin{equation}
+   \sum_{j=1}^N A_{ij} a_j = b_j,
+   \label{eqn:matrix}
+\end{equation}
+where 
+\begin{eqnarray}
+   A_{ij} &=& \delta_{ij} \sum_{k=1}^N w_{jk}\theta_{jk} - w_{ij} \theta_{ij},\label{eqn:aij}\\
+   b_i &=& \sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ji} = -\sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ij}.\label{eqn:bi}
+\end{eqnarray}
+
+We enforce a strong external constraint
+on the solution by keeping the zeropoint fixed 
+for the 
+% 6\,756 
+fields which have already been compared
+and calibrated against APASS.
+We hereafter refer to these fields as \emph{anchors}.
+It is asserted that the zeropoints $a_i$ of the anchor fields 
+are known and not solved for.  However they do appear in the vector 
+$b_j$ as constraints.  In addition to the APASS-based anchors, 
+we selected 3,273 additional anchor fields by hand
+to provide additional constraints in regions not covered by APASS.
+These extra anchors were deemed to have reliable zeropoints 
+based on 
+(i) the information contained in the observing logs,
+(ii) the stability of the standard star zeropoints during the night, and
+(iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
+which is located $\sim$500~m from the INT.
+
+We then solved Eqn.~\ref{eqn:matrix} for the $r$ and $i$ bands
+separately using the least-squares routine 
+in Python's {\sc scipy.sparse} module for sparse matrix algebra.
+This provided us with corrected zeropoints for the remaining fields,
+which were shifted on average by $+0.02\pm0.11$ in $r$ 
+and $+0.01\pm0.12$ in $i$ compared to their provisional calibration.
+
+We then turned to the global calibration of the \ha\ data.
+It is not possible to re-calibrate the narrowband 
+in the same way as the broadbands,
+because the APASS survey does not offer \ha\ photometry.
+We can reasonably assume, however,
+that the corrections required for $r$ and \ha\ are identical,
+much of the time, because the IPHAS data-taking pattern ensured 
+that a field's \ha\ and $r$-band exposures
+were taken consecutively,
+albeit separated by a $\sim$30~s read-out time.
+Hence, we have corrected the \ha\ zeropoints 
+by re-using the zeropoint adjustments that were derived for the $r$ band
+in the earlier steps.
+An exception was made for 3,101 fields
+for which our quality-control routines revealed
+strong zeropoint variations to exist during the night,
+suggesting non-photometric conditions.
+In these cases, the \ha\ zeropoints
+were adjusted by solving Eqn.~\ref{eqn:matrix}
+rather than being linked to the $r$-band shift.
+
+\subsection{Testing the calibration against SDSS}
+
+Having re-calibrated all fields
+to the expected APASS accuracy of 3 per cent,
+we turned to a different survey,
+SDSS Data Release 9 \citep{Ahn2012},
+to validate the results.
+SDSS DR9 includes several strips at low
+Galactic latitudes,
+providing data across 18 per cent of the fields in our data release.
+We cross-matched the IPHAS fields against the subset of
+objects marked as reliable stars in the SDSS catalogue\footnote{
+We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
+to obtain photometry from the SDSS {\sc photoprimary} table 
+with criteria {\sc type = star}, {\sc clean = 1} and {\sc score $>$ 0.7}.}, in much the same way as for APASS,
+with the difference of selecting from the  fainter magnitude ranges of 
+$15<r_{\rm SDSS}<18.0$ and $14.5<i_{\rm SDSS}<17.5$.
+This provided us with a set of 1.2~million cross-matched stars.
+
+Colour transformations were again obtained using a sigma-clipped linear least squares fit:
+\begin{eqnarray}
+r_{\rm IPHAS} = r_{\rm SDSS} - 0.093 - 0.044(r-i)_{\rm SDSS} \label{eqn:sdss_r} \\
+i_{\rm IPHAS} = i_{\rm SDSS} - 0.318 - 0.095(r-i)_{\rm SDSS}. \label{eqn:sdss_i}
+\end{eqnarray}
+The rms residuals of these transformations are 0.045 and 0.073, respectively.
+The equations are similar to the ones
+previously determined for APASS,
+although the colour terms are slightly larger.
+The throughput curve of the SDSS $i$-band filter 
+appears to be somewhat more sensitive at longer wavelengths
+than both the IPHAS and APASS filters.
+
+These global transformations were deemed adequate
+for the purpose of validating our calibration in a statistical sense.
+Separate equations were derived towards different sight-lines
+to investigate the effects of varying reddening regimes.
+The colour term was found 
+to show some variation towards weakly reddened areas,
+where different stellar populations are observed.
+The vast majority of red objects in the global sample
+are those in highly reddened areas, however,
+which agree well with the global transformations
+and dominate the statistical appraisal of our calibration.
+
+Having transformed SDSS magnitudes into the IPHAS system,
+we then computed the median magnitude offset for each IPHAS field
+which contained at least 30 objects with a cross-matched counterpart
+in SDSS.  This was the case for 2,602 fields.
+The median offsets for each of these fields
+are shown in Figs.~\ref{fig:sdss_r} and \ref{fig:sdss_i}.
+Importantly, the mean offset and standard deviation found 
+is $-0.001\pm0.029$~mag in $r$
+and $-0.002\pm0.032$~mag in $i$ (Table~\ref{tbl:offsets_after}).
+In comparison, offsets computed in the identical way
+\emph{before} carrying out the re-calibration showed means
+of $+0.016\pm0.088$~mag in $r$ and $+0.010\pm0.089$~mag in $i$ (Table~\ref{tbl:offsets_after}).
+We conclude that our re-calibration procedure has
+been successful in improving the
+uniformity of the calibration
+by a factor three (i.e. from $\sigma=0.088$ to $\sigma=0.029$ in $r$),
+and as such has achieved our aim of bringing the
+accuracy to the aimed level of 0.03~mag.
+
 \begin{figure*}
@@
 \end{figure*}
-
-\subsubsection{Adjusting fields not covered by APASS}
-
-At the time of writing, the APASS catalogue did not provide 
-sufficient coverage for 7359 of our fields.
-Fortunately, these fields are located mainly 
-at low Galactic longitudes (cf. Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}),
-which were typically observed during the summer months
-when photometric conditions are more prevalent at the telescope.
-These remaining fields have nevertheless 
-been brought onto the same uniform scale 
-by employing an ubercalibration-style scheme,
-which minimises the magnitude offsets between stars
-located in the overlap regions between neighbouring fields.
-
-An algorithm for achieving this minimisation
-has previously been described by~\citet{Glazebrook1994}.
-In brief, there are two fundamental quantities to be
-minimised between each pair of overlapping exposures, 
-denoted by the indices $i$ and $j$. 
-Firstly, the mean magnitude difference between stars in the overlap
-region $\Delta_{ij}=\langle m_i-m_j\rangle=-\Delta_{ji}$ is a local
-constraint. 
-Secondly, to ensure the solution does not stray far 
-from the existing calibration, 
-the difference in zeropoints 
-$\Delta\mathrm{ZP}_{ij}=-\Delta\mathrm{ZP}_{ji}$ 
-between each pair of exposures must also be minimised.
-
-Minimisation of these two quantities is a linear least squares problem 
-because the magnitude $m$ depends linearly on the ZP (Eqn.~\ref{eqn:mag}).
-Hence we can find the ZP shift to be applied to each field 
-by minimising the sum:
-\begin{equation}
-   S = \sum_{i=1}^N \sum_{j=1}^N w_{ij} \theta_{ij} (\Delta_{ij} + a_i - a_j)^2,
-   \label{eqn:chi2}
-\end{equation}
-where $i$ denotes an exposure, 
-$j$ an overlapping exposure, 
-$N$ the number of exposures,
-$a_i$ the ZP to solve for 
-and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$). 
-$w_{ij}$ are weights set to the inverse square of the uncertainty in $\Delta_{ij}$,
-and $\theta_{ij}$ is an overlap function 
-equal to either 1 if exposures $i$ and $j$ overlap or 0 otherwise. 
-Solving for $a_i$ is equivalent to solving $\partial
-S/\partial a_i=0$, which gives the matrix equation:
-\begin{equation}
-   \sum_{j=1}^N A_{ij} a_j = b_j,
-   \label{eqn:matrix}
-\end{equation}
-where 
-\begin{eqnarray}
-   A_{ij} &=& \delta_{ij} \sum_{k=1}^N w_{jk}\theta_{jk} - w_{ij} \theta_{ij},\label{eqn:aij}\\
-   b_i &=& \sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ji} = -\sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ij}.\label{eqn:bi}
-\end{eqnarray}
-
-We enforce a strong external constraint
-on the solution by keeping the zeropoint fixed 
-for the 
-% 6\,756 
-fields which have already been compared
-and calibrated against APASS.
-We hereafter refer to these fields as \emph{anchors}.
-It is asserted that the zeropoints $a_i$ of the anchor fields 
-are known and not solved for.  However they do appear in the vector 
-$b_j$ as constraints.  In addition to the APASS-based anchors, 
-we selected 3\,273 additional anchor fields by hand
-to provide additional constraints in regions not covered by APASS.
-These extra anchors were deemed to have reliable zeropoints 
-based on 
-(i) the information contained in the observing logs,
-(ii) the stability of the standard star zeropoints during the night, and
-(iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
-which is located $\sim$500~m from the INT.
-
-We then solved Eqn.~\ref{eqn:matrix} for the $r$ and $i$ bands
-separately using the least-squares routine 
-in Python's {\sc scipy.sparse} module for sparse matrix algebra.
-This provided us with corrected zeropoints for the remaining fields,
-which were shifted on average by $+0.02\pm0.11$ in $r$ 
-and $+0.01\pm0.12$ in $i$ compared to their provisional calibration.
-The outcome is validated in the next section.
-
-We then turned to the global calibration of the \ha\ data.
-It is not possible to re-calibrate the narrowband 
-in the same way as the broadbands,
-because the APASS survey does not offer \ha\ photometry.
-We can reasonably assume however,
-that the corrections required for $r$ and \ha\ are identical,
-much of the time, because the IPHAS data-taking pattern ensured 
-that a field's \ha\ and $r$-band exposures
-were taken consecutively,
-albeit separated by the $\sim$30~s read-out time.
-Hence, we have corrected the \ha\ zeropoints 
-by re-using the zeropoint adjustments that were derived for the $r$ band
-in the earlier steps.
-An exception was made for a small subset of fields
-% 3\,101 fields
-for which our quality-control routines revealed
-strong zeropoint variations during the night,
-suggesting non-photometric conditions.
-In these cases, the \ha\ zeropoints
-were adjusted by solving Eqn.~\ref{eqn:matrix}
-rather than being linked to the $r$-band shift.
-
-\subsection{Testing the calibration against SDSS}
-
-Having re-calibrated all fields to the expected APASS accuracy of 3 per cent,
-we turned to a different survey, SDSS Data Release 9 \citep{Ahn2012},
-to validate the results.
-SDSS DR9 includes several strips at low
-Galactic latitudes,
-providing data across 18 per cent of the fields in our data release.
-We cross-matched the IPHAS fields against the subset of
-objects marked as reliable stars in the SDSS catalogue\footnote{
-We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
-to obtain photometry from the SDSS {\sc photoprimary} table 
-with criteria {\sc type = star}, {\sc clean = 1} and {\sc score $>$ 0.7}.}
-cross-matching in much the same way as for APASS,
-with the difference of selecting from the  fainter magnitude ranges of 
-$15<r_{\rm SDSS}<18.0$ and $14.5<i_{\rm SDSS}<17.5$.
-This provided us with a set of 1.2~million cross-matched stars.
-
-Colour transformations were again obtained using a sigma-clipped linear least squares fit:
-\begin{eqnarray}
-r_{\rm IPHAS} = r_{\rm SDSS} - 0.093 - 0.044(r-i)_{\rm SDSS} \label{eqn:sdss_r} \\
-i_{\rm IPHAS} = i_{\rm SDSS} - 0.318 - 0.095(r-i)_{\rm SDSS}. \label{eqn:sdss_i}
-\end{eqnarray}
-The rms residuals of these transformations are 0.045 and 0.073, respectively.
-The equations are similar to the ones
-previously determined for APASS,
-although the colour terms are slightly larger.
-The throughput curve of the SDSS $i$-band filter 
-appears to be somewhat more sensitive at longer wavelengths
-than both the IPHAS and APASS filters.
-
-These global transformations were deemed adequate
-for the purpose of validating our calibration in a statistical sense.
-Separate equations were derived towards different sight-lines
-to investigate the effects of varying reddening regimes.
-The colour term was found 
-to show some variation towards weakly reddened areas,
-where different stellar populations are observed.
-The vast majority of red objects in the global sample
-are those in highly reddened areas, however,
-which agree well with the global transformations
-and dominate the statistical appraisal of our calibration.
-
-Having transformed SDSS magnitudes into the IPHAS system,
-we then computed the median magnitude offset for each IPHAS field
-which contained at least 30 objects with a cross-matched counterpart
-in SDSS.  This was the case for 2\,602 fields.
-The median offsets for each of these fields
-are shown in Figs.~\ref{fig:sdss_r} and \ref{fig:sdss_i}.
-Importantly, the mean offset and standard deviation found 
-is $-0.001\pm0.029$~mag in $r$
-and $-0.002\pm0.032$~mag in $i$ (Table~\ref{tbl:offsets_after}).
-In comparison, offsets computed in the identical way
-\emph{before} carrying out our re-calibration showed means
-of $+0.016\pm0.088$~mag in $r$ and $+0.010\pm0.089$~mag in $i$ (Table~\ref{tbl:offsets_after}).
-We conclude that our re-calibration procedure has
-been successful in improving the
-uniformity of the calibration by a factor three
-and has achieved our aim of bringing the
-accuracy down to the level of 0.03~mag (rms).
 
@@
 we discovered towards Sh 2-82 as well.
-Two of our candidates, \#1 and \#4 in Tab.~\ref{tbl:emitters},
+Two of our candidates, \#1 and \#4 in Table~\ref{tbl:emitters},
 have previously been identified

### 2014-04-16 paper/iphas-dr2.tex
Finished polishing Sec 5

@@
 Enhancement of the detection tables by creating new columns
-is the necessary first step because the tables generated by the CASU pipeline
-refer e.g. to source positions in pixel coordinates, to photometric measures 
-in number counts etc., rather than in directly useful astronomical terms.
+is the necessary first step 
+because the tables generated by the CASU pipeline
+refer e.g. to source positions in pixel coordinates,
+to photometric measures in number counts, etc.,
+rather than in directly useful astronomical terms.
 To transform these data into user-friendly quantities,
@@
 which are in use at the 
-WFCAM Science Archive \citep[WSA;][]{Hambly2008}
-and the VISTA Science Archive \citep[VSA;][]{Cross2012}.
+Wide Field Camera (WFCAM) Science Archive \citep[WSA;][]{Hambly2008}
+and the 
+Visible and Infrared Survey Telescope for Astronomy (VISTA)
+Science Archive \citep[VSA;][]{Cross2012}.
 These archives curate the high-resolution near-infrared photometry from both
-the UKIRT Infrared Deep Sky Survey \citep[UKIDSS;][]{Lawrence2007}
+the United Kingdom Infrared Telescope (UKIRT)
+Infrared Deep Sky Survey \citep[UKIDSS;][]{Lawrence2007}
 and the 
@@
 data release and the remainder of the string
-denotes the J2000 coordinates in sexagesimal format.
+denotes the J2000 ICRS coordinates in sexagesimal format.
 For convenience, the coordinates
@@
 
-Photometry is provided based on the 2.3-arcsec diameter circular aperture
+Photometry is provided based on the 2\farcs 3-diameter circular aperture
 by default (columns \emph{r}, \emph{i}, \emph{ha}).
@@
 Each of these magnitude measurements have been
-corrected for the flux lost outside of their respective apertures,
+corrected for the flux lost outside of the respective apertures,
 using a correction term which is inferred from the
@@
 can be used as a criterion for selecting stellar objects
-with reliable photometry from the catalogue.
+with reliable photometry.
 
@@
 aperture size is also used by the CASU pipeline to provide
-a discrete star/galaxy/noise\footnote{For consistency with the terminology used in the CASU pipeline and the WSA archive, extended objects are classified as `galaxies' in the catalogue.
-This class may however include other types of resolved objects.} classification flag
+a discrete star/galaxy\footnote{For consistency with the terminology that is widespread in the CASU pipeline and the WSA/VSA archives, extended objects are classified as `galaxies'.
+This class will flag any type of resolved object, however.}/noise classification flag
 (\emph{rClass}, \emph{iClass}, \emph{haClass})
@@
 
-The resulting band-merged catalogues were found
-to be reliable for the vast majority of fields.
-We do warn that source blending and confusion is unavoidable
-for faint objects in the most crowded regions of the Galactic Plane;
-in \S\ref{sec:discussion} we will show
-that 19 per cent of the sources in our catalogue
-are flagged as blended objects (column \emph{deblend}),
-and their band-merged data should be treated with care
-because they may have fallen victim to source confusion
-during the band-merging step.
+The resulting band-merged catalogues were inspected
+as part of our quality control procedures,
+and were found to be reliable.
 

### 2014-04-16 paper/iphas-dr2.tex
Finished polishing Sec 6

@@
 \emph{rErr2}, \emph{iErr2}, \emph{haErr2}, \emph{errBits2}).
-Second, detections not included in the catalogue
+Second, tables of detections not included in the catalogue
 are made available on our website.
@@
 by cross-matching the overlap regions of all field catalogues,
-again using a maximum matching distance of 1\arcsec.
+again using a maximum matching distance of 1~arcsec.
 The duplicate detections for each unique source are then ranked according to
-(i) filter coverage, (ii) quality score (column \emph{errBits}),
+(i) filter coverage, (ii) quality score,
 and (iii) the average seeing of stars in the CCD frame rounded to 0.2~arcsec.
@@
 selected above were compiled
-into the 98-column source catalogue
-that is described in Appendix~\ref{app:columns}
-and is made available on our website and through Vizier.
+into the final 98-column source catalogue
+that is described in Appendix~\ref{app:columns}.
 The original unweeded list of sources naturally included 
@@
 \begin{table}
-    \caption{Breakdown of catalogue entries as a function of the band(s)
-             in which the source was detected.
-             We also show the fraction of `confirmed' objects,
-             which we define as those sources detected in more than one field
+    \caption{Breakdown of catalogue sources as a function of the band(s)
+             in which the object was detected.
+             We also show the fraction of `confirmed' sources,
+             which we define as those objects
+             detected in more than one field
              (usually the field pair partner).
-             We also show the average magnitudes in each class.}
+             We also show the average magnitudes for each class.}
     \label{tbl:detections}
@@
 
-Table~\ref{tbl:detections} details the number of bands
-in which the catalogue entries are captured;
+Table~\ref{tbl:detections} details the breakdown of these
+sources as a function of the bands in which they are captured.
 159 million sources are detected in all three filters (73 per cent),
@@
 which suggests that a significant fraction of these entries may be spurious detections.
-Hence, we strongly recommend users not to rely on single-band objects
+We recommend users not to rely on single-band objects
 without inspecting the image data by eye.
@@
   The brightness must not exceed the nominal saturation limit
-  and the peak pixel height must not exceed 55\,000 counts.
+  and the peak pixel height must not exceed 55,000 counts.
   Again, this implicitly requires a detection in all three bands.
@@
    216 million (99\%) &
-   There is no star brighter than $V < 4$ within 10 arcmin, 
-   or brighter than $V < 7$ within 5 arcmin.
+   There is no star brighter than $V < 4$ within 10\arcmin, 
+   or brighter than $V < 7$ within 5\arcmin.
    Such very bright stars cause scattered light and diffraction spikes,
@@
 We will explain below that the
-\emph{veryReliable} category is least complete
-at low Galactic longitudes,
+\emph{veryReliable} category suffers most of
+its incompleteness at low Galactic longitudes,
 where source blending can affect
 up to a quarter of the objects.
-The \emph{veryReliable} flag should hence
-only be used in applications in which 
-completeness is not a concern.
 
@@
 For example, by raising the allowed photometric uncertainties
-from 0.1~mag to 0.2~mag,
+from 0.1~mag to 0.2~mag in Table~\ref{tab:reliable},
 42 million candidate sources would be added to the 109 
@@
 which we found to suit many science applications,
-but certainly not all. Users of the data are encouraged 
+but users are nevertheless encouraged 
 to revise the quality criteria according to their needs.
@@
              columns \emph{rErr}, \emph{iErr} and \emph{haErr}
-             in the catalogue 
+             in the catalogue, 
              and the error-bars show the standard deviations.
@@
 We find the typical uncertainty to reach 0.1~mag near $r=$20.5 
-and $i$/\ha$=19.5$.
+and $i$,\ha$=19.5$.
 We note that the fainter depth in $r$ is compensated
 by the fact that most stars have brighter magnitudes in $i$ and \ha;
-the average colours in the catalogue are
+the average colours for sources in the catalogue are
 $\overline{(r-i)}=1.06\pm0.12$ and $\overline{(r-\ha)}=0.44\pm0.03$.
@@
 We also plot the Poissonian uncertainties for comparison (solid red line).
-We find the mean residual to equal $0.03\pm0.04$~mag
+We find the mean residual and standard deviation to be $0.03\pm0.04$~mag
 across the magnitude ranges 13 to 18 ($r$)
-and 12 to 17 ($i$, \ha),
-consistent with the accuracy of our calibration estimated above.
+and 12 to 17 ($i$, \ha).
 Stars fainter than this range appear to be dominated by photon noise (red line),

### 2014-04-16 paper/iphas-dr2.tex
Finished polishing.

@@
          We find that the \emph{veryReliable} quality criteria are successful
-         at removing objects with large uncertainties.
+         at removing objects with large residuals.
     }
@@
 For example, the average source density near $\ell\simeq 30\degr$
-is roughly 300\,000 objects per square degree,
+is roughly 300,000 objects per square degree,
 which is six times more than the density
@@
 taking in the Galactic latitude range $|b|<5\degr$
-at longitudes $\ell=30$ to 215\degr.
+at longitudes $\ell=30\degr$ to 215\degr.
 The new 98-column catalogue provides single-epoch photometry
@@
 for 219~million unique sources.
-To support its exploitation, we provide a list of recommended quality criteria
-that will permit the selection of objects with reliable colours from 
-the catalogue.  The closing demonstrations highlight the use of the 
-survey's unique $(r-\ha,\ r-i)$ diagram for characterising stellar populations
-and selecting emission-line objects.  More comprehensive applications of IPHAS
-can be found in the works of Sale et al (submitted), which applies DR2 to the problem
-of 3D extinction mapping, and of Sabin et al (in preparation), where the results of a search
+To support its exploitation, 
+we provide a list of recommended quality criteria
+that will permit the selection
+of objects with reliable colours from the catalogue.
+The closing demonstrations highlight the use of the 
+survey's unique $(r-\ha,\ r-i)$ diagram
+for characterising stellar populations
+and selecting emission-line objects.
+More comprehensive applications of IPHAS
+can be found in the works of Sale et al (submitted),
+which applies DR2 to the problem of 3D extinction mapping,
+and of Sabin et al (in preparation),
+where the results of a search
 of the image database for new planetary nebulae is presented.
 
-The current plan is to work toward one further major IPHAS source catalogue, in which 
-the remaining gaps in sky coverage will have been eliminated -- observations aimed at 
-replacing data not meeting the IPHAS minimum quality requirements are continuing.  We will also 
-examine options to further improve the global calibration, perhaps tightening the accuracy to
-better than 2\%.  For example, we have in mind investigating the use of the PanSTARRS photometric 
-ladder (Magnier et al 2013) as a reference set, when it becomes available and we will 
+The current plan is to work toward
+one further major IPHAS source catalogue,
+in which the remaining gaps in sky coverage
+will have been eliminated
+-- observations aimed at replacing data
+not meeting the IPHAS minimum quality requirements
+are continuing.
+We will also examine options to further improve
+the global calibration,
+perhaps tightening the accuracy to better than 2\%.
+For example, we have in mind investigating
+the use of the PanSTARRS photometric ladder \citep{Magnier2013}
+as a reference set,
+when it becomes available for the Galactic Plane, and we will 
 explore improving source recovery in 
-the most dense fields via the implementation of PSF fitting in place of aperture photometry.  
-Finally, the next catalogue will detail all the secondary detections to aid time-domain 
-studies.
+the most dense fields via the implementation
+of PSF fitting in place of aperture photometry.  
+Finally, the next catalogue will detail
+all the secondary detections
+to aid time-domain studies.
 

### 2014-04-16 paper/figures/depth/depth.py
Checking of limiting magnitude plots

@@ -1,0 +1,41 @@
+from astropy.io import fits
+import pylab as p
+
+d = fits.getdata('/home/gb/dev/iphas-qc/qcdata/iphas-qc.fits', 1)
+
+mask = d['is_dr2']
+
+
+for band in ['r', 'i', 'h']:
+
+    p.figure()
+    n, bins, patches = p.hist(d[band+'5sig_judged'][mask], 
+                              range=[18,23],
+                              bins=50,
+                              lw=1.0,
+                              color='#dddddd')
+
+
+    p.minorticks_on()
+    p.ylabel('Fields')
+    p.xlabel('$5\sigma$ limiting magnitude')
+    p.xlim([19, 22.0])
+    #p.ylim([0, 14500])
+    #p.legend(loc='lower right')
+
+    if band == 'h':
+        label = '$\\rm H\\alpha$'
+    else:
+        label = '$'+band+'$'
+
+    p.text(0.12, 0.9, label,
+        horizontalalignment='center',
+        verticalalignment='top',
+        transform=p.axes().transAxes,
+        fontsize=16)
+
+    p.tight_layout(pad=0.5)
+    p.ylim(0, 2600)
+    p.savefig('depth_{0}.pdf'.format(band))
+    
+

### 2014-04-16 paper/iphas-dr2.tex
Checking of limiting magnitude plots
Polishing.

@@
-\documentclass[useAMS,usenatbib]{mn2e}
+\documentclass[a4paper,useAMS,usenatbib]{mn2e}
 \usepackage{amssymb,amsmath}
@@
 \usepackage{subcaption}
+
+\voffset=-0.5in
 
@@
 the limited bright samples of Galactic emission line objects 
-previously available into larger, deeper, and more statistically-robust samples that will 
-better inform our understanding 
+previously available into larger, deeper,
+and more statistically-robust samples
+that will better inform our understanding 
 of the early and late stages of stellar evolution.
@@
     \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/depth_r.pdf} 
+        \includegraphics[width=\textwidth]{figures/depth/depth_r.pdf} 
     \end{minipage}
     \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/depth_i.pdf} 
+        \includegraphics[width=\textwidth]{figures/depth/depth_i.pdf} 
     \end{minipage}
     \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{./plots/depth_h.pdf} 
+        \includegraphics[width=\textwidth]{figures/depth/depth_h.pdf} 
     \end{minipage}
@@
 for the \ha\ passband
-and hence there is no formally recognised flux scale 
-for it.
+and hence there is no formally recognised flux scale for it.
 We can specify here, however, 
@@
 less than the flux captured by the much broader
-$r$ band which includes the \ha\ band.
-Hence to assure zero colour relative to the broadbands,
-we set the default zeropoint for the narrowband to be:
+$r$ band \citep{Gonzalez-Solares2008}.
+Hence to assure that $(r-\ha)=0$ for Vega,
+we set the zeropoint for the narrowband to be:
 \begin{equation}
@@
 We refer to \citet{Drew2005} for a more comprehensive
-discussion on the throughput curves of the bands.
+discussion on the filter throughput curves.
 
@@
 gain variations, exposure time, and the re-calibration shift.
-Users will only need to apply an aperture correction, if necessary.
 
@@
 -- observations aimed at replacing data
-not meeting the IPHAS minimum quality requirements
-are continuing.
+not meeting the quality requirements are continuing.
 We will also examine options to further improve
 the global calibration,
-perhaps tightening the accuracy to better than 2\%.
+perhaps tightening the accuracy to better than 2 per cent.
 For example, we have in mind investigating
@@
 
+\footnotesize{
   \bibliographystyle{mn2e}
   \bibliography{iphas-dr2}
+}
 

### 2014-04-16 paper/iphas-dr2.tex

--- 2014-04-16 paper/iphas-dr2.tex
@@
 We present the first quality-controlled and
-globally calibrated source catalogue
+globally-calibrated source catalogue
 derived from the survey,
@@
 across 92 per cent of the footprint.
-The observations were carried out between 2003 and 2013
+The observations were carried out between 2003 and 2012
 at a median seeing of 1.1~arcsec (sampled at 0.33~arcsec/pixel)
@@
 This large-scale programme of observation
--- spanning almost a decade 
+-- spanning a decade so far 
 and using more than 300 nights in competitive open time 
@@
 By increasing the sensitivity 
-with respect to these previous surveys 
+with respect to these preceding surveys 
 by a factor $\sim$1000 (7~magnitudes), 
@@
 which are otherwise too faint or numerous 
-to be targeted by spectroscopic surveys.
+to be targeted by spectroscopic surveys (cf. the use of
+Stromgren $uvbyH\beta$ photometry at blue wavelengths).
 For such science applications to succeed, however, 
@@
         Regions with incomplete data are apparent as white gaps (no data) 
-        or in light grey (denoting that one offset is missing).
+        or in light grey (indicating that one offset is missing).
         The dark vertical strip near $\ell \simeq$125\degr\ 
         is an arbitrary consequence of the tiling pattern,
-        which was populated starting from 0h in right ascension.}
+        which was populated starting from 0h in Right Ascension.}
         \label{fig:footprint}
@@
 The WFC is a mosaic of 4 CCDs 
-that captures a sky area of close to 0.29~deg$^2$.
+that captures a sky area of close to 0.29~deg$^2$ at a pixel scale of
+0.33 arcsec/pixel.
 To cover the Northern Plane with some overlap,
@@
 The catalogue is structured such that it is clear 
-where a contemporaneous observation of both halves of a field pair
-is available.
+where contemporaneous observations of both halves of a field pair
+are available.
 
@@
 which was a 200 deg$^2$ survey programme carried out 
-between 1998 and 2003 after the WFC was commissioned.
+between 1998 and 2003.
 Because IPHAS uses the same telescope and camera combination,
@@
 for the purpose of transforming the raw
-detection tables into a more reliable source catalogue which is 
-(i) quality-controlled,
-(ii) homogeneously calibrated, and 
+detection tables into a more reliable source catalogue which
+(i) is quality-controlled,
+(ii) is homogeneously calibrated, and 
 (iii) contains user-friendly columns and warning flags.
@@
 The great range in sky brightness this
-produced is the main reason for the wide and bi-modal shape
+produced is behind the wide and bi-modal shape
 of the $r$-band limiting magnitude distribution 
@@
 as evaluated by measuring the average PSF Full Width at Half Maximum (FWHM)
-across the detectors,
-which is a proxy for the image quality
-arising from the combined effects of atmospheric and dome seeing.
+across the detectors.
+% Stating the obvious ...a new insertion? which is a proxy for the image quality
+% arising from the combined effects of atmospheric and dome seeing.
 This target is currently attained across 86 per cent of the footprint.
@@
 50 per cent is better than 1.1~arcsec,
-and 10 per cent is better than 0.8~arcsec
-(the WFC pixel scale is 0.33~arcsec/px).
+and 10 per cent is better than 0.8~arcsec.
+% slipped in earlier (the WFC pixel scale is 0.33~arcsec/px).
 In \S\ref{sec:catalogue} we will explain
@@
 showed an inconsistent measurement at the level of 0.2~mag,
-or more than 25 per cent were inconsistent at the level of 0.1~mag.
+or more than 26 per cent were inconsistent at the level of 0.1~mag.
 These limits were set empirically after inspecting
@@
 are observed contemporaneously --  
-normally within 5 minutes of each other.
+nearly always within 5 minutes of each other.
 %An exception was made for 9 fields where the three exposures 
@@
 \textsc{extinct} is the atmospheric extinction coefficient 
-(set in the pipeline at 0.09 for $r$ and 0.05 for $i$
-as representative averages for La Palma),
+(set in the pipeline at 0.09 for $r$ and 0.05 for $i$ as representative
+averages for the telescope site),
 \textsc{airmass} is the normalised optical path length 
-through the atmosphere,
-and \textsc{apcor} is a correction for the flux
-lost outside of the aperture used.
+through the atmosphere and
+\textsc{apcor} is a correction for the flux
+lost outside of the aperture used
+(we adopt a 2\farcs 3-diameter circular aperture by default).
 Finally, \textsc{percorr} is a correction based on the difference
@@
 and as such is an ancillary correction 
-to account for sporadic gain variations;
+to account for sporadic gain variations --
 it is zero for most exposures.
@@
 The broadband zeropoints were determined such that the resulting magnitude system
-refers to the Spectral Energy Distribution (SED) of Vega 
+refers to the spectral energy distribution (SED) of Vega 
 as the zero colour object. 
@@
 in the IPHAS filter system.
-We refer to \citet{Drew2005} for a more comprehensive
-discussion on the throughput curves of the bands.
+Data on the throughput curves of the bands can be obtained from
+the Isaac Newton Group website\footnote{http://catserver.ing.iac.es/filter/list.php?instrument=WFC where the
+filters are named WFCH6568, WFCSloanR and WFCSloanI}.
 
@@
              and CCD response)
-             and $f_\lambda(\lambda)$ is the model SED for Vega
-             by Kurucz (http://kurucz.harvard.edu/stars.html).
+             and $f_\lambda(\lambda)$ is the CALSPEC SED for Vega
+             \citep{Bohlin2014}.
              For reference, we also provide the 
@@
              These notations follow the definitions by \citet{Bessell2012}.
-             When multiplying $\langle f_\lambda\rangle$ by EW,
-             we find that the flux received for Vega
-             is 3.14~mag less in \ha\ than in $r$. 
+             After multiplying $\langle f_\lambda\rangle$ by the EW,
+             we find that the detected flux for Vega in \ha\ is 3.14~magnitudes less
+             than that received in $r$, 
           }
@@
 \ha &  $1.81 \times 10^{-9}$ &   59.6 &  6568  &   6568    \\
-$i$  &  $1.30 \times 10^{-9}$  &  759.9  &  7674  &   7674   \\
+$i$  &  $1.30 \times 10^{-9}$  &  759.9  &  7674  &   7661   \\
     \bottomrule
@@
 where we show the combined colour-colour diagram
-for all 2,801 quality-approved fields
+for the set of 2\,801 quality-approved fields
 which are located towards the Galactic anti-centre
-(defined here as $160\degr < \ell < 200\degr$).
-The main locus of stars is poorly defined
-in the diagram due to the presence of
-incorrectly calibrated fields.
-The calibration errors can exceed 0.1~mag or more
-and need to be corrected for 
-during a global re-calibration procedure.
+($160\degr < \ell < 200\degr$).
+Despite the fact that all the fields included in the diagram
+have been quality-approved,
+the main locus of stars is poorly defined
+due to the presence of calibration errors,
+which regularly exceed 0.1~mag.
 For comparison, Fig.~\ref{fig:ccd_after}
-shows the identical diagram which was constructed
-after our re-calibration procedure.
+shows the identical diagram
+obtained after re-calibrating.
 
@@
 \end{subfigure}
-\caption{IPHAS $(r-\ha,\ r-i)$ colour-colour diagram
-         covering an area of 400~deg$^2$,
-         shown before (panel a) and after (panel b)
-         our global re-calibration was applied.
-         Both figures were created by combining stars
-         detected across all 2,801 quality-approved fields
-         which are located towards the Galactic anti-centre
-         ($160\degr<\ell<200\degr$).
+\caption{IPHAS $(r-\ha,\ r-i)$ colour-colour diagram for an area of 400~deg$^2$, plotted before (panel a) and after (panel b)
+re-calibration.
+Both figures were created by combining the stars detected
+across the entire set of 2\,801 quality-approved fields which are located
+towards the Galactic anti-centre  ($160\degr < \ell < 200\degr$).
 The diagrams are plotted as 2D-histograms
@@
 while bins containing more than 400 sources are bright yellow.
-         The diagrams include all stars
-         brighter than $r<18$ which were
-         classified by the pipeline as being `very reliable'
-         (to be explained in
-         \S\ref{sec:qualitycriteria}).
-         The objects seen to fall above the locus of stars
-         \emph{after} re-calibration 
-         are likely to be genuine \ha\ emission-line objects.}
+We have included all stars which are brighter than $r<18$
+and are classified as `very reliable'
+(these criteria will be explained in \S\ref{sec:qualitycriteria}).}
 \end{figure}
 
-Notable past examples of surveys which required
-global re-calibration include 
-2MASS \citep[][]{Nikolaev2000},
+The calibration variations need to be corrected for 
+during a global re-calibration procedure.
+Notable past examples include the global re-calibration 
+of 2MASS \citep[][]{Nikolaev2000},
 SDSS \citep[][]{Padmanabhan2008}
@@
 that is globally consistent to within 0.01--0.02~mag.
+
 Surveys which observe identical stars at different epochs
@@
 are used to fit the calibration parameters
-\citep{Ivezic2007,Padmanabhan2008,Schlafly2012}.
+\citep[e.g.][]{Ivezic2007,Padmanabhan2008,Schlafly2012}.
 
@@
 Second, IPHAS is not specified as a variability survey, with
-the consequence that stars are not normally observed at more 
+the result that stars are not normally observed at more 
 than one epoch,
@@
 We have found the information contained
-in these narrow overlap regions to be insufficient
+in these overlap regions to be insufficient
 to constrain the calibration parameters.
@@
 Although these systematics are reasonably small within a single field,
-they can combine, during the global calibration routine, to cause artificial zeropoint gradients 
+they can accumulate during the calibration process, causing artificial zeropoint gradients 
 across the survey unless controlled by other external constraints.
 
-For these reasons, we have decided not to depend
+For these reasons, we have not depended
 on an ubercalibration-type scheme alone,
-and have opted to involve an external reference survey
+but have opted to involve an external reference survey
 -- where available --
@@
 
-We have been able to benefit from the
-AAVSO Photometric All-Sky Survey
-(APASS; http://www.aavso.org/apass)
+We have been able to benefit from APASS
+(http://www.aavso.org/apass)
 to bring most of the survey 
@@
 The photometric accuracy of APASS is currently estimated 
-to be at the level of 3 per cent (Henden, private communication),
+to be at the level of 3 per cent,
+%The private communication from Hended is no longer needed :-)
 which is significantly better 
 than the original nightly calibrations of IPHAS
-which are only accurate to $\sim$10 per cent
-when compared to APASS (Table~\ref{tbl:offsets_before}).
+which are accurate to $\sim$10 per cent (Tab.~\ref{tbl:offsets_before}).
 APASS achieves its uniform accuracy 
@@
 which showed a magnitude offset larger than 0.03~mag against APASS.  Experience of re-running
-the calibration and testing the results showed us that it was inadvisable to more
-finely tune the match for IPHAS data obtained in what were generally the most photometric
+the calibration and testing the results showed us that it was inadvisable to tune more
+finely the match for IPHAS data obtained in what were generally the most photometric
 nights.  To this end,
@@
 
-The resulting set of 220,000 cross-matched stars were then used 
+The resulting set of $\sim$220\,000 cross-matched stars were then used 
 to derive APASS-to-IPHAS magnitude transformations
@@
 \end{align}
-The Root Mean Square (rms) residuals of these transformations 
+The root mean square (rms) residuals of these transformations 
 are 0.041 and 0.051, respectively.
@@
 and IPHAS uses Vega-based magnitudes.
-Separate transformations were derived for sight-lines 
+%I wonder if the next bit is stating the too-obvious?
+%Have shortened it a bit.
+Separate transformations were derived for sightlines 
 with varying extinction properties to investigate the robustness
 of the transformations with respect to different reddening regimes.
-The variations at these different
-sight-lines were found to be insignificant.
-This is not surprising because heavily-reddened objects 
-are naturally scarce at $r<16$.
+This sensitivity was found to be insignificant, in keeping with
+the scarcity of heavily-reddened objects at $r<16$.
 
@@
 The offsets follow a near-Gaussian distribution
-with mean and sigma $0.014\pm0.104$~mag in $r$
-and $0.007\pm0.108$~mag in $i$
+centred around $0.014\pm0.104$~mag in $r$ and $0.007\pm0.108$~mag in $i$
 (Table~\ref{tbl:offsets_before}).
-A total of 4,596 IPHAS fields showed a systematic offset
-exceeding $\pm$0.03~mag in either $r$ or $i$
-when compared to APASS.
+A total of 4596 fields showed a systematic offset
+exceeding $\pm$0.03~mag in either $r$ or $i$.
 
@@
 
-We then applied the most important step
-in our re-calibration scheme,
-which is to adjust the provisional zeropoints
-of these 4,596 aberrant fields
+We then applied the most important step in our calibration scheme,
+which is to adjust the provisional zeropoints of these 4596 aberrant fields
 such that their offset is brought to zero.
@@
 correcting the IPHAS zeropoints was repeated a few times to ensure 
-convergence, which was essentially reached after the first iteration.
+convergence, which was closely approached after the first iteration.
 
@@
 
-\subsubsection{Adjusting fields not covered by APASS}
-
-At the time of writing, the APASS catalogue did not provide 
-sufficient coverage for 7,359 of our fields.
-Fortunately, these fields are located mainly 
-at low Galactic longitudes (cf. Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}),
-which were typically observed during the summer months
-when photometric conditions are more prevalent at the telescope.
-These remaining fields have nevertheless 
-been brought onto the same uniform scale 
-by employing an ubercalibration-style scheme,
-which minimises the magnitude offsets between stars
-located in the overlap regions between neighbouring fields.
-
-An algorithm for achieving this minimisation
-has previously been described by~\citet{Glazebrook1994}.
-In brief, there are two fundamental quantities to be
-minimised between each pair of overlapping exposures, 
-denoted by the indices $i$ and $j$. 
-First, the mean magnitude difference between stars in the overlap
-region $\Delta_{ij}=\langle m_i-m_j\rangle=-\Delta_{ji}$ is a local
-constraint. 
-Second, to ensure the solution does not stray far 
-from the existing calibration, 
-the difference in zeropoints 
-$\Delta\mathrm{ZP}_{ij}=-\Delta\mathrm{ZP}_{ji}$ 
-between each pair of exposures must also be minimised.
-
-Minimisation of these two quantities is a linear least squares problem 
-because the magnitude $m$ depends linearly on the ZP (Eqn.~\ref{eqn:mag}).
-Hence we can find the ZP shift to be applied to each field 
-by minimising the sum:
-\begin{equation}
-   S = \sum_{i=1}^N \sum_{j=1}^N w_{ij} \theta_{ij} (\Delta_{ij} + a_i - a_j)^2,
-   \label{eqn:chi2}
-\end{equation}
-where $i$ denotes an exposure, 
-$j$ an overlapping exposure, 
-$N$ the number of exposures,
-$a_i$ the ZP to solve for, 
-and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$). 
-$w_{ij}$ are weights set to the inverse square of the uncertainty in $\Delta_{ij}$,
-and $\theta_{ij}$ is an overlap function 
-equal to either 1 if exposures $i$ and $j$ overlap or 0 otherwise. 
-Solving for $a_i$ is equivalent to solving $\partial
-S/\partial a_i=0$, which gives the matrix equation:
-\begin{equation}
-   \sum_{j=1}^N A_{ij} a_j = b_j,
-   \label{eqn:matrix}
-\end{equation}
-where 
-\begin{eqnarray}
-   A_{ij} &=& \delta_{ij} \sum_{k=1}^N w_{jk}\theta_{jk} - w_{ij} \theta_{ij},\label{eqn:aij}\\
-   b_i &=& \sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ji} = -\sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ij}.\label{eqn:bi}
-\end{eqnarray}
-
-We enforce a strong external constraint
-on the solution by keeping the zeropoint fixed 
-for the 
-% 6\,756 
-fields which have already been compared
-and calibrated against APASS.
-We hereafter refer to these fields as \emph{anchors}.
-It is asserted that the zeropoints $a_i$ of the anchor fields 
-are known and not solved for.  However they do appear in the vector 
-$b_j$ as constraints.  In addition to the APASS-based anchors, 
-we selected 3,273 additional anchor fields by hand
-to provide additional constraints in regions not covered by APASS.
-These extra anchors were deemed to have reliable zeropoints 
-based on 
-(i) the information contained in the observing logs,
-(ii) the stability of the standard star zeropoints during the night, and
-(iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
-which is located $\sim$500~m from the INT.
-
-We then solved Eqn.~\ref{eqn:matrix} for the $r$ and $i$ bands
-separately using the least-squares routine 
-in Python's {\sc scipy.sparse} module for sparse matrix algebra.
-This provided us with corrected zeropoints for the remaining fields,
-which were shifted on average by $+0.02\pm0.11$ in $r$ 
-and $+0.01\pm0.12$ in $i$ compared to their provisional calibration.
-
-We then turned to the global calibration of the \ha\ data.
-It is not possible to re-calibrate the narrowband 
-in the same way as the broadbands,
-because the APASS survey does not offer \ha\ photometry.
-We can reasonably assume, however,
-that the corrections required for $r$ and \ha\ are identical,
-much of the time, because the IPHAS data-taking pattern ensured 
-that a field's \ha\ and $r$-band exposures
-were taken consecutively,
-albeit separated by a $\sim$30~s read-out time.
-Hence, we have corrected the \ha\ zeropoints 
-by re-using the zeropoint adjustments that were derived for the $r$ band
-in the earlier steps.
-An exception was made for 3,101 fields
-for which our quality-control routines revealed
-strong zeropoint variations to exist during the night,
-suggesting non-photometric conditions.
-In these cases, the \ha\ zeropoints
-were adjusted by solving Eqn.~\ref{eqn:matrix}
-rather than being linked to the $r$-band shift.
-
-\subsection{Testing the calibration against SDSS}
-
-Having re-calibrated all fields
-to the expected APASS accuracy of 3 per cent,
-we turned to a different survey,
-SDSS Data Release 9 \citep{Ahn2012},
-to validate the results.
-SDSS DR9 includes several strips at low
-Galactic latitudes,
-providing data across 18 per cent of the fields in our data release.
-We cross-matched the IPHAS fields against the subset of
-objects marked as reliable stars in the SDSS catalogue\footnote{
-We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
-to obtain photometry from the SDSS {\sc photoprimary} table 
-with criteria {\sc type = star}, {\sc clean = 1} and {\sc score $>$ 0.7}.}, in much the same way as for APASS,
-with the difference of selecting from the  fainter magnitude ranges of 
-$15<r_{\rm SDSS}<18.0$ and $14.5<i_{\rm SDSS}<17.5$.
-This provided us with a set of 1.2~million cross-matched stars.
-
-Colour transformations were again obtained using a sigma-clipped linear least squares fit:
-\begin{eqnarray}
-r_{\rm IPHAS} = r_{\rm SDSS} - 0.093 - 0.044(r-i)_{\rm SDSS} \label{eqn:sdss_r} \\
-i_{\rm IPHAS} = i_{\rm SDSS} - 0.318 - 0.095(r-i)_{\rm SDSS}. \label{eqn:sdss_i}
-\end{eqnarray}
-The rms residuals of these transformations are 0.045 and 0.073, respectively.
-The equations are similar to the ones
-previously determined for APASS,
-although the colour terms are slightly larger.
-The throughput curve of the SDSS $i$-band filter 
-appears to be somewhat more sensitive at longer wavelengths
-than both the IPHAS and APASS filters.
-
-These global transformations were deemed adequate
-for the purpose of validating our calibration in a statistical sense.
-Separate equations were derived towards different sight-lines
-to investigate the effects of varying reddening regimes.
-The colour term was found 
-to show some variation towards weakly reddened areas,
-where different stellar populations are observed.
-The vast majority of red objects in the global sample
-are those in highly reddened areas, however,
-which agree well with the global transformations
-and dominate the statistical appraisal of our calibration.
-
-Having transformed SDSS magnitudes into the IPHAS system,
-we then computed the median magnitude offset for each IPHAS field
-which contained at least 30 objects with a cross-matched counterpart
-in SDSS.  This was the case for 2,602 fields.
-The median offsets for each of these fields
-are shown in Figs.~\ref{fig:sdss_r} and \ref{fig:sdss_i}.
-Importantly, the mean offset and standard deviation found 
-is $-0.001\pm0.029$~mag in $r$
-and $-0.002\pm0.032$~mag in $i$ (Table~\ref{tbl:offsets_after}).
-In comparison, offsets computed in the identical way
-\emph{before} carrying out the re-calibration showed means
-of $+0.016\pm0.088$~mag in $r$ and $+0.010\pm0.089$~mag in $i$ (Table~\ref{tbl:offsets_after}).
-We conclude that our re-calibration procedure has
-been successful in improving the
-uniformity of the calibration
-by a factor three (i.e. from $\sigma=0.088$ to $\sigma=0.029$ in $r$),
-and as such has achieved our aim of bringing the
-accuracy to the aimed level of 0.03~mag.
-
 \begin{figure*}
@@
 \end{figure*}
+
+\subsubsection{Adjusting fields not covered by APASS}
+
+At the time of writing, the APASS catalogue did not provide 
+sufficient coverage for 7359 of our fields.
+Fortunately, these fields are located mainly 
+at low Galactic longitudes (cf. Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}),
+which were typically observed during the summer months
+when photometric conditions are more prevalent at the telescope.
+These remaining fields have nevertheless 
+been brought onto the same uniform scale 
+by employing an ubercalibration-style scheme,
+which minimises the magnitude offsets between stars
+located in the overlap regions between neighbouring fields.
+
+An algorithm for achieving this minimisation
+has previously been described by~\citet{Glazebrook1994}.
+In brief, there are two fundamental quantities to be
+minimised between each pair of overlapping exposures, 
+denoted by the indices $i$ and $j$. 
+Firstly, the mean magnitude difference between stars in the overlap
+region $\Delta_{ij}=\langle m_i-m_j\rangle=-\Delta_{ji}$ is a local
+constraint. 
+Secondly, to ensure the solution does not stray far 
+from the existing calibration, 
+the difference in zeropoints 
+$\Delta\mathrm{ZP}_{ij}=-\Delta\mathrm{ZP}_{ji}$ 
+between each pair of exposures must also be minimised.
+
+Minimisation of these two quantities is a linear least squares problem 
+because the magnitude $m$ depends linearly on the ZP (Eqn.~\ref{eqn:mag}).
+Hence we can find the ZP shift to be applied to each field 
+by minimising the sum:
+\begin{equation}
+   S = \sum_{i=1}^N \sum_{j=1}^N w_{ij} \theta_{ij} (\Delta_{ij} + a_i - a_j)^2,
+   \label{eqn:chi2}
+\end{equation}
+where $i$ denotes an exposure, 
+$j$ an overlapping exposure, 
+$N$ the number of exposures,
+$a_i$ the ZP to solve for 
+and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$). 
+$w_{ij}$ are weights set to the inverse square of the uncertainty in $\Delta_{ij}$,
+and $\theta_{ij}$ is an overlap function 
+equal to either 1 if exposures $i$ and $j$ overlap or 0 otherwise. 
+Solving for $a_i$ is equivalent to solving $\partial
+S/\partial a_i=0$, which gives the matrix equation:
+\begin{equation}
+   \sum_{j=1}^N A_{ij} a_j = b_j,
+   \label{eqn:matrix}
+\end{equation}
+where 
+\begin{eqnarray}
+   A_{ij} &=& \delta_{ij} \sum_{k=1}^N w_{jk}\theta_{jk} - w_{ij} \theta_{ij},\label{eqn:aij}\\
+   b_i &=& \sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ji} = -\sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ij}.\label{eqn:bi}
+\end{eqnarray}
+
+We enforce a strong external constraint
+on the solution by keeping the zeropoint fixed 
+for the 
+% 6\,756 
+fields which have already been compared
+and calibrated against APASS.
+We hereafter refer to these fields as \emph{anchors}.
+It is asserted that the zeropoints $a_i$ of the anchor fields 
+are known and not solved for.  However they do appear in the vector 
+$b_j$ as constraints.  In addition to the APASS-based anchors, 
+we selected 3\,273 additional anchor fields by hand
+to provide additional constraints in regions not covered by APASS.
+These extra anchors were deemed to have reliable zeropoints 
+based on 
+(i) the information contained in the observing logs,
+(ii) the stability of the standard star zeropoints during the night, and
+(iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
+which is located $\sim$500~m from the INT.
+
+We then solved Eqn.~\ref{eqn:matrix} for the $r$ and $i$ bands
+separately using the least-squares routine 
+in Python's {\sc scipy.sparse} module for sparse matrix algebra.
+This provided us with corrected zeropoints for the remaining fields,
+which were shifted on average by $+0.02\pm0.11$ in $r$ 
+and $+0.01\pm0.12$ in $i$ compared to their provisional calibration.
+The outcome is validated in the next section.
+
+We then turned to the global calibration of the \ha\ data.
+It is not possible to re-calibrate the narrowband 
+in the same way as the broadbands,
+because the APASS survey does not offer \ha\ photometry.
+We can reasonably assume however,
+that the corrections required for $r$ and \ha\ are identical,
+much of the time, because the IPHAS data-taking pattern ensured 
+that a field's \ha\ and $r$-band exposures
+were taken consecutively,
+albeit separated by the $\sim$30~s read-out time.
+Hence, we have corrected the \ha\ zeropoints 
+by re-using the zeropoint adjustments that were derived for the $r$ band
+in the earlier steps.
+An exception was made for a small subset of fields
+% 3\,101 fields
+for which our quality-control routines revealed
+strong zeropoint variations during the night,
+suggesting non-photometric conditions.
+In these cases, the \ha\ zeropoints
+were adjusted by solving Eqn.~\ref{eqn:matrix}
+rather than by simply applying Eqn.~\ref{eqn:zpha}.
+
+\subsection{Testing the calibration against SDSS}
+
+Having re-calibrated all fields to the expected APASS accuracy of 3 per cent,
+we turned to a different survey, SDSS Data Release 9 \citep{Ahn2012},
+to validate the results.
+SDSS DR9 includes several strips at low
+Galactic latitudes,
+providing data across 18 per cent of the fields in our data release.
+We cross-matched the IPHAS fields against the subset of
+objects marked as reliable stars in the SDSS catalogue\footnote{
+We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
+to obtain photometry from the SDSS {\sc photoprimary} table 
+with criteria {\sc type = star}, {\sc clean = 1} and {\sc score $>$ 0.7}.}
+cross-matching in much the same way as for APASS,
+with the difference of selecting from the  fainter magnitude ranges of 
+$15<r_{\rm SDSS}<18.0$ and $14.5<i_{\rm SDSS}<17.5$.
+This provided us with a set of 1.2~million cross-matched stars.
+
+Colour transformations were again obtained using a sigma-clipped linear least squares fit:
+\begin{eqnarray}
+r_{\rm IPHAS} = r_{\rm SDSS} - 0.093 - 0.044(r-i)_{\rm SDSS} \label{eqn:sdss_r} \\
+i_{\rm IPHAS} = i_{\rm SDSS} - 0.318 - 0.095(r-i)_{\rm SDSS}. \label{eqn:sdss_i}
+\end{eqnarray}
+The rms residuals of these transformations are 0.045 and 0.073, respectively.
+The equations are similar to the ones
+previously determined for APASS,
+although the colour terms are slightly larger.
+The throughput curve of the SDSS $i$-band filter 
+appears to be somewhat more sensitive at longer wavelengths
+than both the IPHAS and APASS filters.
+
+These global transformations were deemed adequate
+for the purpose of validating our calibration in a statistical sense.
+Separate equations were derived towards different sight-lines
+to investigate the effects of varying reddening regimes.
+The colour term was found 
+to show some variation towards weakly reddened areas,
+where different stellar populations are observed.
+The vast majority of red objects in the global sample
+are those in highly reddened areas, however,
+which agree well with the global transformations
+and dominate the statistical appraisal of our calibration.
+
+Having transformed SDSS magnitudes into the IPHAS system,
+we then computed the median magnitude offset for each IPHAS field
+which contained at least 30 objects with a cross-matched counterpart
+in SDSS.  This was the case for 2\,602 fields.
+The median offsets for each of these fields
+are shown in Figs.~\ref{fig:sdss_r} and \ref{fig:sdss_i}.
+Importantly, the mean offset and standard deviation found 
+is $-0.001\pm0.029$~mag in $r$
+and $-0.002\pm0.032$~mag in $i$ (Table~\ref{tbl:offsets_after}).
+In comparison, offsets computed in the identical way
+\emph{before} carrying out our re-calibration showed means
+of $+0.016\pm0.088$~mag in $r$ and $+0.010\pm0.089$~mag in $i$ (Table~\ref{tbl:offsets_after}).
+We conclude that our re-calibration procedure has
+been successful in improving the
+uniformity of the calibration by a factor three
+and has achieved our aim of bringing the
+accuracy down to the level of 0.03~mag (rms).
 
@@
 Enhancement of the detection tables by creating new columns
-is the necessary first step 
-because the tables generated by the CASU pipeline
-refer e.g. to source positions in pixel coordinates,
-to photometric measures in number counts, etc.,
-rather than in directly useful astronomical terms.
+is the necessary first step because the tables generated by the CASU pipeline
+refer to source positions in pixel coordinates, to photometric measures 
+in number counts, and so on, rather than in common astronomical units.
 To transform these data into user-friendly quantities,
@@
 which are in use at the 
-Wide Field Camera (WFCAM) Science Archive \citep[WSA;][]{Hambly2008}
-and the 
-Visible and Infrared Survey Telescope for Astronomy (VISTA)
-Science Archive \citep[VSA;][]{Cross2012}.
+WFCAM Science Archive \citep[WSA;][]{Hambly2008}
+and the VISTA Science Archive \citep[VSA;][]{Cross2012}.
 These archives curate the high-resolution near-infrared photometry from both
-the United Kingdom Infrared Telescope (UKIRT)
-Infrared Deep Sky Survey \citep[UKIDSS;][]{Lawrence2007}
+the UKIRT Infrared Deep Sky Survey \citep[UKIDSS;][]{Lawrence2007}
 and the 
@@
 data release and the remainder of the string
-denotes the J2000 ICRS coordinates in sexagesimal format.
+denotes the J2000 coordinates in sexagesimal format.
 For convenience, the coordinates
@@
 
-Photometry is provided based on the 2\farcs 3-diameter circular aperture
+Photometry is provided based on the 2.3-arcsec diameter circular aperture
 by default (columns \emph{r}, \emph{i}, \emph{ha}).
@@
 Each of these magnitude measurements have been
-corrected for the flux lost outside of the respective apertures,
+corrected for the flux lost outside of the applied aperture,
 using a correction term which is inferred from the
@@
 the four alternative magnitudes are expected
-to be consistent with each other
+to be consistent with each other to
 within the photon noise uncertainty
@@
 can be used as a criterion for selecting stellar objects
-with reliable photometry.
+with reliable photometry from the catalogue.
 
@@
 aperture size is also used by the CASU pipeline to provide
-a discrete star/galaxy\footnote{For consistency with the terminology that is widespread in the CASU pipeline and the WSA/VSA archives, extended objects are classified as `galaxies'.
-This class will flag any type of resolved object, however.}/noise classification flag
+a discrete star/galaxy/noise\footnote{For consistency with the terminology used in the CASU pipeline and the WSA archive, extended objects are classified as `galaxies' in the catalogue.
+This class may however include other types of resolved objects.} classification flag
 (\emph{rClass}, \emph{iClass}, \emph{haClass})
@@
 alternative detections were available in essentially all these situations
-thanks to the IPHAS field pair strategy. In turn, this has meant
-that it has not proved necessary to include these internal warning flags
+thanks to the IPHAS field pair strategy. 
+%In turn, this has meant that it has not proved necessary 
+Hence there has been no need to include these internal warning flags
 in the published source catalogue.
@@
 
-We employed a maximum matching distance of 1~arcsec,
-which was chosen, trading off completeness against reliability.
+We employed a maximum matching distance of 1~arcsec, trading off completeness against reliability.
 On the one hand, a matching distance larger than 1~arcsec 
@@
 further if necessary
-(columns \emph{iXi}, \emph{iEta}, \emph{haXi}, \emph{haEta}).
+(columns \emph{iXi}, \emph{iEta}, \emph{haXi}, \emph{haEta}), or simply to examine
+light centre differences.
+%I think of Dimitri Purbaix here.
 We note that UKIDSS/GPS adopted 
@@
 use these catalogues with caution.
-
-The resulting band-merged catalogues were inspected
-as part of our quality control procedures,
-and were found to be reliable.
+%GEERT.  Is the above too extreme a warning?  Is the r band the 'master', to which the given source 
+%position refers, when available, and hence is the natural choice for astrometric purposes?
+%The caution applies if you mis-associate the position with e.g. an i source, without
+%importing the offsets.  Or do you really believe no position is better than 1 arcsec because
+%we don't know which band it refers to.
+
+The resulting band-merged catalogues were found
+to be reliable for the vast majority of fields.
+We do warn that source blending and confusion is unavoidable
+for faint objects in the most crowded regions of the Galactic Plane;
+in \S\ref{sec:discussion} we will show
+that 19 per cent of the sources in our catalogue
+are flagged as blended objects (column \emph{deblend}),
+and their band-merged data should be treated with care
+because they may have fallen victim to source confusion
+during the band-merging step.
 
@@
 \emph{rErr2}, \emph{iErr2}, \emph{haErr2}, \emph{errBits2}).
-Second, tables of detections not included in the catalogue
+Second, detections not included in the catalogue
 are made available on our website.
@@
 by cross-matching the overlap regions of all field catalogues,
-again using a maximum matching distance of 1~arcsec.
+again using a maximum matching distance of 1\arcsec.
 The duplicate detections for each unique source are then ranked according to
-(i) filter coverage, (ii) quality score,
+(i) filter coverage, (ii) quality score (column \emph{errBits}),
 and (iii) the average seeing of stars in the CCD frame rounded to 0.2~arcsec.
@@
 selected above were compiled
-into the final 98-column source catalogue
-that is described in Appendix~\ref{app:columns}.
+into the 98-column source catalogue
+that is described in Appendix~\ref{app:columns}
+and is made available on our website and through Vizier.
 The original unweeded list of sources naturally included 
@@
 \begin{table}
-    \caption{Breakdown of catalogue sources as a function of the band(s)
-             in which the object was detected.
-             We also show the fraction of `confirmed' sources,
-             which we define as those objects
-             detected in more than one field
+    \caption{Breakdown of catalogue entries as a function of the band(s)
+             in which the source was detected.
+             We also show the fraction of `confirmed' objects,
+             which we define as those sources detected in more than one field
              (usually the field pair partner).
-             We also show the average magnitudes for each class.}
+             We also show the average magnitudes in each class.}
     \label{tbl:detections}
@@
 
-Table~\ref{tbl:detections} details the breakdown of these
-sources as a function of the bands in which they are captured.
+Table~\ref{tbl:detections} details the number of bands
+in which the catalogue entries are captured;
 159 million sources are detected in all three filters (73 per cent),
@@
 which suggests that a significant fraction of these entries may be spurious detections.
-We recommend users not to rely on single-band objects
+Hence, we strongly recommend users not to rely on single-band objects
 without inspecting the image data by eye.
@@
 is markedly better than for $r$ and \ha,
-which is likely explained by the fact that $i$ is least
+which is most likely explained by the fact that $i$ is least
 affected by interstellar extinction, and so the survey can occasionally pick up
@@
 Moreover, objects which are intrinsically very red may also be picked
-up in $i$ alone, while faint objects with strong Balmer emission
+up in $i$ alone, while faint objects with extremely strong Balmer emission
 may appear only in \ha.
@@
   The brightness must not exceed the nominal saturation limit
-  and the peak pixel height must not exceed 55,000 counts.
+  and the peak pixel height must not exceed 55\,000 counts.
   Again, this implicitly requires a detection in all three bands.
@@
    216 million (99\%) &
-   There is no star brighter than $V < 4$ within 10\arcmin, 
-   or brighter than $V < 7$ within 5\arcmin.
+   There is no star brighter than $V < 4$ within 10 arcmin, 
+   or brighter than $V < 7$ within 5 arcmin.
    Such very bright stars cause scattered light and diffraction spikes,
@@
 We will explain below that the
-\emph{veryReliable} category suffers most of
-its incompleteness at low Galactic longitudes,
+\emph{veryReliable} category is least complete
+at low Galactic longitudes,
 where source blending can affect
 up to a quarter of the objects.
+The \emph{veryReliable} flag should hence
+only be used in applications in which 
+completeness is not a concern.
 
@@
 For example, by raising the allowed photometric uncertainties
-from 0.1~mag to 0.2~mag in Table~\ref{tab:reliable},
+from 0.1~mag to 0.2~mag,
 42 million candidate sources would be added to the 109 
@@
 which we found to suit many science applications,
-but users are nevertheless encouraged 
-to revise the quality criteria according to their needs.
+but certainly not all. Users of the data are encouraged to revise 
+the quality criteria according to their needs.
 
@@
 We find the typical uncertainty to reach 0.1~mag near $r=$20.5 
-and $i$,\ha$=19.5$.
+and $i$/\ha$=19.5$.
 We note that the fainter depth in $r$ is compensated
 by the fact that most stars have brighter magnitudes in $i$ and \ha;
-the average colours for sources in the catalogue are
+the average colours in the catalogue are
 $\overline{(r-i)}=1.06\pm0.12$ and $\overline{(r-\ha)}=0.44\pm0.03$.
@@
 We also plot the Poissonian uncertainties for comparison (solid red line).
-We find the mean residual and standard deviation to be $0.03\pm0.04$~mag
+We find the mean residual to equal $0.03\pm0.04$~mag
 across the magnitude ranges 13 to 18 ($r$)
-and 12 to 17 ($i$, \ha).
+and 12 to 17 ($i$, \ha),
+consistent with the accuracy of our calibration estimated above.
 Stars fainter than this range appear to be dominated by photon noise (red line),
@@
 is roughly 300\,000 objects per square degree,
-which is six times more than the density
+which is six times more than
 found near $\ell\simeq 180\degr$.
@@
 variations are also apparent on smaller scales.
-For example, we find a significant drop near the constellations 
+For example, we find a significant drop in source count near the constellations 
 of Aquila ($\ell\simeq40\degr$) and Cygnus ($\ell\simeq80\degr$ and $\ell\simeq90\degr$),
@@
 to be released or for variations in the depth across the included fields.
-For example, the dip in the density near $\ell\simeq140\degr$
+For example, the dip near $\ell\simeq140\degr$
 is an artificial feature caused by gaps
@@
 
-In a forthcoming paper, properly calibrated maps of stellar density
+In a forthcoming paper, properly-calibrated detailed maps of stellar density
 of the northern Galactic Plane will be presented (Farnhill
@@
 corrections based on the statistics of artificial source
-recovery.  Such maps are of interest as tests of detailed models of 
-our Galaxy.
+recovery.  Such maps are of interest as tests of Galactic models.
 
@@
     in the latitude range $-5\degr<b<+5\degr$,
-    such that one may obtain approximate source densities
+    such that one may obtain a rough guide to source density
     by dividing the counts by 10~deg$^2$.
@@
 blend-affected sources correlates with the local source 
-density.  In hard numbers, only 11\% of the sources are blended
+density.  In headline numbers, only 11\% of the sources are blended
 at $\ell>90\degr$, whereas 24\% are blended at $\ell<90\degr$.
@@
 where we present three sets of IPHAS colour/magnitude diagrams
-towards three distinct sight-lines
+towards three distinct sightlines
 located at Galactic longitudes 
@@
 Each of the diagrams reveals a well-defined locus,
-which helps to demonstrate the health of the catalogue
+which helps to further demonstrate the health of the catalogue
 and the global calibration
@@
 of empirical spectra
-transformed into the Vega-based IPHAS system by \cite{Drew2005}.
+synthesised into the Vega-based IPHAS system by \cite{Drew2005}.
 In the colour-magnitude diagrams we only show the reddening vector
@@
 we find a population dominated by lowly-reddened main sequence stars.
-This is consistent with the estimated total sight-line extinction 
+This is consistent with the estimated total sightline extinction 
 of $E(B-V)=0.49$ given by \cite{Schlegel1998},
-following the correction of \cite{Schlafly2011}.
+and applying the 14\% reduction recommended by \cite{Schlafly2011}.
 Looking in more detail we can see
 that the stellar locus is narrower for M-type dwarfs
-than for earlier types;
+than for earlier types:
 we do not observe M dwarfs experiencing 
@@
 of the corresponding phase of stellar evolution,
-would only account for a small proportion of the more nearly \emph{volume-limited} sample seen in the Anticentre direction.
-
-In contrast, closer towards the galactic centre,
-we encounter samples which are \emph{magnitude-limited} instead.
+would only account for a small proportion of this more nearly \emph{volume-limited} sample seen in the Anticentre direction.
+
+In contrast, lines of sight passing into the first Galactic quadrant
+yield samples that are more commonly \emph{magnitude-limited} instead.
 For example, at $\ell=45\degr$ (Fig.~\ref{fig:l45}),
@@
 
-Finally, in one of our earliest sight-lines at $\ell=30\degr$,
+Finally, in one of our lowest-longitude sightlines at $\ell=30\degr$,
 we find a very high number of extremely reddened giants
@@
 exhibit a relatively narrow range of reddenings
-as they lie beyond most or essentially all of the Galactic dust.
+as they lie beyond most of the Galactic dust column.
 At $(l,b)=(30\degr, 0\degr)$ this is not the case:
@@
 
-These are just qualitative vignettes of the information obtainable
+These are just descriptive vignettes of the information obtainable
 from IPHAS colour-colour and colour-magnitude plots.
@@
 A 3-D extinction map based on the DR2 catalogue
-is being released in a separate paper (Sale et al., in preparation).
+is being released in a separate paper \citep{Sale2014}.
+%Sale2014 needs to be 'Sale S. E., et al., 2014, MNRAS, submitted' in the bibliography
 
@@
 across the Galactic Plane.
-\ha\ in emission is a well-known tracer
-for stars in the short-lived pre- or
-post-main sequence stages of their evolution,
-and hence IPHAS allows larger, deeper
-and more statistically robust samples
-of such rare objects to be established.
+%This is repetition
+%\ha\ in emission is a well-known tracer
+%for stars in the short-lived pre- or
+%post-main sequence stages of their evolution,
+%and hence IPHAS allows larger, deeper
+%and more statistically robust samples
+%of such rare objects to be established.
 The survey-wide identification and analysis 
@@
 using model-based colour criteria
-rather than a statistical procedure.
-In what follows we demonstrate this ability 
+rather than an adaptive statistical procedure.
+In what follows we demonstrate this new capability 
 by selecting candidate emission-line objects
@@
 recently noted that the warm dust
-that surrounds Sh~2-82 
+surrounding Sh~2-82 
 appears to contain infrared-bright
@@
 and have been flagged as \emph{reliable}
-in our catalogue.
+in IPHAS DR2.
 The diagram also shows the unreddened main sequence (solid line)
@@
 are marked by red triangles in the colour-colour diagram,
-and yellow triangles in the image mosaic (Fig.~\ref{fig:mosaic_iphas}).
+and by yellow triangles in the image mosaic (Fig.~\ref{fig:mosaic_iphas}).
 Their details are listed in Table~\ref{tbl:emitters}.
@@
 we discovered towards Sh 2-82 as well.
-Two of our candidates, \#1 and \#4 in Table~\ref{tbl:emitters},
+Two of our candidates, \#1 and \#4 in Tab.~\ref{tbl:emitters},
 have previously been identified
@@
 to find intrinsically red objects,
-which is consistent with the presence of a circumstellar disk.
+with SEDs consistent with the presence of a circumstellar disk.
 Although the other four candidate emitters in our sample
@@
 are larger than the
-aperture diameters used in this work.
+aperture diameters specified in this work.
 To enable the analysis of diffuse sources,
@@
 gain variations, exposure time, and the re-calibration shift.
-Users will only need to apply an aperture correction, if necessary.
+As these images still include moonlight and other sources
+of non-astronomical background, they can only support flux measurements
+that include a suitably-chosen local background subtraction.
 
@@
 which is the flux obtained by folding the 
-Vega model SED from Kurucz with the filter transmission curve 
-(uncorrected for the atmosphere and detector quantum efficiency,
-which would otherwise scale down the narrowband flux by 0.707).
+CALSPEC SED with the filter transmission curve only
+(the correction for atmosphere and detector quantum efficiency,
+otherwise scales down the narrowband flux by 0.707).
 This implies that the in-band flux corresponding to
@@
 when the \ha\ magnitude for Vega is set by convention to 0.03 \citep{Fukugita1996}.
+These flux estimates are consistent with the most recent
+version of Vega's SED presented by \citet{Bohlin2014}.
 
@@
 
-A new catalogue of 219 million sources has been
+A new catalogue 
+%of 219 million sources 
+has been
 derived from the INT/WFC Photometric \ha\ Survey
 of the Northern Galactic Plane.
-It is the first catalogue to offer comprehensive CCD photometry
+It is the first to offer comprehensive CCD photometry
 of point sources across the Northern Galactic Plane at visible wavelengths,
@@
 and selecting emission-line objects.  More comprehensive applications of IPHAS
-can be found in the works of Sale et al (submitted), which applies DR2 to the problem
+can be found in the works of \citet{Sale2014}, which applies DR2 to the problem
 of 3D extinction mapping, and of Sabin et al (in preparation), where the results of a search
@@
 replacing data not meeting the IPHAS minimum quality requirements are continuing.  We will also 
-examine options to further improve the global calibration, perhaps tightening the accuracy to
+examine options to further improve the global calibration, with a view to tightening the accuracy to
 better than 2\%.  For example, we have in mind investigating the use of the PanSTARRS photometric 

### 2014-04-16 paper/iphas-dr2.tex

@@
 \begin{figure*}
-        \includegraphics[width=1\linewidth]{figures/footprint/footprint_large.jpg}
+        \includegraphics[width=1\linewidth]{figures/footprint/footprint_small.png}
         \caption{Survey area showing the footprints
@@
 across the detectors.
-% Stating the obvious ...a new insertion? which is a proxy for the image quality
-% arising from the combined effects of atmospheric and dome seeing.
 This target is currently attained across 86 per cent of the footprint.
@@
 and 10 per cent is better than 0.8~arcsec.
-% slipped in earlier (the WFC pixel scale is 0.33~arcsec/px).
 In \S\ref{sec:catalogue} we will explain
@@
              we find that the detected flux for Vega in \ha\ is 3.14~magnitudes less
-             than that received in $r$, 
+             than that received in $r$.
           }
@@
 ($160\degr < \ell < 200\degr$).
-Despite the fact that all the fields included in the diagram
-have been quality-approved,
-the main locus of stars is poorly defined
-due to the presence of calibration errors,
-which regularly exceed 0.1~mag.
+The main locus of stars is poorly defined
+in the diagram due to the presence of
+incorrectly calibrated fields.
+The calibration errors can exceed 0.1~mag
+and need to be corrected for 
+during a global re-calibration procedure.
 For comparison, Fig.~\ref{fig:ccd_after}
@@
 \end{subfigure}
-\caption{IPHAS $(r-\ha,\ r-i)$ colour-colour diagram for an area of 400~deg$^2$, plotted before (panel a) and after (panel b)
+\caption{IPHAS $(r-\ha,\ r-i)$ colour-colour diagram
+         covering an area of 400~deg$^2$,
+         shown before (panel a) and after (panel b)
          re-calibration.
-Both figures were created by combining the stars detected
-across the entire set of 2\,801 quality-approved fields which are located
-towards the Galactic anti-centre  ($160\degr < \ell < 200\degr$).
+         Both figures were created by combining the stars
+         detected across all 2,801 quality-approved fields
+         which are located towards the Galactic anti-centre
+         ($160\degr<\ell<200\degr$).
          The diagrams are plotted as 2D-histograms
@@
          while bins containing more than 400 sources are bright yellow.
-We have included all stars which are brighter than $r<18$
-and are classified as `very reliable'
-(these criteria will be explained in \S\ref{sec:qualitycriteria}).}
+         The diagrams include all stars
+         brighter than $r<18$ which were
+         classified by the pipeline as being `very reliable'
+         (to be explained in
+         \S\ref{sec:qualitycriteria}).
+         The objects seen to fall above the locus of stars
+         \emph{after} re-calibration 
+         are likely to be genuine \ha\ emission-line objects.}
 \end{figure}
 
-The calibration variations need to be corrected for 
-during a global re-calibration procedure.
-Notable past examples include the global re-calibration 
-of 2MASS \citep[][]{Nikolaev2000},
-SDSS \citep[][]{Padmanabhan2008}
+Notable past examples of surveys which required
+global re-calibration include 
+2MASS \citep{Nikolaev2000},
+SDSS \citep{Padmanabhan2008}
 and the Panoramic Survey Telescope 
@@
 to be at the level of 3 per cent,
-%The private communication from Hended is no longer needed :-)
 which is significantly better 
 than the original nightly calibrations of IPHAS
-which are accurate to $\sim$10 per cent (Tab.~\ref{tbl:offsets_before}).
+which are only accurate to $\sim$10 per cent
+when compared to APASS (Table~\ref{tbl:offsets_before}).
 APASS achieves its uniform accuracy 
@@
 and IPHAS uses Vega-based magnitudes.
-%I wonder if the next bit is stating the too-obvious?
-%Have shortened it a bit.
 Separate transformations were derived for sightlines 
@@
 The offsets follow a near-Gaussian distribution
-centred around $0.014\pm0.104$~mag in $r$ and $0.007\pm0.108$~mag in $i$
-(Table~\ref{tbl:offsets_before}).
-A total of 4596 fields showed a systematic offset
-exceeding $\pm$0.03~mag in either $r$ or $i$.
+with mean and sigma $0.014\pm0.104$~mag in $r$
+and $0.007\pm0.108$~mag in $i$ (Table~\ref{tbl:offsets_before}).
+A total of 4,596 IPHAS fields showed a median offset
+exceeding $\pm$0.03~mag in either $r$ or $i$
+when compared to APASS.
 
@@
 
+\subsubsection{Adjusting fields not covered by APASS}
+
+At the time of writing, the APASS catalogue did not provide 
+sufficient coverage for 7,359 of our fields.
+Fortunately, these fields are located mainly 
+at low Galactic longitudes (cf. Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}),
+which were typically observed during the summer months
+when photometric conditions are more prevalent at the telescope.
+These remaining fields have nevertheless 
+been brought onto the same uniform scale 
+by employing an ubercalibration-style scheme,
+which minimises the magnitude offsets between stars
+located in the overlap regions between neighbouring fields.
+
+An algorithm for achieving this minimisation
+has previously been described by~\citet{Glazebrook1994}.
+In brief, there are two fundamental quantities to be
+minimised between each pair of overlapping exposures, 
+denoted by the indices $i$ and $j$. 
+First, the mean magnitude difference between stars in the overlap
+region $\Delta_{ij}=\langle m_i-m_j\rangle=-\Delta_{ji}$ is a local
+constraint. 
+Second, to ensure the solution does not stray far 
+from the existing calibration, 
+the difference in zeropoints 
+$\Delta\mathrm{ZP}_{ij}=-\Delta\mathrm{ZP}_{ji}$ 
+between each pair of exposures must also be minimised.
+
+Minimisation of these two quantities is a linear least squares problem 
+because the magnitude $m$ depends linearly on the ZP (Eqn.~\ref{eqn:mag}).
+Hence we can find the ZP shift to be applied to each field 
+by minimising the sum:
+\begin{equation}
+   S = \sum_{i=1}^N \sum_{j=1}^N w_{ij} \theta_{ij} (\Delta_{ij} + a_i - a_j)^2,
+   \label{eqn:chi2}
+\end{equation}
+where $i$ denotes an exposure, 
+$j$ an overlapping exposure, 
+$N$ the number of exposures,
+$a_i$ the ZP to solve for,
+and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$). 
+$w_{ij}$ are weights set to the inverse square of the uncertainty in $\Delta_{ij}$,
+and $\theta_{ij}$ is an overlap function 
+equal to either 1 if exposures $i$ and $j$ overlap or 0 otherwise. 
+Solving for $a_i$ is equivalent to solving $\partial
+S/\partial a_i=0$, which gives the matrix equation:
+\begin{equation}
+   \sum_{j=1}^N A_{ij} a_j = b_j,
+   \label{eqn:matrix}
+\end{equation}
+where 
+\begin{eqnarray}
+   A_{ij} &=& \delta_{ij} \sum_{k=1}^N w_{jk}\theta_{jk} - w_{ij} \theta_{ij},\label{eqn:aij}\\
+   b_i &=& \sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ji} = -\sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ij}.\label{eqn:bi}
+\end{eqnarray}
+
+We enforce a strong external constraint
+on the solution by keeping the zeropoint fixed 
+for the 
+% 6\,756 
+fields which have already been compared
+and calibrated against APASS.
+We hereafter refer to these fields as \emph{anchors}.
+It is asserted that the zeropoints $a_i$ of the anchor fields 
+are known and not solved for.  However they do appear in the vector 
+$b_j$ as constraints.  In addition to the APASS-based anchors, 
+we selected 3,273 additional anchor fields by hand
+to provide additional constraints in regions not covered by APASS.
+These extra anchors were deemed to have reliable zeropoints 
+based on 
+(i) the information contained in the observing logs,
+(ii) the stability of the standard star zeropoints during the night, and
+(iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
+which is located $\sim$500~m from the INT.
+
+We then solved Eqn.~\ref{eqn:matrix} for the $r$ and $i$ bands
+separately using the least-squares routine 
+in Python's {\sc scipy.sparse} module for sparse matrix algebra.
+This provided us with corrected zeropoints for the remaining fields,
+which were shifted on average by $+0.02\pm0.11$ in $r$ 
+and $+0.01\pm0.12$ in $i$ compared to their provisional calibration.
+
+We then turned to the global calibration of the \ha\ data.
+It is not possible to re-calibrate the narrowband 
+in the same way as the broadbands,
+because the APASS survey does not offer \ha\ photometry.
+We can reasonably assume, however,
+that the corrections required for $r$ and \ha\ are identical,
+much of the time, because the IPHAS data-taking pattern ensured 
+that a field's \ha\ and $r$-band exposures
+were taken consecutively,
+albeit separated by a $\sim$30~s read-out time.
+Hence, we have corrected the \ha\ zeropoints 
+by re-using the zeropoint adjustments that were derived for the $r$ band
+in the earlier steps.
+An exception was made for 3,101 fields
+for which our quality-control routines revealed
+strong zeropoint variations during the night,
+suggesting non-photometric conditions.
+In these cases, the \ha\ zeropoints
+were adjusted by solving Eqn.~\ref{eqn:matrix}
+rather than by simply applying Eqn.~\ref{eqn:zpha}.
+
+\subsection{Testing the calibration against SDSS}
+
+Having re-calibrated all fields to the expected APASS accuracy of 3 per cent,
+we turned to a different survey, SDSS Data Release 9 \citep{Ahn2012},
+to validate the results.
+SDSS DR9 includes several strips at low
+Galactic latitudes,
+providing data across 18 per cent of the fields in our data release.
+We cross-matched the IPHAS fields against the subset of
+objects marked as reliable stars in the SDSS catalogue\footnote{
+We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
+to obtain photometry from the SDSS {\sc photoprimary} table 
+with criteria {\sc type = star}, {\sc clean = 1} and {\sc score $>$ 0.7}.}
+in much the same way as for APASS,
+with the difference of selecting from the  fainter magnitude ranges of 
+$15<r_{\rm SDSS}<18.0$ and $14.5<i_{\rm SDSS}<17.5$.
+This provided us with a set of 1.2~million cross-matched stars.
+
+Colour transformations were again obtained using a sigma-clipped linear least squares fit:
+\begin{eqnarray}
+r_{\rm IPHAS} = r_{\rm SDSS} - 0.093 - 0.044(r-i)_{\rm SDSS} \label{eqn:sdss_r} \\
+i_{\rm IPHAS} = i_{\rm SDSS} - 0.318 - 0.095(r-i)_{\rm SDSS}. \label{eqn:sdss_i}
+\end{eqnarray}
+The rms residuals of these transformations are 0.045 and 0.073, respectively.
+The equations are similar to the ones
+previously determined for APASS,
+although the colour terms are slightly larger.
+The throughput curve of the SDSS $i$-band filter 
+appears to be somewhat more sensitive at longer wavelengths
+than both the IPHAS and APASS filters.
+
+These global transformations were deemed adequate
+for the purpose of validating our calibration in a statistical sense.
+Separate equations were derived towards different sight-lines
+to investigate the effects of varying reddening regimes.
+The colour term was found 
+to show some variation towards weakly reddened areas,
+where different stellar populations are observed.
+The vast majority of red objects in the global sample
+are those in highly reddened areas, however,
+which agree well with the global transformations
+and dominate the statistical appraisal of our calibration.
+
+Having transformed SDSS magnitudes into the IPHAS system,
+we then computed the median magnitude offset for each IPHAS field
+which contained at least 30 objects with a cross-matched counterpart
+in SDSS.  This was the case for 2,602 fields.
+The median offsets for each of these fields
+are shown in Figs.~\ref{fig:sdss_r} and \ref{fig:sdss_i}.
+Importantly, the mean offset and standard deviation found 
+is $-0.001\pm0.029$~mag in $r$
+and $-0.002\pm0.032$~mag in $i$ (Table~\ref{tbl:offsets_after}).
+In comparison, offsets computed in the identical way
+\emph{before} carrying out the re-calibration showed means
+of $+0.016\pm0.088$~mag in $r$ and $+0.010\pm0.089$~mag in $i$ (Table~\ref{tbl:offsets_after}).
+We conclude that our re-calibration procedure has
+been successful in improving the
+uniformity of the calibration
+by a factor three (i.e. from $\sigma=0.088$ to $\sigma=0.029$ in $r$),
+and as such has achieved our aim of bringing the
+accuracy to the aimed level of 0.03~mag.
+
 \begin{figure*}
@@
 \end{figure*}
-
-\subsubsection{Adjusting fields not covered by APASS}
-
-At the time of writing, the APASS catalogue did not provide 
-sufficient coverage for 7359 of our fields.
-Fortunately, these fields are located mainly 
-at low Galactic longitudes (cf. Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}),
-which were typically observed during the summer months
-when photometric conditions are more prevalent at the telescope.
-These remaining fields have nevertheless 
-been brought onto the same uniform scale 
-by employing an ubercalibration-style scheme,
-which minimises the magnitude offsets between stars
-located in the overlap regions between neighbouring fields.
-
-An algorithm for achieving this minimisation
-has previously been described by~\citet{Glazebrook1994}.
-In brief, there are two fundamental quantities to be
-minimised between each pair of overlapping exposures, 
-denoted by the indices $i$ and $j$. 
-Firstly, the mean magnitude difference between stars in the overlap
-region $\Delta_{ij}=\langle m_i-m_j\rangle=-\Delta_{ji}$ is a local
-constraint. 
-Secondly, to ensure the solution does not stray far 
-from the existing calibration, 
-the difference in zeropoints 
-$\Delta\mathrm{ZP}_{ij}=-\Delta\mathrm{ZP}_{ji}$ 
-between each pair of exposures must also be minimised.
-
-Minimisation of these two quantities is a linear least squares problem 
-because the magnitude $m$ depends linearly on the ZP (Eqn.~\ref{eqn:mag}).
-Hence we can find the ZP shift to be applied to each field 
-by minimising the sum:
-\begin{equation}
-   S = \sum_{i=1}^N \sum_{j=1}^N w_{ij} \theta_{ij} (\Delta_{ij} + a_i - a_j)^2,
-   \label{eqn:chi2}
-\end{equation}
-where $i$ denotes an exposure, 
-$j$ an overlapping exposure, 
-$N$ the number of exposures,
-$a_i$ the ZP to solve for 
-and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$). 
-$w_{ij}$ are weights set to the inverse square of the uncertainty in $\Delta_{ij}$,
-and $\theta_{ij}$ is an overlap function 
-equal to either 1 if exposures $i$ and $j$ overlap or 0 otherwise. 
-Solving for $a_i$ is equivalent to solving $\partial
-S/\partial a_i=0$, which gives the matrix equation:
-\begin{equation}
-   \sum_{j=1}^N A_{ij} a_j = b_j,
-   \label{eqn:matrix}
-\end{equation}
-where 
-\begin{eqnarray}
-   A_{ij} &=& \delta_{ij} \sum_{k=1}^N w_{jk}\theta_{jk} - w_{ij} \theta_{ij},\label{eqn:aij}\\
-   b_i &=& \sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ji} = -\sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ij}.\label{eqn:bi}
-\end{eqnarray}
-
-We enforce a strong external constraint
-on the solution by keeping the zeropoint fixed 
-for the 
-% 6\,756 
-fields which have already been compared
-and calibrated against APASS.
-We hereafter refer to these fields as \emph{anchors}.
-It is asserted that the zeropoints $a_i$ of the anchor fields 
-are known and not solved for.  However they do appear in the vector 
-$b_j$ as constraints.  In addition to the APASS-based anchors, 
-we selected 3\,273 additional anchor fields by hand
-to provide additional constraints in regions not covered by APASS.
-These extra anchors were deemed to have reliable zeropoints 
-based on 
-(i) the information contained in the observing logs,
-(ii) the stability of the standard star zeropoints during the night, and
-(iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
-which is located $\sim$500~m from the INT.
-
-We then solved Eqn.~\ref{eqn:matrix} for the $r$ and $i$ bands
-separately using the least-squares routine 
-in Python's {\sc scipy.sparse} module for sparse matrix algebra.
-This provided us with corrected zeropoints for the remaining fields,
-which were shifted on average by $+0.02\pm0.11$ in $r$ 
-and $+0.01\pm0.12$ in $i$ compared to their provisional calibration.
-The outcome is validated in the next section.
-
-We then turned to the global calibration of the \ha\ data.
-It is not possible to re-calibrate the narrowband 
-in the same way as the broadbands,
-because the APASS survey does not offer \ha\ photometry.
-We can reasonably assume however,
-that the corrections required for $r$ and \ha\ are identical,
-much of the time, because the IPHAS data-taking pattern ensured 
-that a field's \ha\ and $r$-band exposures
-were taken consecutively,
-albeit separated by the $\sim$30~s read-out time.
-Hence, we have corrected the \ha\ zeropoints 
-by re-using the zeropoint adjustments that were derived for the $r$ band
-in the earlier steps.
-An exception was made for a small subset of fields
-% 3\,101 fields
-for which our quality-control routines revealed
-strong zeropoint variations during the night,
-suggesting non-photometric conditions.
-In these cases, the \ha\ zeropoints
-were adjusted by solving Eqn.~\ref{eqn:matrix}
-rather than by simply applying Eqn.~\ref{eqn:zpha}.
-
-\subsection{Testing the calibration against SDSS}
-
-Having re-calibrated all fields to the expected APASS accuracy of 3 per cent,
-we turned to a different survey, SDSS Data Release 9 \citep{Ahn2012},
-to validate the results.
-SDSS DR9 includes several strips at low
-Galactic latitudes,
-providing data across 18 per cent of the fields in our data release.
-We cross-matched the IPHAS fields against the subset of
-objects marked as reliable stars in the SDSS catalogue\footnote{
-We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
-to obtain photometry from the SDSS {\sc photoprimary} table 
-with criteria {\sc type = star}, {\sc clean = 1} and {\sc score $>$ 0.7}.}
-cross-matching in much the same way as for APASS,
-with the difference of selecting from the  fainter magnitude ranges of 
-$15<r_{\rm SDSS}<18.0$ and $14.5<i_{\rm SDSS}<17.5$.
-This provided us with a set of 1.2~million cross-matched stars.
-
-Colour transformations were again obtained using a sigma-clipped linear least squares fit:
-\begin{eqnarray}
-r_{\rm IPHAS} = r_{\rm SDSS} - 0.093 - 0.044(r-i)_{\rm SDSS} \label{eqn:sdss_r} \\
-i_{\rm IPHAS} = i_{\rm SDSS} - 0.318 - 0.095(r-i)_{\rm SDSS}. \label{eqn:sdss_i}
-\end{eqnarray}
-The rms residuals of these transformations are 0.045 and 0.073, respectively.
-The equations are similar to the ones
-previously determined for APASS,
-although the colour terms are slightly larger.
-The throughput curve of the SDSS $i$-band filter 
-appears to be somewhat more sensitive at longer wavelengths
-than both the IPHAS and APASS filters.
-
-These global transformations were deemed adequate
-for the purpose of validating our calibration in a statistical sense.
-Separate equations were derived towards different sight-lines
-to investigate the effects of varying reddening regimes.
-The colour term was found 
-to show some variation towards weakly reddened areas,
-where different stellar populations are observed.
-The vast majority of red objects in the global sample
-are those in highly reddened areas, however,
-which agree well with the global transformations
-and dominate the statistical appraisal of our calibration.
-
-Having transformed SDSS magnitudes into the IPHAS system,
-we then computed the median magnitude offset for each IPHAS field
-which contained at least 30 objects with a cross-matched counterpart
-in SDSS.  This was the case for 2\,602 fields.
-The median offsets for each of these fields
-are shown in Figs.~\ref{fig:sdss_r} and \ref{fig:sdss_i}.
-Importantly, the mean offset and standard deviation found 
-is $-0.001\pm0.029$~mag in $r$
-and $-0.002\pm0.032$~mag in $i$ (Table~\ref{tbl:offsets_after}).
-In comparison, offsets computed in the identical way
-\emph{before} carrying out our re-calibration showed means
-of $+0.016\pm0.088$~mag in $r$ and $+0.010\pm0.089$~mag in $i$ (Table~\ref{tbl:offsets_after}).
-We conclude that our re-calibration procedure has
-been successful in improving the
-uniformity of the calibration by a factor three
-and has achieved our aim of bringing the
-accuracy down to the level of 0.03~mag (rms).
 
@@
 which are in use at the
-WFCAM Science Archive \citep[WSA;][]{Hambly2008}
-and the VISTA Science Archive \citep[VSA;][]{Cross2012}.
+Wide Field Camera (WFCAM) Science Archive \citep[WSA;][]{Hambly2008}
+and the 
+Visible and Infrared Survey Telescope for Astronomy (VISTA)
+Science Archive \citep[VSA;][]{Cross2012}.
 These archives curate the high-resolution near-infrared photometry from both
-the UKIRT Infrared Deep Sky Survey \citep[UKIDSS;][]{Lawrence2007}
+the United Kingdom Infrared Telescope (UKIRT)
+Infrared Deep Sky Survey \citep[UKIDSS;][]{Lawrence2007}
 and the 
@@
 
-Photometry is provided based on the 2.3-arcsec diameter circular aperture
+Photometry is provided based on the 2\farcs 3-arcsec diameter circular aperture
 by default (columns \emph{r}, \emph{i}, \emph{ha}).
@@
 Each of these magnitude measurements have been
-corrected for the flux lost outside of the applied aperture,
+corrected for the flux lost outside of the respective apertures,
 using a correction term which is inferred from the
@@
 aperture size is also used by the CASU pipeline to provide
-a discrete star/galaxy/noise\footnote{For consistency with the terminology used in the CASU pipeline and the WSA archive, extended objects are classified as `galaxies' in the catalogue.
-This class may however include other types of resolved objects.} classification flag
+a discrete star/galaxy\footnote{For consistency with the terminology that is used in the CASU pipeline and the WSA/VSA archives, extended objects are classified as `galaxies'.
+This class will flag any type of resolved object, however.}/noise classification flag classification flag
 (\emph{rClass}, \emph{iClass}, \emph{haClass})
@@
 thanks to the IPHAS field pair strategy. 
-%In turn, this has meant that it has not proved necessary 
 Hence there has been no need to include these internal warning flags
@@
 further if necessary
-(columns \emph{iXi}, \emph{iEta}, \emph{haXi}, \emph{haEta}), or simply to examine
-light centre differences.
-%I think of Dimitri Purbaix here.
+(columns \emph{iXi}, \emph{iEta}, \emph{haXi}, \emph{haEta}),
+or simply to examine light centre differences.
 We note that UKIDSS/GPS adopted 
 the same maximum matching distance of 1 arcsec
-for similar reasons \citep{Hambly2008},
-and warn that applications which require absolute
-astrometry with sub-arcsecond accuracy must
-use these catalogues with caution.
-%GEERT.  Is the above too extreme a warning?  Is the r band the 'master', to which the given source 
-%position refers, when available, and hence is the natural choice for astrometric purposes?
-%The caution applies if you mis-associate the position with e.g. an i source, without
-%importing the offsets.  Or do you really believe no position is better than 1 arcsec because
-%we don't know which band it refers to.
-
-The resulting band-merged catalogues were found
-to be reliable for the vast majority of fields.
-We do warn that source blending and confusion is unavoidable
-for faint objects in the most crowded regions of the Galactic Plane;
-in \S\ref{sec:discussion} we will show
-that 19 per cent of the sources in our catalogue
-are flagged as blended objects (column \emph{deblend}),
-and their band-merged data should be treated with care
-because they may have fallen victim to source confusion
-during the band-merging step.
+for similar reasons \citep{Hambly2008}.
+
+The resulting band-merged catalogues were inspected
+as part of our quality control procedures
+and were found to be reliable.
 
@@
 is markedly better than for $r$ and \ha,
-which is most likely explained by the fact that $i$ is least
+which is likely explained by the fact that $i$ is least
 affected by interstellar extinction, and so the survey can occasionally pick up
@@
 Moreover, objects which are intrinsically very red may also be picked
-up in $i$ alone, while faint objects with extremely strong Balmer emission
+up in $i$ alone, while faint objects with very strong Balmer emission
 may appear only in \ha.
@@
 up to a quarter of the objects.
-The \emph{veryReliable} flag should hence
-only be used in applications in which 
-completeness is not a concern.
 
@@
 which we found to suit many science applications,
-but certainly not all. Users of the data are encouraged to revise 
-the quality criteria according to their needs.
+but users are encouraged 
+to revise the quality criteria according to their needs.
 
@@
              columns \emph{rErr}, \emph{iErr} and \emph{haErr}
-             in the catalogue 
+             in the catalogue,
              and the error-bars show the standard deviations.
@@
 We find the typical uncertainty to reach 0.1~mag near $r=$20.5 
-and $i$/\ha$=19.5$.
+and $i$,\ha$=19.5$.
 We note that the fainter depth in $r$ is compensated
@@
 We also plot the Poissonian uncertainties for comparison (solid red line).
-We find the mean residual to equal $0.03\pm0.04$~mag
+We find the mean residual and standard deviation to be $0.03\pm0.04$~mag
 across the magnitude ranges 13 to 18 ($r$)
 and 12 to 17 ($i$, \ha),
-consistent with the accuracy of our calibration estimated above.
+which is consistent with the accuracy of the calibration.
 Stars fainter than this range appear to be dominated by photon noise (red line),
@@
 For example, the average source density near $\ell\simeq 30\degr$
-is roughly 300\,000 objects per square degree,
-which is six times more than
+is roughly 300,000 objects per square degree,
+which is six times more than the density
 found near $\ell\simeq 180\degr$.
@@
 variations are also apparent on smaller scales.
-For example, we find a significant drop in source count near the constellations 
+For example, we find a significant drop near the constellations 
 of Aquila ($\ell\simeq40\degr$) and Cygnus ($\ell\simeq80\degr$ and $\ell\simeq90\degr$),
@@
 is being released in a separate paper \citep{Sale2014}.
-%Sale2014 needs to be 'Sale S. E., et al., 2014, MNRAS, submitted' in the bibliography
 
@@
 across the Galactic Plane.
-%This is repetition
-%\ha\ in emission is a well-known tracer
-%for stars in the short-lived pre- or
-%post-main sequence stages of their evolution,
-%and hence IPHAS allows larger, deeper
-%and more statistically robust samples
-%of such rare objects to be established.
 The survey-wide identification and analysis 
@@
 
-A new catalogue 
-%of 219 million sources 
-has been
+A new catalogue has been
 derived from the INT/WFC Photometric \ha\ Survey
@@
 
-The current plan is to work toward one further major IPHAS source catalogue, in which 
-the remaining gaps in sky coverage will have been eliminated -- observations aimed at 
-replacing data not meeting the IPHAS minimum quality requirements are continuing.  We will also 
-examine options to further improve the global calibration, with a view to tightening the accuracy to
-better than 2\%.  For example, we have in mind investigating the use of the PanSTARRS photometric 
-ladder (Magnier et al 2013) as a reference set, when it becomes available and we will 
+The current plan is to work toward
+one further major IPHAS source catalogue,
+in which the remaining gaps in sky coverage
+will have been eliminated
+-- observations aimed at replacing data
+not meeting the quality requirements are continuing.
+We will also examine options to further improve
+the global calibration,
+perhaps tightening the accuracy to better than 2 per cent.
+For example, we have in mind investigating
+the use of the PanSTARRS photometric ladder \citep{Magnier2013}
+as a reference set,
+when it becomes available for the Galactic Plane, and we will 
 explore improving source recovery in 
-the most dense fields via the implementation of PSF fitting in place of aperture photometry.  
-Finally, the next catalogue will detail all the secondary detections to aid time-domain 
-studies.
+the most dense fields via the implementation of PSF fitting
+in place of aperture photometry.  
+Finally, the next catalogue will detail all the secondary detections
+to aid time-domain studies.
 

### 2014-04-17 paper/iphas-dr2.tex
SUBMITTED VERSION!

@@
 
-\voffset=-0.5in
-
-\graphicspath{ {/home/gb/msc/Dropbox/papers/dr2/} }
+%\voffset=-0.5in
 
@@
 A. A. Henden$^9$,
-M. J. Barlow$^{10}$,
-P. J. Carter$^{11}$,
-R. L. M Corradi$^{6,7}$, \newauthor
-J. J. Drake$^{12}$, 
-J. Eisl\"offel$^{13}$, 
-J. Fabregat$^{14}$, 
-B. T. G\"ansicke$^{11}$,
-N. P. Gentile Fusillo$^{11}$, \newauthor
-S. Greiss$^{11}$, 
-A. S. Hales$^{15}$, 
+A. Aungwerojwit$^{10}$,
+M. J. Barlow$^{11}$,
+P. J. Carter$^{12}$, \newauthor
+R. L. M. Corradi$^{6,7}$, 
+J. J. Drake$^{13}$, 
+J. Eisl\"offel$^{14}$, 
+J. Fabregat$^{15}$, 
+B. T. G\"ansicke$^{12}$, \newauthor
+N. P. Gentile Fusillo$^{12}$, 
+S. Greiss$^{12}$, 
+A. S. Hales$^{16}$, 
 S. Hodgkin$^2$,
-L. Huckvale$^{16}$,
-J. Irwin$^{12}$, 
-C. Knigge$^{17}$, \newauthor
+L. Huckvale$^{17}$, \newauthor
+J. Irwin$^{13}$,
+R. King$^{18}$,
+C. Knigge$^{19}$, 
 T. Kupfer$^{5}$,
-D. J. Lennon$^{18}$,
+E. Lagadec$^{20}$,
+D. J. Lennon$^{21}$, \newauthor
 J. R. Lewis$^{2}$,
 M. Mohr-Smith$^{1}$,
-R. A. H. Morris$^{19}$, \newauthor
-T. Naylor$^{20}$, 
-Q. A. Parker$^{21,22,23}$,
-S. Phillipps$^{19}$, 
-R. Raddi$^{11}$, 
+R. A. H. Morris$^{22}$, 
+T. Naylor$^{18}$, 
+Q. A. Parker$^{23,24,25}$,  \newauthor
+S. Phillipps$^{21}$, 
+S. Pyrzas$^{26}$
+R. Raddi$^{12}$,
+G. H. A. Roelofs$^{13}$, 
 P. Rodr\'\i guez-Gil$^{6,7}$, \newauthor
-L. Sabin$^{24}$, 
-S. Scaringi$^{25,26}$,
-D. Steeghs$^{11}$, 
-Y. C. Unruh$^{27}$, 
-J. van Roestel$^5$, \newauthor
-K. Viironen$^{28}$, 
-J. S. Vink$^{29}$, 
+L. Sabin$^{27}$, 
+S. Scaringi$^{28,29}$,
+D. Steeghs$^{12}$,
+J. Suso$^{14}$,
+R. Tata$^{6,7}$,
+Y. C. Unruh$^{30}$, \newauthor
+J. van Roestel$^5$, 
+K. Viironen$^{31}$, 
+J. S. Vink$^{32}$, 
 N. A. Walton$^{2}$, 
-N. J. Wright$^{1}$,
-A. A. Zijlstra$^{16}$.
+N. J. Wright$^{1}$, \newauthor
+A. A. Zijlstra$^{17}$.
 \\
@@
 $^{2}$Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge, CB3 OHA, U.K.\\
-$^{3}$IGAM, Institute of Physics, University of Graz, Universit\"atsplatz 5, Graz, Austria.\\
+$^{3}$IGAM, Institute of Physics, University of Graz, Universit\"atsplatz 5, 8010 Graz, Austria.\\
 $^{4}$South African Astronomical Observatory, P.O. Box 9, Observatory, 7935 Cape Town, South Africa.\\
@@
 $^{9}$AAVSO, 49 Bay State Road, Cambridge, MA 02138, USA.\\
-$^{10}$University College London, Department of Physics \& Astronomy, 
+$^{10}$Department of Physics, Faculty of Science, Naresuan University, Phitsanulok 65000, Thailand.\\
+$^{11}$University College London, Department of Physics \& Astronomy, 
 Gower Street, London WC1E 6BT, U.K.\\
-$^{11}$Department of Physics, University of Warwick, Gibbet Hill Road, Coventry, CV4 7AL, U.K.\\
-$^{12}$Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, 
+$^{12}$Department of Physics, University of Warwick, Gibbet Hill Road, Coventry, CV4 7AL, U.K.\\
+$^{13}$Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, 
 Cambridge, MA 02138, U.S.A. \\
-$^{13}$Th\"uringer Landessternwarte, Sternwarte 5, D-07778, Tautenburg, Germany. \\
-$^{14}$Observatorio Astron\'omico, Universidad de Valencia,
-Catedr\'atico Jos\'e Beltr\'an 2, 46980 Paterna, Spain.\\
-$^{15}$Joint ALMA Observatory, Alonso de C\'ordova 3107, Vitacura 763-0355, Santiago, Chile. \\
-%$^{16}$Jodrell Bank Centre for Astrophysics, School of Physics \& Astronomy, University of Manchester, Manchester M13 9PL, U.K.\\
-%$^{17}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
-%$^{18}$European Space Astronomy Centre (ESAC), Villafranca del Castillo, Villanueva de la Canada, E-28692 Madrid, Spain.\\
-%$^{19}$School of Physics, Bristol University, Tyndall Avenue, Bristol, BS8 1TL, U.K.\\
-%$^{20}$School of Physics, University of Exeter, Stocker Road, Exeter, EX4 4QL, U.K.\\
-%$^{21}$Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia.\\
-%$^{22}$Australian Astronomical Observatory, PO Box 915, North Ryde, NSW 1670, Australia.\\
-%$^{23}$Research Centre for Astronomy, Astrophysics and Astrophotonics, Macquarie University, Sydney, NSW 2109 Australia.\\
-%$^{24}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
-%$^{25}$Instituut voor Sterrenkunde, K.U. Leuven, Celestijnenlaan 200D, B-3001 Leuven, Belgium.\\
-%$^{26}$Max-Planck-Institut f\"ur extraterrestrische Physik (MPE), Giessenbachstrasse 1, D-85748, Garching bei M\"unchen, Germany. \\
-%$^{27}$Department of Physics, Blackett Laboratory, Imperial College London, Prince Consort Road, London, SW7 2AZ, U.K.\\
-%$^{28}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
-%$^{29}$Armagh Observatory, College Hill, Armagh, Northern Ireland, BT61 9DG, U.K.\\
-$^{16-29}$ (affiliations 16 through 29 are given in the source file)
+%$^{14}$Th\"uringer Landessternwarte, Sternwarte 5, D-07778, Tautenburg, Germany. \\
+%$^{15}$Observatorio Astron\'omico, Universidad de Valencia, Catedr\'atico Jos\'e Beltr\'an 2, 46980 Paterna, Spain.\\
+%$^{16}$Joint ALMA Observatory, Alonso de C\'ordova 3107, Vitacura 763-0355, Santiago, Chile. \\
+%$^{17}$Jodrell Bank Centre for Astrophysics, School of Physics \& Astronomy, University of Manchester, Manchester M13 9PL, U.K.\\
+%$^{18}$School of Physics, University of Exeter, Stocker Road, Exeter, EX4 4QL, U.K.\\
+%$^{19}$School of Physics \& Astronomy, University of Southampton, Southampton, SO17 1BJ, U.K.\\
+%$^{20}$Cornell University, Astronomy Department, Ithaca, NY 14853-6801, USA.\\
+%$^{21}$European Space Astronomy Centre (ESAC), Villafranca del Castillo, Villanueva de la Canada, E-28692 Madrid, Spain.\\
+%$^{22}$School of Physics, Bristol University, Tyndall Avenue, Bristol, BS8 1TL, U.K.\\
+%$^{23}$Department of Physics and Astronomy, Macquarie University, Sydney, NSW 2109, Australia.\\
+%$^{24}$Australian Astronomical Observatory, PO Box 915, North Ryde, NSW 1670, Australia.\\
+%$^{25}$Research Centre for Astronomy, Astrophysics and Astrophotonics, Macquarie University, Sydney, NSW 2109 Australia.\\
+%$^{26}$Instituto de Astronom\'ia, Universidad Cat\'olica del Norte, Avenida Angamos 0610, Antofagasta, Chile.\\
+%$^{27}$Instituto de Astonom\'ia y Meteorolog\'ia, Departamento de F\'isica, CUCEI, Universidad de Guadalajara, Mexico.\\
+%$^{28}$Instituut voor Sterrenkunde, K.U. Leuven, Celestijnenlaan 200D, B-3001 Leuven, Belgium.\\
+%$^{29}$Max-Planck-Institut f\"ur extraterrestrische Physik (MPE), Giessenbachstrasse 1, D-85748, Garching bei M\"unchen, Germany. \\
+%$^{30}$Department of Physics, Blackett Laboratory, Imperial College London, Prince Consort Road, London, SW7 2AZ, U.K.\\
+%$^{31}$Centro de Estudios de F\'isica del Cosmos de Arag\'on, Plaza San Juan 1, Planta 2, Teruel, 44001, Spain. \\
+%$^{32}$Armagh Observatory, College Hill, Armagh, Northern Ireland, BT61 9DG, U.K.\\
+$^{14-32}$ (note to editor: affiliations 14 through 32 are given in the source file)
 }
@@
 
-\begin{abstract} % BACKGROUND, OBJECTIVE, METHODS, RESULTS, CONCLUSIONS
+\begin{abstract}
 The INT/WFC Photometric H$\alpha$ Survey 
@@
 (i) characterise stellar populations and extinction regimes
-towards different Galactic sight-lines
+towards different Galactic sightlines
 and (ii) select \ha\ emission-line objects.
@@
 nearly always within 5 minutes of each other.
-%An exception was made for 9 fields where the three exposures 
-%could not be obtained within the same night
-%but for which the time gap between exposures did not exceed 48 hours.
 We note that the source catalogue details the exact epoch
@@
          \S\ref{sec:qualitycriteria}).
-         The objects seen to fall above the locus of stars
+         The objects which are seen to fall above the locus of stars
          \emph{after} re-calibration 
@@
 from revisiting regions in their footprint to 
-carry out a so-called \emph{ubercalibration}\footnote{`Ubercalibration'
-refers to the name of the code used to re-calibrate SDSS photometry
-using repeat observations. 
+carry out a so-called \emph{ubercalibration}\footnote{`ubercalibration'
+refers to the name of the code used to re-calibrate SDSS photometry. 
 It is an anglicised version of the German word `\"uberkalibration',
-which was reportedly chosen because the initial authors both
-had German-sounding names \citep{Finkbeiner2010}.} procedure,
-in which multiple measurements of stars at different epochs
+which was reportedly chosen because the initial authors, Schlegel and Finkbeiner, both
+have German-sounding names \citep{Finkbeiner2010}.} procedure,
+in which repeat measurements of stars in different nights
 are used to fit the calibration parameters
@@
 We have found the information contained
-in these narrow overlap regions to be insufficient
+in our narrow overlap regions to be insufficient
 to constrain the calibration parameters.
@@
 Although these systematics are reasonably small within a single field,
-they can accumulate during the calibration process, causing artificial zeropoint gradients 
-across the survey unless controlled by other external constraints.
+they can accumulate during a re-calibration process,
+causing artificial zeropoint gradients across the survey
+unless controlled by other external constraints.
 
@@
 \begin{figure*}
-    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_rshift.pdf} 
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_r.pdf} 
+    \includegraphics[width=\textwidth]{figures/calibration/APASS-IPHAS-DR2_rshift.pdf} 
+    \includegraphics[width=\textwidth]{figures/calibration/colourbar_apass_r.pdf} 
     \caption{Median magnitude offset in the $r$ band between IPHAS and APASS,
@@
     \vspace{1cm}
-    \includegraphics[width=\textwidth]{plots/calibration/APASS-IPHAS-DR2_ishift.pdf} 
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_apass_i.pdf} 
+    \includegraphics[width=\textwidth]{figures/calibration/APASS-IPHAS-DR2_ishift.pdf} 
+    \includegraphics[width=\textwidth]{figures/calibration/colourbar_apass_i.pdf} 
     \caption{Same as Fig.~\ref{fig:apass_r} for the $i$-band.}
@@
 on the solution by keeping the zeropoint fixed 
-for the 
-% 6\,756 
-fields which have already been compared
+for the fields which have already been compared
 and calibrated against APASS.
@@
 for the purpose of validating our calibration in a statistical sense.
-Separate equations were derived towards different sight-lines
+Separate equations were derived towards different sightlines
 to investigate the effects of varying reddening regimes.
@@
 \begin{figure*}
-    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_rshift.pdf}
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_r.pdf} 
+    \includegraphics[width=\textwidth]{figures/calibration/SDSS-IPHAS_rshift.pdf}
+    \includegraphics[width=\textwidth]{figures/calibration/colourbar_sdss_r.pdf} 
     \caption{Median magnitude offset between IPHAS and SDSS in the $r$ band
@@
     \vspace{1cm}
-    \includegraphics[width=\textwidth]{plots/calibration/SDSS-IPHAS_ishift.pdf}
-    \includegraphics[width=\textwidth]{plots/calibration/colourbar_sdss_i.pdf} 
+    \includegraphics[width=\textwidth]{figures/calibration/SDSS-IPHAS_ishift.pdf}
+    \includegraphics[width=\textwidth]{figures/calibration/colourbar_sdss_i.pdf} 
     \caption{Same as Fig.~\ref{fig:sdss_r} for the $i$-band.}
@@
 will be discussed in \S\ref{sec:uncertainties}).
-%and we do not recommend reliance on it.
 \item The \emph{deblend} column is used to flag sources 
@@
         \end{tabular}
-        % riha: 158831809 / 144339044
-		% ri: 24558323 / 18878201
-		% iha: 3422905 / 2495655
-		% rha: 1606867 / 1036607
-		% i: 15463660 / 6612351
-		% r: 9263422 / 2478774
-		% ha: 5844538 / 727732
 	\end{center}
@@
 \end{tabular}
-
 \label{tab:veryreliable}
-
 \vspace{1cm}
@@
 and (iii) the object was not located near a very bright star.
-59 million sources (27\%) satisfy
+59 million sources (27 per cent) satisfy
 these stricter criteria 
@@
 
-Fig.~\ref{fig:magdist} compares the $r$-band magnitude
-distributions for \emph{reliable}, \emph{veryReliable}
-and unfiltered objects.
-We find that 81\% of the sources 
-are considered \emph{reliable}
-and 54\% are \emph{veryReliable}
-in the magnitude range $13 < r < 19$.
-We will explain below that the
-\emph{veryReliable} category is least complete
-at low Galactic longitudes,
-where source blending can affect
-up to a quarter of the objects.
-
 \begin{figure}
-    \includegraphics[width=0.5\textwidth]{./plots/magdist/magdist-r.pdf} 
+    \includegraphics[width=0.5\textwidth]{figures/magdist/magdist-r.pdf} 
     \caption{r-band magnitude distribution for all objects in the catalogue 
@@
 
+Fig.~\ref{fig:magdist} compares the $r$-band magnitude
+distributions for \emph{reliable}, \emph{veryReliable}
+and unfiltered objects.
+We find that 81 per cent of the sources 
+are considered \emph{reliable}
+and 54 per cent are \emph{veryReliable}
+in the magnitude range $13 < r < 19$.
+We will explain below that the
+\emph{veryReliable} category is least complete
+at low Galactic longitudes,
+where source blending can affect
+up to a quarter of the objects.
+
 It is easy to see how the quality criteria
@@
 \begin{figure}
-    \includegraphics[width=0.5\textwidth]{figures/uncertainties/uncertainties.pdf} 
+    \includegraphics[width=0.45\textwidth]{figures/uncertainties/uncertainties.pdf} 
     \caption{Mean photometric uncertainties
@@
 
-\begin{figure*}
+In Fig.~\ref{fig:pairmag_reliable} we show 
+a similar comparison between the primary and secondary detections,
+but this time we have only included sources which are flagged
+as \emph{veryReliable} in the catalogue.
+We do not observe an improvement in the average residuals as a function of magnitude,
+but the number of outliers has decreased markedly
+(evidenced by the shorter error bars which denote the standard deviation of the absolute residuals).
+We conclude that the \emph{veryReliable} quality criteria are effective
+at reducing the level of the systematic errors,
+while also removing the unreliable data at the bright and faint end.
+
+\begin{figure}
 \captionsetup[subfigure]{labelformat=empty}
-\begin{subfigure}[b]{0.45\linewidth}
+\begin{subfigure}[b]{\linewidth}
 \centering
-\includegraphics[width=\textwidth]{figures/repeatability/repeatability.pdf}
+\includegraphics[width=0.9\textwidth]{figures/repeatability/repeatability.pdf}
 \caption{}
@@
 \hspace{0.5cm}
-\begin{subfigure}[b]{0.45\linewidth}
+\begin{subfigure}[b]{\linewidth}
 \centering
-\includegraphics[width=\textwidth]{figures/repeatability/repeatability-reliable.pdf}
+\includegraphics[width=0.9\textwidth]{figures/repeatability/repeatability-reliable.pdf}
 \caption{}
@@
     }
-\end{figure*}
-
-In Fig.~\ref{fig:pairmag_reliable} we show 
-a similar comparison between the primary and secondary detections,
-but this time we have only included sources which are flagged
-as \emph{veryReliable} in the catalogue.
-We do not observe an improvement in the average residuals as a function of magnitude,
-but the number of outliers has decreased markedly
-(evidenced by the shorter error bars which denote the standard deviation of the absolute residuals).
-We conclude that the \emph{veryReliable} quality criteria are effective
-at reducing the level of the systematic errors,
-while also removing the unreliable data at the bright and faint end.
-
-\subsection{Source counts and blending}
-\label{sec:densities}
-
-Fig.~\ref{fig:sourcecount} shows the number of sources
-in the catalogue counted in 1\degr-wide strips
-as a function of Galactic longitude (thick blue line).	
-Unsurprisingly, we find the number of sources
-to increase towards the Galactic centre.
-For example, the average source density near $\ell\simeq 30\degr$
-is roughly 300,000 objects per square degree,
-which is six times more than the density
-found near $\ell\simeq 180\degr$.
-In addition to the global trend,
-variations are also apparent on smaller scales.
-For example, we find a significant drop near the constellations 
-of Aquila ($\ell\simeq40\degr$) and Cygnus ($\ell\simeq80\degr$ and $\ell\simeq90\degr$),
-which are regions known to be affected
-by high levels of foreground extinction
-\citep[the extremities of `the Great Rift', e.g.][]{BokBok}.
-However, we warn that the source counts shown 
-have not been corrected for field pairs that have yet
-to be released or for variations in the depth across the included fields.
-For example, the dip near $\ell\simeq140\degr$
-is an artificial feature caused by gaps
-in the footprint coverage (seen in Fig.~\ref{fig:footprint}).
-
-In a forthcoming paper, properly-calibrated detailed maps of stellar density
-of the northern Galactic Plane will be presented (Farnhill
-et al., in preparation).  This will incorporate completeness
-corrections based on the statistics of artificial source
-recovery.  Such maps are of interest as tests of Galactic models.
-
+\end{figure}
 \begin{figure*}
+    \vspace{1cm}
     \includegraphics[width=\textwidth]{figures/sourcecount/sourcecount.pdf} 
@@
 
+\subsection{Source counts and blending}
+\label{sec:densities}
+
+Fig.~\ref{fig:sourcecount} shows the number of sources
+in the catalogue counted in 1\degr-wide strips
+as a function of Galactic longitude (thick blue line).	
+Unsurprisingly, we find the number of sources
+to increase towards the Galactic centre.
+For example, the average source density near $\ell\simeq 30\degr$
+is roughly 300,000 objects per square degree,
+which is six times more than the density
+found near $\ell\simeq 180\degr$.
+In addition to the global trend,
+variations are also apparent on smaller scales.
+For example, we find a significant drop near the constellations 
+of Aquila ($\ell\simeq40\degr$) and Cygnus ($\ell\simeq80\degr$ and $\ell\simeq90\degr$),
+which are regions known to be affected
+by high levels of foreground extinction
+\citep[the extremities of `the Great Rift', e.g.][]{BokBok}.
+However, we warn that the source counts shown 
+have not been corrected for field pairs that have yet
+to be released or for variations in the depth across the included fields.
+For example, the dip near $\ell\simeq140\degr$
+is an artificial feature caused by gaps
+in the footprint coverage (seen in Fig.~\ref{fig:footprint}).
+
+In a forthcoming paper, properly-calibrated detailed maps of stellar density
+of the northern Galactic Plane will be presented (Farnhill
+et al., in preparation).  This will incorporate completeness
+corrections based on the statistics of artificial source
+recovery.  Such maps are of interest as tests of Galactic models.
+
 Fig.~\ref{fig:sourcecount} also shows the number counts
@@
 blend-affected sources correlates with the local source 
-density.  In headline numbers, only 11\% of the sources are blended
-at $\ell>90\degr$, whereas 24\% are blended at $\ell<90\degr$.
+density.  In headline numbers, only 11 per cent of the sources are blended
+at $\ell>90\degr$, whereas 24 per cent are blended at $\ell<90\degr$.
 
@@
 can readily be used to
-(i) characterise the extinction regime along different sight-lines, and
+(i) characterise the extinction regime along different sightlines, and
 (ii) identify \ha\ emission-line objects.
@@
 and the extinction along a line of sight.
-
-This is demonstrated 
-in Fig.~\ref{fig:l180}, \ref{fig:l45} \& \ref{fig:l30},
-where we present three sets of IPHAS colour/magnitude diagrams
-towards three distinct sightlines
-located at Galactic longitudes 
-180\degr, 45\degr\ and 30\degr, respectively,
-which were chosen because they show stellar populations
-with different characteristics.
-Each figure contains all the sources
-flagged as \emph{veryReliable} within a region of one square degree 
-centred on the coordinates indicated in the diagram
-(i.e. within a radius of 0\fdg 564 from the indicated sight-line).
-For clarity, we have imposed the additional criterion
-that the photometric uncertainties
-must be smaller than 0.05 mag in each band,
-corresponding to a cut-off near 19th magnitude.
-
-Each of the diagrams reveals a well-defined locus,
-which helps to further demonstrate the health of the catalogue
-and the global calibration
-for investigating stellar populations across wide areas.
-We have annotated the colour-colour diagrams
-by showing the position 
-of the unreddened main sequence (thin solid line),
-the unreddened giant branch (thick solid line),
-and the reddening track for an A0V-type star (dashed line)
--- all three are based on the \cite{Pickles1998} library 
-of empirical spectra
-synthesised into the Vega-based IPHAS system by \cite{Drew2005}.
-In the colour-magnitude diagrams we only show the reddening vector
-together with the unreddened 1~Gyr isochrone due to \cite{Bressan2012},
-which is made available for the IPHAS system through the
-on-line tool hosted at the Observatory of Padova
-(http://stev.oapd.inaf.it/cmd).
-The isochrone and reddening vector have been placed
-at an arbitrary distance of 2~kpc.
-
-Each of the sightlines reveals a stellar population
-with distinct characteristics.
-Towards the Galactic anti-centre 
-at $\ell=180\degr$ (Fig.~\ref{fig:l180})
-we find a population dominated by lowly-reddened main sequence stars.
-This is consistent with the estimated total sightline extinction 
-of $E(B-V)=0.49$ given by \cite{Schlegel1998},
-and applying the 14\% reduction recommended by \cite{Schlafly2011}.
-Looking in more detail we can see
-that the stellar locus is narrower for M-type dwarfs
-than for earlier types:
-we do not observe M dwarfs experiencing 
-the strongest reddening possible for this sight-line.
-This implies that extinction is still increasing 
-at distances of $\sim$1-2~kpc,
-where M dwarfs become too faint to be contained in the IPHAS catalogue.
-It is also clear that there are no unreddened stars
-earlier than $\sim$K0 visible;
-such stars would be saturated if within a few hundred parsecs.
-This therefore suggests 
-that there is a measurable increase in extinction locally.
-We also note a relative absence of late type giants which,
-due to the relative brevity 
-of the corresponding phase of stellar evolution,
-would only account for a small proportion of this more nearly \emph{volume-limited} sample seen in the Anticentre direction.
-
-In contrast, lines of sight passing into the first Galactic quadrant
-yield samples that are more commonly \emph{magnitude-limited} instead.
-For example, at $\ell=45\degr$ (Fig.~\ref{fig:l45}),
-there is  a wealth of reddened objects and late type giants.
-In the colour-magnitude diagram, it is clear
-that the stars are split into two distinct groups,
-with one significantly redder than the other.
-The bluer group is composed of main sequence stars,
-with the slope of this group in the colour-magnitude diagram
-attributable to the significantly increasing extinction.
-Meanwhile the redder group is principally composed of red giant stars \citep[see][]{Wright2008}.
-As these stars are intrinsically brighter, they will be substantially 
-further away than their main sequence counterparts at the same 
-apparent magnitude.  Given that extinction continues to increase 
-with distance, along this sightline, 
-the red giants we observe will be subject to appreciably more reddening
-than the main sequence stars, pushing them to $(r - i) \sim 1.5$.
-
-Finally, in one of our lowest-longitude sightlines at $\ell=30\degr$,
-we find a very high number of extremely reddened giants
-in addition to an unreddened population
-of foreground dwarfs.
-In contrast to the sight-line at $\ell=45\degr$,
-there is no clear group of giant stars visible
-in the colour-magnitude diagram of Fig.~\ref{fig:l30},
-although the red clump stars are manifest as a track of slight 
-over-density sitting roughly 0.4~mag redder than the A0V reddening track.
-At $(l,b)=(45\degr, +2\degr)$ the giant stars observed
-exhibit a relatively narrow range of reddenings
-as they lie beyond most of the Galactic dust column.
-At $(l,b)=(30\degr, 0\degr)$ this is not the case:
-even at the substantial distances
-at which we can observe reddened giant stars,
-extinction is continuing to rise within the Galactic mid-plane.
-It is also apparent that the $(r-i)$ width of both the M dwarfs
-and early A dwarfs is greater than that in Fig.~\ref{fig:l45}.
-This is indicative of a steeper rise in reddening,
-both within several hundred parsecs (M dwarfs)
-and within a few kpc (early A dwarfs).
-
-These are just descriptive vignettes of the information obtainable
-from IPHAS colour-colour and colour-magnitude plots.
-A more rigorous quantitative analysis of the IPHAS catalogue
-can be undertaken to estimate both the stellar density distribution 
-in the Milky Way \citep{Sale2010} and to create detailed three-dimensional 
-maps of the extinction across several kpc \citep{Sale2009,Sale2012}.
-A 3-D extinction map based on the DR2 catalogue
-is being released in a separate paper \citep{Sale2014}.
 
@@
     both placed at an arbitrary distance of 2~kpc.
-    This is one of the least reddened sight-lines
+    This is one of the least reddened sightlines
     in the survey
@@
     \caption{Same as above for $(l,b)=(45\degr,+2\degr)$,
-    which is one of the highest-density sight-lines in the survey,
+    which is one of the highest-density sightlines in the survey,
     revealing two groups of stars in colour-magnitude space.}
@@
     \caption{Same as above for $(l,b)=(30\degr,0\degr)$,
-    showing one of the most reddened sight-lines in the survey.
+    showing one of the most reddened sightlines in the survey.
     }
@@
 \end{figure*}
+
+This is demonstrated 
+in Fig.~\ref{fig:l180}, \ref{fig:l45} \& \ref{fig:l30},
+where we present three sets of IPHAS colour/magnitude diagrams
+towards three distinct sightlines
+located at Galactic longitudes 
+180\degr, 45\degr\ and 30\degr, respectively,
+which were chosen because they show stellar populations
+with different characteristics.
+Each figure contains all the sources
+flagged as \emph{veryReliable} within a region of one square degree 
+centred on the coordinates indicated in the diagram
+(i.e. within a radius of 0\fdg 564 from the indicated sightline).
+For clarity, we have imposed the additional criterion
+that the photometric uncertainties
+must be smaller than 0.05 mag in each band,
+corresponding to a cut-off near 19th magnitude.
+
+Each of the diagrams reveals a well-defined locus,
+which helps to further demonstrate the health of the catalogue
+and the global calibration
+for investigating stellar populations across wide areas.
+We have annotated the colour-colour diagrams
+by showing the position 
+of the unreddened main sequence (thin solid line),
+the unreddened giant branch (thick solid line),
+and the reddening track for an A0V-type star (dashed line)
+-- all three are based on the \cite{Pickles1998} library 
+of empirical spectra
+synthesised into the Vega-based IPHAS system by \cite{Drew2005}.
+In the colour-magnitude diagrams we only show the reddening vector
+together with the unreddened 1~Gyr isochrone due to \cite{Bressan2012},
+which is made available for the IPHAS system through the
+on-line tool hosted at the Observatory of Padova
+(http://stev.oapd.inaf.it/cmd).
+The isochrone and reddening vector have been placed
+at an arbitrary distance of 2~kpc.
+
+Each of the sightlines reveals a stellar population
+with distinct characteristics.
+Towards the Galactic anti-centre 
+at $\ell=180\degr$ (Fig.~\ref{fig:l180})
+we find a population dominated by lowly-reddened main sequence stars.
+This is consistent with the estimated total sightline extinction 
+of $E(B-V)=0.49$ given by \cite{Schlegel1998},
+and applying the 14 per cent reduction recommended by \cite{Schlafly2011}.
+Looking in more detail we can see
+that the stellar locus is narrower for M-type dwarfs
+than for earlier types:
+we do not observe M dwarfs experiencing 
+the strongest reddening possible for this sightline.
+This implies that extinction is still increasing 
+at distances of $\sim$1-2~kpc,
+where M dwarfs become too faint to be contained in the IPHAS catalogue.
+It is also clear that there are no unreddened stars
+earlier than $\sim$K0 visible;
+such stars would be saturated if within a few hundred parsecs.
+This therefore suggests 
+that there is a measurable increase in extinction locally.
+We also note a relative absence of late type giants which,
+due to the relative brevity 
+of the corresponding phase of stellar evolution,
+would only account for a small proportion of this more nearly \emph{volume-limited} sample seen in the Anticentre direction.
+
+In contrast, lines of sight passing into the first Galactic quadrant
+yield samples that are more commonly \emph{magnitude-limited} instead.
+For example, at $\ell=45\degr$ (Fig.~\ref{fig:l45}),
+there is  a wealth of reddened objects and late type giants.
+In the colour-magnitude diagram, it is clear
+that the stars are split into two distinct groups,
+with one significantly redder than the other.
+The bluer group is composed of main sequence stars,
+with the slope of this group in the colour-magnitude diagram
+attributable to the significantly increasing extinction.
+Meanwhile the redder group is principally composed of red giant stars \citep[see][]{Wright2008}.
+As these stars are intrinsically brighter, they will be substantially 
+further away than their main sequence counterparts at the same 
+apparent magnitude.  Given that extinction continues to increase 
+with distance, along this sightline, 
+the red giants we observe will be subject to appreciably more reddening
+than the main sequence stars, pushing them to $(r - i) \sim 1.5$.
+
+Finally, in one of our lowest-longitude sightlines at $\ell=30\degr$,
+we find a very high number of extremely reddened giants
+in addition to an unreddened population
+of foreground dwarfs.
+In contrast to the sightline at $\ell=45\degr$,
+there is no clear group of giant stars visible
+in the colour-magnitude diagram of Fig.~\ref{fig:l30},
+although the red clump stars are manifest as a track of slight 
+over-density sitting roughly 0.4~mag redder than the A0V reddening track.
+At $(l,b)=(45\degr, +2\degr)$ the giant stars observed
+exhibit a relatively narrow range of reddenings
+as they lie beyond most of the Galactic dust column.
+At $(l,b)=(30\degr, 0\degr)$ this is not the case:
+even at the substantial distances
+at which we can observe reddened giant stars,
+extinction is continuing to rise within the Galactic mid-plane.
+It is also apparent that the $(r-i)$ width of both the M dwarfs
+and early A dwarfs is greater than that in Fig.~\ref{fig:l45}.
+This is indicative of a steeper rise in reddening,
+both within several hundred parsecs (M dwarfs)
+and within a few kpc (early A dwarfs).
+
+These are just descriptive vignettes of the information obtainable
+from IPHAS colour-colour and colour-magnitude plots.
+A more rigorous quantitative analysis of the IPHAS catalogue
+can be undertaken to estimate both the stellar density distribution 
+in the Milky Way \citep{Sale2010} and to create detailed three-dimensional 
+maps of the extinction across several kpc \citep{Sale2009,Sale2012}.
+A 3-D extinction map based on the DR2 catalogue
+is being released in a separate paper \citep{Sale2014}.
 
@@
 \citep{Mayer1973,Lahulla1985,Hunter1990}.
+
+\begin{figure*}
+    \begin{minipage}[b]{0.8\linewidth}
+        \includegraphics[width=\textwidth]{figures/sh2-82/sh2-82-iphas.jpg} 
+    \end{minipage}
+\caption{IPHAS image mosaic of H{\sc ii} region Sh 2-82,
+composed of \ha\ (red channel), $r$ (green channel) and $i$ (blue channel). Yellow triangles show the position of candidate \ha-emitters
+which have been selected from the colour-colour diagram
+in Fig.~\ref{fig:emitters}. 
+}
+\label{fig:mosaic_iphas}
+    \begin{minipage}[b]{0.8\linewidth}
+        \includegraphics[width=\textwidth]{figures/sh2-82/sh2-82-spitzer.jpg} 
+    \end{minipage}
+    \caption{Star-forming region Sh 2-82 as seen in the mid-infrared
+    by the Spitzer Space Telescope. The mosaic is composed of the 24~\micron\ (red), 8.0~\micron\ (green) and 4.5~\micron\ (blue) bands.
+    The image reveals a bubble-shaped structure which surrounds the {\sc Hii} region that is seen in the IPHAS mosaic which spans the same region (Fig.~\ref{fig:mosaic_iphas}). 
+    This structure has previously been labelled as N115 in the 
+catalogue of \citet{Churchwell2006}, and could be a possible site of triggered star formation \citep{Thompson2012,Kendrew2012}.}
+    \label{fig:mosaic_spitzer}
+\end{figure*}
 
@@
 
-\begin{figure*}
-    \begin{minipage}[b]{0.8\linewidth}
-        \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-iphas.pdf} 
-    \end{minipage}
-\caption{IPHAS-based mosaic of H{\sc ii} region Sh 2-82,
-composed of \ha\ (red channel), $r$ (green channel) and $i$ (blue channel). Yellow triangles show the position of candidate \ha-emitters
-which have been selected from the colour-colour diagram
-in Fig.~\ref{fig:emitters}. 
-%The H{\sc ii} region is surrounded by a very faint blue/green reflection nebula and dark cloud filaments.
-}
-\label{fig:mosaic_iphas}
-    \begin{minipage}[b]{0.8\linewidth}
-        \includegraphics[width=\textwidth]{./plots/mosaic/sh2-82-spitzer.pdf} 
-    \end{minipage}
-    \caption{Star-forming region Sh 2-82 as seen in the mid-infrared
-    by the Spitzer Space Telescope. The mosaic is composed of the 24~\micron\ (red), 8.0~\micron\ (green) and 4.5~\micron\ (blue) bands.
-    The image reveals a bubble-shaped structure which surrounds the {\sc Hii} region that is seen in the IPHAS mosaic which spans the same region (Fig.~\ref{fig:mosaic_iphas}). 
-    This structure has previously been labelled as N115 in the 
-catalogue of \citet{Churchwell2006}, and could be a possible site of triggered star formation \citep{Thompson2012,Kendrew2012}.}
-    \label{fig:mosaic_spitzer}
-\end{figure*}
+\begin{figure}
+\begin{center}
+  \includegraphics[width=\linewidth]{figures/sh2-82/sh2-82-ccd.pdf}
+    \caption{$(r-\ha,\ r-i)$ diagram for the rectangular region of 
+    20-by-15 arcmin centred on the H{\sc ii} region Sh 2-82,
+    which is the area shown in Fig.~\ref{fig:mosaic_iphas}.    
+    The diagram shows all objects in the catalogue
+    which have been flagged as \emph{reliable} and are brighter
+    than $r<20$ (grey circles).
+    The unreddened main sequence is indicated by a solid line,
+    while the main sequence for stars with an \ha\ emission line
+    strength of $-10~\rm{\AA}$~EW is indicated by a dashed line
+    \citep[both based on the colour simulations by][]{Barentsen2011a}.
+    Red triangles indicate objects which have been identified as
+    as likely \ha-emitters.}
+    \label{fig:emitters}
+\end{center}
+\end{figure}
 
@@
 
-\begin{figure}
-  \includegraphics[width=0.45\textwidth]{./plots/sh2-82-ccd.pdf}
-    \caption{$(r-\ha,\ r-i)$ diagram for the rectangular region of 
-    20-by-15 arcmin centred on the H{\sc ii} region Sh 2-82,
-    which is the area shown in Fig.~\ref{fig:mosaic_iphas}.    
-    The diagram shows all objects in the catalogue
-    which have been flagged as \emph{reliable} and are brighter
-    than $r<20$ (grey circles).
-    The unreddened main sequence is indicated by a solid line,
-    while the main sequence for stars with an \ha\ emission line
-    strength of $-10~\rm{\AA}$~EW is indicated by a dashed line
-    \citep[both based on the colour simulations by][]{Barentsen2011a}.
-    Red triangles indicate objects which have been identified as
-    as likely \ha-emitters.}
-    \label{fig:emitters}
-\end{figure}
-
 \begin{table}
@@
     \footnotesize
-    \resizebox{0.45\textwidth}{!}{
+    \resizebox{0.47\textwidth}{!}{
         \begin{tabular}{clccc}
@@
 each covering a $5\degr\times5\degr$ tile
-of the footprint.
+of the footprint and comprising 50~gigabyte
+in total.
 
@@
 
+\vspace{0.5cm}
+
 \section{Concluding remarks and future work}
@@
 The new 98-column catalogue provides single-epoch photometry
-across 92\% of the survey area,
+across 92 per cent of the survey area,
 and is the first quality-controlled and globally calibrated
@@
 The INT is operated on the island of La Palma
-by the Isaac Newton Group
+by the Isaac Newton Group (ING)
 in the Spanish Observatorio del Roque de los Muchachos
 of the Instituto de Astrof\'\i sica de Canarias.
-We are deeply indebted to the ING staff and students
-for their ongoing support.
+We are deeply indebted to Ovidiu Vaduvescu,
+Javier M\'endez and the rest of the ING staff and students
+for their ongoing support of the telescope.
 All data were processed 

### 2014-06-26 dr2/images.py
Account for CCDs that cross 0h RA
CCDs crossing 0h RA should now return correct ra_min and ra_max values.

@@ -319,6 +319,10 @@ class SurveyImage(object):
                                        [2048, 1]],
                                       1)
         ra1, ra2 = np.min(corners[:, 0]), np.max(corners[:, 0])
+        # If CCD crosses 0h RA, then need to separate corners on either side
+        if ra2 - ra1 > 180:
+            ra1 = np.max(corners[:, 0][corners[:, 0] < 180])
+            ra2 = np.min(corners[:, 0][corners[:, 0] > 180])
         dec1, dec2 = np.min(corners[:, 1]), np.max(corners[:, 1])
 
         if self.calibrated:

### 2014-06-27 dr2/images.py
-

@@ -4,6 +4,8 @@
 
 This script will edit all images in the data release to ensure they have
 the correct astrometric solution (WCS) and calibration information (PHOTZP).
+
+This script will fail for r376022.fit, which is unfortunately a corrupt file.
 """
 from __future__ import division, print_function, unicode_literals
 from astropy.io import fits
@@ -376,16 +378,16 @@ def prepare_one(run):
     with log.log_to_file(os.path.join(constants.LOGDIR, 'images.log')):
         result = []
         for ccd in constants.EXTENSIONS:
-            try:
-                img = SurveyImage(run, ccd)
-                img.save()
-                result.append(img.get_metadata())
-                img.fits.close()  # avoid memory leak
-            except Exception, e:
-                log.error(str(run) + ': '
-                          + util.get_pid() + ': '
-                          + e.__class__.__name__ + ': '
-                          + str(e))
+            #try:
+            img = SurveyImage(run, ccd)
+            img.save()
+            result.append(img.get_metadata())
+            img.fits.close()  # avoid memory leak
+            #except Exception, e:
+            #    log.error(str(run) + ': '
+            #              + util.get_pid() + ': '
+            #              + e.__class__.__name__ + ': '
+            #              + str(e))
         return result
 
 
@@ -441,4 +443,4 @@ if __name__ == '__main__':
         import os
     prepare_images(client[:])
     """
-    prepare_one(943312)
+    prepare_one(367744)

### 2014-06-27 paper/iphas-dr2-arxiv.tex

--- 2014-04-17 paper/iphas-dr2.tex
@@
 \documentclass[a4paper,useAMS,usenatbib]{mn2e}
+\pdfpagewidth=\paperwidth
+\pdfpageheight=\paperheight
 \usepackage{amssymb,amsmath}
@@
 M. J. Irwin$^2$,
-B. Mizalski$^{4}$,
+B. Miszalski$^{4}$,
 C. Ruhland$^1$,
@@
 S. Phillipps$^{21}$, 
-S. Pyrzas$^{26}$
+S. Pyrzas$^{26}$,
 R. Raddi$^{12}$,
@@
 %$^{32}$Armagh Observatory, College Hill, Armagh, Northern Ireland, BT61 9DG, U.K.\\
-$^{14-32}$ (note to editor: affiliations 14 through 32 are given in the source file)
+$^{14-32}$ (affiliations 14 through 32 are given in the source file)
 }
@@
 
-The catalogue is made available through the
+The catalogue will be made available through the
 Vizier service (http://vizier.u-strasbg.fr),
@@
 This implies that the in-band flux corresponding to
-zero magnitude is $1.56 \time 10^{-7}$ erg\,cm$^{-2}$\,s$^{-1}$,
+zero magnitude is $1.56 \times 10^{-7}$ erg\,cm$^{-2}$\,s$^{-1}$,
 when the \ha\ magnitude for Vega is set by convention to 0.03 \citep{Fukugita1996}.
@@
   \bibliographystyle{mn2e}
-  \bibliography{iphas-dr2}
+  \bibliography{iphas-dr2-arxiv}
 }

### 2014-06-27 dr2/images.py
Resolves ra_min vs ra_max bug for fields crossing the meridian

@@ -320,11 +320,11 @@ class SurveyImage(object):
                                        [2048, 4096],
                                        [2048, 1]],
                                       1)
-        ra1, ra2 = np.min(corners[:, 0]), np.max(corners[:, 0])
+        ra_min, ra_max = np.min(corners[:, 0]), np.max(corners[:, 0])
         # If CCD crosses 0h RA, then need to separate corners on either side
-        if ra2 - ra1 > 180:
-            ra1 = np.max(corners[:, 0][corners[:, 0] < 180])
-            ra2 = np.min(corners[:, 0][corners[:, 0] > 180])
+        if ra_max - ra_min > 180:
+            ra_min = np.min(corners[:, 0][corners[:, 0] > 180])
+            ra_max = np.max(corners[:, 0][corners[:, 0] < 180]) + 360.
         dec1, dec2 = np.min(corners[:, 1]), np.max(corners[:, 1])
 
         if self.calibrated:
@@ -345,8 +345,8 @@ class SurveyImage(object):
                 'in_dr2': in_dr2,
                 'ra': ra,
                 'dec': dec,
-                'ra_min': ra1,
-                'ra_max': ra2,
+                'ra_min': ra_min,
+                'ra_max': ra_max,
                 'dec_min': dec1,
                 'dec_max': dec2,
                 'band': band,
@@ -378,20 +378,15 @@ def get_caldb():
         return CALDB
 
 
-def prepare_one(run):
+def prepare_one(run, save=False):
     with log.log_to_file(os.path.join(constants.LOGDIR, 'images.log')):
         result = []
         for ccd in constants.EXTENSIONS:
-            #try:
             img = SurveyImage(run, ccd)
-            img.save()
+            if save:
+                img.save()
             result.append(img.get_metadata())
             img.fits.close()  # avoid memory leak
-            #except Exception, e:
-            #    log.error(str(run) + ': '
-            #              + util.get_pid() + ': '
-            #              + e.__class__.__name__ + ': '
-            #              + str(e))
         return result
 
 
@@ -407,7 +402,7 @@ def prepare_images(clusterview):
         else:
             idx_band = band
         # [constants.IPHASQC_COND_RELEASE]
-        runs = constants.IPHASQC['run_'+idx_band]
+        runs = constants.IPHASQC['run_'+idx_band][0:10]
         # Prepare each run
         result = clusterview.map(prepare_one, runs, block=True)
         metadata.extend(result)
[…]

### 2014-06-27 paper/iphas-dr2-arxiv.tex
arxiv version of paper

@@
 can be found in the works of \citet{Sale2014}, which applies DR2 to the problem
-of 3D extinction mapping, and of Sabin et al (in preparation), where the results of a search
+of 3D extinction mapping, and of Sabin et al (submitted), where the results of a search
 of the image database for new planetary nebulae is presented.

### 2014-06-27 paper/tar-arxiv.sh

@@ -1,1 +1,3 @@
+# Warning: make sure iphas-dr2-arxiv.bbl exists and is an up-to-date bbl file
 tar -cvf arxiv.tar iphas-dr2-arxiv.tex iphas-dr2-arxiv.bbl mn2e.bst mn2e.cls aas_macros.sty tables/columns.tex figures/footprint/footprint_small.png figures/depth/depth_r.pdf figures/depth/depth_i.pdf figures/depth/depth_h.pdf figures/seeing/seeing_r.pdf figures/seeing/seeing_i.pdf figures/seeing/seeing_ha.pdf figures/caldiagram/ccd-uncalibrated.pdf figures/caldiagram/ccd-calibrated.pdf figures/calibration/APASS-IPHAS-DR2_ishift.pdf figures/calibration/APASS-IPHAS-DR2_rshift.pdf figures/calibration/colourbar_apass_r.pdf figures/calibration/colourbar_apass_i.pdf figures/calibration/SDSS-IPHAS_rshift.pdf figures/calibration/colourbar_sdss_r.pdf figures/calibration/SDSS-IPHAS_ishift.pdf figures/calibration/colourbar_sdss_i.pdf figures/magdist/magdist-r.pdf figures/uncertainties/uncertainties.pdf figures/repeatability/repeatability.pdf figures/repeatability/repeatability-reliable.pdf figures/sourcecount/sourcecount.pdf figures/diagrams/ccd-* figures/diagrams/cmd-* figures/sh2-82/sh2-82-*
+gzip arxiv.tar

### 2014-06-27 dr2/images.py
clean-up after iphas-images regeneration

@@ -378,15 +378,18 @@ def get_caldb():
         return CALDB
 
 
-def prepare_one(run, save=False):
+def prepare_one(run, save=True):
     with log.log_to_file(os.path.join(constants.LOGDIR, 'images.log')):
         result = []
         for ccd in constants.EXTENSIONS:
-            img = SurveyImage(run, ccd)
-            if save:
-                img.save()
-            result.append(img.get_metadata())
-            img.fits.close()  # avoid memory leak
+            try:
+                img = SurveyImage(run, ccd)
+                if save:
+                    img.save()
+                result.append(img.get_metadata())
+                img.fits.close()  # avoid memory leak
+            except Exception, e:
+                log.error(e)
         return result
 
 
@@ -402,7 +405,7 @@ def prepare_images(clusterview):
         else:
             idx_band = band
         # [constants.IPHASQC_COND_RELEASE]
-        runs = constants.IPHASQC['run_'+idx_band][0:10]
+        runs = constants.IPHASQC['run_'+idx_band]
         # Prepare each run
         result = clusterview.map(prepare_one, runs, block=True)
         metadata.extend(result)

### 2014-06-30 paper/iphas-dr2-rev1.tex
Initial commit of paper revision

--- 2014-06-27 paper/iphas-dr2-arxiv.tex
@@
+% This is revision 1 of the IPHAS DR2 paper,
+% created in response to the referee comments received on 24 June 2014
 \documentclass[a4paper,useAMS,usenatbib]{mn2e}
@@
 \usepackage{subcaption}
-
-%\voffset=-0.5in
 
@@
 
+\begin{table}
+    \caption{Overview of the key survey properties.}
+    \label{tbl:survey}
+    \begin{center}
+        \begin{tabular}{ll}
+        \toprule
+        Property & Value \\
+        \midrule
+        Telescope & 2.5m Isaac Newton Telescope, La Palma \\
+        Instrument & Wide Field Camera \\
+        FoV CCD & ... \\
+        Dates & ... \\
+        Detector scale & 0.33 arcsec/pixel \\
+        Filters & $r$, $i$, \ha \\
+        Filter properties & See Table~2 \\
+        Exposure time & 30~s ($r$), 10~s ($i$), 120~s (\ha) \\
+        Magnitude limit (5$\sigma$, Vega) & 21.3 ($r$), 20.1 ($i$), 20.3 (\ha) \\
+        PSF FWHM (median) & 1.1\arcsec\ ($r$), 1.0\arcsec\ ($i$), 1.1\arcsec\ (\ha) \\
+        Survey area & $\sim 1860$~deg$^2$ \\
+        Footprint limits & $ -5\degr < b < +5\degr,\ 29\degr < \ell < 215\degr$ \\
+        \bottomrule
+        \end{tabular}
+	\end{center}
+\end{table}
+
 The detailed properties of the IPHAS observing programme 
@@
   \bibliographystyle{mn2e}
-  \bibliography{iphas-dr2-arxiv}
+  \bibliography{iphas-dr2}
 }

### 2014-07-01 paper/iphas-dr2-rev1.tex
Initial revisions

--- 2014-06-27 paper/iphas-dr2-arxiv.tex
@@
 
-When the initial data release was created in late 2007,
+{\bf When the previous release -- the IDR -- was created in late 2007,
 just over half of the survey footprint was covered
@@
 to support a homogeneously calibrated source catalogue.
-The goal of this paper is to present the next release 
-that takes the coverage up to 92 per cent of the survey area 
-and includes a global photometric re-calibration.
+The goal of this paper is to present the next release, 
+ which supersedes the IDR by providing a global re-calibration
+and taking the coverage up to 92 per cent of the survey area.}
 In this work we
@@
 
+\begin{table}
+    \caption{Key properties of IPHAS DR2.}
+    \label{tbl:survey}
+    \begin{center}
+        \begin{tabular}{ll}
+        \toprule
+        Property & Value \\
+        \midrule
+        Telescope & 2.5m Isaac Newton Telescope, La Palma \\
+        Instrument & Wide Field Camera \\
+        Detectors & Four 2048 $\times$ 4100 pixel CCD's \\
+        %Detector field of view & 11 $\times$ 23 arcmin \\
+        Detector scale & 0.33 arcsec/pixel \\        
+        Filters & $r$, $i$, \ha \\
+        Filter properties & See Table~2 \\
+        Exposure times & 30~s ($r$), 10~s ($i$), 120~s (\ha) \\
+		Saturation limit & TBC \\
+        Detection limit (5$\sigma$, Vega) & 21.3 ($r$), 20.1 ($i$), 20.3 (\ha) \\
+        PSF FWHM (median) & 1.1\arcsec\ ($r$), 1.0\arcsec\ ($i$), 1.1\arcsec\ (\ha) \\
+        Survey area & $\sim 1860$~deg$^2$ \\
+        Footprint boundaries & $ -5\degr < b < +5\degr,\ 29\degr < \ell < 215\degr$ \\
+        DR2 observing period & August 2003 - November 2012 \\
+        Website & www.iphas.org \\
+        \bottomrule
+        \end{tabular}
+	\end{center}
+\end{table}
+
 The detailed properties of the IPHAS observing programme 
@@
 by \citet{Drew2005} and \citet{Gonzalez-Solares2008}. 
-To set the stage for this release, we recap some key points.
+To set the stage for this release,
+we recap some key points {\bf below and in Table~1}.
 IPHAS is an imaging survey of the Galactic Plane north of the celestial equator, 
@@
 to the Cambridge Astronomical Survey Unit (CASU) 
-for initial processing and archival.
+for initial processing and archiving.
 The procedures used by CASU were originally devised
@@
 (iii) contains user-friendly columns and warning flags.
-It is these post-processing steps which are explained next.
+{ \bf It is these post-processing steps which distinguish the present release
+from the IDR, which enforced less stringent quality limits,
+did not offer a global calibration,
+and provided a catalogue in which duplicate detections
+of unique sources were not flagged.
+These new processing steps are explained next.}
+%It is these post-processing steps which are explained next.
 
@@
          while bins containing more than 400 sources are bright yellow.
-         The diagrams include all stars
+         {\bf The diagrams include all stars
          brighter than $r<18$ which were
-         classified by the pipeline as being `very reliable'
-         (to be explained in
-         \S\ref{sec:qualitycriteria}).
+         classified by the pipeline as `bestQualityAllBand'
+         (indicating a high-significance point source detection with
+         accurate photometry in all bands,
+         to be explained in \S\ref{sec:qualitycriteria})}.
          The objects which are seen to fall above the locus of stars
@@
 to validate the results.
-SDSS DR9 includes several strips at low
+{\bf SDSS DR9 includes photometry across several strips at low
 Galactic latitudes,
-providing data across 18 per cent of the fields in our data release.
+which were observed as part of 
+the Sloan Extension for Galactic Understanding and Exploration \citep[SEGUE;][]{segue}.}
+These strips provide data across 18 per cent of the fields in our data release.
 We cross-matched the IPHAS fields against the subset of
@@
 which desire reliable colours
-as well as completeness.
+as well as completeness {\bf down to $\sim$19th magnitude}.
 The listed criteria are designed to 
@@
 (39 per cent) satisfy all the criteria listed in Table~\ref{tab:reliable}
-and are hereafter referred to as `reliable'.
+and are hereafter referred to as {\bf `highQualityAllBand'}.
 For convenience, the catalogue contains a boolean column
-named \emph{reliable} that directly flags these objects.
+named {\bf \emph{highQualityAllBand}} that directly flags these objects.
 
@@
 \caption{Photometric repeatability as a function of magnitude
-         for all sources in the catalogue (left panel)
-         and for the \emph{veryReliable} sources alone (right panel).
+         for all sources in the catalogue (panel a)
+         and for the \emph{veryReliable} sources alone (panel b).
          Black dots show the mean absolute residuals

### 2014-07-01 paper/iphas-dr2-rev1.tex
Added GLIMPSE comparison

@@
 
+{\bf Moreover, IPHAS provides the much-needed counterpart
+to the source catalogues being provided by the recent infrared surveys.
+For example, amongst the 1810 high-quality sources considered in Fig.~\ref{fig:emitters},
+1623 objects (90\%) have a counterpart in the Spitzer catalogue,
+and 371 objects (20\%) have a counterpart in the 
+Wide-field Infrared Survey Explorer catalogue (WISE; Cutri et al. 2013),
+both within a matching radius of 2 arcsec.}
+
 \section{Catalogue and image access}

### 2014-07-02 paper/iphas-dr2-rev1.tex
Added section 7.3

@@
         \midrule
-        Telescope & 2.5m Isaac Newton Telescope, La Palma \\
+        Telescope & 2.5m Isaac Newton Telescope \\
         Instrument & Wide Field Camera \\
@@
         Exposure times & 30~s ($r$), 10~s ($i$), 120~s (\ha) \\
-		Saturation limit & TBC \\
-        Detection limit (5$\sigma$, Vega) & 21.3 ($r$), 20.1 ($i$), 20.3 (\ha) \\
+		Saturation limit & $\sim$13 ($r$), $\sim$12.5 ($i$), $\sim$12.5 (\ha) \\
+        Detection limit (5$\sigma$) & 21.3 ($r$), 20.1 ($i$), 20.3 (\ha) \\
+		Magnitude system & Vega \\
         PSF FWHM (median) & 1.1\arcsec\ ($r$), 1.0\arcsec\ ($i$), 1.1\arcsec\ (\ha) \\
@@
         Footprint boundaries & $ -5\degr < b < +5\degr,\ 29\degr < \ell < 215\degr$ \\
-        DR2 observing period & August 2003 - November 2012 \\
+        Observing period & August 2003 - November 2012 \\
         Website & www.iphas.org \\
@@
 
+\subsection{$r - \ha$ excess as a quantitative measure of H$\alpha$ emission}
+
+{ \bf
+
+As well as discover emission line stars, IPHAS data can provide a
+first estimate of the equivalent width of line emission.
+How this is done has been discussed in previous works 
+-- most notably by \citet{Drew2005}
+and more recently by \cite{Barentsen2011a}.
+Given that the narrowband filter used has a FWHM of 95~\AA ,
+it is to be expected that the appearance
+in a spectrum of line emission corresponding to an equivalent width (EW)
+of 95~\AA\ would increase $r - \ha$ by 0.75 magnitudes
+(i.e. the flux captured within the narrow band doubles).
+Similarly, 10~\AA\ of line emission
+should increase $r - H\alpha$ by 0.11~mag.
+
+This simple reasoning breaks down
+when the $H\alpha$ emission becomes so bright
+that it also dominates the flux captured across the entire $r$ broadband.
+It was noted by \cite{Drew2005}
+that the very bright line limiting value 
+of $r - \ha$ is $\sim$3.1 in the Vega system.
+This fact, on its own, implies that any apparent detection
+of an \ha\ emission line star without a corresponding $r$ 
+detection is only real if the \ha\ magnitude is not brighter 
+by more than 3.1~mag relative to the $r$ detection limit
+(i.e. typically, credible H$\alpha$-only sources
+must be fainter than $\sim$18th mag).
+
+In order to infer the emission EW from the DR2 $r - \ha$ colour
+it is necessary to know the value this colour would have
+in the case of no excess line emission.
+In the {\em general} case this is not possible
+without some prior knowledge of the star's reddening and spectral type
+(i.e. the inferences drawn for an unreddened M dwarf or
+a reddened F star would not be the same -- the absolute value of the 
+equivalent width in the latter case would be greater because the 
+reddened F star {\em without} emission would lie at lower $r -\ha$).
+This is the situation faced in the study of IC 1396 by
+\cite{Barentsen2011a} in which \ha\ EWs were inferred
+relative to a reddened sequence of K-M stars
+appropriate to this star-forming region.
+A different choice of reddening would have resulted
+in different EW estimates.
+
+However it was found by \cite{Drew2005}
+that for earlier type stars in which the spectrum
+is not heavily modulated by line features or molecular bands
+the estimation of net emission EW is more straightforward and unambiguous.  
+It was demonstrated, for spectral slopes ranging
+from the Rayleigh-Jeans limit to that appropriate to an early G star
+and for a range of reddenings,
+that it is possible to draw a unique set of constant equivalent-width 
+lines in the $(r - \ha, r - i)$ diagram
+that can be compared directly with the positions
+of candidate emission line stars,
+thereby predicting emission EW \citep[see fig.~6, ][]{Drew2005}.
+The EW in this case should be understood
+to be the measure of the emission observed
+above the interpolated continuum
+(i.e. no correction is present
+for any assumed infilled photospheric line absorption).
+
+In Fig.~\ref{fig:raddi} we show how well this works in reality:
+the predicted lines of constant emission EW
+in the $(r - \ha, r - i)$ plane
+are compared with the positions of a sample
+of over 200 classical Be stars in Perseus (from Raddi et al, 2014).
+Stars of this type come with the advantage that secular variation,
+whilst itcertainly occurs,
+is not as common as in T Tau stars.
+The objects plotted are a bright selection ($r \lesssim 16.5$)
+for which there are spectroscopic determinations
+of \ha\ emission equivalent width.
+A colour scale has been applied to the data points
+that darkens as EW increases.
+The gradation apparent in the shading of the datapoints
+follows the trend set by the constant EW lines quite well.  
+The experience has been that the EW deduced from
+IPHAS photometry for individual objects is commonly within 10~\AA\
+of subsequent spectroscopic measurement
+\citep[see also fig.~5 of][]{Barentsen2011a}.
+Evidently, the photometry is well-suited to statistical measurement
+across large samples, while for individual objects a useful
+approximation is delivered.
+
+\begin{figure}
+\begin{center}
+  \includegraphics[width=\linewidth]{figures/raddi.pdf}
+    \caption{Photometric predictions of \ha\ equivalent width (EW)
+    compared with spectroscopic measurements
+    of a sample of classical Be stars (from Raddi et al 2014, submitted).
+    The broken lines are predicted lines
+    of constant net emission EW,
+    while the analogous spectroscopic EW is represented
+    by the colouring of the data points.
+    The unreddened main sequence, which normally serves 
+    as the upper bound to the main stellar locus,
+    is drawn as a solid black line.
+    Strongly reddened stars with not-so-strong line emission
+    will fall below it
+    and be hard to pick out without additional information.
+    The two vertical lines drawn are 'curves of growth'
+    showing how $r - H\alpha$ increases 
+    -- and $r - i$ drops a little --
+    as more and more line emission is added,
+    raising the flux in \ha\ narrowband
+    and to a lesser proportion the flux in the $r$ band.
+    The blue line represents the trend
+    for an unreddened Rayleigh-Jeans continuum
+    (with superposed \ha\ emission),
+    while the green line is the trend
+    for the case of an unreddened A0 continuum.}
+    \label{fig:raddi}
+\end{center}
+\end{figure}
+
+}
+
 \section{Catalogue and image access}

### 2014-07-14 paper/iphas-dr2-rev1.tex
Added reliability subsection

@@
 
+\subsection{Reliability}
+\label{sec:reliability}
+
+In producing a point source catalogue, spurious detections 
+will inevitably be made. A fraction of these will mimic stellar 
+sources closely enough that they cannot be excluded using the 
+recorded parameters.
+
+%The majority of sources have been observed in IPHAS more than 
+%once (for objects with S/N $>$ 10 in the $r$-band, 62\% of 
+%objects have secondary detections), however there are regions 
+%observed only once - e.g. survey edges, intersections of inter-CCD 
+%gaps. As such, requiring that objects have a secondary observation 
+%to be considered reliable eliminates a significant fraction of
+%genuine objects..
+
+A statistic that can represent the confidence that objects in the 
+catalogue are genuine is the reliability of the survey. The 
+reliability gives the fraction of objects that can be repeatedly 
+observed; provided the comparison is made to an external source of 
+comparable depth this will provide the ratio of genuine astronomical 
+sources to sources in the catalogue.
+
+The IPHAS catalogue was compared to the SDSS DR7 point source 
+catalogue (table \emph{PhotoObjAll}), in regions of 1 deg$^2$ 
+chosen for contiguous coverage in both IPHAS and SDSS. This limited 
+the choice of comparison regions to the second and third Galactic 
+quadrants as the SDSS strips passing through the early plane are 
+incomplete. DR7 was chosen for this comparison due to a change in 
+the catalogue generation for DR8 which resulted in fewer stars being 
+detected near bright ($r\,<\,10$) sources, which in turn gave an unduly 
+pessimistic values for the reliability of IPHAS sources.
+
+Comparisons were made for a number of subsets of the IPHAS point 
+source catalogue: 
+\begin{enumerate}
+\item the entire (unfiltered) catalogue
+\item objects detected in all three bands with \emph{mergedClass} equals -1, -2 or +1
+\item $rErr\,<\,0.1$ and \emph{rClass} equals -1, -2 or +1
+\item $iErr\,<\,0.1$ and \emph{iClass} equals -1, -2 or +1
+\item \emph{reliable} sources
+\item \emph{veryReliable} sources
+\end{enumerate}
+\noindent with the results presented in Table~\ref{tbl:reliability}. 
+Requiring S/N $>$ 10 eliminates the majority of unreliable detections, 
+as does selecting objects observed in multiple bands. The $i$-band performs 
+marginally better than the $r$-band in terms of reliability, returning a 
+slightly higher ($< 0.5\%$) reliability percentage for detections with S/N 
+$>\,10$. Reliabilities $>99\%$ and $>99.5\%$ are achieved by using the 
+supplied \emph{reliable} and \emph{veryReliable} quality flags.
+
+The majority of unreliable sources in the IPHAS catalogue are 
+located in the vicinity of bright ($r\,\lesssim\,12$) stars, tracing 
+out the area surrounding the saturated source affected by scattered 
+light. These can be eliminated in most cases by requiring detections 
+in multiple bands.
+
+\begin{table}
+    \caption{Reliability values as computed by crossmatching the point 
+             source catalogues of IPHAS and SDSS DR7 for a number of 
+             subsets of IPHAS DR2 across 1 deg$^2$ regions.}
+    \label{tbl:reliability}
+    \begin{center}
+        \begin{tabular}{lcccccc}
+        \toprule
+        Region ($\ell$, $b$) & Raw & nBands == 3 & $rErr\,<\,0.1$ & $iErr\,<\,0.1$ & \emph{reliable} & \emph{veryReliable} \\
+        \midrule
+        (149.39, 4.06) & 92.95 & 98.83 & 98.53 & 98.54 & 99.67 & 99.82 \\
+        (186.59, -2.50) & 89.99 & 98.82 & 98.34 & 98.44 & 99.36 & 99.57 \\
+        (202.70, -1.75) & 89.42 & 98.46 & 97.8 & 98.11 & 99.4 & 99.65 \\
+        \bottomrule
+        \end{tabular}
+    \end{center}
+\end{table}
+
 \subsection{Random and systematic uncertainties}

### 2014-07-16 scripts/release-preparation/augment-catalogue.py
Initial changes due to paper revision

@@ -12,9 +12,16 @@ def augment(filename_origin, filename_target):
     f = fits.open(filename_origin)
 
     for i in np.arange(1, 99, 1):  # Loop over all columns
+        name = f[1].header['TTYPE{0}'.format(i)]
+
+        # Rename reliable/veryReliable into a10/a10point
+        if name == 'reliable':
+            f[1].header['TTYPE{0}'.format(i)] = 'a10'
+        if name == 'veryReliable':
+            f[1].header['TTYPE{0}'.format(i)] = 'a10point'
 
         # Set an appropriate TDISP keyword for floating points
-        name = f[1].header['TTYPE{0}'.format(i)]
+        
         if name in ['ra', 'dec', 'l', 'b']:
             f[1].header['TDISP{0}'.format(i)] = 'F9.5'
         if name in ['posErr', 'pStar', 'pGalaxy', 'pNoise', 'pSaturated',
@@ -85,14 +92,11 @@ if __name__ == '__main__':
 
     for l in np.arange(25, 220, 5):
         for part in ['a', 'b']:
-            origin = DR2+'/light-compressed/iphas-dr2-{0}{1}-light.fits.gz'.format(l, part)
-            target = DR2+'/light-augmented/iphas-dr2-{0}{1}-light.fits.gz'.format(l, part)
-            augment(origin, target)
-
-    """
-    for l in np.arange(25, 220, 5):
-        for part in ['a', 'b']:
             origin = DR2+'/full-compressed/iphas-dr2-{0}{1}.fits.gz'.format(l, part)
             target = DR2+'/full-augmented/iphas-dr2-{0}{1}.fits.gz'.format(l, part)
             augment(origin, target)
-    """
+
+            #origin = DR2+'/light-compressed/iphas-dr2-{0}{1}-light.fits.gz'.format(l, part)
+            #target = DR2+'/light-augmented/iphas-dr2-{0}{1}-light.fits.gz'.format(l, part)
+            #augment(origin, target)
+

### 2014-07-16 scripts/release-preparation/augment-image-metadata.py

@@ -1,18 +1,21 @@
+"""Script to create a user-friendly index of IPHAS image meta data.
+"""
 import numpy as np
 from astropy.table import Table
 from astropy.table import Column
 from dr2.constants import IPHASQC
 
-
-t = Table.read('/home/gb/tmp/dl/iphas-images.fits')
+# Index of images found by the DR2 pipeline
+# ie. produced by dr2.images.prepare_images()
+t = Table.read('iphas-images-pipeline.fits')
 
 # Run 376022 on the disk received from CASU is a corrupt file
 t.remove_row(np.argwhere(t['run'] == 376022)[0][0])
 # Run 367744 appeared twice in iphas-qc.fits
 t.remove_rows(np.argwhere(t['run'] == 367744)[4:])
 
-# Add the full URL of the image location
-urldata = ['www.iphas.org/data/images/'+name[0:4]+'/'+name for name in t['filename']]
+# Add the URL of the image location
+urldata = ['http://www.iphas.org/data/images/'+name[0:4]+'/'+name for name in t['filename']]
 url = Column(name='url', data=urldata)
 t.add_column(url, 0)
 t.remove_column('filename')
@@ -21,8 +24,15 @@ t.remove_column('filename')
 runs = np.concatenate((IPHASQC['run_r'], IPHASQC['run_i'], IPHASQC['run_ha']))
 fields = np.concatenate((IPHASQC['id'], IPHASQC['id'], IPHASQC['id']))
 qflags = np.concatenate((IPHASQC['qflag'], IPHASQC['qflag'], IPHASQC['qflag']))
+qcproblems = np.concatenate((IPHASQC['problems'], IPHASQC['problems'], IPHASQC['problems']))
+depth5sig = np.concatenate((IPHASQC['r5sig_judged'],
+                            IPHASQC['i5sig_judged'],
+                            IPHASQC['h5sig_judged']))
+
 field_dict = dict(zip(runs, fields))
 qflag_dict = dict(zip(runs, qflags))
+qcproblems_dict = dict(zip(runs, qcproblems))
+depth5sig_dict = dict(zip(runs, depth5sig))
 
 # Add the IPHAS field number
 field = Column(name='fieldid', data=[field_dict[r] for r in t['run']])
@@ -30,17 +40,41 @@ t.add_column(field)
 
 # Add the DR2 quality grade
[…]

### 2014-07-16 paper/iphas-dr2-rev1.tex
Initial changes due to paper revision

--- 2014-07-02 paper/iphas-dr2-rev1.tex
@@
 incorrectly calibrated fields.
-The calibration errors can exceed 0.1~mag
+The nightly calibration errors can exceed 0.1~mag
 and need to be corrected for 
@@
 The choice of quality criteria will always tension 
-completeness against accuracy.
+completeness {\bf against reliability,
+i.e. the fraction of spurious sources}.
 To aid users we have listed two sets of
@@
 
+%TODO: This is an example of a high-extinction sightline where
+% optical observations will naturally be extinction-limited.
+% IPHAS will naturally count far fewer sources than (N)IR surveys.
+% In the outer-plane the situation is, in fact, reversed,
+% except where there is UKIDSS GPS coverage (i.e. dec > 60).
+% For example, in the anti-centre region where the dust optical depths are low, IPHAS contains more...
+
 \subsection{$r - \ha$ excess as a quantitative measure of H$\alpha$ emission}

### 2014-07-16 scripts/release-preparation/augment-catalogue.py
-

@@ -5,13 +5,14 @@ and add origin information.
 """
 import numpy as np
 from astropy.io import fits
+from astropy import log
 
 
 def augment(filename_origin, filename_target):
-
+    log.info('Opening {0}'.format(filename_origin))
     f = fits.open(filename_origin)
 
-    for i in np.arange(1, 99, 1):  # Loop over all columns
+    for i in np.arange(1, f[1].header['TFIELDS']+1, 1):  # Loop over all columns
 
         # Set an appropriate TDISP keyword for floating points
         name = f[1].header['TTYPE{0}'.format(i)]
@@ -76,7 +77,7 @@ def augment(filename_origin, filename_target):
     f[1].header['COMMENT'] = 'This catalogue is part of IPHAS DR2.'
     f[1].header['COMMENT'] = 'For more information, visit http://www.iphas.org.'
 
-    f.writeto(filename_target, checksum=True, output_verify='exception',
+    f.writeto(filename_target, checksum=True,
               clobber=True)
 
 
@@ -85,14 +86,10 @@ if __name__ == '__main__':
 
     for l in np.arange(25, 220, 5):
         for part in ['a', 'b']:
-            origin = DR2+'/light-compressed/iphas-dr2-{0}{1}-light.fits.gz'.format(l, part)
-            target = DR2+'/light-augmented/iphas-dr2-{0}{1}-light.fits.gz'.format(l, part)
+            origin = DR2+'/light/iphas-dr2-{0:03d}{1}-light.fits'.format(l, part)
+            target = DR2+'/light-augmented/iphas-dr2-{0:03d}{1}-light.fits.gz'.format(l, part)
+            #augment(origin, target)
+            origin = DR2+'/full/iphas-dr2-{0:03d}{1}.fits'.format(l, part)
+            target = DR2+'/full-augmented/iphas-dr2-{0:03d}{1}.fits'.format(l, part)
             augment(origin, target)
 
-    """
-    for l in np.arange(25, 220, 5):
-        for part in ['a', 'b']:
-            origin = DR2+'/full-compressed/iphas-dr2-{0}{1}.fits.gz'.format(l, part)
-            target = DR2+'/full-augmented/iphas-dr2-{0}{1}.fits.gz'.format(l, part)
-            augment(origin, target)
-    """

### 2014-07-16 scripts/release-preparation/augment-catalogue.py
-

@@ -14,14 +14,7 @@ def augment(filename_origin, filename_target):
     for i in np.arange(1, 99, 1):  # Loop over all columns
         name = f[1].header['TTYPE{0}'.format(i)]
 
-        # Rename reliable/veryReliable into a10/a10point
-        if name == 'reliable':
-            f[1].header['TTYPE{0}'.format(i)] = 'a10'
-        if name == 'veryReliable':
-            f[1].header['TTYPE{0}'.format(i)] = 'a10point'
-
-        # Set an appropriate TDISP keyword for floating points
-        
+        # Set an appropriate TDISP keyword for floating points      
         if name in ['ra', 'dec', 'l', 'b']:
             f[1].header['TDISP{0}'.format(i)] = 'F9.5'
         if name in ['posErr', 'pStar', 'pGalaxy', 'pNoise', 'pSaturated',
@@ -90,7 +83,8 @@ def augment(filename_origin, filename_target):
 if __name__ == '__main__':
     DR2 = '/car-data/gb/iphas-dr2-rc6/concatenated'
 
-    for l in np.arange(25, 220, 5):
+    #for l in np.arange(25, 220, 5):
+    for l in [215]:
         for part in ['a', 'b']:
             origin = DR2+'/full-compressed/iphas-dr2-{0}{1}.fits.gz'.format(l, part)
             target = DR2+'/full-augmented/iphas-dr2-{0}{1}.fits.gz'.format(l, part)

### 2014-07-17 paper/figures/magdist/magdist-query.sh
Initial commit of revised magnitude distribution plot

@@ -1,0 +1,11 @@
+# Get magnitude distributions from the database
+PSQL="ionice -c3 nice -n19 psql -p 5433 -t -A -F',' -c"
+
+for band in r i ha
+do
+    echo "Now querying "$band
+    $PSQL "SELECT ROUND(${band}*2)/2 AS magbin, COUNT(*) FROM iphas2 GROUP BY magbin ORDER BY magbin;" > ${band}.csv
+    $PSQL "SELECT ROUND(${band}*2)/2 AS magbin, COUNT(*) FROM iphas2 WHERE a10 GROUP BY magbin ORDER BY magbin;" > ${band}_a10.csv
+    $PSQL "SELECT ROUND(${band}*2)/2 AS magbin, COUNT(*) FROM iphas2 WHERE a10point GROUP BY magbin ORDER BY magbin;" > ${band}_a10point.csv
+done
+

### 2014-07-17 paper/figures/magdist/plot.py

@@ -1,0 +1,34 @@
+"""Plots magnitude distribution."""
+import matplotlib.pyplot as plt
+from astropy.table import Table
+
+for band in ['r']:
+    r = Table.read(band+'.csv', format='ascii.csv')
+    r2 = Table.read(band+'_reliable.csv', format='ascii.csv')
+    r3 = Table.read(band+'_veryreliable.csv', format='ascii.csv')
+
+
+    plt.figure(figsize=(3.5, 2.5))
+
+    plt.bar(r[band]-0.25, r['count'], width=0.5, 
+            facecolor='#cccccc',
+            label='all sources')
+    plt.bar(r2[band]-0.25, r2['count'], width=0.5,
+            facecolor='#888888',
+            label='\emph{a10}')
+    plt.bar(r3[band]-0.25, r3['count'], width=0.5,
+            facecolor='#444444',
+            label='\emph{a10point}')
+
+    plt.legend(loc='upper left', fontsize=9)
+    plt.xlim([12.25, 22.75])
+    plt.xlabel(band+' magnitude')
+    plt.ylabel('Sources')
+
+    plt.gca().xaxis.set_major_locator(plt.MultipleLocator(1.))
+    #plt.gca().yaxis.set_major_locator(y_majorLocator)
+
+    plt.subplots_adjust(0.15, 0.18, 0.98, 0.9)
+
+    plt.savefig('magdist-'+band+'.pdf')
+    plt.close()

### 2014-07-17 dr2/concatenating.py
Forgot haClass column descriptor

@@ -202,6 +202,7 @@ class Concatenator(object):
                              colmeta -desc "N(0,1) stellarness-of-profile statistic." mergedClassStat;
                              colmeta -desc "1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy." rClass;
                              colmeta -desc "1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy." iClass;
+                             colmeta -desc "1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy." haClass;
                              colmeta -desc "Unique r-band detection identifier (run-ccd-detectionnumber)." rDetectionID;
                              colmeta -desc "Unique i-band detection identifier (run-ccd-detectionnumber)." iDetectionID;
                              colmeta -desc "Unique H-alpha detection identifier (run-ccd-detectionnumber)." haDetectionID;

### 2014-07-17 scripts/postgresql/ingest-iphas-into-postgresql.sh
Update table changes for paper rev1

@@ -4,7 +4,7 @@
 # Author: Geert Barentsen
 
 # Configuration
-DATAPATH="/home/gb/tmp/iphas-dr2-rc6/concatenated/full"  # Path to catalogue FITS tables
+DATAPATH="/home/gb/tmp/iphas-dr2-rc6/concatenated/full-augmented"  # Path to catalogue FITS tables
 TABLENAME="iphas2"
 DBNAME="gb"
 DBPORT=5433
@@ -29,7 +29,6 @@ mergedClassStat real, \
 pStar real, \
 pGalaxy real, \
 pNoise real, \
-pSaturated real, \
 rmi real, \
 rmha real, \
 r real, \
@@ -45,7 +44,8 @@ rEll real, \
 rPA real, \
 rClass smallint, \
 rClassStat real, \
-rErrBits smallint, \
+rDeblend boolean, \
+rSaturated boolean, \
 rMJD double precision, \
 rSeeing real, \
 rDetectionID text, \
@@ -64,7 +64,8 @@ iEll real, \
 iPA real, \
 iClass smallint, \
 iClassStat real, \
-iErrBits smallint, \
+iDeblend boolean, \
+iSaturated boolean, \
 iMJD double precision, \
 iSeeing real, \
 iDetectionID text, \
@@ -85,7 +86,8 @@ haEll real, \
 haPA real, \
 haClass smallint, \
 haClassStat real, \
-haErrBits smallint, \
+haDeblend boolean, \
+haSaturated boolean, \
 haMJD double precision, \
 haSeeing real, \
 haDetectionID text, \
@@ -96,10 +98,9 @@ haEta real, \
 brightNeighb boolean, \
 deblend boolean, \
 saturated boolean, \
-errBits smallint, \
 nBands smallint, \
-reliable boolean, \
-veryReliable boolean, \
+a10 boolean, \
+a10point boolean, \
 fieldID text, \
 fieldGrade text, \
 night integer, \
@@ -134,8 +135,8 @@ echo "Creating iphas2_l_idx"
 $PSQL -c "CREATE INDEX iphas2_l_idx ON iphas2(l);"
 echo "Creating iphas2_b_idx"
 $PSQL -c "CREATE INDEX iphas2_b_idx ON iphas2(b);"
-echo "Creating iphas2_reliable_idx"
-$PSQL -c "CREATE INDEX iphas2_reliable_idx ON iphas2(reliable);"
[…]

### 2014-07-18 paper/iphas-dr2-rev1.tex
introduced a10 and a10point

@@
 $^{3}$IGAM, Institute of Physics, University of Graz, Universit\"atsplatz 5, 8010 Graz, Austria.\\
-$^{4}$South African Astronomical Observatory, P.O. Box 9, Observatory, 7935 Cape Town, South Africa.\\
+$^{4}$South African Astronomical Observatory \& SALT Foundation, P.O. Box 9, Observatory, 7935 Cape Town, South Africa.\\
 $^{5}$Afdeling Sterrenkunde, Radboud Universiteit Nijmegen, Faculteit NWI, Postbus 9010, 6500 GL Nijmegen, The Netherlands.\\
@@
 and recommend a series of quality criteria
-to select the most reliable data
+to select accurate data
 from the catalogue.
@@
 The goal of this paper is to present the next release, 
- which supersedes the IDR by providing a global re-calibration
-and taking the coverage up to 92 per cent of the survey area.}
+ which supersedes the IDR by including a global re-calibration
+and by taking the coverage up to 92 per cent of the survey area.}
 In this work we
@@
         Detectors & Four 2048 $\times$ 4100 pixel CCD's \\
-        %Detector field of view & 11 $\times$ 23 arcmin \\
-        Detector scale & 0.33 arcsec/pixel \\        
+        Pixel scale & 0.33 arcsec/pixel \\        
         Filters & $r$, $i$, \ha \\
         Filter properties & See Table~2 \\
+		Magnitude system & Vega \\
         Exposure times & 30~s ($r$), 10~s ($i$), 120~s (\ha) \\
-		Saturation limit & $\sim$13 ($r$), $\sim$12.5 ($i$), $\sim$12.5 (\ha) \\
-        Detection limit (5$\sigma$) & 21.3 ($r$), 20.1 ($i$), 20.3 (\ha) \\
-		Magnitude system & Vega \\
+		Saturation limit & 13 ($r$), 12 ($i$), 12.5 (\ha) \\
+        Detection limit (5$\sigma$, mean) & 21.2 ($r$), 20.0 ($i$), 20.3 (\ha) \\
         PSF FWHM (median) & 1.1\arcsec\ ($r$), 1.0\arcsec\ ($i$), 1.1\arcsec\ (\ha) \\
@@
 how these have been removed or flagged in the final catalogue.
+{ \bf
 The pipelined data set -- comprising 2.5~terabyte of FITS files --
@@
 for the purpose of transforming the raw
-detection tables into a more reliable source catalogue which
+detection tables into a proper source catalogue which
 (i) is quality-controlled,
@@
 (iii) contains user-friendly columns and warning flags.
-{ \bf It is these post-processing steps which distinguish the present release
+It is these post-processing steps which distinguish the present release
 from the IDR, which enforced less stringent quality limits,
@@
 These new processing steps are explained next.}
-%It is these post-processing steps which are explained next.
 
@@
 have been rejected using seven quality criteria,
-which ensure a reliable and homogeneous level of quality
+which ensure a a10 and homogeneous level of quality
 across the data release:
@@
 for inclusion in the catalogue, in order  
-to deliver the most reliable measurement at a single epoch.
+to deliver the most accurate measurement at a single epoch.
 
@@
          brighter than $r<18$ which were
-         classified by the pipeline as `bestQualityAllBand'
+         classified by the pipeline as `a10point'
          (indicating a high-significance point source detection with
@@
 to provide additional constraints in regions not covered by APASS.
-These extra anchors were deemed to have reliable zeropoints 
+These extra anchors were deemed to have accurate zeropoints 
 based on 
@@
 We cross-matched the IPHAS fields against the subset of
-objects marked as reliable stars in the SDSS catalogue\footnote{
+objects marked as a10 stars in the SDSS catalogue\footnote{
 We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
@@
 can be used as a criterion for selecting stellar objects
-with reliable photometry.
+with accurate photometry.
 
@@
 as part of our quality control procedures
-and were found to be reliable.
+and were found to be a10.
 
@@
 Since the principal aim of this data release is to provide 
-reliable photometry at a single epoch,
+accurate photometry at a single epoch,
 we have focused on providing the magnitudes
@@
 selected above were compiled
-into the final 98-column source catalogue
+into the final 99-column source catalogue
 that is described in Appendix~\ref{app:columns}.
@@
 not have had any known bad pixels in the aperture,
-and not have been on the edge of one of the CCDs
-(this information is contained in the \emph{errBits} quality score column, which is used for internal bookkeeping).
+and not have been on the edge of one of the CCDs.
 \end{enumerate}
@@
 
+{\bf
 Like any other photometric survey,
 the majority of the objects in our catalogue
-are faint sources observed near the detection limits;
-55 per cent of the entries in the catalogue
+are faint sources observed near the detection limits,
+e.g. 55 per cent of the entries in the catalogue
 are fainter than $r > 20$.
@@
 recommended quality criteria 
-in Tables~\ref{tab:reliable} and \ref{tab:veryreliable}.
-
-Table~\ref{tab:reliable} specifies
+in Tables~\ref{tab:a10} and \ref{tab:a10point}.
+
+Table~\ref{tab:a10} specifies
 a set of minimum quality criteria
 which should benefit most applications
-which desire reliable colours
+which desire accurate colours
 as well as completeness {\bf down to $\sim$19th magnitude}.
@@
 A total of 86 million sources out of 219 million 
-(39 per cent) satisfy all the criteria listed in Table~\ref{tab:reliable}
-and are hereafter referred to as {\bf `highQualityAllBand'}.
+(39 per cent) satisfy all the criteria listed in Table~\ref{tab:a10}
+and are hereafter referred to as `a10'.
 For convenience, the catalogue contains a boolean column
-named {\bf \emph{highQualityAllBand}} that directly flags these objects.
+named \emph{a10} that directly flags these objects.
 
 \begin{table*}
-\vspace{2cm}
-\caption{Recommended minimum quality criteria 
-for selecting objects with reliable colours 
-from the IPHAS DR2 source catalogue. 
-86~million entries in the catalogue (39\%)
-satisfy all the criteria listed in this table.
-For convenience, these have been flagged in the catalogue
-using the column named \emph{reliable}.}
+\vspace{2.5cm}
+\caption{\bf Recommended quality criteria 
+for selecting objects with accurate colours.
+These criteria serve to identify objects detected 
+at S/N$\,>\,$10 in all three bands without being saturated,
+and with the added requirement that the photometric measurements
+need to be consistent across different aperture sizes.
+The 86~million objects which satisfy these criteria
+have been flagged in the catalogue using the column named \emph{a10}
+(short for ``all-band 10$\sigma$'').}
 \begin{tabular}{p{8cm}lp{6cm}}
@@
   This implicitly requires a detection in all three bands.  \\
-  $r>13$ {\sc and} $i>12$ {\sc and} \ha\,$>12.5$ {\sc and not} \emph{saturated} &
+  {\sc not} \emph{saturated} &
   158 million (72\%) &
-  The brightness must not exceed the nominal saturation limit
-  and the peak pixel height must not exceed 55,000 counts.
-  Again, this implicitly requires a detection in all three bands.
+  The brightness must not exceed the nominal saturation limits.
   \\
@@
   \hline
-  All of the above (flagged as {\bf\emph{reliable}}) &
+  All of the above (flagged as {\bf\emph{a10}}) &
   86 million (39\%) & \\
@@
 \end{tabular}
-\label{tab:reliable}
-
-\vspace{2cm}
-\caption{Additional quality criteria which are recommended
-for applications which require very reliable colours
-at the expense of completeness. 
-For convenience, the sources which satisfy the criteria listed
-in this table have been flagged in the catalogue
-using the column named \emph{veryReliable}.}
+\label{tab:a10}
+
+\vspace{2.5cm}
+\caption{\bf Additional quality criteria which are recommended
+for applications which require objects to be
+single, unconfused point sources with accurate colours.
+The 59 million sources which satisfy these criteria 
+have been flagged using the column named \emph{a10point}.}
 \begin{tabular}{p{8cm}lp{6cm}}
@@
   \hline
-   \emph{reliable} &
+   \emph{a10} &
    86 million (39\%) &
-   The object must satisfy the criteria listed in Table~\ref{tab:reliable}. \\
+   The object must satisfy the criteria for accurate colours listed in Table~\ref{tab:a10}. \\
 
@@
 
-  All of the above (flagged as {\bf\emph{veryReliable}}) &
+  All of the above (flagged as {\bf\emph{a10point}}) &
   59 million (27\%) & \\
@@
 \end{tabular}
-\label{tab:veryreliable}
+\label{tab:a10point}
 \vspace{1cm}
@@
 For applications which require
-a higher standard of reliability at the expense of completeness,
+a higher standard of {\bf accuracy} at the expense of completeness,
 a further set of additional quality criteria
-are suggested in Table~\ref{tab:veryreliable}.
+are suggested in Table~\ref{tab:a10point}.
 These criteria are designed to ensure that
@@
 these stricter criteria 
-and are hereafter referred to as `very reliable'.
+and are hereafter referred to as `a10point'.
 Again, the catalogue contains a boolean column
-named \emph{veryReliable} which flags these objects.
+named \emph{a10point} which flags these objects.
 
@@
     \includegraphics[width=0.5\textwidth]{figures/magdist/magdist-r.pdf} 
-    \caption{r-band magnitude distribution for all objects in the catalogue 
-    (light grey), for objects flagged as \emph{reliable} 
-    according to the criteria set out in Table~\ref{tab:reliable} (grey),
-    and for objects flagged as \emph{veryReliable} 
-    following Table~\ref{tab:veryreliable} (dark grey).
-    The magnitude distributions for $i$ and \ha\
+    \caption{r-band magnitude distribution
+    for all objects in the catalogue (light grey).
+    Overlaid we also show the distribution for objects
+    detected at S/N $>$ 10 in all bands
+    selected according to the quality criteria 
+    described in Table~\ref{tab:a10} (\emph{a10}, grey),
+    and for the set of unconfused 10$\sigma$ point source detections
+    described in Table~\ref{tab:a10point} (\emph{a10point}, dark grey).
+    The distributions for $i$ and \ha\
     look identical, apart from being shifted
@@
 Fig.~\ref{fig:magdist} compares the $r$-band magnitude
-distributions for \emph{reliable}, \emph{veryReliable}
+distributions for \emph{a10}, \emph{a10point}
 and unfiltered objects.
 We find that 81 per cent of the sources 
-are considered \emph{reliable}
-and 54 per cent are \emph{veryReliable}
+are flagged \emph{a10}
+and 54 per cent are flagged \emph{a10point}
 in the magnitude range $13 < r < 19$.
 We will explain below that the
-\emph{veryReliable} category is least complete
+\emph{a10point} category is least complete
 at low Galactic longitudes,
@@
 For example, by raising the allowed photometric uncertainties
-from 0.1~mag to 0.2~mag in Table~\ref{tab:reliable},
+from 0.1~mag to 0.2~mag in Table~\ref{tab:a10},
 42 million candidate sources would be added to the 109 
@@
 Our choice to adopt 0.1~mag as the cut-off uncertainty
-for the \emph{reliable} category is a pragmatic trade-off
+for the \emph{a10} category is a pragmatic trade-off
 which we found to suit many science applications,
@@
 \item $iErr\,<\,0.1$ and \emph{iClass} equals -1, -2 or +1
-\item \emph{reliable} sources
-\item \emph{veryReliable} sources
+\item \emph{a10} sources
+\item \emph{a10point} sources
 \end{enumerate}
@@
 $>\,10$. Reliabilities $>99\%$ and $>99.5\%$ are achieved by using the 
-supplied \emph{reliable} and \emph{veryReliable} quality flags.
+supplied \emph{a10} and \emph{a10point} quality flags.
 
@@
         \toprule
-        Region ($\ell$, $b$) & Raw & nBands == 3 & $rErr\,<\,0.1$ & $iErr\,<\,0.1$ & \emph{reliable} & \emph{veryReliable} \\
+        Region ($\ell$, $b$) & Raw & nBands == 3 & $rErr\,<\,0.1$ & $iErr\,<\,0.1$ & \emph{a10} & \emph{a10point} \\
         \midrule
@@
 \end{table}
+
+}
 
@@
 
-In Fig.~\ref{fig:pairmag_reliable} we show 
+In Fig.~\ref{fig:pairmag_a10} we show 
 a similar comparison between the primary and secondary detections,
 but this time we have only included sources which are flagged
-as \emph{veryReliable} in the catalogue.
+as \emph{a10point} in the catalogue.
 We do not observe an improvement in the average residuals as a function of magnitude,
@@
 (evidenced by the shorter error bars which denote the standard deviation of the absolute residuals).
-We conclude that the \emph{veryReliable} quality criteria are effective
+We conclude that the \emph{a10point} quality criteria are effective
 at reducing the level of the systematic errors,
-while also removing the unreliable data at the bright and faint end.
+while also removing the inaccurate data at the bright and faint end.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/repeatability/repeatability-reliable.pdf}
+\includegraphics[width=0.9\textwidth]{figures/repeatability/repeatability-a10point.pdf}
 \caption{}
@@
          for all sources in the catalogue (panel a)
-         and for the \emph{veryReliable} sources alone (panel b).
+         and for the \emph{a10point} sources alone (panel b).
          Black dots show the mean absolute residuals
@@
          we derived in Fig.~\ref{fig:uncertainties}.
-         We find that the \emph{veryReliable} quality criteria are successful
+         We find that the \emph{a10point} quality criteria are successful
          at removing objects with large residuals.
@@
     (left and right panel)
-    showing sources flagged as \emph{veryReliable}
+    showing sources flagged as \emph{a10point}
     located in an area of 1~deg$^2$
@@
 Each figure contains all the sources
-flagged as \emph{veryReliable} within a region of one square degree 
+flagged as \emph{a10point} within a region of one square degree 
 centred on the coordinates indicated in the diagram
@@
     The diagram shows all objects in the catalogue
-    which have been flagged as \emph{reliable} and are brighter
+    which have been flagged as \emph{a10} and are brighter
     than $r<20$ (grey circles).
@@
 which are brighter than $r<20$
-and have been flagged as \emph{reliable}
+and have been flagged as \emph{a10}
 in IPHAS DR2.
@@
 at longitudes $\ell=30\degr$ to 215\degr.
-The new 98-column catalogue provides single-epoch photometry
+The new 99-column catalogue provides single-epoch photometry
 across 92 per cent of the survey area,
@@
 To support its exploitation, we provide a list of recommended quality criteria
-that will permit the selection of objects with reliable colours from 
+that will permit the selection of objects with accurate colours from 
 the catalogue.  The closing demonstrations highlight the use of the 

### 2014-07-25 paper/referee/response-revision1.txt
Edits for rev1

MN-14-1011-MJ: Authors' response to the reviewer's comments
===========================================================

We are grateful to the referee for the insightful feedback, which has helped us
to improve the presentation of the data products.  In his/her report, the 
referee identified areas where additional information would (i) strengthen the 
role of the paper as a user's guide to the survey, and (ii) would lessen the 
burden on the reader to track down information from previous works.  
We appreciate the merits of the referee's suggestions in this respect, and have 
acted upon them by adding a total of N new pages, 1 new table, and 1 new figure 
to the manuscript.  These changes have been marked in bold in the manuscript.  
We have also added N new sections of auxillary material to our web site as a 
direct consequence.

In what follows we will start by explaining our response to each of the issues 
which the referee identified as high-level remarks (9 items), followed by our 
response to the more specific remarks (8 items).

High-level items
----------------

[Referee]
> 1. This manuscript presents a large amount of information about IPHAS and the preparation
> of and contents of the data products.  It also relies heavily on previous
> publications for general descriptions of the survey, particulars of the instrumentation,
> data processing algorithms and earlier data releases.   As such, it puts considerable burden 
> on the readers to track down information from many different sources.  It is
> understandable that the authors prefer to not duplicate information that is already
> available elsewhere, but it will make it much easier to understand and use the IPHAS data
> products if at least some of the general information  could be repeated in this
> manuscript.
>
> For example, it would be helpful to include concise summary of the IPHAS survey
> parameters and this data release in tabular form.  This summary would serve as
> a "IPHAS-At-A-Glance" for users, and coule include the survey area, dates, telescope,
> camera and survey bandpasses, etc.  Much of the information is presented in the paper,
> but widely spread over the manuscript.  Without at least a guide to finding the
> information, users will be challenged to find what they need.

[Authors' response]
We have kept the referee's recommendation to lessen the burden on the reader in
mind while revising the paper in response to the items to follow below.
As request by the referee, we have included a concise summary table of the
survey parameters at the start of Section 2, in which the 15 key properties
of the survey are summarised (Table 1).

[Referee]
> 2. It was not immediately obvious if the images and catalogue from the Second IPHAS
> release supersede those in the 2008 Initial Data Release.  I believe the
> Second Release includes a reprocessing of all of the survey data to-date, so it
> is intended to replace the Initial Release, but that should be stated explicitly
> in the Introduction (e.g. in the 3rd paragraph).
>
> If my assumption is correct, there should be a summary statement about why the
> Second Release images and catalogue are in fact better products.  For example,
> better calibrations, processing algorithms, etc.

[Authors' response]
The referee correctly identified that we forgot to state that this new release 
supersedes the previous one.  We have added an explicit statement in this sense
to the third paragraph of Section 1, and in addition we have edited the last
paragraph of Section 3.1 to list the key differences between the past and the
present release.

[Referee]
> 3. As stated in Section 7.2, the aim of IPHAS is to enable the discovery of
> emission line objects in the Galactic Plane.  Consequently, I was very surprised
> that no discussion of the sensitivity to line emission (e.g. H-alpha EW) was
> presented in the manuscript.  Perhaps this will be addressed further in the planned
> Barentsen et al paper, but it seems that at least some rudimentary demonstration
> of the sensitivity to line emission is a requirement for this paper.  The
> general photometric sensitivity in the three bands is discussed in section 6.2, and
> an illustration is given how to find emission-line candidates using the color-color
> diagram in Figure 18.  I strongly recommend that at least some examples be given
> of the r-H-alpha color excess as a function of line EW for a sample of objects
> with published spectra.

[Authors' response]
The referee requested a discussion on the sensitivity of the H-alpha 
photometry to line emission to be included in this work.  In response, we have
added one new section (Sect 7.3) and one new figure (Fig 19) to the paper.
In this new section, which is titled "r-Halpha excess as a quantitative measure
of Halpha emission", we compare the spectroscopic H-alpha EWs for a sample of
classical Be stars from Raddi et al (2014) against the position of these
objects in the IPHAS colour-colour diagram.  A good agreement is found.
Moreover, the accompanying discussion offers a brief guide to the previous 
works on this topic.

[Referee]
> 4. Many of the successful large surveys include a list of known caveats concerning
> released data products to help user's avoid pitfalls in their research.  I suggest that
> the authors generate a short list of the most important things about which users
> should be aware when using IPHAS data.  Users appreciate knowing how to avoid mistakes
> due to misinterpretation of known issues, and do not think less of a product because
> of them.

[Authors' response]
In response to the referee's suggestion, we have (i) introduced a summary of
the most common pitfalls at the start of Section 6, and (ii) have added a
caveats section to our data release web page (www.iphas.org/dr2).
We note that it is not possible, at this stage, to include an exhaustive list
of all the caveats, because we cannot anticipate all forms of use of the data.
Hence in Section 6 we alert the reader to the fact that we offer a caveats
section on the website, which will be updated as user experience accumulates.

[Referee]
> 5. In general, the images do not receive the same level of description in the manuscript
> as does the Catalogue.  Any guidance that can be given to users about them would be
> beneficial.  For example, at the very least include a description of the FITS image
> headers, analogous to the descriptions of the catalog columns that is given in
> the Appendix.  They should also mention that tables containing the image metadata are available at IPHAS web site.

[Authors' response]
Although the paper intended to focus on the catalogue product, we agree with the
referee that it would be beneficial to give users additional guidance on the use 
of the image products.  For this reason, we have made significant improvements 
to our web site by (i) providing a query tool to retrieve IPHAS images by 
coordinates, and (ii) provide more comprehensive documentation on the metadata, 
the FITS header keywords, and the flux calibration (see www.iphas.org/images).
Section 6 of the paper has been edited to advertise these resources, which are 
set to be updated in the future whilst we prepare a data release which focuses
more strongly on image and image mosaic products.

[Referee]
> 6. The fact that only 39% of the catalogue entries satisfy the criteria necessary to be
> classified as "reliable" begs the question "just what are the other 61% of objects."
> The authors should add more information about the general reliability of the
> catalogue with and without their criteria. For example, If a users selects
> "reliable" category sources, what should they expect the net reliability to be?
> Certainly not 100%.  How does reliability vary with source magnitude, SNR, or other 
> measurement parameters?  How much does the reliability vary with source environment
> (e.g. crowding, complex backgrounds)?
>
> Reliability can be gauged in a number of ways, such as comparisons with external,
> deeper catalogues (e.g. SDSS, UKIDSS GPS).

[Authors' response]
The referee questioned the nature of the objects which are marked as "reliable"
in the catalogue, and in particular questioned the nature of the large number of 
sources not flagged as "reliable".  Having considered this comment, we have come 
to conclude that we made a poor choice of terminology by adopting the labels 
"reliable" and "veryReliable" as quality qualifiers.
These classes, which are based on stringent quality criteria set out in 
Tables 6 and 7, aim to help users identify high-quality objects, e.g. by 
demanding a 10-sigma detection in all bands.  It goes without saying that 
there are plenty of objects which do not strictly satisfy these criteria but are 
nevertheless genuine astrophysical objects.  The qualifier "reliable" might 
however confer the incorrect impression that such sources are "unreliable".
As a result, we have decided to change the terminology by renaming the class 
previously known as "reliable" into "a10" (which is short for "all-band 10-sigma"), 
and renaming "veryReliable" into "a10point".  We believe this change in 
terminology will better serve as an operational description of the detection quality.

In addition to this change in terminology, we have followed the referee's suggestion
to gauge the reliability of our objects against an external catalogue.  That is,
we have introduced a new Section 6.2, titled "Source reliability", in which the 
reliability of IPHAS sources as a function of the quality criteria is gauged 
by comparing out catalogue against SDSS.  Moreover, we note that a separate paper 
is in an advanced state of preparation (Farnhill et al.), which will investigate 
the issues surrounding the completeness and reliability of our catalogue, 
e.g. by injecting artificial sources into the images and measuring their recovery 
rates, whilst also comparing source counts between SDSS and IPHAS in more detail.

[Referee]
> 7. The authors have done a good job describing the photometric measurements and
> calibration accuracy.  However, there is relatively little discussion of the
> astrometric position reconstruction other than saying it is done relative to
> the 2MASS Point Source Catalogue, and there is no validation of the astrometric
> quality of the Catalogue other than to mention matching success.  Some quantification
> of the astrometric performance should be added (e.g. mean position residuals as a function
> of magnitude from comparisons with the astrometric reference stars, or external
> astrometric collections, such as UCAC4).
>
> Positions in the IPHAS Catalogue have been reconstructed with respect to 2MASS.
> However, the authors do not state if the proper motion of the 2MASS reference stars 
> was taken into account.  Although small, the net motion of the reference stars can
> collectively imprint on the Catalogue because the IPHAS observations were made at
> different epochs than 2MASS, in some cases as much as 12 years later.   
> Failure to do so can imprint the proper motion onto IPHAS, and also because the 
> IPHAS measurements span nearly 10 years (2003-2012), there can be secular variations.

[Authors' response]
astrometry ...
proper motions ...

(There does not appear to be an item number 8 in the reviewer's comments.)

[Referee]
> 9. The authors point out in the introduction that IPHAS will complement a number of
> new infrared surveys of the Galactic Plane.  The paper would benefit from at least
> one cross-comparison with the catalogue of one of those surveys, such as GLIMPSE
> or MIPSGAL, as an illustration of how the combination of surveys will enable
> new science.  At present, the only tie-in to the infrared surveys is the
> visual comparison between IPHAS and GLIMPSE images of one field.

[Authors' response]
In response to the referee's suggestion to include a cross-comparison between
IPHAS and an infrared survey, we have added a brief discussion on the overlap
between the GLIMPSE and IPHAS catalogues at the end of Section 7.2.  The 
discussion warns the reader that an optical survey will naturally be 
extinction-limited.

### 2014-07-25 paper/iphas-dr2-rev1.tex
Edits for rev1

@@
 
-\subsection{Reliability}
+\subsection{Source reliability}
 \label{sec:reliability}
@@
 \begin{table}
-    \caption{Reliability values as computed by crossmatching the point 
-             source catalogues of IPHAS and SDSS DR7 for a number of 
-             subsets of IPHAS DR2 across 1 deg$^2$ regions.}
+    \caption{Fractions of sources in our catalogue
+             which are also found in the SDSS DR7 catalogue
+             within a cross-matching radius of 0.5~arcsec.
+             These statistics can be interpreted
+             as lower limits
+             for the reliability of our sources,
+             i.e. the probability that an entry in our
+             catalogue is a genuine astrophysical source
+             rather than an instrumental artefact.
+             The recovery fractions were computed
+             in three 1 deg$^2$ circular areas
+             where both surveys overlapped.
+             The recovery rates are shown
+             for IPHAS sources 
+             without selection criteria applied (``all''),
+             using only objects detected in all three bands (``$nBands=3$''),
+             using objects detected at 10$\sigma$
+             in all three bands (``a10''),
+             and finally using the detections
+             which we classified earlier
+             as reliable 10$\sigma$ point sources (``a10point'').
+    		%Reliability values as computed by crossmatching the point 
+             %source catalogues of IPHAS and SDSS DR7 for a number of 
+             %subsets of IPHAS DR2 across 1 deg$^2$ regions.
+             }
     \label{tbl:reliability}
@@
         \toprule
-        Region ($\ell$, $b$) & Raw & nBands == 3 & $rErr\,<\,0.1$ & $iErr\,<\,0.1$ & \emph{a10} & \emph{a10point} \\
+        Region ($\ell$, $b$) & all & $nBands=3$ & \emph{a10} & \emph{a10point} \\
         \midrule
-        (149.39, 4.06) & 92.95 & 98.83 & 98.53 & 98.54 & 99.67 & 99.82 \\
-        (186.59, -2.50) & 89.99 & 98.82 & 98.34 & 98.44 & 99.36 & 99.57 \\
-        (202.70, -1.75) & 89.42 & 98.46 & 97.8 & 98.11 & 99.4 & 99.65 \\
+        (149.39, 4.06) & 93.0\% & 98.8\% & 99.7\% & 99.8\% \\
+        (186.59, -2.50) & 90.0\% & 98.8\% & 99.4\% & 99.6\% \\
+        (202.70, -1.75) & 89.4\% & 98.5\% &99.4\% & 99.7\% \\
         \bottomrule
@@
 \caption{}
-\label{fig:pairmag_reliable}
+\label{fig:pairmag_a10}
 \end{subfigure}
@@
 
-Sh 2-82 is one of a large population of poorly-studied star-forming regions
+Sh 2-82 is one of a large population 
+of poorly-studied star-forming regions
 located at low Galactic latitudes,
@@
 on the extinction, distance
-and stellar contents of these unexplored regions.
-
-{\bf Moreover, IPHAS provides the much-needed counterpart
-to the source catalogues being provided by the recent infrared surveys.
-For example, amongst the 1810 high-quality sources considered in Fig.~\ref{fig:emitters},
-1623 objects (90\%) have a counterpart in the Spitzer catalogue,
-and 371 objects (20\%) have a counterpart in the 
-Wide-field Infrared Survey Explorer catalogue (WISE; Cutri et al. 2013),
-both within a matching radius of 2 arcsec.}
-
+and stellar contents of many of such unexplored regions.
+{\bf For example,
+in the region studied above,
+the IPHAS catalogue and the Spitzer/GLIMPSE archive
+have 4798 sources in common
+within a cross-matching radius of 1 arcsec,
+(out of a total of 10\,739 and 11\,321 entries in the
+IPHAS and Spitzer catalogues, respectively).
+This is a favourable statistic,
+given that this is a high-extinction sightline towards the inner plane
+where an optical survey will naturally be extinction limited,
+and IPHAS can be shown not to reach beyond 4-5\,kpc \citep{Sale2014}.
+The situation will be even more favourable
+in the outer-plane, where dust optical depths are low.
+
+%where IPHAS is not limited by extinction
+%This is an example of a high-extinction sightline where
+%optical observations will naturally be extinction-limited.
+%IPHAS will naturally count far fewer sources than (N)IR surveys.
+%In the outer-plane the situation is, in fact, reversed,
+%except where there is UKIDSS GPS coverage (i.e. dec > 60).
+%For example, in the anti-centre region where the dust optical depths are low, IPHAS contains more...
+
+%Amongst the 1810 high-quality sources which we
+%plotted in Fig.~\ref{fig:emitters},
+%no less than 1561 objects (87\%) have a counterpart
+%in the Spitzer GLIMPSE archive
+%within a matching radius of 1 arcsec.
+%In the other direction, out of the 6753 objects
+%which are in the Spitzer/GLIMPSE catalogue for this region,
+%2984 (44\%) have a counterpart in IPHAS.
+
+% are 6753 region considered in Fig.~\ref{fig:emitters}
+%contains 6753 objects in the
+%Spitzer GLIMPSE catalogue
+
+%371 objects (20\%) also have a counterpart in the 
+%Wide-field Infrared Survey Explorer catalogue (WISE; Cutri et al. 2013),
+%both within a matching radius of 2 arcsec.
+
+}
+
+% reliable & r < 20 & (l > 53.55-10/60.) & (l < 53.55+10/60.)                  & (b > 0.01-7.5/60.) & (b < 0.01+7.5/60.)
+% Vizier II/293
 %TODO: This is an example of a high-extinction sightline where
@@
 In addition, the catalogue can be downloaded
-in its entirety from our website (www.iphas.org)
+in its entirety from our website
 as a collection of binary FITS tables,
@@
 of the footprint and comprising 50~gigabyte
-in total.
+in total (see www.iphas.org/dr2).
 
@@
 aperture diameters specified in this work.
-To enable the analysis of diffuse sources,
-our website provides access to the pipeline-processed
-imaging data. These have been updated to include
-a new FITS header keyword, called {\sc photzp},
+{\bf To enable the analysis of diffuse sources,
+our website provides access to the entire set
+of pipeline-processed imaging data,
+the associated metadata,
+and the documentation describing the FITS header keywords
+(see www.iphas.org/images)}.
+The image headers have been updated to include
+a new keyword, called {\sc photzp},
 which contains the re-calibrated zeropoint.
@@
 
-The image repository on our website
-is accompanied by an archive of detection tables
--- one for each image --
-which may be used to retrieve secondary detections that are not included
-in the DR2 source catalogue.
-We warn that this set of original images and detection tables
-includes data that did not pass quality control,
+We warn that the image repository on our website
+includes data that did not pass quality control
 and has not been globally re-calibrated.
-Such data are flagged and must be used with great caution.
+Such data are flagged {\bf in the metadata table}
+and must be used with great caution.
 

### 2014-07-25 paper/tables/columns.tex
Edits for rev1

--- 2014-04-13 paper/tables/columns.tex
@@
-1 & name & string &  & Sexagesimal, equatorial position-based source name in the form: JHHMMSS.ss+DDMMSS.s.
-The full naming convention for IPHAS DR2 sources has the form "IPHAS2 JHHMMSS.ss+DDMMSS.s", where "J" indicates the position is J2000.
-The "IPHAS2" prefix is not included in the column.
+1 & name & string &  & Position-based source name in the sexagesimal form: "JHHMMSS.ss+DDMMSS.s".
+You need to add the prefix "IPHAS2" followed by a whitespace to obtain the official name "IPHAS2 JHHMMSS.ss+DDMMSS.s" (where "J" indicates that the position is J2000 equatorial and "IPHAS2" indicates DR2).
 \\
-2 & ra & double & degrees & J2000 Right Ascension with respect to the 2MASS PSC reference frame (which is consistent with ICRS to within 0.1 arcsec).
-The coordinate given is obtained from the astrometric measurement in the r'-band exposure.
-If the source is undetected in r', then the i' or H$\alpha$-band coordinate is given.
+2 & ra & double & degrees & J2000 Right Ascension with respect to the 2MASS PSC reference frame, which is consistent with ICRS to within 0.1 arcsec.
+The coordinate given is obtained from the astrometric measurement in the r-band exposure.
+If the source is undetected in r, then the i or H$\alpha$-band coordinate is given.
 \\
 3 & dec & double & degrees & J2000 Declination. See comments above. \\
-4 & sourceID & string &  & Unique identification number of the detection. Identical to rDetectionID if the source was detected in the r-band; identical to iDetectionID or haDetectionID otherwise. \\
-5 & posErr & float & arcsec & Astrometric fit error (RMS). Be aware that the error might be significantly larger than the RMS near CCD edges. \\
+4 & sourceID & string &  & Unique identification number of the detection. Identical to rDetectionID if the source was detected in the r-band. Identical to iDetectionID or haDetectionID otherwise. \\
+5 & posErr & float & arcsec & Astrometric root mean square (RMS) residual measured against 2MASS across the CCD in which the source is detected.
+Be aware that the astrometric error for a source near the corner of a CCD may be significantly larger than the RMS statistic.
+\\
 6 & l & double & degrees & Galactic longitude $\ell$ converted from ra/dec (IAU 1958 system). \\
 7 & b & double & degrees & Galactic latitude $b$ converted from ra/dec (IAU 1958 system). \\
-8 & mergedClass & short &  & Image classification flag based on all bands (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). Computed using the UKIDSS scheme. \\
+8 & mergedClass & short &  & Image classification flag based on all bands: 1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated. Computed using the UKIDSS scheme. \\
 9 & mergedClassStat & float &  & Merged N(0,1) stellarness-of-profile statistic. Computed using the UKIDSS scheme. \\
-10 & pStar & float &  & Probability that the source is a star (value between 0 and 1). \\
-11 & pGalaxy & float &  & Probability that the source is a galaxy (value between 0 and 1). \\
-12 & pNoise & float &  & Probability that the source is noise (value between 0 and 1). \\
-13 & pSaturated & float &  & Probability that the source is saturated (value between 0 and 1). \\
-14 & rmi & float & mag & ($r$ - $i$) colour, formed by subtracting columns r and i.
-Included in the catalogue for convenience only.
-To obtain the uncertainty, take the root of the sum of the squares of columns rErr and iErr.
+10 & pStar & float &  & Probability that the source is a point source (value between 0 and 1). \\
+11 & pGalaxy & float &  & Probability that the source is an extended object, such as a galaxy, or a close blend of two point sources (value between 0 and 1). \\
+12 & pNoise & float &  & Probability that the source is noise, e.g. a cosmic ray (value between 0 and 1). \\
+13 & rmi & float & mag & (r - i) colour, formed by subtracting columns r and i.  To obtain the uncertainty, take the root of the sum of the squares of columns rErr and iErr. \\
+14 & rmha & float & mag & (r - Halpha) colour, formed by subtracting columns r and ha. See comments above. \\
+15 & r & float & mag & Default r-band magnitude using the 2.3 arcsec diameter aperture. Calibrated in the Vega system. \\
+16 & rErr & float & mag & Uncertainty for r. Does not include systematic errors. \\
+17 & rPeakMag & float & mag & Alternative r-band magnitude derived from the peak pixel height (i.e. a 0.3x0.3 arcsec square aperture). Calibrated in the Vega system. \\
+18 & rPeakMagErr & float & mag & Uncertainty in rPeakMag. Does not include systematics. \\
+19 & rAperMag1 & float & mag & Alternative r-band magnitude using the 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
+20 & rAperMag1err & float & mag & Uncertainty in rAperMag1. Does not include systematics. \\
+21 & rAperMag3 & float & mag & Alternative r-band magnitude using the 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
+22 & rAperMag3err & float & mag & Uncertainty in rAperMag3. Does not include systematics. \\
+23 & rGauSig & float & pixels & RMS of axes of ellipse fit in r. \\
+24 & rEll & float &  & Ellipticity in the r-band. \\
+25 & rPA & float & degrees & Position angle in the r-band. \\
+26 & rClass & short &  & Discrete image classification flag: 1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated. \\
+27 & rClassStat & float &  & N(0,1) stellarness-of-profile statistic. \\
+28 & rDeblend & boolean &  & True if the source is blended with a nearby neighbour in the r-band.
+Although a deblending procedure is applied when measuring the photometry, the result may be unreliable (colours should not be trusted in particular).
 \\
-15 & rmha & float & mag & ($r$ - H$\alpha$) colour, formed by subtracting columns r and ha. See comments above. \\
-16 & r & float & mag & Default r-band magnitude using a 2.3 arcsec diameter aperture. Calibrated in the Vega system. \\
-17 & rErr & float & mag & Uncertainty for r. Does not include systematic errors. \\
-18 & rPeakMag & float & mag & Alternative r-band magnitude derived from the peak pixel height (i.e. a 0.3x0.3 arcsec square aperture). Calibrated in the Vega system. \\
-19 & rPeakMagErr & float & mag & Uncertainty in rPeakMag. Does not include systematics. \\
-20 & rAperMag1 & float & mag & Alternative r-band magnitude using a 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
-21 & rAperMag1err & float & mag & Uncertainty in rAperMag1. Does not include systematics. \\
-22 & rAperMag3 & float & mag & Alternative r-band magnitude using a 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
-23 & rAperMag3err & float & mag & Uncertainty in rAperMag3. Does not include systematics. \\
-24 & rGauSig & float & pixels & RMS of axes of ellipse fit in r. \\
-25 & rEll & float &  & Ellipticity in the r-band. \\
-26 & rPA & float & degrees & Position angle in the r-band. \\
-27 & rClass & short &  & Discrete image classification flag (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). \\
-28 & rClassStat & float &  & N(0,1) stellarness-of-profile statistic. \\
-29 & rErrBits & short &  & Bitmask used to flag a bright neighbour (1), source blending (2) and saturation (8). \\
+29 & rSaturated & boolean &  & True if the source is too bright to make an accurate measurement in the r-band (e.g.
+peak pixel $>$ 55000 counts).
+The photometry is likely affected by systematic errors.
+\\
 30 & rMJD & double & days & Modified Julian Date at the start of the r-band exposure. \\
@@
 34 & rY & float & pixels & Pixel coordinate of the source in the r-band exposure, in the coordinate system of the CCD. \\
-35 & i & float & mag & Default i-band magnitude using a 2.3 arcsec diameter aperture. Calibrated in the Vega system. \\
+35 & i & float & mag & Default i-band magnitude using the 2.3 arcsec diameter aperture. Calibrated in the Vega system. \\
 36 & iErr & float & mag & Uncertainty for i. Does not include systematic errors. \\
@@
 38 & iPeakMagErr & float & mag & Uncertainty in iPeakMag. Does not include systematics. \\
-39 & iAperMag1 & float & mag & Alternative i-band magnitude using a 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
+39 & iAperMag1 & float & mag & Alternative i-band magnitude using the 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
 40 & iAperMag1err & float & mag & Uncertainty in iAperMag1. Does not include systematics. \\
-41 & iAperMag3 & float & mag & Alternative i-band magnitude using a 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
+41 & iAperMag3 & float & mag & Alternative i-band magnitude using the 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
 42 & iAperMag3err & float & mag & Uncertainty in iAperMag3. Does not include systematics. \\
@@
 45 & iPA & float & degrees & Position angle. \\
-46 & iClass & short &  & Discrete image classification flag (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). \\
+46 & iClass & short &  & Discrete image classification flag: 1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated. \\
 47 & iClassStat & float &  & N(0,1) stellarness-of-profile statistic. \\
-48 & iErrBits & short &  & Bitmask used to flag a bright neighbour (1), source blending (2) and saturation (8). \\
-49 & iMJD & double & days & Modified Julian Date at the start of the single-band exposure. \\
-50 & iSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
-51 & iDetectionID & string &  & Unique identifier of the r-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e.
+48 & iDeblend & boolean &  & True if the source is blended with a nearby neighbour in the i-band. See comments for rDeblend above. \\
+49 & iSaturated & boolean &  & True if the source is too bright to make an accurate measurement in the i-band. See comments for rSaturated above. \\
+50 & iMJD & double & days & Modified Julian Date at the start of the single-band exposure. \\
+51 & iSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
+52 & iDetectionID & string &  & Unique identifier of the i-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e.
 composed of the INT telescope run number, the CCD number and a sequential source detection number.
 \\
-52 & iX & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
-53 & iY & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
-54 & iXi & float & arcsec & Position offset of the i-band detection relative to the ra column. The original i-band coordinates can be obtained by computing (ra+iXi/3600, dec+iEta/3600). \\
-55 & iEta & float & arcsec & Position offset of the i-band detection relative to the dec column. See comments above. \\
-56 & ha & float & mag & Default H-alpha magnitude using a 2.3 arcsec aperture. Calibrated in the Vega system. \\
-57 & haErr & float & mag & Uncertainty for ha. Does not include systematic errors. \\
-58 & haPeakMag & float & mag & Alternative H-alpha magnitude derived from the peak pixel height (i.e. a 0.3x0.3 arcsec square aperture). Calibrated in the Vega system. \\
-59 & haPeakMagErr & float & mag & Uncertainty in haPeakMag. Does not include systematics. \\
-60 & haAperMag1 & float & mag & Alternative H-alpha magnitude using a 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
-61 & haAperMag1err & float & mag & Uncertainty in haAperMag1. Does not include systematics. \\
-62 & haAperMag3 & float & mag & Alternative H-alpha magnitude using a 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
-63 & haAperMag3err & float & mag & Uncertainty in haAperMag3. Does not include systematics. \\
-64 & haGauSig & float & pixels & RMS of axes of ellipse fit. \\
-65 & haEll & float &  & Ellipticity \\
-66 & haPA & float & degrees & Position angle. \\
-67 & haClass & short &  & Discrete image classification flag (1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated). \\
-68 & haClassStat & float &  & N(0,1) stellarness-of-profile statistic. \\
-69 & haErrBits & short &  & Bitmask used to flag a bright neighbour (1), source blending (2) and saturation (8). \\
-70 & haMJD & double & days & Modified Julian Date at the start of the single-band exposure. \\
-71 & haSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
-72 & haDetectionID & string &  & Unique identifier of the r-band detection in the format "$\#$run-$\#$ccd-$\#$number", i.e.
+53 & iX & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
+54 & iY & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
+55 & iXi & float & arcsec & Position offset of the i-band detection relative to the ra column. The original i-band coordinates can be obtained by computing (ra+iXi/3600, dec+iEta/3600). \\
+56 & iEta & float & arcsec & Position offset of the i-band detection relative to the dec column. See comments above. \\
+57 & ha & float & mag & Default H-alpha magnitude using the 2.3 arcsec aperture. Calibrated in the Vega system. \\
+58 & haErr & float & mag & Uncertainty for ha. Does not include systematic errors. \\
+59 & haPeakMag & float & mag & Alternative H-alpha magnitude derived from the peak pixel height (i.e. a 0.3x0.3 arcsec square aperture). Calibrated in the Vega system. \\
+60 & haPeakMagErr & float & mag & Uncertainty in haPeakMag. Does not include systematics. \\
+61 & haAperMag1 & float & mag & Alternative H-alpha magnitude using the 1.2 arcsec diameter aperture. Calibrated in the Vega system. \\
+62 & haAperMag1err & float & mag & Uncertainty in haAperMag1. Does not include systematics. \\
+63 & haAperMag3 & float & mag & Alternative H-alpha magnitude using the 3.3 arcsec diameter aperture. Calibrated in the Vega system. \\
+64 & haAperMag3err & float & mag & Uncertainty in haAperMag3. Does not include systematics. \\
+65 & haGauSig & float & pixels & RMS of axes of ellipse fit. \\
+66 & haEll & float &  & Ellipticity. \\
+67 & haPA & float & degrees & Position angle. \\
+68 & haClass & short &  & Discrete image classification flag: 1=galaxy, 0=noise, -1=star, -2=probableStar, -3=probableGalaxy, -9=saturated. \\
+69 & haClassStat & float &  & N(0,1) stellarness-of-profile statistic. \\
+70 & haDeblend & boolean &  & True if the source is blended with a nearby neighbour in H-alpha. See comments for rDeblend above. \\
+71 & haSaturated & boolean &  & True if the source is too bright to make an accurate measurement in H-alpha. See comments for rSaturated above. \\
+72 & haMJD & double & days & Modified Julian Date at the start of the single-band exposure. \\
+73 & haSeeing & float & arcsec & Average Full Width at Half Maximum (FWHM) of stars in the same CCD frame. \\
+74 & haDetectionID & string &  & Unique identifier of the H-alpha detection in the format "$\#$run-$\#$ccd-$\#$number", i.e.
 composed of the INT telescope run number, the CCD number and a sequential source detection number.
 \\
-73 & haX & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
-74 & haY & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
-75 & haXi & float & arcsec & Position offset of the H-alpha detection relative to the ra column. The original Ha-band coordinates can be obtained by computing (ra+haXi/3600, dec+haEta/3600). \\
-76 & haEta & float & arcsec & Position offset of the H-alpha relative to the ra column. See comments above. \\
-77 & brightNeighb & boolean &  & True if a very bright star is nearby (i.e. errBits $\geq$ 1). This indicates that the source might be spurious, or the photometry unreliable. \\
-78 & deblend & boolean &  & True if the source was blended with a nearby neighbour due to crowding (i.e.
-errBits $\geq$ 2).
-Although a deblending procedure is applied when measuring the photometry, the result may be unreliable (e.g.
-colours should not be trusted).
+75 & haX & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
+76 & haY & float & pixels & Pixel coordinate of the source, in the coordinate system of the CCD. \\
+77 & haXi & float & arcsec & Position offset of the H-alpha detection relative to the ra column. The original Ha-band coordinates can be obtained by computing (ra+haXi/3600, dec+haEta/3600). \\
+78 & haEta & float & arcsec & Position offset of the H-alpha relative to the ra column. See comments above. \\
+79 & brightNeighb & boolean &  & True if a very bright star is nearby (defined as brighter than V$<$4 within 10 arcmin, or brighter than V$<$7 within 5 arcmin).
+Such very bright stars cause scattered light and diffraction spikes, which may add systematic errors to the photometry or even trigger spurious detections.
 \\
-79 & saturated & boolean &  & True if the source is saturated (i.e.
-peak pixel $>$ 55000 counts) in one or more bands (i.e.
-errBits $\geq$ 8).
-The photometry of saturated stars is affected by systematic errors.
+80 & deblend & boolean &  & True if the source is blended with a nearby neighbour in one or more bands.
+Although a deblending procedure is applied when measuring the photometry, the result may be inaccurate and the colours should not be trusted.
 \\
-80 & errBits & short &  & Maximum value of (rErrBits, iErrBits, haErrBits). \\
-81 & nBands & short &  & Number of bands in which the source is detected (equals 1, 2 or 3). \\
-82 & reliable & boolean &  & True if: errBits\,$\leq$\,2 \& nBands\,==\,3 \& r\,$>$\,13 \& i\,$>$\,12 \& ha\,$>$\,12.5 \& rErr\,$<$\,0.1 \& iErr\,$<$\,0.1 \& haErr\,$<$\,0.1 \& (abs(r-rAperMag1)\,$<$\,3*hypot(rErr,rAperMag1Err)+0.03) \& (abs(i-iAperMag1)\,$<$\,3*hypot(iErr,iAperMag1Err)+0.03) \& (abs(ha-haAperMag1)\,$<$\,3*hypot(haErr,haAperMag1Err)+0.03).
+81 & saturated & boolean &  & True if the source is saturated in one or more bands. The photometry of saturated stars is affected by systematic errors. \\
+82 & nBands & short &  & Number of bands in which the source is detected (equals 1, 2 or 3). \\
+83 & a10 & boolean &  & True if the source is detected at S/N\,$>$\,10 in all bands without being saturated, and if the photometric measurements are consistent across different aperture diameters.
+Algebraic condition: (rErr\,$<$\,0.1 \& iErr\,$<$\,0.1 \& haErr\,$<$\,0.1 \& NOT saturated \& (abs(r-rAperMag1)\,$<$\,3*hypot(rErr,rAperMag1Err)+0.03) \& (abs(i-iAperMag1)\,$<$\,3*hypot(iErr,iAperMag1Err)+0.03) \& (abs(ha-haAperMag1)\,$<$\,3*hypot(haErr,haAperMag1Err)+0.03).
 \\
-83 & veryReliable & boolean &  & True if: reliable \& pStar\,$>$\,0.9 \& errBits\,=\,0. \\
-84 & fieldID & string &  & Human-readable IPHAS field number and observing run (e.g. 0001o\_aug2003). \\
-85 & fieldGrade & string &  & Internal quality control score of the field. One of A, B, C or D. \\
-86 & night & integer &  & Night of the observation (YYYYMMDD). Refers to the UT date at the start of the night. \\
-87 & seeing & float & arcsec & Maximum value of (rSeeing, iSeeing, haSeeing). \\
-88 & ccd & short &  & CCD-chip number on the Wide Field Camera (WFC) of the Isaac Newton Telescope (INT). 1, 2, 3 or 4. \\
-89 & nObs & short &  & Number of repeat observations of this source in the survey. \\
-90 & sourceID2 & string &  & SourceID of the object in the partner exposure (if obtained within 10 minutes of the primary detection). \\
-91 & fieldID2 & string &  & FieldID of the partner detection (e.g. 0001o\_aug2003). \\
-92 & r2 & float & mag & r-band magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
-93 & rErr2 & float & mag & Uncertainty for r2. \\
-94 & i2 & float & mag & i-band magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
-95 & iErr2 & float & mag & Uncertainty for i2. \\
-96 & ha2 & float & mag & H-alpha magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
-97 & haErr2 & float & mag & Uncertainty for ha2. \\
-98 & errBits2 & integer &  & Error bitmask for the partner detection.
+84 & a10point & boolean &  & True if both the a10 quality criteria above are satisfied, and if the object looks like a single, unconfused point source.
+Algebraic condition: a10 \& pStar\,$>$\,0.9 \& NOT deblend \& NOT brightNeighb.
+\\
+85 & fieldID & string &  & Survey field identifier (e.g. 0001\_aug2003). \\
+86 & fieldGrade & string &  & Internal quality control score of the field. One of A, B, C or D. \\
+87 & night & integer &  & Night of the observation (YYYYMMDD). Refers to the UT date at the start of the night. \\
+88 & seeing & float & arcsec & Maximum value of rSeeing, iSeeing, or haSeeing. \\
+89 & ccd & short &  & CCD-chip number on the Wide Field Camera (WFC) of the Isaac Newton Telescope (INT). 1, 2, 3 or 4. \\
+90 & nObs & short &  & Number of repeat observations of this source in the survey. A value larger than 1 indicates that the source is unlikely to be spurious. \\
+91 & sourceID2 & string &  & SourceID of the alternative detection of the object in the partner exposure. \\
+92 & fieldID2 & string &  & FieldID of the partner detection (e.g. 0001o\_aug2003). \\
+93 & r2 & float & mag & r-band magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
+94 & rErr2 & float & mag & Uncertainty for r2. \\
+95 & i2 & float & mag & i-band magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
+96 & iErr2 & float & mag & Uncertainty for i2. \\
+97 & ha2 & float & mag & H-alpha magnitude in the dithered partner field, i.e. the dithered repeat measurement obtained within 10 minutes (if available). \\
+98 & haErr2 & float & mag & Uncertainty for ha2. \\
+99 & errBits2 & integer &  & Error bitmask for the partner detection.
 Used to flag a bright neighbour (1), source blending (2), saturation (8), vignetting (64), truncation (128) and bad pixels (32768).
+Be careful if errBits2\,$>$\,0.
 \\

### 2014-07-25 paper/figures/repeatability/plot.py

@@ -45,7 +45,7 @@ plt.gca().xaxis.set_major_locator(majorLocator)
 plt.gca().yaxis.set_major_locator(y_majorLocator)
 
 if reliable:
-    plt.text(0., 1.05, "(b) \emph{veryReliable} sources", fontsize=11,
+    plt.text(0., 1.05, "(b) \emph{a10point} sources", fontsize=11,
              ha='left', va='bottom', transform=ax.transAxes)
 else:
     plt.text(0., 1.05, "(a) all sources", fontsize=11,
@@ -80,7 +80,7 @@ plt.gca().yaxis.set_major_locator(y_majorLocator)
 plt.xlabel('Magnitude')
 
 if reliable:
-    plt.savefig('repeatability-reliable.pdf')
+    plt.savefig('repeatability-a10point.pdf')
 else:
     plt.savefig('repeatability.pdf')
 

### 2014-07-25 scripts/postgresql/ingest-iphas-into-postgresql.sh

@@ -142,3 +142,6 @@ $PSQL -c "CREATE INDEX iphas2_a10_idx ON iphas2(a10);"
 echo "Analyzing"
 $PSQL -c "ANALYZE iphas2;"
 
+# Clean up
+$PSQL -c "VACUUM;"
+

### 2014-07-25 scripts/website/create-dr2-page.py

@@ -31,9 +31,9 @@ def get_filesize(host, url):
     return result
 
 
-args = {'host': 'stri-cluster.herts.ac.uk',
-        'dir_light': '/~gb/iphas-dr2-catalogue/light-compressed',
-        'dir_full': '/~gb/iphas-dr2-catalogue/full-compressed',
+args = {'host': 'www.iphas.org',
+        'dir_light': '/data/dr2/light',
+        'dir_full': '/data/dr2/full',
         'last_update': datetime.datetime.utcnow().strftime("%Y-%m-%d"),
         'columns': columns}
 

### 2014-07-28 paper/referee/response-revision1.txt
minor rev1 update

@@
 In what follows we will start by explaining our response to each of the issues 
-which the referee identified as high-level remarks (9 items), followed by our 
-response to the more specific remarks (8 items).
+which the referee identified as high-level remarks (9 items), followed by the
+referee's additional comments which affect specific sections (8 items).
 
@@
 extinction-limited.
+
+Section-specific remarks
+------------------------
+
+[Referee]
+> 1. In Section 3.1, paragraph 1, sentence 1:
+> "archival" should be "archiving"
+
+[Authors' response]
+This has been corrected.
+
+[Referee]
+> 2. In the second paragraph of section 3.1, it is stated that a deblending routine
+> is used to remove contaminating sources.  Additional information about this routine
+> should be provided.  Are contaminants removed from the sky background measurement,
+> the source measurement aperture, both?
+
+[Referee]
+> 3. At the beginning of the Data Reduction section (section 3), it would be helpful
+> to provide a summary of how data processing for this release differed from that
+> used for the Initial Data Release.
+
+[Referee]
+> 4.a. It is stated in section 3.2 that the exposures were discarded unless they
+> had 5-sigma limiting magnitudes of r>20 or i>19 and H-apha>19 mag. How were the
+> limiting magnitudes for an exposure estimated?
+
+[Referee]
+> 4.b. Section 3.2 also specifies that an ellipticity limit of eps>0.3 was used.
+> What is the justification of this limit?  For example, what is the potential impact on
+> photometry and sensitivity at eps=0.25?
+
+[Referee]
+> 5.a. Were the atmospheric extinction coefficients of 0.09 (r) and 0.05 (i) simply
+> adopted, or was there any attempt to measure them?  Is there an estimate of
+> how much the extinction may have varied with time, either during a night, or
+> seasonally?
+
+[Referee]
+> 5.b. The PERCORR factor in equation 1 is used to correct for "sporadic gain variation."
+> What was the nature of the variations when non-zero?  How was it measured?
+> What is the size of the correction when it was applied?
+
+[Referee]
+> 5.c. The first paragraph of section 4.2 on Global Calibration states that calibration
+> errors  can be up to 0.1 magnitude in reference to the colour-colour diagram shown in
+> Figure 4a. There are outlier populations in that figure that are up to 0.6 mags
+> from the main locus of points.  These seem to be accurately corrected by
+> the recalibration, but the text seems to understate the severity of the
+> systematics.
+
+[Referee]
+> 5.d.  It is stated in section 4.2, paragraph 2 that 2MASS observed six calibration
+> fields each hour.  That is not correct.  2MASS observed two calibration fields each hour.  However, each observation of a field consisted of six scans.
+
+[Referee]
+> 5.e The authors may wish to specify in Section 4.3 that the SDSS DR9 Galactic Plane data
+> that were used to test the IPHAS calibration were specifically from the SEGUE program
+> of SDSS II, and cite the appropriate reference for SEGUE.
+
+[Referee]
+> 6.a. Section 5.1
+> Would it be possible to provide the aperture corrections applied to the aperture
+> photometry for each image along with the image metadata?
+
+> Shouldn't the "saturated" column flag be specified per band?  Surely it must
+> be possible for extremely red objects to be saturated in the i-band exposure
+> but not r-band.
+
+> 6.b. Section 5.2
+> Little information about the specifics of band-merging is provided other than the
+> reference to Taylor 2006 and TMATCHN.  Users will be forced to track down
+> the reference to understand the process.  The authors should consider giving a short
+> summary of the algorithm.  Some important information that should be provided are:
+
+- How is confusion handled in the band merging process (e.g. what happens when
+more than one i-band detection can be matched to an r-band detection?).
+Source confusion in the band-merging process is mentioned in the final
+paragraph of section 6.3.
+
+- Is there any band-filling when a source is not detected in all bands?  If so, what
+is the algorithm used to provide the flux limits in the non-detected bands?
+
+6.c. Section 5.3 - Table 4
+
+It is not clear what is the specific value is of the average magnitude given for each of
+the different band-detection combinations in Table 4?  Wouldn't something like the peak
+of the log N - log M curves be of more interest to provide some sense of the
+completeness and/or depth of the Catalogue?
+
+6.d. Section 5.4
+
+While the authors provide a thorough description of the criteria used for
+Catalogue selection, no reasons were offered for why those criteria were adopted.
+For example, were there requirements for completeness and reliability that were
+being satisfies?  If not, what were the objectives?
+
+In the first paragraph of section 5.4, it is stated the three criteria were
+enforced to limit the size of the catalogue.  Was this the main driver for establishing
+the top-level criteria?  It seems unusual in this day and age of massive catalogues
+that this would be a driver.
+
+As stated earlier in this review, some description should be offered as to the nature of
+the very large number of detections in the catalogue that are not classified as "reliable."
+It would be useful to describe them to the user, and to offer suggestions of
+how to recognize problematic detections, and also ones that may be real.
+
+7. Section 6.1, par. 2, sentence 1
+
+Should the work "accuracy" be "reliability"?
+
+8. Figure 11 - caption
+
+The reference to the left and right panels should presumably be to the top and bottom panels,
+unless the two figures will be arranged differently in the published version.

### 2014-07-28 paper/referee/response-revision1.txt
-

@@
 
+[Authors' response]
+The second paragraph of 3.1 has been edited to explain the steps of the CASU
+object detection pipeline in more detail.
+
 [Referee]
@@
 > used for the Initial Data Release.
+
+[Authors' response]
+We have edited the final paragraph of Section 3.1 to highlight the key differences
+between the present and previous data release.
 
@@
 
+[Authors' response]
+We have added a footnote to Section 3.2 which details how the 5-sigma limiting 
+magnitude is defined and estimated in our work.
+
 [Referee]
@@
 > photometry and sensitivity at eps=0.25?
+
+[Authors' response]
+We have edited the third paragraph in Section 3.2 to explain that the choice
+of our ellipticity limit at eps>0.3 is based on inspecting examples of 
+high-ellipticity exposures, which revealed no evidence for degraded photometry
+up to ellipticities less than 0.3. In fact high ellipticities are correlated 
+with superior seeing, and therefore somewhat deeper frames.
 
@@
 > seasonally?
+
+[Authors' response]
+TBD
 

### 2014-07-28 paper/iphas-dr2-rev1.tex
minor rev1 update
-

@@
 we have been able to benefit from the existing WFS pipeline.
-A detailed description of the processing steps 
+A description of the processing steps 
 is found in \citet{Irwin2001}.
@@
 
-The reduced images are then stored in multi-extension FITS files
-with a primary header describing the characteristics
-(position, filter, exposure time, etc.) 
-and four image extensions 
-corresponding to each of the four CCDs.
-Source detection and characterisation is then carried out 
-using the \textsc{imcore} tool \citep{Irwin1985,Irwin1997}.
-The flux of each source is measured using both
-the peak pixel height
-(i.e. a square 0\farcs 33$\times$0\farcs 33 aperture)
-and a series of circular apertures of increasing diameter 
+%The reduced images are then stored in multi-extension FITS files
+%with a primary header describing the characteristics
+%(position, filter, exposure time, etc.) 
+%and four image extensions 
+%corresponding to each of the four CCDs.
+
+% called \textsc{imcore},
+
+{ \bf
+Source detection and parametrisation is then carried out
+using the standard methods in use by CASU,
+which can be summarised in four steps
+\citep[a discussion on each of these steps
+and related points can be found in][]{Irwin1985,Irwin1997}:
+\begin{enumerate}
+\item The local sky background is estimated
+by computing the sigma-clipped median intensity
+in a grid of 64$\times$64 pixel bins across the image,
+which is robust against contaminating sources.
+Bilinear interpolation is then used to obtain an estimate 
+of the background level at each pixel.
+\item Objects, or blends of objects, are detected
+by identifying groups of 4 or more neighbouring pixels
+in which the intensity exceeds the background level by $1.25\sigma$.
+\item The object are parametrised:
+positions are obtained by computing 
+the intensity-weighted isophotal centre-of-gravity
+of each object,
+while photometry is obtained 
+by measuring the intensity
+in a series of soft-edged circular apertures of increasing diameter 
 (1\farcs 2, 2\farcs 3, 3\farcs 3, 4\farcs 6 and 6\farcs 6).
-
-The local background levels are estimated 
-by computing the sigma-clipped median
-flux in a grid of 64$\times$64 pixels (21\arcsec$\times$21\arcsec)
-across the image,
-which is then interpolated to obtain an estimate 
-of the background level at each pixel.
-These sky levels are subtracted from the aperture photometry and
--- when required --
-a deblending routine is applied to remove
-the contamination from any nearby sources.
-This approach works very well 
-across the vast majority of the survey area.  Nevertheless,
-the Galactic Plane contains crowded regions 
-with large numbers of overlapping sources
-or rapidly spatially-varying nebulosity,
-where aperture photometry will be compromised by frequent blending
-or poor background subtraction.
-In \S\ref{sec:catalogue} we will explain how affected sources are 
-flagged.
+\item Finally, the parametrisation of overlapping objects
+is refined by simultaneous fitting of
+soft-edged apertures to each blend,
+which is essentially identical 
+to multiple "profile" fitting.
+\end{enumerate}
+A discussion on each of these steps
+can be found in \citep{Irwin1985,Irwin1997}.
+We note that the parametrisation of single, unconfused objects
+is naturally more reliable than that of
+objects which had to be deblended.
+In \S\ref{sec:catalogue} we will explain
+how blended sources are flagged in the catalogue.
+
+%The sentence about deblending could be misleading since the objects are
+%detected using our standard deblending algorithm while the aperture flux
+%estimates for overlapping images are computed by simultaneous fitting of
+%soft-edged apertures.  During this multiple "profile" fitting the sky
+%level is based on the interpolated global detector-level fit
+
+%This approach works very well 
+%across the vast majority of the survey area.  Nevertheless,
+%the Galactic Plane contains crowded regions 
+%with large numbers of overlapping sources
+%or rapidly spatially-varying nebulosity,
+%where aperture photometry will be compromised by frequent blending
+%or poor background subtraction. 
+}
 
@@
 (1) \emph{Depth.} 
-We discarded any exposures for which the $5\sigma$ limiting magnitude 
+We discarded any exposures 
+for which the $5\sigma$ limiting magnitude\footnote{\bf
+We defined the $5\sigma$ limiting magnitude
+as the magnitude which a point source would have,
+when its flux in a 2\farcs 3-diameter aperture
+is five times the level of the noise in the sky background.
+The sky noise was estimated using a robust MAD estimator 
+for noise scaled to equivalent Gaussian rms value, ie. = MAD x 1.48,
+after removing large scale sky background variations.
+(MAD = Median of the Absolute Deviations about the median.)}
 was brighter than 20th magnitude in the $r$-band
@@
 due to mechanical glitches in the telescope tracking system.
-To exclude these, we rejected exposures
-in which the \emph{average} ellipticity
-across the detectors exceeded $e > 0.3$.
-We have also found that, as this threshold is breached,
-the photometric measurements delivered by the pipeline become degraded.
+{ \bf As a result, 3 per cent of our images show an
+average ellipticity across the detectors which is worse
+than $e > 0.2$.
+The inspection of these examples revealed no evidence for
+degraded photometry up to ellipticities less than 0.3.
+In fact high ellipticities are correlated with superior seeing,
+and therefore somewhat deeper frames.
+We did however exclude a small number of exposures
+which exceeded $e > 0.3$.}
+
+%To exclude these, we rejected exposures
+%in which the average ellipticity
+%across the detectors exceeded $e > 0.3$.
+%We have also found that, as this threshold is breached,
+%the photometric measurements delivered by the pipeline become degraded.
 

### 2014-07-29 paper/referee/response-revision1.txt
-

@@
 [Authors' response]
-TBD
+The atmospheric extinction coefficients adopted are standard default values 
+for La Palma.  No direct measures or corrections were attempted, because our 
+re-calibration procedure naturally absorbs extinction variations as part of the
+APASS-based zeropoint estimation.
 
@@
 > What is the size of the correction when it was applied?
+
+[Authors' response]
+We have edited the text to clarify that the PERCORR factor is in fact
+used to correct for the slight difference in internal gain computed using 
+the twilight flats compared to the much redder typical astronomical object.  
+It is computed by making a robust average of the dark sky levels measured on each
+detector during an observing run.  In other words its a few% level update
+on the internal gain values estimated from the flats.
 

### 2014-07-29 paper/iphas-dr2-rev1.tex
-

@@
 { \bf
-Source detection and parametrisation is then carried out
+Object detection and parametrisation is then carried out
 using the standard methods in use by CASU,
@@
 by computing the sigma-clipped median intensity
-in a grid of 64$\times$64 pixel bins across the image,
+in a grid of 64$\times$64 pixel bins across the image.
+Bilinear interpolation is then used to obtain an estimate 
+of the background level at each pixel,
 which is robust against contaminating sources.
-Bilinear interpolation is then used to obtain an estimate 
-of the background level at each pixel.
 \item Objects, or blends of objects, are detected
 by identifying groups of 4 or more neighbouring pixels
-in which the intensity exceeds the background level by $1.25\sigma$.
+in which the intensity exceeds the background level by at least $1.25\sigma$.
 \item The object are parametrised:
@@
 of each object,
-while photometry is obtained 
+and photometry is obtained 
 by measuring the intensity
@@
 (1\farcs 2, 2\farcs 3, 3\farcs 3, 4\farcs 6 and 6\farcs 6).
-\item Finally, the parametrisation of overlapping objects
+Objects are also classified morphologically
+-- stellar, extended or noise --
+based on their intensity as a function of aperture size.
+\item The parametrisation of overlapping objects
 is refined by simultaneous fitting of
 soft-edged apertures to each blend,
-which is essentially identical 
-to multiple "profile" fitting.
+effectively carrying out profile fitting where necessary.
+We note that the parametrisation of these blended objects is naturally
+less reliable than that of single, unconfused sources,
+but they are clearly flagged in the catalogue
+(to be explained in \S\ref{sec:catalogue}).
 \end{enumerate}
-A discussion on each of these steps
-can be found in \citep{Irwin1985,Irwin1997}.
-We note that the parametrisation of single, unconfused objects
-is naturally more reliable than that of
-objects which had to be deblended.
-In \S\ref{sec:catalogue} we will explain
-how blended sources are flagged in the catalogue.
-
-%The sentence about deblending could be misleading since the objects are
-%detected using our standard deblending algorithm while the aperture flux
-%estimates for overlapping images are computed by simultaneous fitting of
-%soft-edged apertures.  During this multiple "profile" fitting the sky
-%level is based on the interpolated global detector-level fit
 
@@
 %or poor background subtraction.
-}
-
-Finally, an astrometric solution is determined
-based on the Two-Micron All Sky Survey (2MASS) point source catalogue \citep{Skrutskie2006},
+
+Having carried out object detection and characterisation,
+an astrometric solution is then fitted
+based on the Two-Micron All Sky Survey (2MASS)
+point source catalogue \citep{Skrutskie2006},
 which itself is calibrated 
 in the International Celestial Reference System (ICRS).
-A provisional photometric calibration is also provided 
+A provisional photometric calibration is also determined 
 based on the average zeropoint
-determined from sets of standard stars observed within the same night.
-Sources are classified morphologically
--- stellar, extended or noise --
-based on the curve-of-growth determined
-from measuring the source intensity in a series of growing apertures.
-Finally, the resulting source detection tables are stored 
-in multi-extension FITS files.
+obtained from sets of standard stars observed within the same night.
 
 At the time of preparing DR2,
-the CASU pipeline had processed
+the pipeline had processed
 74,195 single-band IPHAS exposures 
 in which a total of 1.9~billion \emph{candidate} detections
-were made at the default detection level of 1.25\,$\sigma$.
+were made.
 This total inevitably includes spurious objects, artefacts and
@@
 how these have been removed or flagged in the final catalogue.
-{ \bf
+
 The pipelined data set -- comprising 2.5~terabyte of FITS files --

### 2014-07-29 paper/referee/response-revision1.txt
-

@@
 acted upon them by adding a total of N new pages, 1 new table, and 1 new figure 
-to the manuscript.  These changes have been marked in bold in the manuscript.  
-We have also added N new sections of auxillary material to our web site as a 
-direct consequence.
+to the manuscript (marked in bold).  We have also added N new sections of 
+auxillary material to our web site as a direct consequence of the referee's
+suggestions.
 
@@
 which the referee identified as high-level remarks (9 items), followed by the
-referee's additional comments which affect specific sections (8 items).
+referee's comments on specific sections (8 items).
 
@@
 > manuscript.
->
 > For example, it would be helpful to include concise summary of the IPHAS survey
@@
 [Authors' response]
-We have kept the referee's recommendation to lessen the burden on the reader in
-mind while revising the paper in response to the items to follow below.
-As request by the referee, we have included a concise summary table of the
-survey parameters at the start of Section 2, in which the 15 key properties
-of the survey are summarised (Table 1).
+Whilst revising the manuscript in response to the items that follow below, we
+have kept the referee's recommendation to lessen the burden on the reader in 
+mind.
+We have also added an overview table of the survey parameters at the start of 
+Section 2, in which 15 key properties of the survey are summarised (Table 1).
 
@@
 > in the Introduction (e.g. in the 3rd paragraph).
->
 > If my assumption is correct, there should be a summary statement about why the
@@
 supersedes the previous one.  We have added an explicit statement in this sense
-to the third paragraph of Section 1, and in addition we have edited the last
-paragraph of Section 3.1 to list the key differences between the past and the
-present release.
+to the third paragraph of Section 1.  We have also Section 3.1 to highlight 
+the key differences between the past and the present release (last paragraph).
 
@@
 photometry to line emission to be included in this work.  In response, we have
-added one new section (Sect 7.3) and one new figure (Fig 19) to the paper.
+added one new section (7.3) and one new figure (Fig 19) to the paper.
 In this new section, which is titled "r-Halpha excess as a quantitative measure
@@
 objects in the IPHAS colour-colour diagram.  A good agreement is found.
-Moreover, the accompanying discussion offers a brief guide to the previous 
-works on this topic.
+Moreover, the accompanying discussion offers a guide to the previous works on 
+this topic.
 
@@
 [Authors' response]
-In response to the referee's suggestion, we have (i) introduced a summary of
-the most common pitfalls at the start of Section 6, and (ii) have added a
-caveats section to our data release web page (www.iphas.org/dr2).
-We note that it is not possible, at this stage, to include an exhaustive list
-of all the caveats, because we cannot anticipate all forms of use of the data.
-Hence in Section 6 we alert the reader to the fact that we offer a caveats
-section on the website, which will be updated as user experience accumulates.
+In response to the referee's suggestion, we have added a new Section 6.1
+to alert users to the most common pitfalls.  We have also added a new
+FAQ/Caveats section to our web page (www.iphas.org/dr2), which will be
+updated as user experience accumulates.
 
@@
 coordinates, and (ii) provide more comprehensive documentation on the metadata, 
-the FITS header keywords, and the flux calibration (see www.iphas.org/images).
+the FITS header keywords, the flux calibration, and the astrometric solution
+(see www.iphas.org/images).
 Section 6 of the paper has been edited to advertise these resources, which are 
-set to be updated in the future whilst we prepare a data release which focuses
-more strongly on image and image mosaic products.
+set to be updated in the future.  DR3 is likely to focus more strongly on image
+and image mosaic products, as the survey coverage is approaching 100%.
 
@@
 
-In addition to this change in terminology, we have followed the referee's suggestion
+In addition to this change in terminology, we have adopted the referee's suggestion
 to gauge the reliability of our objects against an external catalogue.  That is,
@@
 [Authors' response]
-We have edited the text to clarify that the PERCORR factor is in fact
-used to correct for the slight difference in internal gain computed using 
-the twilight flats compared to the much redder typical astronomical object.  
+We have edited the text in Section 4.1 to clarify that the PERCORR term is in 
+fact used to correct for the slight difference in internal gain computed using 
+the twilight flats compared to the much redder typical astronomical objects.  
 It is computed by making a robust average of the dark sky levels measured on each
-detector during an observing run.  In other words its a few% level update
-on the internal gain values estimated from the flats.
+detector during an observing run.  The correction is 0.01 +/- 0.01 mag on average
+in the i-band, and 0.00 +/- 0.00 mag on average in r and H-alpha, which is
+now also mentioned in the text.
 
@@
 
+[Authors' response]
+We have edited the first paragraph of Section 4.2 to offer more detail on the
+calibration errors, including the statement that we found the calibration error
+to exceed 0.1 mag in 12 per cent of the fields, and 0.5 mag in 0.7 per cent.
+
 [Referee]
 > 5.d.  It is stated in section 4.2, paragraph 2 that 2MASS observed six calibration
-> fields each hour.  That is not correct.  2MASS observed two calibration fields each hour.  However, each observation of a field consisted of six scans.
+> fields each hour.  That is not correct.  2MASS observed two calibration fields each hour.  
+> However, each observation of a field consisted of six scans.
+
+[Authors' response]
+We have corrected the paragraph accordingly.  Fortunately, our mistake did not
+affect the main point of the sentence, which is that 2MASS observed standard
+fields far more regularly than IPHAS (which observed standards at intervals of 
+2-3 hours, or even less in the presence of clouds).
 
@@
 
+[Authors' response]
+We have added the appropriate citation for SEGUE (Yanny et al. 2009) to the 
+first paragraph of Section 4.3.
+
 [Referee]
@@
 
+[Authors' response]
+Manpower-permitting, we intend to make recommended aperture corrections
+available on our website later this year, as part of a release of the
+single-band detection catalogues and in preparation of the DR3 release.
+
+[Referee]
 > Shouldn't the "saturated" column flag be specified per band?  Surely it must
@@
 
+[Authors' response]
+The saturation flag was already available for each band as part of the 
+"errBits" columns, but as a result of the referee's comment, we have decided 
+to make this information available in a more user-friendly way through the
+introduction of three new boolean columns to the catalogue: "rSaturated", 
+"iSaturated", and "haSaturated".
+Likewise, we have introduced the new user-friendly columns "rDeblend",
+"iDeblend", and "haDeblend" to flag blended sources on a per-band basis. The
+latter change has allowed the original, more obscure columns "rErrBits",
+"iErrBits", and "haErrBits" to be removed, because the information they
+contained is now entirely captured within the six new columns. Our catalogue
+files and column descriptions have been updated accordingly.
+
+[Referee]
 > 6.b. Section 5.2
@@
 > summary of the algorithm.  Some important information that should be provided are:
+> - How is confusion handled in the band merging process (e.g. what happens when
+> more than one i-band detection can be matched to an r-band detection?).
+> Source confusion in the band-merging process is mentioned in the final
+> paragraph of section 6.3.
 
-- How is confusion handled in the band merging process (e.g. what happens when
-more than one i-band detection can be matched to an r-band detection?).
-Source confusion in the band-merging process is mentioned in the final
-paragraph of section 6.3.
+[Authors' response]
+We have edition Section 5.2 to include a summary of the STILTS TMATCHN 
+algorithm, and to add a footnote which details the exact options that were used 
+to run the algorithm.  We have also included a warning that blended objects can
+occasionally fall victim to source confusion, which is a complicated problem
+that we have not attempted to resolve in this release, other than by means of
+clearly flagging blended objects in the catalogue and allowing users to refine
+the inter-band matching distance.  The last paragraph of Section 5.2 now
+reminds users of this more clearly.
 
-- Is there any band-filling when a source is not detected in all bands?  If so, what
-is the algorithm used to provide the flux limits in the non-detected bands?
+[Referee]
+> - Is there any band-filling when a source is not detected in all bands?  If so, what
+> is the algorithm used to provide the flux limits in the non-detected bands?
 
-6.c. Section 5.3 - Table 4
+[Authors' response]
+We do not provide flux limits in the non-detected bands at present, and have
+made this more clear in the first paragraph of Section 5.2.  We do expect to 
+provide flux limits in DR3, however.
 
-It is not clear what is the specific value is of the average magnitude given for each of
-the different band-detection combinations in Table 4?  Wouldn't something like the peak
-of the log N - log M curves be of more interest to provide some sense of the
-completeness and/or depth of the Catalogue?
+[Referee]
+> 6.c. Section 5.3 - Table 4
+> It is not clear what is the specific value is of the average magnitude given for each of
+> the different band-detection combinations in Table 4?  Wouldn't something like the peak
+> of the log N - log M curves be of more interest to provide some sense of the
+> completeness and/or depth of the Catalogue?
 
-6.d. Section 5.4
+[Authors' response]
+Section 5.4 has been edited to make the role of Table 4 clearer, and the average
+magnitudes have been removed since they indeed did not contribute to the key message
+of the Section.
 
-While the authors provide a thorough description of the criteria used for
-Catalogue selection, no reasons were offered for why those criteria were adopted.
-For example, were there requirements for completeness and reliability that were
-being satisfies?  If not, what were the objectives?
+[Referee]
+> 6.d. Section 5.4
+> While the authors provide a thorough description of the criteria used for
+> Catalogue selection, no reasons were offered for why those criteria were adopted.
+> For example, were there requirements for completeness and reliability that were
+> being satisfies?  If not, what were the objectives?
+> In the first paragraph of section 5.4, it is stated the three criteria were
+> enforced to limit the size of the catalogue.  Was this the main driver for establishing
+> the top-level criteria?  It seems unusual in this day and age of massive catalogues
+> that this would be a driver.
 
-In the first paragraph of section 5.4, it is stated the three criteria were
-enforced to limit the size of the catalogue.  Was this the main driver for establishing
-the top-level criteria?  It seems unusual in this day and age of massive catalogues
-that this would be a driver.
+[Authors' response]
+The main driver for imposing the top-level criteria is to avoid the catalogue
+from being dominated by the breadth of spurious detections which the pipeline
+naturally picks up at S/N < 5.  We have edited the first paragraph of Section 
+5.4 to clarify this point.  For IPHAS DR3, we consider following the example
+set by Spitzer/GLIMPSE to offer both a "highly reliable" as well as a 
+"more complete, less reliable" version of the catalogue to the public.
 
-As stated earlier in this review, some description should be offered as to the nature of
-the very large number of detections in the catalogue that are not classified as "reliable."
-It would be useful to describe them to the user, and to offer suggestions of
-how to recognize problematic detections, and also ones that may be real.
+[Referee]
+> As stated earlier in this review, some description should be offered as to the nature of
+> the very large number of detections in the catalogue that are not classified as "reliable."
+> It would be useful to describe them to the user, and to offer suggestions of
+> how to recognize problematic detections, and also ones that may be real.
 
-7. Section 6.1, par. 2, sentence 1
+[Authors' response]
+At this point we acknowledge again that our choice of the word "reliable" as
+the name of a class of high-quality objects was misleading. As explained in our
+reply to "point 6" above, we believe to have resolved this problem by renaming
+the "reliable" class into the "a10" class, to avoid conveying the incorrect 
+impression that objects not in this class are spurious.
 
-Should the work "accuracy" be "reliability"?
+[Referee]
+> 7. Section 6.1, par. 2, sentence 1
+> Should the work "accuracy" be "reliability"?
 
-8. Figure 11 - caption
+[Authors' response]
+Indeed, we have corrected this mistake.  Once again, we would like to thank
+the referee for pointing us to the correct use of the term "reliable" in this
+context.
 
-The reference to the left and right panels should presumably be to the top and bottom panels,
-unless the two figures will be arranged differently in the published version.
+[Referee]
+> 8. Figure 11 - caption
+> The reference to the left and right panels should presumably be to the top and bottom panels,
+> unless the two figures will be arranged differently in the published version.
+
+[Authors' response]
+We have corrected this mistake.

### 2014-07-29 paper/iphas-dr2-rev1.tex
-

@@
 %corresponding to each of the four CCDs.
-
 % called \textsc{imcore},
@@
 
+%%(even though the nominal rms accuracy
+%-- estimated against 2MASS across the focal plane --
+%tends to be better than 0.1~arcsec).
+
 At the time of preparing DR2,
@@
 (we adopt a 2\farcs 3-diameter circular aperture by default).
-Finally, \textsc{percorr} is a correction based on the difference
-between the median dark sky for a CCD against the median for all the CCDs, 
-and as such is an ancillary correction 
-to account for sporadic gain variations --
-it is zero for most exposures.
-All these quantities correspond to header keywords in the 
-multi-extension FITS files produced by the CASU pipeline.
+{\bf
+Finally, \textsc{percorr} is a term used to correct 
+for the small difference in internal gain
+computed using the twilight flats
+compared to the much redder typical astronomical objects. 
+It is estimated by making a robust average of the dark sky levels 
+measured on each detector during an observing run.
+The correction is $0.01\pm0.01$ on average in $i$ and
+$0.00\pm0.00$ in $r$ and \ha.
+All these quantities correspond to header keywords
+in the FITS files produced by the CASU pipeline. }
 
@@
 for the set of 2\,801 quality-approved fields
-which are located towards the Galactic anti-centre
-($160\degr < \ell < 200\degr$).
+which are located towards the Galactic anti-centre.
+{ \bf
 The main locus of stars is poorly defined
 in the diagram due to the presence of
-incorrectly calibrated fields.
-The nightly calibration errors can exceed 0.1~mag
-and need to be corrected for 
-during a global re-calibration procedure.
+incorrectly calibrated fields,
+which need to be corrected
+during a global re-calibration step.
+The application of such a procedure,
+to be explained below,
+has revealed that the error in our initial nightly calibration
+exceeded 0.1~mag in 12 per cent of fields,
+and even exceeded 0.5~mag in 0.7 per cent.
 For comparison, Fig.~\ref{fig:ccd_after}
-shows the identical diagram
-obtained after re-calibrating.
+demonstrates the improvement 
+in the colour-colour diagram
+after re-calibrating. }
 
@@
 can use the repeat measurements to ensure a homogeneous calibration.
-For example, 2MASS attained its global calibration
-by observing six standard fields each hour, 
+{\bf For example, 2MASS attained its global calibration
+by observing two standard fields each hour, }
 allowing zeropoint variations to be tracked 
@@
 rendering the 2MASS approach 
-of observing standards at a very high frequency
+of observing standards at a high frequency
 prohibitively expensive (it does not help that 
-standard fields 
-are very scarce within the Galactic Plane).
-Second, IPHAS is not specified as a variability survey, with
-the result that stars are not normally observed at more 
-than one epoch,
+standard fields are very scarce within the Galactic Plane).
+Second, IPHAS is not specified as a variability survey,
+with the result that stars are not normally observed
+at more than one epoch,
 unless they happen to fall within a narrow overlap region 
@@
 
+{\bf
 The second step in compiling the source catalogue
@@
 We used the \textsc{tmatchn} function 
-of the \textsc{stilts} software for this purpose,
-which allows rows from multiple tables to be matched \citep{Taylor2006}.
+of the \textsc{stilts} software
+for this purpose,
+which allows rows from multiple tables
+to be matched \citep{Taylor2006}.
+In brief, the algorithm\footnote{The exact options used
+to band-merge a field using \textsc{stilts tmatchn} are: matcher=sky params=1.0 multimode=group nin=3 in1=r
+in2=i in3=halpha join1=always join2=always join3=always}
+identifies groups of linked detections such that
+(i) each detection is located within a specified maximum distance from
+one or more members of the group,
+(ii) each detection appears in exactly one group, and
+(iii) the largest possible groups are preferred
+(i.e. preferably containing three detections from all three bands).
 The result of the procedure is a band-merged catalogue
-in which each row corresponds to a group of linked $r$, $i$, and \ha\ detections
-which satisfy a maximum matching distance criterion in a pair-wise sense.
+in which each row corresponds
+to a group of linked $r$, $i$, and \ha\ detections
+which satisfy the matching distance criterion
+in a pair-wise sense.
 Sources for which no counterpart was identified
-are retained in the catalogue as single-band detections.
-
-We employed a maximum matching distance of 1~arcsec, trading off completeness against reliability.
+in all three bands are retained in the catalogue
+as single- or double-band detections, with the columns
+for the missing bands left empty.
+
+We employed a maximum matching distance of 1~arcsec,
+trading off completeness against reliability.
 On the one hand, a matching distance larger than 1~arcsec 
@@
 where the astrometric solution can 
-show local systematic errors which exceed 0.5~arcsec
-(even though the nominal rms accuracy
--- estimated against 2MASS across the focal plane --
-tends to be better than 0.1~arcsec).
+show local systematic errors which exceed 0.5~arcsec.
 The position offsets between the $r$ detection and detections in $i$, and/or \ha\
@@
 We note that UKIDSS/GPS adopted 
-the same maximum matching distance of 1 arcsec
+the same maximum matching distance
 for similar reasons \citep{Hambly2008}.
 
-The resulting band-merged catalogues were inspected
+The resulting band-merged catalogues were inspected by eye
 as part of our quality control procedures
-and were found to be a10.
+and were found to be reliable for the vast majority of objects.
+We do warn that blended objects
+can occasionally fall victim to source confusion
+during the band-merging procedure,
+which is a complicated problem
+that we have not attempted to resolve in this release.
+It is important to bear this in mind when appraising stars
+of seemingly unusual colours (such as candidate emission line stars).
+If blending is flagged, or if the inter-band matching distance
+is unusually large, then the probability that the unusual colour is 
+spurious is greatly increased.
+}
 
@@
 \emph{rErr2}, \emph{iErr2}, \emph{haErr2}, \emph{errBits2}).
-Second, tables of detections not included in the catalogue
+Second, images not included in the catalogue
 are made available on our website.
@@
 that is described in Appendix~\ref{app:columns}.
+{ \bf
 The original unweeded list of sources naturally included 
 a significant number of spurious entries
-as a result of the very sensitive default detection settings
-that are employed by the CASU pipeline.
-To limit the size of the source catalogue,
-we have decided to enforce three basic criteria
+as a result of the sensitive detection level
+that is employed by the CASU pipeline.
+We have decided to enforce three basic criteria
 which must be met for a candidate source
-to be included in the catalogue:
+to be included in the catalogue:}
 \begin{enumerate}
@@
              detected in more than one field
-             (usually the field pair partner).
-             We also show the average magnitudes for each class.}
+             (usually the field pair partner).}
     \label{tbl:detections}
@@
         \toprule
-        Band(s) & Sources & Confirmed & $\overline{r}$ & $\overline{i}$ & $\overline{\ha}$ \\
-        & [$10^6$] & [\emph{nObs}$>$1] & [mag] & [mag] & [mag] \\
+        Band(s) & Sources & Confirmed \\
+        & [$10^6$] & (\emph{nObs}$>$1) \\
         \midrule
-        $r$, $i$, \ha & 159 & 91\% & 19.3 & 18.2 & 18.8  \\
-        $r$, $i$ & 25 & 77\% & 21.2 & 19.9 & - \\
-        $i$, \ha & 3 & 73\% & - & 19.6 & 20.2 \\
-        $r$, \ha & 2 & 65\% & 20.6 & - & 20.1 \\
-        $i$ & 15 & 43\% & - & 19.8 & - \\
-        $r$ & 9 & 27\% & 21.1 & - & - \\
-        \ha & 6 & 12\% & - & - & 19.9 \\
+        $r$, $i$, \ha & 159 & 91\% \\
+        $r$, $i$ & 25 & 77\% \\
+        $i$, \ha & 3 & 73\% \\
+        $r$, \ha & 2 & 65\% \\
+        $i$ & 15 & 43\% \\
+        $r$ & 9 & 27\% \\
+        \ha & 6 & 12\% \\
         \midrule
-        Total  & 219 & 81\% & 19.6 & 18.6 & 18.9 \\
+        Total  & 219 & 81\% \\
         \bottomrule
@@
 and the remaining 30 million are single-band detections.
-Table~\ref{tbl:detections} also presents the average magnitudes
-and the fraction of `confirmed' objects,
-which we define as those sources which have been detected in more than
-one field
-(recall that the majority of the survey area
+{\bf
+Table~\ref{tbl:detections} also presents 
+the fraction of `confirmed' objects,
+which we define as those sources
+which have been detected more than once
+(recall that much of the survey area
 is observed twice due to the field pair strategy).
-We find that the single-band detections tend to show low confirmation rates,
-which suggests that a significant fraction of these entries may be spurious detections.
-We recommend users not to rely on single-band objects
-without inspecting the image data by eye.
+We find that the single-band detections tend
+to show much lower confirmation rates (32 per cent on average)
+than double- and triple-band detections (89 per cent).
+This suggests that a significant fraction of
+these entries may be spurious detections.
 
@@
 may appear only in \ha.
+Nevertheless, we recommend users not to rely on single-band objects
+without inspecting the image data by eye,
+or verifying that the object
+was detected more than once (\emph{nObs}$>$1).
+}
 
@@
 
+{\bf
+
 We now offer an overview of the properties of the
 catalogue by discussing 
-(i) the recommended quality criteria,
-(ii) the photometric uncertainties and repeatability,
-and (iii) the source densities and the frequency of source blending.
-
-\subsection{Recommended quality criteria}
-\label{sec:qualitycriteria}
-
-{\bf
-Like any other photometric survey,
-the majority of the objects in our catalogue
+(i) the known caveats,
+(ii) the recommended quality criteria,
+(iii) the reliability of sources,
+(iv) the photometric uncertainties,
+and (v) the source densities.
+
+\subsection{Known caveats}
+
+Like any other photometric survey
+which operates close
+to the detection and resolution limits,
+our catalogue inevitable contains sources
+with are spurious or have been parametrised inaccurately.
+In what follows, we highlight the most common
+caveats which users of the catalogue might face,
+followed by a discussion on how they can be avoided.
+Known caveats include:
+\begin{itemize}
+\item \emph{Spurious objects.}
+Nebulous sky backgrounds,
+saturation artefacts near bright stars,
+and cosmic rays
+are known to be able to trigger spurious detections.
+The majority of these can be removed by requiring
+that a detection is made in more than one band
+(\emph{nBands}$>$1),
+on more than one occasion (\emph{nObs}$>$1),
+or by ensuring that the object looks like a perfect
+point source (\emph{pStar}$>$0.9).
+\item \emph{Source blending and confusion.}
+Blended objects are known to be parametrised less accurately
+and to be more prone 
+to have fallen victim to source confusion
+during the band-merging procedure.
+We remind the reader that such objects are flagged in the catalogue
+using the columns \emph{rDeblend}, \emph{iDeblend},
+or \emph{haDeblend}.
+\item \emph{Low S/N detections.}
+The majority of the objects in our catalogue
 are faint sources observed near the detection limits,
@@
 for a faint object.
-Most scientific applications will hence require a set of
-quality criteria to be enforced for the purpose
-of removing lower-quality objects.
-
+These objects can be removed by ensuring that an
+object is detected at $S/N>10$
+(\emph{rErr}$<$0.1, \emph{iErr}$<$0.1, \emph{haErr}$<$0.1)
+or has photometry which is consistent across
+the different aperture sizes.
+\item \emph{Saturation.}
+The photometry and astrometry of objects
+brighter than the saturation limit of the instrument
+is subject to systematic errors
+(typically at $r<13$, $i<12$, $\ha<12.5$).
+Such objects can be removed by ensuring that the columns
+\emph{rSaturated}, \emph{iSaturated},
+or \emph{haSaturated} are set to ``False''.
+\end{itemize}
+We note that it is not possible, at this stage,
+to include an exhaustive list of all the caveats,
+because we cannot anticipate all forms of use of the data.
+For this reason, we recommend users to read
+the caveats section on the webpage (www.iphas.org/dr2),
+which will be updated as user experience accumulates.
+
+\subsection{Recommended quality criteria}
+\label{sec:qualitycriteria}
+
+Most applications will require
+a combination of the quality criteria outlined above
+to remove lower-quality objects.
+In what follows we recommend two sets of such criteria.
 The choice of quality criteria will always tension 
@@
 
-In producing a point source catalogue, spurious detections 
-will inevitably be made. A fraction of these will mimic stellar 
-sources closely enough that they cannot be excluded using the 
-recorded parameters.
-
-%The majority of sources have been observed in IPHAS more than 
-%once (for objects with S/N $>$ 10 in the $r$-band, 62\% of 
-%objects have secondary detections), however there are regions 
-%observed only once - e.g. survey edges, intersections of inter-CCD 
-%gaps. As such, requiring that objects have a secondary observation 
-%to be considered reliable eliminates a significant fraction of
-%genuine objects..
-
-A statistic that can represent the confidence that objects in the 
-catalogue are genuine is the reliability of the survey. The 
-reliability gives the fraction of objects that can be repeatedly 
-observed; provided the comparison is made to an external source of 
-comparable depth this will provide the ratio of genuine astronomical 
-sources to sources in the catalogue.
-
-The IPHAS catalogue was compared to the SDSS DR7 point source 
-catalogue (table \emph{PhotoObjAll}), in regions of 1 deg$^2$ 
-chosen for contiguous coverage in both IPHAS and SDSS. This limited 
-the choice of comparison regions to the second and third Galactic 
-quadrants as the SDSS strips passing through the early plane are 
-incomplete. DR7 was chosen for this comparison due to a change in 
-the catalogue generation for DR8 which resulted in fewer stars being 
+Having obtained a set of quality criteria,
+we now assess how these improve the data quality.
+In this section we start by discussing
+the \emph{reliability} of the survey,
+which we define as the fraction of sources
+which are genuine astrophysical objects.
+To estimate the reliability, we cross-matched the IPHAS catalogue
+against SDSS DR7,
+which is a catalogue of similar resolution and depth.\footnote{DR7 was chosen for this comparison due to a change in 
+the catalogue generation for DR8, which resulted in fewer stars being 
 detected near bright ($r\,<\,10$) sources, which in turn gave an unduly 
-pessimistic values for the reliability of IPHAS sources.
-
-Comparisons were made for a number of subsets of the IPHAS point 
-source catalogue: 
+pessimistic values for the reliability of IPHAS sources.}
+By computing the fraction of IPHAS sources which were
+independently detected by SDSS as well,
+we obtain a lower limit for the reliability.
+
+%Such contiguous areas were only found to exist
+%in the second and third Galactic quadrants.
+% as the SDSS strips passing through the early plane are 
+%incomplete.
+
+%, which has a comparable depth.
+
+%observed; provided the comparison is made to an external source of 
+%comparable depth this will provide the ratio of genuine astronomical 
+%sources to sources in the catalogue.
+
+%In producing a point source catalogue, spurious detections 
+%will inevitably be made. A fraction of these will mimic stellar 
+%sources closely enough that they cannot be excluded using the 
+%recorded parameters.
+
+The comparison was carried out across three cone-shaped regions,
+each covering 1 deg$^2$,
+using a strict cross-matching distance of 0.5~arcsec.
+We considered four subsets of the IPHAS catalogue:
 \begin{enumerate}
-\item the entire (unfiltered) catalogue
-\item objects detected in all three bands with \emph{mergedClass} equals -1, -2 or +1
-\item $rErr\,<\,0.1$ and \emph{rClass} equals -1, -2 or +1
-\item $iErr\,<\,0.1$ and \emph{iClass} equals -1, -2 or +1
-\item \emph{a10} sources
-\item \emph{a10point} sources
+\item the entire (unfiltered) catalogue;
+\item objects detected in all three bands;
+\item \emph{a10} sources;
+\item \emph{a10point} sources.
 \end{enumerate}
-\noindent with the results presented in Table~\ref{tbl:reliability}. 
-Requiring S/N $>$ 10 eliminates the majority of unreliable detections, 
-as does selecting objects observed in multiple bands. The $i$-band performs 
-marginally better than the $r$-band in terms of reliability, returning a 
-slightly higher ($< 0.5\%$) reliability percentage for detections with S/N 
-$>\,10$. Reliabilities $>99\%$ and $>99.5\%$ are achieved by using the 
-supplied \emph{a10} and \emph{a10point} quality flags.
-
-The majority of unreliable sources in the IPHAS catalogue are 
-located in the vicinity of bright ($r\,\lesssim\,12$) stars, tracing 
-out the area surrounding the saturated source affected by scattered 
-light. These can be eliminated in most cases by requiring detections 
-in multiple bands.
+
+The results for each of the regions and subsets
+are presented in Table~\ref{tbl:reliability}.
+We find that our catalogue shows a good
+baseline reliability (90.8 per cent),
+which is improved markedly by the simple requirement
+that a source needs to be detected in all three bands
+(98.7 per cent). 
+Reliabilities $>99\%$ and $>99.5\%$ are achieved using the 
+\emph{a10} and \emph{a10point} quality flags.
+
+To understand the nature of the small fraction of objects
+which appear to be ``unreliable'' in this test,
+despite having been flagged as \emph{a10point},
+we investigated the data by eye.
+We found these ``unreliable'' sources to be located
+in the vicinity of moderately bright stars ($r\,\lesssim\,12$),
+tracing out an area which is affected by saturation spikes or
+scattered light in the SDSS images.
+In all cases these remaining objects appear as genuine
+stars in both the IPHAS and SDSS images,
+and hence we think it is likely that the reliability
+of the \emph{a10point} class of objects is close to 100 per cent.
 
@@
              rather than an instrumental artefact.
-             The recovery fractions were computed
-             in three 1 deg$^2$ circular areas
-             where both surveys overlapped.
-             The recovery rates are shown
+             The fractions were computed
+             in three 1 deg$^2$ regions
+             where both surveys
+             show contiguous overlap.
+             The reliability is shown
              for IPHAS sources 
              without selection criteria applied (``all''),
-             using only objects detected in all three bands (``$nBands=3$''),
+             using only objects detected in all three bands ($nBands=3$),
              using objects detected at 10$\sigma$
-             in all three bands (``a10''),
+             in all three bands (\emph{a10}),
              and finally using the detections
-             which we classified earlier
-             as reliable 10$\sigma$ point sources (``a10point'').
-    		%Reliability values as computed by crossmatching the point 
-             %source catalogues of IPHAS and SDSS DR7 for a number of 
-             %subsets of IPHAS DR2 across 1 deg$^2$ regions.
+             classified as reliable 10$\sigma$ point sources (\emph{a10point}).
              }
@@
         (186.59, -2.50) & 90.0\% & 98.8\% & 99.4\% & 99.6\% \\
-        (202.70, -1.75) & 89.4\% & 98.5\% &99.4\% & 99.7\% \\
+        (202.70, -1.75) & 89.4\% & 98.5\% &99.4\% & 99.7\% \\ \midrule
+        mean & 90.8\% & 98.7\% & 99.5\% & 99.7\% \\
         \bottomrule
@@
 
+We are grateful to the anonymous referee for the insightful feedback,
+which has helped us to improve the presentation of the data products.
+
 \label{lastpage}

### 2014-07-30 paper/referee/response-revision1.txt
Improved astrometry section

@@
 > astrometric collections, such as UCAC4).
->
 > Positions in the IPHAS Catalogue have been reconstructed with respect to 2MASS.
@@
 [Authors' response]
-astrometry ...
-proper motions ...
-
-(There does not appear to be an item number 8 in the reviewer's comments.)
+We have edited Section 3.1 to offer the reader more details on the astrometric 
+calibration.  This section now also includes the results of an external validation of 
+our astrometry against UCAC4, which showed a mean residual of 0.09 +/- 0.07 arcsec at the 
+bright end of our catalogue.  We have also added a statement to specify that
+no proper motion information was used in arriving at our astrometry, which is
+of limited effect because hundreds 2MASS positional reference stars were available 
+in each CCD to fit the parameters of the WCS solution.
 
 [Referee]
+(There does not appear to be an item number 8 in the referee report.)
 > 9. The authors point out in the introduction that IPHAS will complement a number of

### 2014-07-30 paper/iphas-dr2-rev1.tex
Improved astrometry section

@@
 by computing the sigma-clipped median intensity
-in a grid of 64$\times$64 pixel bins across the image.
+in a grid of 64$\times$64 pixel bins across the image,
+which is robust against contaminating sources.
 Bilinear interpolation is then used to obtain an estimate 
-of the background level at each pixel,
-which is robust against contaminating sources.
+of the background level at each pixel.
 \item Objects, or blends of objects, are detected
@@
 soft-edged apertures to each blend,
-effectively carrying out profile fitting where necessary.
+effectively carrying out profile fitting where it is necessary.
 We note that the parametrisation of these blended objects is naturally
 less reliable than that of single, unconfused sources,
-but they are clearly flagged in the catalogue
+which is why they are flagged in the catalogue
 (to be explained in \S\ref{sec:catalogue}).
@@
 Having carried out object detection and characterisation,
-an astrometric solution is then fitted
-based on the Two-Micron All Sky Survey (2MASS)
-point source catalogue \citep{Skrutskie2006},
-which itself is calibrated 
-in the International Celestial Reference System (ICRS).
-A provisional photometric calibration is also determined 
-based on the average zeropoint
-obtained from sets of standard stars observed within the same night.
-
-%%(even though the nominal rms accuracy
-%-- estimated against 2MASS across the focal plane --
-%tends to be better than 0.1~arcsec).
+the astrometric calibration is then determined.
+The solution starts with a rough 
+World Coordinate System (WCS)
+based on the known telescope and camera geometry,
+which uses a Zenithal Polynomial projection
+\citep[ZPN;][]{Calabretta2002}
+to model the field distortion of the WFC.
+The parameters of this solution
+are then progressively refined using the
+Two-Micron All Sky Survey catalogue \citep[2MASS; ][]{Skrutskie2006}.
+We did not include proper motion information during this step,
+because each CCD typically contains
+hundreds of 2MASS reference stars.
+The resulting fit has previously been shown
+to deliver results which are internally consistent
+to better than 0\farcs 1 across the detector
+\citep{Gonzalez-Solares2008}.
+
+An external validation has been carried out
+by comparing our positions against the
+United States Naval Observatory CCD Astrograph Catalog 
+\citep[UCAC4;][]{UCAC4}.
+Fig.~\ref{fig:astrometry} shows the distribution
+of the astrometric residuals
+between IPHAS and UCAC4
+for N million stars in the magnitude
+range $13 < r < 15$,
+which is the range where both surveys overlap
+and where the formal mean error of UCAC4
+is better than 50~mas \citep{UCAC3}.
+We find the mean residual to be $0\farcs 09\pm0\farcs 07$,
+which is satisfactory for our purposes.
+
+\begin{figure}
+    \includegraphics[width=0.5\textwidth]{figures/astrometry/astrometry.pdf} 
+    \caption{Distribution of position residuals
+            between IPHAS DR2 and UCAC4.
+            The residuals were computed for stars
+            in the magnitude range $13 < r < 15$ which were
+            not blended and not saturated.}
+    \label{fig:astrometry}
+\end{figure}
+
+%which itself is calibrated
+%in the International Celestial Reference System (ICRS).
+%(Figure: note that blended stars were excluded).
+
+%A provisional photometric calibration is also determined 
+%based on the average zeropoint
+%obtained from sets of standard stars observed within the same night.
 
@@
 
-The pipelined data set -- comprising 2.5~terabyte of FITS files --
+The entire data set -- comprising 2.5~terabyte of FITS files --
 was then transferred to the University of Hertfordshire
@@
 It is these post-processing steps which distinguish the present release
-from the IDR, which enforced less stringent quality limits,
+from the IDR; which enforced less stringent quality limits,
 did not offer a global calibration,
@@
 as the magnitude which a point source would have,
-when its flux in a 2\farcs 3-diameter aperture
-is five times the level of the noise in the sky background.
+when its flux in the 2\farcs 3-diameter aperture
+equals five times the level of the noise in the sky background.
 The sky noise was estimated using a robust MAD estimator 
-for noise scaled to equivalent Gaussian rms value, ie. = MAD x 1.48,
+for noise scaled to equivalent Gaussian standard deviation,
+ie. = MAD x 1.48,
 after removing large scale sky background variations.

### 2014-07-30 paper/figures/astrometry/astrometry.py

@@ -1,0 +1,19 @@
+import pylab as p
+from astropy.table import Table
+
+
+t = Table.read('smallmatch.fits')
+
+p.figure()
+p.hist(t['angDist'], range=[0.0, 0.5],
+                   bins=50, 
+                   lw=1.0,
+                   color='#dddddd')
+p.minorticks_on()
+                             
+
+p.xlim([0,0.5])
+p.xlabel('{\sc iphas} - {\sc ucac4} position offset [arcsec]')
+p.ylabel('Stars')
+#tight_layout(pad=0.5)
+p.savefig('astrometry.pdf')

### 2014-07-31 paper/figures/astrometry/astrometry.py
Fine-tuned astrometry section

@@ -1,19 +1,14 @@
+"""Plots a histogram showing positional residuals between IPHAS and UCAC4."""
 import pylab as p
 from astropy.table import Table
 
-
-t = Table.read('smallmatch.fits')
+t = Table.read('iphas-x-ucac4.fits')
 
 p.figure()
-p.hist(t['angDist'], range=[0.0, 0.5],
-                   bins=50, 
-                   lw=1.0,
-                   color='#dddddd')
+p.hist(t['Separation']*3600., range=[0.0, 0.5], bins=50, lw=1.0, color='#dddddd')
 p.minorticks_on()
-                             
-
 p.xlim([0,0.5])
-p.xlabel('{\sc iphas} - {\sc ucac4} position offset [arcsec]')
-p.ylabel('Stars')
-#tight_layout(pad=0.5)
+p.xlabel('Astrometric residuals against UCAC4 [arcsec]', fontsize=9)
+p.ylabel('Stars', fontsize=9)
+p.tight_layout(pad=0.5)
 p.savefig('astrometry.pdf')

### 2014-07-31 paper/figures/fieldpair/plot-fieldpair.py
Added fieldpair plot

@@ -1,0 +1,363 @@
+"""Plots the IPHAS survey footprint."""
+import abc
+import numpy as np
+import matplotlib.pyplot as plt
+from astropy import wcs
+from astropy.io import fits
+
+def equ2gal(equ):
+    """Convert single Equatorial J2000d to Galactic coordinates."""
+    import math as m
+    from math import sin, cos, atan, asin, floor
+    
+    ra, dec = equ
+    OB = m.radians(23.4333334);
+    dec = m.radians(dec)
+    ra = m.radians(ra)
+    a = 27.128251 # The RA of the North Galactic Pole
+    d = 192.859481 # The declination of the North Galactic Pole
+    l = 32.931918 # The ascending node of the Galactic plane on the equator
+    sdec = sin(dec)
+    cdec = cos(dec)
+    sa = sin(m.radians(a))
+    ca = cos(m.radians(a))
+
+    GT = asin(cdec * ca * cos(ra - m.radians(d)) + sdec * sa)
+    GL = m.degrees(atan((sdec - sin(GT) * sa) / (cdec * sin(ra - m.radians(d)) * ca)))
+    TP = sdec - sin(GT) * sa
+    BT = cdec * sin(ra - m.radians(d)) * ca
+    if (BT < 0):
+        GL += 180
+    else:
+        if (TP < 0):
+            GL += 360
+           
+    GL += l
+    if (GL > 360):
+        GL -= 360
+   
+    LG = floor(GL)
+    LM = floor((GL - floor(GL)) * 60)
+    LS = ((GL - floor(GL)) * 60 - LM) * 60
+    GT = m.degrees(GT)
+
+    D = abs(GT)
+    if (GT > 0):
+        BG = floor(D)
+    else:
+        BG = -1*floor(D)
+        
+    BM = floor((D - floor(D)) * 60)
+    BS = ((D - floor(D)) * 60 - BM) * 60
+    if (GT < 0):
+        BM = -BM
+        BS = -BS
+    
+    #if GL > 180:
+    #    GL -= 360
+    return [GL,GT]
+    
+def equ2gal_array(equ):
+    """Convert Equatorial J2000d to Galactic coordinates."""
+    out = []
+    for e in equ:
+        out.append( equ2gal(e) )
+    return np.array(out)
+
+
+class Survey(object):
+    """Common base class for photometric surveys."""
+    
+    def __init__(self, name):
+        self.name = name
+    
+    
+    @abc.abstractmethod
[…]

### 2014-07-31 paper/referee/response-revision1.txt
First draft for resubmission

@@
 We appreciate the merits of the referee's suggestions in this respect, and have 
-acted upon them by adding a total of N new pages, 1 new table, and 1 new figure 
-to the manuscript (marked in bold).  We have also added N new sections of 
-auxillary material to our web site as a direct consequence of the referee's
-suggestions.
+acted upon them by adding a total of 4 new pages (3 new sections, 3 new figures,
+and 1 new table) to the manuscript.  Changes have been marked in bold.
+We have also ~doubled the auxillary documentation that is available on our web
+site a direct consequence of the referee's suggestions, and will continue to
+expand the on line documentation as user experience accumulates.
 
@@
 [Authors' response]
-Whilst revising the manuscript in response to the items that follow below, we
-have kept the referee's recommendation to lessen the burden on the reader in 
-mind.
-We have also added an overview table of the survey parameters at the start of 
-Section 2, in which 15 key properties of the survey are summarised (Table 1).
+Whilst revising the manuscript in response to the items that follow below, we have 
+kept the referee's recommendation to lessen the burden on the reader in mind.
+As an immediate response to this specific item, we have extended Section 2 to 
+(i) include an overview table of the survey parameters, summarising 15 key properties 
+of the survey (Table 1), and (ii) included a plot reminding the reader of the footprint
+of a single IPHAS pointing and its offset position (Figure 1).
 
@@
 photometry to line emission to be included in this work.  In response, we have
-added one new section (7.3) and one new figure (Fig 19) to the paper.
+added one new section (7.4) and one new figure (Fig 21) to the paper.
 In this new section, which is titled "r-Halpha excess as a quantitative measure
@@
 In response to the referee's suggestion to include a cross-comparison between
-IPHAS and an infrared survey, we have added a brief discussion on the overlap
-between the GLIMPSE and IPHAS catalogues at the end of Section 7.2.  The 
-discussion warns the reader that an optical survey will naturally be 
-extinction-limited.
+IPHAS and an infrared survey, we have added a new Section 7.3 which provides
+a brief discussion on the overlap between the GLIMPSE and IPHAS catalogues.  
+The discussion reminds the user that an optical survey will naturally be 
+extinction-limited, by showing the sources which IPHAS and Spitzer do not have
+in common as a function of their position on the Spitzer colour-colour diagram.
 

### 2014-07-31 paper/iphas-dr2-rev1.tex
First draft for resubmission

@@
 
-\begin{figure*}
-        \includegraphics[width=1\linewidth]{figures/footprint/footprint_small.png}
-        \caption{Survey area showing the footprints
-        of all the quality-approved IPHAS fields
-        which have been included in this data release.
-        The area covered by each field has been coloured black
-        with a semi-transparent opacity of 20 per cent,
-        such that regions where fields overlap are darker.
-        The IPHAS strategy is to observe each field twice
-        with a small offset,
-        and hence the vast majority of the area 
-        is covered twice (dominant grey colour).
-        There are small overlaps between all the neighbouring fields
-        which can be seen as a honeycomb pattern
-        of dark grey lines across the survey area.
-        Regions with incomplete data are apparent as white gaps (no data) 
-        or in light grey (indicating that one offset is missing).
-        The dark vertical strip near $\ell \simeq$125\degr\ 
-        is an arbitrary consequence of the tiling pattern,
-        which was populated starting from 0h in Right Ascension.}
-        \label{fig:footprint}
-\end{figure*}
-
 \begin{table}
@@
 
+\begin{figure}
+    \includegraphics[width=\linewidth]{figures/fieldpair/fieldpair.pdf} 
+    \caption{
+    Example footprint of a pointing and its offset position.
+    Each field in the survey is accompanied by an offset field
+    at 5\arcmin\ West and South
+	to deal with gaps between the CCDs.
+	This unit of observation is called a \emph{field pair}.
+	The WFC is a mosaic of four CCDs,
+	and hence a field pair is composed of eight CCD frames.
+	In this example, each CCD is plotted in semi-transparent grey
+	to highlight the overlap regions.
+	}
+    \label{fig:fieldpair}
+\end{figure}
+
 The WFC is a mosaic of 4 CCDs 
@@
 and to enable quality checks.
+An example of the footprint of a pointing
+with its offset position is shown in Fig.~\ref{fig:fieldpair}.
 Hence, the basic unit of observation
@@
 were obtained in November 2012.
+
+\begin{figure*}
+        \includegraphics[width=1\linewidth]{figures/footprint/footprint_small.png}
+        \caption{Survey area showing the footprints
+        of all the quality-approved IPHAS fields
+        which have been included in this data release.
+        The area covered by each field has been coloured black
+        with a semi-transparent opacity of 20 per cent,
+        such that regions where fields overlap are darker.
+        The IPHAS strategy is to observe each field twice
+        with a small offset,
+        and hence the vast majority of the area 
+        is covered twice (dominant grey colour).
+        There are small overlaps between all the neighbouring fields
+        which can be seen as a honeycomb pattern
+        of dark grey lines across the survey area.
+        Regions with incomplete data are apparent as white gaps (no data) 
+        or in light grey (indicating that one offset is missing).
+        The dark vertical strip near $\ell \simeq$125\degr\ 
+        is an arbitrary consequence of the tiling pattern,
+        which was populated starting from 0h in Right Ascension.}
+        \label{fig:footprint}
+\end{figure*}
 
@@
 
-%This approach works very well 
-%across the vast majority of the survey area.  Nevertheless,
-%the Galactic Plane contains crowded regions 
-%with large numbers of overlapping sources
-%or rapidly spatially-varying nebulosity,
-%where aperture photometry will be compromised by frequent blending
-%or poor background subtraction.
-
-Having carried out object detection and characterisation,
+Having carried out object detection,
 the astrometric calibration is then determined.
@@
 The parameters of this solution
-are then progressively refined using the
-Two-Micron All Sky Survey catalogue \citep[2MASS; ][]{Skrutskie2006}.
+are then progressively refined by fitting against the
+Two-Micron All Sky Survey \citep[2MASS; ][]{Skrutskie2006}.
 We did not include proper motion information during this step,
@@
 
-An external validation has been carried out
-by comparing our positions against the
+An external validation of the astrometry 
+has been carried out by comparing our positions against the
 United States Naval Observatory CCD Astrograph Catalog 
@@
 Fig.~\ref{fig:astrometry} shows the distribution
-of the astrometric residuals
-between IPHAS and UCAC4
-for N million stars in the magnitude
-range $13 < r < 15$,
+of the position offsets against UCAC4
+computed for 1.3 million stars
+in the magnitude range $13 < r < 15$,
 which is the range where both surveys overlap
@@
 is better than 50~mas \citep{UCAC3}.
-We find the mean residual to be $0\farcs 09\pm0\farcs 07$,
+We find the mean position offset to be $0\farcs 09\pm0\farcs 07$,
 which is satisfactory for our purposes.
+We note that both UCAC4 and 2MASS -- and hence also IPHAS --
+are calibrated in the International Celestial Reference System (ICRS).
+%(Figure: note that blended stars were excluded).
+%We note that the mean residual is  
+%is $-0.006 \pm 0.086$ in R.A.
+%and $0.012\pm0.073$ in Dec.
 
 \begin{figure}
-    \includegraphics[width=0.5\textwidth]{figures/astrometry/astrometry.pdf} 
-    \caption{Distribution of position residuals
-            between IPHAS DR2 and UCAC4.
-            The residuals were computed for stars
-            in the magnitude range $13 < r < 15$ which were
-            not blended and not saturated.}
+    \includegraphics[width=\linewidth]{figures/astrometry/astrometry.pdf} 
+    \caption{Distribution of the distance
+    		between the positions of stars
+    		which appear both in IPHAS DR2 and UCAC4
+    		within a cross-matching distance of 1\arcsec.
+            The residuals were computed for 
+            the 1.3 million stars in the IPHAS catalogue
+            which are not blended, not saturated,
+            and fall in the magnitude range $13 < r < 15$.}
     \label{fig:astrometry}
@@
 They are likely to be YSOs exhibiting a mild infrared excess.
-
-Sh 2-82 is one of a large population 
+The recovery of \ha-emitters in Spitzer
+data illustrates how IPHAS can complement infrared surveys.
+
+\subsection{IPHAS as a complement to infrared data}
+
+{\bf Towards Sh2-82,
+the IPHAS catalogue and the Spitzer/GLIMPSE archive
+have 4798 entries in common,
+out of a total of 10\,739 and 11\,321 entries, respectively.
+1356 objects of these objects were detected in all
+four GLIMPSE bands.
+Fig.~\ref{fig:spitzer-ccd} presents their colour-colour diagram,
+in which we have indicated the 845 objects (62 per cent)
+that have a counterpart in IPHAS (grey circles),
+as well as the 511 objects that don't (red crosses).
+
+\begin{figure}
+    \includegraphics[width=\linewidth]{figures/spitzer-ccd/spitzer-ccd.pdf} 
+    \caption{Spitzer/GLIMPSE colour-colour diagram towards Sh~2-82,
+    showing the objects for which a counterpart exists in IPHAS
+    (845 objects, grey circles) 
+    and those for which no counterpart was identified 
+    (511 objects, red crosses).
+    A cross-matching radius of 1 arcsec
+    was used to identify counterparts.
+    The arrow illustrates the reddening vector corresponding
+    to $A_{Ks}=3$ following the reddening law due to \citet{Flaherty2007}.
+    It is apparent that the most heavily reddened objects
+    tend to have no counterpart.
+    Three heavily-reddened objects without counterpart
+    are located above the diagram but are not shown for clarity. }
+    \label{fig:spitzer-ccd}
+\end{figure}
+
+We find that objects located in the upper half of the diagram,
+where highly-reddened objects are expected to sit,
+are less likely to appear in both surveys.
+This is not surprising,
+because Sh2-82 is located 
+towards a high-extinction sightline 
+near the inner plane
+where an optical survey will naturally be extinction-limited.
+Nevertheless, the extinction map due to \citet{Sale2014}
+has demonstrated that IPHAS can probe stars as far away 
+as 4-5\,kpc towards this region.
+The situation will be even more favourable
+in the outer-plane, where dust optical depths are low.
+
+Finally, we note that Sh 2-82 is one of a large population 
 of poorly-studied star-forming regions
@@
 and stellar contents of many of such unexplored regions.
-{\bf For example,
-in the region studied above,
-the IPHAS catalogue and the Spitzer/GLIMPSE archive
-have 4798 sources in common
-within a cross-matching radius of 1 arcsec,
-(out of a total of 10\,739 and 11\,321 entries in the
-IPHAS and Spitzer catalogues, respectively).
-This is a favourable statistic,
-given that this is a high-extinction sightline towards the inner plane
-where an optical survey will naturally be extinction limited,
-and IPHAS can be shown not to reach beyond 4-5\,kpc \citep{Sale2014}.
-The situation will be even more favourable
-in the outer-plane, where dust optical depths are low.
+}
 
@@
 %both within a matching radius of 2 arcsec.
-
-}
 

### 2014-07-31 paper/referee/response-revision1.txt
First rev1 draft ready

@@
 We have also ~doubled the auxillary documentation that is available on our web
-site a direct consequence of the referee's suggestions, and will continue to
+site as a direct consequence of the referee's suggestions, and will continue to
 expand the on line documentation as user experience accumulates.
@@
 [Authors' response]
-Whilst revising the manuscript in response to the items that follow below, we have 
-kept the referee's recommendation to lessen the burden on the reader in mind.
-As an immediate response to this specific item, we have extended Section 2 to 
-(i) include an overview table of the survey parameters, summarising 15 key properties 
-of the survey (Table 1), and (ii) included a plot reminding the reader of the footprint
-of a single IPHAS pointing and its offset position (Figure 1).
+Whilst revising the manuscript in response to the items that will follow below,
+we have kept the referee's recommendation to lessen the burden on the reader in
+mind.  In immediate response to this comment, we have extended Section 2 to 
+(i) include an overview table of the survey parameters, which summarises 15 key 
+properties of the survey, and (ii) included an extra figure to remind the reader
+of the footprint of an IPHAS field.
 
@@
 supersedes the previous one.  We have added an explicit statement in this sense
-to the third paragraph of Section 1.  We have also Section 3.1 to highlight 
-the key differences between the past and the present release (last paragraph).
+to the third paragraph of Section 1.  We have also edited the last paragraph
+of Section 3.1 to highlight the key differences between the past and the present
+release.
 
@@
 photometry to line emission to be included in this work.  In response, we have
-added one new section (7.4) and one new figure (Fig 21) to the paper.
+added one new section (7.4) and one new figure (Fig 21).
 In this new section, which is titled "r-Halpha excess as a quantitative measure
@@
 [Authors' response]
-Although the paper intended to focus on the catalogue product, we agree with the
+Although the paper does intend to focus on the catalogue product, we agree with the
 referee that it would be beneficial to give users additional guidance on the use 
-of the image products.  For this reason, we have made significant improvements 
-to our web site by (i) providing a query tool to retrieve IPHAS images by 
-coordinates, and (ii) provide more comprehensive documentation on the metadata, 
-the FITS header keywords, the flux calibration, and the astrometric solution
-(see www.iphas.org/images).
-Section 6 of the paper has been edited to advertise these resources, which are 
-set to be updated in the future.  DR3 is likely to focus more strongly on image
-and image mosaic products, as the survey coverage is approaching 100%.
+of the images.  For this reason, we have made significant improvements to our 
+web site by (i) providing a query tool to retrieve IPHAS images by coordinates, 
+and (ii) provide more comprehensive documentation on the metadata, the header 
+keywords, the flux calibration, and the WCS astrometric solution (see www.iphas.org/images).
+Moreover, Section 6 of the paper has been edited to advertise these resources.
+We note that the next release, DR3, intends to focus more strongly on our image
+and image mosaic products, as the survey is approaching ~100% coverage.
 
@@
 demanding a 10-sigma detection in all bands.  It goes without saying that 
-there are plenty of objects which do not strictly satisfy these criteria but are 
+there are plenty of objects which do not satisfy such strict criteria but are 
 nevertheless genuine astrophysical objects.  The qualifier "reliable" might 
@@
 previously known as "reliable" into "a10" (which is short for "all-band 10-sigma"), 
-and renaming "veryReliable" into "a10point".  We believe this change in 
-terminology will better serve as an operational description of the detection quality.
+and renaming "veryReliable" into "a10point" (short for "all-band 10-sigma and
+perfectly point-like").  We believe this change in terminology will better serve 
+as an operational description of the detection quality.
 
@@
 reliability of IPHAS sources as a function of the quality criteria is gauged 
-by comparing out catalogue against SDSS.  Moreover, we note that a separate paper 
-is in an advanced state of preparation (Farnhill et al.), which will investigate 
-the issues surrounding the completeness and reliability of our catalogue, 
-e.g. by injecting artificial sources into the images and measuring their recovery 
-rates, whilst also comparing source counts between SDSS and IPHAS in more detail.
+against SDSS.  Moreover, we note that a separate paper is in an advanced state 
+of preparation (Farnhill et al.), which will investigate the issues surrounding 
+the completeness and reliability of our catalogue, e.g. by injecting artificial 
+sources into the images and measuring their recovery rates, whilst also comparing 
+source counts between SDSS and IPHAS in more detail.
 
@@
 We have edited Section 3.1 to offer the reader more details on the astrometric 
-calibration.  This section now also includes the results of an external validation of 
-our astrometry against UCAC4, which showed a mean residual of 0.09 +/- 0.07 arcsec at the 
-bright end of our catalogue.  We have also added a statement to specify that
-no proper motion information was used in arriving at our astrometry, which is
-of limited effect because hundreds 2MASS positional reference stars were available 
-in each CCD to fit the parameters of the WCS solution.
+calibration.  We have also added the results of an external validation of our 
+astrometry against UCAC4, which shows a mean residual of 0.09 +/- 0.07 arcsec at
+the bright end of our catalogue.  Finally, we have also added a statement to specify
+that no proper motion information was used in arriving at our astrometry, because
+because hundreds of 2MASS positional reference stars were typically available in 
+each CCD to fit the parameters of the WCS solution.
 
@@
 In response to the referee's suggestion to include a cross-comparison between
-IPHAS and an infrared survey, we have added a new Section 7.3 which provides
+IPHAS and an infrared survey, we have added a new section (7.3) which provides
 a brief discussion on the overlap between the GLIMPSE and IPHAS catalogues.  
 The discussion reminds the user that an optical survey will naturally be 
-extinction-limited, by showing the sources which IPHAS and Spitzer do not have
-in common as a function of their position on the Spitzer colour-colour diagram.
+extinction-limited, by showing where sources which IPHAS and GLIMPE have in common
+are positioned in a colour-colour diagram.
 
@@
 We have edited the third paragraph in Section 3.2 to explain that the choice
-of our ellipticity limit at eps>0.3 is based on inspecting examples of 
-high-ellipticity exposures, which revealed no evidence for degraded photometry
-up to ellipticities less than 0.3. In fact high ellipticities are correlated 
-with superior seeing, and therefore somewhat deeper frames.
+of our ellipticity limit is based on inspecting examples of high-ellipticity 
+exposures, which revealed no evidence for degraded photometry up to ellipticities 
+less than 0.3.  In fact we have found high ellipticities to be correlated with 
+superior seeing, and therefore somewhat deeper frames.
 
@@
 The atmospheric extinction coefficients adopted are standard default values 
-for La Palma.  No direct measures or corrections were attempted, because our 
-re-calibration procedure naturally absorbs extinction variations as part of the
+for La Palma.  No direct measures or corrections were attempted, in part because 
+our re-calibration procedure naturally absorbs extinction variations as part of the
 APASS-based zeropoint estimation.
@@
 affect the main point of the sentence, which is that 2MASS observed standard
-fields far more regularly than IPHAS (which observed standards at intervals of 
-2-3 hours, or even less in the presence of clouds).
+fields more regularly than IPHAS.
 
@@
 Manpower-permitting, we intend to make recommended aperture corrections
-available on our website later this year, as part of a release of the
-single-band detection catalogues and in preparation of the DR3 release.
+available on our website later this year, as part of the preparations for the 
+future DR3 release, which intends to focus on the image data more strongly.
 
@@
 Likewise, we have introduced the new user-friendly columns "rDeblend",
-"iDeblend", and "haDeblend" to flag blended sources on a per-band basis. The
-latter change has allowed the original, more obscure columns "rErrBits",
-"iErrBits", and "haErrBits" to be removed, because the information they
-contained is now entirely captured within the six new columns. Our catalogue
-files and column descriptions have been updated accordingly.
+"iDeblend", and "haDeblend" to flag blended sources on a per-band basis.
+These changes have allowed the original, more obscure columns "rErrBits",
+"iErrBits", and "haErrBits" to be removed from the catalogue, because the 
+information they contained is now entirely captured within the six new columns.
+Our catalogue files and column descriptions have been updated accordingly.
 
@@
 [Authors' response]
-We have edition Section 5.2 to include a summary of the STILTS TMATCHN 
-algorithm, and to add a footnote which details the exact options that were used 
-to run the algorithm.  We have also included a warning that blended objects can
-occasionally fall victim to source confusion, which is a complicated problem
-that we have not attempted to resolve in this release, other than by means of
-clearly flagging blended objects in the catalogue and allowing users to refine
-the inter-band matching distance.  The last paragraph of Section 5.2 now
-reminds users of this more clearly.
+We have edition Section 5.2 to include a summary of the STILTS TMATCHN algorithm.
+We have also included a warning that blended objects can occasionally fall victim 
+to source confusion, which is a complicated problem that we have not attempted 
+to resolve in this release, other than by means of clearly flagging blended objects 
+in the catalogue and allowing users to refine the inter-band matching distance.
+The last paragraph of Section 5.2 now reminds users of this more clearly.
 
@@
 [Authors' response]
-We do not provide flux limits in the non-detected bands at present, and have
-made this more clear in the first paragraph of Section 5.2.  We do expect to 
-provide flux limits in DR3, however.
+We do not provide flux limits in the non-detected bands at present.  This has
+now been clarified in the first paragraph of Section 5.2.  We are investigating
+the possibility to introduce flux limits in DR3, however.
 
@@
 [Authors' response]
-Section 5.4 has been edited to make the role of Table 4 clearer, and the average
-magnitudes have been removed since they indeed did not contribute to the key message
-of the Section.
+Section 5.4 has been edited to make the role of Table 4 clearer.  The average
+magnitudes have been removed from this table since they did not contribute to
+the message which the Section attempts to convey.
 
@@
 The main driver for imposing the top-level criteria is to avoid the catalogue
-from being dominated by the breadth of spurious detections which the pipeline
+from being dominated by the wealth of spurious detections which the pipeline
 naturally picks up at S/N < 5.  We have edited the first paragraph of Section 
-5.4 to clarify this point.  For IPHAS DR3, we consider following the example
-set by Spitzer/GLIMPSE to offer both a "highly reliable" as well as a 
-"more complete, less reliable" version of the catalogue to the public.
+5.4 to clarify this point.  For DR3, we consider following the example set by 
+Spitzer/GLIMPSE to offer both a "highly reliable" as well as a  "more complete,
+less reliable" version of the catalogue to the public.
 
@@
 [Authors' response]
-At this point we acknowledge again that our choice of the word "reliable" as
-the name of a class of high-quality objects was misleading. As explained in our
-reply to "point 6" above, we believe to have resolved this problem by renaming
-the "reliable" class into the "a10" class, to avoid conveying the incorrect 
-impression that objects not in this class are spurious.
+As we acknowledged during our response to "point 6", we believe that our choice
+of the word "reliable" as the name of a class of high-quality objects was 
+misleading. We believe to have resolved this problem by renaming the "reliable"
+class into the "a10" class, which avoids conveying the incorrect impression that 
+objects not in this class are spurious by default.
 
@@
 [Authors' response]
-Indeed, we have corrected this mistake.  Once again, we would like to thank
-the referee for pointing us to the correct use of the term "reliable" in this
-context.
+Indeed, we have corrected this mistake. Again, this is a result of our previously
+incorrect usage of the term "reliable".
 
@@
 We have corrected this mistake.
+
+---
+
+Again, we thank the referee for the insightful feedback, which has not only
+helped to benefit the DR2 paper, but will also be taken into account during the
+preparation of DR3.
+
+Kind regards,
+The authors.

### 2014-07-31 paper/iphas-dr2-rev1.tex
First rev1 draft ready

@@
 towards different Galactic sightlines
-and (ii) select \ha\ emission-line objects.
+and (ii) select and quantify \ha\ emission-line objects.
 IPHAS is the first survey to offer comprehensive CCD photometry
@@
 a preliminary catalogue of candidate emission-line objects \citep{Witham2008};
-discoveries of new symbiotic stars \citep{Corradi2008, Corradi2010, Corradi2011}; 
+discoveries of new symbiotic stars \citep{Corradi2008, Corradi2010, Corradi2011, Rodriguez2014}; 
 new cataclysmic variables \citep{Witham2007,Wesson2008,Aungwerojwit2012}; 
@@
 and new and remarkable planetary nebulae 
-\citep{Mampaso2006, Viironen2009a, Viironen2009b, Sabin2010, Corradi2011, Viironen2011}.
+\citep{Mampaso2006, Viironen2009a, Viironen2009b, Sabin2010, Corradi2011, Viironen2011,Sabin2014}.
 
@@
 and the probabilistic inference of stellar properties
-\citep{Sale2009, Sale2010, Giammanco2011, Sale2012, Barentsen2013}. 
+\citep{Sale2009, Sale2010, Giammanco2011, Sale2012, Barentsen2013, Sale2014}. 
 In effect, the availability of narrowband \ha\
@@
 Finally, in \S\ref{sec:dataaccess} we discuss access
-to the catalogue, an online library of reduced images and relevant source code,
+to the catalogue and an online library of reduced images.
 The paper ends with conclusions in \S\ref{sec:conclusions} where we also outline
@@
         Telescope & 2.5m Isaac Newton Telescope \\
-        Instrument & Wide Field Camera \\
+        Instrument & Wide Field Camera (WFC) \\
         Detectors & Four 2048 $\times$ 4100 pixel CCD's \\
@@
 To set the stage for this release,
-we recap some key points {\bf below and in Table~1}.
+we recap some key points in this section {\bf (see Table~1 for a summary)}.
 IPHAS is an imaging survey of the Galactic Plane north of the celestial equator, 
@@
 	to deal with gaps between the CCDs.
-	This unit of observation is called a \emph{field pair}.
+	This unit of observation is called a \emph{field pair},
+	which is observed in all three filters within a span of 10 minutes.
 	The WFC is a mosaic of four CCDs,
 	and hence a field pair is composed of eight CCD frames.
-	In this example, each CCD is plotted in semi-transparent grey
+	In this example, each CCD is plotted as a semi-transparent grey rectangle
 	to highlight the overlap regions.
+	Note that the L-shaped arrangement of the CCD mosaic allows
+	nearly the entire field to be captured using just two pointings,
+	apart from two $\sim10\arcsec$-wide squares
+	which are located where inter-CCD gaps overlap.	
 	}
@@
 and to enable quality checks.
-An example of the footprint of a pointing
-with its offset position is shown in Fig.~\ref{fig:fieldpair}.
+{ \bf An example footprint of a pointing
+and its offset position is shown in Fig.~\ref{fig:fieldpair}.}
 Hence, the basic unit of observation
@@
 30~s ($r$) and 10~s ($i$)
-in the vast majority of the survey observations.\footnote{The $r$-band exposure time was 10~s instead of 30~s in the first months of data taking. Since October 2010, the $i$-band exposure time 
-has been increased from 10~s to 20~s to by-pass a sporadic exposure timing bug that affects the WFC.}
+in the vast majority of the survey observations.\footnote{The $r$-band exposure time was 10s instead of 30s in the first months of data taking. Since October 2010, the $i$-band exposure time 
+has been increased from 10s to 20s to by-pass a sporadic exposure timing bug that affects the WFC.}
 
@@
 for the INT Wide Field imaging Survey \citep[WFS;][]{McMahon2001,Irwin2005},
-which was a 200 deg$^2$ survey programme carried out 
+which was a 200 deg$^2$ {\bf extragalactic} survey programme carried out 
 between 1998 and 2003.
@@
 gain correction and de-fringing.
-
-%The reduced images are then stored in multi-extension FITS files
-%with a primary header describing the characteristics
-%(position, filter, exposure time, etc.) 
-%and four image extensions 
-%corresponding to each of the four CCDs.
-% called \textsc{imcore},
 
@@
 based on the known telescope and camera geometry,
-which uses a Zenithal Polynomial projection
+using a Zenithal Polynomial projection
 \citep[ZPN;][]{Calabretta2002}
-to model the field distortion of the WFC.
+to model the field distortion of the camera.
 The parameters of this solution
@@
 are calibrated in the International Celestial Reference System (ICRS).
-%(Figure: note that blended stars were excluded).
 %We note that the mean residual is  
@@
     \includegraphics[width=\linewidth]{figures/astrometry/astrometry.pdf} 
-    \caption{Distribution of the distance
-    		between the positions of stars
+    \caption{Distribution of the position offsets
+    		of stars
     		which appear both in IPHAS DR2 and UCAC4
@@
             which are not blended, not saturated,
-            and fall in the magnitude range $13 < r < 15$.}
+            and fall in the magnitude range $13 < r < 15$.
+            The mean and standard deviation of this distribution 
+            equals $0\farcs 09\pm0\farcs 07$.}
     \label{fig:astrometry}
 \end{figure}
-
-%which itself is calibrated
-%in the International Celestial Reference System (ICRS).
-%(Figure: note that blended stars were excluded).
-
-%A provisional photometric calibration is also determined 
-%based on the average zeropoint
-%obtained from sets of standard stars observed within the same night.
 
@@
 (iii) contains user-friendly columns and warning flags.
-It is these post-processing steps which distinguish the present release
-from the IDR; which enforced less stringent quality limits,
-did not offer a global calibration,
-and provided a catalogue in which duplicate detections
+It is these post-processing steps which distinguish this release
+from the IDR, which (i) enforced less stringent quality limits,
+(ii) did not offer a global calibration,
+and (iii) provided a catalogue in which duplicate detections
 of unique sources were not flagged.
-These new processing steps are explained next.}
+These improvements are explained next.}
 
@@
 have been rejected using seven quality criteria,
-which ensure a a10 and homogeneous level of quality
+which ensure a good and homogeneous level of quality
 across the data release:
@@
 We defined the $5\sigma$ limiting magnitude
-as the magnitude which a point source would have,
-when its flux in the 2\farcs 3-diameter aperture
-equals five times the level of the noise in the sky background.
+as the magnitude which an imaginary point source would have
+if its flux equalled five times the level of the noise in the sky background.
 The sky noise was estimated using a robust MAD estimator 
@@
 degraded photometry up to ellipticities less than 0.3.
-In fact high ellipticities are correlated with superior seeing,
-and therefore somewhat deeper frames.
+In fact high ellipticities are naturally correlated with superior seeing.
 We did however exclude a small number of exposures
 which exceeded $e > 0.3$.}
-
-%To exclude these, we rejected exposures
-%in which the average ellipticity
-%across the detectors exceeded $e > 0.3$.
-%We have also found that, as this threshold is breached,
-%the photometric measurements delivered by the pipeline become degraded.
 
@@
 and excluded from the release 
-by placing them on a black-list.
+if deemed necessary.
 
@@
 for e.g. time-domain studies of bright stars.
-The images and detection tables of the discarded data are made
-available through our website (www.iphas.org),
+The discarded data is
+available through our website,
 but will be ignored in the remainder of this work.
@@
 It is estimated by making a robust average of the dark sky levels 
-measured on each detector during an observing run.
-The correction is $0.01\pm0.01$ on average in $i$ and
-$0.00\pm0.00$ in $r$ and \ha.
+measured on each detector during an observing run
+(the correction is $0.01\pm0.01$ on average in $i$ and
+$0.00\pm0.00$ in $r$ and \ha).
 All these quantities correspond to header keywords
@@
 in the timing of exposures).
+{ \bf
 This is demonstrated in Fig.~\ref{fig:ccd_before},
 where we show the combined colour-colour diagram
-for the set of 2\,801 quality-approved fields
-which are located towards the Galactic anti-centre.
-{ \bf
+for nearly 3,000 fields across an area of 400~deg$^2$.
 The main locus of stars is poorly defined
@@
 has revealed that the error in our initial nightly calibration
-exceeded 0.1~mag in 12 per cent of fields,
+exceeded 0.1~mag in 12 per cent of the fields,
 and even exceeded 0.5~mag in 0.7 per cent.
@@
          while bins containing more than 400 sources are bright yellow.
-         {\bf The diagrams include all stars
+         The diagrams include all stars
          brighter than $r<18$ which were
@@
          accurate photometry in all bands,
-         to be explained in \S\ref{sec:qualitycriteria})}.
+         to be explained in \S\ref{sec:qualitycriteria}).
          The objects which are seen to fall above the locus of stars
@@
 which all achieved photometry 
-that is globally consistent to within 0.01--0.02~mag.
+that is globally consistent to within 0.01--0.02~mag
+after re-calibration.
 
@@
 allowing zeropoint variations to be tracked 
-over very short timescales \citep{Nikolaev2000}.
+over short timescales \citep{Nikolaev2000}.
 Alternatively, the SDSS and PanSTARRS surveys could benefit
@@
 We cross-matched the IPHAS fields against the subset of
-objects marked as a10 stars in the SDSS catalogue\footnote{
+objects marked as reliable stars in the SDSS catalogue\footnote{
 We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
@@
 to link sources between the three bands.
-We used the \textsc{tmatchn} function 
-of the \textsc{stilts} software
+We used the TMATCHN function 
+of the STILTS software
 for this purpose,
@@
 to be matched \citep{Taylor2006}.
-In brief, the algorithm\footnote{The exact options used
-to band-merge a field using \textsc{stilts tmatchn} are: matcher=sky params=1.0 multimode=group nin=3 in1=r
-in2=i in3=halpha join1=always join2=always join3=always}
+In brief, the algorithm
 identifies groups of linked detections such that
@@
 in all three bands are retained in the catalogue
-as single- or double-band detections, with the columns
-for the missing bands left empty.
+as single- or double-band detections,
+with empty columns for the missing bands.
 
@@
 is unusually large, then the probability that the unusual colour is 
-spurious is greatly increased.
+spurious due to source confusion is greatly increased.
 }
@@
 
-Not all the single-band detections are spurious sources, however.
+Not all the single-band detections are spurious, however.
 We note that the confirmation rate for $i$-band detections
@@
 to the detection and resolution limits,
-our catalogue inevitable contains sources
+our catalogue inevitably contains sources
 with are spurious or have been parametrised inaccurately.
@@
 and to be more prone 
-to have fallen victim to source confusion
+to source confusion
 during the band-merging procedure.
@@
 using the columns \emph{rDeblend}, \emph{iDeblend},
-or \emph{haDeblend}.
+\emph{haDeblend}, and \emph{deblend}.
 \item \emph{Low S/N detections.}
@@
 object is detected at $S/N>10$
-(\emph{rErr}$<$0.1, \emph{iErr}$<$0.1, \emph{haErr}$<$0.1)
-or has photometry which is consistent across
+and has photometry which is consistent across
 the different aperture sizes.
+Such objects are flagged in the catalogue
+using the \emph{a10} column.
 \item \emph{Saturation.}
@@
 brighter than the saturation limit of the instrument
-is subject to systematic errors
-(typically at $r<13$, $i<12$, $\ha<12.5$).
+is subject to systematic errors.
 Such objects can be removed by ensuring that the columns
-\emph{rSaturated}, \emph{iSaturated},
-or \emph{haSaturated} are set to ``False''.
+\emph{rSaturated}, \emph{iSaturated}, \emph{haSaturated}
+or \emph{saturated} are set to ``False''.
 \end{itemize}
@@
 For this reason, we recommend users to read
-the caveats section on the webpage (www.iphas.org/dr2),
+the FAQ section on our website,
 which will be updated as user experience accumulates.
@@
 Most applications will require
-a combination of the quality criteria outlined above
-to remove lower-quality objects.
-In what follows we recommend two sets of such criteria.
-The choice of quality criteria will always tension 
-completeness {\bf against reliability,
-i.e. the fraction of spurious sources}.
+a combination of quality criteria
+to be used
+to avoid the caveats outlined above.
+The choice of criteria will always tension 
+completeness against reliability,
+i.e. the fraction of spurious sources.
 To aid users we have listed two sets of
@@
 
-Table~\ref{tab:a10} specifies
+First, Table~\ref{tab:a10} specifies
 a set of minimum quality criteria
@@
 (i) remove low-S/N sources, 
-(ii) remove saturated sources,
-and (iii) remove objects for which the 2\farcs 3-diameter
+(ii) remove saturated objects,
+and (iii) remove sources for which the 2\farcs 3-diameter
 aperture magnitude is inconsistent 
-with its alternative 1\farcs 2-diameter measurement
-within each filter.
-This last criterion is a proxy
+with its alternative 1\farcs 2-diameter measurement.
+This latter criterion is a proxy
 for identifying objects which are affected
 by poor background subtraction
-or failed source deblending.
+or failed deblending.
 A total of 86 million sources out of 219 million 
@@
 For convenience, the catalogue contains a boolean column
-named \emph{a10} that directly flags these objects.
+named \emph{a10} that flags these objects.
 
@@
 a higher standard of {\bf accuracy} at the expense of completeness,
-a further set of additional quality criteria
+a set of additional quality criteria
 are suggested in Table~\ref{tab:a10point}.
@@
 Fig.~\ref{fig:magdist} compares the $r$-band magnitude
-distributions for \emph{a10}, \emph{a10point}
-and unfiltered objects.
-We find that 81 per cent of the sources 
+distributions for objects tagged \emph{a10} and \emph{a10point}
+against the unfiltered catalogue.
+In the magnitude range $13 < r < 19$,
+we find that 81 per cent of the sources 
 are flagged \emph{a10}
-and 54 per cent are flagged \emph{a10point}
-in the magnitude range $13 < r < 19$.
+and 54 per cent are flagged \emph{a10point}.
 We will explain below that the
@@
 which are genuine astrophysical objects.
-To estimate the reliability, we cross-matched the IPHAS catalogue
-against SDSS DR7,
-which is a catalogue of similar resolution and depth.\footnote{DR7 was chosen for this comparison due to a change in 
-the catalogue generation for DR8, which resulted in fewer stars being 
-detected near bright ($r\,<\,10$) sources, which in turn gave an unduly 
-pessimistic values for the reliability of IPHAS sources.}
+To estimate the reliability, we cross-matched IPHAS
+against SDSS DR7\footnote{SDSS DR7 was chosen for this comparison
+due to an apparent change in 
+the catalogue generation for DR8 and DR9,
+which resulted in fewer genuine stars being
+detected near bright ($r\,<\,12$) sources,
+which in turn gave an unduly pessimistic view
+on our reliability.},
+which is a catalogue of similar resolution and depth.
 By computing the fraction of IPHAS sources which were
-independently detected by SDSS as well,
-we obtain a lower limit for the reliability.
-
-%Such contiguous areas were only found to exist
-%in the second and third Galactic quadrants.
-% as the SDSS strips passing through the early plane are 
-%incomplete.
-
-%, which has a comparable depth.
-
-%observed; provided the comparison is made to an external source of 
-%comparable depth this will provide the ratio of genuine astronomical 
-%sources to sources in the catalogue.
+also independently detected by SDSS,
+we can obtain a good lower limit on the reliability.
 
@@
 (98.7 per cent). 
-Reliabilities $>99\%$ and $>99.5\%$ are achieved using the 
-\emph{a10} and \emph{a10point} quality flags.
-
-To understand the nature of the small fraction of objects
-which appear to be ``unreliable'' in this test,
-despite having been flagged as \emph{a10point},
+Reliabilities of 99.5 and 99.7 per cent are achieved
+using the \emph{a10} and \emph{a10point} quality flags.
+
+To understand the nature of the small fraction
+of \emph{a10point}-flagged sources
+which appear to be unreliable in this test,
 we investigated the data by eye.
-We found these ``unreliable'' sources to be located
+We found these `unreliable' objects to be located
 in the vicinity of moderately bright stars ($r\,\lesssim\,12$),
@@
 scattered light in the SDSS images.
-In all cases these remaining objects appear as genuine
-stars in both the IPHAS and SDSS images,
-and hence we think it is likely that the reliability
-of the \emph{a10point} class of objects is close to 100 per cent.
+In all cases we found these remaining objects 
+to appear as genuine stars in both the IPHAS and SDSS images.
+It is hence likely that the reliability
+of the \emph{a10point} class is close to 100 per cent.
 
@@
 {\bf Towards Sh2-82,
-the IPHAS catalogue and the Spitzer/GLIMPSE archive
+the IPHAS and Spitzer/GLIMPSE catalogues
 have 4798 entries in common,
-out of a total of 10\,739 and 11\,321 entries, respectively.
+out of 10\,739 and 11\,321 entries in total, respectively.
 1356 objects of these objects were detected in all
@@
 that have a counterpart in IPHAS (grey circles),
-as well as the 511 objects that don't (red crosses).
+and the 511 objects that do not (red crosses).
 
@@
 
-We find that objects located in the upper half of the diagram,
-where highly-reddened objects are expected to sit,
-are less likely to appear in both surveys.
+We find that objects located in the upper half of the Spitzer diagram
+are less likely to have an an IPHAS counterpart,
+which is where highly-reddened objects are expected to sit.
 This is not surprising,
@@
 towards a high-extinction sightline 
-near the inner plane
+near the inner plane,
 where an optical survey will naturally be extinction-limited.
@@
 
-Finally, we note that Sh 2-82 is one of a large population 
+We note that Sh 2-82 is one of a large population 
 of poorly-studied star-forming regions
@@
 
-% are 6753 region considered in Fig.~\ref{fig:emitters}
-%contains 6753 objects in the
-%Spitzer GLIMPSE catalogue
-
-%371 objects (20\%) also have a counterpart in the 
-%Wide-field Infrared Survey Explorer catalogue (WISE; Cutri et al. 2013),
-%both within a matching radius of 2 arcsec.
-
 % reliable & r < 20 & (l > 53.55-10/60.) & (l < 53.55+10/60.)                  & (b > 0.01-7.5/60.) & (b < 0.01+7.5/60.)
-% Vizier II/293
 %TODO: This is an example of a high-extinction sightline where
@@
 { \bf
-
 As well as discover emission line stars, IPHAS data can provide a
@@
 our website provides access to the entire set
-of pipeline-processed imaging data,
-the associated metadata,
-and the documentation describing the FITS header keywords
+of pipeline-processed imaging data and
+associated meta data
 (see www.iphas.org/images)}.
@@
 and has not been globally re-calibrated.
-Such data are flagged {\bf in the metadata table}
+Such data are flagged {\bf in the on line meta data table,
+which is available from our website,}
 and must be used with great caution.
@@
 can be found in the works of \citet{Sale2014}, which applies DR2 to the problem
-of 3D extinction mapping, and of Sabin et al (submitted), where the results of a search
+of 3D extinction mapping, and of \citet{Sabin2014}, where the results of a search
 of the image database for new planetary nebulae is presented.

### 2014-07-31 paper/iphas-dr2-rev1.tex
-

@@
     to $A_{Ks}=3$ following the reddening law due to \citet{Flaherty2007}.
-    It is apparent that the most heavily reddened objects
+    It is apparent that objects near the top of the diagram,
+    where heavily reddened objects are expected to sit,
     tend to have no counterpart.
-    Three heavily-reddened objects without counterpart
-    are located above the diagram but are not shown for clarity. }
+    In contrast, many of the objects on the right hand side of the diagram,
+    where YSOs are expected to sit, do have counterparts.
+    %We note that three heavily-reddened objects without counterpart
+    %are located above the diagram but are not shown for clarity. 
+    }
     \label{fig:spitzer-ccd}

### 2014-08-04 paper/referee/response-revision1.txt
Changed suggested by JED

@@
 and 1 new table) to the manuscript.  Changes have been marked in bold.
-We have also ~doubled the auxillary documentation that is available on our web
-site as a direct consequence of the referee's suggestions, and will continue to
-expand the on line documentation as user experience accumulates.
+We have also ~doubled the auxiliary documentation that is available on our web
+site prompted by the referee's suggestions.  It is our intention to
+expand the on-line documentation as user experience accumulates.
 
@@
 Whilst revising the manuscript in response to the items that will follow below,
-we have kept the referee's recommendation to lessen the burden on the reader in
-mind.  In immediate response to this comment, we have extended Section 2 to 
+we have kept in mind the referee's recommendation to lessen the burden on the
+reader.
+In immediate response to this comment, we have extended Section 2 to  
 (i) include an overview table of the survey parameters, which summarises 15 key 
 properties of the survey, and (ii) included an extra figure to remind the reader
-of the footprint of an IPHAS field.
+of the footprint of an IPHAS field pair.
 
@@
 [Authors' response]
-Although the paper does intend to focus on the catalogue product, we agree with the
-referee that it would be beneficial to give users additional guidance on the use 
-of the images.  For this reason, we have made significant improvements to our 
-web site by (i) providing a query tool to retrieve IPHAS images by coordinates, 
+Although the paper itself specifically targets the catalogue product, we
+agree with the referee that it would be beneficial to give users
+additional guidance on the use of the images.  For this reason, we
+have made significant improvements to our web site by (i) providing a
+query tool to retrieve IPHAS images by coordinates, 
 and (ii) provide more comprehensive documentation on the metadata, the header 
@@
 We note that the next release, DR3, intends to focus more strongly on our image
-and image mosaic products, as the survey is approaching ~100% coverage.
+and image mosaic products, when the survey approaches ~100% coverage.
 
@@
 in the catalogue, and in particular questioned the nature of the large number of 
-sources not flagged as "reliable".  Having considered this comment, we have come 
-to conclude that we made a poor choice of terminology by adopting the labels 
+sources not flagged as "reliable".  Having considered this comment, we
+have concluded that we made a poor choice of terminology by adopting the labels 
 "reliable" and "veryReliable" as quality qualifiers.
@@
 and renaming "veryReliable" into "a10point" (short for "all-band 10-sigma and
-perfectly point-like").  We believe this change in terminology will better serve 
-as an operational description of the detection quality.
+perfectly point-like").  We believe this change in terminology works better in
+providing an operational description of the detection quality.
 
-In addition to this change in terminology, we have adopted the referee's suggestion
+In addition to this change, we have adopted the referee's suggestion
 to gauge the reliability of our objects against an external catalogue.  That is,
@@
 against SDSS.  Moreover, we note that a separate paper is in an advanced state 
-of preparation (Farnhill et al.), which will investigate the issues surrounding 
+of preparation (Farnhill et al.), that investigates the issues surrounding 
 the completeness and reliability of our catalogue, e.g. by injecting artificial 
-sources into the images and measuring their recovery rates, whilst also comparing 
-source counts between SDSS and IPHAS in more detail.
+sources into the images and measuring their recovery rates, whilst
+also comparing source counts between SDSS and IPHAS in more detail.
 
@@
 astrometry against UCAC4, which shows a mean residual of 0.09 +/- 0.07 arcsec at
-the bright end of our catalogue.  Finally, we have also added a statement to specify
-that no proper motion information was used in arriving at our astrometry, because
-because hundreds of 2MASS positional reference stars were typically available in 
-each CCD to fit the parameters of the WCS solution.
+the brighter end of our catalogue.  Finally, we have also added a statement
+to specify that no proper motion information was used in arriving at our 
+astrometry, because hundreds of 2MASS positional reference stars were typically
+available in each CCD to fit the parameters of the WCS solution.
 
@@
 The discussion reminds the user that an optical survey will naturally be 
-extinction-limited, by showing where sources which IPHAS and GLIMPE have in common
+extinction-limited, by showing where sources in common between IPHAS and GLIMPSE
 are positioned in a colour-colour diagram.
@@
 exposures, which revealed no evidence for degraded photometry up to ellipticities 
-less than 0.3.  In fact we have found high ellipticities to be correlated with 
-superior seeing, and therefore somewhat deeper frames.
+of 0.3.
 
@@
 naturally picks up at S/N < 5.  We have edited the first paragraph of Section 
-5.4 to clarify this point.  For DR3, we consider following the example set by 
-Spitzer/GLIMPSE to offer both a "highly reliable" as well as a  "more complete,
-less reliable" version of the catalogue to the public.
+5.4 to clarify this point.  For DR3, we are considering following the
+example set by Spitzer/GLIMPSE to offer both a "highly reliable" as
+well as a  "more complete, less reliable" version of the catalogue to
+the community.
 
@@
 of the word "reliable" as the name of a class of high-quality objects was 
-misleading. We believe to have resolved this problem by renaming the "reliable"
-class into the "a10" class, which avoids conveying the incorrect impression that 
+misleading. We believe we have resolved this problem by renaming the "reliable"
+class as the "a10" class, which avoids conveying the incorrect impression that 
 objects not in this class are spurious by default.
@@
 [Authors' response]
-Indeed, we have corrected this mistake. Again, this is a result of our previously
-incorrect usage of the term "reliable".
+Indeed, we have corrected this mistake. Again, this is a result of our 
+previously incorrect usage of the term "reliable".
 
@@
 Again, we thank the referee for the insightful feedback, which has not only
-helped to benefit the DR2 paper, but will also be taken into account during the
+benefitted the DR2 paper, but will also be taken into account during the
 preparation of DR3.

### 2014-08-04 paper/iphas-dr2-rev1.tex
Changed suggested by JED

@@
 The goal of this paper is to present the next release, 
- which supersedes the IDR by including a global re-calibration
+ which supercedes the IDR by including a global re-calibration
 and by taking the coverage up to 92 per cent of the survey area.}
@@
 To set the stage for this release,
-we recap some key points in this section {\bf (see Table~1 for a summary)}.
+we recap some key points in this section 
+{\bf (see Table~1 for a quick-reference overview)}.
 IPHAS is an imaging survey of the Galactic Plane north of the celestial equator, 
@@
 We defined the $5\sigma$ limiting magnitude
-as the magnitude which an imaginary point source would have
+as the magnitude which a point source would have
 if its flux equalled five times the level of the noise in the sky background.
@@
 than $e > 0.2$.
-The inspection of these examples revealed no evidence for
-degraded photometry up to ellipticities less than 0.3.
-In fact high ellipticities are naturally correlated with superior seeing.
-We did however exclude a small number of exposures
-which exceeded $e > 0.3$.}
+The inspection of these examples revealed no evidence
+for degraded photometry up to ellipticities of 0.3.
+We have excluded a small number of exposures
+which exceeded $e = 0.3$.}
 
@@
 in our narrow overlap regions to be insufficient
-to constrain the calibration parameters.
+to constrain the calibration parameters well enough.
 This is because photometry at the extreme edges of the WFC
--- where neighbouring fields overlap -- 
+-- where neighbouring field pairs overlap -- 
 is prone to systematics at the level of 1--2~per cent.
@@
 greater than 20 per cent;
-\item the source must not have been detected in one of the strongly
-vignetted corners of the detector, 
+\item the source must not have been detected
+in one of the strongly vignetted corners of the instrument, 
 not have had any known bad pixels in the aperture,
@@
 is markedly better than for $r$ and \ha,
-which is likely explained by the fact that $i$ is least
+which is likely to be explained by the fact that $i$ is least
 affected by interstellar extinction, and so the survey can occasionally pick up
@@
 Like any other photometric survey
-which operates close
+in which a majority of detected objects are close
 to the detection and resolution limits,
 our catalogue inevitably contains sources
-with are spurious or have been parametrised inaccurately.
+that are spurious or have been parametrised inaccurately.
 In what follows, we highlight the most common
@@
 followed by a discussion on how they can be avoided.
-Known caveats include:
+These caveats include:
 \begin{itemize}
@@
 the different aperture sizes.
-Such objects are flagged in the catalogue
-using the \emph{a10} column.
+Objects detected at $S/N>10$ in all bands
+are flagged in the catalogue using the \emph{a10} column.
 \item \emph{Saturation.}
@@
 
-Most applications will require
+Many applications will require
 a combination of quality criteria
 to be used
-to avoid the caveats outlined above.
+to avoid the issues identified above.
 The choice of criteria will always tension 
@@
 which desire accurate colours
-as well as completeness {\bf down to $\sim$19th magnitude}.
-The listed criteria are designed to 
-(i) remove low-S/N sources, 
-(ii) remove saturated objects,
-and (iii) remove sources for which the 2\farcs 3-diameter
-aperture magnitude is inconsistent 
-with its alternative 1\farcs 2-diameter measurement.
+as well as completeness down to $\sim$19th magnitude.
+The listed criteria are designed
+to ensure that each band offers photometry
+at a high S/N which is consistent
+across different aperture sizes.
 This latter criterion is a proxy
@@
 (39 per cent) satisfy all the criteria listed in Table~\ref{tab:a10}
-and are hereafter referred to as `a10'.
+and are hereafter referred to as `a10'
+(short for ``all-band $10\sigma$'').
 For convenience, the catalogue contains a boolean column
@@
 For applications which require
-a higher standard of {\bf accuracy} at the expense of completeness,
+a higher standard of accuracy at the expense of completeness,
 a set of additional quality criteria
@@
 These criteria are designed to ensure that
-(i) the object appeared as a perfect point source,
+(i) the object appeared as a perfect point source in all bands,
 (ii) the object was not blended with a nearby neighbour,
@@
 out of 10\,739 and 11\,321 entries in total, respectively.
-1356 objects of these objects were detected in all
-four GLIMPSE bands.
-Fig.~\ref{fig:spitzer-ccd} presents their colour-colour diagram,
-in which we have indicated the 845 objects (62 per cent)
-that have a counterpart in IPHAS (grey circles),
+Fig.~\ref{fig:spitzer-ccd} presents the Spitzer colour-colour diagram
+of the region, containing the 1356 objects that were detected 
+in all four GLIMPSE bands.
+We have indicated the 845 objects (62 per cent)
+that have a counterpart in IPHAS (grey circles)
 and the 511 objects that do not (red crosses).
@@
     where YSOs are expected to sit, do have counterparts.
-    %We note that three heavily-reddened objects without counterpart
-    %are located above the diagram but are not shown for clarity. 
     }
@@
 We find that objects located in the upper half of the Spitzer diagram
-are less likely to have an an IPHAS counterpart,
+are less likely to have an IPHAS counterpart,
 which is where highly-reddened objects are expected to sit.
@@
 Stars of this type come with the advantage that secular variation,
-whilst itcertainly occurs,
+whilst it certainly occurs,
 is not as common as in T Tau stars.

### 2014-08-05 paper/iphas-dr2-rev1.tex
Included edit by MJI

@@
 
-All raw IPHAS data were transferred
-to the Cambridge Astronomical Survey Unit (CASU) 
-for initial processing and archiving.
-The procedures used by CASU were originally devised
-for the INT Wide Field imaging Survey \citep[WFS;][]{McMahon2001,Irwin2005},
-which was a 200 deg$^2$ {\bf extragalactic} survey programme carried out 
-between 1998 and 2003.
-Because IPHAS uses the same telescope and camera combination,
-we have been able to benefit from the existing WFS pipeline.
-A description of the processing steps 
-is found in \citet{Irwin2001}.
-Its application to IPHAS has previously been described
+All raw IPHAS data were transferred to the Cambridge Astronomical Survey Unit 
+(CASU) for initial processing and archiving.  The procedures used by CASU 
+were originally devised for the INT Wide Field imaging Survey 
+\citep[WFS;][]{McMahon2001,Irwin2005}, which was a 200 deg$^2$ 
+{\bf extragalactic} survey programme carried out between 1998 and 2003.
+Because IPHAS uses the same telescope and camera combination, we have been 
+able to benefit from the existing WFS pipeline. A description of the 
+processing steps can be found in \citet{Irwin2001}. Its application to IPHAS 
+has previously been described
 by \citet{Drew2005} and \citet{Gonzalez-Solares2008},
-and the source code is available
-on-line\footnote{http://casu.ast.cam.ac.uk/surveys-projects/software-release}. 
-In brief, the pipeline takes care of bias subtraction,
-linearity correction, flat-fielding,
-gain correction and de-fringing.
+and some elements of the the source code are available 
+on line\footnote{http://casu.ast.cam.ac.uk/surveys-projects/software-release}. 
+In brief, the imaging processing part of the pipeline takes care of bias 
+subtraction, the linearity correction, flat-fielding with internal gain
+correction, and de-fringing for the i-band.
 
@@
 Object detection and parametrisation is then carried out
-using the standard methods in use by CASU,
+using the standard methods developed by CASU,
 which can be summarised in four steps
@@
 \begin{enumerate}
-\item The local sky background is estimated
-by computing the sigma-clipped median intensity
-in a grid of 64$\times$64 pixel bins across the image,
-which is robust against contaminating sources.
-Bilinear interpolation is then used to obtain an estimate 
-of the background level at each pixel.
+\item The local sky background is estimated by first computing an iteratively
+sigma-clipped median intensity on a grid of 64$\times$64 pixel bins across 
+the image from each detector.  This is usually robust against contaminating 
+sources corrupting the background level. The resulting array of background
+values is then filtered to further reduce the effect of large objects on
+the local background level. Bilinear interpolation is then used to obtain an 
+estimate of the background level at the orginal image pixel scale.
+\item To improve faint object detection, each image is smoothed with a matched
+detection filter which is used in conjunction with the unsmoothed image
+for object detection and parameterisation.
 \item Objects, or blends of objects, are detected
-by identifying groups of 4 or more neighbouring pixels
-in which the intensity exceeds the background level by at least $1.25\sigma$.
-\item The object are parametrised:
-positions are obtained by computing 
-the intensity-weighted isophotal centre-of-gravity
-of each object,
-and photometry is obtained 
-by measuring the intensity
-in a series of soft-edged circular apertures of increasing diameter 
-(1\farcs 2, 2\farcs 3, 3\farcs 3, 4\farcs 6 and 6\farcs 6).
-Objects are also classified morphologically
--- stellar, extended or noise --
-based on their intensity as a function of aperture size.
-\item The parametrisation of overlapping objects
-is refined by simultaneous fitting of
-soft-edged apertures to each blend,
-effectively carrying out profile fitting where it is necessary.
+by identifying groups of 4 or more neighbouring contiguous pixels
+in which the intensity exceeds the background level by at least $1.25\sigma$
+on the matched filtered image. Objects are deblended using a sequence of
+successively higher detection levels.
+\item The objects are parametrised using the unsmoothed image at these pixel 
+locations : positions are obtained based on an intensity-weighted isophotal 
+centre-of-gravity of each object; whilst photometry is derived by measuring 
+the intensity in a series of soft-edged circular apertures covering a range 
+of diameters (1\farcs 2, 2\farcs 3, 3\farcs 3, 4\farcs 6, 6\farcs 6 and 
+9\farcs 2).  Objects are also classified morphologically
+-- stellar, extended or noise -- based on their intensity as a function of 
+aperture size and on their intensity-weighted second moments, where the latter
+are used to derive an estimate of object ellipticity.
+\item The parametrisation of overlapping objects is refined by simultaneous 
+fitting of soft-edged apertures to each blend,
+effectively carrying out ``top hat'' profile fitting where it is necessary.
 We note that the parametrisation of these blended objects is naturally
@@
 \citep[ZPN;][]{Calabretta2002}
-to model the field distortion of the camera.
+to model the (fixed) field distortion of the camera.
 The parameters of this solution
@@
 to deliver results which are internally consistent
-to better than 0\farcs 1 across the detector
+to better than 0\farcs 1 across the detector array
 \citep{Gonzalez-Solares2008}.
@@
 which is satisfactory for our purposes.
-We note that both UCAC4 and 2MASS -- and hence also IPHAS --
-are calibrated in the International Celestial Reference System (ICRS).
+%We note that both UCAC4 and 2MASS -- and hence also IPHAS --
+%not 100% true due to bulk pm of calibrators - Mike
+%are calibrated in the International Celestial Reference System (ICRS).
 %We note that the mean residual is  
@@
 across the data release:
-
-\begin{figure}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{figures/depth/depth_r.pdf} 
-    \end{minipage}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{figures/depth/depth_i.pdf} 
-    \end{minipage}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{figures/depth/depth_h.pdf} 
-    \end{minipage}
-    \caption{Distribution of the 5$\sigma$ limiting magnitude
-             across all quality-approved survey fields
-             for $r$ (top), $i$ (middle) and \ha\ (bottom).
-             Fields with a limiting magnitude brighter than
-             20th ($r$) or 19th (\ha, $i$) were rejected
-             from the data release.
-             The $r$-band depth is most sensitive 
-             to the presence of the moon above the horizon: 
-             this is the main reason for the wide, bi-modal character
-             of its distribution.}
-    \label{fig:depth}
-\end{figure}
 
@@
 It is expected to be zero (circular)
-in a perfect imaging system,
+in a perfect noise-free imaging system,
 but it is slightly non-zero in any real telescope data 
-due to optical distortions and tracking errors.
+due to optical distortions, tracking errors and photon plus readout noise.
 Indeed, it is worth noting
@@
 is available as a column in the catalogue.
+
+\begin{figure}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=\textwidth]{figures/depth/depth_r.pdf} 
+    \end{minipage}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=\textwidth]{figures/depth/depth_i.pdf} 
+    \end{minipage}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=\textwidth]{figures/depth/depth_h.pdf} 
+    \end{minipage}
+    \caption{Distribution of the 5$\sigma$ limiting magnitude
+             across all quality-approved survey fields
+             for $r$ (top), $i$ (middle) and \ha\ (bottom).
+             Fields with a limiting magnitude brighter than
+             20th ($r$) or 19th (\ha, $i$) were rejected
+             from the data release.
+             The $r$-band depth is most sensitive 
+             to the presence of the moon above the horizon: 
+             this is the main reason for the wide, bi-modal character
+             of its distribution.}
+    \label{fig:depth}
+\end{figure}
 
@@
 for the small difference in internal gain
-computed using the twilight flats
+computed using the relatively blue twilight flats
 compared to the much redder typical astronomical objects. 

### 2014-08-05 paper/iphas-dr2-rev1.tex
Tinkered with figure placement

@@
 Fig.~\ref{fig:astrometry} shows the distribution
-of the position offsets against UCAC4
+of the offsets against UCAC4
 computed for 1.3 million stars
@@
 for the purpose of transforming the raw
-detection tables into a proper source catalogue which
+detection tables into a source catalogue which
 (i) is quality-controlled,
@@
 
+\begin{figure}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=\textwidth]{figures/depth/depth_r.pdf} 
+    \end{minipage}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=\textwidth]{figures/depth/depth_i.pdf} 
+    \end{minipage}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=\textwidth]{figures/depth/depth_h.pdf} 
+    \end{minipage}
+    \caption{Distribution of the 5$\sigma$ limiting magnitude
+             across all quality-approved survey fields
+             for $r$ (top), $i$ (middle) and \ha\ (bottom).
+             Fields with a limiting magnitude brighter than
+             20th ($r$) or 19th (\ha, $i$) were rejected
+             from the data release.
+             The $r$-band depth is most sensitive 
+             to the presence of the moon above the horizon: 
+             this is the main reason for the wide, bi-modal character
+             of its distribution.}
+    \label{fig:depth}
+\end{figure}
+
 (2) \emph{Ellipticity.} 
@@
 is available as a column in the catalogue.
-
-\begin{figure}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{figures/depth/depth_r.pdf} 
-    \end{minipage}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{figures/depth/depth_i.pdf} 
-    \end{minipage}
-    \begin{minipage}[b]{\linewidth}
-        \includegraphics[width=\textwidth]{figures/depth/depth_h.pdf} 
-    \end{minipage}
-    \caption{Distribution of the 5$\sigma$ limiting magnitude
-             across all quality-approved survey fields
-             for $r$ (top), $i$ (middle) and \ha\ (bottom).
-             Fields with a limiting magnitude brighter than
-             20th ($r$) or 19th (\ha, $i$) were rejected
-             from the data release.
-             The $r$-band depth is most sensitive 
-             to the presence of the moon above the horizon: 
-             this is the main reason for the wide, bi-modal character
-             of its distribution.}
-    \label{fig:depth}
-\end{figure}
 
@@
 
+%\newpage
+
 \section{Source catalogue generation}
@@
 
+A total of 219 million primary detections satisfied
+the above criteria and have been included in the catalogue.
+
+Table~\ref{tbl:detections} details the breakdown of these
+sources as a function of the bands in which they are captured.
+159 million sources are detected in all three filters (73 per cent),
+30 million are detected in two filters (14 per cent),
+and the remaining 30 million are single-band detections.
+{\bf
+Table~\ref{tbl:detections} also presents 
+the fraction of `confirmed' objects,
+which we define as those sources
+which have been detected more than once
+(recall that much of the survey area
+is observed twice due to the field pair strategy).
+We find that the single-band detections tend
+to show much lower confirmation rates (32 per cent on average)
+than double- and triple-band detections (89 per cent).
+This suggests that a significant fraction of
+these entries may be spurious detections.
+
 \begin{table}
@@
 
-A total of 219 million primary detections satisfied
-the above criteria and have been included in the catalogue.
-
-Table~\ref{tbl:detections} details the breakdown of these
-sources as a function of the bands in which they are captured.
-159 million sources are detected in all three filters (73 per cent),
-30 million are detected in two filters (14 per cent),
-and the remaining 30 million are single-band detections.
-{\bf
-Table~\ref{tbl:detections} also presents 
-the fraction of `confirmed' objects,
-which we define as those sources
-which have been detected more than once
-(recall that much of the survey area
-is observed twice due to the field pair strategy).
-We find that the single-band detections tend
-to show much lower confirmation rates (32 per cent on average)
-than double- and triple-band detections (89 per cent).
-This suggests that a significant fraction of
-these entries may be spurious detections.
-
 Not all the single-band detections are spurious, however.
@@
 }
+
+%\newpage
 
@@
 \label{tab:a10}
-
 \vspace{2.5cm}
@@
 
-%In producing a point source catalogue, spurious detections 
-%will inevitably be made. A fraction of these will mimic stellar 
-%sources closely enough that they cannot be excluded using the 
-%recorded parameters.
-
 The comparison was carried out across three cone-shaped regions,
@@
 using the \emph{a10} and \emph{a10point} quality flags.
-
-To understand the nature of the small fraction
-of \emph{a10point}-flagged sources
-which appear to be unreliable in this test,
-we investigated the data by eye.
-We found these `unreliable' objects to be located
-in the vicinity of moderately bright stars ($r\,\lesssim\,12$),
-tracing out an area which is affected by saturation spikes or
-scattered light in the SDSS images.
-In all cases we found these remaining objects 
-to appear as genuine stars in both the IPHAS and SDSS images.
-It is hence likely that the reliability
-of the \emph{a10point} class is close to 100 per cent.
 
@@
 
+To understand the nature of the small fraction
+of \emph{a10point}-flagged sources
+which appear to be unreliable in this test,
+we investigated the data by eye.
+We found these `unreliable' objects to be located
+in the vicinity of moderately bright stars ($r\,\lesssim\,12$),
+tracing out an area which is affected by saturation spikes or
+scattered light in the SDSS images.
+In all cases we found these remaining objects 
+to appear as genuine stars in both the IPHAS and SDSS images.
+It is hence likely that the reliability
+of the \emph{a10point} class is close to 100 per cent.
+
 }
@@
 \label{sec:uncertainties}
+
+Fig.~\ref{fig:uncertainties} shows the mean photometric
+uncertainties (\emph{rErr}, \emph{iErr}, \emph{haErr})
+as a function of magnitude.
+We find the typical uncertainty to reach 0.1~mag near $r=$20.5 
+and $i$,\ha$=19.5$.
+We note that the fainter depth in $r$ is compensated
+by the fact that most stars have brighter magnitudes in $i$ and \ha;
+the average colours in the catalogue are
+$\overline{(r-i)}=1.06\pm0.12$ and $\overline{(r-\ha)}=0.44\pm0.03$.
+We warn that the statistics shown in Fig.~\ref{fig:uncertainties}
+are the random errors based on the expected Poissonian photon noise.
+Systematics, such as calibration and deblending errors,
+are not included.
 
@@
 
-Fig.~\ref{fig:uncertainties} shows the mean photometric
-uncertainties (\emph{rErr}, \emph{iErr}, \emph{haErr})
-as a function of magnitude.
-We find the typical uncertainty to reach 0.1~mag near $r=$20.5 
-and $i$,\ha$=19.5$.
-We note that the fainter depth in $r$ is compensated
-by the fact that most stars have brighter magnitudes in $i$ and \ha;
-the average colours in the catalogue are
-$\overline{(r-i)}=1.06\pm0.12$ and $\overline{(r-\ha)}=0.44\pm0.03$.
-We warn that the statistics shown in Fig.~\ref{fig:uncertainties}
-are the random errors based on the expected Poissonian photon noise.
-Systematics, such as calibration and deblending errors,
-are not included.
-
 To appraise the level at which our photometry is affected by such systematics,
@@
 (i.e. \emph{r} vs \emph{r2}, \emph{i} vs \emph{i2}, \emph{ha} vs \emph{ha2}).
-In Fig.~\ref{fig:pairmag} we show the mean absolute residuals between
+In Fig.~\ref{fig:pairmag}a we show the mean absolute residuals between
 these primary and secondary magnitudes as a function of magnitude (black dots).
@@
 
-In Fig.~\ref{fig:pairmag_a10} we show 
-a similar comparison between the primary and secondary detections,
-but this time we have only included sources which are flagged
-as \emph{a10point} in the catalogue.
-We do not observe an improvement in the average residuals as a function of magnitude,
-but the number of outliers has decreased markedly
-(evidenced by the shorter error bars which denote the standard deviation of the absolute residuals).
-We conclude that the \emph{a10point} quality criteria are effective
-at reducing the level of the systematic errors,
-while also removing the inaccurate data at the bright and faint end.
-
-\begin{figure}
-\captionsetup[subfigure]{labelformat=empty}
-\begin{subfigure}[b]{\linewidth}
-\centering
-\includegraphics[width=0.9\textwidth]{figures/repeatability/repeatability.pdf}
-\caption{}
-\label{fig:pairmag}
-\end{subfigure}
-\hspace{0.5cm}
-\begin{subfigure}[b]{\linewidth}
-\centering
-\includegraphics[width=0.9\textwidth]{figures/repeatability/repeatability-a10point.pdf}
-\caption{}
-\label{fig:pairmag_a10}
-\end{subfigure}
+\begin{figure*}
+	\vspace{1cm}
+    \begin{minipage}[b]{\linewidth}
+        \includegraphics[width=0.5\textwidth]{figures/repeatability/repeatability.pdf} 
+        \includegraphics[width=0.5\textwidth]{figures/repeatability/repeatability-a10point.pdf}
+    \end{minipage}
     \caption{Photometric repeatability as a function of magnitude
@@
          We find that the \emph{a10point} quality criteria are successful
-         at removing objects with large residuals.
-    }
-\end{figure}
-\begin{figure*}
-    \vspace{1cm}
+         at removing objects with large residuals.}
+    \label{fig:pairmag}
+
+    \vspace{2cm}
     \includegraphics[width=\textwidth]{figures/sourcecount/sourcecount.pdf} 
@@
 
+In Fig.~\ref{fig:pairmag}b we show 
+a similar comparison between the primary and secondary detections,
+but this time we have only included sources which are flagged
+as \emph{a10point} in the catalogue.
+We do not observe an improvement in the average residuals as a function of magnitude,
+but the number of outliers has decreased markedly
+(evidenced by the shorter error bars which denote the standard deviation of the absolute residuals).
+We conclude that the \emph{a10point} quality criteria are effective
+at reducing the level of the systematic errors,
+while also removing the inaccurate data at the bright and faint end.
+
 \subsection{Source counts and blending}
@@
 spurious is greatly increased.
+
+%\newpage
 
@@
 
-\begin{figure}
-\begin{center}
-  \includegraphics[width=\linewidth]{figures/sh2-82/sh2-82-ccd.pdf}
-    \caption{$(r-\ha,\ r-i)$ diagram for the rectangular region of 
-    20-by-15 arcmin centred on the H{\sc ii} region Sh 2-82,
-    which is the area shown in Fig.~\ref{fig:mosaic_iphas}.    
-    The diagram shows all objects in the catalogue
-    which have been flagged as \emph{a10} and are brighter
-    than $r<20$ (grey circles).
-    The unreddened main sequence is indicated by a solid line,
-    while the main sequence for stars with an \ha\ emission line
-    strength of $-10~\rm{\AA}$~EW is indicated by a dashed line
-    \citep[both based on the colour simulations by][]{Barentsen2011a}.
-    Red triangles indicate objects which have been identified as
-    as likely \ha-emitters.}
-    \label{fig:emitters}
-\end{center}
-\end{figure}
-
 Fig.~\ref{fig:emitters} presents
@@
 Their details are listed in Table~\ref{tbl:emitters}.
-
-\begin{table}
-    \caption{Candidate \ha-emitters towards Sh 2-82.}
-    \label{tbl:emitters}
-    \footnotesize
-    \resizebox{0.47\textwidth}{!}{
-        \begin{tabular}{clccc}
-        \toprule
-        \# & Name [IPHAS2 ...] & $r$ & $i$ & \ha  \\
-        \midrule
-        1 & J192954.40+181026.1& $17.69\pm0.01$ & $16.12\pm0.01$ & $16.19\pm0.01$ \\
-        2 & J193011.01+182051.2& $18.55\pm0.02$ & $16.95\pm0.02$ & $17.31\pm0.02$ \\
-        3 & J193021.52+181954.5& $19.72\pm0.05$ & $17.94\pm0.03$ & $18.47\pm0.04$ \\
-        4 & J193024.45+181938.3& $19.31\pm0.04$ & $17.55\pm0.02$ & $17.99\pm0.03$ \\
-        5 & J193033.00+181609.3& $18.25\pm0.01$ & $16.91\pm0.01$ & $16.92\pm0.01$ \\
-        6 & J193042.48+182317.4& $19.96\pm0.03$ & $18.11\pm0.03$ & $18.48\pm0.03$ \\
-        \bottomrule
-        \end{tabular}
-    }
-\end{table}
 
@@
 data illustrates how IPHAS can complement infrared surveys.
+
+\begin{figure}
+\begin{center}
+  \includegraphics[width=\linewidth]{figures/sh2-82/sh2-82-ccd.pdf}
+    \caption{$(r-\ha,\ r-i)$ diagram for the rectangular region of 
+    20-by-15 arcmin centred on the H{\sc ii} region Sh 2-82,
+    which is the area shown in Fig.~\ref{fig:mosaic_iphas}.    
+    The diagram shows all objects in the catalogue
+    which have been flagged as \emph{a10} and are brighter
+    than $r<20$ (grey circles).
+    The unreddened main sequence is indicated by a solid line,
+    while the main sequence for stars with an \ha\ emission line
+    strength of $-10~\rm{\AA}$~EW is indicated by a dashed line
+    \citep[both based on the colour simulations by][]{Barentsen2011a}.
+    Red triangles indicate objects which have been identified as
+    as likely \ha-emitters.}
+    \label{fig:emitters}
+\end{center}
+\end{figure}
+
+\begin{table}
+    \caption{Candidate \ha-emitters towards Sh 2-82.}
+    \label{tbl:emitters}
+    \footnotesize
+    \resizebox{0.47\textwidth}{!}{
+        \begin{tabular}{clccc}
+        \toprule
+        \# & Name [IPHAS2 ...] & $r$ & $i$ & \ha  \\
+        \midrule
+        1 & J192954.40+181026.1& $17.69\pm0.01$ & $16.12\pm0.01$ & $16.19\pm0.01$ \\
+        2 & J193011.01+182051.2& $18.55\pm0.02$ & $16.95\pm0.02$ & $17.31\pm0.02$ \\
+        3 & J193021.52+181954.5& $19.72\pm0.05$ & $17.94\pm0.03$ & $18.47\pm0.04$ \\
+        4 & J193024.45+181938.3& $19.31\pm0.04$ & $17.55\pm0.02$ & $17.99\pm0.03$ \\
+        5 & J193033.00+181609.3& $18.25\pm0.01$ & $16.91\pm0.01$ & $16.92\pm0.01$ \\
+        6 & J193042.48+182317.4& $19.96\pm0.03$ & $18.11\pm0.03$ & $18.48\pm0.03$ \\
+        \bottomrule
+        \end{tabular}
+    }
+\end{table}
 
@@
 
-\vspace{0.5cm}
+%\vspace{0.5cm}
 

### 2014-08-06 paper/iphas-dr2-rev1.tex
Revision submitted 2014-08-06!

@@
 The goal of this paper is to present the next release, 
- which supercedes the IDR by including a global re-calibration
+which supersedes the IDR by including a global re-calibration
 and by taking the coverage up to 92 per cent of the survey area.}
@@
 which is satisfactory for our purposes.
-%We note that both UCAC4 and 2MASS -- and hence also IPHAS --
-%not 100% true due to bulk pm of calibrators - Mike
-%are calibrated in the International Celestial Reference System (ICRS).
-%We note that the mean residual is  
-%is $-0.006 \pm 0.086$ in R.A.
-%and $0.012\pm0.073$ in Dec.
 
@@
     \includegraphics[width=\linewidth]{figures/astrometry/astrometry.pdf} 
-    \caption{Distribution of the position offsets
-    		of stars
+    \caption{Distribution of the astrometric residuals of stars
     		which appear both in IPHAS DR2 and UCAC4
@@
 and even exceeded 0.5~mag in 0.7 per cent.
-For comparison, Fig.~\ref{fig:ccd_after}
+Fig.~\ref{fig:ccd_after}
 demonstrates the improvement 
@@
 
-%\newpage
-
 \section{Source catalogue generation}
@@
 }
-
-%\newpage
 
@@
 to ensure that each band offers photometry
-at a high S/N which is consistent
+at S/N\ $>10$ that is self-consistent
 across different aperture sizes.
-This latter criterion is a proxy
-for identifying objects which are affected
-by poor background subtraction
-or failed deblending.
+%This latter criterion is a proxy
+%for identifying objects which are affected
+%by poor background subtraction
+%or failed deblending.
 A total of 86 million sources out of 219 million 
@@
 against the unfiltered catalogue.
-In the magnitude range $13 < r < 19$,
-we find that 81 per cent of the sources 
-are flagged \emph{a10}
+we find that 81 per cent of sources 
+in the magnitude range $13 < r < 19$
+are flagged \emph{a10},
 and 54 per cent are flagged \emph{a10point}.
-We will explain below that the
+We will show below that the
 \emph{a10point} category is least complete
@@
 To estimate the reliability, we cross-matched IPHAS
-against SDSS DR7\footnote{SDSS DR7 was chosen for this comparison
+against SDSS DR7\footnote{SDSS DR7 was chosen
 due to an apparent change in 
-the catalogue generation for DR8 and DR9,
+the catalogue preparation for DR8 and DR9,
 which resulted in fewer genuine stars being
@@
 each covering 1 deg$^2$,
-using a strict cross-matching distance of 0.5~arcsec.
+using a strict cross-matching distance of 0\farcs 5.
 We considered four subsets of the IPHAS catalogue:
@@
 
-The results for each of the regions and subsets
-are presented in Table~\ref{tbl:reliability}.
-We find that our catalogue shows a good
-baseline reliability (90.8 per cent),
-which is improved markedly by the simple requirement
-that a source needs to be detected in all three bands
-(98.7 per cent). 
-Reliabilities of 99.5 and 99.7 per cent are achieved
-using the \emph{a10} and \emph{a10point} quality flags.
-
 \begin{table}
     \caption{Fractions of sources in our catalogue
-             which are also found in the SDSS DR7 catalogue
-             within a cross-matching radius of 0.5~arcsec.
-             These statistics can be interpreted
+             which are also found in SDSS DR7
+             within a cross-matching radius of 0\farcs 5.
+             These percentages can be interpreted
              as lower limits
@@
              The reliability is shown
-             for IPHAS sources 
-             without selection criteria applied (``all''),
+             without quality criteria applied (``all''),
              using only objects detected in all three bands ($nBands=3$),
@@
              in all three bands (\emph{a10}),
-             and finally using the detections
+             and finally using objects
              classified as reliable 10$\sigma$ point sources (\emph{a10point}).
@@
 \end{table}
+
+The results for each of the regions and subsets
+are presented in Table~\ref{tbl:reliability}.
+We find that our catalogue shows a good
+baseline reliability (90.8 per cent),
+which is improved markedly by the simple requirement
+that a source needs to be detected in all three bands
+(98.7 per cent). 
+Reliabilities of 99.5 and 99.7 per cent are achieved
+using the \emph{a10} and \emph{a10point} quality flags.
 
@@
 spurious is greatly increased.
-
-%\newpage
 
@@
 the IPHAS and Spitzer/GLIMPSE catalogues
-have 4798 entries in common,
-out of 10\,739 and 11\,321 entries in total, respectively.
+have 4,798 entries in common,
+out of 10,739 and 11,321 entries in total, respectively.
 Fig.~\ref{fig:spitzer-ccd} presents the Spitzer colour-colour diagram
-of the region, containing the 1356 objects that were detected 
+of the region, showing the 1,356 objects that were detected 
 in all four GLIMPSE bands.
@@
 which is where highly-reddened objects are expected to sit.
+Moreover, amongst the objects with a counterpart,
+we note that more have been detected in $i$ (843 objects)
+than in $r$ (634) or \ha\ (600).
 This is not surprising,
@@
 }
-
-%where IPHAS is not limited by extinction
-%This is an example of a high-extinction sightline where
-%optical observations will naturally be extinction-limited.
-%IPHAS will naturally count far fewer sources than (N)IR surveys.
-%In the outer-plane the situation is, in fact, reversed,
-%except where there is UKIDSS GPS coverage (i.e. dec > 60).
-%For example, in the anti-centre region where the dust optical depths are low, IPHAS contains more...
-
-%Amongst the 1810 high-quality sources which we
-%plotted in Fig.~\ref{fig:emitters},
-%no less than 1561 objects (87\%) have a counterpart
-%in the Spitzer GLIMPSE archive
-%within a matching radius of 1 arcsec.
-%In the other direction, out of the 6753 objects
-%which are in the Spitzer/GLIMPSE catalogue for this region,
-%2984 (44\%) have a counterpart in IPHAS.
-
-% reliable & r < 20 & (l > 53.55-10/60.) & (l < 53.55+10/60.)                  & (b > 0.01-7.5/60.) & (b < 0.01+7.5/60.)
-%TODO: This is an example of a high-extinction sightline where
-% optical observations will naturally be extinction-limited.
-% IPHAS will naturally count far fewer sources than (N)IR surveys.
-% In the outer-plane the situation is, in fact, reversed,
-% except where there is UKIDSS GPS coverage (i.e. dec > 60).
-% For example, in the anti-centre region where the dust optical depths are low, IPHAS contains more...
 
@@
 
-%\vspace{0.5cm}
-
 \section{Concluding remarks and future work}

### 2014-08-11 paper/iphas-dr2-rev2.tex
Revision 2, submitted 11 Aug 2014

--- 2014-08-06 paper/iphas-dr2-rev1.tex
@@
-% This is revision 1 of the IPHAS DR2 paper,
-% created in response to the referee comments received on 24 June 2014
+% This is revision 2 of the IPHAS DR2 paper,
+% created in response to the referee comments received on 11 August 2014
 \documentclass[a4paper,useAMS,usenatbib]{mn2e}
@@
 
-{\bf When the previous release -- the IDR -- was created in late 2007,
+When the previous release -- the IDR -- was created in late 2007,
 just over half of the survey footprint was covered
@@
 which supersedes the IDR by including a global re-calibration
-and by taking the coverage up to 92 per cent of the survey area.}
+and by taking the coverage up to 92 per cent of the survey area.
 In this work we
@@
 we recap some key points in this section 
-{\bf (see Table~1 for a quick-reference overview)}.
+(see Table~1 for a quick-reference overview).
 IPHAS is an imaging survey of the Galactic Plane north of the celestial equator, 
@@
 and to enable quality checks.
-{ \bf An example footprint of a pointing
-and its offset position is shown in Fig.~\ref{fig:fieldpair}.}
+An example footprint of a pointing
+and its offset position is shown in Fig.~\ref{fig:fieldpair}.
 Hence, the basic unit of observation
@@
 \citep[WFS;][]{McMahon2001,Irwin2005}, which was a 200 deg$^2$ 
-{\bf extragalactic} survey programme carried out between 1998 and 2003.
+extragalactic survey programme carried out between 1998 and 2003.
 Because IPHAS uses the same telescope and camera combination, we have been 
@@
 
-{ \bf
 Object detection and parametrisation is then carried out
@@
 are then progressively refined by fitting against the
-Two-Micron All Sky Survey \citep[2MASS; ][]{Skrutskie2006}.
-We did not include proper motion information during this step,
-because each CCD typically contains
-hundreds of 2MASS reference stars.
+Two-Micron All Sky Survey \citep[2MASS; ][]{Skrutskie2006},
+{ \bf
+albeit without correcting for the $\sim$10 years of proper motion
+between the IPHAS and 2MASS epochs. }
 The resulting fit has previously been shown
@@
 Fig.~\ref{fig:astrometry} shows the distribution
-of the offsets against UCAC4
+of the astrometric offsets
 computed for 1.3 million stars
@@
 is better than 50~mas \citep{UCAC3}.
-We find the mean position offset to be $0\farcs 09\pm0\farcs 07$,
+We find the mean difference in position 
+between IPHAS and UCAC4 to be 94~mas,
 which is satisfactory for our purposes.
+{ \bf
+The remaining residuals are in part due to
+the motion of the Earth through our galaxy,
+which we did not account for.
+}
 
@@
             The mean and standard deviation of this distribution 
-            equals $0\farcs 09\pm0\farcs 07$.}
+            equals $94\pm65$~mas.}
     \label{fig:astrometry}
@@
 of unique sources were not flagged.
-These improvements are explained next.}
+These improvements are explained next.
 
@@
 due to mechanical glitches in the telescope tracking system.
-{ \bf As a result, 3 per cent of our images show an
+As a result, 3 per cent of our images show an
 average ellipticity across the detectors which is worse
@@
 We have excluded a small number of exposures
-which exceeded $e = 0.3$.}
+which exceeded $e = 0.3$.
 
@@
 (we adopt a 2\farcs 3-diameter circular aperture by default).
-{\bf
 Finally, \textsc{percorr} is a term used to correct 
@@
 All these quantities correspond to header keywords
-in the FITS files produced by the CASU pipeline. }
+in the FITS files produced by the CASU pipeline.
 
@@
 in the timing of exposures).
-{ \bf
 This is demonstrated in Fig.~\ref{fig:ccd_before},
@@
 in the colour-colour diagram
-after re-calibrating. }
+after re-calibrating.
 
@@
 can use the repeat measurements to ensure a homogeneous calibration.
-{\bf For example, 2MASS attained its global calibration
-by observing two standard fields each hour, }
+For example, 2MASS attained its global calibration
+by observing two standard fields each hour,
 allowing zeropoint variations to be tracked 
@@
 to validate the results.
-{\bf SDSS DR9 includes photometry across several strips at low
+SDSS DR9 includes photometry across several strips at low
 Galactic latitudes,
 which were observed as part of 
-the Sloan Extension for Galactic Understanding and Exploration \citep[SEGUE;][]{segue}.}
+the Sloan Extension for Galactic Understanding and Exploration \citep[SEGUE;][]{segue}.
 These strips provide data across 18 per cent of the fields in our data release.
@@
 
-{\bf
 The second step in compiling the source catalogue
@@
 spurious due to source confusion is greatly increased.
-}
 
@@
 that is described in Appendix~\ref{app:columns}.
-{ \bf
 The original unweeded list of sources naturally included 
@@
 which must be met for a candidate source
-to be included in the catalogue:}
+to be included in the catalogue:
 \begin{enumerate}
@@
 and the remaining 30 million are single-band detections.
-{\bf
 Table~\ref{tbl:detections} also presents 
@@
 was detected more than once (\emph{nObs}$>$1).
-}
 
@@
 \label{sec:discussion}
-
-{\bf
 
@@
 across different aperture sizes.
-%This latter criterion is a proxy
-%for identifying objects which are affected
-%by poor background subtraction
-%or failed deblending.
 A total of 86 million sources out of 219 million 
@@
 
-}
-
 \subsection{Random and systematic uncertainties}
@@
 
-{\bf Towards Sh2-82,
+Towards Sh2-82,
 the IPHAS and Spitzer/GLIMPSE catalogues
@@
 and stellar contents of many of such unexplored regions.
-}
 
@@
 
-{ \bf
 As well as discover emission line stars, IPHAS data can provide a
@@
 
-In order to infer the emission EW from the DR2 $r - \ha$ colour
+In order to infer the emission EW from the $r - \ha$ colour
 it is necessary to know the value this colour would have
@@
 
-}
-
 \section{Catalogue and image access}
@@
 aperture diameters specified in this work.
-{\bf To enable the analysis of diffuse sources,
+To enable the analysis of diffuse sources,
 our website provides access to the entire set
@@
 associated meta data
-(see www.iphas.org/images)}.
+(see www.iphas.org/images).
 The image headers have been updated to include
@@
 and has not been globally re-calibrated.
-Such data are flagged {\bf in the on line meta data table,
-which is available from our website,}
+Such data are flagged in the on line meta data table,
+which is available from our website,
 and must be used with great caution.

### 2014-08-12 paper/iphas-dr2-rev2-arxiv.tex
Rev2 as uploaded to arxiv

--- 2014-08-11 paper/iphas-dr2-rev2.tex
@@
-% This is revision 2 of the IPHAS DR2 paper,
-% created in response to the referee comments received on 11 August 2014
 \documentclass[a4paper,useAMS,usenatbib]{mn2e}
@@
 \author[G. Barentsen
-et. al]{Geert Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
+et al.]{Geert Barentsen$^{1}$\thanks{E-mail:geert@barentsen.be},
 H. J. Farnhill$^1$,
@@
 Two-Micron All Sky Survey \citep[2MASS; ][]{Skrutskie2006},
-{ \bf
 albeit without correcting for the $\sim$10 years of proper motion
-between the IPHAS and 2MASS epochs. }
+between the IPHAS and 2MASS epochs.
 The resulting fit has previously been shown
@@
 which is satisfactory for our purposes.
-{ \bf
-The remaining residuals are in part due to
-the motion of the Earth through our galaxy,
+These residuals are in part due to
+the motion of the Earth through our Galaxy,
 which we did not account for.
-}
 
@@
 We discarded any exposures 
-for which the $5\sigma$ limiting magnitude\footnote{\bf
+for which the $5\sigma$ limiting magnitude\footnote{
 We defined the $5\sigma$ limiting magnitude
@@
 \vspace{2.5cm}
-\caption{\bf Recommended quality criteria 
+\caption{Recommended quality criteria 
 for selecting objects with accurate colours.
@@
 \vspace{2.5cm}
-\caption{\bf Additional quality criteria which are recommended
+\caption{Additional quality criteria which are recommended
 for applications which require objects to be
@@
    145 million (66\%) &
-   % pStar > 0.89 is identical to mergedClass == -1
    The object must appear as a perfect point source,
@@
 
-For applications which require
-a higher standard of accuracy at the expense of completeness,
-a set of additional quality criteria
-are suggested in Table~\ref{tab:a10point}.
-These criteria are designed to ensure that
-(i) the object appeared as a perfect point source in all bands,
-(ii) the object was not blended with a nearby neighbour,
-and (iii) the object was not located near a very bright star.
-59 million sources (27 per cent) satisfy
-these stricter criteria 
-and are hereafter referred to as `a10point'.
-Again, the catalogue contains a boolean column
-named \emph{a10point} which flags these objects.
-
 \begin{figure}
-    \includegraphics[width=0.5\textwidth]{figures/magdist/magdist-r.pdf} 
+    \includegraphics[width=0.45\textwidth]{figures/magdist/magdist-r.pdf} 
     \caption{r-band magnitude distribution
@@
 \end{figure}
+
+For applications which require
+a higher standard of accuracy at the expense of completeness,
+a set of additional quality criteria
+are suggested in Table~\ref{tab:a10point}.
+These criteria are designed to ensure that
+(i) the object appeared as a perfect point source in all bands,
+(ii) the object was not blended with a nearby neighbour,
+and (iii) the object was not located near a very bright star.
+59 million sources (27 per cent) satisfy
+these stricter criteria 
+and are hereafter referred to as `a10point'.
+Again, the catalogue contains a boolean column
+named \emph{a10point} which flags these objects.
 
@@
 
+In Fig.~\ref{fig:pairmag}b we show 
+a similar comparison between the primary and secondary detections,
+but this time we have only included sources which are flagged
+as \emph{a10point} in the catalogue.
+We do not observe an improvement in the average residuals as a function of magnitude,
+but the number of outliers has decreased markedly
+(evidenced by the shorter error bars which denote the standard deviation of the absolute residuals).
+We conclude that the \emph{a10point} quality criteria are effective
+at reducing the level of the systematic errors,
+while also removing the inaccurate data at the bright and faint end.
+
 \begin{figure*}
@@
 
-    \vspace{2cm}
+    \vspace{1.5cm}
     \includegraphics[width=\textwidth]{figures/sourcecount/sourcecount.pdf} 
@@
 \end{figure*}
-
-In Fig.~\ref{fig:pairmag}b we show 
-a similar comparison between the primary and secondary detections,
-but this time we have only included sources which are flagged
-as \emph{a10point} in the catalogue.
-We do not observe an improvement in the average residuals as a function of magnitude,
-but the number of outliers has decreased markedly
-(evidenced by the shorter error bars which denote the standard deviation of the absolute residuals).
-We conclude that the \emph{a10point} quality criteria are effective
-at reducing the level of the systematic errors,
-while also removing the inaccurate data at the bright and faint end.
 
@@
 
+% woof woof ／人 ◕‿‿◕ 人＼ woof woof
+
 \section*{Acknowledgments}
